### 2011-03-23 ANTsTBSS/paper.tex
Add papers and figures.

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%

%\documentclass[preprint,authoryear,review,12pt]{elsarticle}
\documentclass[final,5p,times,twocolumn]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{color}
\usepackage{multirow,booktabs,ctable,array}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{ulem}
\usepackage{setspace}
\usepackage{listings}
\usepackage{float}

\floatstyle{plain}
\newfloat{command}{thp}{lop}
\floatname{command}{Command}

%\usepackage[nomarkers,notablist]{endfloat}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
% \usepackage{amsthm}

 \usepackage{makecell}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\providecommand{\OO}[1]{\operatorname{O}\bigl(#1\bigr)}

\graphicspath{{./Figures/}
                          }

\long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

    \usepackage{color}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.955}
    \definecolor{listwhite}{gray}{1.0}

\newcommand{\lstsetcpp}
{
\lstset{frame = tb,
        framerule = 0.25pt,
        float,
        fontadjust,
        backgroundcolor={\color{listlightgray}},
        basicstyle = {\ttfamily\scriptsize},
        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
        identifierstyle = {\ttfamily},
        commentstyle = {\ttfamily\color{listcomment}\textit},
        stringstyle = {\ttfamily},
        showstringspaces = false,
        showtabs = false,
        numbers = none,
        numbersep = 6pt,
        numberstyle={\ttfamily\color{listnumbers}},
        tabsize = 2,
        language=[ANSI]C++,
        floatplacement=!h,
        caption={},
        captionpos=b,
        }
}

\journal{Neuroimage}

\begin{document}

\begin{frontmatter}

\title{ANTs-Flavored Tract-Based Spatial Statistics}

\author[label1]{Nicholas J.~Tustison\fnref{label0}}
%  \ead{ntustison@virginia.edu}
  \fntext[label0]{\scriptsize Corresponding author:  PO Box 801339, Charlottesville, VA 22908; T:  434-924-7730; email address:  ntustison@virginia.edu }
\author[label2]{Brian B.~Avants}
\author[label2]{Philip A.~Cook}
\author[label3]{Junghoon Kim}
\author[label3]{John Whyte}
\author[label2]{James C.~Gee}
\author[label1]{James R.~Stone}

\address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
\address[label2]{Penn Image Computing and Science Laboratory, University of Pennsylvania,
                Philadelphia, PA}
\address[label3]{Moss Rehabilitation Research Institute, Albert Einstein Healthcare Network, Philadelphia, PA}

%\maketitle

%\linenumbers

\begin{abstract}

%% Text of abstract

We propose a modification of the popular tract-based spatial statistics (TBSS) framework \citep{Smith2006} for investigating diffusion tensor derived scalar measures across populations.   Whereas the current crucial alignment step is performed typically using the derived fractional anisotropy images, we propose using the Symmetric Group Normalization strategy for anatomical unbiased template construction from T1-weighted images which serves as the coordinate system for statistical inference.   A potential benefit of this modified approach is reduction of the general confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.  Alignment performance is evaluated using open data sets from the International Neuroimaging Data-sharing Initiative.  This assessment demonstrated improved alignment in both fractional anisotropy (FA) and mean diffusivity (MD) images versus standard TBSS.  Further analyses evidencing increased regional sensitivity in peripheral white matter to FA differences include a DTI-based assessment of a cohort of 16 traumatic brain injury survivors compared with 17 control subjects.  Similar to the Oxford Centre for Functional MRI of the Brain (FMRIB) who provide TBSS to the public, the software tools discussed herein are provided to the public as part of the open source Advanced Normalization Tools (ANTs) repository. 
\end{abstract}

\begin{keyword}
Advanced Normalization Tools (ANTs) \sep diffusion imaging \sep fractional anisotropy \sep template construction \sep tract-based spatial statistics \sep traumatic brain injury
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

\end{frontmatter}
%
%
\newpage

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text

\section{Introduction}
%State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results.

Ever since the seminal work of Basser et al. \cite{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique, continued quantitative development has furthered its applicability to probing clinical questions of interest particularly pertaining to the microstructure of white matter \cite{Basser1996,Assaf2008}.  Given this sensitivity to brain architecture, DTI has found multiple uses in assessing neuro-structural differences in cross-population studies \cite{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010}
including subjects with traumatic brain injury (TBI) \cite{Caeyenberghs2010,Jiang2010,Warner2010}.  One such statistical framework which has significantly facilitated quantitative analysis of DTI data is the well-known tract-based spatial statistics (TBSS) originally formulated in \cite{Smith2006} and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).%
\footnote{
http://www.fmrib.ox.ac.uk/fsl/tbss/index.html
} 
Based solely on the number of citations in the literature,%
\footnote{
As of the beginning of the year 2011, the original Neuroimage 2006 TBSS paper has been cited more than 360 times making it the most cited in the journal's library.
}
usage of TBSS has certainly permeated the neuroimaging community and, in combination with its accessibility and ease of use, has become a de facto quantitative tool for DTI analysis.

\begin{figure*}
\begin{tabular}{c}
  \includegraphics[width=170mm]{workflow.pdf}
\end{tabular}
\caption{Proposed ANTs-TBSS workflow.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
(transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.  The FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.  Each individual T1 is then nonrigidly registered to the anatomical template.  Optionally, one can register the T1 template to the MNI152 template which resides in the same space as the FMRIB58 template.  Thus, the composite transform, carefully constructed without any FA-to-FA image registration, can take an individual subject's FA map (or other DTI-derived measure) to the desired standardized coordinate system.    }
\label{fig:workflow}
\end{figure*}

Prior to the advent of TBSS, voxel-based approaches, such as statistical parametric mapping \cite{Ashburner2000}, were used in assessing voxel-wise population differences \cite{Rugg-Gunn2001,Watkins2001,Kubicki2005,Park2004} using DTI data. In addition to concerns regarding the general application of SPM \cite{Bookstein2001,Ashburner2001,Davatzikos2004}, particularly the requirement for stringent spatial normalization,
issues specific to the application to DTI-derived measures, such as the effects of size variation of the smoothing kernel and the potential non-normality of the residuals \cite{Jones2005}, encourage caution in the application of SPM to DTI-derived data.  Although the latter can be addressed via the use of non-parametric testing \cite{Chung2008}, addressing the former in a principled way is elusive.

It was precisely due to these concerns that the TBSS framework was developed constituting an entire processing pipeline spanning from the reconstruction of the DTI tensorial data
%to the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA)), to the nonrigid alignment of the population to a standard coordinate system and skeletonization, and, finally, 
through to the statistical analysis for discerning differences across populations.  Major intermediate steps include the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA) \cite{Basser1996}), nonrigid alignment of the population to a standard coordinate system, and subsequent ``skeletonization'' of the mean of the aligned FA maps.  The alignment to a standard space is performed in such a way to remove the smoothing step associated with SPM.  Skeletonization and its projection across all registered DTI-derived scalar images is meant to accommodate the possibility of imperfectly aligned data.  However, as shown in a recent study \cite{Zalesky2011}, although the alignment error is reduced with the skeleton projection step, misalignments still remain post-projection.  Added to this is the potential confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.

Because spatial normalization is critical to obtaining valid results, we propose a modification of the standard TBSS pipeline directed at some of these residual concerns which we refer to as ANTs-TBSS.  Instead of DTI-derived scalar alignment via the images themselves (either random subject selection or locating an approximation to the ``mean subject'') or to a standard template (e.g. FMRIB58\_FA\_1mm.nii.gz which we denote as ``FMRIB58'' in the remainder of the text), we evaluate the substitution of an anatomical, i.e. T1-weighted, unbiased shape/intensity template constructed from the population images.  DTI-derived images can then be mapped to their corresponding T1-weighted image in a highly constrained manner and subsequently mapped to the template via nonrigid registration of the T1-weighted image.  This eliminates any possible complications associated with direct FA-to-FA image registration.

We recognize that using the anatomical images in the alignment process has been previously proposed (in fact, such a suggestion occurs in the {\it Future directions} section of the original TBSS paper).  Our contribution stems from the specific proposal of using a high performance \cite{Klein2009}, nonrigid registration algorithm \cite{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \cite{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.  This results in a few advantages in using ANTs-TBSS vs standard TBSS:
\begin{itemize}
\item As explained in \cite{Avants2010}, ANTs optimal template construction creates a true Frech\'et mean from the population sample with high resolution which facilitates within-group registration.  This obviates the need for an externally-derived template or an $\OO{n^2}$ search for an approximation to the mean subject or the bias resulting from random subject selection.
\item Since the creation of the standardized coordinate system is performed using the anatomical images, nonrigid registration of the DTI-derived scalar images is 
minimized thus mitigating the confounds described previously.  Each DTI-based image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \cite{Wu2008}.  
\item Although ANTs-TBSS increases the total number of registrations to be performed, this makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
\item Instead of Gaussian smoothing as with SPM, TBSS uses the warping process to FMRIB58 by default to provide an indirect smoothing.  Such resampling occurs in ANTs-TBSS as a by-product of the registration of the DTI-derived image to the corresponding T1 image in a subject-specific manner.
%\item Instead of using thresholding to create the white matter segmentation from which the skeleton is created, an anatomically-based segmentation, either in the template space or in each subject space followed by a label fusion technique (e.g. majority voting), can be used to ultimately provide a more accurate skeleton.
\item Oftentimes TBSS is only one of several neuroimage analyses performed in a population-based study.  Other investigations, such as voxel- and tensor-based morphometry, cortical thickness measurements, etc. can take advantage of the anatomical template used in ANTs-TBSS.
\end{itemize}

Of practical significance, we note that all software components described in this paper, similar to FMRIB's software library (FSL), of which TBSS is a component, are publicly available as open source primarily through the Advanced Normalization Tools (ANTs) repository.%
\footnote{
http://www.picsl.upenn.edu/ANTs
}
In addition to registration and template building, ANTs also includes brain extraction \cite{Avants2010a}, bias correction \cite{Tustison2010}, $n$-tissue segmentation \cite{Avants2011a}, and cortical thickness estimation solutions \cite{Das2009}.  %As part of this work, we also explicitly show how one can merge the ANTs and TBSS components into a single modified ANTs-flavored TBSS pipeline.
\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=80mm]{NKI.pdf}
\end{tabular}
\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.  (a) The standard FMRIB58 template. (b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).  
(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).  Note the visual alignment difference particularly in the peripheral white matter complex.
}
\label{fig:NKI}
\end{figure}

We first describe in greater detail the proposed ANTs-TBSS framework with an emphasis on comparison with standard TBSS.  We then describe the two open data sets used to evaluate the proposed methodology in addition to a description of diffuse TBI data 
also used to illustrate concepts discussed in this work.
%and analysis previously reported in \cite{Avants2008}.   
%The complementary assessment performed in this paper is used to provide additional analyses to these previous findings.

%A crucial component of TBSS is the alignment of the DTI-derived scalar images \cite{Zalesky2011}. 
%
%Voxel-based morphometry vs. tractography vs. tract-specific analysis
%
%Talk about {\em skeleton projection} and how our work is different from \cite{Zalesky2011}.
%Also talk about the advantages over building a dti-based template as discussed in 
%\cite{Hecke2010}.

\section{Material and Methods}

%In order to understand the differences in approaches, we first briefly describe
%the current TBSS protocol as it is implemented.  This is followed by a
%brief outline of the proposed modification using ANTs.  Afterwards
%we give a description of the data sets used to demonstrate the methods.

\subsection{Tract-Based Spatial Statistics}

FSL hosts a large number of excellent algorithms for computational neuroanalysis.  These include, among others, well-known tools for image segmentation (FAST---\cite{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), as well as TBSS.
For the comparative evaluation studies used in this manuscript, TBSSv1.2%
\footnote{
Build 416 downloaded from the FMRIB web site http://www.fmrib.ox.ac.uk/fsldownloads/ on November 18, 2010.
}
was used.  The actual code pipeline constituting TBSS prior to statistical analysis consists of the following 4 shell scripts:
\begin{description}
  \item[tbss\_1\_preproc]  Input is a directory of FA images which are slightly eroded and then organized into a separate FA directory.  A mask is also generated for each input FA image for use during the spatial normalization process.
  \item[tbss\_2\_reg]  The organized FA data and corresponding masks produced in the previous step are warped to one of three possible coordinate systems using FLIRT/ FNIRT:  1)  the atlas included with the FSL download---FMRIB58---(recommended), 2) a user-specified target image or 3) an exhaustive search for the ``optimal'' target image.  In \cite{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.  However, even though it would be faster to simply choose a single subject at random, the complexities of nonrigid image registration would seemingly promote the exhaustive search over random selection.   As a compromise between these two options, the first option in the script is to use FMRIB58 which also resides in the space of the MNI152 T1 template.  It should also be noted that this particular script invokes another FSL script, {\bf fsl\_reg}.  An option associated with {\bf fsl\_reg} permits the use of an FNIRT configuration file which has been optimized for use with the FMRIB58 atlas.  
  \item[tbss\_3\_postreg] The mean FA image is generated from the aligned FA images which is subsequently skeletonized.
  \item[tbss\_4\_prestats] The skeleton mask is created by thresholding the mean FA skeleton.  The skeleton distance map is then generated and all FA data are projected onto the skeleton.
\end{description}

\subsection{ANTs-Flavored TBSS}
\label{sec:ANTs-TBSS}

Our proposed modification to standard TBSS is illustrated in Figure \ref{fig:workflow}.  Instead of direct FA-to-FA registration, our ANTs-based modification can be described as follows:

\begin{enumerate}
  \item Build the anatomical template from  the population sample or sample subset using SyGN.
  \item If only using a sample subset for template creation, warp the remaining
        anatomical data to the template and save the resulting transformation for each         
        subject indexed by $i$ ($T^i_{template}$).
  \item Find the optimal transformation from each
        subject's FA to the anatomical template:
  \begin{enumerate}
  \item For each anatomical image, segment the white matter (in ANTs one can use N4 bias correction and Atropos $n$-tissue segmentation).
  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical
        image ($R^i_{dwi}$).
  \item Perform distortion correction on the FA to anatomical mapping 
        by localized nonrigid registration
        of rigidly transformed FA map ($=R^i_{dwi} \mathrm{FA}$) using
        the white matter mask obtained in step 3 ($T^i_{FA}$).
  \item For each subject, concatenate the transforms which maps each
        subject's FA to the template ($T^i_{total} = T^i_{template} \circ 
        T^i_{FA} \circ R^i_{dwi} \mathrm{FA}$).
  \item Warp each subject's FA or MD map to the template using $T^i_{total}$.
  \end{enumerate}      
  \item Perform the remaining portions of the standard TBSS pipeline:
    \begin{enumerate}
      \item Create the mean FA image from all template-aligned FA images.
      \item Produce the white matter skeleton from the mean FA image.
      \item Create skeleton distance map and project all FA data onto the skeleton.
      \item Run the statistical analysis.
    \end{enumerate}
\end{enumerate}

%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.

%\subsection{Symmetric Group Normalization Template Creation}

%Due to the nonlinearity inherent in morphometric neuromaging studies, choice of coordinate
%system can affect quantitative outcome \citep{Thompson2000,Senjen2005}.  
%This dependence has motivated significant work in defining coordinate systems
%using population-derived templates \cite{Joshi2004,Lorenzen2006,Avants2010}. 
%
%based on \cite{Klein2009}
%symmetric group-wise normalization scheme for template co
%
%nstruction was presented in \cite{Avants2010}.  
%
%
%\subsection{Benefits of our approach}
%\begin{itemize}
%   \item White matter segmentations can be performed on the T1 template 
%\end{itemize}
%
%
%\subsection{ANTs-Flavored TBSS}

%\subsection{ANTs-Flavored Tract-Based Spatial Statistics}
%
%\begin{enumerate}
%  \item Create the anatomical template from sample T1 images, i.e.
%\begin{command}
%\lstsetcpp
%\begin{lstlisting}
%$ buildtemplateparallel.sh -d 3 -o T_ -c 0 -g 0.25 -i 4 \
%    -m 50x50x20 -n 1 -r 1 -s CC -t GR T1*.nii.gz
%\end{lstlisting}
%\end{command} 
%\newline
%  where the following flags are used:
%  \begin{description}
%    \item[-d] is the image dimension,
%    \item[-o] is the output prefix where the final template will be named \verb#T_template.nii.gz#,
%    \item[-c] specifies the computational parallelization method,
%    \item[-g] is the gradient step size, 
%    \item[-i] is the number of template-building iterations,
%    \item[-m] is the maximum number of iterations per pairwise registration,
%    \item[-n] is a boolean flag denoting whether or not N4 bias correction is used,
%    \item[-r] is a boolean flag denoting whether or not a rigid registration is used to initialize the template,
%    \item[-s] specifies the similarity metric used (\verb#CC# = cross correlation), and
%    \item[-t] specifies the transformation model used (\verb#GR# = greedy SyN).
%  \end{description}
%The list of files \verb#T1*.nii.gz# denote the set of T1 images used to create the template.  The output consists of the template and the transforms mapping between the subject and template.  Although these parameters generally work quite well, additional information can be found in the help menu of \verb#buildtemplateparallel.sh#.
%\item For each subject, use N4 and Atropos to segment the white matter in each subject's T1 following skull stripping (options include BET \cite{Smith2002} or Atropos \cite{Avants2010a}).  
%\begin{command}
%\lstsetcpp
%\begin{lstlisting}
%$ N4BiasFieldCorrection -d 3 -i T1_subX_brain.nii.gz \
%    -b [200] -c [50x50x50x30,1e-5] \
%    -x T1_subX_mask.nii.gz -o T1_subX_n4.nii.gz
%\end{lstlisting}
%\end{command} 
%\newline
%  where \verb#T1_subX_mask.nii.gz# is the output of the skull stripping step 
%  with brain voxel values of `1' and `0' elsewhere.  The following flags are used:
%  \begin{description}
%    \item[-d] is the image dimension,
%    \item[-i] defines the input image,
%    \item[-b] specifies the ``spline distance'',
%    \item[-c] defines the iterations and convergence threshold, and
%    \item[-o] specifies the output image.
%  \end{description}
%Other options can be explored with the help menu (use \verb#--help#).
%\begin{command}
%\lstsetcpp
%\begin{lstlisting}
%$ Atropos -d 3 -a T1_subX_n4.nii.gz -i kmeans[3] \
%    -m [0.2,1x1x1] -c [5,0.0000001] -k Gaussian \
%    -x T1_subX_mask.nii.gz -o T1_subX_atropos.nii.gz
%\end{lstlisting}
%\end{command} 
%\newline
%  the following flags are used:
%  \begin{description}
%    \item[-d] is the image dimension,
%    \item[-a] defines the input image,
%    \item[-i] specifies the initialization method (here we are using K-means with 3 classes),
%    \item[-m] is the 
%    \item[-c] defines the iterations and convergence threshold, and
%    \item[-o] specifies the output labeled image.
%  \end{description}
%Other options can be explored with the help menu (use \verb#--help#).
%\newline
%This step provides the white matter ma
%\end{enumerate}

\subsection{Imaging Data}

The tensorial data associated with each of the following data sets were reconstructed from the diffusion weighted sequence using Camino%
\footnote{
http://web4.cs.ucl.ac.uk/research/medic/camino/pmwiki/
}---an open source toolkit for diffusion MRI processing and analysis \cite{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.

\subsubsection{Data from the International Neuroimaging Data-sharing Initiative (INDI)}

In support of open science, the 1000 Functional Connectomes Project (FCP)%
\footnote{
ttp://fcon\_1000.projects.nitrc.org
} 
was initiated
on December 11, 2009 by various members of the MRI community \cite{Biswal2010}.  Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
Commitments from institutions such as they NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC).%
\footnote{
http://www.nitrc.org
}
Data from these two institutions were used to demonstrate the alignment properties obtained using SyGN for DTI-derived data normalization.

\paragraph{NKI/Rockland Sample}
Data from the first 14 weeks of prospective data-sharing were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
Due to various technical issues (e.g. lack of corresponding T1 image, failed DTI reconstruction, and inadequate FSL registration performance) only 57 subjects of the NKI data set were used in the image alignment assessment.  
Germane to our evaluation, each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).  Images were anonymized including defacing of the T1 images prior to upload.

\paragraph{NYU Institute for Pediatric Neuroscience Sample}
Data from the first 3 parts of prospective data-sharing from NYU were also downloaded on January 15, 2011 comprised of 49 subjects (average age in years: 30.29 $\pm$ 8.81).  All subjects were used in the image alignment assessment.  Two 64-directional DTI scans were performed (parameters: conventional single-shot spin echo
EPI pulse sequence, TR = 5200 ms, TE = 78 ms, axial acquisition, 
voxel size = $3 \times 3 \times 3$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2530 ms, TE = 3.25 ms, flip angle = 7$^\circ$, 256 contiguous slices of 1.33 mm, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1.3 mm$^3$).  Images were anonymized including defacing of the T1 images prior to upload.

%
%The 1000 Functional Connectomes Project (FCP)%
%\footnote{
%http://fcon\_1000.projects.nitrc.org/index.html
%}
%
%{\color{red}{Quote}}  INDI is a next-generation data-sharing effort, with the goal of sharing phenotypically rich imaging datasets, on both retrospective and prospective bases. On 10/8/10, the Nathan Kline Institute (NKI) / Rockland Sample led the way for INDI-Prospective, starting weekly uploads of 5-7 subjects / week (R-fMRI, DTI, and extensive psychiatric and cognitive phenotypic data, including IQ), which continue to progress (currently n = 55 and growing!). More recently, the NYU Institute Pediatric Neuroscience released its sample examining cocaine dependence in adults via INDI-Retro. The list of sites pledging data from around the world is growing, and we are working to generate common phenotypic variables across datasets as possible (notice the frequency with which IQ is starting to appear Ð itÕs just one of our targets!). 
%
%\subsubsection{Nathan Kline Institute (NKI) / Rockland Sample}
%
%{\color{red}{Quote}}  The NKI/Rockland Sample is intended to be a phenotypically rich neuroimaging sample, consisting of data obtained from individuals between the ages of 6 and 85 year-old. All individuals to be included in the sample undergo semi-structured diagnostic psychiatric interviews, and complete a battery of psychiatric, cognitive and behavioral assessments in order to provide comprehensive phenotypic information for the purpose of exploring brain/behavior relationships.
%
%Each subject consists of the following data:
%\begin{itemize}
%\item 10 minute resting state fMRI scan (R-fMRI) 
%\item 6-direction diffusion tensor imaging (DTI) scan 
%\item 64-direction diffusion tensor imaging scan (2mm isotropic)
%\item MPRAGE anatomical scan
%\item A variety of psychiatric, cognitive and behavioral assessments
%\end{itemize}
%
%\subsubsection{NYU Institute for Pediatric Neuroscience}
%
%{\color{red}{Quote}}  Summary: The NYU Phyllis Green and Randolph C?wen Institute for Pediatric Neuroscience (IPN) sample will be a collection of past and present scans obtained from psychiatrically screened individuals ranging in age from 6 to 55 years old. The initial release consists of datasets from 49 psychiatrically neurotypical adults, with age, gender and intelligence quotient (IQ) information provided. Future releases will include more comprehensive phenotypic information, and child and adolescent datasets, as well as individuals from clinical populations.
%
%Each subject consists of the following data:
%\begin{itemize}
%\item At least one 6-minute resting state fMRI scan (R-fMRI)%
%\footnote{
% Most participants have 2 R-fMRI scans, collected less than 1 hour apart in the same scanning session. Rest$_1$ is always collected first.
%}
%\item One high-resolution T1-weighted mprage, defaced to protect patient confidentiality
%\item Two 64-direction diffusion tensor imaging scans
%\item Demographic information (age, gender) and IQ-measures (Verbal, Performance, and Composite; Weschler Abbreviated Scale of Intelligence [WASI])
%\end{itemize}
%
\subsubsection{Diffuse Traumatic Brain Injury Data}

A third data set was obtained as port of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in chronic survivors of TBI (principal investigator: John Whyte). A more detailed data description can be found in \cite{Avants2008}. The following is a brief description relevant to our particular study.  
17 controls and 16 patients with TBI were used for the analysis presented in this
work.  Each patient had a history of non-penetrating traumatic brain injury 
of at least moderate severity defined by significant and well-documented
loss or alteration of consciousness following injury in addition to meeting
several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and
years of education.
High resolution T1-weighted anatomic images were obtained using 3D MP-RAGE 
imaging sequence using the following acquisition parameters: TR = 1620 ms, 
TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 
1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 
1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
DTI images were also obtained

%
%\subsubsection{Participants}
%{\color{red}{Quote:}} The data were collected as part of a larger study investigating the neural correlates of attention deficits and treatment responses of various psychoactive drugs in the survivors of TBI (principal investigator: J.W.). Thirty individuals with TBI and 20 healthy volunteers were recruited. We planned to recruit more TBI participants because data of TBI survivors are more likely to be discarded in a typical functional neuroimaging study due to movements in the scanner and poor behavioral performance. TBI participants were recruited from a variety of clinical services at MossRehab and through a consent-based registry of individuals with TBI who are interested in participating in rehabilitation research. To be included, participants had to be between the ages of 16 and 60, and to have a history of non-penetrating traumatic brain injury of at least moderate severity at least 3 months prior to enrollment. Severity level was defined by significant and well-documented loss or alteration of consciousness following injury (i.e., lowest Glasgow Coma Scale (GCS) score of less than 12, or prospectively documented post-traumatic amnesia (PTA) of greater than 1 hour), or focal abnormality on a neuroimaging study that was attributable to traumatic injury. A subjective complaint of attention difficulties by the participant, treating clinician, or caregiver was also required. Potential participants were excluded if they had a history of prem orbid neurologic disease, psychosis, major affective disorder, mental retardation, Attention Deficit Hyperactivity Disorder, or if they were currently abusing alcohol or recreational drugs. Persons who were taking psychoactive medications other than anticonvulsants were also excluded. Participants and/or their involved caregivers (depending on the participantÕs cognitive capacity) provided informed consent. The study protocol was approved by the Albert Einstein Healthcare Network and the University of Pennsylvania IRBs. Twenty healthy volunteers, matched to patients for age, gender, handedness, years of education, and ethnicity, participated in the study. Control participants were recruited based on the same inclusion/exclusion criteria as patients, with the exception that they never had a TBI resulting in loss or alteration of consciousness, nor suffered attention complaints. Control participants were recruited through the family and friendship networks of the participants with TBI, and through public advertising.
%Data from one TBI participant whose MRI scan showed a large area of encephalom alacia over almost the entire right hemisphere were excluded. The remaining 29 subjects with TBI included 21 men and 8 women aged between 18 and 58 years (mean age = 36.9, SD = 11.4) with a mean education of 13.1 years (SD = 2.8). Twenty three of them were right-handed (Edinburgh Handedness Inventory, Oldfield, 1971). Fourteen of them were Caucasians, 10 African Americans, 4 Hispanics, and 1 Asian. Selected demographic and clinical characteristics of the TBI survivors are reported in Table 1. Control participants included 17 men and 3 women aged between 21 and 50 years (mean age = 34.9, SD = 9.8) with a mean education of 13.1 years (SD = 1.7). Sixteen of them were right-handed. Eleven of them were Caucasians, 7 African Americans, 1 Asian, and 1 unknown. The two groups did not differ significantly in terms of age, gender, ethnicity, handedness or years of education (tested with t-test or FisherÕs exact test, as appropriate).
%
%\subsubsection{Image acquisition}
%The functional imaging was conducted on a Siemens 3.0 T Trio whole-body scanner (Siemens AG, Erlangen, Germany), using a standard Transmit/Receive head coil. High resolution T1-weighted anatomic images were obtained using 3D M PRAGE imaging sequence using the following acquisition parameters: TR = 1620 ms, TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 1NEX with a scan time of 6 minutes. The resulting voxel size was 1 mm$^3$.
%Lesion assessment
%To quantify the volume of lesions, a trained observer (J.P.) manually segmented the lesion area under supervision of a neurologist (H.B.C.) with extensive experience in lesion assessment. Focal lesions included any cystic cavities and other focal regions of abnormal signal in the white or gray matter. The ITK-SNAP software \citep{Yushkevich2006} was used for a 3D-based segmentation. Total lesion volume was calculated using a stand-alone utility provided by VoxBo software (Center for Functional Neuroimaging, Philadelphia, PA, http://www.voxbo.org).

\section{Results and Discussion}
%Results should be clear and concise.

\subsection{Alignment Assessment of the INDI Data}

\begin{table}
  \caption{Volume (in voxels) of thresholded mean FA skeletons.}
  \label{table:vol}
  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
  \hline
  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
  \cline{2-5}
  {} & inner & outer & inner & outer \\
  \hline
  FMRIB & 84899 & 122306 & 72765 & 76769 \\
  ANTs$_5$ & 87572 & 134421 & 73650 & 94177 \\
  ANTs$_{10}$ & 86908 & 130569 & 73419 & 98014 \\
  ANTs$_{20}$ & 86523 & 129729 & 73577 & 95484 \\
  ANTs$_{30}$ & 86971 & 128892 & 73443 & 94862 \\
  \hline
  \hline
  \end{tabular*}
\end{table}

\begin{table}
  \caption{Number of connected components of thresholded mean FA skeletons.}
  \label{table:cc}
  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
  \hline
  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
  \cline{2-5}
  {} & $\geq$ 1 voxel & $\geq$ 2 voxels & $\geq$ 1 voxel & $\geq$ 2 voxels \\
  \hline
  FMRIB & 4340 & 1802 & 2431 & 993 \\
  ANTs$_5$ & 3561 & 1384 & 2278 & 916 \\
  ANTs$_{10}$ & 3413 & 1302 & 2232 & 899 \\
  ANTs$_{20}$ & 3283 & 1252 & 2093 & 885 \\
  ANTs$_{30}$ & 3114 & 1243 & 2128 & 867 \\
  \hline
  \hline
  \end{tabular*}
\end{table}

Since our proposed modification of TBSS impacts strictly the
critical alignment step, the first set of evaluations demonstrates
the difference between the two methods, i.e. spatial normalization 
via registration of the FA maps to the FMRIB58 template
or spatial normalization via ANTs template construction.  It is important to note that we did not consider normalization either to a random subject
or to the subject best representative of the mean as defined in 
\cite{Smith2006}. The former was not investigated due to the possible bias
in selecting a random subject as described in \cite{Smith2006}.  
Neither was the latter option pursued due to the computational complexity
required for such an $\OO{n^2}$ search over the relatively large NKI and NYU data sets.  In addition, normalization to
the FMRIB58 template is both the default and recommended option
with the parameters of FNIRT optimally tuned for said option.  To warp the MD images following FA registration we used FSL's {\bf applywarp} function using the FA-derived transform.

%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{NYU.pdf}
%\end{tabular}
%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.  (a) The standard FMRIB58\_FA\_1mm template. (b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).  
%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).  Note the visual alignment difference particularly in the peripheral white matter complex.
%}
%\label{fig:NYU}
%\end{figure}

To explore alignment with ANTs-TBSS, 4 templates were constructed for each of the 2 data sets.  Each within-group template was generated from $\{5,10,20,30\}$ randomly selected subjects to see if the number of subjects comprising the template affected the alignment performance. In addition
to the composite transforms described in Section \ref{sec:ANTs-TBSS} to take each subject's FA or MD image to the population-specific anatomical template, we also registered
each ANTs T1 template to the MNI152 template.  This permitted a
standard space for qualitative and quantitative comparison.

Sample axial slices for all resulting FA mean images for the NKI data set are provided in Figure \ref{fig:NKI}.  We also include the same axial slice of the FMRIB58 template in Figure \ref{fig:NKI} (a) for comparison.  In both sets of sample slices, visual inspection would seemingly indicate better performance with the ANTs-based FA mean images compared with the standard TBSS-produced FA mean image.  Also, there is little discernible difference between the different ANTs results.  Alignment differences between the two approaches are particularly apparent in the peripheral white matter where the anatomical structure in the ANTs template-based FA mean images is visually differentiable.

Relative performance is quantified in the plots produced in Figure \ref{fig:NKIplots} and Figure \ref{fig:NYUplots}.  For both data sets, the mean FA image derived from the standard TBSS approach was thresholded at FA $\geq 0.2$.  These two binary masks obtained in both the NKI and NYU cohorts provided the standard group masks for fair comparisons for all evaluations.    For the aligned images the coefficient of variation (CoV) was calculated in a voxel-wise fashion within the mask.  Intuitively, one would expect a smaller variance across better aligned sets of voxel values thus producing a lower CoV for a given mean value.  By calculating the cumulative distribution function (CDF) of the set of CoV values within the mask, quantitative assessment of alignment performance can be visualized.  In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.  This division was accomplished by applying a distance transform \cite{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.  Note that only those voxels that were also within the thresholded FA mask were considered in the inner/outer analysis.

There is a visible improvement in alignment for both data sets using the ANTs-based approach over individual subject FA registration to the FMRIB template.  There is little difference between the ANTs results with only a slight trend in improved performance with the larger template subject sample sizes.  This was the case for all three regional CoV calculations for both data sets.  Also, as expected, there is greater misalignment in the outer region versus the inner core of the white matter mask regardless of methodology.  However, the performance tally was the same in both the inner and outer regions as was the case with the entire white matter mask although the disparity in performance was much less for the NYU data set as the NKI data set most likely due to the differences in resolution of the original DTI.

The improvements seen in the CDF plots are also supported by comparative quantities derived from the skeletons of the different mean FA images.  Since the skeletonization process is meant to determine the center of the various white matter tracts \cite{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.  In addition, as described in \cite{Zalesky2011}, the number of connected components of the white matter skeleton is directly related to the contiguity and thus inversely related to the level of structure.  Table \ref{table:vol} provides the volume of the mean FA skeletons (thresholded at FA $\geq 0.2$) for standard TBSS protocol using the FMRIB template and the ANTs approach using the various templates.  In agreement with the CDF plots illustrating improved alignment with ANTs, the voxel volumes of the skeletons are greater using ANTs templates with the greatest difference seen in the outer portion of the white matter.
Related, in Table \ref{table:cc}, we provide the number of connected components for both data sets for all isolated components greater in size than 1 voxel and 2 voxels.  Even though the skeletal volumes associated with the ANTs approach are greater, the number of connected components is less which would also support the observation of improved alignment.

%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{inner.png}
%\end{tabular}
%\caption{
%}
%\label{fig:inner}
%\end{figure}
%
%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{outer.png}
%\end{tabular}
%\caption{Three-canonical view of the inner
%}
%\label{fig:outer}
%\end{figure}

\begin{figure*}
\begin{tabular}{c}
  \includegraphics[width=180mm]{NKIplots2.pdf}
\end{tabular}
\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Figure \ref{fig:NKI}(b)).  In addition to the whole brain analysis (first column), the coefficient of variation was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).  The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
}
\label{fig:NKIplots}
\end{figure*}

\begin{figure*}
\begin{tabular}{c}
  \includegraphics[width=180mm]{NYUplots2.pdf}
\end{tabular}
\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Figure \ref{fig:NKI}(b)).  In addition to the whole brain analysis (first column), the coefficient of variation was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).  The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
}
\label{fig:NYUplots}
\end{figure*}

\subsection{TBSS and ANTs-TBSS of TBI Cohort}

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=80mm]{meanFA.pdf}
\end{tabular}
\caption{Comparison of mean FA images used in TBSS taken from the TBI data described in Section 2.  The mean FA image illustrated in (a) is created by mapping the sample FA population to the default FMRIB58 template which is the standard protocol for TBSS.  In comparison, the mean template represented in (c) is created from alignment of the population sample to the optimal T1 template as described in the paper. The respective skeleton masks in (b) and (d) are created by thresholding the resulting skeleton values $\geq 0.2$.  The yellow area points to a peripheral difference in findings between the two analyses.
}
\label{fig:meanFA}
\end{figure}

A patient template and a control template were  constructed 
using the ANTs software from all subjects of the respective populations.  These two templates were then used to create a third template which provided the standard space for the ANTs-based TBSS assessment.  Standard TBSS was also performed.

The resulting mean FA images and thresholded skeletons are represented in Figure \ref{fig:meanFA} which demonstrate appreciable in alignment.  These images support earlier findings in improved alignment with the ANTs-based approach.  The skeleton masks were produced by thresholding the skeleton values at FA $\geq 0.2$ and it would seem that the improved alignment produces a more extensive skeletal network for greater localization testing of significant differences, particularly in the outer white matter regions.  Given that TBSS is intrinsically a statistical method for finding local differences, improved alignment would presumably provide increased regional sensitivity to differences.

Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.  The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.

%As an additional quantitative comparison, we calculated the volumetric ratio between areas of significance and the volumetric brain mask for both $p = 0.01$ and $p = 0.05$.  For standard TBSS, the MNI 152 "first brain mask" was used and for the ANTs template, Atropos was used to obtain a mask by skull stripping of the final template.  These values are given in Table \ref{table:hoon}.
%
%
%\begin{table}
%  \caption{Volumetric ratio differences between ANTsTBSS and TBSS.}
%  \label{table:hoon}
%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccc}
%  \hline
%  {} & $\frac{V(\mathrm{mean\,FA\,mask})}{V(\mathrm{brain\,mask})}$ & $\frac{V(\mathrm{sig.\,areas})}{V(\mathrm{mean\,FA\,mask})}$ & $\frac{V(\mathrm{sig.\,areas})}{V(\mathrm{brain\,mask})}$ \\
%  \hline
%  $p \leq 0.01$ & {} & {} & {} \\
%  $p \leq 0.05$ & {} & {} & {} \\
%  \hline
%  \hline
%  \end{tabular*}
%\end{table}

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=80mm]{hoonData2.pdf}
\end{tabular}
\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).  Axial slices were chosen based on visual assessment of correspondence in both the ANTs template and MNI 152 template.
On the left is the thresholded skeleton overlaid on the mean FA image.  
On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).  The yellow arrow points to an area of significant difference in the peripheral white matter found using ANTs-TBSS.
}
\label{fig:hoonData}
\end{figure}

%\section{Discussion}
%This should explore the significance of the results of the work, not repeat them. A combined Results and Discussion section is often appropriate. Avoid extensive citations and discussion of published literature.

\section{Conclusions}
%The main conclusions of the study may be presented in a short Conclusions section, which may stand alone or form a subsection of a Discussion or Results and Discussion section.

The popularity of FSL's TBSS is due, in large part, to its ease of use.  Anybody interested in performing a neuroimaging study in which DTI-derived measures are compared across populations can simply download FSL and immediately apply it to their data.  The proposed work is a substantial improvement to the core TBSS framework which dovetails with FSL.  In addition, it is comparably easy to download and apply what we have proposed in the form of ANTs-TBSS.

The proposed improvement potentially has many benefits which were listed in the introduction.  Such benefits range from the theoretical in the form of eliminating the confounding effects of registering data which is to be statistically differentiated to the practical---reusing ANTs template construction for other neuroimaging analyses.
In performing the various analyses described in the work, it was pleasantly surprising to discover how well the alignments were for the ANTs template data considering that alignment was based on FA-to-FA correspondence and no such image registrations were performed to align FA directly with FA as with TBSS.  Furthermore, since TBSS is a local quantification method with increased statistical power through the use of the skeletonization and projection step, improved alignments can increase the regional sensitivity to local white matter differences particularly in the peripheral tracts.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}          ==>>  [#]
%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
%%   \citet{key}         ==>>  Author [#]

%% References with bibTeX database:

\section*{Acknowledgments}
All visualizations were performed using ITK-SNAP%
\footnote{
http://www.itksnap.org/
}
\cite{Yushkevich2006} and
DTI-TK.%
\footnote{
http://www.nitrc.org/projects/dtitk/
}

\section*{References}

\bibliographystyle{elsarticle-harv}
\bibliography{references}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model1-num-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-1-num.tex'.

### 2011-03-24 ANTsTBSS/paper.tex
Minor edits, new figure.

@@
 
-Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.
+%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.
 The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.
@@
 
-\begin{figure}
-\begin{tabular}{c}
-  \includegraphics[width=80mm]{hoonData2.pdf}
-\end{tabular}
-\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).
+%\begin{figure}
+%\begin{tabular}{c}
+%  \includegraphics[width=80mm]{hoonData2.pdf}
+%\end{tabular}
+%\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).
 Axial slices were chosen based on visual assessment of correspondence in both the ANTs template and MNI 152 template.
-On the left is the thresholded skeleton overlaid on the mean FA image.  
-On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).
+%On the left is the thresholded skeleton overlaid on the mean FA image.  
+%On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).
 The yellow arrow points to an area of significant difference in the peripheral white matter found using ANTs-TBSS.
-}
-\label{fig:hoonData}
-\end{figure}
+%}
+%\label{fig:hoonData}
+%\end{figure}
 

### 2011-03-25 ANTsTBSS/paper.tex
Minor edits, new figure.
Including Phil's suggestions.

@@
         of rigidly transformed FA map ($=R^i_{dwi} \mathrm{FA}$) using
-        the white matter mask obtained in step 3 ($T^i_{FA}$).
+        the white matter mask obtained in step 3(a) ($T^i_{FA}$).
   \item For each subject, concatenate the transforms which maps each
@@
 \footnote{
-http://web4.cs.ucl.ac.uk/research/medic/camino/pmwiki/
+http://camino.org.uk/
 }---an open source toolkit for diffusion MRI processing and analysis \cite{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.
@@
 
-A third data set was obtained as port of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in chronic survivors of TBI (principal investigator: John Whyte).
+A third data set was obtained as port of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in long-term survivors of TBI (principal investigator: John Whyte).
 A more detailed data description can be found in \cite{Avants2008}.
@@
 
+%
 %Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.

### 2011-06-06 ANTsTBSS/paper.tex
ENH: Major revision.

@@
 
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% Use the options 1p,two column; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
 %% for a journal layout:
@@
 
-\title{ANTs-Flavored Tract-Based Spatial Statistics}
+\title{Normalization Circularity in Population-Based FA Analysis}
 
@@
 %% Text of abstract
-
-We propose a modification of the popular tract-based spatial statistics (TBSS) framework \citep{Smith2006} for investigating diffusion tensor derived scalar measures across populations.
+Critical to many studies investigating white matter differences across populations 
+is normalization of fractional anisotropy (FA) images to a common coordinate system
+for statistical analysis.  Although variations exist, common to several current 
+approaches is direct FA-to-FA registration whether it be to a chosen individual or
+a generated template.  This type of circular analysis 
+is susceptible to the general confounding 
+effect of minimizing data differences with nonrigid registration which are potentially 
+the precise differences to be statistically quantified.  We demonstrate that
+population normalization to a template explicitly optimizes certain voxelwise statistical 
+testing measures as opposed to maximizing anatomical alignment.  
+As a corrective, we advocate the use of the Symmetric Group Normalization strategy for 
+unbiased template construction from anatomical (e.g. T1-weighted) images which provides 
+the coordinate system for statistical inference of FA differences.
+Following template construction, individual FA images are warped to the template via 
+the corresponding anatomical image.  Performance is evaluated using open data from the International 
+Neuroimaging Data-sharing Initiative and a cohort of 16 traumatic brain injury survivors compared with 
+17 control subjects.  Alignment comparisons with the normalization component of the 
+popular tract-based spatial statistics (TBSS)
+show that our anatomically-based template approach produces good alignments despite
+the absence of any direct FA-to-FA registration.  Of practical consideration we note that the 
+software tools discussed in this work are provided to the public as part of the open source 
+Advanced Normalization Tools (ANTs) repository.
+
+%We propose a modification of the popular tract-based spatial statistics (TBSS) framework \citep{Smith2006} for investigating diffusion tensor derived scalar measures across populations.
 Whereas the current crucial alignment step is performed typically using the derived fractional anisotropy images, we propose using the Symmetric Group Normalization strategy for anatomical unbiased template construction from T1-weighted images which serves as the coordinate system for statistical inference.
@@
 \begin{keyword}
-Advanced Normalization Tools (ANTs) \sep diffusion imaging \sep fractional anisotropy \sep template construction \sep tract-based spatial statistics \sep traumatic brain injury
+Advanced Normalization Tools (ANTs) \sep diffusion tensor imaging \sep fractional anisotropy \sep template construction \sep traumatic brain injury
 %% keywords here, in the form: keyword \sep keyword
@@
 
-Ever since the seminal work of Basser et al.
-\cite{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique, continued quantitative development has furthered its applicability to probing clinical questions of interest particularly pertaining to the microstructure of white matter \cite{Basser1996,Assaf2008}.
+The seminal work of Basser et al.
+\cite{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique.
+Continued quantitative development has furthered its applicability to probing clinical questions of interest particularly pertaining to the microstructure of white matter \cite{Basser1996,Assaf2008}.
 Given this sensitivity to brain architecture, DTI has found multiple uses in assessing neuro-structural differences in cross-population studies \cite{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010}
 including subjects with traumatic brain injury (TBI) \cite{Caeyenberghs2010,Jiang2010,Warner2010}.
-One such statistical framework which has significantly facilitated quantitative analysis of DTI data is the well-known tract-based spatial statistics (TBSS) originally formulated in \cite{Smith2006} and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).%
+
+Two popular comprehensive software packages for assessing population FA differences include 
+the SPM Matlab-based toolkit%
+\footnote{
+http://www.fil.ion.ucl.ac.uk/spm/
+}
+for voxel-based morphometry (VBM)
+and the tract-based spatial statistics (TBSS) framework \cite{Smith2006}.%
 \footnote{
@@
 } 
-Based solely on the number of citations in the literature,%
+Although concerns regarding the general 
+application of VBM \cite{Bookstein2001,Ashburner2001,Davatzikos2004}
+and its application to FA data \cite{Jones2005,Chung2008} has been discussed 
+in the literature, it continues to find application in FA population 
+studies \cite{Kakeda2010,Takao2010}.
+The TBSS framework was partially developed in response to these concerns
+in which voxel values are projected onto a template white matter skeleton for 
+increased statistical power.  
+%(e.g. \cite{Arnone2008,Bodini2009}).
+
+Although there are substantive differences in normalization schemes
+between these and other image registration approaches, a core
+commonality includes direct FA-to-FA registration.
+Potential difficulties associated with direct FA-to-FA normalization is exemplified by 
+a standard population study in which all subject FA images are mapped to a representative 
+template using the popular sum of squared differences (SSD) image similarity metric.%
 \footnote{
-As of the beginning of the year 2011, the original Neuroimage 2006 TBSS paper has been cited more than 360 times making it the most cited in the journal's library.
-}
-usage of TBSS has certainly permeated the neuroimaging community and, in combination with its accessibility and ease of use, has become a de facto quantitative tool for DTI analysis.
+In addition to TBSS and SPM, other popular packages in the neuroimaging community 
+which use the SSD metric includes AIR \cite{Woods1998} and LDDMM \cite{Beg2005}.
+}  
+Given a set of $M$ subject images, $\{I_1, I_2, \ldots, I_M\}$, and
+a template image, $J$, population normalization concerns finding the set of $M$ transformations, 
+$\boldsymbol{\mathcal{T}} = \{\mathcal{T}_1, \mathcal{T}_2,\ldots,\mathcal{T}_M \}$, which minimize the SSD metric over the population, i.e.
+\begin{align}
+  \boldsymbol{\mathcal{T}} = \underset{\mathcal{T}_1, \ldots, \mathcal{T}_M}{\operatorname{argmin}}
+    \sum_{m=1}^M\sum_{n=1}^N \left( I^n_m( \mathcal{T}_m(\mathbf{x}) ) - J \right)^2.
+\end{align}
+Switching the order of the summations (since the transformations are mutually independent) and 
+recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the 
+mean of the aligned images, it becomes apparent that the transformation solution
+\begin{align}\label{eq:variance}
+  \boldsymbol{\mathcal{T}} = \underset{\mathcal{T}_1, \ldots, \mathcal{T}_M}{\operatorname{argmin}}
+    \sum_{n=1}^N \underbrace{\sum_{m=1}^M \left( I^n_m( \mathcal{T}_m(\mathbf{x}) ) - J \right)^2}_{
+    \propto\mathrm{\,voxelwise\,variance}}
+\end{align}
+is that which minimizes the voxelwise variance or, equivalently,
+maximizes various voxelwise statistical quantities including the common Student's $t$-test.%
+Note that many of the standard similarity metrics are also going to exhibit this
+effect to some degree as they minimize some local measure of intensity-based distance 
+between the subject and the template.
+
+Thus, instead of normalization based on anatomical alignment followed by
+statistical testing, direct FA normalization 
+explicitly conflates anatomical alignment with optimizing certain statistical testing results
+evoking a sense of circularity in the analysis \cite{Kriegeskorte2010}.
+According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on
+data that were selected for showing the effect of interest or a related effect.''  As illustrated with
+Eqn. \ref{eq:variance}, direct FA-to-FA registration causes voxels to be matched (or selected) according to the
+the statistical testing result (or the effect of interest) instead of anatomical alignment.
+%This results in a significant increase in the number of false positives for this type
+%of testing which we illustrate with the following example.
+
+\paragraph{Simulated Data Example}
+
+\begin{figure}
+\begin{tabular}{c}
+  \includegraphics[width=85mm]{simulatedData.pdf}
+\end{tabular}
+\caption{(a) Axial slice from the discrete BrainWeb data set.  From (a) 
+we create the (b) simulated template and sample (c) control and (d) patient images with both 
+additive noise and subsequent Gaussian smoothing. }
+\label{fig:simulated_data}
+\end{figure}
+
+\begin{figure}
+\begin{tabular}{c}
+  \includegraphics[width=85mm]{simulatedDataWarped.pdf}
+\end{tabular}
+\caption{Warping the image in Fig. \ref{fig:simulated_data}(c) to the template in 
+  Fig. \ref{fig:simulated_data}(b) using the SSD similarity metric produces the image shown in
+  (a).  The corresponding deformation is given in (b).  Although the images were anatomically
+  aligned prior to invoking ANTs registration, deformation is driven by the similarity metric
+  causing the accumulation of type 1 errors.}
+\label{fig:simulated_data_warped}
+\end{figure}
+
+Axial slice number 159 was extracted from \verb#subject_04# of the discrete BrainWeb database
+\cite{Aubert-Broche2006} which is shown in Fig. \ref{fig:simulated_data}.  After thresholding out 
+the white matter and assigning all white matter voxels a value of 0.7, 10\% of these white 
+matter voxels were randomly selected to be assigned a value of 0.6 in the simulated control 
+template and a value of 0.3 in the simulated patient template.  
+Note that these values were arbitrarily selected to be more or less in the range of typical
+FA values.  Gaussian noise (mean = 0.0, standard deviation = 0.1) was added to the white
+matter template followed by a Gaussian smoothing (standard deviation = 1.5 mm) shown in Fig. \ref{fig:simulated_data}(b).  A simulated control cohort was created consisting of 20 subjects.  
+Starting with the simulated control template, each of the 20 subjects is created by adding
+Gaussian noise (mean = 0.0, standard deviation = 0.1) followed by 
+Gaussian smoothing (standard deviation = 1.5 mm).  A simulated patient cohort of 20 subjects
+was created similarly using the simulated patient template.  A sample control subject and sample
+patient subject are given, respectively, in Fig. \ref{fig:simulated_data}(c) and (d).
 
 \begin{figure*}
+\begin{center}
+\begin{tabular}{c}
+  \includegraphics[width=180mm]{simulatedTesting.pdf}
+\end{tabular}
+\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the simulated experiment adjusted for multiple comparisons (FDR).
+The true significance curves are drawn in black.  Elevated statistical power is seen with all 4 
+         metrics for both tests although the effect is less in the more robust Wilcoxon results  Not shown are the significance curves for the Brunner-Munzel testing \cite{Rorden2007} which showed similar effects to the results on the left.
+        }
+\label{fig:simulated_testing}
+\end{center}        
+\end{figure*}
+
+After creation of the simulated data set, we performed voxelwise testing (both $t$-testing
+and Wilcoxon rank-sum testing, one-tailed, $\alpha = 0.95$) with false discovery
+rate correction within the white matter to get ``true'' statistical significance 
+curves.  These curves are plotted in black in Fig. \ref{fig:simulated_testing}.  To illustrate the 
+effect of similarity metric on statistical significance, we performed ANTs registration
+\cite{Avants2011} of each subject to the template
+(Fig. \ref{fig:simulated_data}(b)) using cross correlation (CC), mutual information (MI), 
+Demons \cite{Thirion1998}, and sum of squared differences (SSD). A representative deformed image
+using the SSD metric and its corresponding deformed grid is given in 
+Fig. \ref{fig:simulated_data_warped}. 
+The resulting statistical significance curves are plotted with the true significance curves in Fig. 
+\ref{fig:simulated_testing}.
+
+This example demonstrates the undesirable phenomenon previously described.  Even though these
+images were anatomically aligned prior to registration, the resulting deformations and 
+increased false statistical power can be explained in the context of Eqn. (\ref{eq:variance}).  
+Common similarity metric functions can induce false positives by minimizing local intensity 
+differences between the subjects and template which explicitly maximizes statistical
+testing results in contrast to maximizing anatomical alignment.  This effect is strongest in 
+the SSD and Demons metrics and weakest in the MI and CC metrics likely due
+to the degree of correspondence to Eqn. (\ref{eq:variance}).   This is also supported by contrasting
+the $t$-test results given in Fig. \ref{fig:simulated_testing}(a) with the non-parametric Wilcoxon
+rank-sum testing (Fig. \ref{fig:simulated_testing}(b)) where as large of an effect for the latter would not be expected although an effect is
+still apparent.
+
+Although various processing choices mitigate these effects, such as Gaussian smoothing, nonparametric
+testing, and
+the use of more robust similarity metrics, such choices are often made on an ad hoc basis.  
+In contrast, in this work we strongly advocate the general application
+%A more principled approach is the use of an anatomical, i.e. T1-weighted, unbiased shape/intensity 
+%template constructed from the population images.  DTI-derived scalar images can then be mapped 
+%to their corresponding T1-weighted image in a highly constrained manner and subsequently 
+%mapped to the template via nonrigid registration of the T1-weighted image.  This minimizes 
+%possible complications associated with direct FA-to-FA image registration.
+%We recognize that using the anatomical images in the alignment process has been previously proposed (in fact, such a suggestion occurs in the {\it Future directions} section of the original TBSS paper).
+of a high performance \cite{Klein2009}, nonrigid registration algorithm \cite{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \cite{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
+
+\begin{figure*}
+\begin{center}
 \begin{tabular}{c}
@@
 \end{tabular}
-\caption{Proposed ANTs-TBSS workflow.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
+\caption{Possible ANTs-based TBSS workflow.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
 (transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.
-The FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.
-Each individual T1 is then nonrigidly registered to the anatomical template.
+Each FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.
+The individual T1-weighted image is then nonrigidly registered to the anatomical template.
 Optionally, one can register the T1 template to the MNI152 template which resides in the same space as the FMRIB58 template.
@@
 \label{fig:workflow}
+\end{center}
 \end{figure*}
 
-Prior to the advent of TBSS, voxel-based approaches, such as statistical parametric mapping \cite{Ashburner2000}, were used in assessing voxel-wise population differences \cite{Rugg-Gunn2001,Watkins2001,Kubicki2005,Park2004} using DTI data.
-In addition to concerns regarding the general application of SPM \cite{Bookstein2001,Ashburner2001,Davatzikos2004}, particularly the requirement for stringent spatial normalization,
-issues specific to the application to DTI-derived measures, such as the effects of size variation of the smoothing kernel and the potential non-normality of the residuals \cite{Jones2005}, encourage caution in the application of SPM to DTI-derived data.
-Although the latter can be addressed via the use of non-parametric testing \cite{Chung2008}, addressing the former in a principled way is elusive.
-
-It was precisely due to these concerns that the TBSS framework was developed constituting an entire processing pipeline spanning from the reconstruction of the DTI tensorial data
-%to the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA)), to the nonrigid alignment of the population to a standard coordinate system and skeletonization, and, finally,
-through to the statistical analysis for discerning differences across populations.
+We first briefly describe the SyGN algorithm used to create anatomical templates.  
+We then explain the the step-by-step process for aligning subject FA images to 
+the anatomical template (which is facilitated with the various tools available 
+in ANTs).  After describing the real image data used in this work, we show the 
+issues raised affect common analysis scenarios.
+
+%
+%
+%
+%It was precisely due to these concerns that the TBSS framework was developed constituting an entire processing pipeline spanning from the reconstruction of the DTI tensorial data
+%%to the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA)), to the nonrigid alignment of the population to a standard coordinate system and skeletonization, and, finally,
+%through to the statistical analysis for discerning differences across populations.
 Major intermediate steps include the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA) \cite{Basser1996}), nonrigid alignment of the population to a standard coordinate system, and subsequent ``skeletonization'' of the mean of the aligned FA maps.
@@
 
-Because spatial normalization is critical to obtaining valid results, we propose a modification of the standard TBSS pipeline directed at some of these residual concerns which we refer to as ANTs-TBSS.
+%Because spatial normalization is critical to obtaining valid results, we propose a modification of the standard TBSS pipeline directed at some of these residual concerns which we refer to as ANTs-TBSS.
 Instead of DTI-derived scalar alignment via the images themselves (either random subject selection or locating an approximation to the ``mean subject'') or to a standard template (e.g.
@@
 
-We recognize that using the anatomical images in the alignment process has been previously proposed (in fact, such a suggestion occurs in the {\it Future directions} section of the original TBSS paper).
-Our contribution stems from the specific proposal of using a high performance \cite{Klein2009}, nonrigid registration algorithm \cite{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \cite{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
-This results in a few advantages in using ANTs-TBSS vs standard TBSS:
-\begin{itemize}
-\item As explained in \cite{Avants2010}, ANTs optimal template construction creates a true Frech\'et mean from the population sample with high resolution which facilitates within-group registration.
-This obviates the need for an externally-derived template or an $\OO{n^2}$ search for an approximation to the mean subject or the bias resulting from random subject selection.
-\item Since the creation of the standardized coordinate system is performed using the anatomical images, nonrigid registration of the DTI-derived scalar images is 
-minimized thus mitigating the confounds described previously.
-Each DTI-based image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \cite{Wu2008}.
-\item Although ANTs-TBSS increases the total number of registrations to be performed, this makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
-\item Instead of Gaussian smoothing as with SPM, TBSS uses the warping process to FMRIB58 by default to provide an indirect smoothing.
-Such resampling occurs in ANTs-TBSS as a by-product of the registration of the DTI-derived image to the corresponding T1 image in a subject-specific manner.
-%\item Instead of using thresholding to create the white matter segmentation from which the skeleton is created, an anatomically-based segmentation, either in the template space or in each subject space followed by a label fusion technique (e.g.
-majority voting), can be used to ultimately provide a more accurate skeleton.
-\item Oftentimes TBSS is only one of several neuroimage analyses performed in a population-based study.
-Other investigations, such as voxel- and tensor-based morphometry, cortical thickness measurements, etc.
-can take advantage of the anatomical template used in ANTs-TBSS.
-\end{itemize}
-
-Of practical significance, we note that all software components described in this paper, similar to FMRIB's software library (FSL), of which TBSS is a component, are publicly available as open source primarily through the Advanced Normalization Tools (ANTs) repository.%
-\footnote{
-http://www.picsl.upenn.edu/ANTs
-}
-In addition to registration and template building, ANTs also includes brain extraction \cite{Avants2010a}, bias correction \cite{Tustison2010}, $n$-tissue segmentation \cite{Avants2011a}, and cortical thickness estimation solutions \cite{Das2009}.
-%As part of this work, we also explicitly show how one can merge the ANTs and TBSS components into a single modified ANTs-flavored TBSS pipeline.
-\begin{figure}
-\begin{tabular}{c}
-  \includegraphics[width=80mm]{NKI.pdf}
-\end{tabular}
-\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.
-(a) The standard FMRIB58 template.
-(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
-(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).
-Note the visual alignment difference particularly in the peripheral white matter complex.
-}
-\label{fig:NKI}
-\end{figure}
-
-We first describe in greater detail the proposed ANTs-TBSS framework with an emphasis on comparison with standard TBSS.
+%We first describe in greater detail the proposed ANTs-TBSS framework with an emphasis on comparison with standard TBSS.
 We then describe the two open data sets used to evaluate the proposed methodology in addition to a description of diffuse TBI data
-also used to illustrate concepts discussed in this work.
+%also used to illustrate concepts discussed in this work.
 %and analysis previously reported in \cite{Avants2008}.   
@@
 \section{Material and Methods}
+
+%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.
 
@@
 
-\subsection{Tract-Based Spatial Statistics}
-
-FSL hosts a large number of excellent algorithms for computational neuroanalysis.
-These include, among others, well-known tools for image segmentation (FAST---\cite{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), as well as TBSS.
-For the comparative evaluation studies used in this manuscript, TBSSv1.2%
-\footnote{
-Build 416 downloaded from the FMRIB web site http://www.fmrib.ox.ac.uk/fsldownloads/ on November 18, 2010.
-}
-was used.  The actual code pipeline constituting TBSS prior to statistical analysis consists of the following 4 shell scripts:
-\begin{description}
-  \item[tbss\_1\_preproc]  Input is a directory of FA images which are slightly eroded and then organized into a separate FA directory.
-A mask is also generated for each input FA image for use during the spatial normalization process.
-  \item[tbss\_2\_reg]  The organized FA data and corresponding masks produced in the previous step are warped to one of three possible coordinate systems using FLIRT/ FNIRT:  1)  the atlas included with the FSL download---FMRIB58---(recommended), 2) a user-specified target image or 3) an exhaustive search for the ``optimal'' target image.
-In \cite{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.
-However, even though it would be faster to simply choose a single subject at random, the complexities of nonrigid image registration would seemingly promote the exhaustive search over random selection.
-As a compromise between these two options, the first option in the script is to use FMRIB58 which also resides in the space of the MNI152 T1 template.
-It should also be noted that this particular script invokes another FSL script, {\bf fsl\_reg}.
-An option associated with {\bf fsl\_reg} permits the use of an FNIRT configuration file which has been optimized for use with the FMRIB58 atlas.
-  \item[tbss\_3\_postreg] The mean FA image is generated from the aligned FA images which is subsequently skeletonized.
-  \item[tbss\_4\_prestats] The skeleton mask is created by thresholding the mean FA skeleton.  The skeleton distance map is then generated and all FA data are projected onto the skeleton.
-\end{description}
-
-\subsection{ANTs-Flavored TBSS}
-\label{sec:ANTs-TBSS}
-
-Our proposed modification to standard TBSS is illustrated in Figure \ref{fig:workflow}.  Instead of direct FA-to-FA registration, our ANTs-based modification can be described as follows:
+\subsection{Anatomical Image Template Construction}
+
+An important step in reducing intersubject variability is
+normalization to a standardized coordinate system for
+cross-population comparison.  Various approaches
+exist for determination of the normalized space such as selection of a
+pre-existing template based on a single subject, e.g. the Talairach atlas \cite{Talairach1988}), or 
+a publicly available averaged group of subjects, e.g. the MNI \cite{Collins1994} or 
+ICBM \cite{Mazziotta1995} templates.
+
+Recent trends, however, indicate a growing popularity of generating population-specific 
+templates for minimizing potential confounds.  This includes the
+diffeomorphic approach of Joshi et al. who employ the SSD metric and a shape 
+distance to bring the subject group of images into alignment \cite{Joshi2004}.  Variants
+include extension to multiple modalities \cite{Lorenzen2006} and small deformations
+\cite{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
+towards a representative image template \cite{Learned-Miller2006}.
+
+In contrast, Avants et al. explicitly model the geometric component of the 
+normalized space during optimization.  Coupling the intrinsic symmetry of 
+SyN pairwise registration \cite{Avants2011} and an
+optimized sharpening/averaging of the template appearance, Symmetric Group
+Normalization (SyGN) is a powerful framework for producing optimal population-specific
+templates \cite{Avants2010}. Given a set of representative images, 
+$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
+diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
+\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
+the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
+which minimize the following cost function:
+\begin{align}
+  \sum_{m=1}^M \left[
+    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
+    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
+    \right]
+\end{align}
+where $D$ is the diffeomorphic shape distance, 
+\begin{align}
+  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
+\end{align}
+dependent upon the choice of the linear operator, $L$, and
+$v$ is the diffeomorphism-generating velocity field, 
+\begin{align}
+  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
+\end{align}
+$\Pi$ is the choice of similarity metric, often cross-correlation \cite{Avants2008a}, calculated in the 
+virtual domain midway between each individual image and the current estimate of the template.
+
+With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
+to identity, iterative optimization
+involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
+updating of $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
+%For the interested reader, a set of scripts has been created and is distributed with the ANTs library for facilitating template creation.
+
+\subsection{Anatomically-Based Population Normalization}
+
+An example of our proposed workflow applied to anatomically-based population normalization prior to TBSS analysis  is illustrated in Fig.
+\ref{fig:workflow}.
+Instead of direct FA-to-FA registration, our ANTs-based modification can be described as follows:
 
 \begin{enumerate}
-  \item Build the anatomical template from  the population sample or sample subset using SyGN.
-  \item If only using a sample subset for template creation, warp the remaining
+  \item Build the anatomical template from  the population sample or sample subset using SyGN as described in
+  the previous section.
+  \item If only using a sample subset for template creation, register the remaining
         anatomical data to the template and save the resulting transformation for each         
-        subject indexed by $i$ ($T^i_{template}$).
+        subject indexed by $i$ (i.e. $\mathcal{T}^i_{template}$).
   \item Find the optimal transformation from each
@@
   \begin{enumerate}
-  \item For each anatomical image, segment the white matter (in ANTs one can use N4 bias correction and Atropos $n$-tissue segmentation).
+  \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \cite{Tustison2010} and Atropos $n$-tissue segmentation \cite{Avants2011a}).
   \item Rigidly transform each subject's average DWI to the bias-corrected anatomical
-        image ($R^i_{dwi}$).
+        image ($\mathcal{R}^i_{dwi}$).
   \item Perform distortion correction on the FA to anatomical mapping 
         by localized nonrigid registration
-        of rigidly transformed FA map ($=R^i_{dwi} \mathrm{FA}$) using
-        the white matter mask obtained in step 3(a) ($T^i_{FA}$).
+        of rigidly transformed FA map ($=\mathcal{R}^i_{dwi} \mathrm{FA}$) using
+        the white matter mask obtained in step 3(a) ($\mathcal{T}^i_{FA}$).
   \item For each subject, concatenate the transforms which maps each
-        subject's FA to the template ($T^i_{total} = T^i_{template} \circ 
-        T^i_{FA} \circ R^i_{dwi} \mathrm{FA}$).
-  \item Warp each subject's FA or MD map to the template using $T^i_{total}$.
+        subject's FA to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
+        \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
+  \item Warp each subject's FA or MD map to the template using $\mathcal{T}^i_{total}$.
   \end{enumerate}      
-  \item Perform the remaining portions of the standard TBSS pipeline:
-    \begin{enumerate}
-      \item Create the mean FA image from all template-aligned FA images.
-      \item Produce the white matter skeleton from the mean FA image.
-      \item Create skeleton distance map and project all FA data onto the skeleton.
-      \item Run the statistical analysis.
+%  \item Perform the remaining portions of the standard TBSS pipeline:
+%    \begin{enumerate}
+%      \item Create the mean FA image from all template-aligned FA images.
+%      \item Produce the white matter skeleton from the mean FA image.
+%      \item Create skeleton distance map and project all FA data onto the skeleton.
+%      \item Run the statistical analysis.
+%    \end{enumerate}
 \end{enumerate}
-\end{enumerate}
-
-%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.
-
-%\subsection{Symmetric Group Normalization Template Creation}
-
-%Due to the nonlinearity inherent in morphometric neuromaging studies, choice of coordinate
-%system can affect quantitative outcome \citep{Thompson2000,Senjen2005}.  
-%This dependence has motivated significant work in defining coordinate systems
-%using population-derived templates \cite{Joshi2004,Lorenzen2006,Avants2010}. 
-%
-%based on \cite{Klein2009}
-%symmetric group-wise normalization scheme for template co
-%
-%nstruction was presented in \cite{Avants2010}.  
-%
-%
-%\subsection{Benefits of our approach}
-%\begin{itemize}
-%   \item White matter segmentations can be performed on the T1 template 
-%\end{itemize}
-%
-%
-%\subsection{ANTs-Flavored TBSS}
-
-%\subsection{ANTs-Flavored Tract-Based Spatial Statistics}
-%
-%\begin{enumerate}
-%  \item Create the anatomical template from sample T1 images, i.e.
-%\begin{command}
-%\lstsetcpp
-%\begin{lstlisting}
-%$ buildtemplateparallel.sh -d 3 -o T_ -c 0 -g 0.25 -i 4 \
-%    -m 50x50x20 -n 1 -r 1 -s CC -t GR T1*.nii.gz
-%\end{lstlisting}
-%\end{command} 
-%\newline
-%  where the following flags are used:
-%  \begin{description}
-%    \item[-d] is the image dimension,
-%    \item[-o] is the output prefix where the final template will be named \verb#T_template.nii.gz#,
-%    \item[-c] specifies the computational parallelization method,
-%    \item[-g] is the gradient step size, 
-%    \item[-i] is the number of template-building iterations,
-%    \item[-m] is the maximum number of iterations per pairwise registration,
-%    \item[-n] is a boolean flag denoting whether or not N4 bias correction is used,
-%    \item[-r] is a boolean flag denoting whether or not a rigid registration is used to initialize the template,
-%    \item[-s] specifies the similarity metric used (\verb#CC# = cross correlation), and
-%    \item[-t] specifies the transformation model used (\verb#GR# = greedy SyN).
-%  \end{description}
-%The list of files \verb#T1*.nii.gz# denote the set of T1 images used to create the template.
-The output consists of the template and the transforms mapping between the subject and template.
-Although these parameters generally work quite well, additional information can be found in the help menu of \verb#buildtemplateparallel.sh#.
-%\item For each subject, use N4 and Atropos to segment the white matter in each subject's T1 following skull stripping (options include BET \cite{Smith2002} or Atropos \cite{Avants2010a}).  
-%\begin{command}
-%\lstsetcpp
-%\begin{lstlisting}
-%$ N4BiasFieldCorrection -d 3 -i T1_subX_brain.nii.gz \
-%    -b [200] -c [50x50x50x30,1e-5] \
-%    -x T1_subX_mask.nii.gz -o T1_subX_n4.nii.gz
-%\end{lstlisting}
-%\end{command} 
-%\newline
-%  where \verb#T1_subX_mask.nii.gz# is the output of the skull stripping step 
-%  with brain voxel values of `1' and `0' elsewhere.  The following flags are used:
-%  \begin{description}
-%    \item[-d] is the image dimension,
-%    \item[-i] defines the input image,
-%    \item[-b] specifies the ``spline distance'',
-%    \item[-c] defines the iterations and convergence threshold, and
-%    \item[-o] specifies the output image.
-%  \end{description}
-%Other options can be explored with the help menu (use \verb#--help#).
-%\begin{command}
-%\lstsetcpp
-%\begin{lstlisting}
-%$ Atropos -d 3 -a T1_subX_n4.nii.gz -i kmeans[3] \
-%    -m [0.2,1x1x1] -c [5,0.0000001] -k Gaussian \
-%    -x T1_subX_mask.nii.gz -o T1_subX_atropos.nii.gz
-%\end{lstlisting}
-%\end{command} 
-%\newline
-%  the following flags are used:
-%  \begin{description}
-%    \item[-d] is the image dimension,
-%    \item[-a] defines the input image,
-%    \item[-i] specifies the initialization method (here we are using K-means with 3 classes),
-%    \item[-m] is the 
-%    \item[-c] defines the iterations and convergence threshold, and
-%    \item[-o] specifies the output labeled image.
-%  \end{description}
-%Other options can be explored with the help menu (use \verb#--help#).
-%\newline
-%This step provides the white matter ma
-%\end{enumerate}
 
@@
 
-The tensorial data associated with each of the following data sets were reconstructed from the diffusion weighted sequence using Camino%
+Tensorial data associated with each of the following data sets were reconstructed from the diffusion weighted sequences using Camino%
 \footnote{
@@
 
-\subsubsection{Data from the International Neuroimaging Data-sharing Initiative (INDI)}
-
-In support of open science, the 1000 Functional Connectomes Project (FCP)%
-\footnote{
-ttp://fcon\_1000.projects.nitrc.org
-} 
-was initiated
-on December 11, 2009 by various members of the MRI community \cite{Biswal2010}.
-Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
-Commitments from institutions such as they NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
-currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC).%
-\footnote{
-http://www.nitrc.org
-}
-Data from these two institutions were used to demonstrate the alignment properties obtained using SyGN for DTI-derived data normalization.
-
-\paragraph{NKI/Rockland Sample}
-Data from the first 14 weeks of prospective data-sharing were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
-Due to various technical issues (e.g.
-lack of corresponding T1 image, failed DTI reconstruction, and inadequate FSL registration performance) only 57 subjects of the NKI data set were used in the image alignment assessment.
-Germane to our evaluation, each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
-EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
-voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
-TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
-1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).  Images were anonymized including defacing of the T1 images prior to upload.
-
-\paragraph{NYU Institute for Pediatric Neuroscience Sample}
-Data from the first 3 parts of prospective data-sharing from NYU were also downloaded on January 15, 2011 comprised of 49 subjects (average age in years: 30.29 $\pm$ 8.81).
-All subjects were used in the image alignment assessment.
-Two 64-directional DTI scans were performed (parameters: conventional single-shot spin echo
-EPI pulse sequence, TR = 5200 ms, TE = 78 ms, axial acquisition, 
-voxel size = $3 \times 3 \times 3$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2530 ms, TE = 3.25 ms, flip angle = 7$^\circ$, 256 contiguous slices of 1.33 mm, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1.3 mm$^3$).
-Images were anonymized including defacing of the T1 images prior to upload.
-
-%
-%The 1000 Functional Connectomes Project (FCP)%
-%\footnote{
-%http://fcon\_1000.projects.nitrc.org/index.html
-%}
-%
-%{\color{red}{Quote}}  INDI is a next-generation data-sharing effort, with the goal of sharing phenotypically rich imaging datasets, on both retrospective and prospective bases.
-On 10/8/10, the Nathan Kline Institute (NKI) / Rockland Sample led the way for INDI-Prospective, starting weekly uploads of 5-7 subjects / week (R-fMRI, DTI, and extensive psychiatric and cognitive phenotypic data, including IQ), which continue to progress (currently n = 55 and growing!).
-More recently, the NYU Institute Pediatric Neuroscience released its sample examining cocaine dependence in adults via INDI-Retro.
-The list of sites pledging data from around the world is growing, and we are working to generate common phenotypic variables across datasets as possible (notice the frequency with which IQ is starting to appear Ð itÕs just one of our targets!).
-%
-%\subsubsection{Nathan Kline Institute (NKI) / Rockland Sample}
-%
-%{\color{red}{Quote}}  The NKI/Rockland Sample is intended to be a phenotypically rich neuroimaging sample, consisting of data obtained from individuals between the ages of 6 and 85 year-old.
-All individuals to be included in the sample undergo semi-structured diagnostic psychiatric interviews, and complete a battery of psychiatric, cognitive and behavioral assessments in order to provide comprehensive phenotypic information for the purpose of exploring brain/behavior relationships.
-%
-%Each subject consists of the following data:
-%\begin{itemize}
-%\item 10 minute resting state fMRI scan (R-fMRI) 
-%\item 6-direction diffusion tensor imaging (DTI) scan 
-%\item 64-direction diffusion tensor imaging scan (2mm isotropic)
-%\item MPRAGE anatomical scan
-%\item A variety of psychiatric, cognitive and behavioral assessments
-%\end{itemize}
-%
-%\subsubsection{NYU Institute for Pediatric Neuroscience}
-%
-%{\color{red}{Quote}}  Summary: The NYU Phyllis Green and Randolph C?wen Institute for Pediatric Neuroscience (IPN) sample will be a collection of past and present scans obtained from psychiatrically screened individuals ranging in age from 6 to 55 years old.
-The initial release consists of datasets from 49 psychiatrically neurotypical adults, with age, gender and intelligence quotient (IQ) information provided.
-Future releases will include more comprehensive phenotypic information, and child and adolescent datasets, as well as individuals from clinical populations.
-%
-%Each subject consists of the following data:
-%\begin{itemize}
-%\item At least one 6-minute resting state fMRI scan (R-fMRI)%
-%\footnote{
-% Most participants have 2 R-fMRI scans, collected less than 1 hour apart in the same scanning session. Rest$_1$ is always collected first.
-%}
-%\item One high-resolution T1-weighted mprage, defaced to protect patient confidentiality
-%\item Two 64-direction diffusion tensor imaging scans
-%\item Demographic information (age, gender) and IQ-measures (Verbal, Performance, and Composite; Weschler Abbreviated Scale of Intelligence [WASI])
-%\end{itemize}
-%
 \subsubsection{Diffuse Traumatic Brain Injury Data}
 
-A third data set was obtained as port of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in long-term survivors of TBI (principal investigator: John Whyte).
+The TBI data used in this study is part of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in long-term survivors of TBI (principal investigator: John Whyte).
 A more detailed data description can be found in \cite{Avants2008}.
@@
 1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
-DTI images were also obtained
+DTI images were also obtained.
 
@@
 
+\subsubsection{Data from the International Neuroimaging Data-sharing Initiative (INDI)}
+
+In support of open science, the 1000 Functional Connectomes Project (FCP)%
+\footnote{
+ttp://fcon\_1000.projects.nitrc.org
+} 
+was initiated
+on December 11, 2009 by various members of the MRI community \cite{Biswal2010}.
+Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
+Commitments from institutions such as they NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
+currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC).%
+\footnote{
+http://www.nitrc.org
+}
+%Data from these two institutions were used to demonstrate the alignment properties obtained using SyGN for DTI-derived data normalization.
+
+\paragraph{NKI/Rockland Sample}
+Data from the first 14 weeks of prospective data-sharing were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
+Due to various issues (e.g.
+lack of corresponding T1 image, failed DTI reconstruction, and age matching requirements) only 57 subjects were used for the first alignment analysis and 27 subjects of the NKI data set were used for the second analysis consisting of 9 females (age: $29.3 \pm 6.6$ years) and 18 males (age: $ 26.6 \pm 6.9$ years).
+Germane to our evaluation, each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
+EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
+voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
+TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
+1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).  Images were anonymized including defacing of the T1 images prior to upload.
+
+%\paragraph{NYU Institute for Pediatric Neuroscience Sample}
+%Data from the first 3 parts of prospective data-sharing from NYU were also downloaded on January 15, 2011 comprised of 49 subjects (average age in years: 30.29 $\pm$ 8.81).
+All subjects were used in the image alignment assessment.
+Two 64-directional DTI scans were performed (parameters: conventional single-shot spin echo
+%EPI pulse sequence, TR = 5200 ms, TE = 78 ms, axial acquisition, 
+%voxel size = $3 \times 3 \times 3$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2530 ms, TE = 3.25 ms, flip angle = 7$^\circ$, 256 contiguous slices of 1.33 mm, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1.3 mm$^3$).
+Images were anonymized including defacing of the T1 images prior to upload.
+
+%
+%The 1000 Functional Connectomes Project (FCP)%
+%\footnote{
+%http://fcon\_1000.projects.nitrc.org/index.html
+%}
+%
+%{\color{red}{Quote}}  INDI is a next-generation data-sharing effort, with the goal of sharing phenotypically rich imaging datasets, on both retrospective and prospective bases.
+On 10/8/10, the Nathan Kline Institute (NKI) / Rockland Sample led the way for INDI-Prospective, starting weekly uploads of 5-7 subjects / week (R-fMRI, DTI, and extensive psychiatric and cognitive phenotypic data, including IQ), which continue to progress (currently n = 55 and growing!).
+More recently, the NYU Institute Pediatric Neuroscience released its sample examining cocaine dependence in adults via INDI-Retro.
+The list of sites pledging data from around the world is growing, and we are working to generate common phenotypic variables across datasets as possible (notice the frequency with which IQ is starting to appear Ð itÕs just one of our targets!).
+%
+%\subsubsection{Nathan Kline Institute (NKI) / Rockland Sample}
+%
+%{\color{red}{Quote}}  The NKI/Rockland Sample is intended to be a phenotypically rich neuroimaging sample, consisting of data obtained from individuals between the ages of 6 and 85 year-old.
+All individuals to be included in the sample undergo semi-structured diagnostic psychiatric interviews, and complete a battery of psychiatric, cognitive and behavioral assessments in order to provide comprehensive phenotypic information for the purpose of exploring brain/behavior relationships.
+%
+%Each subject consists of the following data:
+%\begin{itemize}
+%\item 10 minute resting state fMRI scan (R-fMRI) 
+%\item 6-direction diffusion tensor imaging (DTI) scan 
+%\item 64-direction diffusion tensor imaging scan (2mm isotropic)
+%\item MPRAGE anatomical scan
+%\item A variety of psychiatric, cognitive and behavioral assessments
+%\end{itemize}
+%
+%\subsubsection{NYU Institute for Pediatric Neuroscience}
+%
+%{\color{red}{Quote}}  Summary: The NYU Phyllis Green and Randolph C?wen Institute for Pediatric Neuroscience (IPN) sample will be a collection of past and present scans obtained from psychiatrically screened individuals ranging in age from 6 to 55 years old.
+The initial release consists of datasets from 49 psychiatrically neurotypical adults, with age, gender and intelligence quotient (IQ) information provided.
+Future releases will include more comprehensive phenotypic information, and child and adolescent datasets, as well as individuals from clinical populations.
+%
+%Each subject consists of the following data:
+%\begin{itemize}
+%\item At least one 6-minute resting state fMRI scan (R-fMRI)%
+%\footnote{
+% Most participants have 2 R-fMRI scans, collected less than 1 hour apart in the same scanning session. Rest$_1$ is always collected first.
+%}
+%\item One high-resolution T1-weighted mprage, defaced to protect patient confidentiality
+%\item Two 64-direction diffusion tensor imaging scans
+%\item Demographic information (age, gender) and IQ-measures (Verbal, Performance, and Composite; Weschler Abbreviated Scale of Intelligence [WASI])
+%\end{itemize}
+%
+
+\subsection{FMRIB Software Library (FSL)}
+
+FSL hosts a large number of algorithms for computational neuroanalysis.
+These include, among others, well-known tools for image segmentation (FAST---\cite{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), and TBSS and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).
+Based solely on the number of citations in the literature,%
+\footnote{
+As of the beginning of the year 2011, the original Neuroimage 2006 TBSS paper has been cited more than 360 times making it the most cited in the journal's library.
+}
+usage of TBSS has become a de facto quantitative tool for DTI analysis based, at least in part, on its accessibility and ease of use.
+Due to its large presence in the neuroimaging community, many of the comparisons that are performed in the next section use the FMRIB58 template (shown in the upper left of Figure \ref{fig:workflow}).
+This template serves as the final coordinate system for most of the comparisons whether it be based on direct normalization or T1 normalization to the MNI152 template which resides in the FMRIB58 space.
+We also used the normalization component of TBSS composed of FLIRT and FNIRT registration steps optimized for normalization to FMRIB58 to provide an external comparative baseline (which we denote in subsequent sections as ``tbss'').
+
+%For the comparative evaluation studies used in this manuscript, TBSSv1.2%
+%\footnote{
+%Build 416 downloaded from the FMRIB web site http://www.fmrib.ox.ac.uk/fsldownloads/ on November 18, 2010.
+%}
+%was used.  The actual code pipeline constituting TBSS prior to statistical analysis consists of the following 4 shell scripts:
+%\begin{description}
+%  \item[tbss\_1\_preproc]  Input is a directory of FA images which are slightly eroded and then organized into a separate FA directory.
+A mask is also generated for each input FA image for use during the spatial normalization process.
+%  \item[tbss\_2\_reg]  The organized FA data and corresponding masks produced in the previous step are warped to one of three possible coordinate systems using FLIRT/ FNIRT:  1)  the atlas included with the FSL download---FMRIB58---(recommended), 2) a user-specified target image or 3) an exhaustive search for the ``optimal'' target image.
+In \cite{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.
+However, even though it would be faster to simply choose a single subject at random, the complexities of nonrigid image registration would seemingly promote the exhaustive search over random selection.
+As a compromise between these two options, the first option in the script is to use FMRIB58 which also resides in the space of the MNI152 T1 template.
+It should also be noted that this particular script invokes another FSL script, {\bf fsl\_reg}.
+An option associated with {\bf fsl\_reg} permits the use of an FNIRT configuration file which has been optimized for use with the FMRIB58 atlas.
+%  \item[tbss\_3\_postreg] The mean FA image is generated from the aligned FA images which is subsequently skeletonized.
+%  \item[tbss\_4\_prestats] The skeleton mask is created by thresholding the mean FA skeleton.  The skeleton distance map is then generated and all FA data are projected onto the skeleton.
+%\end{description}
+
 \section{Results and Discussion}
@@
 
-\subsection{Alignment Assessment of the INDI Data}
+\subsection{TBI Cohort}
+
+\begin{figure*}
+\begin{center}
+\begin{tabular}{c}
+  \includegraphics[width=180mm]{tbiTesting.pdf}
+\end{tabular}
+\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
+Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn.
+(\ref{eq:variance}), namely the SSD and Demons metric which shrinks drastically for the Wilcoxon rank-sum test.
+Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.
+TBSS and the 4 ANTs metric
+results were all obtained using direct registration involving subject FA and the FMRIB58 template.
+The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
+common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
+\label{fig:tbi_testing}
+\end{center}        
+\end{figure*}
+
+A patient template and a control template were  constructed 
+from the T1-weighted images of the two groups.  These two 
+templates were then used to create a third common template.  
+To compare with TBSS normalization, the common template was
+registered to the MNI152 template.  Thus, the transformation
+which takes the FA image of an individual subject to the FMRIB58
+template is the composition of transforms: 
+subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
+common T1 template $\rightarrow$ MNI152 (FMRIB58) template.
+We evaluated this strategy relative to direct FA-to-FA
+ANTs registration with the common similarity metrics used in
+the simulated data example described earlier.
+Standard TBSS alignment was also performed.
+
+Statistical power curves (one-tailed, $\alpha = 0.95$)
+are shown in Fig. \ref{fig:tbi_testing}.
+Similar to previous findings with the simulated data, the SSD and 
+Demons metric show elevated statistical power using the Student's 
+$t$-test over the other alignment protocols which all exhibit similar 
+strength.  However, this discrepancy decreases significantly with the 
+nonparametric testing.  The Demons and SSD power decreases while the others
+remain fairly similar.  Interestingly, the ANTs template-based approach
+trends towards better performance relative to the remaining 3 strategies.
+It should be noted that although TBSS uses
+the SSD metric, the fixed and moving images are smoothed and subsampled
+even at the highest resolution level which substantially decreases the local effect
+of the metric.
 
 \begin{table}
-  \caption{Volume (in voxels) of thresholded mean FA skeletons.}
-  \label{table:vol}
-  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
+  \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
+  \label{table:lom}
+  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccccc}
   \hline
-  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-  \cline{2-5}
-  {} & inner & outer & inner & outer \\
+%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
+%  \cline{2-5}
+  {Norm.} & Target & Jaccard & Dice  & False $-$ & False $+$\\
   \hline
-  FMRIB & 84899 & 122306 & 72765 & 76769 \\
-  ANTs$_5$ & 87572 & 134421 & 73650 & 94177 \\
-  ANTs$_{10}$ & 86908 & 130569 & 73419 & 98014 \\
-  ANTs$_{20}$ & 86523 & 129729 & 73577 & 95484 \\
-  ANTs$_{30}$ & 86971 & 128892 & 73443 & 94862 \\
   \hline
+  antsT1 & 0.917 & 0.621 & 0.767 & 0.082 & 0.341 \\
+  antsCC & 0.780 & 0.570 & 0.727 & 0.220 & 0.320 \\
+  antsMI & 0.776 & 0.569 & 0.725 & 0.223 & 0.320 \\
+  antsDemons & 0.780 & 0.566 & 0.723 & 0.220 & 0.326 \\
+  antsSSD & 0.783 & 0.562 & 0.720 &  0.217 & 0.334 \\
+  tbss & 0.716 & 0.550 & 0.710 & 0.283 & 0.295 \\
   \hline
@@
 
-\begin{table}
-  \caption{Number of connected components of thresholded mean FA skeletons.}
-  \label{table:cc}
-  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
-  \hline
-  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-  \cline{2-5}
-  {} & $\geq$ 1 voxel & $\geq$ 2 voxels & $\geq$ 1 voxel & $\geq$ 2 voxels \\
-  \hline
-  FMRIB & 4340 & 1802 & 2431 & 993 \\
-  ANTs$_5$ & 3561 & 1384 & 2278 & 916 \\
-  ANTs$_{10}$ & 3413 & 1302 & 2232 & 899 \\
-  ANTs$_{20}$ & 3283 & 1252 & 2093 & 885 \\
-  ANTs$_{30}$ & 3114 & 1243 & 2128 & 867 \\
-  \hline
-  \hline
-  \end{tabular*}
-\end{table}
-
-Since our proposed modification of TBSS impacts strictly the
-critical alignment step, the first set of evaluations demonstrates
-the difference between the two methods, i.e. spatial normalization 
-via registration of the FA maps to the FMRIB58 template
-or spatial normalization via ANTs template construction.  It is important to note that we did not consider normalization either to a random subject
-or to the subject best representative of the mean as defined in 
-\cite{Smith2006}. The former was not investigated due to the possible bias
-in selecting a random subject as described in \cite{Smith2006}.  
-Neither was the latter option pursued due to the computational complexity
-required for such an $\OO{n^2}$ search over the relatively large NKI and NYU data sets.  In addition, normalization to
-the FMRIB58 template is both the default and recommended option
-with the parameters of FNIRT optimally tuned for said option.  To warp the MD images following FA registration we used FSL's {\bf applywarp} function using the FA-derived transform.
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{NYU.pdf}
-%\end{tabular}
-%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.
-(a) The standard FMRIB58\_FA\_1mm template.
-(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
-%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).
-Note the visual alignment difference particularly in the peripheral white matter complex.
-%}
-%\label{fig:NYU}
-%\end{figure}
-
-To explore alignment with ANTs-TBSS, 4 templates were constructed for each of the 2 data sets.
-Each within-group template was generated from $\{5,10,20,30\}$ randomly selected subjects to see if the number of subjects comprising the template affected the alignment performance.
-In addition
-to the composite transforms described in Section \ref{sec:ANTs-TBSS} to take each subject's FA or MD image to the population-specific anatomical template, we also registered
-each ANTs T1 template to the MNI152 template.  This permitted a
-standard space for qualitative and quantitative comparison.
-
-Sample axial slices for all resulting FA mean images for the NKI data set are provided in Figure \ref{fig:NKI}.
-We also include the same axial slice of the FMRIB58 template in Figure \ref{fig:NKI} (a) for comparison.
-In both sets of sample slices, visual inspection would seemingly indicate better performance with the ANTs-based FA mean images compared with the standard TBSS-produced FA mean image.
-Also, there is little discernible difference between the different ANTs results.
-Alignment differences between the two approaches are particularly apparent in the peripheral white matter where the anatomical structure in the ANTs template-based FA mean images is visually differentiable.
-
-Relative performance is quantified in the plots produced in Figure \ref{fig:NKIplots} and Figure \ref{fig:NYUplots}.
-For both data sets, the mean FA image derived from the standard TBSS approach was thresholded at FA $\geq 0.2$.
-These two binary masks obtained in both the NKI and NYU cohorts provided the standard group masks for fair comparisons for all evaluations.
-For the aligned images the coefficient of variation (CoV) was calculated in a voxel-wise fashion within the mask.
-Intuitively, one would expect a smaller variance across better aligned sets of voxel values thus producing a lower CoV for a given mean value.
-By calculating the cumulative distribution function (CDF) of the set of CoV values within the mask, quantitative assessment of alignment performance can be visualized.
-In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
-This division was accomplished by applying a distance transform \cite{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.
-Note that only those voxels that were also within the thresholded FA mask were considered in the inner/outer analysis.
-
-There is a visible improvement in alignment for both data sets using the ANTs-based approach over individual subject FA registration to the FMRIB template.
-There is little difference between the ANTs results with only a slight trend in improved performance with the larger template subject sample sizes.
-This was the case for all three regional CoV calculations for both data sets.
-Also, as expected, there is greater misalignment in the outer region versus the inner core of the white matter mask regardless of methodology.
-However, the performance tally was the same in both the inner and outer regions as was the case with the entire white matter mask although the disparity in performance was much less for the NYU data set as the NKI data set most likely due to the differences in resolution of the original DTI.
-
-The improvements seen in the CDF plots are also supported by comparative quantities derived from the skeletons of the different mean FA images.
-Since the skeletonization process is meant to determine the center of the various white matter tracts \cite{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.
-In addition, as described in \cite{Zalesky2011}, the number of connected components of the white matter skeleton is directly related to the contiguity and thus inversely related to the level of structure.
-Table \ref{table:vol} provides the volume of the mean FA skeletons (thresholded at FA $\geq 0.2$) for standard TBSS protocol using the FMRIB template and the ANTs approach using the various templates.
-In agreement with the CDF plots illustrating improved alignment with ANTs, the voxel volumes of the skeletons are greater using ANTs templates with the greatest difference seen in the outer portion of the white matter.
-Related, in Table \ref{table:cc}, we provide the number of connected components for both data sets for all isolated components greater in size than 1 voxel and 2 voxels.
-Even though the skeletal volumes associated with the ANTs approach are greater, the number of connected components is less which would also support the observation of improved alignment.
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{inner.png}
-%\end{tabular}
-%\caption{
-%}
-%\label{fig:inner}
-%\end{figure}
-%
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{outer.png}
-%\end{tabular}
-%\caption{Three-canonical view of the inner
-%}
-%\label{fig:outer}
-%\end{figure}
-
-\begin{figure*}
+The CDF of the voxelwise FA coefficient of variation of the aligned
+images is shown for each case in Figure \ref{fig:tbi_cov}.  This is
+very much related to earlier findings in that strategies evincing
+greater alignment, as quantified by the coefficient of variation,
+will trend towards smaller variance thus exhibiting a left-shifted
+CDF.  Although direct FA-to-FA registration using ANTs with common
+similarity metrics gives ``better alignment'' such alignment is 
+produced at a cost of significant type 1 errors.  Common subsampling and
+smoothing strategies and local neighborhood metrics such as CC and MI reduce the 
+Type 1 error problem (as evidenced in Fig. \ref{fig:tbi_testing}) 
+but also show a decrease in alignment.  This is further supported
+by the overlap measures \cite{Crum2006} given in Table \ref{table:lom}.
+For each alignment strategy, we thresholded the derived mean FA image (FA $
+\geq$ 0.2) and compared it with white matter mask of the MNI152 template
+segmented using Atropos \cite{Avants2011a}.  The template-based
+strategy showed the best overall overlap with the white matter MNI152
+segmentation.
+
+Finally, qualitative assessment is provided in Fig. \ref{fig:meanFA} 
+in which the mean FA image derived from the template-based approach
+and the post-normalization component of the TBSS pipeline (Fig. \ref{fig:meanFA}(c)) 
+is juxtaposed 
+with the mean FA image produced by the TBSS pipeline alone (Fig.\ref{fig:meanFA}(a)).
+The skeleton masks were produced by thresholding the skeleton values at 
+FA $\geq 0.2$ and it would seem that the improved alignment produces 
+a more extensive skeletal network for greater localization testing of 
+significant differences, particularly in the outer white matter regions
+(compare Fig. \ref{fig:meanFA}(b) and Fig. \ref{fig:meanFA}(d)).
+%Given that TBSS is intrinsically a statistical method for finding local 
+%differences, improved alignment would presumably provide increased regional sensitivity to differences.
+
+\begin{figure}
+  \begin{center}
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{NKIplots2.pdf}
+  \includegraphics[width=85mm]{tbiCoV.pdf}
 \end{tabular}
-\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Figure \ref{fig:NKI}(b)).
-In addition to the whole brain analysis (first column), the coefficient of variation was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-}
-\label{fig:NKIplots}
-\end{figure*}
-
-\begin{figure*}
+\caption{Comparison of the different alignment strategies in normalizing the TBI data to the FMRIB58 template.
+Assessment is quantitated using the voxelwise coefficient of variation of the FA.
+TBSS and the 4 ANTs metric
+results were all obtained using direct registration involving subject FA and the FMRIB58 template.
+The curve denoted as ``antsT1'' was also obtained in FMRIB58 space with the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
+common T1 template $\rightarrow$ MNI152 (FMRIB58) template. }
+\label{fig:tbi_cov}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{NYUplots2.pdf}
+  \includegraphics[width=85mm]{meanFA.pdf}
 \end{tabular}
-\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Figure \ref{fig:NKI}(b)).
-In addition to the whole brain analysis (first column), the coefficient of variation was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-}
-\label{fig:NYUplots}
-\end{figure*}
-
-\subsection{TBSS and ANTs-TBSS of TBI Cohort}
-
-\begin{figure}
-\begin{tabular}{c}
-  \includegraphics[width=80mm]{meanFA.pdf}
-\end{tabular}
-\caption{Comparison of mean FA images used in TBSS taken from the TBI data described in Section 2.
+\caption{Qualitative comparison of mean FA images used in TBSS taken from the TBI data described in Section 2.
 The mean FA image illustrated in (a) is created by mapping the sample FA population to the default FMRIB58 template which is the standard protocol for TBSS.
@@
 The respective skeleton masks in (b) and (d) are created by thresholding the resulting skeleton values $\geq 0.2$.
-The yellow area points to a peripheral difference in findings between the two analyses.
 }
 \label{fig:meanFA}
+\end{center}
 \end{figure}
 
-A patient template and a control template were  constructed 
-using the ANTs software from all subjects of the respective populations.
-These two templates were then used to create a third template which provided the standard space for the ANTs-based TBSS assessment.
-Standard TBSS was also performed.
-
-The resulting mean FA images and thresholded skeletons are represented in Figure \ref{fig:meanFA} which demonstrate appreciable in alignment.
-These images support earlier findings in improved alignment with the ANTs-based approach.
-The skeleton masks were produced by thresholding the skeleton values at FA $\geq 0.2$ and it would seem that the improved alignment produces a more extensive skeletal network for greater localization testing of significant differences, particularly in the outer white matter regions.
-Given that TBSS is intrinsically a statistical method for finding local differences, improved alignment would presumably provide increased regional sensitivity to differences.
-
-%
-%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.
+%
+%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Fig.
+\ref{fig:hoonData}.
 The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.
@@
 
+\subsection{INDI NKI/Rockland Sample}
+In this section, further exploration of relevant issues employs 
+the NKI/Rockland sample of the INDI database.  
+The first is one of experimental set-up consideration in which 
+we investigate how template-based
+alignment is affected by the number of subject images used to 
+create the template.  Both FA and mean diffusivity (MD) images were used.
+This is followed by statistical analysis
+of discriminating power of gender FA differences using 
+the various alignment strategies.
+
+\subsubsection{Alignment Assessment of Template Group Size}
+
+ \begin{figure}
+  \begin{center}
+\begin{tabular}{c}
+  \includegraphics[width=85mm]{NKI2.pdf}
+\end{tabular}
+\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.
+(a) The standard FMRIB58 template.
+(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
+(c) and (d) The mean FA image produced using the ANTs template-based strategy employing 5 and 30 randomly chosen subjects, respectively, warped to the space of (a) followed by T1 registration of all subjects to the resulting anatomical template.
+Note the visual alignment difference particularly in the peripheral white matter complex.
+}
+\label{fig:NKI}
+\end{center}
+\end{figure}
+
+Four templates were generated from $\{5,10,20,30\}$ randomly selected subjects 
+to see if the number of subjects comprising the template affected the alignment 
+performance. In addition to the composite transforms
+ to take each subject's FA or MD image to the population-specific anatomical template, we also registered
+each ANTs T1 template to the MNI152 template.  
+This permitted a
+standard space for qualitative and quantitative comparison with TBSS-aligned
+FA and MD images.  
+After template construction, the remaining
+images were registered to the resulting template and, ultimately, 
+to the MNI152 (FMRIB58) template via the template-to-MNI152 transformation.
+
+\begin{figure}
+\begin{tabular}{c}
+  \includegraphics[width=85mm]{NKIplots3.pdf}
+\end{tabular}
+\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
+\ref{fig:NKI}(b)).
+The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
+}
+\label{fig:NKIplots}
+\end{figure}
+
+Sample axial slices for the FA FMRIB58-aligned mean images for the NKI data 
+set of template sizes of 5 and 30 images are provided in Fig. \ref{fig:NKI}.  
+We also include the same axial slice  of the FMRIB58 template in Fig. \ref{fig:NKI} 
+(a) for comparison.  In both sets of sample slices, visual inspection would seemingly 
+indicate better 
+performance with the ANTs-based FA mean images compared with the standard 
+TBSS-produced FA mean image.  Also, there is little discernible difference
+ between the different ANTs results.  Alignment differences between the 
+ approaches are particularly apparent in the peripheral white matter where 
+ the anatomical structure in the ANTs template-based FA mean images is visually differentiable.
+
+Relative performance is quantified in the plots produced in Fig. \ref{fig:NKIplots}. 
+The mean FA image derived from the standard TBSS approach was thresholded at FA $\geq 0.2$. 
+ This binary mask obtained in the NKI cohort provided the standard group mask for 
+ fair comparisons for all evaluations.    For all FMRIB58-aligned images the coefficient of 
+ variation (CoV) was calculated in a voxel-wise fashion within the mask.  Intuitively, 
+ one would expect a smaller variance across better aligned sets of voxel values thus 
+ producing a lower CoV for a given mean value.  By calculating the cumulative distribution 
+ function (CDF) of the set of CoV values within the mask, quantitative assessment of alignment performance can be visualized.  
+% In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
+This division was accomplished by applying a distance transform \cite{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.
+Note that only those voxels that were also within the thresholded FA mask were considered in the inner/outer analysis.
+
+There is a visible improvement in alignment for both data sets using the ANTs-based approach over individual subject FA registration to the FMRIB template.
+There is little difference between the ANTs results with only a slight trend in improved performance with the larger template subject sample sizes.
+%This was the case for all three regional CoV calculations for both data sets.
+Also, as expected, there is greater misalignment in the outer region versus the inner core of the white matter mask regardless of methodology.
+However, the performance tally was the same in both the inner and outer regions as was the case with the entire white matter mask.% although the disparity in performance was much less for the NYU data set as the NKI data set most likely due to the differences in resolution of the original DTI.
+
+%The improvements seen in the CDF plots are also supported by comparative quantities derived from the skeletons of the different mean FA images.
+Since the skeletonization process is meant to determine the center of the various white matter tracts \cite{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.
+In addition, as described in \cite{Zalesky2011}, the number of connected components of the white matter skeleton is directly related to the contiguity and thus inversely related to the level of structure.
+Table \ref{table:vol} provides the volume of the mean FA skeletons (thresholded at FA $\geq 0.2$) for standard TBSS protocol using the FMRIB template and the ANTs approach using the various templates.
+In agreement with the CDF plots illustrating improved alignment with ANTs, the voxel volumes of the skeletons are greater using ANTs templates with the greatest difference seen in the outer portion of the white matter.
+%Related, in Table \ref{table:cc}, we provide the number of connected components for both data sets for all isolated components greater in size than 1 voxel and 2 voxels.
+Even though the skeletal volumes associated with the ANTs approach are greater, the number of connected components is less which would also support the observation of improved alignment.
+
+%\begin{table}
+%  \caption{Volume (in voxels) of thresholded mean FA skeletons.}
+%  \label{table:vol}
+%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
+%  \hline
+%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
+%  \cline{2-5}
+%  {} & inner & outer & inner & outer \\
+%  \hline
+%  FMRIB & 84899 & 122306 & 72765 & 76769 \\
+%  ANTs$_5$ & 87572 & 134421 & 73650 & 94177 \\
+%  ANTs$_{10}$ & 86908 & 130569 & 73419 & 98014 \\
+%  ANTs$_{20}$ & 86523 & 129729 & 73577 & 95484 \\
+%  ANTs$_{30}$ & 86971 & 128892 & 73443 & 94862 \\
+%  \hline
+%  \hline
+%  \end{tabular*}
+%\end{table} 
+%
+%\begin{table}
+%  \caption{Number of connected components of thresholded mean FA skeletons.}
+%  \label{table:cc}
+%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
+%  \hline
+%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
+%  \cline{2-5}
+%  {} & $\geq$ 1 voxel & $\geq$ 2 voxels & $\geq$ 1 voxel & $\geq$ 2 voxels \\
+%  \hline
+%  FMRIB & 4340 & 1802 & 2431 & 993 \\
+%  ANTs$_5$ & 3561 & 1384 & 2278 & 916 \\
+%  ANTs$_{10}$ & 3413 & 1302 & 2232 & 899 \\
+%  ANTs$_{20}$ & 3283 & 1252 & 2093 & 885 \\
+%  ANTs$_{30}$ & 3114 & 1243 & 2128 & 867 \\
+%  \hline
+%  \hline
+%  \end{tabular*}
+%\end{table}
+
+%Since our proposed modification of TBSS impacts strictly the
+%critical alignment step, the first set of evaluations demonstrates
+%the difference between the two methods, i.e. spatial normalization 
+%via registration of the FA maps to the FMRIB58 template
+%or spatial normalization via ANTs template construction.
+
+%It is important to note that we did not consider normalization either to a random subject
+%or to the subject best representative of the mean as defined in 
+%\cite{Smith2006}. The former was not investigated due to the possible bias
+%in selecting a random subject as described in \cite{Smith2006}.  
+%Neither was the latter option pursued due to the computational complexity
+%required for such an $\OO{n^2}$ search over the relatively large NKI data set.  In addition, normalization to
+%the FMRIB58 template is both the default and recommended option
+%with the parameters of FNIRT optimally tuned for said option.
+
+%To warp the MD images following FA registration we used FSL's {\bf applywarp} function using the FA-derived transform.
+
+%\begin{figure}
+%\begin{tabular}{c}
+%  \includegraphics[width=80mm]{NYU.pdf}
+%\end{tabular}
+%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.
+(a) The standard FMRIB58\_FA\_1mm template.
+(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
+%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).
+Note the visual alignment difference particularly in the peripheral white matter complex.
+%}
+%\label{fig:NYU}
+%\end{figure}
+
+%\begin{figure}
+%\begin{tabular}{c}
+%  \includegraphics[width=80mm]{inner.png}
+%\end{tabular}
+%\caption{
+%}
+%\label{fig:inner}
+%\end{figure}
+%
+%\begin{figure}
+%\begin{tabular}{c}
+%  \includegraphics[width=80mm]{outer.png}
+%\end{tabular}
+%\caption{Three-canonical view of the inner
+%}
+%\label{fig:outer}
+%\end{figure}
+
+%\begin{figure*}
+%\begin{tabular}{c}
+%  \includegraphics[width=180mm]{NYUplots2.pdf}
+%\end{tabular}
+%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
+\ref{fig:NKI}(b)).
+In addition to the whole brain analysis (first column), the coefficient of variation was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
+The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
+%}
+%\label{fig:NYUplots}
+%\end{figure*}
+
+\subsubsection{Statistical Power Related to Gender-based FA Differences}
+
+For completeness, the second NKI data subset was used to explore the 
+power of the different alignment strategies in discriminating FA 
+differences in a control group divided by gender.  As described 
+earlier, this group consisted of 27 subjects 
+comprised of 9 females (age: $29.3 \pm 6.6$ years) and  18 males (age: $26.6 
+\pm 6.9$ years).  Each subject was registered to the template made from 
+30 randomly chosen subjects, described in the previous section, and
+subsequently warped to the MNI152 template space.
+
+Results are shown in Fig. \ref{fig:nki_testing} in which all but the more
+robust metrics (MI and CC) exhibit similar power.  Note that the magnitude
+of the effects is much smaller than for TBI.  Again, the template-based approach
+performs comparatively well without the negative effects associated with
+the issues perviously described.
+
+\begin{figure*}
+\begin{center}
+\begin{tabular}{c}
+  \includegraphics[width=180mm]{nkiTesting.pdf}
+\end{tabular}
+\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results 
+(one-tailed, $\alpha = 0.95$) for the NKI FA data aligned to the FMRIB58 template 
+adjusted for multiple comparisons (FDR).  
+Similar statistical power is seen in all alignment scenarios except for the more robust
+metrics (CC and MI) in the $t$-test which show decreased power.   }
+\label{fig:nki_testing}
+\end{center}        
+\end{figure*}
+
 %\section{Discussion}
@@
 
-The popularity of FSL's TBSS is due, in large part, to its ease of use.
+%The popularity of FSL's TBSS is due, in large part, to its ease of use.
 Anybody interested in performing a neuroimaging study in which DTI-derived measures are compared across populations can simply download FSL and immediately apply it to their data.
@@
 
-The proposed improvement potentially has many benefits which were listed in the introduction.
-Such benefits range from the theoretical in the form of eliminating the confounding effects of registering data which is to be statistically differentiated to the practical---reusing ANTs template construction for other neuroimaging analyses.
-In performing the various analyses described in the work, it was pleasantly surprising to discover how well the alignments were for the ANTs template data considering that alignment was based on FA-to-FA correspondence and no such image registrations were performed to align FA directly with FA as with TBSS.
+With the increasing use of DTI for white matter assessment and the
+increasing number of researchers that are taking advantage of the numerous
+software tools that are available to facilitate this research, guidance
+is necessary to ensure the proper use of these tools and the
+calculation and reporting of findings \cite{Ridgway2008}.  The normalization 
+circularity issue described in this work is the inducement of substantial Type 1
+errors caused by FA-to-FA normalization using standard similarity metrics.
+To reduce these types of errors, we advocate the use of an anatomical template-building
+strategy whereby the transformation of FA images to a common space is performed
+using each individual's anatomical image.
+This proposed improvement potentially has many benefits:
+%which were listed in the 
+%introduction.  Such benefits range from the theoretical in the form of eliminating 
+%the confounding effects of registering data which is to be statistically 
+%differentiated to the practical, e.g. reusing the ANTs-based template for other 
+%neuroimaging analyses.  
+\begin{itemize}
+\item As explained in \cite{Avants2010}, ANTs optimal template construction creates a true Frech\'et mean from the population sample with high resolution which facilitates within-group registration.
+This obviates the need for an externally-derived template or an $\OO{n^2}$ search for an approximation to the mean subject or the bias resulting from random subject selection.
+\item Since the creation of the standardized coordinate system is performed using the anatomical images, nonrigid registration of the DTI-derived scalar images is 
+minimized thus mitigating the confounds described previously.
+Each DTI-based image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \cite{Wu2008}.
+\item Although the total number of registrations to be performed is increased, this makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
+%\item Instead of Gaussian smoothing as with SPM, TBSS uses the warping process to FMRIB58 by default to provide an indirect smoothing.
+Such resampling occurs in ANTs-TBSS as a by-product of the registration of the DTI-derived image to the corresponding T1 image in a subject-specific manner.
+%\item Instead of using thresholding to create the white matter segmentation from which the skeleton is created, an anatomically-based segmentation, either in the template space or in each subject space followed by a label fusion technique (e.g.
+majority voting), can be used to ultimately provide a more accurate skeleton.
+\item Oftentimes FA analysis is only one of several components performed in a population-based study.
+Other investigations, such as voxel- and tensor-based morphometry, cortical thickness measurements, etc.
+can take advantage of the anatomical template.
+\item The proposed modification is compatible with other frameworks such as TBSS where the normalization step is
+performed using ANTs-based tools followed by skeletonization and projection using the FMRIB software (see Fig. \ref{fig:workflow}).
+\end{itemize}
+
+Of practical significance, we emphasize that all software components described 
+in this paper are publicly available as open source primarily through the 
+Advanced Normalization Tools (ANTs) repository.%
+\footnote{
+http://www.picsl.upenn.edu/ANTs
+}
+In addition to registration and template building, ANTs also includes brain 
+extraction \cite{Avants2010a}, bias correction \cite{Tustison2010}, $n$-tissue 
+segmentation \cite{Avants2011a}, and cortical thickness estimation solutions \cite{Das2009}.
+
+%In performing the various analyses described in the work, it was pleasantly surprising to discover how well the alignments were for the ANTs template data considering that alignment was based on FA-to-FA correspondence and no such image registrations were performed to align FA directly with FA as with TBSS.
 Furthermore, since TBSS is a local quantification method with increased statistical power through the use of the skeletonization and projection step, improved alignments can increase the regional sensitivity to local white matter differences particularly in the peripheral tracts.
@@
 }
+We also gratefully acknowledge Dr. Niels van Strien of the Norwegian University of Science and Technology
+who packaged the template construction algorithm in the very useful script \verb#buildtemplateparallel.sh#
+which is publicly available in ANTs.
 

### 2011-06-06 ANTsTBSS/paper.tex
ENH: compile script + revised abstract

@@
 \begin{abstract}
-
-%% Text of abstract
-Critical to many studies investigating white matter differences across populations 
-is normalization of fractional anisotropy (FA) images to a common coordinate system
-for statistical analysis.  Although variations exist, common to several current 
-approaches is direct FA-to-FA registration whether it be to a chosen individual or
-a generated template.  This type of circular analysis 
-is susceptible to the general confounding 
-effect of minimizing data differences with nonrigid registration which are potentially 
-the precise differences to be statistically quantified.  We demonstrate that
-population normalization to a template explicitly optimizes certain voxelwise statistical 
-testing measures as opposed to maximizing anatomical alignment.  
-As a corrective, we advocate the use of the Symmetric Group Normalization strategy for 
-unbiased template construction from anatomical (e.g. T1-weighted) images which provides 
-the coordinate system for statistical inference of FA differences.
-Following template construction, individual FA images are warped to the template via 
-the corresponding anatomical image.  Performance is evaluated using open data from the International 
-Neuroimaging Data-sharing Initiative and a cohort of 16 traumatic brain injury survivors compared with 
-17 control subjects.  Alignment comparisons with the normalization component of the 
-popular tract-based spatial statistics (TBSS)
-show that our anatomically-based template approach produces good alignments despite
-the absence of any direct FA-to-FA registration.  Of practical consideration we note that the 
-software tools discussed in this work are provided to the public as part of the open source 
-Advanced Normalization Tools (ANTs) repository.
+Template-based normalization plays a critical role in many population
+studies based on diffusion tensor imaging (DTI).  Several current
+approaches compute the fractional anisotropy (FA) from the DTI and then
+directly register the FA image to a FA template before performing
+statistical comparison between different groups within a population.  In
+this paper, we demonstrate that this analysis strategy induces selection
+bias that varies in strength with the similarity metric used in
+registration.  Specifically, we show that the popular sum of squared
+difference intensity metric maximizes effect size, rather than
+anatomical alignment, and consistently overestimates statistical
+significance.  Consequently, we advocate a more traditional approach
+that normalizes a population of FA images through the associated T1
+anatomical images.  We show that the Symmetric Group Normalization
+strategy for unbiased template construction from anatomical (e.g.
+T1-weighted) images provides a valid and unbiased coordinate system for
+statistical inference of FA differences and yet retains detection power
+in comparison to FA-based normalization methods.  We establish this
+outcome by using open data in normal controls from the International
+Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
+brain injury survivors compared with 17 control subjects.  Alignment
+comparisons with the normalization component of the popular tract-based
+spatial statistics (TBSS) show that our anatomically-based template
+approach produces good alignments despite the absence of any direct
+FA-to-FA registration.  Of practical consideration we note that the
+software tools discussed in this work are provided to the public as part
+of the open source Advanced Normalization Tools (ANTs) repository.
 
@@
 including subjects with traumatic brain injury (TBI) \cite{Caeyenberghs2010,Jiang2010,Warner2010}.
+
+Normalization (also known as image registration) maps each image in the population to a common coordinate system thus enabling statistical comparison across patient groups or association with other variables such as genetics, cognition or age.
 

### 2011-06-06 ANTsTBSS/paper.tex
ENH: compile script + revised abstract
ENH: try to explain some of the issues --- edited to before section 2.

@@
 of the open source Advanced Normalization Tools (ANTs) repository.
+In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the ultimate statistical functions used to assess population effects.
 
@@
 \cite{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique.
-Continued quantitative development has furthered its applicability to probing clinical questions of interest particularly pertaining to the microstructure of white matter \cite{Basser1996,Assaf2008}.
-Given this sensitivity to brain architecture, DTI has found multiple uses in assessing neuro-structural differences in cross-population studies \cite{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010}
-including subjects with traumatic brain injury (TBI) \cite{Caeyenberghs2010,Jiang2010,Warner2010}.
-
-Normalization (also known as image registration) maps each image in the population to a common coordinate system thus enabling statistical comparison across patient groups or association with other variables such as genetics, cognition or age.
-
-Two popular comprehensive software packages for assessing population FA differences include 
-the SPM Matlab-based toolkit%
+DTI's sensitivity to brain architecture \cite{Basser1996,Assaf2008} enables multiple uses in assessing neuro-structural differences in cross-population studies \cite{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010} including subjects with traumatic brain injury (TBI) \cite{Caeyenberghs2010,Jiang2010,Warner2010}.
+Two popular comprehensive software packages for assessing population FA differences include the SPM Matlab-based toolkit%
 \footnote{
@@
 } 
-Although concerns regarding the general 
-application of VBM \cite{Bookstein2001,Ashburner2001,Davatzikos2004}
-and its application to FA data \cite{Jones2005,Chung2008} has been discussed 
-in the literature, it continues to find application in FA population 
-studies \cite{Kakeda2010,Takao2010}.
-The TBSS framework was partially developed in response to these concerns
-in which voxel values are projected onto a template white matter skeleton for 
-increased statistical power.  
+VBM\cite{Ashburner2001} continues to find application to FA studies\cite{Kakeda2010,Takao2010} despite concerns about its validity\cite{Bookstein2001,Davatzikos2004} in DTI \cite{Jones2005,Chung2008}.
+The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.
 %(e.g. \cite{Arnone2008,Bodini2009}).
@@
 commonality includes direct FA-to-FA registration.
-Potential difficulties associated with direct FA-to-FA normalization is exemplified by 
-a standard population study in which all subject FA images are mapped to a representative 
-template using the popular sum of squared differences (SSD) image similarity metric.%
+The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and an image gradient) and is therefore widely used.
+Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and provide reasonable performance levels \cite{Klein2009}.
 \footnote{
@@
 }  
-Given a set of $M$ subject images, $\{I_1, I_2, \ldots, I_M\}$, and
-a template image, $J$, population normalization concerns finding the set of $M$ transformations, 
+Now, suppose the SSD metric is optimized to find a set of $M$ transformations that map a population of $M$ FA images, $\{I_1, I_2, \ldots, I_M\}$, to a representative FA template, $J$.
+Then, the algorithm will compute transformations
 $\boldsymbol{\mathcal{T}} = \{\mathcal{T}_1, \mathcal{T}_2,\ldots,\mathcal{T}_M \}$, which minimize the SSD metric over the population, i.e.
@@
 \end{align}
-Switching the order of the summations (since the transformations are mutually independent) and 
-recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the 
-mean of the aligned images, it becomes apparent that the transformation solution
+Switching the order of the summations (since the transformations are mutually independent) and recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the mean of the aligned images, it becomes apparent that the transformation solution
 \begin{align}\label{eq:variance}
@@
 \end{align}
-is that which minimizes the voxelwise variance or, equivalently,
-maximizes various voxelwise statistical quantities including the common Student's $t$-test.%
-Note that many of the standard similarity metrics are also going to exhibit this
-effect to some degree as they minimize some local measure of intensity-based distance 
-between the subject and the template.
+is that which minimizes the voxelwise variance.
+Reducing variance will also directly affect the value of a standard population statistic, the Student's $t$-test.
+Other standard similarity metrics will also exhibit this
+trend to some degree as they minimize some local measure of intensity-based distance between the subject and the template.
 
 Thus, instead of normalization based on anatomical alignment followed by
-statistical testing, direct FA normalization 
-explicitly conflates anatomical alignment with optimizing certain statistical testing results
-evoking a sense of circularity in the analysis \cite{Kriegeskorte2010}.
-According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on
-data that were selected for showing the effect of interest or a related effect.''  As illustrated with
-Eqn. \ref{eq:variance}, direct FA-to-FA registration causes voxels to be matched (or selected) according to the
-the statistical testing result (or the effect of interest) instead of anatomical alignment.
+statistical testing, direct FA normalization explicitly conflates
+anatomical alignment with optimizing the very statistical testing
+results that one will ultimately use to assess hypotheses.  This immediately evokes a sense of circularity in the analysis \cite{Kriegeskorte2010}.
+According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on data that were selected for showing the effect of interest or a related effect.''  As illustrated with
+Eqn.
+\ref{eq:variance}, direct FA-to-FA registration will find transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's test).
+While the SSD metric is able to provide visually attractive alignments across subjects, the voxel-level details (that is, which voxels are matched where) are, as we show in the following simulation study, strikingly important in terms of the effect on detection power.
+
 %This results in a significant increase in the number of false positives for this type
@@
 Gaussian noise (mean = 0.0, standard deviation = 0.1) followed by 
-Gaussian smoothing (standard deviation = 1.5 mm).  A simulated patient cohort of 20 subjects
+Gaussian smoothing (standard deviation = 1.5 mm).
+Thus, each subject is differentiated only by a unique random draw from the noise distribution and there is no real shape difference between subjects.
+A simulated patient cohort of 20 subjects
 was created similarly using the simulated patient template.  A sample control subject and sample
@@
 The true significance curves are drawn in black.  Elevated statistical power is seen with all 4 
-         metrics for both tests although the effect is less in the more robust Wilcoxon results  Not shown are the significance curves for the Brunner-Munzel testing \cite{Rorden2007} which showed similar effects to the results on the left.
+         metrics for both tests although the effect is less in the more robust Wilcoxon results.
+Not shown are the significance curves for the Brunner-Munzel testing \cite{Rorden2007} which showed similar effects to the results on the left.
         }
@@
 and Wilcoxon rank-sum testing, one-tailed, $\alpha = 0.95$) with false discovery
-rate correction within the white matter to get ``true'' statistical significance 
-curves.  These curves are plotted in black in Fig. \ref{fig:simulated_testing}.  To illustrate the 
-effect of similarity metric on statistical significance, we performed ANTs registration
-\cite{Avants2011} of each subject to the template
+rate correction within the white matter to get ``true'' cumulative distribution functions (cdfs).
+The cdf vertical axis shows the cumulative number of voxels with corrected p-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
+From here, we denote these graphs as "statistical significance curves."  These curves are plotted in black in Fig.
+\ref{fig:simulated_testing}.
+To illustrate the
+effect of similarity metric on statistical significance, we performed ANTs registration \cite{Avants2011} of each subject to the template
 (Fig. \ref{fig:simulated_data}(b)) using cross correlation (CC), mutual information (MI), 
@@
 
-This example demonstrates the undesirable phenomenon previously described.  Even though these
-images were anatomically aligned prior to registration, the resulting deformations and 
-increased false statistical power can be explained in the context of Eqn. (\ref{eq:variance}).  
-Common similarity metric functions can induce false positives by minimizing local intensity 
-differences between the subjects and template which explicitly maximizes statistical
+This example demonstrates the undesirable phenomenon previously described.
+Even though these images were anatomically aligned prior to registration, the resulting deformations (falsely) increased detection power.
+This effect can be explained in the context of Eqn.
+(\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the t-test).
+Note, however, that this effect is increased by the fact that noise was added to the data.
+{\textcolor{red}{FIXME --- does more / less noise increase / decrease the exaggerated power?
+}}
+Common similarity metric functions can therefore induce false positives by minimizing local intensity 
+variance between the subjects and template which explicitly maximizes statistical
 testing results in contrast to maximizing anatomical alignment.  This effect is strongest in 
@@
 the $t$-test results given in Fig. \ref{fig:simulated_testing}(a) with the non-parametric Wilcoxon
-rank-sum testing (Fig. \ref{fig:simulated_testing}(b)) where as large of an effect for the latter would not be expected although an effect is
-still apparent.
-
-Although various processing choices mitigate these effects, such as Gaussian smoothing, nonparametric
-testing, and
-the use of more robust similarity metrics, such choices are often made on an ad hoc basis.  
+rank-sum testing (Fig. \ref{fig:simulated_testing}(b)).  The rank test does not explicitly use variance in its assessment and thus exhibits far less statistical exaggeration.
+
+Although various processing choices mitigate these effects, such as Gaussian smoothing, nonparametric testing, and the use of more robust similarity metrics, such choices are often made on an ad hoc basis.
 In contrast, in this work we strongly advocate the general application
@@
 of a high performance \cite{Klein2009}, nonrigid registration algorithm \cite{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \cite{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
+The key point, here, is that the FA images are mapped to the common reference space via intra-subject DT-to-T1 mapping followed by T1-to-T1 alignment.
+The T1-to-T1 step is (statistically) independent of the FA values which are the ultimate statistical target of hypothesis testing in FA population studies.
+Below, we show that this same pattern of effects is reproduced in real data and detail our proposed solution.
 

### 2011-06-07 ANTsTBSS/paper.tex
ENH: minor changes + important discussion points

@@
 and Wilcoxon rank-sum testing, one-tailed, $\alpha = 0.95$) with false discovery
-rate correction within the white matter to get ``true'' cumulative distribution functions (cdfs).
-The cdf vertical axis shows the cumulative number of voxels with corrected p-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
+rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).
+The CDF vertical axis shows the cumulative number of voxels with corrected p-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
 From here, we denote these graphs as "statistical significance curves."  These curves are plotted in black in Fig.
@@
 \subsection{Anatomical Image Template Construction}
-
-An important step in reducing intersubject variability is
-normalization to a standardized coordinate system for
-cross-population comparison.  Various approaches
-exist for determination of the normalized space such as selection of a
-pre-existing template based on a single subject, e.g. the Talairach atlas \cite{Talairach1988}), or 
-a publicly available averaged group of subjects, e.g. the MNI \cite{Collins1994} or 
-ICBM \cite{Mazziotta1995} templates.
-
-Recent trends, however, indicate a growing popularity of generating population-specific 
-templates for minimizing potential confounds.  This includes the
-diffeomorphic approach of Joshi et al. who employ the SSD metric and a shape 
-distance to bring the subject group of images into alignment \cite{Joshi2004}.  Variants
+In general, we normalize images to to a standard coordinate system to
+reduce intersubject variability in population studies.  Various
+approaches exist for determining the normalized space such as selection
+of a pre-existing template based on a single subject, e.g. the Talairach
+atlas \cite{Talairach1988}), or a publicly available averaged group of
+subjects, e.g. the MNI \cite{Collins1994} or ICBM \cite{Mazziotta1995}
+templates.
+One challenge with standard templates is that they may inadvertantly bias one's results by enabling better normalization of subjects to which the template is more similar.
+This issue is exacerbated when dealing with populations that have high variance (e.g.
+due to disease) and/or when one's normalization method is low-dimensional (not flexible enough to capture large shape differences).
+
+Population-specific templates (or optimal templates) alleviate some of
+these issues by deriving a most representative image from the population
+\cite{FIXME}.  Large deformation registration algorithms also reduce
+this confound by being less sensitive to the deformation distance
+between subject and target.  Some approaches combine both advantages,
+for instance, the diffeomorphic approach of Joshi et al. who employ the
+SSD metric and a shape distance to bring the subject group of images
+into alignment \cite{Joshi2004}.  Variants
 include extension to multiple modalities \cite{Lorenzen2006} and small deformations
@@
 Normalization (SyGN) is a powerful framework for producing optimal population-specific
-templates \cite{Avants2010}. Given a set of representative images, 
+templates \cite{Avants2010} with arbitrary similarity metric choice. Given a set of representative images, 
 $\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
@@
 involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
-updating of $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
+updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
 %For the interested reader, a set of scripts has been created and is distributed with the ANTs library for facilitating template creation.
@@
 \subsection{Anatomically-Based Population Normalization}
-
-An example of our proposed workflow applied to anatomically-based population normalization prior to TBSS analysis  is illustrated in Fig.
+We now detail a standard approach that uses SyGN and intra-subject inter-modality mapping to provide an unbiased strategy for FA normalization.
+An example of our proposed workflow applied to anatomically-based population normalization prior to FA analysis is illustrated in Fig.
 \ref{fig:workflow}.
@@
   \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \cite{Tustison2010} and Atropos $n$-tissue segmentation \cite{Avants2011a}).
-  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical
-        image ($\mathcal{R}^i_{dwi}$).
+  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
   \item Perform distortion correction on the FA to anatomical mapping 
@@
         \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
-  \item Warp each subject's FA or MD map to the template using $\mathcal{T}^i_{total}$.
+  \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.
+Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any other scalar derivable from DTI to the template.
+One other detail is that we prefer to warp the subject's DTI to the template and then, in template space, derive the scalar images of interest.
   \end{enumerate}      
@@
 \end{enumerate}
+The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender and age effects in a control population.
 
 \subsection{Imaging Data}
-
-Tensorial data associated with each of the following data sets were reconstructed from the diffusion weighted sequences using Camino%
+Diffusion tensor data associated with each of the following data sets were reconstructed from the diffusion weighted sequences using Camino%
 \footnote{
@@
 The following is a brief description relevant to our particular study.
-17 controls and 16 patients with TBI were used for the analysis presented in this
-work.  Each patient had a history of non-penetrating traumatic brain injury 
+17 controls and 16 patients with TBI were used for the analysis presented in this work.  Each patient had a history of non-penetrating traumatic brain injury 
 of at least moderate severity defined by significant and well-documented
 loss or alteration of consciousness following injury in addition to meeting
-several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and
-years of education.
-High resolution T1-weighted anatomic images were obtained using 3D MP-RAGE 
-imaging sequence using the following acquisition parameters: TR = 1620 ms, 
+several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and years of education.
+High resolution T1-weighted anatomic images were obtained using a 3D MP-RAGE 
+imaging sequence with the following acquisition parameters: TR = 1620 ms, 
 TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 
@@
 DTI images were also obtained. 
-
 %
@@
 } 
-was initiated
-on December 11, 2009 by various members of the MRI community \cite{Biswal2010}.
+was initiated on December 11, 2009 by various members of the MRI community \cite{Biswal2010}.
 Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
@@
 
-A patient template and a control template were  constructed 
-from the T1-weighted images of the two groups.  These two 
+We constructed a patient template and a control template from the T1-weighted images of the two groups.  These two 
 templates were then used to create a third common template.  
@@
 
-Statistical power curves (one-tailed, $\alpha = 0.95$)
+Cumulative distribution functions (one-tailed, $\alpha = 0.95$)
 are shown in Fig. \ref{fig:tbi_testing}.
@@
 trends towards better performance relative to the remaining 3 strategies.
-It should be noted that although TBSS uses
-the SSD metric, the fixed and moving images are smoothed and subsampled
-even at the highest resolution level which substantially decreases the local effect
-of the metric.
+It should be noted that although TBSS uses the SSD metric, the fixed and moving images are smoothed and subsampled even at the highest resolution level which substantially decreases the local effect of the metric.
 
@@
 
-The CDF of the voxelwise FA coefficient of variation of the aligned
+The CDF of the voxelwise FA coefficient of variation {\textcolor{red}{ FIXME define coefficient of variation and explain what it means --- how does it relate to variance?}} of the aligned
 images is shown for each case in Figure \ref{fig:tbi_cov}.  This is
@@
  one would expect a smaller variance across better aligned sets of voxel values thus 
- producing a lower CoV for a given mean value.  By calculating the cumulative distribution 
- function (CDF) of the set of CoV values within the mask, quantitative assessment of alignment performance can be visualized.  
+ producing a lower CoV for a given mean value.  By calculating the CDF of the set of CoV values within the mask, quantitative assessment of alignment performance can be visualized.  
 % In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
@@
 Avoid extensive citations and discussion of published literature.
+
+\subsection{Discussion} Prior evaluation work using anatomically labeled
+data \cite{bunch_of_stuff} indicates that mutual information,
+cross-correlation and the SSD/Demons metric are all capable of producing
+high quality anatomical alignment.  All of this work, though, compares
+the quality of alignment at a relatively coarse scale of major gyri,
+lobes and regions.  At the voxel level, it is nearly impossible to
+determine the ground truth correspondence.  Indeed, there is no reason
+to believe that subtle differences between a set of alignment strategies
+(induced, in the experiments of this paper, by changing the similarity
+metrics used in image registration) that all apparently work well should
+be considered significant.
+This work, however, highlights a dramatic impact of similarity metric choice on detection power in template-based FA studies.
+Our contention is that this dramatic "improvement" in detection power is not due to better anatomical alignment.
+Rather, it is a result of circularity/non-independence of the normalization and statistical estimation strategy.
+Consequently, we recommend that future work should take greater care when pairing the normalization strategy with the statistical framework.
+Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.
+That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}
 

### 2011-06-07 ANTsTBSS/paper.tex
ENH: one specific sentence in discussion

@@
 Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.
-That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}
+That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}  In particular, one should avoid combining the SSD metric for normalizing FA with Student's t-test for FA differences.
 

### 2011-06-07 ANTsTBSS/paper.tex
ENH: one specific sentence in discussion

--- 2011-06-07 ANTsTBSS/paper.tex
@@
 Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.
-That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}
+That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.}}}  In particular, one should avoid combining the SSD metric for normalizing FA with Student's t-test for FA differences.
 

### 2011-06-07 ANTsTBSS/paper.tex
Minor edits.

--- 2011-06-07 ANTsTBSS/paper.tex
@@
 in comparison to FA-based normalization methods.  We establish this
-outcome by using open data in normal controls from the International
+outcome by using open data comprised of normal controls from the International
 Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
@@
 } 
-VBM\cite{Ashburner2001} continues to find application to FA studies\cite{Kakeda2010,Takao2010} despite concerns about its validity\cite{Bookstein2001,Davatzikos2004} in DTI \cite{Jones2005,Chung2008}.
+VBM \cite{Ashburner2001} continues to find application to FA studies \cite{Kakeda2010,Takao2010} despite concerns about its validity \cite{Bookstein2001,Davatzikos2004} in DTI \cite{Jones2005,Chung2008}.
 The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.
@@
 The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and an image gradient) and is therefore widely used.
-Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and provide reasonable performance levels \cite{Klein2009}.
+Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and produce reasonable performance levels \cite{Klein2009}.%
 \footnote{
 In addition to TBSS and SPM, other popular packages in the neuroimaging community 
-which use the SSD metric includes AIR \cite{Woods1998} and LDDMM \cite{Beg2005}.
-}  
-Now, suppose the SSD metric is optimized to find a set of $M$ transformations that map a population of $M$ FA images, $\{I_1, I_2, \ldots, I_M\}$, to a representative FA template, $J$.
-Then, the algorithm will compute transformations
+which use the SSD metric include AIR \cite{Woods1998} and LDDMM \cite{Beg2005}.
+}
+
+However, suppose the SSD metric is optimized to find a set of $M$ transformations that map a population of $M$ FA images, $\{I_1, I_2, \ldots, I_M\}$, to a representative FA template, $J$.
+The algorithm will then compute the transformations,
 $\boldsymbol{\mathcal{T}} = \{\mathcal{T}_1, \mathcal{T}_2,\ldots,\mathcal{T}_M \}$, which minimize the SSD metric over the population, i.e.
@@
 Eqn.
-\ref{eq:variance}, direct FA-to-FA registration will find transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's test).
+\ref{eq:variance}, direct FA-to-FA registration will find transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
+$t$-test).
 While the SSD metric is able to provide visually attractive alignments across subjects, the voxel-level details (that is, which voxels are matched where) are, as we show in the following simulation study, strikingly important in terms of the effect on detection power.
@@
 rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).
-The CDF vertical axis shows the cumulative number of voxels with corrected p-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
+The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
 From here, we denote these graphs as "statistical significance curves."  These curves are plotted in black in Fig.
@@
 This effect can be explained in the context of Eqn.
-(\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the t-test).
+(\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the $t$-test).
 Note, however, that this effect is increased by the fact that noise was added to the data.
-{\textcolor{red}{FIXME --- does more / less noise increase / decrease the exaggerated power?
-}}
+Decreasing the standard deviation of the noise causes an increase in separation between  the SSD and Demon's metric pair with the MI and CC metric pair with the latter pair closer to the the true significance curve.
+Increasing the noise yields the same ranking with a more even spread between significance curves.
+
 Common similarity metric functions can therefore induce false positives by minimizing local intensity 
@@
 the $t$-test results given in Fig. \ref{fig:simulated_testing}(a) with the non-parametric Wilcoxon
-rank-sum testing (Fig. \ref{fig:simulated_testing}(b)).  The rank test does not explicitly use variance in its assessment and thus exhibits far less statistical exaggeration.
+rank-sum testing (Fig.
+\ref{fig:simulated_testing}(b)).
+The rank test does not explicitly use variance in its assessment and thus exhibits far less statistical exaggeration even though it offers roughly 95\%
+the power of the $t$-test assuming certain assumptions are met \cite{Siegel1988,Rorden2007}.
 
@@
 these issues by deriving a most representative image from the population
-\cite{FIXME}.  Large deformation registration algorithms also reduce
+\cite{Good2001}.  Large deformation registration algorithms also reduce
 this confound by being less sensitive to the deformation distance
@@
 
-The CDF of the voxelwise FA coefficient of variation {\textcolor{red}{ FIXME define coefficient of variation and explain what it means --- how does it relate to variance?}} of the aligned
-images is shown for each case in Figure \ref{fig:tbi_cov}.  This is
-very much related to earlier findings in that strategies evincing
-greater alignment, as quantified by the coefficient of variation,
-will trend towards smaller variance thus exhibiting a left-shifted
-CDF.  Although direct FA-to-FA registration using ANTs with common
+The CDF of the voxelwise FA coefficient of variation of the aligned
+images is shown for each case in Figure \ref{fig:tbi_cov}.  
+The coefficient of variation at each voxel is defined as the variance
+divided by the mean over the population of FA values.  Thus,
+those metrics which best minimize the voxelwise variance would
+trend towards greater alignment as quantified by the coefficient of
+variation.
+
+%This is very much related to earlier findings in that strategies evincing
+%greater alignment, as quantified by the coefficient of variation,
+%will trend towards smaller variance thus exhibiting a left-shifted
+%CDF.
+
+Although direct FA-to-FA registration using ANTs with common
 similarity metrics gives ``better alignment'' such alignment is 
@@
 to the MNI152 (FMRIB58) template via the template-to-MNI152 transformation.
-
-\begin{figure}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{NKIplots3.pdf}
-\end{tabular}
-\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
-\ref{fig:NKI}(b)).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-}
-\label{fig:NKIplots}
-\end{figure}
 
@@
 
-\subsubsection{Statistical Power Related to Gender-based FA Differences}
+\begin{figure}
+\begin{tabular}{c}
+  \includegraphics[width=85mm]{NKIplots3.pdf}
+\end{tabular}
+\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
+\ref{fig:NKI}(b)).
+The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
+}
+\label{fig:NKIplots}
+\end{figure}
+
+\subsubsection{Statistical Power Related to Gender-Based FA Differences}
 
@@
 
-\subsection{Discussion} Prior evaluation work using anatomically labeled
-data \cite{bunch_of_stuff} indicates that mutual information,
+\section{Discussion} Prior evaluation work using anatomically labeled
+data indicates that mutual information,
 cross-correlation and the SSD/Demons metric are all capable of producing
-high quality anatomical alignment.  All of this work, though, compares
+high quality anatomical alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
 the quality of alignment at a relatively coarse scale of major gyri,
@@
 Our contention is that this dramatic "improvement" in detection power is not due to better anatomical alignment.
-Rather, it is a result of circularity/non-independence of the normalization and statistical estimation strategy.
+Rather, it is a result of circularity/nonindependence of the normalization and statistical estimation strategy.
 Consequently, we recommend that future work should take greater care when pairing the normalization strategy with the statistical framework.
 Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.
-That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}
+That is, {the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.
+In particular, one should avoid combining the SSD metric for normalizing FA with Student's t-test for FA differences.
 

### 2011-06-08 ANTsTBSS/paper.tex
to jpg
Minor edits.

@@
 } 
-VBM \cite{Ashburner2001} continues to find application to FA studies \cite{Kakeda2010,Takao2010} despite concerns about its validity \cite{Bookstein2001,Davatzikos2004} in DTI \cite{Jones2005,Chung2008}.
+VBM \cite{Ashburner2001} continues to find application to FA studies \cite{Kakeda2010,Takao2010} despite concerns about its general validity \cite{Bookstein2001,Davatzikos2004} and with respect to DTI \cite{Jones2005,Chung2008}.
 The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.
@@
 
-Although there are substantive differences in normalization schemes
-between these and other image registration approaches, a core
+Although there are substantive differences in normalization
+between these and other analysis approaches, a core
 commonality includes direct FA-to-FA registration.
-The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and an image gradient) and is therefore widely used.
-Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and produce reasonable performance levels \cite{Klein2009}.%
+The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and gradient) and is therefore widely used.
+Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and yield reasonable performance levels \cite{Klein2009}.%
 \footnote{
@@
 
-However, suppose the SSD metric is optimized to find a set of $M$ transformations that map a population of $M$ FA images, $\{I_1, I_2, \ldots, I_M\}$, to a representative FA template, $J$.
-The algorithm will then compute the transformations,
+However, suppose the SSD metric is used to find a set of $M$ transformations that map a population of $M$ FA images, $\{I_1, I_2, \ldots, I_M\}$, to a representative FA template, $J$, for statistical testing between groups.
+This results in the transformations,
 $\boldsymbol{\mathcal{T}} = \{\mathcal{T}_1, \mathcal{T}_2,\ldots,\mathcal{T}_M \}$, which minimize the SSD metric over the population, i.e.
@@
 Eqn.
-\ref{eq:variance}, direct FA-to-FA registration will find transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
+\ref{eq:variance}, direct FA-to-FA registration will produce transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
 $t$-test).
@@
   aligned prior to invoking ANTs registration, deformation is driven by the similarity metric
-  causing the accumulation of type 1 errors.}
+  causing the accumulation of Type 1 errors.}
 \label{fig:simulated_data_warped}
@@
 The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
-From here, we denote these graphs as "statistical significance curves."  These curves are plotted in black in Fig.
-\ref{fig:simulated_testing}.
+We denote these graphs as "statistical significance curves."
+The true curves for the simulated experiment are plotted in black in Fig. \ref{fig:simulated_testing}.  
+We note that the amount of smoothing increases the proportion of significant voxels beyond
+the prescribed 10\%.
+
 To illustrate the 
@@
 This effect can be explained in the context of Eqn.
-(\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the $t$-test).
-Note, however, that this effect is increased by the fact that noise was added to the data.
-Decreasing the standard deviation of the noise causes an increase in separation between  the SSD and Demon's metric pair with the MI and CC metric pair with the latter pair closer to the the true significance curve.
-Increasing the noise yields the same ranking with a more even spread between significance curves.
+(\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the $t$-test statistics).
+Whereas  decreasing the standard deviation of the noise causes an increase in separation between  the SSD and Demon's metric pair with the MI and CC metric pair with the latter pair closer to the the true significance curve, increasing the noise yields the same ranking with a more even spread between significance curves.
 
@@
 \end{tabular}
-\caption{Possible ANTs-based TBSS workflow.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
+\caption{Possible ANTs-based TBSS workflow using SyGN.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
 (transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.
@@
 between subject and target.  Some approaches combine both advantages,
-for instance, the diffeomorphic approach of Joshi et al. who employ the
+for instance, the diffeomorphic approach of Joshi et al. employs the
 SSD metric and a shape distance to bring the subject group of images
@@
 We now detail a standard approach that uses SyGN and intra-subject inter-modality mapping to provide an unbiased strategy for FA normalization.
-An example of our proposed workflow applied to anatomically-based population normalization prior to FA analysis is illustrated in Fig.
+An example of our proposed workflow applied to anatomically-based population normalization prior to FA analysis using TBSS is illustrated in Fig.
 \ref{fig:workflow}.
@@
   \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.  
-Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any other scalar derivable from DTI to the template.
-One other detail is that we prefer to warp the subject's DTI to the template and then, in template space, derive the scalar images of interest.
+  Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any 
+  other scalar derivable from DTI to the template.  
+%  One other detail is that we prefer to warp the subject's DTI to 
+%  the template and then, in template space, derive the scalar images of interest.  
   \end{enumerate}      
@@
 Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
-Commitments from institutions such as they NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
+Commitments from institutions such as the NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
 currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC).%
@@
 
-The CDF of the voxelwise FA coefficient of variation of the aligned
+The CDF of the voxelwise FA coefficient of variation (CV) of the aligned
 images is shown for each case in Figure \ref{fig:tbi_cov}.  
-The coefficient of variation at each voxel is defined as the variance
+The CV at each voxel is defined as the variance
 divided by the mean over the population of FA values.  Thus,
 those metrics which best minimize the voxelwise variance would
-trend towards greater alignment as quantified by the coefficient of
-variation.
+trend towards greater alignment as quantified by the CV.
 
 %This is very much related to earlier findings in that strategies evincing
-%greater alignment, as quantified by the coefficient of variation,
+%greater alignment, as quantified by the CV,
 %will trend towards smaller variance thus exhibiting a left-shifted
@@
 similarity metrics gives ``better alignment'' such alignment is 
-produced at a cost of significant type 1 errors.  Common subsampling and
+produced at a cost of significant Type 1 errors.  Common subsampling and
 smoothing strategies and local neighborhood metrics such as CC and MI reduce the 
@@
 The skeleton masks were produced by thresholding the skeleton values at 
-FA $\geq 0.2$ and it would seem that the improved alignment produces 
+FA $\geq 0.2$ and it would appear that the improved alignment produces 
 a more extensive skeletal network for greater localization testing of 
@@
 \end{tabular}
-\caption{Comparison of the different alignment strategies in normalizing the TBI data to the FMRIB58 template.
-Assessment is quantitated using the voxelwise coefficient of variation of the FA.
-TBSS and the 4 ANTs metric
+\caption{Comparison of the different alignment strategies in normalizing the TBI data to the FMRIB58 template.  Assessment is quantitated using the voxelwise CV of the FA.  TBSS and the 4 ANTs metric
 results were all obtained using direct registration involving subject FA and the FMRIB58 template.
@@
 images were registered to the resulting template and, ultimately, 
-to the MNI152 (FMRIB58) template via the template-to-MNI152 transformation.
+to the MNI152 (FMRIB58) template via the template-to-MNI152 transformations.
 
@@
 We also include the same axial slice  of the FMRIB58 template in Fig. \ref{fig:NKI} 
-(a) for comparison.  In both sets of sample slices, visual inspection would seemingly 
+(a) for comparison.  In both sets of sample slices, visual inspection would  
 indicate better 
@@
  This binary mask obtained in the NKI cohort provided the standard group mask for 
- fair comparisons for all evaluations.    For all FMRIB58-aligned images the coefficient of 
- variation (CoV) was calculated in a voxel-wise fashion within the mask.  Intuitively, 
+ fair comparisons for all evaluations.    For all FMRIB58-aligned images, the CV 
+ was calculated in a voxelwise fashion within the mask.  Intuitively, 
  one would expect a smaller variance across better aligned sets of voxel values thus 
- producing a lower CoV for a given mean value.  By calculating the CDF of the set of CoV values within the mask, quantitative assessment of alignment performance can be visualized.  
+ producing a lower CV for a given mean value.  By calculating the CDF of the set of CV values within the mask, quantitative assessment of alignment performance can be visualized.  
 % In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
@@
 There is little difference between the ANTs results with only a slight trend in improved performance with the larger template subject sample sizes.
-%This was the case for all three regional CoV calculations for both data sets.
+%This was the case for all three regional CV calculations for both data sets.
 Also, as expected, there is greater misalignment in the outer region versus the inner core of the white matter mask regardless of methodology.
@@
 %\end{tabular}
-%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
+%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
 \ref{fig:NKI}(b)).
-In addition to the whole brain analysis (first column), the coefficient of variation was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
+In addition to the whole brain analysis (first column), the CV was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
 The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
@@
 \end{tabular}
-\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the coefficient of variation at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
+\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
 \ref{fig:NKI}(b)).
@@
 data indicates that mutual information,
-cross-correlation and the SSD/Demons metric are all capable of producing
+cross-correlation, Demons, and SSD metrics are all capable of producing
 high quality anatomical alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
@@
 That is, {the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.
-In particular, one should avoid combining the SSD metric for normalizing FA with Student's t-test for FA differences.
+In particular, one should avoid combining the SSD metric for normalizing FA with Student's $t$-test for FA differences.
 
@@
 minimized thus mitigating the confounds described previously.
-Each DTI-based image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \cite{Wu2008}.
-\item Although the total number of registrations to be performed is increased, this makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
+Each DTI-based scalar image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \cite{Wu2008}.
+\item Although the total number of registrations to be performed is increased, this strategy makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
 %\item Instead of Gaussian smoothing as with SPM, TBSS uses the warping process to FMRIB58 by default to provide an indirect smoothing.
@@
 
-Of practical significance, we emphasize that all software components described 
+All software components described 
 in this paper are publicly available as open source primarily through the 

### 2011-06-08 ANTsTBSS/paper.tex
Minor edit.
ENH: adding the caveats paragraph

@@
 
-\title{Normalization Circularity in Population-Based FA Analysis}
+\title{On Population-Based FA Analysis with Non-independent Normalization and Statistics}
 
@@
 significance.  Consequently, we advocate a more traditional approach
-that normalizes a population of FA images through the associated T1
+that normalizes a population of FA images through the associated 
 anatomical images.  We show that the Symmetric Group Normalization
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedData.pdf}
+  \includegraphics[width=85mm]{simulatedData.jpg}
 \end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedDataWarped.pdf}
+  \includegraphics[width=85mm]{simulatedDataWarped.jpg}
 \end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{simulatedTesting.pdf}
+  \includegraphics[width=180mm]{simulatedTesting.jpg}
 \end{tabular}
@@
 After creation of the simulated data set, we performed voxelwise testing (both $t$-testing
-and Wilcoxon rank-sum testing, one-tailed, $\alpha = 0.95$) with false discovery
-rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).
+and Wilcoxon rank-sum testing, aka the Mann-Whitney test, one-tailed,
+$\alpha = 0.95$) with false discovery rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).
 The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
@@
 
-To illustrate the 
-effect of similarity metric on statistical significance, we performed ANTs registration \cite{Avants2011} of each subject to the template
+To illustrate the effect of similarity metric on statistical significance, we performed ANTs registration \cite{Avants2011} of each subject to the template
 (Fig. \ref{fig:simulated_data}(b)) using cross correlation (CC), mutual information (MI), 
@@
 
-This example demonstrates the undesirable phenomenon previously described.
-Even though these images were anatomically aligned prior to registration, the resulting deformations (falsely) increased detection power.
-This effect can be explained in the context of Eqn.
-(\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the $t$-test statistics).
-Whereas  decreasing the standard deviation of the noise causes an increase in separation between  the SSD and Demon's metric pair with the MI and CC metric pair with the latter pair closer to the the true significance curve, increasing the noise yields the same ranking with a more even spread between significance curves.
+This example demonstrates the undesirable phenomenon previously
+described.  Even though these images were anatomically aligned prior
+to registration, the resulting deformations (falsely) increased
+detection power.  This effect can be explained in the context of
+Eqn. (\ref{eq:variance}) and by the fact that we added noise to the
+underlying distribution which allows the deformable mapping algorithms
+the freedom to transform the positions of the most "powerful" voxels
+to minimize variance across the population (and increase the $t$-test
+statistics).  Although we do not show the results, additional
+experiments show that decreasing the standard deviation of the noise
+causes an increase in separation between the SSD/Demon's metrics and
+the MI and CC metrics.  MI and CC are closer to the the true
+significance curve.  Alternatively, if we increase the noise level,
+the same ranking occurs with even more spread between significance curves.
 
@@
 \begin{tabular}{c}
-  \includegraphics[width=170mm]{workflow.pdf}
+  \includegraphics[width=170mm]{workflow.jpg}
 \end{tabular}
@@
 
-\section{Results and Discussion}
+\section{Results}
 %Results should be clear and concise.
@@
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{tbiTesting.pdf}
+  \includegraphics[width=180mm]{tbiTesting.jpg}
 \end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{tbiCoV.pdf}
+  \includegraphics[width=85mm]{tbiCoV.jpg}
 \end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{meanFA.pdf}
+  \includegraphics[width=85mm]{meanFA.jpg}
 \end{tabular}
@@
 %\begin{tabular}{c}
-%  \includegraphics[width=80mm]{hoonData2.pdf}
+%  \includegraphics[width=80mm]{hoonData2.jpg}
 %\end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{NKI2.pdf}
+  \includegraphics[width=85mm]{NKI2.jpg}
 \end{tabular}
@@
 %\begin{tabular}{c}
-%  \includegraphics[width=80mm]{NYU.pdf}
+%  \includegraphics[width=80mm]{NYU.jpg}
 %\end{tabular}
@@
 %\begin{tabular}{c}
-%  \includegraphics[width=180mm]{NYUplots2.pdf}
+%  \includegraphics[width=180mm]{NYUplots2.jpg}
 %\end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{NKIplots3.pdf}
+  \includegraphics[width=85mm]{NKIplots3.jpg}
 \end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{nkiTesting.pdf}
+  \includegraphics[width=180mm]{nkiTesting.jpg}
 \end{tabular}
@@
 \section{Discussion} Prior evaluation work using anatomically labeled
-data indicates that mutual information,
-cross-correlation, Demons, and SSD metrics are all capable of producing
-high quality anatomical alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
+data indicates that mutual information, cross-correlation, Demons, and
+SSD metrics are all capable of producing high quality anatomical
+alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
 the quality of alignment at a relatively coarse scale of major gyri,
@@
 determine the ground truth correspondence.  Indeed, there is no reason
-to believe that subtle differences between a set of alignment strategies
-(induced, in the experiments of this paper, by changing the similarity
-metrics used in image registration) that all apparently work well should
-be considered significant.
-This work, however, highlights a dramatic impact of similarity metric choice on detection power in template-based FA studies.
-Our contention is that this dramatic "improvement" in detection power is not due to better anatomical alignment.
-Rather, it is a result of circularity/nonindependence of the normalization and statistical estimation strategy.
-Consequently, we recommend that future work should take greater care when pairing the normalization strategy with the statistical framework.
-Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.
-That is, {the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.
-In particular, one should avoid combining the SSD metric for normalizing FA with Student's $t$-test for FA differences.
+to believe that subtle differences between a set of alignment
+strategies (induced, in the experiments of this paper, by changing the
+similarity metrics used in image registration) that all apparently
+work well should be considered significant.  This work, however,
+highlights a dramatic impact of similarity metric choice on detection
+power in template-based FA studies.  Our contention is that this
+dramatic "improvement" in detection power is not due to better
+anatomical alignment.  Rather, it is a result of
+circularity/nonindependence of the normalization and statistical
+estimation strategy.  Consequently, we recommend that future work
+should take greater care when pairing the normalization strategy with
+the statistical framework.  Furthermore, we recommend that researchers
+should diligently seek to maintain independence between these two
+critical stages of a population study.  That is, the features that
+drive the minimization of a similarity metric should be independent of
+the features that will be used in hypothesis testing.  In particular,
+one should avoid combining the SSD metric for normalizing FA with
+Student's $t$-test for FA differences.  A final point is that
+cumulative distribution functions and other power assessment methods
+cannot be regarded as meaningful comparison techniques in the presence
+of normalization/statistical non-independence.  Similar issues are
+raised in \cite{Kriegeskorte2010}.
+
+Caveats with our analysis include a lack of ground truth in the
+clinical component of the study.   Additionally, we are not addressing
+our results' biological plausibility, nor do we
+intend to (in this paper) enter into discussion on the mechanisms of
+TBI, gender difference and DTI that may lead to the detected results.
+Rather, we focus on the technical issue of circularity induced by
+normalizing the image feature of interest with functions that maximize
+the statistic used in hypothesis testing.  We also cannot argue that our analysis
+invalidates previous results that used a circular strategy.  Instead, we encourage the community to
+reexamine FA normalization and analysis methods and to consider the
+potential confounds highlighted in this paper.  Rank tests
+provide a more conservative alternative as does the methodology
+presented here based on T1 to DT intrasubject mapping and T1
+normalization.
 

### 2011-06-08 ANTsTBSS/paper.tex
ENH: adding why the paper hasnt been written before sentences

@@
 
-\section{Discussion} Prior evaluation work using anatomically labeled
+\section{Discussion} 
+Prior evaluation work using anatomically labeled
 data indicates that mutual information, cross-correlation, Demons, and
@@
 presented here based on T1 to DT intrasubject mapping and T1
-normalization.
+normalization.  To our surprise, we could not find a prior publication that directly
+addresses this issue.  Two reasons may be that the majority of
+technical work using DTI has focused on more cutting-edge issues while
+the clinical work using DTI is more concerned with generating
+statistically significant results.  An additional reason may be that
+only recently have toolkits emerged to enable easy comparison of
+statistical techniques (via {\bf R}), normalization methods (ANTs) and
+diffusion tensor processing (Camino).
 

### 2011-06-08 ANTsTBSS/paper.tex
ENH: adding why the paper hasnt been written before sentences

@@
 We also gratefully acknowledge Dr. Niels van Strien of the Norwegian University of Science and Technology
-who packaged the template construction algorithm in the very useful script \verb#buildtemplateparallel.sh#
+who assisted in packaging the template construction algorithm in the very useful script \verb#buildtemplateparallel.sh#
 which is publicly available in ANTs.

### 2011-06-08 ANTsTBSS/paper.tex
First draft submission.

@@
 
-%\documentclass[preprint,authoryear,review,12pt]{elsarticle}
-\documentclass[final,5p,times,twocolumn]{elsarticle}
+\documentclass[preprint,authoryear,review,12pt]{elsarticle}
+%\documentclass[final,5p,times,twocolumn]{elsarticle}
 
@@
 
-\title{On Population-Based FA Analysis with Non-independent Normalization and Statistics}
+\title{Circularity in Population-Based FA Analysis}
 
@@
 anatomical alignment, and consistently overestimates statistical
-significance.  Consequently, we advocate a more traditional approach
+significance.  Consequently, we advocate a more cautious approach
 that normalizes a population of FA images through the associated 
@@
 of the open source Advanced Normalization Tools (ANTs) repository.
-In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the ultimate statistical functions used to assess population effects.
+In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the statistical functions used to assess population effects.
 
@@
 \begin{keyword}
-Advanced Normalization Tools (ANTs) \sep diffusion tensor imaging \sep fractional anisotropy \sep template construction \sep traumatic brain injury
+Advanced Normalization Tools (ANTs) \sep diffusion tensor imaging \sep fractional anisotropy \sep template construction
 %% keywords here, in the form: keyword \sep keyword
@@
 The seminal work of Basser et al.
-\cite{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique.
-DTI's sensitivity to brain architecture \cite{Basser1996,Assaf2008} enables multiple uses in assessing neuro-structural differences in cross-population studies \cite{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010} including subjects with traumatic brain injury (TBI) \cite{Caeyenberghs2010,Jiang2010,Warner2010}.
+\citep{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique.
+DTI's sensitivity to brain architecture \citep{Basser1996,Assaf2008} enables promising analysis possibilities assessing neuro-structural differences in cross-population studies \citep{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010} including subjects with traumatic brain injury (TBI) \citep{Caeyenberghs2010,Jiang2010,Warner2010}.
 Two popular comprehensive software packages for assessing population FA differences include the SPM Matlab-based toolkit%
@@
 for voxel-based morphometry (VBM)
-and the tract-based spatial statistics (TBSS) framework \cite{Smith2006}.%
+and the tract-based spatial statistics (TBSS) framework \citep{Smith2006}.%
 \footnote{
@@
 } 
-VBM \cite{Ashburner2001} continues to find application to FA studies \cite{Kakeda2010,Takao2010} despite concerns about its general validity \cite{Bookstein2001,Davatzikos2004} and with respect to DTI \cite{Jones2005,Chung2008}.
+VBM \citep{Ashburner2001} continues to find application to FA studies \citep{Kakeda2010,Takao2010} despite concerns about its general validity \citep{Bookstein2001,Davatzikos2004} and with respect to DTI \citep{Jones2005,Chung2008}.
 The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.
-%(e.g. \cite{Arnone2008,Bodini2009}).
+%(e.g. \citep{Arnone2008,Bodini2009}).
 
@@
 The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and gradient) and is therefore widely used.
-Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and yield reasonable performance levels \cite{Klein2009}.%
+Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and yield reasonable performance levels \citep{Klein2009}.%
 \footnote{
 In addition to TBSS and SPM, other popular packages in the neuroimaging community 
-which use the SSD metric include AIR \cite{Woods1998} and LDDMM \cite{Beg2005}.
+which use the SSD metric include AIR \citep{Woods1998} and LDDMM \citep{Beg2005}.
 }
@@
 anatomical alignment with optimizing the very statistical testing
-results that one will ultimately use to assess hypotheses.  This immediately evokes a sense of circularity in the analysis \cite{Kriegeskorte2010}.
+results that one will ultimately use to assess hypotheses.  This immediately evokes a sense of circularity in the analysis \citep{Kriegeskorte2010}.
 According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on data that were selected for showing the effect of interest or a related effect.''  As illustrated with
@@
 \begin{figure}
+\begin{center}
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedData.jpg}
+  \includegraphics[width=85mm]{simulatedData.pdf}
 \end{tabular}
@@
 \label{fig:simulated_data}
+\end{center}
 \end{figure}
@@
 \begin{figure}
+\begin{center}
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedDataWarped.jpg}
+  \includegraphics[width=85mm]{simulatedDataWarped.pdf}
 \end{tabular}
@@
 \label{fig:simulated_data_warped}
+\end{center}
 \end{figure}
@@
 Axial slice number 159 was extracted from \verb#subject_04# of the discrete BrainWeb database
-\cite{Aubert-Broche2006} which is shown in Fig. \ref{fig:simulated_data}.  After thresholding out 
+\citep{Aubert-Broche2006} which is shown in Fig. \ref{fig:simulated_data}.  After thresholding out 
 the white matter and assigning all white matter voxels a value of 0.7, 10\% of these white 
@@
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{simulatedTesting.jpg}
+%  \includegraphics[width=180mm]{simulatedTesting.pdf}
+  \includegraphics[width=130mm]{simulatedTesting.pdf}
 \end{tabular}
@@
          metrics for both tests although the effect is less in the more robust Wilcoxon results.
-Not shown are the significance curves for the Brunner-Munzel testing \cite{Rorden2007} which showed similar effects to the results on the left.
+Not shown are the significance curves for the Brunner-Munzel testing \citep{Rorden2007} which showed similar effects to the results on the left.
         }
@@
 
-To illustrate the effect of similarity metric on statistical significance, we performed ANTs registration \cite{Avants2011} of each subject to the template
+To illustrate the effect of similarity metric on statistical significance, we performed ANTs registration \citep{Avants2011} of each subject to the template
 (Fig. \ref{fig:simulated_data}(b)) using cross correlation (CC), mutual information (MI), 
-Demons \cite{Thirion1998}, and sum of squared differences (SSD). A representative deformed image
+Demons \citep{Thirion1998}, and sum of squared differences (SSD). A representative deformed image
 using the SSD metric and its corresponding deformed grid is given in 
@@
 
-Common similarity metric functions can therefore induce false positives by minimizing local intensity 
+Common similarity metric functions can therefore cause false positives by minimizing local intensity 
 variance between the subjects and template which explicitly maximizes statistical
@@
 The rank test does not explicitly use variance in its assessment and thus exhibits far less statistical exaggeration even though it offers roughly 95\%
-the power of the $t$-test assuming certain assumptions are met \cite{Siegel1988,Rorden2007}.
+the power of the $t$-test assuming certain assumptions are met \citep{Siegel1988,Rorden2007}.
 
@@
 %We recognize that using the anatomical images in the alignment process has been previously proposed (in fact, such a suggestion occurs in the {\it Future directions} section of the original TBSS paper).
-of a high performance \cite{Klein2009}, nonrigid registration algorithm \cite{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \cite{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
+of a high performance \citep{Klein2009}, nonrigid registration algorithm \citep{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \citep{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
 The key point, here, is that the FA images are mapped to the common reference space via intra-subject DT-to-T1 mapping followed by T1-to-T1 alignment.
@@
 \begin{tabular}{c}
-  \includegraphics[width=170mm]{workflow.jpg}
+%  \includegraphics[width=170mm]{workflow.pdf}
+  \includegraphics[width=130mm]{workflow.pdf}
 \end{tabular}
@@
 %through to the statistical analysis for discerning differences across populations.
-Major intermediate steps include the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA) \cite{Basser1996}), nonrigid alignment of the population to a standard coordinate system, and subsequent ``skeletonization'' of the mean of the aligned FA maps.
+Major intermediate steps include the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA) \citep{Basser1996}), nonrigid alignment of the population to a standard coordinate system, and subsequent ``skeletonization'' of the mean of the aligned FA maps.
 The alignment to a standard space is performed in such a way to remove the smoothing step associated with SPM.
 Skeletonization and its projection across all registered DTI-derived scalar images is meant to accommodate the possibility of imperfectly aligned data.
-However, as shown in a recent study \cite{Zalesky2011}, although the alignment error is reduced with the skeleton projection step, misalignments still remain post-projection.
+However, as shown in a recent study \citep{Zalesky2011}, although the alignment error is reduced with the skeleton projection step, misalignments still remain post-projection.
 Added to this is the potential confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.
@@
 
-%A crucial component of TBSS is the alignment of the DTI-derived scalar images \cite{Zalesky2011}. 
+%A crucial component of TBSS is the alignment of the DTI-derived scalar images \citep{Zalesky2011}. 
 %
@@
 %
-%Talk about {\em skeleton projection} and how our work is different from \cite{Zalesky2011}.
+%Talk about {\em skeleton projection} and how our work is different from \citep{Zalesky2011}.
 %Also talk about the advantages over building a dti-based template as discussed in 
-%\cite{Hecke2010}.
+%\citep{Hecke2010}.
 
@@
 of a pre-existing template based on a single subject, e.g. the Talairach
-atlas \cite{Talairach1988}), or a publicly available averaged group of
-subjects, e.g. the MNI \cite{Collins1994} or ICBM \cite{Mazziotta1995}
+atlas \citep{Talairach1988}, or a publicly available averaged group of
+subjects, e.g. the MNI \citep{Collins1994} or ICBM \citep{Mazziotta1995}
 templates.
@@
 these issues by deriving a most representative image from the population
-\cite{Good2001}.  Large deformation registration algorithms also reduce
+\citep{Good2001}.  Large deformation registration algorithms also reduce
 this confound by being less sensitive to the deformation distance
@@
 SSD metric and a shape distance to bring the subject group of images
-into alignment \cite{Joshi2004}.  Variants
-include extension to multiple modalities \cite{Lorenzen2006} and small deformations
-\cite{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
-towards a representative image template \cite{Learned-Miller2006}.
+into alignment \citep{Joshi2004}.  Variants
+include extension to multiple modalities \citep{Lorenzen2006} and small deformations
+\citep{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
+towards a representative image template \citep{Learned-Miller2006}.
 
@@
 normalized space during optimization.  Coupling the intrinsic symmetry of 
-SyN pairwise registration \cite{Avants2011} and an
+SyN pairwise registration \citep{Avants2011} and an
 optimized sharpening/averaging of the template appearance, Symmetric Group
 Normalization (SyGN) is a powerful framework for producing optimal population-specific
-templates \cite{Avants2010} with arbitrary similarity metric choice. Given a set of representative images, 
+templates \citep{Avants2010} with arbitrary similarity metric choice. Given a set of representative images, 
 $\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
@@
 \end{align}
-$\Pi$ is the choice of similarity metric, often cross-correlation \cite{Avants2008a}, calculated in the 
+$\Pi$ is the choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
 virtual domain midway between each individual image and the current estimate of the template.
@@
   \begin{enumerate}
-  \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \cite{Tustison2010} and Atropos $n$-tissue segmentation \cite{Avants2011a}).
+  \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \citep{Tustison2010} and Atropos $n$-tissue segmentation \citep{Avants2011a}).
   \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
@@
 \end{enumerate}
-The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender and age effects in a control population.
+The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender effects in a control population.
 
@@
 http://camino.org.uk/
-}---an open source toolkit for diffusion MRI processing and analysis \cite{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.
+}---an open source toolkit for diffusion MRI processing and analysis \citep{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.
 
@@
 } 
-was initiated on December 11, 2009 by various members of the MRI community \cite{Biswal2010}.
+was initiated on December 11, 2009 by various members of the MRI community \citep{Biswal2010}.
 Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
@@
 FSL hosts a large number of algorithms for computational neuroanalysis.
-These include, among others, well-known tools for image segmentation (FAST---\cite{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), and TBSS and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).
+These include, among others, well-known tools for image segmentation (FAST---\citep{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), and TBSS and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).
 Based solely on the number of citations in the literature,%
@@
 %  \item[tbss\_2\_reg]  The organized FA data and corresponding masks produced in the previous step are warped to one of three possible coordinate systems using FLIRT/ FNIRT:  1)  the atlas included with the FSL download---FMRIB58---(recommended), 2) a user-specified target image or 3) an exhaustive search for the ``optimal'' target image.
-In \cite{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.
+In \citep{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.
 However, even though it would be faster to simply choose a single subject at random, the complexities of nonrigid image registration would seemingly promote the exhaustive search over random selection.
@@
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{tbiTesting.jpg}
+%  \includegraphics[width=180mm]{tbiTesting.pdf}
+  \includegraphics[width=130mm]{tbiTesting.pdf}
 \end{tabular}
@@
 \begin{table}
+\begin{center}
   \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
   \label{table:lom}
-  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccccc}
+  \begin{tabular*}{130mm}{@{\extracolsep{\fill}} cccccc}
   \hline
@@
   \end{tabular*}
+\end{center}  
 \end{table}
+
+%\begin{table}
+%\begin{center}
+%  \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
+%  \label{table:lom}
+%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccccc}
+%  \hline
+%  {Norm.} & Target & Jaccard & Dice  & False $-$ & False $+$\\
+%  \hline
+%  \hline
+%  antsT1 & 0.917 & 0.621 & 0.767 & 0.082 & 0.341 \\
+%  antsCC & 0.780 & 0.570 & 0.727 & 0.220 & 0.320 \\
+%  antsMI & 0.776 & 0.569 & 0.725 & 0.223 & 0.320 \\
+%  antsDemons & 0.780 & 0.566 & 0.723 & 0.220 & 0.326 \\
+%  antsSSD & 0.783 & 0.562 & 0.720 &  0.217 & 0.334 \\
+%  tbss & 0.716 & 0.550 & 0.710 & 0.283 & 0.295 \\
+%  \hline
+%  \end{tabular*}
+%\end{center}  
+%\end{table}
 
@@
 but also show a decrease in alignment.  This is further supported
-by the overlap measures \cite{Crum2006} given in Table \ref{table:lom}.
+by the overlap measures \citep{Crum2006} given in Table \ref{table:lom}.
 For each alignment strategy, we thresholded the derived mean FA image (FA $
 \geq$ 0.2) and compared it with white matter mask of the MNI152 template
-segmented using Atropos \cite{Avants2011a}.  The template-based
+segmented using Atropos \citep{Avants2011a}.  The template-based
 strategy showed the best overall overlap with the white matter MNI152
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{tbiCoV.jpg}
+  \includegraphics[width=85mm]{tbiCoV.pdf}
 \end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{meanFA.jpg}
+  \includegraphics[width=85mm]{meanFA.pdf}
 \end{tabular}
@@
 %\begin{tabular}{c}
-%  \includegraphics[width=80mm]{hoonData2.jpg}
+%  \includegraphics[width=80mm]{hoonData2.pdf}
 %\end{tabular}
@@
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{NKI2.jpg}
+  \includegraphics[width=85mm]{NKI2.pdf}
 \end{tabular}
@@
 % In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
-This division was accomplished by applying a distance transform \cite{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.
+This division was accomplished by applying a distance transform \citep{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.
 Note that only those voxels that were also within the thresholded FA mask were considered in the inner/outer analysis.
@@
 %The improvements seen in the CDF plots are also supported by comparative quantities derived from the skeletons of the different mean FA images.
-Since the skeletonization process is meant to determine the center of the various white matter tracts \cite{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.
+Since the skeletonization process is meant to determine the center of the various white matter tracts \citep{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.
 In addition, as described in \cite{Zalesky2011}, the number of connected components of the white matter skeleton is directly related to the contiguity and thus inversely related to the level of structure.
@@
 %or to the subject best representative of the mean as defined in 
-%\cite{Smith2006}. The former was not investigated due to the possible bias
+%\citep{Smith2006}. The former was not investigated due to the possible bias
 %in selecting a random subject as described in \cite{Smith2006}.  
@@
 %\begin{tabular}{c}
-%  \includegraphics[width=80mm]{NYU.jpg}
+%  \includegraphics[width=80mm]{NYU.pdf}
 %\end{tabular}
@@
 %\begin{tabular}{c}
-%  \includegraphics[width=180mm]{NYUplots2.jpg}
+%  \includegraphics[width=180mm]{NYUplots2.pdf}
 %\end{tabular}
@@
 \begin{figure}
+\begin{center}
 \begin{tabular}{c}
-  \includegraphics[width=85mm]{NKIplots3.jpg}
+  \includegraphics[width=85mm]{NKIplots3.pdf}
 \end{tabular}
@@
 \label{fig:NKIplots}
+\end{center}
 \end{figure}
@@
 \begin{tabular}{c}
-  \includegraphics[width=180mm]{nkiTesting.jpg}
+%  \includegraphics[width=180mm]{nkiTesting.pdf}
+  \includegraphics[width=130mm]{nkiTesting.pdf}
 \end{tabular}
@@
 SSD metrics are all capable of producing high quality anatomical
-alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
+alignment (e.g. \citep{Klein2009}).  All of this work, though, compares
 the quality of alignment at a relatively coarse scale of major gyri,
@@
 power in template-based FA studies.  Our contention is that this
-dramatic "improvement" in detection power is not due to better
-anatomical alignment.  Rather, it is a result of
+dramatic ``improvement'' in detection power is not due to better
+anatomical alignment.  Rather, it is a result of the
 circularity/nonindependence of the normalization and statistical
@@
 cannot be regarded as meaningful comparison techniques in the presence
-of normalization/statistical non-independence.  Similar issues are
+of normalization/statistical nonindependence.  Similar issues are
 raised in \cite{Kriegeskorte2010}.
@@
 clinical component of the study.   Additionally, we are not addressing
-our results' biological plausibility, nor do we
+the biological plausibility or our results nor do we
 intend to (in this paper) enter into discussion on the mechanisms of
 TBI, gender difference and DTI that may lead to the detected results.
-Rather, we focus on the technical issue of circularity induced by
+Rather, we focus on the technical issue of circularity caused by
 normalizing the image feature of interest with functions that maximize
 the statistic used in hypothesis testing.  We also cannot argue that our analysis
-invalidates previous results that used a circular strategy.  Instead, we encourage the community to
+invalidates previous studies in which circularity might be an issue.  
+Finally, we do not address the very promising work concerning
+whole tensor DTI normalization \citep{Zhang2007,Hecke2007} and DTI template-based
+strategies \citep{Mori2009,Hecke2011}.  Although investigation is beyond the scope
+of this work, these whole-tensor approaches might provide the necessary
+anatomical information to overcome the deficiencies addressed in this
+work.
+
+Instead, we encourage the community to
 reexamine FA normalization and analysis methods and to consider the
@@
 only recently have toolkits emerged to enable easy comparison of
-statistical techniques (via {\bf R}), normalization methods (ANTs) and
+statistical techniques (via {\bf R} \citep{RDevelopmentCoreTeam2011}), normalization methods (ANTs), and
 diffusion tensor processing (Camino).
@@
 is necessary to ensure the proper use of these tools and the
-calculation and reporting of findings \cite{Ridgway2008}.  The normalization 
+calculation and reporting of findings \citep{Ridgway2008}.  The normalization 
 circularity issue described in this work is the inducement of substantial Type 1
@@
 minimized thus mitigating the confounds described previously.
-Each DTI-based scalar image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \cite{Wu2008}.
+Each DTI-based scalar image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \citep{Wu2008}.
 \item Although the total number of registrations to be performed is increased, this strategy makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
@@
 In addition to registration and template building, ANTs also includes brain 
-extraction \cite{Avants2010a}, bias correction \cite{Tustison2010}, $n$-tissue 
-segmentation \cite{Avants2011a}, and cortical thickness estimation solutions \cite{Das2009}.
+extraction \citep{Avants2010a}, bias correction \citep{Tustison2010}, $n$-tissue 
+segmentation \citep{Avants2011a}, and cortical thickness estimation solutions \citep{Das2009}.
 
@@
 %% Usage of \cite is as follows:
-%%   \cite{key}          ==>>  [#]
+%%   \citep{key}          ==>>  [#]
 %%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
@@
 }
-\cite{Yushkevich2006} and
+\citep{Yushkevich2006} and
 DTI-TK.%

### 2011-06-10 ANTsTBSS/paper.tex
Minor edit.

@@
 between these and other analysis approaches, a core
-commonality includes direct FA-to-FA registration.
+commonality includes direct FA-to-FA registration {\color{red}{using various similarity metrics}}.
 The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and gradient) and is therefore widely used.

### 2011-06-18 Posters/NNS2011/myposter.tex
NNS2011 poster

%\documentclass[portrait,final,a0paper]{baposter}
\documentclass[paperwidth=48in,paperheight=48in,portrait,final]{baposter}
% Usa a4shrink for an a4 sized paper.

\tracingstats=2

\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{bm}

\usepackage{graphicx}
\usepackage{multicol}

\usepackage{pgfbaselayers}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\usepackage{times}
%\usepackage{helvet}
%\usepackage{bookman}
\usepackage{palatino}

\newcommand{\captionfont}{\footnotesize}

\selectcolormodel{cmyk}

\graphicspath{{images/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some math symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Format 
\newcommand{\Matrix}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\Vector}[1]{\Matrix{#1}}
\newcommand*{\SET}[1]  {\ensuremath{\mathcal{#1}}}
\newcommand*{\MAT}[1]  {\ensuremath{\mathbf{#1}}}
\newcommand*{\VEC}[1]  {\ensuremath{\bm{#1}}}
\newcommand*{\CONST}[1]{\ensuremath{\mathit{#1}}}
\newcommand*{\norm}[1]{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand*{\abs}[1]{\mathopen| #1 \mathclose|}% use instead of $\|x\|$
\newcommand*{\absLR}[1]{\left| #1 \right|}% use instead of $\|x\|$

\def\norm#1{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand{\normLR}[1]{\left\| #1 \right\|}% use instead of $\|x\|$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{0.7em}
\setlength{\columnseprule}{0mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{%
\setlength{\itemsep}{1pt}%
\setlength{\parskip}{0pt}%
\setlength{\parsep}{0pt}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Define some colors
\definecolor{silver}{cmyk}{0,0,0,0.3}
\definecolor{yellow}{cmyk}{0,0,0.9,0.0}
\definecolor{reddishyellow}{cmyk}{0,0.22,1.0,0.0}
\definecolor{black}{cmyk}{0,0,0.0,1.0}
\definecolor{darkYellow}{cmyk}{0,0,1.0,0.5}
\definecolor{darkSilver}{cmyk}{0,0,0,0.1}

\definecolor{lightyellow}{cmyk}{0,0,0.3,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}

% http://www.creativecolorschemes.com/resources/free-color-schemes/blue-tone-color-scheme.shtml

\definecolor{colorA}{cmyk}{1,0.25,0.1,0.5}
\definecolor{colorB}{cmyk}{1,0.3,0.05,0.2}
\definecolor{colorC}{cmyk}{0.6,0,0.05,0}
\definecolor{colorD}{cmyk}{0.95,0.6,0.05,0.15}
\definecolor{colorE}{cmyk}{0.6,0.3,0.05,0.05}
\definecolor{colorF}{cmyk}{0.4,0.15,0,0}
\definecolor{colorf}{cmyk}{0.1,0.0375,0,0}
\definecolor{colorG}{cmyk}{0.75,0.3,0.05,0.1}
\definecolor{colorH}{cmyk}{0.65,0.15,0,0.05}
\definecolor{colorI}{cmyk}{0.3,0.05,0,0}
\definecolor{colorJ}{cmyk}{0.65,0,1,0}
\definecolor{colorK}{cmyk}{0,0.5,1,0}
\definecolor{colork}{cmyk}{0,0.125,0.25,0}
\definecolor{colorL}{cmyk}{0.65,0.8,0,0}
\definecolor{colorM}{cmyk}{0.7,0.4,0.2,0.6}
\definecolor{colorN}{cmyk}{0.3,0.15,0.1,0.3}
\definecolor{colorn}{cmyk}{0.075,0.0375,0.025,0.075}
\definecolor{colorO}{cmyk}{0.25,0.05,0.1,0.15}

%%
\typeout{Poster Starts}
\background{
  \begin{tikzpicture}[remember picture,overlay]%
    \draw (current page.north west)+(-2em,2em) node[anchor=north west] {\includegraphics[height=1.1\textheight]{silhouettes_background}};
  \end{tikzpicture}%
}

\newlength{\leftimgwidth}
\begin{poster}%
  % Poster Options
  {
  % Show grid to help with alignment
  grid=no,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=colorf,
  bgColorTwo=white,
  borderColor=colorK,
  headerColorOne=colork,
  headerColorTwo=white,
  headerFontColor=colorM,
  boxColorOne=white,
  boxColorTwo=colorC,
  % Format of textbox
  textborder=roundedleft,
%  textborder=rectangle,
  % Format of text header
  eyecatcher=no,
  headerborder=closed,
  headerheight=0.11\textheight,
  headershape=roundedright,
  headershade=shade-lr,
  headerfont=\Large, %Sans Serif
  boxshade=plain,
%  background=shade-tb,
  background=plain,
  linewidth=2pt
  }
  % Eye Catcher
  {\includegraphics[width=10em]{D1077}} % No eye catcher for this poster. (eyecatcher=no above). If an eye catcher is present, the title is centered between eye-catcher and logo.
  % Title
  { %Sans Serif
  %\bf% Serif
  Multivariate analysis of diffusion tensor imaging and cortical thickness maps in a traumatic brain injury (TBI) cohort using Advanced Normalization Tools (ANTs)
  \vspace{0.1em}}
  % Authors
  { %Sans Serif
  % Serif
  Nicholas J. Tustison$^1$, Brian B. Avants$^2$, Philip A. Cook$^2$, Junghoon Kim$^3$, John Whyte$^3$, James C. Gee$^2$, Stephen  Ahlers$^4$, James R. Stone$^1$ \newline
  % Affiliations
  {\normalsize $^1$University of Virginia, $^2$University of Pennsylvania, $^3$Rehabilitation Research Institute, $^4$Naval Medical Research Center \\
  P309
  }

  }
  % University logo
  {% The makebox allows the title to flow into the logo, this is a hack because of the L shaped logo.
%    \makebox[8em][r]{%
%      \begin{minipage}{16em}
%        \hfill
%%        \includegraphics[height=2em]{msrlogo}
%        \includegraphics[height=7.0em]{logo}
%      \end{minipage}
%    }
  }

  \tikzstyle{light shaded}=[top color=baposterBGtwo!30!white,bottom color=baposterBGone!30!white,shading=axis,shading angle=30]

  % Width of left inset image
     \setlength{\leftimgwidth}{0.78em+8.0em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position 
%%% towards an already declared box. So if you have a box attached to the 
%%% bottom, one to the top and a third one which should be in between, you 
%%% have to specify the top and bottom boxes before you specify the middle 
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %
    % A coloured circle useful as a bullet with an adjustably strong filling
    \newcommand{\colouredcircle}[1]{%
      \tikz{\useasboundingbox (-0.2em,-0.32em) rectangle(0.2em,0.32em); \draw[draw=black,fill=baposterBGone!80!black!#1!white,line width=0.03em] (0,0) circle(0.18em);}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Introduction}{name=introduction,column=0,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

TBI is a complex disease process involving mechanical disruption
of neurological tissues and activation of secondary injury cascades, 
culminating in loss of function.  Multiple neuroimaging approaches 
have recently been developed to detect white matter (WM) and gray 
matter injury following TBI. 
%Diffusion tensor imaging (DTI) affords 
%quantification of microstructural white matter injury. High-resolution 
%T1-weighted sequences allow for computation of cortical thickness maps. 
It is hypothesized that a multivariate analysis of these two imaging 
modalities may provide insight into the alteration of cortical and WM 
circuits following TBI.  To address this hypothesis, we employ a novel 
multivariate technique, sparse canonical correlation analysis (SCCA), 
to quantify traumatically induced disruption of WM and cortical networks.

  \vspace{0.3em}
 }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Diffuse Traumatic Brain Injury Data}{name=data,column=0,below=introduction}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The TBI data used in this study is part of a larger effort investigating 
the relationship between various neuroimaging indices and cognitive and 
functional abilities in long-term survivors of TBI (principal 
investigator: John Whyte).   
\paragraph{Description of cohort: }
17 controls and 16 patients with TBI were used.  
Each patient had a history of non-penetrating traumatic brain injury 
of at least moderate severity defined by significant and well-documented
loss or alteration of consciousness following injury in addition to meeting
several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and years of education.
\paragraph{Image acquisition:} 
High resolution T1-weighted anatomic images were obtained using a 3D MP-RAGE 
imaging sequence with the following acquisition parameters: TR = 1620 ms, 
TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 
1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 
1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
DTI images were also obtained.

      \vspace{0.3em}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Advanced Normalization Tools}{name=ants,column=0,span=1,below=data,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{itemize}
%\item{\bf Processing}
%\vspace{-10mm}\\
\begin{center}
{\bf Processing} \\
\vspace{2mm}
\begin{tabular}{cc}
\includegraphics[height=20mm]{normalization.pdf} & 
\includegraphics[height=25mm]{template.pdf} \\
SyN normalization \cite{Avants2011} & SyGN template building \cite{Avants2010}\\
\vspace{-1mm}\\
\includegraphics[height=20mm]{n4.pdf} & 
\includegraphics[height=20mm]{atropos.pdf} \\
N4 bias correction \cite{Tustison2010} & Atropos $n$-tissue segmentation \cite{Avants2011a} \\  
\vspace{-1mm}\\
\includegraphics[height=20mm]{r16.pdf} & 
\includegraphics[height=20mm]{glamglue.pdf} \\
DiReCT cortical thickness \cite{Das2009} & topological well-composedness \cite{Tustison2011} \\
\vspace{-1mm}\\
\end{tabular}
\end{center}

\begin{center}
{\bf Statistics}\\
\vspace{4mm}
\begin{tabular}{c}
\includegraphics[height=25mm]{MUvPSCCAN.pdf} \\
sparse canonical correlation analysis \cite{Avants2010b}
\end{tabular}
\end{center}

\smaller
\begin{center}
  http://wwww.picsl.upenn.edu/ANTs
\end{center}

%\vspace{0.3em}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{ANTs TBI Multivariate Processing Workflow}{name=workflow,column=1,span=2}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \vspace{0.3em}

\begin{center}
\begin{tabular}{c}
\includegraphics[height=80mm]{workflow.pdf} 
\end{tabular}
\end{center}

  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Results}{name=results,column=1,span=1,below=workflow}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \vspace{0.3em}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Conclusions}{name=conclusions,column=1,below=results}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 SCCA demonstrates significant differences between the control and patient groups in both the FA ($p < 0.002$) and gray matter ($p < 0.01$)	that	are	widespread	but	largely	focus	on	thalamocortical networks related to the limbic system. Specific regional differences in- cluded the medial thalamic nuclei, hypothalamus, amygdala, hippo- campus, anterior cingulate cortex, orbitofrontal cortex and fornix. Using these SCCA-identified regions, we demonstrate a strong correlation of the degree of injury in WM and GM within the patient group.

  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{References}{name=references,column=2,below=workflow}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \smaller

      \vspace{-0.4em}
      \renewcommand{\refname}{\vspace{-0.8em}}
      \bibliographystyle{abbrv}
      \bibliography{references}

%    \bibliographystyle{ieee}
%    \renewcommand{\section}[2]{\vskip 0.05em}
%      \begin{thebibliography}{1}\itemsep=-0.01em
%      \setlength{\baselineskip}{0.4em}
%      \bibitem{amberg07:nonrigid}
%        B.~Amberg, S.~Romdhani, T. Vetter.
%        \newblock {O}ptimal {S}tep {N}onrigid {ICP} {A}lgorithms for {S}urface {R}egistration
%        \newblock In {\em Computer Vision and Pattern Recognition 2007}
%      \bibitem{amberg08:recognition}
%        B.~Amberg, R.~Knothe, T. Vetter.
%        \newblock Expression Invariant Face Recognition with a 3D Morphable Model
%        \newblock In {\em Automated Face and Gesture Recognition 2008}
%      \end{thebibliography}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \headerbox{Expression Neutralization}{name=results neutralization,column=1,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \begin{tikzpicture}[x=0.3333\linewidth,y=-0.42\linewidth]
%    \path [use as bounding box] (-0.5,-0.5) rectangle(2.5,1.7);
%    \path
%    (0,0) node{\includegraphics[width=0.42\linewidth]{D1077}}
%    (1,0) node{\includegraphics[width=0.47\linewidth]{D1077_fit_expression}}
%    (2,0) node{\includegraphics[width=0.47\linewidth]{D1077_fit}}
%
%    (0,1) node{\includegraphics[width=0.42\linewidth]{D1360}}
%    (1,1) node{\includegraphics[width=0.47\linewidth]{D1360_fit_expression}}
%    (2,1) node{\includegraphics[width=0.47\linewidth]{D1360_fit}}
%
%    (0,1.6) node {\smaller a) Target}
%    (1,1.6) node {\smaller b) Fit}
%    (2,1.6) node {\smaller c) Normalized};
%  \end{tikzpicture}
%  \vspace{0.5em}
%
%  Expression normalisation for two scans of the same individual.  
%  The robust fitting gives a good estimate (b) of the true face surface given
%  the noisy measurement (a). It fills in holes and removes artifacts using
%  prior knowledge from the face model. The pose and expression normalized faces
%  (c) are used for face recognition.
%  \vspace{0.5em}
%  }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \headerbox{Funding}{name=funding,column=1,span=2,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \smaller 
%  \hspace{1em}This work was supported in part by Microsoft Research through the European PhD Scholarship Programme.
%  }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \headerbox{Robustness}{name=robustness,column=2,row=0,above=results,span=1}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \begin{tikzpicture}[x=0.3333\linewidth,y=-0.42\linewidth]
%    \path [use as bounding box] (-0.5,-0.5) rectangle(2.5,1.7);
%    \path
%    (0,0) node{\includegraphics[width=0.42\linewidth]{D1160}}
%    (1,0) node{\includegraphics[width=0.42\linewidth]{D1425}}
%    (2,0) node{\includegraphics[width=0.42\linewidth]{D1205}}
%
%    (0,1) node{\includegraphics[width=0.28\linewidth]{D1160_fit_expression}}
%    (1,1) node{\includegraphics[width=0.28\linewidth]{D1425_fit_expression}}
%    (2,1) node{\includegraphics[width=0.28\linewidth]{D1205_fit_expression}}
%
%    (1,0.5) node {\smaller a) Target}
%    (1,1.6) node {\smaller b) Robust Reconstruction};
%  \end{tikzpicture}
%  \vspace{0.5em}
%
%  The reconstruction (b) is robust against scans (a) with artifacts, noise, and
%  holes.
%  
%  This is achieved by a robust iteratively reweighted ICP algorithm and outlier
%  rejection based on angle comparisions between corresponding points.
%  }

\end{poster}

\end{document}

### 2011-06-25 ANTsTBSS/paper.tex
Added a single reference.

@@
 } 
-VBM \citep{Ashburner2001} continues to find application to FA studies \citep{Kakeda2010,Takao2010} despite concerns about its general validity \citep{Bookstein2001,Davatzikos2004} and with respect to DTI \citep{Jones2005,Chung2008}.
+VBM \citep{Ashburner2001} continues to find application to FA studies \citep{Kakeda2010,Takao2010,Preziosa2011} despite concerns about its general validity \citep{Bookstein2001,Davatzikos2004} and with respect to DTI \citep{Jones2005,Chung2008}.
 The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.

### 2011-07-01 Posters/NNS2011/myposter.tex
DOC: added bibs

@@
 \usepackage{amssymb}
+\usepackage{latexsym}
 \usepackage{relsize}
@@
   % Affiliations
-  {\normalsize $^1$University of Virginia, $^2$University of Pennsylvania, $^3$Rehabilitation Research Institute, $^4$Naval Medical Research Center \\
+  {\normalsize $^1$University of Virginia, $^2$University of Pennsylvania, $^3$Moss Rehabilitation Research Institute, $^4$Naval Medical Research Center \\
   P309
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Diffuse Traumatic Brain Injury Data}{name=data,column=0,below=introduction}{
+  \headerbox{Diffuse traumatic brain injury data}{name=data,column=0,below=introduction}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\vspace{-6mm}
+
 \begin{center}
-{\bf Statistics}\\
-\vspace{4mm}
-\begin{tabular}{c}
-\includegraphics[height=25mm]{MUvPSCCAN.pdf} \\
-sparse canonical correlation analysis \cite{Avants2010b}
-\end{tabular}
+{\bf Multivariate analysis using SCCA}\\
+\end{center}
+\vspace{-1mm}
+Sparse canonical correlation analysis (SCCA) adapts classical CCA to
+situations where the number of predictors is much greater than the number
+of subjects for inference of correlative relationships between different
+``views'' of some underlying phenomenon. For gray matter density ($\mathbf{G}$) and
+DTI-derived FA  ($\mathbf{F}$) views, SCCA solves the following
+criterion \cite{Avants2010b}:
+\vspace{-7mm}
+\begin{center}
+\begin{align*}
+\underset{\omega_{\mathbf{G}}, \omega_{\mathbf{F}}}{\operatorname{argmax}} \left\{\omega_{\mathbf{G}} \mathbf{G}^\mathrm{T}\mathbf{F}\omega_{\mathbf{F}} - \lambda_{\mathbf{G}}\sum_{v_{\mathbf{G}}} |\omega_{\mathbf{G}}|_1 - \lambda_{\mathbf{F}}\sum_{v_{\mathbf{F}}} |\omega_{\mathbf{F}}|_1\right\}: \|\omega_{\mathbf{G}}\|, \|\omega_{\mathbf{F}}\| \leq 1.
+\end{align*}
 \end{center}
 
-\smaller
-\begin{center}
-  http://wwww.picsl.upenn.edu/ANTs
-\end{center}
+%\smaller
+%\begin{center}
+%  http://wwww.picsl.upenn.edu/ANTs
+%\end{center}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{ANTs TBI Multivariate Processing Workflow}{name=workflow,column=1,span=2}{
+  \headerbox{Materials and methods:  ANTs subject and template processing}{name=workflow,column=1,span=2}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{center}
-\begin{tabular}{c}
-\includegraphics[height=80mm]{workflow.pdf} 
+\begin{tabular}{p{11cm}p{11cm}}
+{\makebox[11cm][c]{\bf Individual subject processing}\vspace{-3mm} } & 
+{\makebox[11cm][c]{\bf Template generation}} \\
+{\makebox[11cm][c]{\includegraphics[height=60mm]{individualProcessing.pdf}}}\vspace{3mm} &
+{\makebox[11cm][c]{\includegraphics[height=60mm]{templateBuilding.pdf}}} \\
+\small Processing of each anatomical (T1-weighted) image consists of the
+following steps: 
+\begin{itemize}
+  \item bias correction,
+  \item skull stripping using Atropos coupled with anatomical priors, and
+  \item prior-based csf, white matter, and gray matter segmentation.  The
+  latter yields the gray matter density image for each subject.
+\end{itemize}
+FA images are mapped to each anatomical image using the transform
+obtained from registration of the average diffusion image to the 
+corresponding T1. 
+& 
+\small Statistical analysis requires normalization to a standard space provided
+by the total template.  All 17 controls and all 16 TBI patients were used
+to generate their respective unbiased templates.  These two templates were then
+used to generate the final/standard template.  Thus the mapping from the individual 
+gray matter density (GMD) image to the total template is given by
+\begin{align*}
+  \mathrm{GMD}_{subject}\,\,\mathbf{Id}\,\, \mathrm{T1}_{subject}\leadsto \mathrm{template}_{population}\leadsto \mathrm{template}_{total}.
+\end{align*}
+Similarly, the mapping for each subject's FA image to the total template is given
+by the 
+transform composition
+\begin{align*}
+  \mathrm{FA}_{subject}\rightarrow\mathrm{T1}_{subject}\leadsto \mathrm{template}_{population}\leadsto \mathrm{template}_{total}.
+\end{align*}
+\\
 \end{tabular}
@@
 
+\vspace{-10mm}
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Results}{name=results,column=1,span=1,below=workflow}{
+  \headerbox{Multivariate analysis results}{name=results,column=1,span=2,below=workflow}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \vspace{0.3em}
+
+\begin{center}
+\begin{tabular}{p{6cm}cc} \vspace{-70mm}
+{\makebox[6cm][c]{\includegraphics[height=50mm]{scca.pdf} }} \
+\small The significant correlations derived from the GMD/FA views
+vs. diagnostic view provide significant regions for finding
+correlations between GMD and FA. & {           } &
+{\includegraphics[height=80mm]{results.pdf}}
+\end{tabular}
+\end{center}
+
   }
@@
 
- SCCA demonstrates significant differences between the control and patient groups in both the FA ($p < 0.002$) and gray matter ($p < 0.01$)	that	are	widespread	but	largely	focus	on	thalamocortical networks related to the limbic system.
-Specific regional differences in- cluded the medial thalamic nuclei, hypothalamus, amygdala, hippo- campus, anterior cingulate cortex, orbitofrontal cortex and fornix.
-Using these SCCA-identified regions, we demonstrate a strong correlation of the degree of injury in WM and GM within the patient group.
+ SCCA demonstrates significant differences between the control and patient groups in both the FA ($p < 0.002$) and gray matter ($p < 0.04$)	that	are	widespread	but	largely	focus	on	thalamocortical networks related to the limbic system.
+Specific regional differences included the medial thalamic nuclei, hypothalamus, amygdala, hippocampus, anterior cingulate cortex, orbitofrontal cortex and fornix.
+Using these SCCA-identified regions, we demonstrate a strong correlation ($p < 0.08$) of the degree of injury in WM and GM within the patient group.
 
+\vspace{0.3em}
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{References}{name=references,column=2,below=workflow}{
+  \headerbox{References}{name=references,column=2,below=results,above=bottom}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \smaller
 
+\vspace{0.3em}
+    \scriptsize
       \vspace{-0.4em}
@@
 %      \end{thebibliography}
+  }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \headerbox{Open source software}{name=software,column=1,below=conclusions,above=bottom}{
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{1mm}
+  \begin{center}
+  \begin{tabular}{cc}
+  \includegraphics[height=18mm]{ants_logo.png} &
+  \includegraphics[height=18mm]{itkLogo.jpg} \\
+%  {Advanced Normalization Tools} & {Insight Toolkit} \\
+  {\tiny http://www.picsl.upenn.edu/ANTs} &
+%  {\tiny http://www.cs.ucl.ac.uk/research/medic/camino} &
+  {\tiny http://www.itk.org/} \\
+  \end{tabular}
+  \end{center}
+
   }

### 2011-07-02 Posters/NNS2011/myposter.tex
ENH: Minor edits.

@@
 
-\vspace{0.3em}
-    \scriptsize
+\vspace{2mm}
+    \tiny % \scriptsize
       \vspace{-0.4em}
@@
       \bibliography{references}
+
+  \begin{center}
+  {\normalsize \bf Open source software} \\
+{\vspace{3mm}}
+  \begin{tabular}{cc}
+  \includegraphics[height=12mm]{ants_logo.png} &
+  \includegraphics[height=11mm]{itkLogo.jpg} \\
+%  {Advanced Normalization Tools} & {Insight Toolkit} \\
+  {\tiny http://www.picsl.upenn.edu/ANTs} &
+%  {\tiny http://www.cs.ucl.ac.uk/research/medic/camino} &
+  {\tiny http://www.itk.org/} \\
+  \end{tabular}
+  \end{center}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Open source software}{name=software,column=1,below=conclusions,above=bottom}{
+  \headerbox{Acknowledgments}{name=acknowledgments,column=1,below=conclusions,above=bottom}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \vspace{1mm}
-  \begin{center}
-  \begin{tabular}{cc}
-  \includegraphics[height=18mm]{ants_logo.png} &
-  \includegraphics[height=18mm]{itkLogo.jpg} \\
-%  {Advanced Normalization Tools} & {Insight Toolkit} \\
-  {\tiny http://www.picsl.upenn.edu/ANTs} &
-%  {\tiny http://www.cs.ucl.ac.uk/research/medic/camino} &
-  {\tiny http://www.itk.org/} \\
-  \end{tabular}
-  \end{center}
-
+  \scriptsize
+  \begin{itemize}
+  \item Partial funding support from NZ Breacher study
+  \item This project is also funded, in part, under a grant with the Pennsylvania 
+  Department of Health. The Department specifically disclaims responsibility 
+  for any analyses, interpretations, or conclusions.
+  \end{itemize}
   }

### 2011-07-11 ANTsTBSS/paper.tex
ENH: Minor edits.
Added refernce.

@@
 
-\documentclass[preprint,authoryear,review,12pt]{elsarticle}
-%\documentclass[final,5p,times,twocolumn]{elsarticle}
+%\documentclass[preprint,authoryear,review,12pt]{elsarticle}
+\documentclass[final,5p,times,twocolumn]{elsarticle}
 
@@
 of a high performance \citep{Klein2009}, nonrigid registration algorithm \citep{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \citep{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
-The key point, here, is that the FA images are mapped to the common reference space via intra-subject DT-to-T1 mapping followed by T1-to-T1 alignment.
+The key point is that the FA images are mapped to the common reference space via intra-subject DT-to-T1 mapping followed by T1-to-T1 alignment.
 The T1-to-T1 step is (statistically) independent of the FA values which are the ultimate statistical target of hypothesis testing in FA population studies.
@@
   \label{table:lom}
-  \begin{tabular*}{130mm}{@{\extracolsep{\fill}} cccccc}
+  \begin{tabular*}{85mm}{@{\extracolsep{\fill}} cccccc}
   \hline
@@
 Finally, we do not address the very promising work concerning
-whole tensor DTI normalization \citep{Zhang2007,Hecke2007} and DTI template-based
-strategies \citep{Mori2009,Hecke2011}.  Although investigation is beyond the scope
+whole tensor DTI normalization \citep{Zhang2007,Hecke2007}, DTI template-based
+strategies \citep{Mori2009,Hecke2011}, {\color{red}{or other related methods \cite{jbabdi2010}}}.  Although investigation is beyond the scope
 of this work, these whole-tensor approaches might provide the necessary

### 2011-10-17 ANTsTBSS/paper.tex
ENH: Adding revision stuff

@@
 clinical component of the study.   Additionally, we are not addressing
-the biological plausibility or our results nor do we
+the biological plausibility of our results nor do we
 intend to (in this paper) enter into discussion on the mechanisms of

### 2011-10-17 ANTsTBSS/paperRevision.tex
ENH: Adding revision stuff

--- 2011-10-17 ANTsTBSS/paper.tex
@@
         floatplacement=!h,
-        caption={},
+        caption={\small {\ttfamily CreateDTICohort} which implements the a ground truth DTI simulator \cite{van-hecke2009}.
+The short command line menu which is invoked using the  `{\ttfamily {-}h}'  option.
+The long command line menu is obtained by typing `{\ttfamily {-}{-}help}'},
         captionpos=b,
@@
 
-\title{Circularity in Population-Based FA Analysis}
+\title{Circularity in Voxel-Based Analysis}
 
@@
 \begin{abstract}
-Template-based normalization plays a critical role in many population
-studies based on diffusion tensor imaging (DTI).  Several current
-approaches compute the fractional anisotropy (FA) from the DTI and then
-directly register the FA image to a FA template before performing
-statistical comparison between different groups within a population.  In
-this paper, we demonstrate that this analysis strategy induces selection
-bias that varies in strength with the similarity metric used in
+Recent discussions within the neuroimaging community have highlighted the
+problematic presence of selection bias in experimental design.  Although 
+initially centering around the selection of voxels during the course of
+fMRI studies, we demonstrate how this concept of circularity bias can 
+potentially corrupt voxel-based analyses with a particular emphasis on
+population studies of fractional anisotropy (FA).  
+For such studies, template-based registration plays a critical role
+in which a representative template serves as the normalized space for 
+group alignment.  Following 
+computation of FA, each subject's FA map is registered to the FA template prior to 
+performing statistical comparisons between different groups within a population.
+We illustrate how this analysis strategy can induce selection
+bias which varies in strength with the similarity metric used in
 registration.  Specifically, we show that the popular sum of squared
@@
 that normalizes a population of FA images through the associated 
-anatomical images.  We show that the Symmetric Group Normalization
+anatomical images.  One possibility, the Symmetric Group Normalization
 strategy for unbiased template construction from anatomical (e.g.
-T1-weighted) images provides a valid and unbiased coordinate system for
-statistical inference of FA differences and yet retains detection power
-in comparison to FA-based normalization methods.  We establish this
-outcome by using open data comprised of normal controls from the International
+T1-weighted) images, provides a valid and unbiased coordinate system for
+statistical inference of FA differences.  
+We establish this
+outcome by using simulated data derived from normal controls from the International
 Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
-brain injury survivors compared with 17 control subjects.  Alignment
+brain injury survivors compared with 17 control subjects.  
+Alignment
 comparisons with the normalization component of the popular tract-based
@@
 approach produces good alignments despite the absence of any direct
-FA-to-FA registration.  Of practical consideration we note that the
-software tools discussed in this work are provided to the public as part
-of the open source Advanced Normalization Tools (ANTs) repository.
-In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the statistical functions used to assess population effects.
-
-%We propose a modification of the popular tract-based spatial statistics (TBSS) framework \citep{Smith2006} for investigating diffusion tensor derived scalar measures across populations.
-Whereas the current crucial alignment step is performed typically using the derived fractional anisotropy images, we propose using the Symmetric Group Normalization strategy for anatomical unbiased template construction from T1-weighted images which serves as the coordinate system for statistical inference.
-A potential benefit of this modified approach is reduction of the general confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.
-Alignment performance is evaluated using open data sets from the International Neuroimaging Data-sharing Initiative.
-This assessment demonstrated improved alignment in both fractional anisotropy (FA) and mean diffusivity (MD) images versus standard TBSS.
-Further analyses evidencing increased regional sensitivity in peripheral white matter to FA differences include a DTI-based assessment of a cohort of 16 traumatic brain injury survivors compared with 17 control subjects.
-Similar to the Oxford Centre for Functional MRI of the Brain (FMRIB) who provide TBSS to the public, the software tools discussed herein are provided to the public as part of the open source Advanced Normalization Tools (ANTs) repository.
+FA-to-FA registration.  
+%Of practical consideration we note that the
+%software tools discussed in this work are provided to the public as part
+%of the open source Advanced Normalization Tools (ANTs) repository.  
+In conclusion, we recommend against circular analysis strategies that 
+select image similarity metrics that are tied directly to the statistical 
+functions used to assess population effects. 
 \end{abstract}
@@
 \begin{keyword}
-Advanced Normalization Tools (ANTs) \sep diffusion tensor imaging \sep fractional anisotropy \sep template construction
+Advanced Normalization Tools (ANTs) \sep circularity\ diffusion tensor imaging \sep fractional anisotropy \sep template construction \sep voxel-based analysis
 %% keywords here, in the form: keyword \sep keyword
@@
 \section{Introduction}
+Motivating the findings of a recent paper, the authors related their
+puzzlement at the unusually high number of papers reporting 
+significant correlations in fMRI studies \cite{Vul2009}. During the 
+resulting investigation, the authors discovered the permeating
+presence of a particular bias characterized by such related terms as ``circularity''
+and ``non-independence''.
+While related literature has focused on the presence of such bias in 
+fMRI studies \cite{Vul2009,Vul2010} and, more generally, in neuroscience research 
+\cite{Kriegeskorte2009}, we show how this can potentially extend to what has
+become a fundamental neuroimaging analysis technique---voxel-based analysis.
+
+Similarly, we were perplexed by the disparity in statistical results of one of 
+our recent analyses \cite{Stone2011} using the same DTI data and same 
+voxel-based analysis technique but different normalization approaches.  Despite 
+distinctively better alignments with one approach, there was little 
+difference in the statistical results between the two approaches.  
+It was eventually discovered that the 
+incommensurability between alignment quality and effect size in our
+analysis was also a due to a subtle, yet significant circularity bias
+which we explain below in the context of FA population studies.  
+%Due to the public availability and widespread use of
+%these voxel-based analysis tools we show 
+%how this source of bias can enter into voxel-based studies focusing on
+%population-based FA .
+
+%Whereas the authors
+%of \cite{Vul2009} were able to estimate reliability measures of
+%both personality tests and fMRI and, thus, provide an upper bound 
+%for possible correlations, the nature of the the analysis provided
+%no indication as to the source of the disparity.
+
 %State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results.
@@
 \citep{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique.
-DTI's sensitivity to brain architecture \citep{Basser1996,Assaf2008} enables promising analysis possibilities assessing neuro-structural differences in cross-population studies \citep{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010} including subjects with traumatic brain injury (TBI) \citep{Caeyenberghs2010,Jiang2010,Warner2010}.
+DTI's sensitivity to brain architecture \citep{Basser1996,Assaf2008} enables promising analysis possibilities assessing neuro-structural differences in cross-population studies \citep{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010}.
+% including subjects with traumatic brain injury (TBI) \citep{Caeyenberghs2010,Jiang2010,Warner2010}.  
 Two popular comprehensive software packages for assessing population FA differences include the SPM Matlab-based toolkit%
@@
 between these and other analysis approaches, a core
-commonality includes direct FA-to-FA registration {\color{red}{using various similarity metrics}}.
-The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and gradient) and is therefore widely used.
-Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and yield reasonable performance levels \citep{Klein2009}.%
-\footnote{
-In addition to TBSS and SPM, other popular packages in the neuroimaging community 
-which use the SSD metric include AIR \citep{Woods1998} and LDDMM \citep{Beg2005}.
-}
+commonality includes direct FA-to-FA registration using various similarity metrics.
+The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image intensity difference to zero), is computationally efficient (just compute the image difference and gradient), and is therefore widely used.
+Some of the most popular registration methods (e.g.
+SPM,
+Demons \cite{Thirion1998}, FSL's FNIRT, AIR \citep{Woods1998}, and LDDMM \citep{Beg2005}) rely on this metric and yield reasonable performance levels \citep{Klein2009}.%
+%\footnote{
+%In addition to TBSS and SPM, other popular packages in the neuroimaging community 
+%which use the SSD metric include AIR \citep{Woods1998} and LDDMM \citep{Beg2005}.
+%}
 
@@
 \end{align}
-is that which minimizes the voxelwise variance.
-Reducing variance will also directly affect the value of a standard population statistic, the Student's $t$-test.
+is that which minimizes the voxelwise group variance, or at least an approximation thereof.
+Reducing the group variance will also directly affect the value of a standard population statistic, the Student's $t$-test.
 Other standard similarity metrics will also exhibit this
@@
 Thus, instead of normalization based on anatomical alignment followed by
-statistical testing, direct FA normalization explicitly conflates
+statistical testing, direct normalization of the images to be statistically compared 
+explicitly conflates
 anatomical alignment with optimizing the very statistical testing
@@
 Eqn.
-\ref{eq:variance}, direct FA-to-FA registration will produce transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
+\ref{eq:variance}, direct FA-to-FA template registration will produce transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
 $t$-test).
-While the SSD metric is able to provide visually attractive alignments across subjects, the voxel-level details (that is, which voxels are matched where) are, as we show in the following simulation study, strikingly important in terms of the effect on detection power.
-
-%This results in a significant increase in the number of false positives for this type
-%of testing which we illustrate with the following example.
-
-\paragraph{Simulated Data Example}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedData.pdf}
-\end{tabular}
-\caption{(a) Axial slice from the discrete BrainWeb data set.  From (a) 
-we create the (b) simulated template and sample (c) control and (d) patient images with both 
-additive noise and subsequent Gaussian smoothing. }
-\label{fig:simulated_data}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedDataWarped.pdf}
-\end{tabular}
-\caption{Warping the image in Fig. \ref{fig:simulated_data}(c) to the template in 
-  Fig. \ref{fig:simulated_data}(b) using the SSD similarity metric produces the image shown in
-  (a).  The corresponding deformation is given in (b).  Although the images were anatomically
-  aligned prior to invoking ANTs registration, deformation is driven by the similarity metric
-  causing the accumulation of Type 1 errors.}
-\label{fig:simulated_data_warped}
-\end{center}
-\end{figure}
-
-Axial slice number 159 was extracted from \verb#subject_04# of the discrete BrainWeb database
-\citep{Aubert-Broche2006} which is shown in Fig. \ref{fig:simulated_data}.  After thresholding out 
-the white matter and assigning all white matter voxels a value of 0.7, 10\% of these white 
-matter voxels were randomly selected to be assigned a value of 0.6 in the simulated control 
-template and a value of 0.3 in the simulated patient template.  
-Note that these values were arbitrarily selected to be more or less in the range of typical
-FA values.  Gaussian noise (mean = 0.0, standard deviation = 0.1) was added to the white
-matter template followed by a Gaussian smoothing (standard deviation = 1.5 mm) shown in Fig. \ref{fig:simulated_data}(b).  A simulated control cohort was created consisting of 20 subjects.  
-Starting with the simulated control template, each of the 20 subjects is created by adding
-Gaussian noise (mean = 0.0, standard deviation = 0.1) followed by 
-Gaussian smoothing (standard deviation = 1.5 mm).
-Thus, each subject is differentiated only by a unique random draw from the noise distribution and there is no real shape difference between subjects.
-A simulated patient cohort of 20 subjects
-was created similarly using the simulated patient template.  A sample control subject and sample
-patient subject are given, respectively, in Fig. \ref{fig:simulated_data}(c) and (d).
+While the SSD metric is able to provide visually attractive alignments across subjects, the voxel-level details (that is, which voxels are matched where) are strikingly important in terms of the effect on detection power.
+
+Although various processing choices mitigate the effects of such bias,
+such as Gaussian smoothing, nonparametric testing \citep{Rorden2007}, 
+statistical analysis based on orthogonal projections onto the white matter skeleton 
+\cite{Smith2006}, multiple comparisons corrections \cite{Nichols2003}, and the use of more robust similarity metrics, 
+such choices are often made ad hoc and/or post hoc and, to our 
+knowledge, not with the realization of the potential
+for circularity bias described by Eqn. (\ref{eq:variance}).
+
+Instead of direct FA-to-FA template mapping for 
+alignment to the normalized space, we advocate a statistically 
+independent alignment protocol in which the FA images are mapped to 
+the common reference space via intra-subject FA-to-T1 mapping 
+followed by T1-to-T1 template alignment thus eliminating all direct
+FA-to-FA registrations. The key point is that the intensity value differences used 
+to assess alignment (i.e. the similarity metric) in FA-to-T1 and T1-to-T1 
+normalizations are statistically independent of the intensity value differences used for 
+the post-alignment numerical analysis.  It should be noted that this discussion
+is certainly not limited to studies involving FA.  From 
+Eqn. (\ref{eq:variance}), it is evident that such bias can occur anytime the 
+similarity metric (and the values employed) are non-independent of 
+the statistical analysis.
+
+Using simulated data, we first establish that this effect is real 
+and warrants concern on the part of researchers using 
+these techniques.  We then show how the general application
+of a high performance \citep{Klein2009}, nonrigid registration algorithm 
+\citep{Avants2011}, which forms the underlying machinery for an optimal 
+template construction algorithm \citep{Avants2010}, can facilitate alignment 
+between DTI-derived scalar images for cross-population studies.  
+Although this type of alignment procedure is not uncommon 
+(e.g. \cite{Lipton2008}), to our knowledge, this is the 
+first time a connection has been made with the concept of 
+circularity.
+
+%The rest of the paper establishes the severity of 
+%this bias in situations
+% although scientific best 
+%practices would dictate the reasonable removal of all 
+%potential sources of bias (including circularity) regardless of 
+%the general degree of severity.
+
+\section{Methods and Materials}
+% Provide sufficient detail to allow the work to be reproduced. 
+% Methods already published should be indicated by a reference: 
+% only relevant modifications should be described.
+
+\subsection{Template Construction for Anatomically-Based Population Normalization}
 
@@
 \begin{tabular}{c}
-%  \includegraphics[width=180mm]{simulatedTesting.pdf}
-  \includegraphics[width=130mm]{simulatedTesting.pdf}
-\end{tabular}
-\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the simulated experiment adjusted for multiple comparisons (FDR).
-The true significance curves are drawn in black.  Elevated statistical power is seen with all 4 
-         metrics for both tests although the effect is less in the more robust Wilcoxon results.
-Not shown are the significance curves for the Brunner-Munzel testing \citep{Rorden2007} which showed similar effects to the results on the left.
-        }
-\label{fig:simulated_testing}
-\end{center}        
-\end{figure*}
-
-After creation of the simulated data set, we performed voxelwise testing (both $t$-testing
-and Wilcoxon rank-sum testing, aka the Mann-Whitney test, one-tailed,
-$\alpha = 0.95$) with false discovery rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).
-The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
-We denote these graphs as "statistical significance curves."
-The true curves for the simulated experiment are plotted in black in Fig. \ref{fig:simulated_testing}.  
-We note that the amount of smoothing increases the proportion of significant voxels beyond
-the prescribed 10\%.
-
-To illustrate the effect of similarity metric on statistical significance, we performed ANTs registration \citep{Avants2011} of each subject to the template
-(Fig. \ref{fig:simulated_data}(b)) using cross correlation (CC), mutual information (MI), 
-Demons \citep{Thirion1998}, and sum of squared differences (SSD). A representative deformed image
-using the SSD metric and its corresponding deformed grid is given in 
-Fig. \ref{fig:simulated_data_warped}. 
-The resulting statistical significance curves are plotted with the true significance curves in Fig. 
-\ref{fig:simulated_testing}.
-
-This example demonstrates the undesirable phenomenon previously
-described.  Even though these images were anatomically aligned prior
-to registration, the resulting deformations (falsely) increased
-detection power.  This effect can be explained in the context of
-Eqn. (\ref{eq:variance}) and by the fact that we added noise to the
-underlying distribution which allows the deformable mapping algorithms
-the freedom to transform the positions of the most "powerful" voxels
-to minimize variance across the population (and increase the $t$-test
-statistics).  Although we do not show the results, additional
-experiments show that decreasing the standard deviation of the noise
-causes an increase in separation between the SSD/Demon's metrics and
-the MI and CC metrics.  MI and CC are closer to the the true
-significance curve.  Alternatively, if we increase the noise level,
-the same ranking occurs with even more spread between significance curves.
-
-Common similarity metric functions can therefore cause false positives by minimizing local intensity 
-variance between the subjects and template which explicitly maximizes statistical
-testing results in contrast to maximizing anatomical alignment.  This effect is strongest in 
-the SSD and Demons metrics and weakest in the MI and CC metrics likely due
-to the degree of correspondence to Eqn. (\ref{eq:variance}).   This is also supported by contrasting
-the $t$-test results given in Fig. \ref{fig:simulated_testing}(a) with the non-parametric Wilcoxon
-rank-sum testing (Fig.
-\ref{fig:simulated_testing}(b)).
-The rank test does not explicitly use variance in its assessment and thus exhibits far less statistical exaggeration even though it offers roughly 95\%
-the power of the $t$-test assuming certain assumptions are met \citep{Siegel1988,Rorden2007}.
-
-Although various processing choices mitigate these effects, such as Gaussian smoothing, nonparametric testing, and the use of more robust similarity metrics, such choices are often made on an ad hoc basis.
-In contrast, in this work we strongly advocate the general application
-%A more principled approach is the use of an anatomical, i.e. T1-weighted, unbiased shape/intensity 
-%template constructed from the population images.  DTI-derived scalar images can then be mapped 
-%to their corresponding T1-weighted image in a highly constrained manner and subsequently 
-%mapped to the template via nonrigid registration of the T1-weighted image.  This minimizes 
-%possible complications associated with direct FA-to-FA image registration.
-%We recognize that using the anatomical images in the alignment process has been previously proposed (in fact, such a suggestion occurs in the {\it Future directions} section of the original TBSS paper).
-of a high performance \citep{Klein2009}, nonrigid registration algorithm \citep{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \citep{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
-The key point is that the FA images are mapped to the common reference space via intra-subject DT-to-T1 mapping followed by T1-to-T1 alignment.
-The T1-to-T1 step is (statistically) independent of the FA values which are the ultimate statistical target of hypothesis testing in FA population studies.
-Below, we show that this same pattern of effects is reproduced in real data and detail our proposed solution.
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-%  \includegraphics[width=170mm]{workflow.pdf}
-  \includegraphics[width=130mm]{workflow.pdf}
+  \includegraphics[width=170mm]{workflow.pdf}
+%  \includegraphics[width=130mm]{workflow.pdf}
 \end{tabular}
@@
 
-We first briefly describe the SyGN algorithm used to create anatomical templates.  
-We then explain the the step-by-step process for aligning subject FA images to 
-the anatomical template (which is facilitated with the various tools available 
-in ANTs).  After describing the real image data used in this work, we show the 
-issues raised affect common analysis scenarios.
-
-%
-%
-%
-%It was precisely due to these concerns that the TBSS framework was developed constituting an entire processing pipeline spanning from the reconstruction of the DTI tensorial data
-%%to the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA)), to the nonrigid alignment of the population to a standard coordinate system and skeletonization, and, finally,
-%through to the statistical analysis for discerning differences across populations.
-Major intermediate steps include the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA) \citep{Basser1996}), nonrigid alignment of the population to a standard coordinate system, and subsequent ``skeletonization'' of the mean of the aligned FA maps.
-The alignment to a standard space is performed in such a way to remove the smoothing step associated with SPM.
-Skeletonization and its projection across all registered DTI-derived scalar images is meant to accommodate the possibility of imperfectly aligned data.
-However, as shown in a recent study \citep{Zalesky2011}, although the alignment error is reduced with the skeleton projection step, misalignments still remain post-projection.
-Added to this is the potential confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.
-
-%Because spatial normalization is critical to obtaining valid results, we propose a modification of the standard TBSS pipeline directed at some of these residual concerns which we refer to as ANTs-TBSS.
-Instead of DTI-derived scalar alignment via the images themselves (either random subject selection or locating an approximation to the ``mean subject'') or to a standard template (e.g.
-FMRIB58\_FA\_1mm.nii.gz which we denote as ``FMRIB58'' in the remainder of the text), we evaluate the substitution of an anatomical, i.e.
-T1-weighted, unbiased shape/intensity template constructed from the population images.
-DTI-derived images can then be mapped to their corresponding T1-weighted image in a highly constrained manner and subsequently mapped to the template via nonrigid registration of the T1-weighted image.
-This eliminates any possible complications associated with direct FA-to-FA image registration.
-
-%We first describe in greater detail the proposed ANTs-TBSS framework with an emphasis on comparison with standard TBSS.
-We then describe the two open data sets used to evaluate the proposed methodology in addition to a description of diffuse TBI data
-%also used to illustrate concepts discussed in this work.
-%and analysis previously reported in \cite{Avants2008}.   
-%The complementary assessment performed in this paper is used to provide additional analyses to these previous findings.
-
-%A crucial component of TBSS is the alignment of the DTI-derived scalar images \citep{Zalesky2011}. 
-%
-%Voxel-based morphometry vs. tractography vs. tract-specific analysis
-%
-%Talk about {\em skeleton projection} and how our work is different from \citep{Zalesky2011}.
-%Also talk about the advantages over building a dti-based template as discussed in 
-%\citep{Hecke2010}.
-
-\section{Material and Methods}
-
-%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.
-
-%In order to understand the differences in approaches, we first briefly describe
-%the current TBSS protocol as it is implemented.  This is followed by a
-%brief outline of the proposed modification using ANTs.  Afterwards
-%we give a description of the data sets used to demonstrate the methods.
-
-\subsection{Anatomical Image Template Construction}
-In general, we normalize images to to a standard coordinate system to
-reduce intersubject variability in population studies.  Various
+In general, normalization to a standard coordinate system to
+reduces intersubject variability in population studies.  Various
 approaches exist for determining the normalized space such as selection
@@
 templates.  
-One challenge with standard templates is that they may inadvertantly bias one's results by enabling better normalization of subjects to which the template is more similar.
+%One challenge with standard templates is that they may inadvertantly bias one's results by enabling better normalization of subjects to which the template is more similar.
 This issue is exacerbated when dealing with populations that have high variance (e.g.
 due to disease) and/or when one's normalization method is low-dimensional (not flexible enough to capture large shape differences).
-
 Population-specific templates (or optimal templates) alleviate some of
@@
 towards a representative image template \citep{Learned-Miller2006}.
-
 In contrast, Avants et al. explicitly model the geometric component of the 
@@
 Normalization (SyGN) is a powerful framework for producing optimal population-specific
-templates \citep{Avants2010} with arbitrary similarity metric choice. Given a set of representative images, 
-$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
-diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
-\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
-the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
-which minimize the following cost function:
-\begin{align}
-  \sum_{m=1}^M \left[
-    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
-    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
-    \right]
-\end{align}
-where $D$ is the diffeomorphic shape distance, 
-\begin{align}
-  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
-\end{align}
-dependent upon the choice of the linear operator, $L$, and
-$v$ is the diffeomorphism-generating velocity field, 
-\begin{align}
-  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
-\end{align}
-$\Pi$ is the choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
-virtual domain midway between each individual image and the current estimate of the template.
-
-With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
-to identity, iterative optimization
-involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
-updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
+templates \citep{Avants2010} with arbitrary similarity metric choice.
+
+%Given a set of representative images, 
+%$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
+%diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
+%\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
+%the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
+%which minimize the following cost function:
+%\begin{align}
+%  \sum_{m=1}^M \left[
+%    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
+%    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
+%    \right]
+%\end{align}
+%where $D$ is the diffeomorphic shape distance, 
+%\begin{align}
+%  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
+%\end{align}
+%dependent upon the choice of the linear operator, $L$, and
+%$v$ is the diffeomorphism-generating velocity field, 
+%\begin{align}
+%  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
+%\end{align}
+%$\Pi$ is the choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
+%virtual domain midway between each individual image and the current estimate of the template. 
+%
+%With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
+%to identity, iterative optimization
+%involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
+%updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
 %For the interested reader, a set of scripts has been created and is distributed with the ANTs library for facilitating template creation.
 
-\subsection{Anatomically-Based Population Normalization}
-We now detail a standard approach that uses SyGN and intra-subject inter-modality mapping to provide an unbiased strategy for FA normalization.
-An example of our proposed workflow applied to anatomically-based population normalization prior to FA analysis using TBSS is illustrated in Fig.
+Our proposed workflow applied to anatomically-based population normalization is illustrated in Fig.
 \ref{fig:workflow}.
@@
 \end{enumerate}
-The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender effects in a control population.
+%The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender effects in a control population.
 
 \subsection{Imaging Data}
+
+\subsubsection{NKI/Rockland Data for Generation of the Simulated DTI Gender Cohort}
+Data from the first 14 weeks of prospective data-sharing initiative sponsored by the 1000 Functional Connectomes Project (FCP)%
+\footnote{
+In support of open science, the 1000 Functional Connectomes Project (FCP---http://fcon\_1000.projects.nitrc.org) 
+was initiated on December 11, 2009 by various members of the MRI community \citep{Biswal2010}.
+Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
+Commitments from institutions such as the NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have resulted in prospective data repositories
+currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC---http://www.nitrc.org).
+}
+were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
+Due to various issues (e.g.
+lack of corresponding T1 image, failed DTI reconstruction, and age matching requirements) only 62 subjects (21 females and 41 males).
+Each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
+EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
+voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
+TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
+1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).
+Images were anonymized including defacing of the T1 images prior to upload.
 Diffusion tensor data associated with each of the following data sets were reconstructed from the diffusion weighted sequences using Camino%
@@
 http://camino.org.uk/
-}---an open source toolkit for diffusion MRI processing and analysis \citep{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.
+}---an open source toolkit for diffusion MRI processing and 
+analysis \citep{Cook2006} in combination with ANTs-based 
+registration for motion correction of the DTI sequence.
 
@@
 1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
-DTI images were also obtained. 
+DTI images were also obtained. Diffusion tensor data were reconstructed using Camino.
+
+\subsection{Simulated DTI Gender Cohort Creation}
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{cc}
+  \includegraphics[width=41mm]{B0Average_slice85.png} & 
+  \includegraphics[width=41mm]{DTIAverageFA_slice85.png} \\
+  (a) & (b) 
+\end{tabular}
+\caption{Mid-axial slice of the (a) average b0 image and (b) FA image derived from the DTI atlas used
+to create the simulated gender data sets.}
+\label{fig:simulated}
+\end{center}        
+\end{figure}
+
+To provide realistic ground truth FA data, we implemented 
+the work of Van Hecke et al. \cite{van-hecke2009} which 
+we have also made available to the public
+\cite{Tustison2011}.%
+\footnote{
+http://hdl.handle.net/10380/3315
+}  
+Given an input DTI atlas, b0 image, scheme file indicating
+the gradient directions, and 
+a cohort of aligned DTI data (to model intersubject variability), the
+user is able to create 
+simulated data sets with specified pathology and Rician noise.
+For the simulated experiments described in the next section we
+created a T1 template from 30 randomly chosen T1 images.  The
+template was then registered to the MNI152 template.  
+The remaining 32 T1 images were then registered to the
+template.  Each DTI and b0 image was mapped to the MNI152 map as 
+described above using the T1 to MNI composite mapping.  DTI 
+were reoriented using the PPD method \cite{alexander2001}.  The 62
+aligned DTI were averaged using the log-euclidean framework 
+\cite{arsigny2006} to create the DTI atlas.  Similarly, the 62 aligned
+b0 images were averaged to create the representative b0 image
+for simulated DTI creation.  The derived FA and b0 images are shown in 
+Fig. \ref{fig:simulated}.  An FA mask was created by thresholding 
+the FA image of the DTI atlas at 0.2.
+
+The 41 aligned male DTI were used to generate 18 ``Control'' 30-direction
+diffusion-weighted images
+with no simulated noise nor pathology.  Similarly, the 21 aligned female
+DTI were used to generate 19 ``Experimental'' diffusion-weighted images.  
+DTI were reconstructed using Camino which were then use to create
+individual FA images were
+created from the DTI images to use in the simulated experiments.
+The number of subjects in each
+group was chosen based on the median control and experimental
+subject numbers from a sampling of FA studies.
+
+%Drawing upon recent
+%literature, Van Hecke et al. point out that white matter pathology,
+%typically due to demyelination and axonal degeneration, is manifest 
+%in DTI by a decrease in longitudinal diffusivity and an increase
+%in transverse diffusivity.  This change corresponds to a decrease
+%in the first eigenvalue and increase in the second and third
+%eigenvalues, respectively, of the diffusion tensor.  These values
+%are specified via the mask option so that user-specified pathologies
+%can be tailored to different regions. One decision that
+%we made, which is not explicit in the text was how to
+%distribute the increase in eigenvalues 2 and 3 in the case
+%of pathology.  To address this omission, we assumed a constant
+%ratio before and after increasing the eigenvalues corresponding
+%to transverse diffusion.
+
+%We have included the short command
+%line help in Listing 1 for the interested reader who would like to 
+%reproduce our results.  
 %
-%\subsubsection{Participants}
-%{\color{red}{Quote:}} The data were collected as part of a larger study investigating the neural correlates of attention deficits and treatment responses of various psychoactive drugs in the survivors of TBI (principal investigator: J.W.).
-Thirty individuals with TBI and 20 healthy volunteers were recruited.
-We planned to recruit more TBI participants because data of TBI survivors are more likely to be discarded in a typical functional neuroimaging study due to movements in the scanner and poor behavioral performance.
-TBI participants were recruited from a variety of clinical services at MossRehab and through a consent-based registry of individuals with TBI who are interested in participating in rehabilitation research.
-To be included, participants had to be between the ages of 16 and 60, and to have a history of non-penetrating traumatic brain injury of at least moderate severity at least 3 months prior to enrollment.
-Severity level was defined by significant and well-documented loss or alteration of consciousness following injury (i.e., lowest Glasgow Coma Scale (GCS) score of less than 12, or prospectively documented post-traumatic amnesia (PTA) of greater than 1 hour), or focal abnormality on a neuroimaging study that was attributable to traumatic injury.
-A subjective complaint of attention difficulties by the participant, treating clinician, or caregiver was also required.
-Potential participants were excluded if they had a history of prem orbid neurologic disease, psychosis, major affective disorder, mental retardation, Attention Deficit Hyperactivity Disorder, or if they were currently abusing alcohol or recreational drugs.
-Persons who were taking psychoactive medications other than anticonvulsants were also excluded.
-Participants and/or their involved caregivers (depending on the participantÕs cognitive capacity) provided informed consent.
-The study protocol was approved by the Albert Einstein Healthcare Network and the University of Pennsylvania IRBs.
-Twenty healthy volunteers, matched to patients for age, gender, handedness, years of education, and ethnicity, participated in the study.
-Control participants were recruited based on the same inclusion/exclusion criteria as patients, with the exception that they never had a TBI resulting in loss or alteration of consciousness, nor suffered attention complaints.
-Control participants were recruited through the family and friendship networks of the participants with TBI, and through public advertising.
-%Data from one TBI participant whose MRI scan showed a large area of encephalom alacia over almost the entire right hemisphere were excluded.
-The remaining 29 subjects with TBI included 21 men and 8 women aged between 18 and 58 years (mean age = 36.9, SD = 11.4) with a mean education of 13.1 years (SD = 2.8).
-Twenty three of them were right-handed (Edinburgh Handedness Inventory, Oldfield, 1971).
-Fourteen of them were Caucasians, 10 African Americans, 4 Hispanics, and 1 Asian.
-Selected demographic and clinical characteristics of the TBI survivors are reported in Table 1.
-Control participants included 17 men and 3 women aged between 21 and 50 years (mean age = 34.9, SD = 9.8) with a mean education of 13.1 years (SD = 1.7).
-Sixteen of them were right-handed.
-Eleven of them were Caucasians, 7 African Americans, 1 Asian, and 1 unknown.
-The two groups did not differ significantly in terms of age, gender, ethnicity, handedness or years of education (tested with t-test or FisherÕs exact test, as appropriate).
+%\begin{figure}
+%\lstsetcpp
+%\begin{lstlisting}
+%COMMAND: 
+%     CreateDTICohort
 %
-%\subsubsection{Image acquisition}
-%The functional imaging was conducted on a Siemens 3.0 T Trio whole-body scanner (Siemens AG, Erlangen, Germany), using a standard Transmit/Receive head coil.
-High resolution T1-weighted anatomic images were obtained using 3D M PRAGE imaging sequence using the following acquisition parameters: TR = 1620 ms, TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 1NEX with a scan time of 6 minutes.
-The resulting voxel size was 1 mm$^3$.
-%Lesion assessment
-%To quantify the volume of lesions, a trained observer (J.P.) manually segmented the lesion area under supervision of a neurologist (H.B.C.) with extensive experience in lesion assessment.
-Focal lesions included any cystic cavities and other focal regions of abnormal signal in the white or gray matter.
-The ITK-SNAP software \citep{Yushkevich2006} was used for a 3D-based segmentation.
-Total lesion volume was calculated using a stand-alone utility provided by VoxBo software (Center for Functional Neuroimaging, Philadelphia, PA, http://www.voxbo.org).
-
-\subsubsection{Data from the International Neuroimaging Data-sharing Initiative (INDI)}
-
-In support of open science, the 1000 Functional Connectomes Project (FCP)%
-\footnote{
-ttp://fcon\_1000.projects.nitrc.org
-} 
-was initiated on December 11, 2009 by various members of the MRI community \citep{Biswal2010}.
-Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
-Commitments from institutions such as the NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
-currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC).%
-\footnote{
-http://www.nitrc.org
-}
-%Data from these two institutions were used to demonstrate the alignment properties obtained using SyGN for DTI-derived data normalization.
-
-\paragraph{NKI/Rockland Sample}
-Data from the first 14 weeks of prospective data-sharing were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
-Due to various issues (e.g.
-lack of corresponding T1 image, failed DTI reconstruction, and age matching requirements) only 57 subjects were used for the first alignment analysis and 27 subjects of the NKI data set were used for the second analysis consisting of 9 females (age: $29.3 \pm 6.6$ years) and 18 males (age: $ 26.6 \pm 6.9$ years).
-Germane to our evaluation, each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
-EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
-voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
-TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
-1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).  Images were anonymized including defacing of the T1 images prior to upload.
-
-%\paragraph{NYU Institute for Pediatric Neuroscience Sample}
-%Data from the first 3 parts of prospective data-sharing from NYU were also downloaded on January 15, 2011 comprised of 49 subjects (average age in years: 30.29 $\pm$ 8.81).
-All subjects were used in the image alignment assessment.
-Two 64-directional DTI scans were performed (parameters: conventional single-shot spin echo
-%EPI pulse sequence, TR = 5200 ms, TE = 78 ms, axial acquisition, 
-%voxel size = $3 \times 3 \times 3$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2530 ms, TE = 3.25 ms, flip angle = 7$^\circ$, 256 contiguous slices of 1.33 mm, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1.3 mm$^3$).
-Images were anonymized including defacing of the T1 images prior to upload.
-
-%
-%The 1000 Functional Connectomes Project (FCP)%
-%\footnote{
-%http://fcon\_1000.projects.nitrc.org/index.html
-%}
-%
-%{\color{red}{Quote}}  INDI is a next-generation data-sharing effort, with the goal of sharing phenotypically rich imaging datasets, on both retrospective and prospective bases.
-On 10/8/10, the Nathan Kline Institute (NKI) / Rockland Sample led the way for INDI-Prospective, starting weekly uploads of 5-7 subjects / week (R-fMRI, DTI, and extensive psychiatric and cognitive phenotypic data, including IQ), which continue to progress (currently n = 55 and growing!).
-More recently, the NYU Institute Pediatric Neuroscience released its sample examining cocaine dependence in adults via INDI-Retro.
-The list of sites pledging data from around the world is growing, and we are working to generate common phenotypic variables across datasets as possible (notice the frequency with which IQ is starting to appear Ð itÕs just one of our targets!).
-%
-%\subsubsection{Nathan Kline Institute (NKI) / Rockland Sample}
-%
-%{\color{red}{Quote}}  The NKI/Rockland Sample is intended to be a phenotypically rich neuroimaging sample, consisting of data obtained from individuals between the ages of 6 and 85 year-old.
-All individuals to be included in the sample undergo semi-structured diagnostic psychiatric interviews, and complete a battery of psychiatric, cognitive and behavioral assessments in order to provide comprehensive phenotypic information for the purpose of exploring brain/behavior relationships.
-%
-%Each subject consists of the following data:
-%\begin{itemize}
-%\item 10 minute resting state fMRI scan (R-fMRI) 
-%\item 6-direction diffusion tensor imaging (DTI) scan 
-%\item 64-direction diffusion tensor imaging scan (2mm isotropic)
-%\item MPRAGE anatomical scan
-%\item A variety of psychiatric, cognitive and behavioral assessments
-%\end{itemize}
-%
-%\subsubsection{NYU Institute for Pediatric Neuroscience}
-%
-%{\color{red}{Quote}}  Summary: The NYU Phyllis Green and Randolph C?wen Institute for Pediatric Neuroscience (IPN) sample will be a collection of past and present scans obtained from psychiatrically screened individuals ranging in age from 6 to 55 years old.
-The initial release consists of datasets from 49 psychiatrically neurotypical adults, with age, gender and intelligence quotient (IQ) information provided.
-Future releases will include more comprehensive phenotypic information, and child and adolescent datasets, as well as individuals from clinical populations.
-%
-%Each subject consists of the following data:
-%\begin{itemize}
-%\item At least one 6-minute resting state fMRI scan (R-fMRI)%
-%\footnote{
-% Most participants have 2 R-fMRI scans, collected less than 1 hour apart in the same scanning session. Rest$_1$ is always collected first.
-%}
-%\item One high-resolution T1-weighted mprage, defaced to protect patient confidentiality
-%\item Two 64-direction diffusion tensor imaging scans
-%\item Demographic information (age, gender) and IQ-measures (Verbal, Performance, and Composite; Weschler Abbreviated Scale of Intelligence [WASI])
-%\end{itemize}
-%
-
-\subsection{FMRIB Software Library (FSL)}
-
-FSL hosts a large number of algorithms for computational neuroanalysis.
-These include, among others, well-known tools for image segmentation (FAST---\citep{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), and TBSS and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).
-Based solely on the number of citations in the literature,%
-\footnote{
-As of the beginning of the year 2011, the original Neuroimage 2006 TBSS paper has been cited more than 360 times making it the most cited in the journal's library.
-}
-usage of TBSS has become a de facto quantitative tool for DTI analysis based, at least in part, on its accessibility and ease of use.
-Due to its large presence in the neuroimaging community, many of the comparisons that are performed in the next section use the FMRIB58 template (shown in the upper left of Figure \ref{fig:workflow}).
-This template serves as the final coordinate system for most of the comparisons whether it be based on direct normalization or T1 normalization to the MNI152 template which resides in the FMRIB58 space.
-We also used the normalization component of TBSS composed of FLIRT and FNIRT registration steps optimized for normalization to FMRIB58 to provide an external comparative baseline (which we denote in subsequent sections as ``tbss'').
-
-%For the comparative evaluation studies used in this manuscript, TBSSv1.2%
-%\footnote{
-%Build 416 downloaded from the FMRIB web site http://www.fmrib.ox.ac.uk/fsldownloads/ on November 18, 2010.
-%}
-%was used.  The actual code pipeline constituting TBSS prior to statistical analysis consists of the following 4 shell scripts:
-%\begin{description}
-%  \item[tbss\_1\_preproc]  Input is a directory of FA images which are slightly eroded and then organized into a separate FA directory.
-A mask is also generated for each input FA image for use during the spatial normalization process.
-%  \item[tbss\_2\_reg]  The organized FA data and corresponding masks produced in the previous step are warped to one of three possible coordinate systems using FLIRT/ FNIRT:  1)  the atlas included with the FSL download---FMRIB58---(recommended), 2) a user-specified target image or 3) an exhaustive search for the ``optimal'' target image.
-In \citep{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.
-However, even though it would be faster to simply choose a single subject at random, the complexities of nonrigid image registration would seemingly promote the exhaustive search over random selection.
-As a compromise between these two options, the first option in the script is to use FMRIB58 which also resides in the space of the MNI152 T1 template.
-It should also be noted that this particular script invokes another FSL script, {\bf fsl\_reg}.
-An option associated with {\bf fsl\_reg} permits the use of an FNIRT configuration file which has been optimized for use with the FMRIB58 atlas.
-%  \item[tbss\_3\_postreg] The mean FA image is generated from the aligned FA images which is subsequently skeletonized.
-%  \item[tbss\_4\_prestats] The skeleton mask is created by thresholding the mean FA skeleton.  The skeleton distance map is then generated and all FA data are projected onto the skeleton.
-%\end{description}
+%OPTIONS: 
+%     -d, --image-dimensionality 2/3
+%     -a, --dti-atlas inputDTIAtlasFileName
+%     -x, --label-mask-image maskImageFileName
+%                            lowerThresholdValue
+%     -n, --noise-sigma <noiseSigma=18>
+%     -p, --pathology label[<percentageChangeEig1=-0.05>,
+%             <percentageChangeAvgEig2andEig3=0.05>,
+%             <numberOfVoxels=all or percentageOfvoxels>]
+%     -w, --dwi-parameters [B0Image,directionFile,bvalue]
+%                          [B0Image,schemeFile]
+%     -r, --registered-population dtiFileNames.txt
+%     -o, --output [outputDirectory,fileNameSeriesRoot,
+%             <numberOfControls=10>,
+%             <numberOfExperimentals=10>]
+%     -h 
+%     --help 
+%\end{lstlisting} 
+%\end{figure}
 
@@
 
-\subsection{TBI Cohort}
+\subsection{Simulated data set}
 
@@
 \begin{center}
-\begin{tabular}{c}
+\begin{tabular}{cc}
+  \includegraphics[width=86mm]{ISV_SpeSenAlreadyAligned.pdf} &
+  \includegraphics[width=86mm]{ISV_SpeSen_projectionAlreadyAligned.pdf}
+\end{tabular}
+\caption{ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
+different simulated experiments for both (a) voxelwise analysis over the
+entire FA mask and (b) using TBSS projections onto the mean FA skeleton.
+Degree of locality and correspondence to 
+Eqn. (\ref{eq:variance}) resulted in performance disparity between the
+different metrics.  Since SSD and the SSD-like Demon's formulation 
+minimize voxelwise group variance, we see the greatest bias in those 
+two metrics.  In contrast, MI and CC are less susceptible to bias with
+MI performing quite well.  Comparing voxelwise and projection results, 
+it is apparent that the TBSS projection strategy only slightly improves 
+the effects of bias for the simulated data.
+}
+\label{fig:simulated_plots}
+\end{center}        
+\end{figure*}
+
+To demonstrate the significance of circularity bias on voxelwise 
+analysis and to demonstrate its correspondence with different
+common similarity metrics, we used the simulated DTI images 
+described earlier to quantify such effects in FA.
+Since each simulated DTI and derived FA map was generated in the
+MNI152 template space, the cohort is already anatomically
+aligned.  However, as described in the introduction, localized
+image registration using common similarity metrics drive the 
+alignment towards minimizing local intensity differences (not necessarily
+anatomical differences) thereby
+inducing false positives.  To test this, we applied ANTs registration
+using default regularization and transformation parameters with
+four common similarity metrics:  neighborhood cross-correlation (CC), 
+Demons, mutual information (MI), and sum of squared differences (SSD).
+We limited optimization to 10 iterations at the full resolution 
+level.%
+\footnote{
+Specifically, the ANTs command line parameters common to each metric were:
+{\tt -i 10 -t SyN[0.25] -r Gauss[3,0]}.
+}
+Since neither noise nor pathology was introduced, only the differences
+associated with the gender cohort modeling is responsible for driving
+the registration.
+
+Voxelwise analysis included performing Student's t-test within the
+DTI masked region (FA $\geq 0.2$) for both the ground truth (i.e.
+anatomically aligned) data set and performing the same calculations
+for each of the four registration scenarios.  We used a significance
+level cutoff of $p = 0.05$ to determine ground truth significant
+voxels in the already aligned data set.  We then varied the 
+significance level cutoff for each of the registered data sets to plot
+the corresponding ROC curves given in Fig. \ref{fig:simulated_plots}.
+Visual comparison of the significant regions associated with the 
+voxelwise analysis for ground truth data and with the Demons registration data for
+$p \leq 0.05$ shown in Fig. \ref{fig:simulated_significance} demonstrates 
+the potential severity metric choice can have on results.
+
+We also performed a similar assessment
+where instead of performing voxelwise calculations over the entire
+white matter mask, we performed the FA skeleton projection characteristic
+of TBSS using the relevant portions of the TBSS scripts found in the 
+FMRIB software library (FSL).%
+\footnote{
+http://www.fmrib.ox.ac.uk/fsl/
+}
+The FA 
+image produced from the
+DTI atlas was used to produce a common skeleton for all comparisons.
+The associated ROC curves are given
+in the right plot of Fig. \ref{fig:simulated_plots}.  Although projection
+does have slightly ameliorating effects, the associated bias is still 
+significant across certain metrics.
+
+Despite the experimental set-up ensuring anatomical alignment in the
+ground truth data, the resulting bias effects accord with our earlier 
+description of the problem with these similarity metrics and direct
+FA-to-FA template normalization.  
+Whereas MI shows very little effects of bias due to it's statistical
+formulation and relative non-locality, the SSD and SSD-like Demons
+metrics are formulated such that they explicitly minimize voxelwise
+group variance and thus induce such bias effects.
+
+It should be noted that all the effects described are absent
+any multiple comparisons corrections or smoothing prior to statistical
+testing.  Although such techniques can minimize 
+or even possibly eliminate these detrimental confounds, our contention
+is (and these results demonstrate) that normalization choices can significantly
+impact the results prior to any such
+post hoc corrections and scientific best practices dictate that, where possible, 
+experimental adjustments eliminating or minimizing potential sources of bias should
+be made prior to experimental execution and analysis.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{cc}
+  \includegraphics[width=41mm]{no_registration_voxelwise_slice90.png} &
+  \includegraphics[width=41mm]{demons_voxelwise_slice90.png} \\
+  (a) & (b)
+\end{tabular}
+\caption{(a) Mid-axial slice of the DTI atlas-derived FA map where the 
+significant differences ($p < 0.05$) between control (male) and 
+experimental (female) groups are labeled in red.  (b) Same mid-axial 
+slice showing an increase in significant regions following ANTs 
+registration employing the Demons metric.
+}
+\label{fig:simulated_significance}
+\end{center}        
+\end{figure}
+
+\subsection{TBI Cohort}
+
+%\begin{figure*}
+%\begin{center}
+%\begin{tabular}{c}
 %  \includegraphics[width=180mm]{tbiTesting.pdf}
-  \includegraphics[width=130mm]{tbiTesting.pdf}
-\end{tabular}
-\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
+%\end{tabular}
+%\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
 Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn.
@@
 TBSS and the 4 ANTs metric
-results were all obtained using direct registration involving subject FA and the FMRIB58 template.
+%results were all obtained using direct registration involving subject FA and the FMRIB58 template.
 The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
-\label{fig:tbi_testing}
-\end{center}        
-\end{figure*}
-
-We constructed a patient template and a control template from the T1-weighted images of the two groups.  These two 
+%common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
+%\label{fig:tbi_testing}
+%\end{center}        
+%\end{figure*}
+
+Even under ideal conditions (i.e. no noise, no pathology, simple modeling of 
+intersubject variation, and precise anatomical alignment), the simulated
+data demonstrated the presence of circularity bias that varied with similarity
+metric.  In this section, we use data that was analyzed in \cite{Stone2011}
+and show how bias is potentially manifested in real data.
+
+We constructed a patient template and a control template from the 
+T1-weighted images of the two groups.  These two 
 templates were then used to create a third common template.  
@@
 
-Cumulative distribution functions (one-tailed, $\alpha = 0.95$)
-are shown in Fig. \ref{fig:tbi_testing}.
-Similar to previous findings with the simulated data, the SSD and 
-Demons metric show elevated statistical power using the Student's 
-$t$-test over the other alignment protocols which all exhibit similar 
-strength.  However, this discrepancy decreases significantly with the 
-nonparametric testing.  The Demons and SSD power decreases while the others
+Cumulative distribution functions (CDF) of the $p$-values (one-tailed, $\alpha = 0.95$)
+over the entire FA mask (FA $\geq 0.2$) for the different normalization
+strategies are shown in Fig. \ref{fig:tbi_testing}.
+A bias which (falsely) decreases $p$-values will demonstrate a left-shifted
+CDF. 
+  The Demons and SSD power decreases while the others
 remain fairly similar.  Interestingly, the ANTs template-based approach
@@
 
-\begin{table}
+%This is very much related to earlier findings in that strategies evincing
+%greater alignment, as quantified by the CV,
+%will trend towards smaller variance thus exhibiting a left-shifted
+%CDF.
+
+%Similar to previous findings with the simulated data, the SSD and 
+%Demons metric show an elevated effect using the Student's 
+%$t$-test over the other alignment protocols which all exhibit similar 
+%strength.
+
+\begin{figure}
 \begin{center}
-  \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
-  \label{table:lom}
-  \begin{tabular*}{85mm}{@{\extracolsep{\fill}} cccccc}
-  \hline
-%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-%  \cline{2-5}
-  {Norm.} & Target & Jaccard & Dice  & False $-$ & False $+$\\
-  \hline
-  \hline
-  antsT1 & 0.917 & 0.621 & 0.767 & 0.082 & 0.341 \\
-  antsCC & 0.780 & 0.570 & 0.727 & 0.220 & 0.320 \\
-  antsMI & 0.776 & 0.569 & 0.725 & 0.223 & 0.320 \\
-  antsDemons & 0.780 & 0.566 & 0.723 & 0.220 & 0.326 \\
-  antsSSD & 0.783 & 0.562 & 0.720 &  0.217 & 0.334 \\
-  tbss & 0.716 & 0.550 & 0.710 & 0.283 & 0.295 \\
-  \hline
-  \end{tabular*}
+\begin{tabular}{c}
+  \includegraphics[width=85mm]{tbittest.pdf}
+\end{tabular}
+\caption{Student's $t$-test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
+Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn.
+(\ref{eq:variance}), namely the SSD and Demons metric.
+Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.
+TBSS and the 4 ANTs metric
+results were all obtained using direct registration involving subject FA and the FMRIB58 template.
+The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
+common T1 template $\rightarrow$ MNI152 (FMRIB58) template.  
+}
+\label{fig:tbi_testing}
 \end{center}        
-\end{table}
-
-%\begin{table}
-%\begin{center}
-%  \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
-%  \label{table:lom}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccccc}
-%  \hline
-%  {Norm.} & Target & Jaccard & Dice  & False $-$ & False $+$\\
-%  \hline
-%  \hline
-%  antsT1 & 0.917 & 0.621 & 0.767 & 0.082 & 0.341 \\
-%  antsCC & 0.780 & 0.570 & 0.727 & 0.220 & 0.320 \\
-%  antsMI & 0.776 & 0.569 & 0.725 & 0.223 & 0.320 \\
-%  antsDemons & 0.780 & 0.566 & 0.723 & 0.220 & 0.326 \\
-%  antsSSD & 0.783 & 0.562 & 0.720 &  0.217 & 0.334 \\
-%  tbss & 0.716 & 0.550 & 0.710 & 0.283 & 0.295 \\
-%  \hline
-%  \end{tabular*}
-%\end{center}  
-%\end{table}
+\end{figure}
 
@@
 trend towards greater alignment as quantified by the CV.
-
-%This is very much related to earlier findings in that strategies evincing
-%greater alignment, as quantified by the CV,
-%will trend towards smaller variance thus exhibiting a left-shifted
-%CDF.
-
 Although direct FA-to-FA registration using ANTs with common
@@
 Type 1 error problem (as evidenced in Fig. \ref{fig:tbi_testing}) 
-but also show a decrease in alignment.  This is further supported
-by the overlap measures \citep{Crum2006} given in Table \ref{table:lom}.
-For each alignment strategy, we thresholded the derived mean FA image (FA $
-\geq$ 0.2) and compared it with white matter mask of the MNI152 template
-segmented using Atropos \citep{Avants2011a}.  The template-based
-strategy showed the best overall overlap with the white matter MNI152
-segmentation.
+but also show a decrease in alignment.
 
@@
 is juxtaposed 
-with the mean FA image produced by the TBSS pipeline alone (Fig.\ref{fig:meanFA}(a)).
+with the mean FA image produced by the TBSS pipeline alone (Fig. \ref{fig:meanFA}(a)).
 The skeleton masks were produced by thresholding the skeleton values at 
@@
 
-%
-%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Fig.
-\ref{fig:hoonData}.
-The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.
-
-%As an additional quantitative comparison, we calculated the volumetric ratio between areas of significance and the volumetric brain mask for both $p = 0.01$ and $p = 0.05$.
-For standard TBSS, the MNI 152 "first brain mask" was used and for the ANTs template, Atropos was used to obtain a mask by skull stripping of the final template.
-These values are given in Table \ref{table:hoon}.
-%
-%
-%\begin{table}
-%  \caption{Volumetric ratio differences between ANTsTBSS and TBSS.}
-%  \label{table:hoon}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccc}
-%  \hline
-%  {} & $\frac{V(\mathrm{mean\,FA\,mask})}{V(\mathrm{brain\,mask})}$ & $\frac{V(\mathrm{sig.\,areas})}{V(\mathrm{mean\,FA\,mask})}$ & $\frac{V(\mathrm{sig.\,areas})}{V(\mathrm{brain\,mask})}$ \\
-%  \hline
-%  $p \leq 0.01$ & {} & {} & {} \\
-%  $p \leq 0.05$ & {} & {} & {} \\
-%  \hline
-%  \hline
-%  \end{tabular*}
-%\end{table}
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{hoonData2.pdf}
-%\end{tabular}
-%\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).
-Axial slices were chosen based on visual assessment of correspondence in both the ANTs template and MNI 152 template.
-%On the left is the thresholded skeleton overlaid on the mean FA image.  
-%On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).
-The yellow arrow points to an area of significant difference in the peripheral white matter found using ANTs-TBSS.
-%}
-%\label{fig:hoonData}
-%\end{figure}
-
-\subsection{INDI NKI/Rockland Sample}
-In this section, further exploration of relevant issues employs 
-the NKI/Rockland sample of the INDI database.  
-The first is one of experimental set-up consideration in which 
-we investigate how template-based
-alignment is affected by the number of subject images used to 
-create the template.  Both FA and mean diffusivity (MD) images were used.
-This is followed by statistical analysis
-of discriminating power of gender FA differences using 
-the various alignment strategies.
-
-\subsubsection{Alignment Assessment of Template Group Size}
-
- \begin{figure}
-  \begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{NKI2.pdf}
-\end{tabular}
-\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.
-(a) The standard FMRIB58 template.
-(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
-(c) and (d) The mean FA image produced using the ANTs template-based strategy employing 5 and 30 randomly chosen subjects, respectively, warped to the space of (a) followed by T1 registration of all subjects to the resulting anatomical template.
-Note the visual alignment difference particularly in the peripheral white matter complex.
-}
-\label{fig:NKI}
-\end{center}
-\end{figure}
-
-Four templates were generated from $\{5,10,20,30\}$ randomly selected subjects 
-to see if the number of subjects comprising the template affected the alignment 
-performance. In addition to the composite transforms
- to take each subject's FA or MD image to the population-specific anatomical template, we also registered
-each ANTs T1 template to the MNI152 template.  
-This permitted a
-standard space for qualitative and quantitative comparison with TBSS-aligned
-FA and MD images.  
-After template construction, the remaining
-images were registered to the resulting template and, ultimately, 
-to the MNI152 (FMRIB58) template via the template-to-MNI152 transformations.
-
-Sample axial slices for the FA FMRIB58-aligned mean images for the NKI data 
-set of template sizes of 5 and 30 images are provided in Fig. \ref{fig:NKI}.  
-We also include the same axial slice  of the FMRIB58 template in Fig. \ref{fig:NKI} 
-(a) for comparison.  In both sets of sample slices, visual inspection would  
-indicate better 
-performance with the ANTs-based FA mean images compared with the standard 
-TBSS-produced FA mean image.  Also, there is little discernible difference
- between the different ANTs results.  Alignment differences between the 
- approaches are particularly apparent in the peripheral white matter where 
- the anatomical structure in the ANTs template-based FA mean images is visually differentiable.
-
-Relative performance is quantified in the plots produced in Fig. \ref{fig:NKIplots}. 
-The mean FA image derived from the standard TBSS approach was thresholded at FA $\geq 0.2$. 
- This binary mask obtained in the NKI cohort provided the standard group mask for 
- fair comparisons for all evaluations.    For all FMRIB58-aligned images, the CV 
- was calculated in a voxelwise fashion within the mask.  Intuitively, 
- one would expect a smaller variance across better aligned sets of voxel values thus 
- producing a lower CV for a given mean value.  By calculating the CDF of the set of CV values within the mask, quantitative assessment of alignment performance can be visualized.  
-% In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
-This division was accomplished by applying a distance transform \citep{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.
-Note that only those voxels that were also within the thresholded FA mask were considered in the inner/outer analysis.
-
-There is a visible improvement in alignment for both data sets using the ANTs-based approach over individual subject FA registration to the FMRIB template.
-There is little difference between the ANTs results with only a slight trend in improved performance with the larger template subject sample sizes.
-%This was the case for all three regional CV calculations for both data sets.
-Also, as expected, there is greater misalignment in the outer region versus the inner core of the white matter mask regardless of methodology.
-However, the performance tally was the same in both the inner and outer regions as was the case with the entire white matter mask.% although the disparity in performance was much less for the NYU data set as the NKI data set most likely due to the differences in resolution of the original DTI.
-
-%The improvements seen in the CDF plots are also supported by comparative quantities derived from the skeletons of the different mean FA images.
-Since the skeletonization process is meant to determine the center of the various white matter tracts \citep{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.
-In addition, as described in \cite{Zalesky2011}, the number of connected components of the white matter skeleton is directly related to the contiguity and thus inversely related to the level of structure.
-Table \ref{table:vol} provides the volume of the mean FA skeletons (thresholded at FA $\geq 0.2$) for standard TBSS protocol using the FMRIB template and the ANTs approach using the various templates.
-In agreement with the CDF plots illustrating improved alignment with ANTs, the voxel volumes of the skeletons are greater using ANTs templates with the greatest difference seen in the outer portion of the white matter.
-%Related, in Table \ref{table:cc}, we provide the number of connected components for both data sets for all isolated components greater in size than 1 voxel and 2 voxels.
-Even though the skeletal volumes associated with the ANTs approach are greater, the number of connected components is less which would also support the observation of improved alignment.
-
-%\begin{table}
-%  \caption{Volume (in voxels) of thresholded mean FA skeletons.}
-%  \label{table:vol}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
-%  \hline
-%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-%  \cline{2-5}
-%  {} & inner & outer & inner & outer \\
-%  \hline
-%  FMRIB & 84899 & 122306 & 72765 & 76769 \\
-%  ANTs$_5$ & 87572 & 134421 & 73650 & 94177 \\
-%  ANTs$_{10}$ & 86908 & 130569 & 73419 & 98014 \\
-%  ANTs$_{20}$ & 86523 & 129729 & 73577 & 95484 \\
-%  ANTs$_{30}$ & 86971 & 128892 & 73443 & 94862 \\
-%  \hline
-%  \hline
-%  \end{tabular*}
-%\end{table} 
-%
-%\begin{table}
-%  \caption{Number of connected components of thresholded mean FA skeletons.}
-%  \label{table:cc}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
-%  \hline
-%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-%  \cline{2-5}
-%  {} & $\geq$ 1 voxel & $\geq$ 2 voxels & $\geq$ 1 voxel & $\geq$ 2 voxels \\
-%  \hline
-%  FMRIB & 4340 & 1802 & 2431 & 993 \\
-%  ANTs$_5$ & 3561 & 1384 & 2278 & 916 \\
-%  ANTs$_{10}$ & 3413 & 1302 & 2232 & 899 \\
-%  ANTs$_{20}$ & 3283 & 1252 & 2093 & 885 \\
-%  ANTs$_{30}$ & 3114 & 1243 & 2128 & 867 \\
-%  \hline
-%  \hline
-%  \end{tabular*}
-%\end{table}
-
-%Since our proposed modification of TBSS impacts strictly the
-%critical alignment step, the first set of evaluations demonstrates
-%the difference between the two methods, i.e. spatial normalization 
-%via registration of the FA maps to the FMRIB58 template
-%or spatial normalization via ANTs template construction.
-
-%It is important to note that we did not consider normalization either to a random subject
-%or to the subject best representative of the mean as defined in 
-%\citep{Smith2006}. The former was not investigated due to the possible bias
-%in selecting a random subject as described in \cite{Smith2006}.  
-%Neither was the latter option pursued due to the computational complexity
-%required for such an $\OO{n^2}$ search over the relatively large NKI data set.  In addition, normalization to
-%the FMRIB58 template is both the default and recommended option
-%with the parameters of FNIRT optimally tuned for said option.
-
-%To warp the MD images following FA registration we used FSL's {\bf applywarp} function using the FA-derived transform.
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{NYU.pdf}
-%\end{tabular}
-%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.
-(a) The standard FMRIB58\_FA\_1mm template.
-(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
-%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).
-Note the visual alignment difference particularly in the peripheral white matter complex.
-%}
-%\label{fig:NYU}
-%\end{figure}
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{inner.png}
-%\end{tabular}
-%\caption{
-%}
-%\label{fig:inner}
-%\end{figure}
-%
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{outer.png}
-%\end{tabular}
-%\caption{Three-canonical view of the inner
-%}
-%\label{fig:outer}
-%\end{figure}
-
-%\begin{figure*}
-%\begin{tabular}{c}
-%  \includegraphics[width=180mm]{NYUplots2.pdf}
-%\end{tabular}
-%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
-\ref{fig:NKI}(b)).
-In addition to the whole brain analysis (first column), the CV was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-%}
-%\label{fig:NYUplots}
-%\end{figure*}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{NKIplots3.pdf}
-\end{tabular}
-\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
-\ref{fig:NKI}(b)).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-}
-\label{fig:NKIplots}
-\end{center}
-\end{figure}
-
-\subsubsection{Statistical Power Related to Gender-Based FA Differences}
-
-For completeness, the second NKI data subset was used to explore the 
-power of the different alignment strategies in discriminating FA 
-differences in a control group divided by gender.  As described 
-earlier, this group consisted of 27 subjects 
-comprised of 9 females (age: $29.3 \pm 6.6$ years) and 18 males (age: $26.6 
-\pm 6.9$ years).  Each subject was registered to the template made from 
-30 randomly chosen subjects, described in the previous section, and
-subsequently warped to the MNI152 template space.
-
-Results are shown in Fig. \ref{fig:nki_testing} in which all but the more
-robust metrics (MI and CC) exhibit similar power.  Note that the magnitude
-of the effects is much smaller than for TBI.  Again, the template-based approach
-performs comparatively well without the negative effects associated with
-the issues perviously described.
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-%  \includegraphics[width=180mm]{nkiTesting.pdf}
-  \includegraphics[width=130mm]{nkiTesting.pdf}
-\end{tabular}
-\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results 
-(one-tailed, $\alpha = 0.95$) for the NKI FA data aligned to the FMRIB58 template 
-adjusted for multiple comparisons (FDR).  
-Similar statistical power is seen in all alignment scenarios except for the more robust
-metrics (CC and MI) in the $t$-test which show decreased power.   }
-\label{fig:nki_testing}
-\end{center}        
-\end{figure*}
-
 %\section{Discussion}
@@
 
-\section{Discussion} 
+\section{Discussion and Conclusions} 
 Prior evaluation work using anatomically labeled
@@
 intend to (in this paper) enter into discussion on the mechanisms of
-TBI, gender difference and DTI that may lead to the detected results.
+TBI, gender difference, and DTI that may lead to the detected results.
 Rather, we focus on the technical issue of circularity caused by
@@
 whole tensor DTI normalization \citep{Zhang2007,Hecke2007}, DTI template-based
-strategies \citep{Mori2009,Hecke2011}, {\color{red}{or other related methods \cite{jbabdi2010}}}.  Although investigation is beyond the scope
-of this work, these whole-tensor approaches might provide the necessary
-anatomical information to overcome the deficiencies addressed in this
-work.
+strategies \citep{Mori2009,Hecke2011}, or other related methods \cite{jbabdi2010}.  
+%Although investigation is beyond the scope
+%of this work, these whole-tensor approaches might provide the necessary
+%anatomical information to overcome the deficiencies addressed in this
+%work.
 
@@
 reexamine FA normalization and analysis methods and to consider the
-potential confounds highlighted in this paper.  Rank tests
+potential confounds highlighted in this paper and others
+\cite{Ridgway2008}.  Rank tests
 provide a more conservative alternative as does the methodology
@@
 
-\section{Conclusions}
-%The main conclusions of the study may be presented in a short Conclusions section, which may stand alone or form a subsection of a Discussion or Results and Discussion section.
-
-%The popularity of FSL's TBSS is due, in large part, to its ease of use.
-Anybody interested in performing a neuroimaging study in which DTI-derived measures are compared across populations can simply download FSL and immediately apply it to their data.
-The proposed work is a substantial improvement to the core TBSS framework which dovetails with FSL.
-In addition, it is comparably easy to download and apply what we have proposed in the form of ANTs-TBSS.
-
-With the increasing use of DTI for white matter assessment and the
-increasing number of researchers that are taking advantage of the numerous
-software tools that are available to facilitate this research, guidance
-is necessary to ensure the proper use of these tools and the
-calculation and reporting of findings \citep{Ridgway2008}.  The normalization 
-circularity issue described in this work is the inducement of substantial Type 1
-errors caused by FA-to-FA normalization using standard similarity metrics.
-To reduce these types of errors, we advocate the use of an anatomical template-building
-strategy whereby the transformation of FA images to a common space is performed
-using each individual's anatomical image.
-This proposed improvement potentially has many benefits:
-%which were listed in the 
-%introduction.  Such benefits range from the theoretical in the form of eliminating 
-%the confounding effects of registering data which is to be statistically 
-%differentiated to the practical, e.g. reusing the ANTs-based template for other 
-%neuroimaging analyses.  
-\begin{itemize}
-\item As explained in \cite{Avants2010}, ANTs optimal template construction creates a true Frech\'et mean from the population sample with high resolution which facilitates within-group registration.
-This obviates the need for an externally-derived template or an $\OO{n^2}$ search for an approximation to the mean subject or the bias resulting from random subject selection.
-\item Since the creation of the standardized coordinate system is performed using the anatomical images, nonrigid registration of the DTI-derived scalar images is 
-minimized thus mitigating the confounds described previously.
-Each DTI-based scalar image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \citep{Wu2008}.
-\item Although the total number of registrations to be performed is increased, this strategy makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
-%\item Instead of Gaussian smoothing as with SPM, TBSS uses the warping process to FMRIB58 by default to provide an indirect smoothing.
-Such resampling occurs in ANTs-TBSS as a by-product of the registration of the DTI-derived image to the corresponding T1 image in a subject-specific manner.
-%\item Instead of using thresholding to create the white matter segmentation from which the skeleton is created, an anatomically-based segmentation, either in the template space or in each subject space followed by a label fusion technique (e.g.
-majority voting), can be used to ultimately provide a more accurate skeleton.
-\item Oftentimes FA analysis is only one of several components performed in a population-based study.
-Other investigations, such as voxel- and tensor-based morphometry, cortical thickness measurements, etc.
-can take advantage of the anatomical template.
-\item The proposed modification is compatible with other frameworks such as TBSS where the normalization step is
-performed using ANTs-based tools followed by skeletonization and projection using the FMRIB software (see Fig. \ref{fig:workflow}).
-\end{itemize}
-
-All software components described 
-in this paper are publicly available as open source primarily through the 
-Advanced Normalization Tools (ANTs) repository.%
-\footnote{
-http://www.picsl.upenn.edu/ANTs
-}
-In addition to registration and template building, ANTs also includes brain 
-extraction \citep{Avants2010a}, bias correction \citep{Tustison2010}, $n$-tissue 
-segmentation \citep{Avants2011a}, and cortical thickness estimation solutions \citep{Das2009}.
-
-%In performing the various analyses described in the work, it was pleasantly surprising to discover how well the alignments were for the ANTs template data considering that alignment was based on FA-to-FA correspondence and no such image registrations were performed to align FA directly with FA as with TBSS.
-Furthermore, since TBSS is a local quantification method with increased statistical power through the use of the skeletonization and projection step, improved alignments can increase the regional sensitivity to local white matter differences particularly in the peripheral tracts.
-
 %% The Appendices part is started with the command \appendix;

### 2011-10-17 ANTsTBSS/paperRevision.tex
ENH:  Changed coption for Figure 1.

@@
 \end{tabular}
-\caption{Possible ANTs-based TBSS workflow using SyGN.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
+\caption{Possible ANTs-based workflow using SyGN for normalization of images for statistical analysis.  In contrast to standard  FA-to-FA template registration
 (transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.

### 2011-10-19 ANTsTBSS/paperRevision.tex
ENH:  Removed unnecessary phrase.
Take a break, Driver 8

@@
 fMRI studies, we demonstrate how this concept of circularity bias can 
-potentially corrupt voxel-based analyses with a particular emphasis on
+potentially corrupt voxel-based analysis with a particular emphasis on
 population studies of fractional anisotropy (FA).  
@@
 bias which varies in strength with the similarity metric used in
-registration.  Specifically, we show that the popular sum of squared
-difference intensity metric  maximizes effect size, rather than
-anatomical alignment, and consistently overestimates statistical
-significance.  Consequently, we advocate a more cautious approach
-that normalizes a population of FA images through the associated 
-anatomical images.  One possibility, the Symmetric Group Normalization
-strategy for unbiased template construction from anatomical (e.g.
-T1-weighted) images, provides a valid and unbiased coordinate system for
-statistical inference of FA differences.  
-We establish this
-outcome by using simulated data derived from normal controls from the International
-Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
-brain injury survivors compared with 17 control subjects.  
-Alignment
-comparisons with the normalization component of the popular tract-based
-spatial statistics (TBSS) show that our anatomically-based template
-approach produces good alignments despite the absence of any direct
-FA-to-FA registration.  
-%Of practical consideration we note that the
-%software tools discussed in this work are provided to the public as part
-%of the open source Advanced Normalization Tools (ANTs) repository.  
-In conclusion, we recommend against circular analysis strategies that 
-select image similarity metrics that are tied directly to the statistical 
-functions used to assess population effects. 
+registration.  Specifically, we analytically demonstrate that the popular sum of squared
+difference (SSD) intensity metric explicitly maximizes effect size rather than
+anatomical alignment---an experimental design flaw referred to as `circularity bias.'  
+Through the use of both simulated diffusion data and real data involving
+traumatic brain injury survivors, we explore the magnitude of this bias
+and how it varies with similarity metric under the hypothesis that while
+SSD-related metrics, such as Demons, will manifest similar effects, 
+other metrics which 
+are not formulated based on absolute intensity differences 
+(e.g. mutual information (MI) and cross correlation (CC)) will produce
+less of an effect. 
+Consequently, given the variability in voxel-based analysis outcomes 
+with similarity metric choice,
+we caution researchers specifically in the use of SSD 
+and SSD-related measures where normalization and statistical 
+analysis involve the same image set and, instead, advocate a more 
+cautious approach
+where normalization of the individual subject images
+to the statistical reference space occurs through the 
+corresponding (statistically independent) anatomical images to 
+an anatomical template.
+
+%One possibility, the Symmetric Group Normalization
+%strategy for unbiased template construction from anatomical (e.g.
+%T1-weighted) images, provides a valid and unbiased coordinate system for
+%statistical inference of statistical differences.  
+%
+%
+%
+%The first set of experiments explores the different outcomes 
+%
+%
+%We establish this
+%outcome by using simulated data derived from normal controls from the International
+%Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
+%brain injury survivors compared with 17 control subjects.  
+%Alignment
+%comparisons with the normalization component of the popular tract-based
+%spatial statistics (TBSS) show that our anatomically-based template
+%approach produces good alignments despite the absence of any direct
+%FA-to-FA registration.  
+%
+%
+%In conclusion, we recommend against circular analysis strategies that 
+%select image similarity metrics that are tied directly to the statistical 
+%functions used to assess population effects. 
 \end{abstract}
@@
 \end{align}
-is that which minimizes the voxelwise group variance, or at least an approximation thereof.
+is that which minimizes the voxelwise group variance.
 Reducing the group variance will also directly affect the value of a standard population statistic, the Student's $t$-test.
+Based on this derivation, one would also expect the popular Demons metric to be susceptible to this erroneous increase in statistical power due to identical objective functions although a more aggressive displacement/gradient at each iteration will relatively exacerbate this effect for a given number of iterations.
 Other standard similarity metrics will also exhibit this
-trend to some degree as they minimize some local measure of intensity-based distance between the subject and the template.
+trend to some degree as they minimize some local measure of intensity-based distance between the subject and the template although the effect will be minimized relative to the similarity to Eqn.
+(\ref{eq:variance}).
 
@@
 Although various processing choices mitigate the effects of such bias,
-such as Gaussian smoothing, nonparametric testing \citep{Rorden2007}, 
+such as Gaussian smoothing during and following normalization, 
+nonparametric testing \citep{Rorden2007}, 
 statistical analysis based on orthogonal projections onto the white matter skeleton 
@@
 to assess alignment (i.e. the similarity metric) in FA-to-T1 and T1-to-T1 
-normalizations are statistically independent of the intensity value differences used for 
+normalizations are statistic gally independent of the intensity value differences used for 
 the post-alignment numerical analysis.  It should be noted that this discussion
-is certainly not limited to studies involving FA.  From 
-Eqn. (\ref{eq:variance}), it is evident that such bias can occur anytime the 
-similarity metric (and the values employed) are non-independent of 
-the statistical analysis.
-
-Using simulated data, we first establish that this effect is real 
-and warrants concern on the part of researchers using 
-these techniques.  We then show how the general application
-of a high performance \citep{Klein2009}, nonrigid registration algorithm 
-\citep{Avants2011}, which forms the underlying machinery for an optimal 
-template construction algorithm \citep{Avants2010}, can facilitate alignment 
-between DTI-derived scalar images for cross-population studies.  
-Although this type of alignment procedure is not uncommon 
-(e.g. \cite{Lipton2008}), to our knowledge, this is the 
-first time a connection has been made with the concept of 
-circularity.
-
+is certainly not limited to studies involving FA.  It should be apparent
+that such bias can occur anytime the images used for normalization
+are non-independent of the statistical analysis as defined by the similarity
+metric and statistical testing.
+
+Two sets of experiments establish that the similarity metric
+can significantly affect voxel-based analysis outcomes and that,
+based on the previous description of circularity bias,
+the relationship in performance variability and similarity metric
+is to be expected.  Specifically, we show that whereas SSD and Demons
+substantially overestimate statistical significance, mutual
+information and cross correlation are less susceptible although the
+effect is not
+negligible.  Furthermore, we also show that this effects persists
+even with the common practice of post-normalization smoothing prior
+to mass univariate statistical testing.
+
+Given this general overview, the experimental sets can be briefly
+characterized as follows:
+\begin{itemize}
+  \item {\it Simulated DTI Data}:  Using a simulated DTI generator,
+  FA images based on male/female populations were generated in situ
+  thus guaranteeing anatomical alignment prior to processing.
+  The simulated images in each group were then ``registered'' to the FA image
+  of the defining DTI 
+  template using four common similarity metrics (CC, Demons, MI, and SSD).  
+  ROC curves based on voxel-based analysis for each of these scenarios demonstrate that 
+  effect size is maximized over anatomical alignment and that
+  this effect size varies with similarity metric.  Additional
+  experiments establish that this trend persists even for commonly used smoothing
+  kernels and the projection step of TBSS.
+  \item {\it Real TBI Data}:  FA data of TBI survivors and controls were 
+  normalized to a template using four common similarity metrics
+  (CC, Demons, MI, and SSD).  Additional normalization included normalization via
+  the subjects' anatomical T1-weighted images as well as the default registration for TBSS.
+  Voxelwise statistical testing was also performed with
+  and without smoothing.  Even though ground truth is not available for such
+  data, the SSD and Demons metrics showed increased statistical power as evidenced
+  by a general decrease in voxelwise $p$-values.
+\end{itemize}
+All major processing steps were performed using the open source Advanced Normalization Tools 
+(ANTs)%
+\footnote{
+http://www.picsl.upenn.edu/ANTs
+}
+software repository which should facilitate reproducibility for the interested 
+reader.
+
+% We then show how the general application
+%of a high performance \citep{Klein2009}, nonrigid registration algorithm 
+%\citep{Avants2011}, which forms the underlying machinery for an optimal 
+%template construction algorithm \citep{Avants2010}, can facilitate alignment 
+%between DTI-derived scalar images for cross-population studies.  
+%Although this type of alignment procedure is not uncommon 
+%(e.g. \cite{Lipton2008}), to our knowledge, this is the 
+%first time a connection has been made with the concept of 
+%circularity.  
 %The rest of the paper establishes the severity of 
@@
 
-\section{Methods and Materials}
+\section{Materials and Methods}
 % Provide sufficient detail to allow the work to be reproduced. 
@@
 
-\subsection{Template Construction for Anatomically-Based Population Normalization}
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=170mm]{workflow.pdf}
-%  \includegraphics[width=130mm]{workflow.pdf}
-\end{tabular}
-\caption{Possible ANTs-based workflow using SyGN for normalization of images for statistical analysis.  In contrast to standard  FA-to-FA template registration
-(transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.
-Each FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.
+%\subsection{Template Construction for Anatomically-Based Population Normalization}
+%
+%\begin{figure*}
+%\begin{center}
+%\begin{tabular}{c}
+%  \includegraphics[width=170mm]{workflow.pdf}
+%%  \includegraphics[width=130mm]{workflow.pdf}
+%\end{tabular}
+%\caption{Possible ANTs-based workflow using SyGN for normalization of images for statistical analysis.  In contrast to standard  FA-to-FA template registration
+%(transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.
+Each FA image is then rigidly mapped, with distortion correction, to the individual T1-weighted using the white matter mask as shown on the left.
 The individual T1-weighted image is then nonrigidly registered to the anatomical template.
-Optionally, one can register the T1 template to the MNI152 template which resides in the same space as the FMRIB58 template.
+Optionally, one can register the T1-weighted template to the MNI152 template which resides in the same space as the FMRIB58 template.
 Thus, the composite transform, carefully constructed without any FA-to-FA image registration, can take an individual subject's FA map (or other DTI-derived measure) to the desired standardized coordinate system.
 }
-\label{fig:workflow}
-\end{center}
-\end{figure*}
-
-In general, normalization to a standard coordinate system to
-reduces intersubject variability in population studies.  Various
-approaches exist for determining the normalized space such as selection
-of a pre-existing template based on a single subject, e.g. the Talairach
-atlas \citep{Talairach1988}, or a publicly available averaged group of
-subjects, e.g. the MNI \citep{Collins1994} or ICBM \citep{Mazziotta1995}
-templates.  
-%One challenge with standard templates is that they may inadvertantly bias one's results by enabling better normalization of subjects to which the template is more similar.
+%\label{fig:workflow}
+%\end{center}
+%\end{figure*}
+%
+%
+%
+%In general, normalization to a standard coordinate system to
+%reduces intersubject variability in population studies.  Various
+%approaches exist for determining the normalized space such as selection
+%of a pre-existing template based on a single subject, e.g. the Talairach
+%atlas \citep{Talairach1988}, or a publicly available averaged group of
+%subjects, e.g. the MNI \citep{Collins1994} or ICBM \citep{Mazziotta1995}
+%templates.  
+%%One challenge with standard templates is that they may inadvertantly bias one's results by enabling better normalization of subjects to which the template is more similar.
 This issue is exacerbated when dealing with populations that have high variance (e.g.
 due to disease) and/or when one's normalization method is low-dimensional (not flexible enough to capture large shape differences).
-Population-specific templates (or optimal templates) alleviate some of
-these issues by deriving a most representative image from the population
-\citep{Good2001}.  Large deformation registration algorithms also reduce
-this confound by being less sensitive to the deformation distance
-between subject and target.  Some approaches combine both advantages,
-for instance, the diffeomorphic approach of Joshi et al. employs the
-SSD metric and a shape distance to bring the subject group of images
-into alignment \citep{Joshi2004}.  Variants
-include extension to multiple modalities \citep{Lorenzen2006} and small deformations
-\citep{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
-towards a representative image template \citep{Learned-Miller2006}.
-In contrast, Avants et al. explicitly model the geometric component of the 
-normalized space during optimization.  Coupling the intrinsic symmetry of 
-SyN pairwise registration \citep{Avants2011} and an
-optimized sharpening/averaging of the template appearance, Symmetric Group
-Normalization (SyGN) is a powerful framework for producing optimal population-specific
-templates \citep{Avants2010} with arbitrary similarity metric choice.
-
-%Given a set of representative images, 
-%$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
-%diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
-%\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
-%the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
-%which minimize the following cost function:
-%\begin{align}
-%  \sum_{m=1}^M \left[
-%    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
-%    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
-%    \right]
-%\end{align}
-%where $D$ is the diffeomorphic shape distance, 
-%\begin{align}
-%  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
-%\end{align}
-%dependent upon the choice of the linear operator, $L$, and
-%$v$ is the diffeomorphism-generating velocity field, 
-%\begin{align}
-%  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
-%\end{align}
-%$\Pi$ is the choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
-%virtual domain midway between each individual image and the current estimate of the template. 
-%
-%With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
-%to identity, iterative optimization
-%involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
-%updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
-%For the interested reader, a set of scripts has been created and is distributed with the ANTs library for facilitating template creation.
-
-Our proposed workflow applied to anatomically-based population normalization is illustrated in Fig.
+%Population-specific templates (or optimal templates) alleviate some of
+%these issues by deriving a most representative image from the population
+%\citep{Good2001}.  Large deformation registration algorithms also reduce
+%this confound by being less sensitive to the deformation distance
+%between subject and target.  Some approaches combine both advantages,
+%for instance, the diffeomorphic approach of Joshi et al. employs the
+%SSD metric and a shape distance to bring the subject group of images
+%into alignment \citep{Joshi2004}.  Variants
+%include extension to multiple modalities \citep{Lorenzen2006} and small deformations
+%\citep{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
+%towards a representative image template \citep{Learned-Miller2006}.
+%In contrast, Avants et al. explicitly model the geometric component of the 
+%normalized space during optimization.  Coupling the intrinsic symmetry of 
+%SyN pairwise registration \citep{Avants2011} and an
+%optimized sharpening/averaging of the template appearance, Symmetric Group
+%Normalization (SyGN) is a powerful framework for producing optimal population-specific
+%templates \citep{Avants2010} with arbitrary similarity metric choice. 
+%
+%%Given a set of representative images, 
+%%$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
+%%diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
+%%\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
+%%the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
+%%which minimize the following cost function:
+%%\begin{align}
+%%  \sum_{m=1}^M \left[
+%%    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
+%%    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
+%%    \right]
+%%\end{align}
+%%where $D$ is the diffeomorphic shape distance, 
+%%\begin{align}
+%%  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
+%%\end{align}
+%%dependent upon the choice of the linear operator, $L$, and
+%%$v$ is the diffeomorphism-generating velocity field, 
+%%\begin{align}
+%%  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
+%%\end{align}
+%%$\Pi$ is the choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
+%%virtual domain midway between each individual image and the current estimate of the template. 
+%%
+%%With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
+%%to identity, iterative optimization
+%%involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
+%%updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
+%%For the interested reader, a set of scripts has been created and is distributed with the ANTs library for facilitating template creation.
+%
+%Our proposed workflow applied to anatomically-based population normalization is illustrated in Fig.
 \ref{fig:workflow}.
 Instead of direct FA-to-FA registration, our ANTs-based modification can be described as follows:
-
-\begin{enumerate}
-  \item Build the anatomical template from  the population sample or sample subset using SyGN as described in
-  the previous section.
-  \item If only using a sample subset for template creation, register the remaining
-        anatomical data to the template and save the resulting transformation for each         
-        subject indexed by $i$ (i.e. $\mathcal{T}^i_{template}$).
-  \item Find the optimal transformation from each
-        subject's FA to the anatomical template:
-  \begin{enumerate}
-  \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \citep{Tustison2010} and Atropos $n$-tissue segmentation \citep{Avants2011a}).
-  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
-  \item Perform distortion correction on the FA to anatomical mapping 
-        by localized nonrigid registration
-        of rigidly transformed FA map ($=\mathcal{R}^i_{dwi} \mathrm{FA}$) using
-        the white matter mask obtained in step 3(a) ($\mathcal{T}^i_{FA}$).
-  \item For each subject, concatenate the transforms which maps each
-        subject's FA to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
-        \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
-  \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.  
-  Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any 
-  other scalar derivable from DTI to the template.  
-%  One other detail is that we prefer to warp the subject's DTI to 
-%  the template and then, in template space, derive the scalar images of interest.  
-  \end{enumerate}      
-%  \item Perform the remaining portions of the standard TBSS pipeline:
+%
+%\begin{enumerate}
+%  \item Build the anatomical template from  the population sample or sample subset using SyGN as described in
+%  the previous section.
+%  \item If only using a sample subset for template creation, register the remaining
+%        anatomical data to the template and save the resulting transformation for each         
+%        subject indexed by $i$ (i.e. $\mathcal{T}^i_{template}$).
+%  \item Find the optimal transformation from each
+%        subject's FA to the anatomical template:
 %  \begin{enumerate}
-%      \item Create the mean FA image from all template-aligned FA images.
-%      \item Produce the white matter skeleton from the mean FA image.
-%      \item Create skeleton distance map and project all FA data onto the skeleton.
-%      \item Run the statistical analysis.
+%  \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \citep{Tustison2010} and Atropos $n$-tissue segmentation \citep{Avants2011a}).
+%  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
+%  \item Perform distortion correction on the FA to anatomical mapping 
+%        by localized nonrigid registration
+%        of rigidly transformed FA map ($=\mathcal{R}^i_{dwi} \mathrm{FA}$) using
+%        the white matter mask obtained in step 3(a) ($\mathcal{T}^i_{FA}$).
+%  \item For each subject, concatenate the transforms which maps each
+%        subject's FA to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
+%        \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
+%  \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.  
+%  Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any 
+%  other scalar derivable from DTI to the template.  
+%%  One other detail is that we prefer to warp the subject's DTI to 
+%%  the template and then, in template space, derive the scalar images of interest.  
 %  \end{enumerate}      
-\end{enumerate}
-%The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender effects in a control population.
+%%  \item Perform the remaining portions of the standard TBSS pipeline:
+%%    \begin{enumerate}
+%%      \item Create the mean FA image from all template-aligned FA images.
+%%      \item Produce the white matter skeleton from the mean FA image.
+%%      \item Create skeleton distance map and project all FA data onto the skeleton.
+%%      \item Run the statistical analysis.
+%%    \end{enumerate}
+%\end{enumerate}
+%%The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender effects in a control population.
 
@@
 Due to various issues (e.g.
-lack of corresponding T1 image, failed DTI reconstruction, and age matching requirements) only 62 subjects (21 females and 41 males).
+lack of corresponding T1-weighted image, failed DTI reconstruction, and age matching requirements) only 62 subjects (21 females and 41 males).
 Each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
 EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
-voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
+voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1-weighted anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
 TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
 1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).
-Images were anonymized including defacing of the T1 images prior to upload.
+Images were anonymized including defacing of the T1-weighted images prior to upload.
 Diffusion tensor data associated with each of the following data sets were reconstructed from the diffusion weighted sequences using Camino%
@@
 \subsection{Simulated DTI Gender Cohort Creation}
+\label{sec:sim}
 
@@
 To provide realistic ground truth FA data, we implemented 
-the work of Van Hecke et al. \cite{van-hecke2009} which 
+the method of Van Hecke et al. \cite{van-hecke2009} which 
 we have also made available to the public
@@
 }  
-Given an input DTI atlas, b0 image, scheme file indicating
+Given a DTI atlas, a b0 image, a scheme file indicating
 the gradient directions, and 
 a cohort of aligned DTI data (to model intersubject variability), the
-user is able to create 
-simulated data sets with specified pathology and Rician noise.
+algorithm generates
+simulated data sets with specified pathology and noise.
 For the simulated experiments described in the next section we
-created a T1 template from 30 randomly chosen T1 images.  The
+created a T1-weighted template from 30 randomly chosen T1-weighted images of
+the NKI/Rockland data set described above.  
+The resulting
 template was then registered to the MNI152 template.  
-The remaining 32 T1 images were then registered to the
-template.  Each DTI and b0 image was mapped to the MNI152 map as 
-described above using the T1 to MNI composite mapping.  DTI 
+The remaining 32 T1-weighted images were also registered to the
+template resulting in a composite transform for each subject describing
+the transform from the individual subject's T1-weighted image domain to 
+the MNI152 template.
+
+To obtain the mappings from the DTI space to the population-specific
+template, the following steps were taken:
+\begin{itemize}
+  \item For each anatomical T1-weighted image, segment the white matter (in ANTs, one can use N4 bias correction \citep{Tustison2010} and Atropos $n$-tissue segmentation \citep{Avants2011a}).
+  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
+  \item Perform distortion correction on the FA to anatomical mapping 
+        by localized nonrigid registration
+        of rigidly transformed FA map ($=\mathcal{R}^i_{dwi} \mathrm{FA}$) using
+        the white matter mask obtained in step 3(a) ($\mathcal{T}^i_{FA}$).
+  \item For each subject, concatenate the transforms which maps each
+        subject's FA to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
+        \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
+  \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.  
+%  Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any 
+%  other scalar derivable from DTI to the template.  
+\end{itemize}
+DTI 
 were reoriented using the PPD method \cite{alexander2001}.  The 62
@@
 b0 images were averaged to create the representative b0 image
-for simulated DTI creation.  The derived FA and b0 images are shown in 
+for simulated DTI creation.  The derived FA image from the DTI atlas and 
+the b0 image are shown in 
 Fig. \ref{fig:simulated}.  An FA mask was created by thresholding 
@@
 with no simulated noise nor pathology.  Similarly, the 21 aligned female
-DTI were used to generate 19 ``Experimental'' diffusion-weighted images.  
-DTI were reconstructed using Camino which were then use to create
-individual FA images were
-created from the DTI images to use in the simulated experiments.
+DTI were used to generate 19 ``Experimental'' diffusion-weighted image sets.
+DTI were reconstructed using Camino which were then used to create
+individual FA images.
 The number of subjects in each
-group was chosen based on the median control and experimental
-subject numbers from a sampling of FA studies.
+group was chosen based on the median number of control and experimental
+subjects from the sampling of FA studies described in 
+\cite{van-hecke2009}.
 
@@
 \caption{ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
-different simulated experiments for both (a) voxelwise analysis over the
+different similarity metrics for both (a) voxelwise analysis over the
 entire FA mask and (b) using TBSS projections onto the mean FA skeleton.
@@
 \end{figure*}
+
+\begin{figure*}
+\begin{center}
+\begin{tabular}{ccc}
+  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth0.pdf} &
+  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth1.pdf} &
+  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth2.pdf} \\
+  \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth0.pdf} &
+  \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth1.pdf} &
+  \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth2.pdf} 
+\end{tabular}
+\caption{Post-normalization smoothing analysis of the simulated data for
+fwhm kernel sizes of 4 mm, 8 mm, and 16 mm.  Top row:  
+ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
+different similarity metrics for voxelwise analysis.  Smoothing
+tends to ameliorate the bias effect but the relative strength
+continues to correlate with similarity metric.  Bottom row:
+A relative 
+increase in (false) statistical power induced by SSD and Demons, as 
+described in the text, resulted in more voxels having lower 
+$p$-values.  This will be manifested as a left-shifted CDF 
+curve which is apparent in the SSD and Demons metrics for all selected smoothing kernel sizes.
+}
+\label{fig:simulated_plots_smooth}
+\end{center}        
+\end{figure*}
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{cc}
+  \includegraphics[width=41mm]{smooth0_truth.png} &
+  \includegraphics[width=41mm]{smooth0_demons.png} \\
+  (a) & (b)
+\end{tabular}
+\caption{(a) Mid-axial slice of the DTI atlas-derived FA map where the 
+significant differences ($p \leq 0.05$, smoothing = 4 mm fwhm) between control (male) and 
+experimental (female) groups are labeled in red.  (b) Same mid-axial 
+slice showing an increase in significant regions following ANTs 
+registration employing the Demons metric.
+}
+\label{fig:simulated_significance}
+\end{center}        
+\end{figure}
 
@@
 in the right plot of Fig. \ref{fig:simulated_plots}.  Although projection
-does have slightly ameliorating effects, the associated bias is still 
+does have slightly mitigating effects, the associated bias is still 
 significant across certain metrics.
+
+Common practice is to smooth the images following normalization and
+prior to statistical testing where the kernel size is typically dependent upon
+one's estimate of the registration accuracy.  We tested the application
+of common kernel sizes of 4 mm, 8 mm, and 16 mm full width at half
+maximum (fwhm).  These plots are shown in Fig. \ref{fig:simulated_plots_smooth}
+where we have produced similar ROC curves as in Fig. \ref{fig:simulated_plots}.
+We also included $p$-value cumulative distribution functions (CDFs) \cite{yanovsky2009}.
+Whereas ROC curves incorporate false/true positives and false/true
+negatives, the $p$-value CDFs directly illustrate that certain metrics 
+induce false positives.  False elevation of statistical power is caused
+by a general decrease in $p$-values.  This decrease in $p$-values causes
+a left-shift in the CDF which is observed for both the SSD and Demons metrics.
 
@@
 FA-to-FA template normalization.  
-Whereas MI shows very little effects of bias due to it's statistical
+Whereas MI shows very little effects of bias due to its statistical
 formulation and relative non-locality, the SSD and SSD-like Demons
 metrics are formulated such that they explicitly minimize voxelwise
-group variance and thus induce such bias effects.
-
-It should be noted that all the effects described are absent
-any multiple comparisons corrections or smoothing prior to statistical
-testing.  Although such techniques can minimize 
-or even possibly eliminate these detrimental confounds, our contention
-is (and these results demonstrate) that normalization choices can significantly
-impact the results prior to any such
-post hoc corrections and scientific best practices dictate that, where possible, 
-experimental adjustments eliminating or minimizing potential sources of bias should
-be made prior to experimental execution and analysis.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{cc}
-  \includegraphics[width=41mm]{no_registration_voxelwise_slice90.png} &
-  \includegraphics[width=41mm]{demons_voxelwise_slice90.png} \\
-  (a) & (b)
-\end{tabular}
-\caption{(a) Mid-axial slice of the DTI atlas-derived FA map where the 
-significant differences ($p < 0.05$) between control (male) and 
-experimental (female) groups are labeled in red.  (b) Same mid-axial 
-slice showing an increase in significant regions following ANTs 
-registration employing the Demons metric.
-}
-\label{fig:simulated_significance}
-\end{center}        
-\end{figure}
+group variance and thus cause an elevation in false positives.
 
@@
 (\ref{eq:variance}), namely the SSD and Demons metric which shrinks drastically for the Wilcoxon rank-sum test.
-Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.
+Results for the other metrics including TBSS and the T1-weighted template-based strategy (``antsT1'') remain fairly stable between the two tests.
 TBSS and the 4 ANTs metric
 %results were all obtained using direct registration involving subject FA and the FMRIB58 template.
-The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
-%common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
+The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1-weighted $\rightarrow$ group T1-weighted template $\rightarrow$
+%common T1-weighted template $\rightarrow$ MNI152 (FMRIB58) template.    }
 %\label{fig:tbi_testing}
@@
 metric.  In this section, we use data that was analyzed in \cite{Stone2011}
-and show how bias is potentially manifested in real data.
-
-We constructed a patient template and a control template from the 
-T1-weighted images of the two groups.  These two 
-templates were then used to create a third common template.  
-To compare with TBSS normalization, the common template was
-registered to the MNI152 template.  Thus, the transformation
-which takes the FA image of an individual subject to the FMRIB58
-template is the composition of transforms: 
-subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template.
-We evaluated this strategy relative to direct FA-to-FA
-ANTs registration with the common similarity metrics used in
-the simulated data example described earlier.
-Standard TBSS alignment was also performed.
-
-Cumulative distribution functions (CDF) of the $p$-values (one-tailed, $\alpha = 0.95$)
-over the entire FA mask (FA $\geq 0.2$) for the different normalization
-strategies are shown in Fig. \ref{fig:tbi_testing}.
-A bias which (falsely) decreases $p$-values will demonstrate a left-shifted
+and show how bias is potentially manifested in real FA analysis.  We investigated
+the following FA-to-FA template normalization strategies for each subject where we used the  FMRIB58 template distributed with the FSL toolkit as the FA template:
+\begin{description}
+  \item[antsCC:] Direct normalization using ANTs with the CC similarity metric,
+  \item[antsDemons:] Direct normalization using ANTs with the Demons similarity metric,
+  \item[antsMI:] Direct normalization using ANTs with the MI similarity metric,
+  \item[antsSSD:] Direct normalization using ANTs with the SSD similarity metric,
+  \item[tbss:] Direct normalization using FNIRT (also available in FSL) with the
+        optimal configuration parameters for the FMRIB58 template encapsulated
+        in the config file, \verb#FA_2_FMRIB58_1mm.cnf# also distributed
+        with the FSL toolkit, and
+  \item[antsT1:] Indirect normalization where a subject's FA image is mapped to the individual
+        T1-weighted image.  This transform is composed with the T1-weighted
+        transform to the T1-weighted template and with a third transform describing the
+        T1-template to MNI152 template mapping (which resides in the same space as the FMRIB58
+        template). Additional details specifically related to the first transform
+        are described formally in
+        Section \ref{sec:sim}.  For the population-specific template, we constructed a patient                  
+        template and a control template from the T1-weighted images of the two groups.  
+        These two templates were then used to create a third common template which was 
+        registered to the MNI152 template.  
+\end{description}
+
+For each of these scenarios, the normalized images were smoothed using
+a kernel size of 3.5 mm (fwhm).  
+Voxelwise analysis was then performed
+within the region defined by thresholding the rescaled FMRIB58 template
+(rescaled to $[0,1]$, thresholded at FA $\geq$ 0.2). Application of
+ Student's t-test 
+(one-tailed, $\alpha = 0.95$)
+ at each voxel between the smoothed experimental and control group
+ FA images
+resulted in a set of $p$-values for each experiment.  Since standard
+practice typically applies multiple comparisons correction to this
+mass univariate statistical testing, we used
+false discovery rate (FDR) correction to produce a set of corrected
+$p$-values.
+
+Given the lack 
+of ground truth, we are unable to provide ROC curves for comparison
+of bias significance. However, as with the simulated data, the CDFs of the 
+corrected $p$-values 
+for the different normalization
+strategies facilitates visualization of the correlated increase in statistical
+power with choice of normalization strategy.
+A bias which (falsely) decreases $p$-values will yield a left-shifted
 CDF.
-  The Demons and SSD power decreases while the others
-remain fairly similar.  Interestingly, the ANTs template-based approach
-trends towards better performance relative to the remaining 3 strategies.
-It should be noted that although TBSS uses the SSD metric, the fixed and moving images are smoothed and subsampled even at the highest resolution level which substantially decreases the local effect of the metric.
-
-%This is very much related to earlier findings in that strategies evincing
-%greater alignment, as quantified by the CV,
-%will trend towards smaller variance thus exhibiting a left-shifted
-%CDF.
-
-%Similar to previous findings with the simulated data, the SSD and 
-%Demons metric show an elevated effect using the Student's 
-%$t$-test over the other alignment protocols which all exhibit similar 
-%strength.
+
+%As was evident with the simulated data, the SSD and Demons metrics 
+%significantly increase statistical power which would agree with our
+%earlier discussions about the effects of circularity bias intrinsic
+%with these metrics. MI and CC, which do not explicitly minimize
+%the voxelwise group variance, produce statistical power with the ANTs T1-weighted
+%template-based approach and the results produced by FNIRT.  Although
+%the FNIRT settings used in TBSS also employ SSD, the images are
+%subsampled at each resolution level with the highest resolution being
+%half that of the original image.  Simultaneously, the images are smoothed
+%at each resolution level.  These additional factors would seem to 
+%minimize the influence of bias.
 
@@
 \end{tabular}
-\caption{Student's $t$-test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
+\caption{Student's $t$-test results (one-tailed, $\alpha = 0.95$) for the smoothed (kernel size = )
+TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
 Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn.
 (\ref{eq:variance}), namely the SSD and Demons metric.
-Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.
+Results for the other metrics including TBSS and the T1-weighted template-based strategy (``antsT1'') remain fairly stable between the two tests.
 TBSS and the 4 ANTs metric
 results were all obtained using direct registration involving subject FA and the FMRIB58 template.
-The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template.  
+The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1-weighted $\rightarrow$ group T1-weighted template $\rightarrow$
+common T1-weighted template $\rightarrow$ MNI152 (FMRIB58) template.  
 }
@@
 \end{figure}
+
+%To measure alignment in FA, the CDF of the coefficient of variation (CV) 
+%measured at each voxel is often used \cite{van-hecke2011}.  The
+%curves for the different normalization experiments are plotted 
+%in Fig. \ref{fig:tbi_cov}.
+
+%Finally, qualitative assessment is provided in Fig. \ref{fig:meanFA} 
+%in which the mean FA image derived from the template-based approach
+%and the post-normalization component of the TBSS pipeline (Fig. \ref{fig:meanFA}(c)) 
+%is juxtaposed 
+%with the mean FA image produced by the TBSS pipeline alone (Fig. \ref{fig:meanFA}(a)).
+%The skeleton masks were produced by thresholding the skeleton values at 
+%FA $\geq 0.2$ and it would appear that the improved alignment produces 
+%a more extensive skeletal network for greater localization testing of 
+%significant differences, particularly in the outer white matter regions
+%(compare Fig. \ref{fig:meanFA}(b) and Fig. \ref{fig:meanFA}(d)).
+%Given that TBSS is intrinsically a statistical method for finding local 
+%differences, improved alignment would presumably provide increased regional sensitivity to differences.
 
@@
 those metrics which best minimize the voxelwise variance would
-trend towards greater alignment as quantified by the CV.
-Although direct FA-to-FA registration using ANTs with common
+trend towards greater alignment as quantified by the CV \cite{van-hecke2011}.
+Since minimization of the groupwise variance
+will also minimize the CV, it is expected that the SSD and Demons 
+metrics will produce relatively superior alignments to the
+other strategies 
+where alignment assessment is defined by the CV whereas the MI and CC metrics 
+produce lower CV alignments.  Despite the fact that no direct FA-to-FA 
+registrations were ever performed, the `antsT1' normalization produced better alignments
+than `tbss'.  These can be visually assessed in Fig. \ref{fig:meanFA} where
+we plotted the resulting mean FA images for both sets of normalizations.
+Although direct FA-to-FA registration with common
 similarity metrics gives ``better alignment'' such alignment is 
-produced at a cost of significant Type 1 errors.  Common subsampling and
-smoothing strategies and local neighborhood metrics such as CC and MI reduce the 
-Type 1 error problem (as evidenced in Fig. \ref{fig:tbi_testing}) 
-but also show a decrease in alignment.
-
-Finally, qualitative assessment is provided in Fig. \ref{fig:meanFA} 
-in which the mean FA image derived from the template-based approach
-and the post-normalization component of the TBSS pipeline (Fig. \ref{fig:meanFA}(c)) 
-is juxtaposed 
-with the mean FA image produced by the TBSS pipeline alone (Fig. \ref{fig:meanFA}(a)).
-The skeleton masks were produced by thresholding the skeleton values at 
-FA $\geq 0.2$ and it would appear that the improved alignment produces 
-a more extensive skeletal network for greater localization testing of 
-significant differences, particularly in the outer white matter regions
-(compare Fig. \ref{fig:meanFA}(b) and Fig. \ref{fig:meanFA}(d)).
-%Given that TBSS is intrinsically a statistical method for finding local 
-%differences, improved alignment would presumably provide increased regional sensitivity to differences.
+produced at a cost of significant type 1 errors.
+
+All of these issues point to fundamental considerations concerning the
+clinical interpretability of results in which circularity bias is shown 
+to have a significant impact with commonly chosen registration 
+parameters.  Given the fundamental nature of normalization for
+neuroimaging, it is important that proper experimental set-up be
+given priority over obtaining statistical significance.  Since
+we have shown, both analytically and empirically,  that statistical 
+power can be artificially inflated based on similarity metric choice,
+and that these are some of the most commonly used metrics, we encourage
+investigators to exercise caution in making such normalization choices.
 
@@
 results were all obtained using direct registration involving subject FA and the FMRIB58 template.
-The curve denoted as ``antsT1'' was also obtained in FMRIB58 space with the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template. }
+The curve denoted as ``antsT1'' was also obtained in FMRIB58 space with the following transformation composition: subject FA $\rightarrow$ subject T1-weighted $\rightarrow$ group T1-weighted template $\rightarrow$
+common T1-weighted template $\rightarrow$ MNI152 (FMRIB58) template. }
 \label{fig:tbi_cov}
@@
 The mean FA image illustrated in (a) is created by mapping the sample FA population to the default FMRIB58 template which is the standard protocol for TBSS.
-In comparison, the mean template represented in (c) is created from alignment of the population sample to the optimal T1 template as described in the paper.
+In comparison, the mean template represented in (c) is created from alignment of the population sample to the optimal T1-weighted template as described in the paper.
 The respective skeleton masks in (b) and (d) are created by thresholding the resulting skeleton values $\geq 0.2$.
@@
 Prior evaluation work using anatomically labeled
-data indicates that mutual information, cross-correlation, Demons, and
+data indicates that MI, CC, Demons, and
 SSD metrics are all capable of producing high quality anatomical
@@
 provide a more conservative alternative as does the methodology
-presented here based on T1 to DT intrasubject mapping and T1
+presented here based on T1-weighted to DT intrasubject mapping and T1
 normalization.  To our surprise, we could not find a prior publication that directly

### 2011-10-19 ANTsTBSS/paperRevision.tex
ENH:  MInor edits.

@@
 initially centering around the selection of voxels during the course of
-fMRI studies, we demonstrate how this concept of circularity bias can 
+fMRI studies, we demonstrate how this bias can 
 potentially corrupt voxel-based analysis with a particular emphasis on
@@
 computation of FA, each subject's FA map is registered to the FA template prior to 
-performing statistical comparisons between different groups within a population.
+performing statistical comparisons between different groups.
 We illustrate how this analysis strategy can induce selection
@@
 to the statistical reference space occurs through the 
-corresponding (statistically independent) anatomical images to 
-an anatomical template.
+corresponding (statistically independent) anatomical images.
 
@@
 \begin{keyword}
-Advanced Normalization Tools (ANTs) \sep circularity\ diffusion tensor imaging \sep fractional anisotropy \sep template construction \sep voxel-based analysis
+Advanced Normalization Tools (ANTs) \sep circularity \sep diffusion tensor imaging \sep fractional anisotropy \sep template construction \sep voxel-based analysis
 %% keywords here, in the form: keyword \sep keyword
@@
 % including subjects with traumatic brain injury (TBI) \citep{Caeyenberghs2010,Jiang2010,Warner2010}.  
-Two popular comprehensive software packages for assessing population FA differences include the SPM Matlab-based toolkit%
+Two popular comprehensive software packages for assessing population differences include the SPM Matlab-based toolkit%
 \footnote{
@@
 \end{align}
-Switching the order of the summations (since the transformations are mutually independent) and recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the mean of the aligned images, it becomes apparent that the transformation solution
+Switching the order of the summations (since the transformations are mutually independent) and recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the mean of the aligned images, it becomes apparent that the transformation solution,
 \begin{align}\label{eq:variance}
@@
     \sum_{n=1}^N \underbrace{\sum_{m=1}^M \left( I^n_m( \mathcal{T}_m(\mathbf{x}) ) - J \right)^2}_{
-    \propto\mathrm{\,voxelwise\,variance}}
+    \propto\mathrm{\,voxelwise\,variance}},
 \end{align}
-is that which minimizes the voxelwise group variance.
-Reducing the group variance will also directly affect the value of a standard population statistic, the Student's $t$-test.
+is that which minimizes the voxelwise group variance.  Reducing the group variance will also directly affect the value of a standard population statistic, the Student's $t$-test.%
+\footnote{
 Based on this derivation, one would also expect the popular Demons metric to be susceptible to this erroneous increase in statistical power due to identical objective functions although a more aggressive displacement/gradient at each iteration will relatively exacerbate this effect for a given number of iterations.
 Other standard similarity metrics will also exhibit this
-trend to some degree as they minimize some local measure of intensity-based distance between the subject and the template although the effect will be minimized relative to the similarity to Eqn.
-(\ref{eq:variance}).
+trend to some degree as they minimize a local measure of intensity-based distance between the subject and the template.
+}
 
 Thus, instead of normalization based on anatomical alignment followed by
-statistical testing, direct normalization of the images to be statistically compared 
+statistical testing, direct normalization of the images to be statistically compared using SSD
 explicitly conflates
@@
 results that one will ultimately use to assess hypotheses.  This immediately evokes a sense of circularity in the analysis \citep{Kriegeskorte2010}.
-According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on data that were selected for showing the effect of interest or a related effect.''  As illustrated with
-Eqn.
-\ref{eq:variance}, direct FA-to-FA template registration will produce transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
+According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on data that were selected for showing the effect of interest or a related effect.''  Direct FA-to-FA template registration will produce transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
 $t$-test).
@@
 to assess alignment (i.e. the similarity metric) in FA-to-T1 and T1-to-T1 
-normalizations are statistic gally independent of the intensity value differences used for 
+normalizations are statistically independent of the intensity value differences used for 
 the post-alignment numerical analysis.  It should be noted that this discussion
-is certainly not limited to studies involving FA.  It should be apparent
-that such bias can occur anytime the images used for normalization
+is certainly not limited to studies involving FA.  
+Such bias can occur anytime the images used for normalization
 are non-independent of the statistical analysis as defined by the similarity
@@
 is to be expected.  Specifically, we show that whereas SSD and Demons
-substantially overestimate statistical significance, mutual
-information and cross correlation are less susceptible although the
+substantially overestimate statistical significance, MI and CC are less susceptible although the
 effect is not
@@
   experiments establish that this trend persists even for commonly used smoothing
-  kernels and the projection step of TBSS.
+  kernels and the TBSS projection step.
   \item {\it Real TBI Data}:  FA data of TBI survivors and controls were 
   normalized to a template using four common similarity metrics
-  (CC, Demons, MI, and SSD).  Additional normalization included normalization via
+  (CC, Demons, MI, and SSD).  Additional exploration included normalization via
   the subjects' anatomical T1-weighted images as well as the default registration for TBSS.
@@
 The TBI data used in this study is part of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in long-term survivors of TBI (principal investigator: John Whyte).
-A more detailed data description can be found in \cite{Avants2008}.
-The following is a brief description relevant to our particular study.
 17 controls and 16 patients with TBI were used for the analysis presented in this work.  Each patient had a history of non-penetrating traumatic brain injury 
@@
 DTI images were also obtained. Diffusion tensor data were reconstructed using Camino.
+A more detailed data description can be found in \cite{Avants2008}.
 
@@
 For the simulated experiments described in the next section we
-created a T1-weighted template from 30 randomly chosen T1-weighted images of
-the NKI/Rockland data set described above.  
+created a T1-weighted population-specific template from 30 randomly chosen T1-weighted images of
+the NKI/Rockland data set previously described.  
 The resulting
 template was then registered to the MNI152 template.  
-The remaining 32 T1-weighted images were also registered to the
+The remaining 32 T1-weighted images were also registered to the population-specific 
 template resulting in a composite transform for each subject describing
-the transform from the individual subject's T1-weighted image domain to 
+the mapping from the individual subject's T1-weighted image domain to 
 the MNI152 template.
@@
         of rigidly transformed FA map ($=\mathcal{R}^i_{dwi} \mathrm{FA}$) using
-        the white matter mask obtained in step 3(a) ($\mathcal{T}^i_{FA}$).
+        the white matter mask obtained in a previous step ($\mathcal{T}^i_{FA}$).
   \item For each subject, concatenate the transforms which maps each
@@
 \begin{center}
-\begin{tabular}{cc}
-  \includegraphics[width=86mm]{ISV_SpeSenAlreadyAligned.pdf} &
-  \includegraphics[width=86mm]{ISV_SpeSen_projectionAlreadyAligned.pdf}
-\end{tabular}
-\caption{ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
-different similarity metrics for both (a) voxelwise analysis over the
-entire FA mask and (b) using TBSS projections onto the mean FA skeleton.
-Degree of locality and correspondence to 
-Eqn. (\ref{eq:variance}) resulted in performance disparity between the
-different metrics.  Since SSD and the SSD-like Demon's formulation 
-minimize voxelwise group variance, we see the greatest bias in those 
-two metrics.  In contrast, MI and CC are less susceptible to bias with
-MI performing quite well.  Comparing voxelwise and projection results, 
-it is apparent that the TBSS projection strategy only slightly improves 
-the effects of bias for the simulated data.
-}
-\label{fig:simulated_plots}
-\end{center}        
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
 \begin{tabular}{ccc}
+  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAligned.pdf} &
   \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth0.pdf} &
-  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth1.pdf} &
-  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth2.pdf} \\
+  \includegraphics[width=57mm]{ISV_SpeSenAlreadyAlignedSmooth1.pdf} \\
+  \includegraphics[width=57mm]{ISV_CDFAlreadyAligned.pdf} &
   \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth0.pdf} &
-  \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth1.pdf} &
-  \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth2.pdf} 
+  \includegraphics[width=57mm]{ISV_CDFAlreadyAlignedSmooth1.pdf} \\
 \end{tabular}
 \caption{Post-normalization smoothing analysis of the simulated data for
-fwhm kernel sizes of 4 mm, 8 mm, and 16 mm.  Top row:  
+fwhm kernel sizes of 0 mm, 4 mm, and 8 mm.  Top row:  
 ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
@@
 }
-\label{fig:simulated_significance}
+\label{fig:simulated_projection}
+\end{center}        
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+  \includegraphics[width=81mm]{ISV_SpeSen_projectionAlreadyAligned.pdf}
+\end{tabular}
+\caption{ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
+different similarity metrics using TBSS projections onto the mean FA skeleton.
+Degree of locality and correspondence to 
+Eqn. (\ref{eq:variance}) resulted in performance disparity between the
+different metrics.  Since SSD and the SSD-like Demon's formulation 
+minimize voxelwise group variance, we see the greatest bias in those 
+two metrics.  In contrast, MI and CC are less susceptible to bias with
+MI performing quite well.  Comparing voxelwise (cf Fig. \ref{fig:simulated_plots_smooth}) 
+and projection results, 
+it is apparent that the TBSS projection strategy only slightly improves 
+the effects of bias for the simulated data.
+}
+\label{fig:simulated_plots}
 \end{center}        
@@
 
-Voxelwise analysis included performing Student's t-test within the
+Voxelwise analysis included performing Student's $t$-test within the
 DTI masked region (FA $\geq 0.2$) for both the ground truth (i.e.
@@
 voxels in the already aligned data set.  We then varied the 
-significance level cutoff for each of the registered data sets to plot
-the corresponding ROC curves given in Fig. \ref{fig:simulated_plots}.
+significance level cutoff for each of the registered data sets.
 Visual comparison of the significant regions associated with the 
@@
 the potential severity metric choice can have on results.
+
+Standard practice is to smooth the images following normalization and
+prior to statistical testing where the kernel size is typically dependent upon
+one's estimate of the registration accuracy.  We tested the application
+of common kernel sizes of 4 mm and 8 mm full width at half
+maximum (fwhm) and juxtaposed those results with no smoothing in
+Fig. \ref{fig:simulated_plots_smooth}.  ROC curves are given
+for each of the three cases.
+We also included $p$-value cumulative distribution functions (CDFs) \cite{yanovsky2009}.
+Whereas ROC curves incorporate false/true positives and false/true
+negatives, the $p$-value CDFs directly illustrate that certain metrics 
+induce false positives.  False elevation of statistical power is caused
+by a general decrease in $p$-values.  This decrease in $p$-values causes
+a left-shift in the CDF which is observed for both the SSD and Demons metrics.
 
@@
 significant across certain metrics.
-
-Common practice is to smooth the images following normalization and
-prior to statistical testing where the kernel size is typically dependent upon
-one's estimate of the registration accuracy.  We tested the application
-of common kernel sizes of 4 mm, 8 mm, and 16 mm full width at half
-maximum (fwhm).  These plots are shown in Fig. \ref{fig:simulated_plots_smooth}
-where we have produced similar ROC curves as in Fig. \ref{fig:simulated_plots}.
-We also included $p$-value cumulative distribution functions (CDFs) \cite{yanovsky2009}.
-Whereas ROC curves incorporate false/true positives and false/true
-negatives, the $p$-value CDFs directly illustrate that certain metrics 
-induce false positives.  False elevation of statistical power is caused
-by a general decrease in $p$-values.  This decrease in $p$-values causes
-a left-shift in the CDF which is observed for both the SSD and Demons metrics.
 
@@
 (rescaled to $[0,1]$, thresholded at FA $\geq$ 0.2). Application of
- Student's t-test 
+ Student's $t$-test 
 (one-tailed, $\alpha = 0.95$)
@@
 \end{tabular}
-\caption{Student's $t$-test results (one-tailed, $\alpha = 0.95$) for the smoothed (kernel size = )
+\caption{Student's $t$-test results (one-tailed, $\alpha = 0.95$) for the smoothed (kernel size = 3.5 mm fwhm)
 TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
@@
 metrics will produce relatively superior alignments to the
-other strategies 
-where alignment assessment is defined by the CV whereas the MI and CC metrics 
-produce lower CV alignments.  Despite the fact that no direct FA-to-FA 
-registrations were ever performed, the `antsT1' normalization produced better alignments
+other strategies.  Despite the fact that no direct FA-to-FA 
+registrations were performed, the `antsT1' normalization produced better alignments
 than `tbss'.  These can be visually assessed in Fig. \ref{fig:meanFA} where

### 2011-12-27 WBIR2012/paper.tex
Initial commit.

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithmic,eqparbox,array}
\usepackage{algorithm}

%\floatname{algorithm}{Procedure}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\renewcommand{\algorithmiccomment}[1]{\hfill\eqparbox{COMMENT}{// #1}}
%
\begin{document}
%
\frontmatter          % for the preliminaries
%

\mainmatter              % start of the contributions
%
\title{Diffeomorphic Directly Manipulated Free-Form Deformation Image Registration via Vector Field Flows}
%
\titlerunning{Diffeomorphic DMFFD Image Registration}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nicholas J. Tustison\inst{1} \and Brian B. Avants\inst{2}}
%
\authorrunning{N. Tustison et al.} % abbreviated author list (for running head)
%

\institute{University of Virginia, Charlottesville VA 22903, USA,\\
\email{ntustison@virginia.edu}
\and
University of Pennsylvania,
Philadelphia PA  18104,USA,\\
\email{avants@picsl.upenn.edu}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Motivated by previous work \cite{tustison2009} and 
recent diffeomorphic image registration developments in which the 
characteristic velocity field is represented by spatiotemporal B-splines 
\cite{de-craene2011}, we present a diffeomorphic B-spline-based
image registration algorithm combining and extending these techniques.  
The advancements of the proposed framework
over previous work include analytical integration of
the continuous B-spline velocity field, potential spatial weighting 
of the gradient permitting (among other things) enforcement of 
stationary boundary conditions, multiresolution in terms of the B-spline
mesh, facilitating the incorporation of landmark information, 
and the possibility of modeling temporal periodicity.  In addition to
theoretical and practical discussions of our contribution, we also describe 
its parallelized implementation as open source in the Insight Toolkit.  
\keywords{B-splines, DMFFD, diffeomorphism, ITK}
\end{abstract}

\section{Introduction}
Significant algorithmic developments characterizing modern intensity-based
image registration research include the B-spline parameterized approach
(so called {\em free-form deformation}) with early contributions 
including \cite{szeliski1997,thevenaz1998,rueckert1999}.  Amongst the 
variant extensions, the 
{\em directly manipulated free-form deformation} approach  
 addressed the hemstitching issue associated with the
problematic energy topographies caused by the distribution of the uniform 
B-spline shape functions over the transformation domain \cite{tustison2009}.

Other important image registration research
reflected increased emphasis on topological transformation considerations
in modeling biological/physical systems where topology is 
consistent throughout the course of deformation or a 
homeomorphic relationship is assumed between image domains.
Methods such as LDDMM \cite{beg2005} optimize time-varying velocity field 
flows to yield diffeomorphic transformations.  Alternatively, the FFD 
variant reported in \cite{rueckert2006} enforced diffeomorphic transforms
by concatenating multiple FFD transforms, each of which is constrained
to describe a one-to-one mapping.
Recently, the work of 
\cite{de-craene2011} combined these registration concepts into a single
framework called {\em temporal free-form deformation} in which the 
time-varying velocity field characteristic of LDDMM-style algorithms
is modeled using a 4-D B-spline object (3-D + time).  Integration of 
the velocity field yields the mapping between parameterized time points.

In this work, we describe our extension to these methods.  Similar to 
\cite{de-craene2011}, we also use an $N$-D + time B-spline object to 
represent the characteristic velocity field.  However, we use the 
directly manipulated free-form deformation optimization formulation to improve 
convergence during the course of optimization.  This also facilitates
modeling temporal periodicity and
the enforcement of stationary boundaries consistent with diffeomorphic
transforms.
We also incorporate B-spline mesh multi-resolution capabilities
for increased control during registration progress.  Of note, whereas the 
previous work described in \cite{de-craene2011} discretized the velocity field before 
performing numerical integration (which reduces the advantages of the continuous
B-spline representation including the decrease in accuracy associated with
numerical integration and increased computational time) we show how the continuous B-spline 
representation permits analytical integration for increased accuracy.  
Most importantly, we also describe
our parallelized algorithmic implementation as open source available through the Insight Toolkit.%
\footnote{
http://www.itk.org/
}

We first describe the methodology by laying out a mathematical description 
of the various algorithmic elements coupled with implementation details
where appropriate.  This is followed by several registration examples
to provide insight into the details of our contribution.

\section{Methods: Formulae  and Implementation}

In this section, we explain the underlying theory focusing on
differences with previous work.  We first explain how B-spline
velocity fields can be used to produce diffeomorphisms through
analytical integration involving the B-spline basis functions. 
We then show how our previous work involving optimization in 
B-spline vector spaces \cite{tustison2009} can be used for
optimization of diffeomorphisms.  Additional insight is then
gleaned by illustrating correspondence between theory and 
implementation.

\subsection{B-spline velocity field transform}

Briefly, as with other diffeomorphic formulations based on vector flows, we
assume the diffeomorphism, $\phi$, is defined on the image domain, $\Omega$, 
with stationary boundaries such that $\phi( \partial \Omega) = \mathbf{Id}$.
$\phi$ is generated as the solution of the ordinary differential
equation 
\begin{align}
  \label{eq:ode}
  \frac{d\phi(\mathbf{x}, t)}{dt} = v(\phi(\mathbf{x}, t), t),\,\,\phi( \mathbf{x}, 0 ) = \mathbf{x}
\end{align}
where $v$ is a (potentially) time-dependent smooth field, $v : \Omega \times t 
\rightarrow \mathrm{R}^d$ parameterized by $t \in [0,1]$.  Diffeomorphic mappings
between parameterized time points $\{t_a,t_b\} \in [0,1]$ 
are obtained from  Eq. (\ref{eq:ode}) through the following integration
\begin{align}
  \label{eq:integral}
\phi(\mathbf{x},t_b) &= \phi(\mathbf{x},t_a) + \int_{t_a}^{t_b} v(\phi(\mathbf{x}), t) dt.
\end{align}

In the case of 
$d$-dimensional registration, we can represent the time-dependent velocity field  
as a $(d + 1)$-dimensional B-spline object
\begin{align}
v(\phi(\mathbf{x},t), t) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} B_{i_t}(t) \prod_{j=1}^d B_{i_j}(x_j)
\end{align}
where $v_{i_1,\ldots,i_d,i_t}$ is a $(d+1)$-dimensional control point lattice
characterizing the velocity field and $B(\cdot)$ are the univariate B-spline
basis functions separately modulating the solution in each parametric dimension.
Combining Eqns. (\ref{eq:ode}) and (\ref{eq:integral}) we obtain
\begin{align}
\phi(\mathbf{x},t_b) = \phi(\mathbf{x},t_a) + \int_{t_a}^{t_b} \left( \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T 
    v_{i_1,\ldots,i_d,i_t} B_{i_t}(t)\prod_{j=1}^d B_{j}(x_j) \right) dt 
\end{align}
Moving the integral inside the summations
\begin{align}
\label{eq:velocityfield}
\phi(\mathbf{x},t_b) = \phi(\mathbf{x},t_a) + \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} 
        \underbrace{\left( \sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} \int_{t_a}^{t_b} B_{i_t}(t) dt \right)}_{\phi'_{i_1,\ldots,i_d}} \prod_{j=1}^d B_{i_j}(x_j) 
\end{align}
demonstrates the sole dependency on the B-spline basis functions in
the temporal parametric dimension.  The integral is easily calculated
as the B-spline basis functions are piecewise polynomials.
Also, it is seen that the 
quantity inside the parentheses comprises the $d$-dimensional 
control point lattice defining the diffeomorphic mapping
defined between time points $[t_a,t_b]$ obtained by 
integration.  Furthermore, supposing the initial diffeomorphism,
$\phi(\mathbf{x}, t_a)$, is similarly parameterized, i.e.
\begin{align}
  \phi(\mathbf{x}, t_a) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} \phi^{t_a}_{i_1,\ldots,i_d}\prod_{j=1}^d B_{i_j}(x_j), 
\end{align}
the diffeomorphism at $t_b$ is obtained by simple addition of the two control point 
lattices,
\begin{align}
\phi(\mathbf{x},t_b) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} \left(\phi^{t_a}+
\phi' \right)_{i_1,\ldots,i_d} \prod_{j=1}^d B_{i_j}(x_j).
\end{align}

\subsection{Directly manipulated free-form deformation optimization of the B-spline velocity field}

In \cite{tustison2009} 
it was observed that optimization of FFD registration with gradient descent is 
intrinsically susceptible to problematic energy topographies.  However, a
well-understood preconditioned gradient was proposed based on the work 
described in \cite{tustison2006} which substantially 
improves performance which we refer to as DMFFD image registration.  
Similarly, we propose the following velocity field control
point lattice preconditioned gradient,
$\delta v_{i_1,\ldots,i_d,i_t}$, given the similarity metric, $\Pi$,%
\footnote
{
Current options include neighborhood cross correlation (CC), mutual information (MI), and
Demons-style sum of squared differences (Demons).
}
\begin{align}
\label{eq:dmffd}
  \delta v_{i_1,\ldots,i_d,i_t} &= \left( \sum_{c=1}^{N_{\Omega} \times N_t} \left( \frac{\partial \Pi}{\partial \mathbf{x}} \right)_c B_{i_t}(t^c)\prod_{j=1}^d B_{i_j}(x_j^c)  \right. \nonumber \\
  &\cdot \left. \frac{B_{i_t}^2(t^c) \prod_{j=1}^d B_{i_j}^2 (x_j^c)} 
  {\sum_{k_1=1}^{r+1}\ldots\sum_{k_d=1}^{r+1} \sum_{k_t=1}^{r+1} B_{k_t}^2(t^c) 
  \prod_{j=1}^d B_{k_j}^2 (x_j^c)} \right) \nonumber \\
  &\cdot\left({\sum_{c=1}^{N_{\Omega}\times N_t}B_{i_t}^2(t^c) \prod_{j=1}^d B_{i_j}^2 (x_j^c)} \right) ^{-1}
\end{align}
which is a slight modification of Eqn. (21)
in \cite{tustison2009} which takes into account the temporal locations of the 
dense gradient field sampled in $t \in [0,1]$. $N_t$ and $N_\Omega$ are the number
of time point samples and the number of voxels in the reference image domain, respectively.
$r$ is the spline order in all dimensions%
\footnote{
Spline orders can be specified separately for each dimension but, for simplicity,
we only specify a single spline order.  However, the framework can easily 
accommodate different spline orders.
}
and $c$ indexes the spatio-temporal dense metric
gradient sample.
Combining DMFFD optimization with the B-spline velocity field transform discussed in
the previous subsection, we outline the iterative routine for optimizing the 
velocity field control in Algorithm \ref{alg1}.

\begin{algorithm}                    
\caption{Gradient descent optimization of the B-spline velocity field}        
\label{alg1}      
\begin{algorithmic}[1]
\REQUIRE images $\mathcal{I}$ and $\mathcal{J}$
\REQUIRE  user-specified similarity metric $\Pi$ 
\REQUIRE  velocity field mesh size and multiresolution schedule, $R$
\REQUIRE  number of time point samples, $N_t$ (default: 4)
\REQUIRE  spline order (default: 3)
\REQUIRE  gradient step size, $\lambda$ (default: 0.25)
\ENSURE velocity field control point lattice, $v$
\STATE $v \gets \mathbf{0}$ \COMMENT{Initialize velocity field}
\FOR{number of resolution levels} 
\FOR{number of iterations for current level} 
  \STATE $G \gets []$  \COMMENT{Initialize dense gradient storage array}
\FOR{$t = 1 \to N_t$ } 
  \STATE $t' \gets \frac{t-1}{N_t-1}$  \COMMENT{Scale time point to $[0,1]$}
  \STATE $\mathcal{I}' \gets \mathcal{I} \circ \phi(\mathbf{x}, t')$  \COMMENT{Warp $\mathcal{I}$ to the current time point}
  \STATE $\mathcal{J}' \gets \mathcal{J} \circ \phi^{-1}(\mathbf{x}, 1-t')$ \COMMENT{Warp $\mathcal{J}$ to the current time point}
  \STATE $G[t] \gets \frac{\partial \Pi(\mathcal{I}',\mathcal{J}')}{\partial \mathbf{x}}$  \COMMENT{Store dense similarity metric gradient}
\ENDFOR  
  \STATE $\delta v \gets B(G)$ \COMMENT{Calculate velocity lattice gradient (Eqn. (\ref{eq:dmffd}))}
  \STATE $v \gets v + \lambda\delta v$ \COMMENT{Take a step in the gradient direction}
\ENDFOR  
  \STATE $v \gets R(v)$         \COMMENT{Refine lattice based on schedule}
%  \IF {converged}
%    \RETURN $v$
%  \ENDIF
\ENDFOR  
\end{algorithmic}
\end{algorithm}

\subsection{Implementation}

As mentioned previously, the registration algorithm has been implemented in the
Insight Toolkit and consists of the following major classes: 
\begin{itemize}
  \item \verb#itk::TimeVaryingBSplineVelocityFieldIntegrationImageFilter#
  \item \verb#itk::TimeVaryingBSplineVelocityFieldTransform#
  \item \verb#itk::TimeVaryingBSplineVelocityFieldImageRegistrationMethod#
  \item \verb#itk::TimeVaryingBSplineVelocityFieldTransformParametersAdaptor#
  \item \verb#itk::BSplineScatteredDataPointSetToImageFilter#
\end{itemize}
The \verb#TimeVaryingBSplineVelocityFieldIntegrationImageFilter# class implements
the integration described by Eqn. (\ref{eq:velocityfield}).  Given a 
velocity field control point lattice as 
input and the lower and upper integration limits, integration is performed 
in a multi-threaded fashion (since each control point in $\phi'_{i_1,\ldots,i_d}$
can be integrated separately).  The \verb#TimeVaryingBSplineVelocityFieldTransform# 
is derived from the base transform class which handles the mapping of geometric primitives
for warping images.  The DMFFD gradient calculation  is handled by the class \verb#BSplineScatteredDataPointSetToImageFilter# (cf. Eqn. \ref{eq:dmffd} and line 11
of Algorithm \ref{alg1}) which takes as input the dense similarity metric and an optional
weighting for each gradient sample.  This permits enforcement of stationary physical
boundaries by specifying a zero gradient on the boundaries and a large weighting.
Coordinating all the elements of image registration is the
\verb#TimeVaryingBSplineVelocityFieldImageRegistrationMethod# class which
encapsulates Algorithm \ref{alg1}.  The resolution scheduling is determined
by the \verb#TimeVaryingBSplineVelocityFieldTransformParametersAdaptor# class.
We provide access to the new ITK 
registration framework (including the B-spline velocity field transform) 
through the command line module \verb#hormigita# available both in ANTs%
\footnote{
http://www.picsl.upenn.edu/ANTs
}
and accompanied by a technical report offered through the Insight Journal
\cite{tustison2012}.%
\footnote{
http://www.insight-journal.org
}
The interested reader should consult \cite{tustison2012} for further details on 
actual usage.

\section{Registration Examples}

\section{Discussion and Conclusions}
This work constitutes an advantageous combination of the continuous aspects of
B-splines with the diffeomorphic registration framework via vector field 
flows.  In contrast to previous work, we take advantage of the continuous 
aspect of the velocity field representation by analytical integration for
a more accurate diffeomorphic transform derivation.  We also incorporate
DMFFD optimization of the B-spline velocity field which facilitates 
convergence and permits enforcement of stationary boundary conditions.  
While not discussed, further advantages include incorporation of temporal 
periodicity in dealing with the possibility of multiple images describing
periodic motion (e.g. cardiac or pulmonary motion).  Although not discussed
in this work, a future publication will explore these possibilities in 
greater detail.

\bibliographystyle{splncs03}
\bibliography{references}

\end{document}

### 2012-01-05 ANTsTBSS/paperRevision.tex
New Safety Suit album---good stuff.

@@
 
+  \usepackage{lineno}
+  \usepackage{makecell}
+\usepackage{amssymb}
 \usepackage{color}
@@
 \usepackage{amsmath}
-\usepackage{lineno}
 \usepackage{ulem}
@@
 %% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
 %% The amsthm package provides extended theorem environments
 % \usepackage{amsthm}
-
- \usepackage{makecell}
 
@@
 in which a representative template serves as the normalized space for 
-group alignment.  Following 
-computation of FA, each subject's FA map is registered to the FA template prior to 
-performing statistical comparisons between different groups.
+group alignment.  A standard approach maps each subject's FA image
+to a FA template before performing statistical comparisons between different groups.
 We illustrate how this analysis strategy can induce selection
@@
 to the statistical reference space occurs through the 
-corresponding (statistically independent) anatomical images.
+corresponding (statistically independent) anatomical images.  A second
+alternative is to use similarity terms that are less sensitive to this bias.
 
@@
 puzzlement at the unusually high number of papers reporting 
-significant correlations in fMRI studies \cite{Vul2009}. During the 
+significant correlations in fMRI studies \citep{Vul2009}. During the 
 resulting investigation, the authors discovered the permeating
@@
 While related literature has focused on the presence of such bias in 
-fMRI studies \cite{Vul2009,Vul2010} and, more generally, in neuroscience research 
-\cite{Kriegeskorte2009}, we show how this can potentially extend to what has
+fMRI studies \citep{Vul2009,Vul2010} and, more generally, in neuroscience research 
+\citep{Kriegeskorte2009}, we show how this can potentially extend to what has
 become a fundamental neuroimaging analysis technique---voxel-based analysis.
@@
 Similarly, we were perplexed by the disparity in statistical results of one of 
-our recent analyses \cite{Stone2011} using the same DTI data and same 
+our recent analyses \citep{Stone2011} using the same DTI data and same 
 voxel-based analysis technique but different normalization approaches.  Despite 
@@
 } 
-VBM \citep{Ashburner2001} continues to find application to FA studies \citep{Kakeda2010,Takao2010,Preziosa2011} despite concerns about its general validity \citep{Bookstein2001,Davatzikos2004} and with respect to DTI \citep{Jones2005,Chung2008}.
+VBM \citep{Ashburner2001} continues to find application to FA studies
+\citep{Kakeda2010,Takao2010,Preziosa2011} despite concerns about its
+general validity \citep{Bookstein2001,Davatzikos2004} in particular with respect to DTI \citep{Jones2005,Chung2008}.
 The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.
@@
 commonality includes direct FA-to-FA registration using various similarity metrics.
-The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image intensity difference to zero), is computationally efficient (just compute the image difference and gradient), and is therefore widely used.
+The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image intensity difference to zero), is computationally efficient (compute the image difference and gradient), and is therefore widely used.
 Some of the most popular registration methods (e.g.
 SPM,
-Demons \cite{Thirion1998}, FSL's FNIRT, AIR \citep{Woods1998}, and LDDMM \citep{Beg2005}) rely on this metric and yield reasonable performance levels \citep{Klein2009}.%
+Demons \cite{Thirion1998}, FSL's FNIRT and LDDMM \citep{Beg2005}) rely on this metric and yield reasonable performance levels \citep{Klein2009}.%
 %\footnote{
@@
 \end{align}
-Switching the order of the summations (since the transformations are mutually independent) and recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the mean of the aligned images, it becomes apparent that the transformation solution,
+Switching the order of the summations and recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the mean of the aligned images, it becomes apparent that the transformation solution,
 \begin{align}\label{eq:variance}
@@
 
-Although various processing choices mitigate the effects of such bias,
-such as Gaussian smoothing during and following normalization, 
+Various processing choices may mitigate the effects of such bias.  Gaussian smoothing during and following normalization, 
 nonparametric testing \citep{Rorden2007}, 
 statistical analysis based on orthogonal projections onto the white matter skeleton 
-\cite{Smith2006}, multiple comparisons corrections \cite{Nichols2003}, and the use of more robust similarity metrics, 
-such choices are often made ad hoc and/or post hoc and, to our 
+\citep{Smith2006}, multiple comparisons corrections \citep{Nichols2003},
+and the use of more robust similarity metrics may all impact the
+outcome.  However, such choices are often made ad hoc and/or post hoc and, to our 
 knowledge, not with the realization of the potential
@@
 are non-independent of the statistical analysis as defined by the similarity
-metric and statistical testing.
+metric and statistical testing. {\color{red}{ For example, following Eqn. (\ref{eq:variance}),
+circularity could theoretically be an issue for VBM if one segmented 
+the template image to obtain the gray matter density (GMD)
+template image and then registered individual subject GMD
+images to the GMD template image using SSD-related similarity measures 
+(e.g. \cite{Good2001}, FSL-VBM,%
+\footnote{
+{\color{red}{http://www.fmrib.ox.ac.uk/fsl/fslvbm/index.html}}
+}%
+and DARTEL VBM%
+\footnote{
+{\color{red}{http://www.fil.ion.ucl.ac.uk/~john/misc/VBMclass10.pdf}}
+}%
+).}}
 
@@
 effect is not
-negligible.  Furthermore, we also show that this effects persists
+negligible.  Furthermore, we also show that this effect persists
 even with the common practice of post-normalization smoothing prior
@@
 \end{itemize}
-All major processing steps were performed using the open source Advanced Normalization Tools 
+All major processing steps were performed using publicly available software including the open source Advanced Normalization Tools 
 (ANTs)%
@@
 }
-software repository which should facilitate reproducibility for the interested 
+which should facilitate reproducibility for the interested 
 reader.
@@
 were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
-Due to various issues (e.g.
-lack of corresponding T1-weighted image, failed DTI reconstruction, and age matching requirements) only 62 subjects (21 females and 41 males).
-Each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
+Due to various issues (e.g. lack of corresponding T1-weighted image,
+failed DTI reconstruction, and age matching requirements) only 62
+subjects (21 females and 41 males) were available.  Each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
 EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
@@
 To provide realistic ground truth FA data, we implemented 
-the method of Van Hecke et al. \cite{van-hecke2009} which 
+the method of Van Hecke et al. \citep{van-hecke2009} which 
 we have also made available to the public
-\cite{Tustison2011}.%
+\citep{Tustison2011}.%
 \footnote{
@@
   \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
-  \item Perform distortion correction on the FA to anatomical mapping 
+  \item Perform distortion correction on the {\color{red}{DTI}} to anatomical mapping 
         by localized nonrigid registration
@@
   \item For each subject, concatenate the transforms which maps each
-        subject's FA to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
+        subject's {\color{red}{DTI}} to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
         \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
-  \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.  
+  \item Warp each subject's {\color{red}{DTI}} map to the template using $\mathcal{T}^i_{total}$.  
 %  Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any 
@@
 DTI were used to generate 19 ``Experimental'' diffusion-weighted image sets.
-DTI were reconstructed using Camino which were then used to create
+These simulated DTI were reconstructed using Camino which yielded
 individual FA images.
@@
 
+\subsection{Statistical methods} 
+We employ two complementary tools for examining the impact of image 
+registration strategy on statistical outcome.  The receiver operating
+characteristic (ROC) and the cumulative distribution function (CDF).
+We provide a brief summary of these tools and their purpose in this study.
+\subsubsection{ROC analysis} 
+The ROC may be used to look at the performance of a continuous model
+against a known classification when the threshold on the model is
+varied.  In this study, the ground-truth is determined by the voxels
+that---via simulation---we know have a significant difference between
+groups.  We choose $p<0.05$ to determine the
+ground truth.  This leads to two classes of voxels: those that are
+significantly different between groups and those that are not.  As we
+vary the threshold on the statistical map that is generated by a
+registration algorithm, we can compare the identified voxels against
+this ground truth and draw the ROC \citep{Fitzpatrick1998,Wenzel2010}.
+
+\subsubsection{Cumulative distribution functions} 
+Empirically constructed CDFs have been used with maps of $p$-values,
+derived from registration, to visually quantify detection power
+\citep{Brun2008,Hua2009}.  In this case, one examines the area under the
+CDF that is to the left of some statistical cutoff, usually $p<0.05$.  If one assumes
+that the method is not producing biased results, then the method that
+has a greater area under the CDF, left of cutoff, has greater
+detection power.
+
 \section{Results}
@@
 
+%\begin{landscape}
 \begin{figure*}
@@
 \caption{Post-normalization smoothing analysis of the simulated data for
-fwhm kernel sizes of 0 mm, 4 mm, and 8 mm.  Top row:  
+fwhm kernel sizes of 0 mm, 4 mm, and 8 mm.  The ground truth data is
+smoothed in the same way as the normalized data.  Top row:  
 ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
@@
 $p$-values.  This will be manifested as a left-shifted CDF 
-curve which is apparent in the SSD and Demons metrics for all selected smoothing kernel sizes.
+curve which is apparent in the SSD and Demons metrics for all selected
+smoothing kernel sizes.  In summary, improvements in detection power
+(if judged by the area under the CDF that is below 0.05) do not imply results that are closer to the
+truth (which is shown by examining total area under the ROC curve).
 }
@@
 \end{figure*}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{cc}
-  \includegraphics[width=41mm]{smooth0_truth.png} &
-  \includegraphics[width=41mm]{smooth0_demons.png} \\
-  (a) & (b)
-\end{tabular}
-\caption{(a) Mid-axial slice of the DTI atlas-derived FA map where the 
-significant differences ($p \leq 0.05$, smoothing = 4 mm fwhm) between control (male) and 
-experimental (female) groups are labeled in red.  (b) Same mid-axial 
-slice showing an increase in significant regions following ANTs 
-registration employing the Demons metric.
-}
-\label{fig:simulated_projection}
-\end{center}        
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=81mm]{ISV_SpeSen_projectionAlreadyAligned.pdf}
-\end{tabular}
-\caption{ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
-different similarity metrics using TBSS projections onto the mean FA skeleton.
-Degree of locality and correspondence to 
-Eqn. (\ref{eq:variance}) resulted in performance disparity between the
-different metrics.  Since SSD and the SSD-like Demon's formulation 
-minimize voxelwise group variance, we see the greatest bias in those 
-two metrics.  In contrast, MI and CC are less susceptible to bias with
-MI performing quite well.  Comparing voxelwise (cf Fig. \ref{fig:simulated_plots_smooth}) 
-and projection results, 
-it is apparent that the TBSS projection strategy only slightly improves 
-the effects of bias for the simulated data.
-}
-\label{fig:simulated_plots}
-\end{center}        
-\end{figure}
+%\end{landscape}
 
@@
 Visual comparison of the significant regions associated with the 
-voxelwise analysis for ground truth data and with the Demons registration data for
+voxelwise analysis for ground truth data {\color{red}{and the data using 
+ the different similarity metrics 
+for}}
 $p \leq 0.05$ shown in Fig. \ref{fig:simulated_significance} demonstrates 
@@
 The associated ROC curves are given
-in the right plot of Fig. \ref{fig:simulated_plots}.  Although projection
+in  Fig. \ref{fig:simulated_projection}.  Although projection
 does have slightly mitigating effects, the associated bias is still 
@@
 group variance and thus cause an elevation in false positives.
+
+\begin{figure*}
+\begin{center}
+\begin{tabular}{cc}
+  \includegraphics[width=175mm]{falsepositives_axialslices.pdf}
+\end{tabular}
+\caption{{\color{red}{Top row:  Mid-axial slices of the DTI atlas-derived FA map 
+showing the 
+ground truth significant differences ($p \leq 0.05$, smoothing = 4 mm fwhm,
+uncorrected for multiple comparisons) 
+between control (male) and 
+experimental (female) groups.  Rows 2--4: Same mid-axial 
+slices showing varying changes in significant regions following ANTs 
+registration employing different similarity metrics.  The key point is that there
+is no anatomical misalignment between the groups before the
+registration algorithm is run.  Although detection power ``improves'',
+(particularly with the SSD and Demons metrics)
+this is not due to improved anatomical alignment but rather due to
+local effects related to circularity bias induced by the relationship
+between the test statistic and the similarity metric.}} }
+\label{fig:simulated_significance}
+\end{center}        
+\end{figure*}
+
+%\begin{figure}
+%\begin{center}
+%\begin{tabular}{cc}
+%  \includegraphics[width=41mm]{smooth0_truth.png} &
+%  \includegraphics[width=41mm]{smooth0_demons.png} \\
+%  (a) & (b)
+%\end{tabular}
+%\caption{(a) Mid-axial slice of the DTI atlas-derived FA map where the
+%  ground truth significant differences ($p \leq 0.05$, smoothing = 4 mm fwhm) between control (male) and 
+%experimental (female) groups are labeled in red.  (b) Same mid-axial 
+%slice showing an increase in significant regions following ANTs 
+%registration employing the Demons metric.  The key point is that there
+%is no anatomical misalignment between the groups before the
+%registration algorithm is run.  Although detection power ``improves'',
+%this is not due to improved anatomical alignment but rather due to
+%local effects related to circularity bias induced by the relationship
+%between the test statistic and the similarity metric. 
+%}
+%\label{fig:simulated_significance}
+%\end{center}        
+%\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+  \includegraphics[width=81mm]{ISV_SpeSen_projectionAlreadyAligned.pdf}
+\end{tabular}
+\caption{ROC curves plotting $(1 - specificity)$ vs. $sensitivity$ for the
+different similarity metrics using TBSS projections onto the mean FA skeleton.
+Degree of locality and correspondence to 
+Eqn. (\ref{eq:variance}) resulted in performance disparity between the
+different metrics.  Since SSD and the SSD-like Demon's formulation 
+minimize voxelwise group variance, we see the greatest bias in those 
+two metrics.  In contrast, MI and CC are less susceptible to bias with
+MI performing quite well.  Comparing voxelwise (cf Fig. \ref{fig:simulated_plots_smooth}) 
+and projection results, 
+it is apparent that the TBSS projection strategy only slightly improves 
+the effects of bias for the simulated data.
+}
+\label{fig:simulated_projection}
+\end{center}        
+\end{figure}
 
@@
 those metrics which best minimize the voxelwise variance would
-trend towards greater alignment as quantified by the CV \cite{van-hecke2011}.
+trend towards greater alignment as quantified by the CV \citep{van-hecke2011}.
 Since minimization of the groupwise variance
@@
 \section{Discussion and Conclusions} 
+% Again, cross-correlation was used as objective function at each level of the non-linear alignment.  
 Prior evaluation work using anatomically labeled
@@
 SSD metrics are all capable of producing high quality anatomical
-alignment (e.g. \citep{Klein2009}).  All of this work, though, compares
+alignment \citep[see][]{Klein2009}.  All of this work, though, compares
 the quality of alignment at a relatively coarse scale of major gyri,
@@
 the features that will be used in hypothesis testing.  In particular,
-one should avoid combining the SSD metric for normalizing FA with
-Student's $t$-test for FA differences.  A final point is that
+one should avoid combining the SSD metric for normalization with
+Student's $t$-test for assessing {\color{red}{image-derived}} differences.  A final point is that
 cumulative distribution functions and other power assessment methods
@@
 raised in \cite{Kriegeskorte2010}.
+
+Perhaps the closest work to ours is \cite{Wenzel2010}.  The authors
+used ROC curves to examine the impact of registration strategy on the
+ability to identify effects in PET.  They found that
+their own method, which uses cross-correlation, performed
+better than the SSD-based SPM2 with respect to simulated data.  Our
+findings suggest that the difference in performance is due, in part,
+to the choice of similarity metric.  However, this cannot be verified
+without further investigation into methodological specifics. 
+{\color{red}{Also, the work of \cite{Freire2001} demonstrated
+that rigid body motion correction algorithms for alignment of
+fMRI data show a similar trend in that more robust similarity
+metrics, such as MI, are less susceptible to bias caused by 
+intensity outliers (e.g. activation) than SSD. This could 
+potentially provide additional
+motivation for favoring more robust similarity metrics
+in normalization.
+}}
 
@@
 whole tensor DTI normalization \citep{Zhang2007,Hecke2007}, DTI template-based
-strategies \citep{Mori2009,Hecke2011}, or other related methods \cite{jbabdi2010}.  
+strategies \citep{Mori2009,Hecke2011}, or other related methods \citep{jbabdi2010}.  
 %Although investigation is beyond the scope
@@
 Instead, we encourage the community to
-reexamine FA normalization and analysis methods and to consider the
+reexamine normalization and analysis methods and to consider the
 potential confounds highlighted in this paper and others
-\cite{Ridgway2008}.  Rank tests
+\citep{Ridgway2008}.  Rank tests
 provide a more conservative alternative as does the methodology

### 2012-01-05 Posters/NNS2011/myposter.tex
New Safety Suit album---good stuff.

@@
 ``views'' of some underlying phenomenon. For gray matter density ($\mathbf{G}$) and
-DTI-derived FA  ($\mathbf{F}$) views, SCCA solves the following
+DTI-derived  ($\mathbf{F}$) views, SCCA solves the following
 criterion \cite{Avants2010b}:
@@
   \item skull stripping using Atropos coupled with anatomical priors, and
-  \item prior-based csf, white matter, and gray matter segmentation.  The
-  latter yields the gray matter density image for each subject.
+  \item prior-based cerebrospinal fluid, white and gray matter segmentation.  The
+  latter yields the gray matter density (GMD) image for each subject.
 \end{itemize}
-FA images are mapped to each anatomical image using the transform
+DTI-derived fractional anisotropy (FA) images are mapped to each anatomical image using the transform
 obtained from registration of the average diffusion image to the 
@@
 used to generate the final/standard template.  Thus the mapping from the individual 
-gray matter density (GMD) image to the total template is given by
+GMD image to the total template is given by
 \begin{align*}
@@
 correlations between GMD and FA. & {           } &
-{\includegraphics[height=80mm]{results.pdf}}
+{\includegraphics[height=80mm]{resultsPartial.pdf}}
 \end{tabular}
@@
 Specific regional differences included the medial thalamic nuclei, hypothalamus, amygdala, hippocampus, anterior cingulate cortex, orbitofrontal cortex and fornix.
-Using these SCCA-identified regions, we demonstrate a strong correlation ($p < 0.08$) of the degree of injury in WM and GM within the patient group.
+Using these SCCA-identified regions, we demonstrate a strong correlation ($p < 0.01$) of the degree of injury in WM and GM within the patient group.
 
@@
   \begin{center}
-  {\normalsize \bf Open source software} \\
+  {\small \bf Open source software} \\
 {\vspace{3mm}}
@@
   \vspace{1mm}
-  \scriptsize
-  \begin{itemize}
-  \item Partial funding support from NZ Breacher study
-  \item This project is also funded, in part, under a grant with the Pennsylvania 
-  Department of Health. The Department specifically disclaims responsibility 
+  \smaller
+   Partial funding support was provided by US ARMY Medical Research
+   and Materiel Command grant W81XWH-09-2-0055.
+   This project is also funded, in part, under a grant with the Pennsylvania 
+  Department of Health. The Pennsylvania Department of Health specifically disclaims responsibility 
   for any analyses, interpretations, or conclusions.
-  \end{itemize}
   }

### 2012-01-05 WBIR2012/paper.tex
ENH:  Added reference.
New Safety Suit album---good stuff.

@@
 The advancements of the proposed framework
-over previous work include analytical integration of
-the continuous B-spline velocity field, potential spatial weighting 
+over previous work include a preconditioned gradient descent algorithm, 
+potential spatial weighting 
 of the gradient permitting (among other things) enforcement of 
@@
 by concatenating multiple FFD transforms, each of which is constrained
-to describe a one-to-one mapping.
+to describe a one-to-one mapping.  Another FFD registration
+incorporated the recent log-Euclidean framework for enforcing diffeomorphic
+transformations \cite{Modat2011}.
 Recently, the work of 
@@
 We also incorporate B-spline mesh multi-resolution capabilities
-for increased control during registration progress.  Of note, whereas the 
-previous work described in \cite{de-craene2011} discretized the velocity field before 
-performing numerical integration (which reduces the advantages of the continuous
-B-spline representation including the decrease in accuracy associated with
-numerical integration and increased computational time) we show how the continuous B-spline 
-representation permits analytical integration for increased accuracy.  
+for increased control during registration progress.  
 Most importantly, we also describe
@@
 between parameterized time points $\{t_a,t_b\} \in [0,1]$ 
-are obtained from  Eq. (\ref{eq:ode}) through the following integration
+are obtained from  Eq. (\ref{eq:ode}) through integration of the transport
+equation, viz.
 \begin{align}
@@
 \begin{align}
-v(\phi(\mathbf{x},t), t) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} B_{i_t}(t) \prod_{j=1}^d B_{i_j}(x_j)
+v(\mathbf{x}, t) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} B_{i_t}(t) \prod_{j=1}^d B_{i_j}(x_j)
 \end{align}
@@
 basis functions separately modulating the solution in each parametric dimension.
-Combining Eqns. (\ref{eq:ode}) and (\ref{eq:integral}) we obtain
+
+Although various methods exist for solving Eqns. (\ref{eq:ode}) and (\ref{eq:integral}),
+we use $4^{th}$-order Runge-Kutta, i.e.
 \begin{align}
-\phi(\mathbf{x},t_b) = \phi(\mathbf{x},t_a) + \int_{t_a}^{t_b} \left( \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T 
-    v_{i_1,\ldots,i_d,i_t} B_{i_t}(t)\prod_{j=1}^d B_{j}(x_j) \right) dt 
+  \phi_{n+1} &= \phi_{n} + \frac{1}{6}\left( k_1 + 2k_2 + 2k_3 + k_4 \right) \\
+  t_{n+1} &= t_{n} + \Delta t \\ 
+  \phi_0 &= \phi(t_0) 
 \end{align}
-Moving the integral inside the summations
+where
 \begin{align}
-\label{eq:velocityfield}
-\phi(\mathbf{x},t_b) = \phi(\mathbf{x},t_a) + \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} 
-        \underbrace{\left( \sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} \int_{t_a}^{t_b} B_{i_t}(t) dt \right)}_{\phi'_{i_1,\ldots,i_d}} \prod_{j=1}^d B_{i_j}(x_j) 
+  k_1 &= v\left( \phi_{n}, t_{n} \right)\Delta t \\
+  k_2 &= v\left( \phi_{n} + \frac{k_1}{2}, t_{n} + \frac{\Delta t}{2} \right)\Delta t \\
+  k_3 &= v\left( \phi_{n} + \frac{k_2}{2}, t_{n} + \frac{\Delta t}{2} \right)\Delta t \\
+  k_4 &= v\left( \phi_{n} + k_3, t_{n} + \Delta t \right)\Delta t
 \end{align}  
-demonstrates the sole dependency on the B-spline basis functions in
-the temporal parametric dimension.  The integral is easily calculated
-as the B-spline basis functions are piecewise polynomials.
-Also, it is seen that the 
-quantity inside the parentheses comprises the $d$-dimensional 
-control point lattice defining the diffeomorphic mapping
-defined between time points $[t_a,t_b]$ obtained by 
-integration.  Furthermore, supposing the initial diffeomorphism,
-$\phi(\mathbf{x}, t_a)$, is similarly parameterized, i.e.
-\begin{align}
-  \phi(\mathbf{x}, t_a) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} \phi^{t_a}_{i_1,\ldots,i_d}\prod_{j=1}^d B_{i_j}(x_j), 
-\end{align}
-the diffeomorphism at $t_b$ is obtained by simple addition of the two control point 
-lattices,
-\begin{align}
-\phi(\mathbf{x},t_b) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} \left(\phi^{t_a}+
-\phi' \right)_{i_1,\ldots,i_d} \prod_{j=1}^d B_{i_j}(x_j).
-\end{align}
+which provides a less computationally expensive, more stable alternative than other numerical
+methods.
+
+%Combining Eqns. (\ref{eq:ode}) and (\ref{eq:integral}) we obtain
+%\begin{align}
+%\phi(\mathbf{x},t_b) = \phi(\mathbf{x},t_a) + \int_{t_a}^{t_b} \left( \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T 
+%    v_{i_1,\ldots,i_d,i_t} B_{i_t}(t)\prod_{j=1}^d B_{j}(x_j) \right) dt 
+%\end{align}
+%Moving the integral inside the summations
+%\begin{align}
+%\label{eq:velocityfield}
+%\phi(\mathbf{x},t_b) = \phi(\mathbf{x},t_a) + \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} 
+%        \underbrace{\left( \sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} \int_{t_a}^{t_b} B_{i_t}(t) dt \right)}_{\phi'_{i_1,\ldots,i_d}} \prod_{j=1}^d B_{i_j}(x_j) 
+%\end{align}
+%demonstrates the sole dependency on the B-spline basis functions in
+%the temporal parametric dimension.  The integral is easily calculated
+%as the B-spline basis functions are piecewise polynomials.
+%Also, it is seen that the 
+%quantity inside the parentheses comprises the $d$-dimensional 
+%control point lattice defining the diffeomorphic mapping
+%defined between time points $[t_a,t_b]$ obtained by 
+%integration.  Furthermore, supposing the initial diffeomorphism,
+%$\phi(\mathbf{x}, t_a)$, is similarly parameterized, i.e.
+%\begin{align}
+%  \phi(\mathbf{x}, t_a) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} \phi^{t_a}_{i_1,\ldots,i_d}\prod_{j=1}^d B_{i_j}(x_j), 
+%\end{align}
+%the diffeomorphism at $t_b$ is obtained by simple addition of the two control point 
+%lattices,
+%\begin{align}
+%\phi(\mathbf{x},t_b) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d} \left(\phi^{t_a}+
+%\phi' \right)_{i_1,\ldots,i_d} \prod_{j=1}^d B_{i_j}(x_j).
+%\end{align}
 
@@
 \begin{itemize}
-  \item \verb#itk::TimeVaryingBSplineVelocityFieldIntegrationImageFilter#
+  \item \verb#itk::TimeVaryingVelocityFieldIntegrationImageFilter#
   \item \verb#itk::TimeVaryingBSplineVelocityFieldTransform#
@@
 \end{itemize}
-The \verb#TimeVaryingBSplineVelocityFieldIntegrationImageFilter# class implements
-the integration described by Eqn. (\ref{eq:velocityfield}).  Given a 
+The \verb#TimeVaryingVelocityFieldIntegrationImageFilter# class implements
+the Runge-Kutta integration described earlier.  Given a 
 velocity field control point lattice as 
 input and the lower and upper integration limits, integration is performed 
-in a multi-threaded fashion (since each control point in $\phi'_{i_1,\ldots,i_d}$
+in a multi-threaded fashion (since each point in the domain
 can be integrated separately).  The \verb#TimeVaryingBSplineVelocityFieldTransform# 
@@
 registration framework (including the B-spline velocity field transform) 
-through the command line module \verb#hormigita# available both in ANTs%
+through the command line module \verb#antsRegistration# available both in ANTs%
 \footnote{
@@
 B-splines with the diffeomorphic registration framework via vector field 
-flows.  In contrast to previous work, we take advantage of the continuous 
-aspect of the velocity field representation by analytical integration for
-a more accurate diffeomorphic transform derivation.  We also incorporate
+flows.  We incorporate
 DMFFD optimization of the B-spline velocity field which facilitates 

### 2012-01-28 ANTsTBSS/paperRevision.tex
ENH: More stuff.

@@
 statistical testing, direct normalization of the images to be statistically compared using SSD
-explicitly conflates
+explicitly conflates {\color{red}{attempts at}}
 anatomical alignment with optimizing the very statistical testing

### 2012-01-28 WBIR2012/paper.tex
ENH: More stuff.

@@
 \usepackage{algorithm}
+\usepackage{graphicx}
+\usepackage{listings}
+\usepackage{color}
+
+    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
+    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
+    \definecolor{listnumbers}{gray}{0.65}
+    \definecolor{listlightgray}{gray}{0.955}
+    \definecolor{listwhite}{gray}{1.0}
 
@@
 its parallelized implementation as open source in the Insight Toolkit.  
-\keywords{B-splines, DMFFD, diffeomorphism, ITK}
+\keywords{B-splines, DMFFD, diffeomorphisms, image registration, ITK}
 \end{abstract}
@@
 of the various algorithmic elements coupled with implementation details
-where appropriate.  This is followed by several registration examples
-to provide insight into the details of our contribution.
+where appropriate.  This is followed by evaluation on a public data set.
 
@@
 Spline orders can be specified separately for each dimension but, for simplicity,
-we only specify a single spline order.  However, the framework can easily 
-accommodate different spline orders.
+we only specify a single spline order.  
 }
@@
 \footnote{
-http://www.insight-journal.org
+http://hdl.handle.net/10380/3334
 }
@@
 
-\section{Registration Examples}
+\section{LPBA40 Evaluation}
+
+As a preliminary evaluation for our algorithm we used a portion 
+of the approach of Klein et al. \cite{Klein2009} in which 
+various open brain data sets were used to evaluate different
+segmentation algorithms.  Specifically, we used the first 20 
+subjects of the LPBA40 data
+set \cite{shattuck2008}%
+\footnote{
+http://www.loni.ucla.edu/Atlases/LPBA40
+}
+publicly available from the Laboratory of Neuro Imaging at
+the University of California, Los Angeles.  Preprocessing included
+brain extraction using the provided masks and N4 bias correction
+\cite{tustison2010}.
+
+Batch registration was facilitated by the inclusion of the proposed
+method in our ANTs interface to the new Insight Toolkit registration
+framework called \verb#antsRegistration# which is detailed in 
+\cite{tustison2012}.  An active development version is found in 
+the ANTS Sourceforge repository.%
+\footnote{
+http://sourceforge.net/projects/advants/
+}
+A portion of the batch script used to produce the registrations is 
+found in Listing \ref{listing:script}.  Each image was histogram-
+matched prior to application of an initial affine transform using
+the Demons metric.  The resulting transform was composed with the
+proposed work using a neighborhood cross correlation similarity metric
+\cite{avants2008} of radius = 4.  A multi-resolution strategy was 
+employed where the initial B-spline velocity field mesh size was 
+$24\times24\times12\times1$ which was doubled at each level.  
+Other registration parameters can be gleaned from the Listing \ref{listing:script}.
+Batch processing was performed on the linux cluster at the 
+University of Virginia with each job allocated 4 nodes for
+multi-threading purposes.  Each registration took approximately
+3 hours.
+
+\vspace{2mm}
+\lstset{frame = htb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{listlightgray}},
+        basicstyle = {\ttfamily\scriptsize},
+        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{listcomment}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=bash,
+        floatplacement=!h,
+        caption={Representative script used for the LPBA40 
+        evaluation.},
+        captionpos=b,
+        label=listing:script
+        }
+\begin{lstlisting}
+# Register the n4 corrected moving LPBA subject to the 
+# n4 corrected fixed LPBA subject.
+
+antsRegistration --dimensionality 3 \
+                 --output ${prefix} \
+                 --use-histogram-matching 1 \
+                 --transform Affine[0.5] \                # affine stage
+                 --metric Demons[${fixed_n4},${moving_n4},1,0,Regular,0.01] \ 
+                 --iterations 100x100x100 \
+                 --smoothing-sigmas 4.0x3.0x2.0 \
+                 --shrink-factors 4x3x2 \
+                 --transform tvdmffd[0.75,24x24x12x1,4] \ # tv dmffd stage
+                 --metric CC[${fixed_n4},${moving_n4},1,4] \ 
+                 --iterations 40x50x2 \
+                 --smoothing-sigmas 1.0x0.5x0.0 \
+                 --shrink-factors 3x2x1
+
+# Apply the resulting transforms (affine + tvdmffd) to the 
+# moving labels.
+
+antsApplyTransforms --dimensionality 3 \
+                    --input ${moving_labels} \
+                    --reference-image ${fixed_n4} \
+                    --output ${moving_warped_labels} \
+                    --n NearestNeighbor \
+                    --transform ${prefix}1Warp.nii.gz \
+                    --transform ${prefix}0Affine.mat \
+                    --default-value 0
+\end{lstlisting}
+
+Registration assessment used the average target overlap 
+\cite{tustison2009a} for each of the 56 LPBA40 labeled regions.
+The numerical values of this assessment are given in Table 1.
+These values were projected onto subject 1 for display in Fig. 1
+which can be compared to a similar display given in \cite{Klein2009}
+with the caveat that the colormap is slightly different.
+
+\begin{figure}
+\centering
+\begin{tabular}{c}
+  \includegraphics[width=120mm]{Figures/leftHemisphere.png} \\
+  \includegraphics[width=120mm]{Figures/rightHemisphere.png}
+\end{tabular}
+\caption{Average regional total overlap measures mapped onto Subject 1
+of the LPBA40 data set.}
+\label{fig:lpba40}
+\end{figure}
+
+\setlength{\tabcolsep}{4pt}
+
+\begin{table}[ht]
+\begin{minipage}[b]{0.5\linewidth}\centering
+\begin{tabular}{lc}
+{\bf Region} & {\bf Overlap} \\
+\hline
+L superior frontal gyrus & 0.802 \\
+R superior frontal gyrus & 0.810 \\
+L middle frontal gyrus & 0.790 \\
+R middle frontal gyrus & 0.778 \\
+L inferior frontal gyrus & 0.753 \\
+R inferior frontal gyrus & 0.728 \\
+L precentral gyrus & 0.687 \\
+R precentral gyrus & 0.674 \\
+L middle orbitofrontal gyrus & 0.687 \\
+R middle orbitofrontal gyrus & 0.671 \\
+L lateral orbitofrontal gyrus & 0.669 \\
+R lateral orbitofrontal gyrus & 0.607 \\
+L gyrus rectus & 0.654 \\
+R gyrus rectus & 0.652 \\
+L postcentral gyrus & 0.604 \\
+R postcentral gyrus & 0.620 \\
+L superior parietal gyrus & 0.709 \\
+R superior parietal gyrus & 0.701 \\
+L supramarginal gyrus & 0.619 \\
+R supramarginal gyrus & 0.623 \\
+L angular gyrus & 0.594 \\
+R angular gyrus & 0.620 \\
+L precuneus & 0.668 \\
+R precuneus & 0.659 \\
+L superior occipital gyrus & 0.548 \\
+R superior occipital gyrus & 0.561 \\
+L middle occipital gyrus & 0.671 \\
+R middle occipital gyrus & 0.696 \\
+\hline
+\end{tabular}
+\end{minipage}
+\begin{minipage}[b]{0.5\linewidth}
+\centering
+\begin{tabular}{lc}
+{\bf Region} & {\bf Overlap} \\
+\hline
+L inferior occipital gyrus & 0.621 \\
+R inferior occipital gyrus & 0.637 \\
+L cuneus & 0.622 \\
+R cuneus & 0.624 \\
+L superior temporal gyrus & 0.750 \\
+R superior temporal gyrus & 0.737 \\
+L middle temporal gyrus & 0.660 \\
+R middle temporal gyrus & 0.644 \\
+L inferior temporal gyrus & 0.662 \\
+R inferior temporal gyrus & 0.674 \\
+L parahippocampal gyrus & 0.711 \\
+R parahippocampal gyrus & 0.707 \\
+L lingual gyrus & 0.694 \\
+R lingual gyrus & 0.681 \\
+L fusiform gyrus & 0.703 \\
+R fusiform gyrus & 0.719 \\
+L insular cortex & 0.812 \\
+R insular cortex & 0.768 \\
+L cingulate gyrus & 0.691 \\
+R cingulate gyrus & 0.676 \\
+L caudate & 0.704 \\
+R caudate & 0.717 \\
+L putamen & 0.773 \\
+R putamen & 0.761 \\
+L hippocampus & 0.753 \\
+R hippocampus & 0.742 \\
+cerebellum & 0.935 \\
+brainstem & 0.899 \\
+\hline
+\end{tabular}
+\end{minipage}
+\vspace{1.0mm}
+\caption{Average target overlap values for cross-registration of the first 
+20 subjects of the LPBA40 data set.  A visual depiction of these values is
+given in Figure \ref{fig:lpba40}.
+}
+\end{table}
 

### 2012-05-22 KapowskiChronicles/intro.tex
ENH:  Adding Kapowski chronicles.

\section{Introduction}

Neuroscientific investigations into cortical morphological changes/differences have illuminated
interesting correlations with normal and pathological neurodevelopment in 
addition to cognitive function.
Historically rooted in the meticulous work of von Economo \citep{economo2008},
computational methods for modeling the
cortical gray matter have resulted in numerous studies 
exploring such relationships.  
These analysis techniques
have been instrumental in demonstrating 
the presence of regional cortical abnormalities in such conditions as 
Huntington's disease \citep{rosas2002,rosas2005,selemon2004}, 
schizophrenia \citep{nesvag2008}, bipolar disorder \cite{lyoo2006}, Alzheimer's disease and frontotemporal
dementia \citep{du2007,dickerson2009}, Parkinson's disease \citep{jubault2011},
multiple sclerosis \citep{ramasamy2009}, autism \citep{chung2005,hardan2006},
migraines \citep{dasilva2007}, chronic smoking \citep{kuhn2010}, alcoholism \citep{fortier2011},
cocaine addiction \citep{makris2008}, Tourette syndrome in children \citep{sowell2008},
scoliosis in
female adolescents \citep{wang2012}, obsessive compulsive
disorder \citep{shin2007}, ADHD \citep{almeida-montes2012}, obesity \citep{raji2010}, and heritable \citep{peterson2009}
and elderly \citep{ballmaier2004} depression.  Further study 
topics exploring correlations with cortical thickness include
normal aging \citep{kochunov2011},
gender differences \citep{luders2006a},
anatomical asymmetries 
\citep{luders2006,amunts2007}, intelligence \citep{shaw2006}, athletic
ability \citep{wei2011}, musical ability \citep{bermudez2009,foster2010}, 
white collar criminals \citep{raine2011}, and Tetris-playing
ability in female adolescents \citep{haier2009}.  Additionally,
recent studies have demonstrated possible functional 
connectivity relationships using cortical thickness measures
\cite{worsley2005,lerch2006,he2007}.
Although these study findings
are subject to debate and interpretation \citep{gernsbacher2007}, 
the availability of quantitative
computational methods for extracting such information
has proven invaluable for developing and refining fundamental 
neuroscience hypotheses.

Broadly characterized as surface mesh-based or volumetric \citep{scott2009,clarkson2011}, such 
techniques have been introduced and extensively developed over 
the last $10+$ years.  Representative of the former is the
Freesurfer%
\footnote{
http://surfer.nmr.mgh.harvard.edu/
}
cortical modeling software package \citep{dale1999,fischl1999,fischl2000,fischl2002,fischl2004}
which owes its popularity to public availability, excellent documentation, 
good performance, and  integration with other toolkits, such as the extensive FMRIB software 
library (FSL) \citep{smith2004}.  Similar to other surface
approaches (e.g. \cite{davatzikos1996,magnotta1999,macdonald2000,kim2005}), the pial
and white matter surfaces from individual subject MR data are modeled with polygonal meshes  
which are then used to determine local cortical thickness based on a specified correspondence between 
the surface models.

Image volumetric (or meshless) techniques are varied both in algorithmic terms as well as
the underlying definition of cortical thickness.  An early, foundational technique is the 
method of \cite{jones2000} in which the inner and outer surface geometry is used to determine the
solution to Laplace's equation in which thickness is measured by integrating along the 
tangents of the resulting field lines spanning the boundary surfaces.  Subsequent contributions
improved upon the original formulation.  For example, in \cite{yezzi2003}, an Eulerian PDE approach
was proposed to facilitate the computation of correspondence paths.  Extending the surface-based
work of \cite{macdonald2000}, the hybrid approach of
\cite{kim2005} uses the discrete Laplacian field to deform the white matter surface mesh towards the 
pial surface.    Although the Laplacian-based approach has several advantages
including generally lower computational times and 
non-crossing correspondence paths, direct correlative assessments with histology
are difficult as the quantified distances 
are not necessarily Euclidean.  Other volumetric algorithms have employed coupled
level sets \citep{zeng1999}, model-free intelligent search strategies either normal to 
the gray-white matter interface \citep{scott2009} or using a min-max rule \citep{clement-vachet2011}.
Most relevant to this work is the DiReCT (Diffeomorphic Registration-based 
Cortical Thickness) algorithm proposed by \cite{das2009} in which a 
a registration-based approach proposed where the derived diffeomorphic mapping between the 
white and pial matter surfaces is used to propagate thickness values 
through the cortical gray matter.

Despite the numerous techniques that have been proposed in the literature (of which
only a fraction has been cited), several requisite processing components are 
common to many of them which are prior to quantification of cortical thickness.
These include inhomogeneity correction, skull stripping, and $n$-tissue segmentation 
for differentiating the gray and white matters.  For post-processing statistical analysis 
across populations, construction of population-specific unbiased templates
is also  beneficial.
In addition, some of these steps might include a crucial registration component, e.g. 
propagating template-based tissue priors for improved segmentation.  The requisite
additional components coupled with the general lack of availability of published
algorithms \cite{kovacevic2006} inhibits performing studies by external researchers 
and makes comparative evaluations difficult.  For example, one recent evaluation 
study \citep{clarkson2011} compared
Freesurfer (a surface-based method) with two volumetric methods \citep{jones2000,das2009}.
Whereas the entire Freesurfer processing pipeline has been made publicly available, 
tuned by the original authors in terms of implementation, and described in great detail 
(specifically in terms of suggested parameters); both volumetric methods were 
implemented solely by the authors of the evaluation (not the actual software developers) using 
unspecified parameters making the comparisons less than ideal.

In this work we describe our cortical thickness software pipeline which 
we have made open source as part of our Advanced Normalization Tools (ANTs) 
repository.  This includes all the necessary preprocessing steps consisting
of well-vetted previously published algorithms for bias correction \cite{tustison2010},
brain extraction \cite{avants2010a}, $n$-tissue segmentation \cite{avants2011a},
template construction \cite{avants2010}, and image normalization \cite{avants2011}.
We also describe improvements made to the original DiReCT algorithm \cite{das2009}.
Equally as important, we demonstrate how to coordinate
these pipeline components and provide a set of useful parameters
which are used to analyze the publicly available IXI data set. This is 
encapsulated in a well-documented shell script which is also available in ANTS. 
Furthermore, we provide all the derived image data and other processing scripts on 
a github repository specifically meant for this publication.  This permits
the set of results described in this work to be fully reproducible thus
permitting other researchers to use a complete volumetric pipeline for measuring
cortical thickness.

%This might explain why there have been few evaluation studies.  For example, in
%comparing volumetric and surface-based methods, \cite{clarkson2011} use the
%Freesurfer implementation but rely on their own implementations of published 
%literature which might not be an unbiased evaluation particularly given the 
%complexity of the underlying registration algorithmic work in \cite{das2009}.
%Based solely on the number of citations in the literature,
%the Freesurfer%
%\footnote{
%http://surfer.nmr.mgh.harvard.edu/
%} 
%cortical modeling software package is perhaps the most ubiquitous 
%with several publications detailing various stages of development and
%methodology \citep{dale1999,fischl1999,fischl2000,fischl2002,fischl2004}.  
%Public availability, excellent documentation, good performance, and 
%integration with other toolkits, such as the extensive FMRIB software 
%library (FSL) \cite{smith2004},
%have contributed to its popularity.

%Freesurfer's individual brain processing pipeline begins with segmentation
%and surface modeling of the gray white matter interface.  Gradients

%\begin{table*}
%\caption{Component-wise breakdown of reported cortical thickness methods reported in the literature.}
%\begin{tabular*}{0.95\textwidth}{@{\extracolsep{\fill}} c c c c c p{4.5cm} }
%\hline
%{\bf Algorithm} & {\bf V/S} & {\bf Bias correction} & {\bf Brain extraction} & {\bf Segmentation} & \multicolumn{1}{c}{\bf Additional notes} \\
%\hline
%BRAINSURF \cite{magnotta1999} & S & {} & none & Discriminate analysis \cite{harris1999} & {} \\
%ASP \cite{macdonald2000} & S & {} & N3$^\dagger$ \cite{Sled1998} & Neural-nets \cite{ozkan1993} & {}\\
%Laplace \cite{Jones2000} & V & {} & none & simple thresholding & {} \\
%CLASP \cite{kim2005} & V/S & {} & N3$^\dagger$ \cite{Sled1998} & K-NN \cite{cocosco2003} & { Partial volume classification of GM/CSF \cite{Choi1991} is used to facilitate reconstructing the pial surface. }\\
%\cite{hutton2005} & V & {} & {} & {} & {} \\
%DiReCT \cite{das2009} & V & {} & none & FAST$^\dagger$ \cite{zhang2001} & {} \\
%\cite{scott2009} & V & {} & none & E-M Bayes \cite{pokric2001}  & {} \\
%\cite{acosta2009} & V & {} & \cite{van-leemput1999a} & E-M Bayes. \cite{van-leemput1999} & {}\\
%\cite{lerch2005}$^{BrainVisa}$ & {} & {} & {} & {} & {} \\
%CRUISE\cite{han2004,tosun2006} & S& {}  & TOADS$^\dagger$ \cite{bazin2007} & {} \\
%
%CLADA\cite{nakamura2011} & S& {}  & PABIC$^\dagger$ \cite{styner2000} & \cite{nakamura2009} & {longitudinal analysis}\\
%%SIENA$^\dagger$\cite{smith2002} & {} & {} & FAST$^\dagger$ \cite{zhang2001} & {longitudinal analysis}\\
%\hline
%\end{tabular*}
%\end{table*}
%
%Open source packages:
%\begin{itemize}
%\item (http://www.ncbi.nlm.nih.gov/pubmed/15957597) TINA - be sure to read the reviews which aren't very good.
%\item (http://www.nitrc.org/projects/arctic/ %http://www.na-mic.org/Wiki/index.php/UNC_ARCTIC_Tutorial) 
%ARCTIC (Automatic Regional Cortical ThICkness)
%\item Brain Voyager (Goebel?)
%\item TOADS-CRUISE (Tosun et al.) http://www.nitrc.org/projects/toads-cruise/
%\item GAMBIT
%\item http://www.bic.mni.mcgill.ca/thickness\_population\_simulation/ \cite{lerch2005}
%\item Be sure to email Vincent Magnotta to see if they have open source tools in Brains for estimation of cortical thickness
%\end{itemize}

### 2012-05-22 KapowskiChronicles/methods.tex
ENH:  Adding Kapowski chronicles.

\section{Methods and Materials}

\subsection{ANTs volumetric-based cortical thickness estimation pipeline}

The ANTs-based cortical thickness estimation workflow is illustrated 
in Figure \ref{fig:pipeline}.  The steps are as follows:
\begin{enumerate}
  \item initial N4 bias correction on input anatomical MRI,
  \item brain extraction using a hybrid segmentation/template-based strategy,
  \item alternating between prior-based segmentation and white matter posterior
        probability weighted bias correction,
  \item DiReCT-based cortical thickness estimation, and
  \item optional normalization to specified template.
\end{enumerate}
Each component, including both software and data, is briefly detailed 
below with the relevant references for additional information.

%We also note that each component is publicly available with all ANTs 
%algorithms available as open source.%
%\footnote{
%http://www.picsl.upenn.edu/ANTS
%}
Additionally, the coordination of all the algorithmic components is
encapsulated in the shell script \verb#antsCorticalThickness.sh#.  This includes
optimal parameters for each of the algorithmic components which has worked
well for our processing and which are used to acquire the results 
described in this work.

%\lstset{frame = tb,
%        framerule = 0.25pt,
%        float,
%        fontadjust,
%        backgroundcolor={\color{listlightgray}},
%        basicstyle = {\ttfamily\scriptsize},
%        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
%        identifierstyle = {\ttfamily},
%        commentstyle = {\ttfamily\color{listcomment}\textit},
%        stringstyle = {\ttfamily},
%        showstringspaces = false,
%        showtabs = false,
%        numbers = none,
%        numbersep = 6pt,
%        numberstyle={\ttfamily\color{listnumbers}},
%        tabsize = 2,
%        language=,
%        floatplacement=!h,
%        caption={\small \baselineskip 12pt DiReCT long command line menu which is invoked using the `{\ttfamily {-}{-}help}' option.  The short command line menu is obtained by typing `{\ttfamily {-}h}'}.,
%        captionpos=b,
%        label=listing:long
%        }
%\lstsetcpplong
%\begin{lstlisting}
%This script, apb.sh, performs T1 anatomical brain 
%processing where the following steps are currently 
%applied:
%
%  1. Brain extraction
%  2. Brain 3-tissue segmentation
%  3. Cortical thickness
%  4. (Optional) registration to a template
%
%Usage:
%
%abp.sh -d ImageDimension
%       -i T1Image.nii.gz
%       -e BrainExtractionTemplate
%       -m BrainExtractionProbabilityMask
%       -l BrainParcellationTemplate
%       -p BrainParcellationProbabilityMask
%       <OPTARGS>
%       -o OutputPrefix
%
%Example:
%
%abp.sh -d 3 
%       -i t1.nii.gz 
%       -e brainWithSkullTemplate.nii.gz 
%       -m brainPrior.nii.gz 
%       -l corticalLabels.nii.gz 
%       -p corticalLabelPriors.nii.gz 
%       -o output
%
%Compulsory arguments:
%
%     -d:  ImageDimension                        2 or 3 (for 2 or 3 dimensional single image)
%     -a:  Anatomical T1 image                   typically T1.
%     -e:  Brain extraction template             Anatomical template created using e.g. LPBA40 data set with
%                                                buildtemplateparallel.sh in ANTs.
%     -m:  Brain extraction probability mask     Brain probability mask created using e.g. LPBA40 data set which
%                                                have brain masks defined, and warped to anatomical template and
%                                                averaged resulting in a probability image.
%     -l   Brain segmentation template           Anatomical template for brain segmentation.  E.g. NIREP template
%                                                with labels.
%     -p   Brain segmentationpriors              Label probability priors corresponding to the image specified
%                                                with the -l option.  Specified using c-style formatting, e.g.
%                                                -p labelsPriors\%02d.nii.gz.
%     -o:  OutputPrefix                          The following images are created using the specified prefix:
%                                                  * /Users/ntustison/Data//tmp13243//tmpN4Corrected.nii.gz
%                                                  * /Users/ntustison/Data//tmp13243//tmpExtractedBrain.nii.gz
%                                                  * /Users/ntustison/Data//tmp13243//tmp3TissueBrainSegmentation.nii.gz
%                                                  * /Users/ntustison/Data//tmp13243//tmpCorticalThickness.nii.gz
%                                                  * /Users/ntustison/Data//tmp13243//tmpSurfaceCurvature.nii.gz
%
%Optional arguments:
%
%     -s:  image file suffix                     Any of the standard ITK IO formats e.g. nrrd, nii.gz (default), mhd
%     -t:  template for t1 registration
%     -k:  keep temporary files                  Keep brain extraction/segmentation warps, etc (default = false).
%     -w:  white matter label                    white matter label for segmentation (default = 3).
%     -g:  gray matter label                     cortical gray matter label for segmentation (default = 2)
%     -i:  max iterations for registration       ANTS registration max iterations (default = 50x100x20)
%\end{lstlisting}

\begin{figure*}
  \centering
  \includegraphics[width=130mm]{Figures/Kapowski_pipeline.pdf}
  \caption{The ANTs T1 processing workflow containing all elements for 
  determining cortical thickness. Not shown is the optional single subject
  to template registration.}
  \label{fig:pipeline}
\end{figure*}

\subsubsection{Anatomical template construction}

Normalizing images to a standard coordinate system
reduces intersubject variability in population studies.  Various
approaches exist for determining the normalized space such as the selection
of a pre-existing template based on a single subject, e.g. the Talairach
atlas \citep{Talairach1988}, or a publicly available averaged group of
subjects, e.g. the MNI \citep{Collins1994} or ICBM \citep{Mazziotta1995}
templates.  Additionally, mean templates constructed from labeled
data can be used to construct spatial priors for improving segmentation
algorithms.
The work of \cite{avants2010} explicitly models the geometric component of the 
normalized space during optimization to produce such mean templates.  Coupling the intrinsic symmetry of 
SyN pairwise registration \citep{avants2011} and an
optimized sharpening/averaging of the template appearance, Symmetric Group
Normalization (SyGN) is a powerful framework for producing optimal population-specific
templates \citep{avants2010} with arbitrary similarity metric choice.

%One challenge with standard templates is that they may inadvertently bias one's results by enabling better normalization of subjects to which the template is more similar.  This issue is exacerbated when dealing with populations that have high variance (e.g. due to disease) and/or when one's normalization method is low-dimensional (not flexible enough to capture large shape differences).

%Population-specific templates alleviate some of
%the issues with other template approaches by deriving a most representative image from the population
%\citep{Good2001}.  Large deformation registration algorithms also reduce
%this confound by being less sensitive to the deformation distance
%between subject and target.  Some approaches combine both advantages,
%for instance, the diffeomorphic approach of Joshi et al. employs the
%SSD metric and a shape distance to bring the subject group of images
%into alignment \citep{Joshi2004}.  Variants
%include extension to multiple modalities \citep{Lorenzen2006} and small deformations
%\citep{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
%towards a representative image template \citep{Learned-Miller2006}.

The ANTs implementation of this technique is currently available as a shell script, 
\verb#buildtemplateparallel.sh#, and a multivariate version,
\verb#antsMultivariateTemplateConstruction.sh#, both of which are distributed as part of
 the ANTs repository.
The latter script permits the construction of multimodal templates (e.g. 
T1-weighted, T2-weighted, and proton density MRI as described in the 
Evaluation section).  Both scripts accommodate a variety of computational resources
for facilitating template construction.  These computational resource possibilities include:
\begin{itemize}
  \item serial processing on a single workstation, 
  \item parallelized processing on a single workstation with multiple cores using \verb#pexec#%
  \footnote{http://www.gnu.org/software/pexec/pexec.1.html},
  \item parallelized processing using Apple's XGrid technology%
  \footnote{https://developer.apple.com/hardwaredrivers/hpc/xgrid\_intro.html}, 
  \item parallelized processing using Sun Grid Engine for cluster-based systems%
  \footnote{http://www.oracle.com/technetwork/oem/grid-engine-166852.html}, and 
  \item parallelized processing using the Portable Batch System for cluster-based systems%
  \footnote{http://www.pbsworks.com/}.
\end{itemize}
Within this work multiple templates were created for all stages of 
image processing and analysis.  The creation of these templates are described
in the corresponding data section.

%Given a set of representative images, 
%$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
%diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
%\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
%the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
%which minimize the following cost function:
%\begin{align}
%  \sum_{m=1}^M \left[
%    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
%    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
%    \right]
%\end{align}
%where $D$ is the diffeomorphic shape distance, 
%\begin{align}
%  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
%\end{align}
%dependent upon the choice of the linear operator, $L$, and
%$v$ is the diffeomorphism-generating velocity field, 
%\begin{align}
%  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
%\end{align}
%$\Pi$ is th
%e choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
%virtual domain midway between each individual image and the current estimate of the template. 
%
%With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
%to identity, iterative optimization
%involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
%updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.

\subsubsection{N4 Bias field correction}

Critical to quantitative processing of MRI is the minimization of
field inhomogeneity effects which causes artificial low frequency 
intensity variation across the image.  Large-scale studies, such
as the Alzheimer's Disease Neuroimaging Initiative (ADNI), employ
perhaps the most widely used bias correction algorithm, N3 \cite{sled1998}, 
as part of their standard protocol \citep{boyes2008}.

In \cite{tustison2010}, we introduced an extension of N3, denoted as
N4, which demonstrates improved performance and convergence behavior
on a variety of data.  This improvement is a result of an enhanced 
fitting routine (which includes multi-resolution capabilities) and a modified optimization 
formulation.  For our workflow, the additional possibility of specifying
a weighted mask in N4 permits the use of the current white matter probability map 
calculated during the segmentation pipeline for further improvement of 
bias field estimation.  In addition to its public availability 
through ANTs and the Insight Toolkit, it has also been included in
the popular open source Slicer software package for visualization and medical
image computing \cite{fedorov2011}.

N4 is used in two places during the individual subject processing (cf. Figure
\ref{fig:pipeline}).  Following conversion of the raw dicom T1-weighted image
to Nifti format using our related \verb#Neuropipedream# set of raw image conversion
and organization tools%
\footnote{
http://sourceforge.net/projects/neuropipedream/
}, N4 is used to generate an initial bias corrected image for use in
brain extraction.  The input mask is created by adaptively thresholding 
the background from the foreground using Otsu's algorithm \cite{otsu1979}.
Following brain extraction, the three-tissue segmentation involves iterating
between bias field correction using the current white matter posterior 
probability as a weight mask and then using that bias corrected image
as input to the Atropos segmentation step (described in subsequent sections).

\subsubsection{Atropos 3-tissue segmentation}

In \cite{avants2011a} we presented an open source $n$-tissue segmentation software tool
(which we denote as ``Atropos'') attempting to distill 20+ years of active research in this area
particularly some of its most seminal work (e.g. \cite{zhang2001,ashburner2005}). 
Specification of prior probabilities includes spatially varying Markov Random Field modeling, 
prior label maps, and prior probability maps typically derived from our template building 
process.  Additional
capabilities include handling of multivariate data, 
partial volume modeling \cite{shattuck2001}, a memory-minimization mode,
label propagation, a plug-n-play architecture for incorporation of novel likelihood models
which includes both parametric and non-parametric models for both scalar and tensorial
images, and alternative posterior formulations for different segmentation tasks.

\subsubsection{Brain extraction}

Brain extraction using ANTs combines template building, high-performance
brain image registration \citep{avants2011}, and Atropos with topological refinements.  
An optimal template for brain extraction is 
generated offline using labeled brain data.  For example, in this work we use the LPBA40 data 
for generating a brain extraction template and a corresponding brain probability mask which is
available on the website associated with this submission.

  The warped template probability map is thresholded at 0.5 and the resulting mask is dilated
with a radius of 2.  Atropos is used to generate an initial 3-tissue segmentation estimate within the mask
region.  Each of the three tissue masks undergo specific morphological operations which are then
combined to create a brain extraction mask for use in the rest of the cortical thickness workflow.

A comparison using open access brain data with publicly available brain extraction algorithms 
including  AFNI's \verb#3dIntracranial# \citep{ward1999}, FSL's \verb#BET2# \citep{smith2002}, 
Freesurfer's \verb#mri_watershed# \citep{segonne2004}, and BrainSuite \citep{dogdas2005}  demonstrated
that our combined registration/segmentation approach \citep{avants2010a} performs at the top level
alongside BrainSuite (tuned) and FreeSurfer.

\subsubsection{DiReCT Cortical Thickness Estimation}

Although the basic formulation of DiReCT as reported in this work is as it 
was introduced in \cite{das2009}, we have made several 
improvements (perhaps the most significant being that this particular 
ITK-compatible implementation has been significantly multi-threaded,
is written in ITK coding style, and has been made publicly available through 
ANTs complete with a unique user interface design developed specifically for 
ANTs tools.

\subsection{Public Data Resources}

\subsubsection{LPBA40 Data for Skull Stripping}

For the brain extraction step
we used the data from the LPBA40 repository \citep{shattuck2008}.
These data consist of 40 high-resolution 3D Spoiled Gradient Echo
(SPGR) MRI acquisitions which were manually labeled delineating
56 brain structures.  Additional post processing 
included automated brain extraction using FSL's brain extraction tool 
\citep{smith2002} which was followed by manual corrections.  These
40 brain masks are also included with the database.

The first 20 subjects were used to create a population-specific
unbiased average template \cite{avants2010}.  The brain masks corresponding 
to the 20 subjects were warped to the template space using the 
transforms derived during the template building process.  A template 
probability mask was created by averaging the warped brain masks.
For brain extraction of any single individual this ANTS-based LPBA40 template 
is coarsely registered to the single subject brain.  The template
probability brain mask is warped to the individual subject and 
is used as the initial brain mask estimate.  As mentioned previously,
Atropos and binary morphological operations are used to refine the
brain mask estimate.

\subsubsection{NIREP Data for 3-Tissue Segmentation}

The nonrigid image registration evaluation project is an ongoing 
framework for evaluating image registration algorithms \citep{christensen2006}.
The initial data set introduced into the project consists of 
16 (8 male and 8 female) high resolution skull-stripped brain 
data with 32 cortical labels manually drawn using published protocol.
Given the gray matter labels, the white matter and CSF were identified 
for each of the 16 subjects using Atropos.  Similar to the LPBA40
data set, a NIREP template was created from 8 subjects and each the
warped labels were used to create probabilistic estimates of the 
labeled region boundaries. These probability maps were used as 
spatial prior probabilities during the 3-tissue segmentation component
of the pipeline.  Using SyN, the NIREP template is registered to the
extracted individual subject brain which is followed by a warping of the 
NIREP priors to the space of the individual subject.  The initial warped 
white matter probability map is used as the weighted confidence mask 
in the follow-up bias correction step.

\subsection{IXI Data for Pipeline Evaluation}

The IXI data%
\footnote{
http://biomedic.doc.ic.ac.uk/brain-development/
}
consists of approximately 600 healthy subjects imaged at three sites 
using several modalities (T1-weighted, T2-weighted, proton density, magnetic 
resonance angiography, and diffusion tensor imaging).  The 
database also consists of  demographic information such as age, weight,
height, ethnicity, occupation category, educational level, and marital status.
The number of subjects spanning a range of demographic characteristics makes
this a rich data set for validating and exploring correlations with cortical 
thickness measured using the ANTs pipeline.

### 2012-05-22 KapowskiChronicles/paper.tex
ENH:  Adding Kapowski chronicles.

--- 2011-10-17 ANTsTBSS/paper.tex
@@
 \usepackage{float}
+\usepackage{listings}
+\usepackage{color}
+
+    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
+    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
+    \definecolor{listnumbers}{gray}{0.65}
+    \definecolor{listlightgray}{gray}{0.955}
+    \definecolor{listwhite}{gray}{1.0}
+
+\newcommand{\lstsetcpplong}
+{
+\lstset{frame = tb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{listlightgray}},
+        basicstyle = {\ttfamily\scriptsize},
+        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{listcomment}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false,
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=,
+        floatplacement=!h,
+        caption={\small \baselineskip 12pt DiReCT long command line menu which is invoked using the `{\ttfamily {-}{-}help}' option.
+The short command line menu is obtained by typing `{\ttfamily {-}h}'},
+        captionpos=b,
+        label=listing:long
+        }
+}
 
@@
 
-    \usepackage{color}
-
-    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
-    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
-    \definecolor{listnumbers}{gray}{0.65}
-    \definecolor{listlightgray}{gray}{0.955}
-    \definecolor{listwhite}{gray}{1.0}
-
-\newcommand{\lstsetcpp}
-{
-\lstset{frame = tb,
-        framerule = 0.25pt,
-        float,
-        fontadjust,
-        backgroundcolor={\color{listlightgray}},
-        basicstyle = {\ttfamily\scriptsize},
-        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
-        identifierstyle = {\ttfamily},
-        commentstyle = {\ttfamily\color{listcomment}\textit},
-        stringstyle = {\ttfamily},
-        showstringspaces = false,
-        showtabs = false,
-        numbers = none,
-        numbersep = 6pt,
-        numberstyle={\ttfamily\color{listnumbers}},
-        tabsize = 2,
-        language=[ANSI]C++,
-        floatplacement=!h,
-        caption={},
-        captionpos=b,
-        }
-}
-
-\journal{Neuroimage}
+\journal{NeuroImage}
 
@@
 
-\title{Circularity in Population-Based FA Analysis}
+\title{The Kapowski Chronicles:  A Complete, Volumetric Cortical Thickness Open Source Pipeline with Evaluation on Public Data}
 
-\author[label1]{Nicholas J.~Tustison\fnref{label0}}
-%  \ead{ntustison@virginia.edu}
-  \fntext[label0]{\scriptsize Corresponding author:  PO Box 801339, Charlottesville, VA 22908; T:  434-924-7730; email address:  ntustison@virginia.edu }
-\author[label2]{Brian B.~Avants}
+\author[label1]{Nicholas J.~Tustison$^\dagger$
+  \fnref{label0}}
+  \fntext[label0]{\scriptsize Corresponding author:  PO Box 801339, Charlottesville, VA 22908; T:  434-924-7730; email address:  ntustison@virginia.edu.\\ \\
+  $\dagger$The first two authors contributed equally to this work.}
+\author[label2]{Brian B.~Avants$^\dagger$}
 \author[label2]{Philip A.~Cook}
-\author[label3]{Junghoon Kim}
-\author[label3]{John Whyte}
+\author[label2]{Gang Song}
+\author[label2]{Sandhitsu R.~Das}
+\author[label2]{Jeffrey R.~Duda}
+\author[label3]{Niels van Strien}
+\author[label1]{James R.~Stone}
 \author[label2]{James C.~Gee}
-\author[label1]{James R.~Stone}
-
 \address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
@@
                 Philadelphia, PA}
-\address[label3]{Moss Rehabilitation Research Institute, Albert Einstein Healthcare Network, Philadelphia, PA}
+\address[label3]{Department of Circulation and Medical Imaging, Norwegian University of Science and Technology, Trondheim, Norway}
 
@@
 \begin{abstract} 
-Template-based normalization plays a critical role in many population
-studies based on diffusion tensor imaging (DTI).  Several current
-approaches compute the fractional anisotropy (FA) from the DTI and then
-directly register the FA image to a FA template before performing
-statistical comparison between different groups within a population.  In
-this paper, we demonstrate that this analysis strategy induces selection
-bias that varies in strength with the similarity metric used in
-registration.  Specifically, we show that the popular sum of squared
-difference intensity metric maximizes effect size, rather than
-anatomical alignment, and consistently overestimates statistical
-significance.  Consequently, we advocate a more cautious approach
-that normalizes a population of FA images through the associated 
-anatomical images.  We show that the Symmetric Group Normalization
-strategy for unbiased template construction from anatomical (e.g.
-T1-weighted) images provides a valid and unbiased coordinate system for
-statistical inference of FA differences and yet retains detection power
-in comparison to FA-based normalization methods.  We establish this
-outcome by using open data comprised of normal controls from the International
-Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
-brain injury survivors compared with 17 control subjects.  Alignment
-comparisons with the normalization component of the popular tract-based
-spatial statistics (TBSS) show that our anatomically-based template
-approach produces good alignments despite the absence of any direct
-FA-to-FA registration.  Of practical consideration we note that the
-software tools discussed in this work are provided to the public as part
-of the open source Advanced Normalization Tools (ANTs) repository.
-In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the statistical functions used to assess population effects.
-
-%We propose a modification of the popular tract-based spatial statistics (TBSS) framework \citep{Smith2006} for investigating diffusion tensor derived scalar measures across populations.
-Whereas the current crucial alignment step is performed typically using the derived fractional anisotropy images, we propose using the Symmetric Group Normalization strategy for anatomical unbiased template construction from T1-weighted images which serves as the coordinate system for statistical inference.
-A potential benefit of this modified approach is reduction of the general confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.
-Alignment performance is evaluated using open data sets from the International Neuroimaging Data-sharing Initiative.
-This assessment demonstrated improved alignment in both fractional anisotropy (FA) and mean diffusivity (MD) images versus standard TBSS.
-Further analyses evidencing increased regional sensitivity in peripheral white matter to FA differences include a DTI-based assessment of a cohort of 16 traumatic brain injury survivors compared with 17 control subjects.
-Similar to the Oxford Centre for Functional MRI of the Brain (FMRIB) who provide TBSS to the public, the software tools discussed herein are provided to the public as part of the open source Advanced Normalization Tools (ANTs) repository.
+Numerous studies have explored the relationship between cortical
+structure and brain development, cognitive 
+function, and functional connectivity.
+The highly convoluted cortical topography makes manual measurements
+arduous and often impractical given the population sizes necessary 
+for sufficient statistical power.  Computational techniques 
+have permitted large-scale studies as they
+provide robust and reliable localized measurements characterizing
+the cortex with little or no human intervention.  Particularly useful
+to the neuroscience community are publicly available tools, such as the popular 
+surface-based Freesurfer, which facilitate the testing and refinement of 
+hypotheses.  In this paper, we introduce the volume-based 
+Advanced Normalization Tools
+(ANTS) cortical thickness automated pipeline
+comprised entirely of well-vetted components developed as part of the ANTS 
+toolkit such as multivariate template construction,
+SyN (image registration), N4 (bias
+correction),
+Atropos ($n$-tissue segmentation), and DiReCT (cortical thickness) 
+all developed as part of the ANTS open science effort.  
+Complementing the open source aspect of ANTS we confirm 
+previous findings of
+gender differences, age correlation, and hemispherical asymmetry 
+in cortical thickness using the publicly available IXI data
+consisting of multimodal structural MRI of approximately 600 subjects.
+To further promote open science and use of the proposed tools,
+ all scripts and templates used to produce the results
+are hosted on the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC) website.
 \end{abstract}
@@
 \begin{keyword}
-Advanced Normalization Tools (ANTs) \sep diffusion tensor imaging \sep fractional anisotropy \sep template construction
+advanced normalization tools \sep bias correction \sep cortical thickness \sep open source software \sep segmentation \sep skull stripping
 %% keywords here, in the form: keyword \sep keyword
@@
 
-%% main text
-
-\section{Introduction}
-%State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results.
-
-The seminal work of Basser et al.
-\citep{Basser1994a,Basser1994} established diffusion tensor imaging (DTI) as a viable investigatory MRI technique.
-DTI's sensitivity to brain architecture \citep{Basser1996,Assaf2008} enables promising analysis possibilities assessing neuro-structural differences in cross-population studies \citep{Kubicki2005,Arnone2008,Kantarci2010,Rametti2010} including subjects with traumatic brain injury (TBI) \citep{Caeyenberghs2010,Jiang2010,Warner2010}.
-Two popular comprehensive software packages for assessing population FA differences include the SPM Matlab-based toolkit%
-\footnote{
-http://www.fil.ion.ucl.ac.uk/spm/
-}
-for voxel-based morphometry (VBM)
-and the tract-based spatial statistics (TBSS) framework \citep{Smith2006}.%
-\footnote{
-http://www.fmrib.ox.ac.uk/fsl/tbss/index.html
-} 
-VBM \citep{Ashburner2001} continues to find application to FA studies \citep{Kakeda2010,Takao2010,Preziosa2011} despite concerns about its general validity \citep{Bookstein2001,Davatzikos2004} and with respect to DTI \citep{Jones2005,Chung2008}.
-The TBSS framework was partially developed in response to these concerns in which voxel values are projected onto a template white matter skeleton for increased statistical power.
-%(e.g. \citep{Arnone2008,Bodini2009}).
-
-Although there are substantive differences in normalization
-between these and other analysis approaches, a core
-commonality includes direct FA-to-FA registration {\color{red}{using various similarity metrics}}.
-The sum of squared differences (SSD) image similarity metric is perhaps the easiest to interpret (it drives the image difference to zero), is computationally efficient (just compute the image difference and gradient) and is therefore widely used.
-Some of the most popular registration methods (SPM, Demons, FSL's FNIRT) rely on this metric and yield reasonable performance levels \citep{Klein2009}.%
-\footnote{
-In addition to TBSS and SPM, other popular packages in the neuroimaging community 
-which use the SSD metric include AIR \citep{Woods1998} and LDDMM \citep{Beg2005}.
-}
-
-However, suppose the SSD metric is used to find a set of $M$ transformations that map a population of $M$ FA images, $\{I_1, I_2, \ldots, I_M\}$, to a representative FA template, $J$, for statistical testing between groups.
-This results in the transformations,
-$\boldsymbol{\mathcal{T}} = \{\mathcal{T}_1, \mathcal{T}_2,\ldots,\mathcal{T}_M \}$, which minimize the SSD metric over the population, i.e.
-\begin{align}
-  \boldsymbol{\mathcal{T}} = \underset{\mathcal{T}_1, \ldots, \mathcal{T}_M}{\operatorname{argmin}}
-    \sum_{m=1}^M\sum_{n=1}^N \left( I^n_m( \mathcal{T}_m(\mathbf{x}) ) - J \right)^2.
-\end{align}
-Switching the order of the summations (since the transformations are mutually independent) and recognizing that a principal criterion for the selection of $J$ is such that it be a good approximation of the mean of the aligned images, it becomes apparent that the transformation solution
-\begin{align}\label{eq:variance}
-  \boldsymbol{\mathcal{T}} = \underset{\mathcal{T}_1, \ldots, \mathcal{T}_M}{\operatorname{argmin}}
-    \sum_{n=1}^N \underbrace{\sum_{m=1}^M \left( I^n_m( \mathcal{T}_m(\mathbf{x}) ) - J \right)^2}_{
-    \propto\mathrm{\,voxelwise\,variance}}
-\end{align}
-is that which minimizes the voxelwise variance.
-Reducing variance will also directly affect the value of a standard population statistic, the Student's $t$-test.
-Other standard similarity metrics will also exhibit this
-trend to some degree as they minimize some local measure of intensity-based distance between the subject and the template.
-
-Thus, instead of normalization based on anatomical alignment followed by
-statistical testing, direct FA normalization explicitly conflates
-anatomical alignment with optimizing the very statistical testing
-results that one will ultimately use to assess hypotheses.  This immediately evokes a sense of circularity in the analysis \citep{Kriegeskorte2010}.
-According to Kriegeskrote et al., ``An analysis is circular (or nonindependent) if it is based on data that were selected for showing the effect of interest or a related effect.''  As illustrated with
-Eqn.
-\ref{eq:variance}, direct FA-to-FA registration will produce transformations that align (or select) voxels such that the statistical testing result (or the effect of interest) is increased (assuming one is using Student's
-$t$-test).
-While the SSD metric is able to provide visually attractive alignments across subjects, the voxel-level details (that is, which voxels are matched where) are, as we show in the following simulation study, strikingly important in terms of the effect on detection power.
-
-%This results in a significant increase in the number of false positives for this type
-%of testing which we illustrate with the following example.
-
-\paragraph{Simulated Data Example}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedData.pdf}
-\end{tabular}
-\caption{(a) Axial slice from the discrete BrainWeb data set.  From (a) 
-we create the (b) simulated template and sample (c) control and (d) patient images with both 
-additive noise and subsequent Gaussian smoothing. }
-\label{fig:simulated_data}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{simulatedDataWarped.pdf}
-\end{tabular}
-\caption{Warping the image in Fig. \ref{fig:simulated_data}(c) to the template in 
-  Fig. \ref{fig:simulated_data}(b) using the SSD similarity metric produces the image shown in
-  (a).  The corresponding deformation is given in (b).  Although the images were anatomically
-  aligned prior to invoking ANTs registration, deformation is driven by the similarity metric
-  causing the accumulation of Type 1 errors.}
-\label{fig:simulated_data_warped}
-\end{center}
-\end{figure}
-
-Axial slice number 159 was extracted from \verb#subject_04# of the discrete BrainWeb database
-\citep{Aubert-Broche2006} which is shown in Fig. \ref{fig:simulated_data}.  After thresholding out 
-the white matter and assigning all white matter voxels a value of 0.7, 10\% of these white 
-matter voxels were randomly selected to be assigned a value of 0.6 in the simulated control 
-template and a value of 0.3 in the simulated patient template.  
-Note that these values were arbitrarily selected to be more or less in the range of typical
-FA values.  Gaussian noise (mean = 0.0, standard deviation = 0.1) was added to the white
-matter template followed by a Gaussian smoothing (standard deviation = 1.5 mm) shown in Fig. \ref{fig:simulated_data}(b).  A simulated control cohort was created consisting of 20 subjects.  
-Starting with the simulated control template, each of the 20 subjects is created by adding
-Gaussian noise (mean = 0.0, standard deviation = 0.1) followed by 
-Gaussian smoothing (standard deviation = 1.5 mm).
-Thus, each subject is differentiated only by a unique random draw from the noise distribution and there is no real shape difference between subjects.
-A simulated patient cohort of 20 subjects
-was created similarly using the simulated patient template.  A sample control subject and sample
-patient subject are given, respectively, in Fig. \ref{fig:simulated_data}(c) and (d).
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-%  \includegraphics[width=180mm]{simulatedTesting.pdf}
-  \includegraphics[width=130mm]{simulatedTesting.pdf}
-\end{tabular}
-\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the simulated experiment adjusted for multiple comparisons (FDR).
-The true significance curves are drawn in black.  Elevated statistical power is seen with all 4 
-         metrics for both tests although the effect is less in the more robust Wilcoxon results.
-Not shown are the significance curves for the Brunner-Munzel testing \citep{Rorden2007} which showed similar effects to the results on the left.
-        }
-\label{fig:simulated_testing}
-\end{center}        
-\end{figure*}
-
-After creation of the simulated data set, we performed voxelwise testing (both $t$-testing
-and Wilcoxon rank-sum testing, aka the Mann-Whitney test, one-tailed,
-$\alpha = 0.95$) with false discovery rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).
-The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.
-We denote these graphs as "statistical significance curves."
-The true curves for the simulated experiment are plotted in black in Fig. \ref{fig:simulated_testing}.  
-We note that the amount of smoothing increases the proportion of significant voxels beyond
-the prescribed 10\%.
-
-To illustrate the effect of similarity metric on statistical significance, we performed ANTs registration \citep{Avants2011} of each subject to the template
-(Fig. \ref{fig:simulated_data}(b)) using cross correlation (CC), mutual information (MI), 
-Demons \citep{Thirion1998}, and sum of squared differences (SSD). A representative deformed image
-using the SSD metric and its corresponding deformed grid is given in 
-Fig. \ref{fig:simulated_data_warped}. 
-The resulting statistical significance curves are plotted with the true significance curves in Fig. 
-\ref{fig:simulated_testing}.
-
-This example demonstrates the undesirable phenomenon previously
-described.  Even though these images were anatomically aligned prior
-to registration, the resulting deformations (falsely) increased
-detection power.  This effect can be explained in the context of
-Eqn. (\ref{eq:variance}) and by the fact that we added noise to the
-underlying distribution which allows the deformable mapping algorithms
-the freedom to transform the positions of the most "powerful" voxels
-to minimize variance across the population (and increase the $t$-test
-statistics).  Although we do not show the results, additional
-experiments show that decreasing the standard deviation of the noise
-causes an increase in separation between the SSD/Demon's metrics and
-the MI and CC metrics.  MI and CC are closer to the the true
-significance curve.  Alternatively, if we increase the noise level,
-the same ranking occurs with even more spread between significance curves.
-
-Common similarity metric functions can therefore cause false positives by minimizing local intensity 
-variance between the subjects and template which explicitly maximizes statistical
-testing results in contrast to maximizing anatomical alignment.  This effect is strongest in 
-the SSD and Demons metrics and weakest in the MI and CC metrics likely due
-to the degree of correspondence to Eqn. (\ref{eq:variance}).   This is also supported by contrasting
-the $t$-test results given in Fig. \ref{fig:simulated_testing}(a) with the non-parametric Wilcoxon
-rank-sum testing (Fig.
-\ref{fig:simulated_testing}(b)).
-The rank test does not explicitly use variance in its assessment and thus exhibits far less statistical exaggeration even though it offers roughly 95\%
-the power of the $t$-test assuming certain assumptions are met \citep{Siegel1988,Rorden2007}.
-
-Although various processing choices mitigate these effects, such as Gaussian smoothing, nonparametric testing, and the use of more robust similarity metrics, such choices are often made on an ad hoc basis.
-In contrast, in this work we strongly advocate the general application
-%A more principled approach is the use of an anatomical, i.e. T1-weighted, unbiased shape/intensity 
-%template constructed from the population images.  DTI-derived scalar images can then be mapped 
-%to their corresponding T1-weighted image in a highly constrained manner and subsequently 
-%mapped to the template via nonrigid registration of the T1-weighted image.  This minimizes 
-%possible complications associated with direct FA-to-FA image registration.
-%We recognize that using the anatomical images in the alignment process has been previously proposed (in fact, such a suggestion occurs in the {\it Future directions} section of the original TBSS paper).
-of a high performance \citep{Klein2009}, nonrigid registration algorithm \citep{Avants2011}, which forms the underlying machinery for an optimal template construction algorithm \citep{Avants2010}, in facilitating alignment between DTI-derived scalar images for cross-population studies.
-The key point is that the FA images are mapped to the common reference space via intra-subject DT-to-T1 mapping followed by T1-to-T1 alignment.
-The T1-to-T1 step is (statistically) independent of the FA values which are the ultimate statistical target of hypothesis testing in FA population studies.
-Below, we show that this same pattern of effects is reproduced in real data and detail our proposed solution.
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-%  \includegraphics[width=170mm]{workflow.pdf}
-  \includegraphics[width=130mm]{workflow.pdf}
-\end{tabular}
-\caption{Possible ANTs-based TBSS workflow using SyGN.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
-(transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.
-Each FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.
-The individual T1-weighted image is then nonrigidly registered to the anatomical template.
-Optionally, one can register the T1 template to the MNI152 template which resides in the same space as the FMRIB58 template.
-Thus, the composite transform, carefully constructed without any FA-to-FA image registration, can take an individual subject's FA map (or other DTI-derived measure) to the desired standardized coordinate system.
-}
-\label{fig:workflow}
-\end{center}
-\end{figure*}
-
-We first briefly describe the SyGN algorithm used to create anatomical templates.  
-We then explain the the step-by-step process for aligning subject FA images to 
-the anatomical template (which is facilitated with the various tools available 
-in ANTs).  After describing the real image data used in this work, we show the 
-issues raised affect common analysis scenarios.
-
-%
-%
-%
-%It was precisely due to these concerns that the TBSS framework was developed constituting an entire processing pipeline spanning from the reconstruction of the DTI tensorial data
-%%to the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA)), to the nonrigid alignment of the population to a standard coordinate system and skeletonization, and, finally,
-%through to the statistical analysis for discerning differences across populations.
-Major intermediate steps include the calculation of the corresponding DTI-derived scalar measures (typically fractional anisotropy (FA) \citep{Basser1996}), nonrigid alignment of the population to a standard coordinate system, and subsequent ``skeletonization'' of the mean of the aligned FA maps.
-The alignment to a standard space is performed in such a way to remove the smoothing step associated with SPM.
-Skeletonization and its projection across all registered DTI-derived scalar images is meant to accommodate the possibility of imperfectly aligned data.
-However, as shown in a recent study \citep{Zalesky2011}, although the alignment error is reduced with the skeleton projection step, misalignments still remain post-projection.
-Added to this is the potential confounding effect of minimizing data differences via nonrigid registration which are potentially the precise differences to be statistically quantified.
-
-%Because spatial normalization is critical to obtaining valid results, we propose a modification of the standard TBSS pipeline directed at some of these residual concerns which we refer to as ANTs-TBSS.
-Instead of DTI-derived scalar alignment via the images themselves (either random subject selection or locating an approximation to the ``mean subject'') or to a standard template (e.g.
-FMRIB58\_FA\_1mm.nii.gz which we denote as ``FMRIB58'' in the remainder of the text), we evaluate the substitution of an anatomical, i.e.
-T1-weighted, unbiased shape/intensity template constructed from the population images.
-DTI-derived images can then be mapped to their corresponding T1-weighted image in a highly constrained manner and subsequently mapped to the template via nonrigid registration of the T1-weighted image.
-This eliminates any possible complications associated with direct FA-to-FA image registration.
-
-%We first describe in greater detail the proposed ANTs-TBSS framework with an emphasis on comparison with standard TBSS.
-We then describe the two open data sets used to evaluate the proposed methodology in addition to a description of diffuse TBI data
-%also used to illustrate concepts discussed in this work.
-%and analysis previously reported in \cite{Avants2008}.   
-%The complementary assessment performed in this paper is used to provide additional analyses to these previous findings.
-
-%A crucial component of TBSS is the alignment of the DTI-derived scalar images \citep{Zalesky2011}. 
-%
-%Voxel-based morphometry vs. tractography vs. tract-specific analysis
-%
-%Talk about {\em skeleton projection} and how our work is different from \citep{Zalesky2011}.
-%Also talk about the advantages over building a dti-based template as discussed in 
-%\citep{Hecke2010}.
-
-\section{Material and Methods}
-
-%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.
-
-%In order to understand the differences in approaches, we first briefly describe
-%the current TBSS protocol as it is implemented.  This is followed by a
-%brief outline of the proposed modification using ANTs.  Afterwards
-%we give a description of the data sets used to demonstrate the methods.
-
-\subsection{Anatomical Image Template Construction}
-In general, we normalize images to to a standard coordinate system to
-reduce intersubject variability in population studies.  Various
-approaches exist for determining the normalized space such as selection
-of a pre-existing template based on a single subject, e.g. the Talairach
-atlas \citep{Talairach1988}, or a publicly available averaged group of
-subjects, e.g. the MNI \citep{Collins1994} or ICBM \citep{Mazziotta1995}
-templates.
-One challenge with standard templates is that they may inadvertantly bias one's results by enabling better normalization of subjects to which the template is more similar.
-This issue is exacerbated when dealing with populations that have high variance (e.g.
-due to disease) and/or when one's normalization method is low-dimensional (not flexible enough to capture large shape differences).
-
-Population-specific templates (or optimal templates) alleviate some of
-these issues by deriving a most representative image from the population
-\citep{Good2001}.  Large deformation registration algorithms also reduce
-this confound by being less sensitive to the deformation distance
-between subject and target.  Some approaches combine both advantages,
-for instance, the diffeomorphic approach of Joshi et al. employs the
-SSD metric and a shape distance to bring the subject group of images
-into alignment \citep{Joshi2004}.  Variants
-include extension to multiple modalities \citep{Lorenzen2006} and small deformations
-\citep{Geng2009}.  These approaches iteratively minimize group difference in ``congealing''
-towards a representative image template \citep{Learned-Miller2006}.
-
-In contrast, Avants et al. explicitly model the geometric component of the 
-normalized space during optimization.  Coupling the intrinsic symmetry of 
-SyN pairwise registration \citep{Avants2011} and an
-optimized sharpening/averaging of the template appearance, Symmetric Group
-Normalization (SyGN) is a powerful framework for producing optimal population-specific
-templates \citep{Avants2010} with arbitrary similarity metric choice. Given a set of representative images, 
-$\{I_1, I_2, \ldots, I_M\}$, optimization involves finding the set of paired
-diffeomorphic transformations, $\left\{\left(\phi^1_1, \phi^1_2\right), 
-\left(\phi^2_1, \phi^2_2\right), \ldots, \left(\phi^M_1, \phi^M_2\right) \right\}$,
-the optimal template appearance, $J^*$, and corresponding coordinate system, $\psi(\mathbf{x})$,
-which minimize the following cost function:
-\begin{align}
-  \sum_{m=1}^M \left[
-    D\left( \psi(\mathbf{x}), \phi^m_1(\mathbf{x}, 1) \right) + 
-    \Pi \left( I_m\left(\phi^m_2(\mathbf{x}, 0.5) \right), J^*\left(\phi^m_1(\mathbf{x}, 0.5) \right) \right)
-    \right]
-\end{align}
-where $D$ is the diffeomorphic shape distance, 
-\begin{align}
-  D\left(\phi(\mathbf{x}, 0), \phi(\mathbf{x}, 1)\right) = \int_{0}^1 \| v(\mathbf{x}, t) \|_L dt, 
-\end{align}
-dependent upon the choice of the linear operator, $L$, and
-$v$ is the diffeomorphism-generating velocity field, 
-\begin{align}
-  v\left(\phi(\mathbf{x}, t), t \right) = \frac{d\phi(\mathbf{x}, t)}{dt},\,\,\, \phi(\mathbf{x}, 0) = \mathbf{x}.
-\end{align}
-$\Pi$ is the choice of similarity metric, often cross-correlation \citep{Avants2008a}, calculated in the 
-virtual domain midway between each individual image and the current estimate of the template.
-
-With initial assignment of $\left\{\left(\phi^m_1, \phi^m_2\right)\right\}$ and $\psi(\mathbf{x})$ 
-to identity, iterative optimization
-involves estimating the pairwise transformations, estimation of the optimal template appearance, and 
-updating $\psi(\mathbf{x})$ by averaging the current estimate of $\left\{\phi_1^m\right\}$.  Additional details are given in \cite{Avants2010}.  
-%For the interested reader, a set of scripts has been created and is distributed with the ANTs library for facilitating template creation.
-
-\subsection{Anatomically-Based Population Normalization}
-We now detail a standard approach that uses SyGN and intra-subject inter-modality mapping to provide an unbiased strategy for FA normalization.
-An example of our proposed workflow applied to anatomically-based population normalization prior to FA analysis using TBSS is illustrated in Fig.
-\ref{fig:workflow}.
-Instead of direct FA-to-FA registration, our ANTs-based modification can be described as follows:
-
-\begin{enumerate}
-  \item Build the anatomical template from  the population sample or sample subset using SyGN as described in
-  the previous section.
-  \item If only using a sample subset for template creation, register the remaining
-        anatomical data to the template and save the resulting transformation for each         
-        subject indexed by $i$ (i.e. $\mathcal{T}^i_{template}$).
-  \item Find the optimal transformation from each
-        subject's FA to the anatomical template:
-  \begin{enumerate}
-  \item For each anatomical image, segment the white matter (in ANTs, one can use N4 bias correction \citep{Tustison2010} and Atropos $n$-tissue segmentation \citep{Avants2011a}).
-  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical image ($\mathcal{R}^i_{dwi}$).
-  \item Perform distortion correction on the FA to anatomical mapping 
-        by localized nonrigid registration
-        of rigidly transformed FA map ($=\mathcal{R}^i_{dwi} \mathrm{FA}$) using
-        the white matter mask obtained in step 3(a) ($\mathcal{T}^i_{FA}$).
-  \item For each subject, concatenate the transforms which maps each
-        subject's FA to the template ($\mathcal{T}^i_{total} = \mathcal{T}^i_{template} \circ 
-        \mathcal{T}^i_{FA} \circ \mathcal{R}^i_{dwi} \mathrm{FA}$).
-  \item Warp each subject's FA map to the template using $\mathcal{T}^i_{total}$.  
-  Equivalently, one could warp the mean diffusion (MD), axial diffusion (AD) or any 
-  other scalar derivable from DTI to the template.  
-%  One other detail is that we prefer to warp the subject's DTI to 
-%  the template and then, in template space, derive the scalar images of interest.  
-  \end{enumerate}      
-%  \item Perform the remaining portions of the standard TBSS pipeline:
-%    \begin{enumerate}
-%      \item Create the mean FA image from all template-aligned FA images.
-%      \item Produce the white matter skeleton from the mean FA image.
-%      \item Create skeleton distance map and project all FA data onto the skeleton.
-%      \item Run the statistical analysis.
-%    \end{enumerate}
-\end{enumerate}
-The next section compares this framework to FA-based normalization strategies with respect to both a patient-control study and a study of gender effects in a control population.
-
-\subsection{Imaging Data}
-Diffusion tensor data associated with each of the following data sets were reconstructed from the diffusion weighted sequences using Camino%
-\footnote{
-http://camino.org.uk/
-}---an open source toolkit for diffusion MRI processing and analysis \citep{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.
-
-\subsubsection{Diffuse Traumatic Brain Injury Data}
-
-The TBI data used in this study is part of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in long-term survivors of TBI (principal investigator: John Whyte).
-A more detailed data description can be found in \cite{Avants2008}.
-The following is a brief description relevant to our particular study.
-17 controls and 16 patients with TBI were used for the analysis presented in this work.  Each patient had a history of non-penetrating traumatic brain injury 
-of at least moderate severity defined by significant and well-documented
-loss or alteration of consciousness following injury in addition to meeting
-several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and years of education.
-High resolution T1-weighted anatomic images were obtained using a 3D MP-RAGE 
-imaging sequence with the following acquisition parameters: TR = 1620 ms, 
-TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 
-1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 
-1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
-DTI images were also obtained. 
-%
-%\subsubsection{Participants}
-%{\color{red}{Quote:}} The data were collected as part of a larger study investigating the neural correlates of attention deficits and treatment responses of various psychoactive drugs in the survivors of TBI (principal investigator: J.W.).
-Thirty individuals with TBI and 20 healthy volunteers were recruited.
-We planned to recruit more TBI participants because data of TBI survivors are more likely to be discarded in a typical functional neuroimaging study due to movements in the scanner and poor behavioral performance.
-TBI participants were recruited from a variety of clinical services at MossRehab and through a consent-based registry of individuals with TBI who are interested in participating in rehabilitation research.
-To be included, participants had to be between the ages of 16 and 60, and to have a history of non-penetrating traumatic brain injury of at least moderate severity at least 3 months prior to enrollment.
-Severity level was defined by significant and well-documented loss or alteration of consciousness following injury (i.e., lowest Glasgow Coma Scale (GCS) score of less than 12, or prospectively documented post-traumatic amnesia (PTA) of greater than 1 hour), or focal abnormality on a neuroimaging study that was attributable to traumatic injury.
-A subjective complaint of attention difficulties by the participant, treating clinician, or caregiver was also required.
-Potential participants were excluded if they had a history of prem orbid neurologic disease, psychosis, major affective disorder, mental retardation, Attention Deficit Hyperactivity Disorder, or if they were currently abusing alcohol or recreational drugs.
-Persons who were taking psychoactive medications other than anticonvulsants were also excluded.
-Participants and/or their involved caregivers (depending on the participantÕs cognitive capacity) provided informed consent.
-The study protocol was approved by the Albert Einstein Healthcare Network and the University of Pennsylvania IRBs.
-Twenty healthy volunteers, matched to patients for age, gender, handedness, years of education, and ethnicity, participated in the study.
-Control participants were recruited based on the same inclusion/exclusion criteria as patients, with the exception that they never had a TBI resulting in loss or alteration of consciousness, nor suffered attention complaints.
-Control participants were recruited through the family and friendship networks of the participants with TBI, and through public advertising.
-%Data from one TBI participant whose MRI scan showed a large area of encephalom alacia over almost the entire right hemisphere were excluded.
-The remaining 29 subjects with TBI included 21 men and 8 women aged between 18 and 58 years (mean age = 36.9, SD = 11.4) with a mean education of 13.1 years (SD = 2.8).
-Twenty three of them were right-handed (Edinburgh Handedness Inventory, Oldfield, 1971).
-Fourteen of them were Caucasians, 10 African Americans, 4 Hispanics, and 1 Asian.
-Selected demographic and clinical characteristics of the TBI survivors are reported in Table 1.
-Control participants included 17 men and 3 women aged between 21 and 50 years (mean age = 34.9, SD = 9.8) with a mean education of 13.1 years (SD = 1.7).
-Sixteen of them were right-handed.
-Eleven of them were Caucasians, 7 African Americans, 1 Asian, and 1 unknown.
-The two groups did not differ significantly in terms of age, gender, ethnicity, handedness or years of education (tested with t-test or FisherÕs exact test, as appropriate).
-%
-%\subsubsection{Image acquisition}
-%The functional imaging was conducted on a Siemens 3.0 T Trio whole-body scanner (Siemens AG, Erlangen, Germany), using a standard Transmit/Receive head coil.
-High resolution T1-weighted anatomic images were obtained using 3D M PRAGE imaging sequence using the following acquisition parameters: TR = 1620 ms, TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 1NEX with a scan time of 6 minutes.
-The resulting voxel size was 1 mm$^3$.
-%Lesion assessment
-%To quantify the volume of lesions, a trained observer (J.P.) manually segmented the lesion area under supervision of a neurologist (H.B.C.) with extensive experience in lesion assessment.
-Focal lesions included any cystic cavities and other focal regions of abnormal signal in the white or gray matter.
-The ITK-SNAP software \citep{Yushkevich2006} was used for a 3D-based segmentation.
-Total lesion volume was calculated using a stand-alone utility provided by VoxBo software (Center for Functional Neuroimaging, Philadelphia, PA, http://www.voxbo.org).
-
-\subsubsection{Data from the International Neuroimaging Data-sharing Initiative (INDI)}
-
-In support of open science, the 1000 Functional Connectomes Project (FCP)%
-\footnote{
-ttp://fcon\_1000.projects.nitrc.org
-} 
-was initiated on December 11, 2009 by various members of the MRI community \citep{Biswal2010}.
-Motivated by the absence of neuroimaging data for research, this initiative seeks to form collaborative partnerships with imaging institutions for sharing well-documented multi-modal image sets accompanied by phenotypic data.
-Commitments from institutions such as the NYU Institute for Pediatric Neuroscience and the Nathan Kline Institute (NKI)--Rockland have already resulted in prospective data repositories
-currently distributed to the public via the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC).%
-\footnote{
-http://www.nitrc.org
-}
-%Data from these two institutions were used to demonstrate the alignment properties obtained using SyGN for DTI-derived data normalization.
-
-\paragraph{NKI/Rockland Sample}
-Data from the first 14 weeks of prospective data-sharing were downloaded on January 15, 2011 comprised of 74 subjects (average age in years: 32.4 $\pm$ 17.8).  
-Due to various issues (e.g.
-lack of corresponding T1 image, failed DTI reconstruction, and age matching requirements) only 57 subjects were used for the first alignment analysis and 27 subjects of the NKI data set were used for the second analysis consisting of 9 females (age: $29.3 \pm 6.6$ years) and 18 males (age: $ 26.6 \pm 6.9$ years).
-Germane to our evaluation, each imaging session for each subject produced a 64-directional diffusion tensor imaging scan (parameters: conventional single-shot spin echo
-EPI pulse sequence, TR = 10000 ms, TE = 91 ms, axial acquisition, 
-voxel size = $2 \times 2 \times 2$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2500 ms, b = 1000 s/mm$^2$ for each direction, 58 contiguous slices of 2.0 mm thickness,
-TI = 1200 ms, TE = 3.5 ms, flip angle = 8$^\circ$, 192 contiguous slices of 
-1.0 mm thickness, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1 mm$^3$).  Images were anonymized including defacing of the T1 images prior to upload.
-
-%\paragraph{NYU Institute for Pediatric Neuroscience Sample}
-%Data from the first 3 parts of prospective data-sharing from NYU were also downloaded on January 15, 2011 comprised of 49 subjects (average age in years: 30.29 $\pm$ 8.81).
-All subjects were used in the image alignment assessment.
-Two 64-directional DTI scans were performed (parameters: conventional single-shot spin echo
-%EPI pulse sequence, TR = 5200 ms, TE = 78 ms, axial acquisition, 
-%voxel size = $3 \times 3 \times 3$ mm$^3$) and a T1 anatomical scan (parameters:  TR = 2530 ms, TE = 3.25 ms, flip angle = 7$^\circ$, 256 contiguous slices of 1.33 mm, FOV = 256 $\times$ 256 mm$^2$, and voxel size = 1.3 mm$^3$).
-Images were anonymized including defacing of the T1 images prior to upload.
-
-%
-%The 1000 Functional Connectomes Project (FCP)%
-%\footnote{
-%http://fcon\_1000.projects.nitrc.org/index.html
-%}
-%
-%{\color{red}{Quote}}  INDI is a next-generation data-sharing effort, with the goal of sharing phenotypically rich imaging datasets, on both retrospective and prospective bases.
-On 10/8/10, the Nathan Kline Institute (NKI) / Rockland Sample led the way for INDI-Prospective, starting weekly uploads of 5-7 subjects / week (R-fMRI, DTI, and extensive psychiatric and cognitive phenotypic data, including IQ), which continue to progress (currently n = 55 and growing!).
-More recently, the NYU Institute Pediatric Neuroscience released its sample examining cocaine dependence in adults via INDI-Retro.
-The list of sites pledging data from around the world is growing, and we are working to generate common phenotypic variables across datasets as possible (notice the frequency with which IQ is starting to appear Ð itÕs just one of our targets!).
-%
-%\subsubsection{Nathan Kline Institute (NKI) / Rockland Sample}
-%
-%{\color{red}{Quote}}  The NKI/Rockland Sample is intended to be a phenotypically rich neuroimaging sample, consisting of data obtained from individuals between the ages of 6 and 85 year-old.
-All individuals to be included in the sample undergo semi-structured diagnostic psychiatric interviews, and complete a battery of psychiatric, cognitive and behavioral assessments in order to provide comprehensive phenotypic information for the purpose of exploring brain/behavior relationships.
-%
-%Each subject consists of the following data:
-%\begin{itemize}
-%\item 10 minute resting state fMRI scan (R-fMRI) 
-%\item 6-direction diffusion tensor imaging (DTI) scan 
-%\item 64-direction diffusion tensor imaging scan (2mm isotropic)
-%\item MPRAGE anatomical scan
-%\item A variety of psychiatric, cognitive and behavioral assessments
-%\end{itemize}
-%
-%\subsubsection{NYU Institute for Pediatric Neuroscience}
-%
-%{\color{red}{Quote}}  Summary: The NYU Phyllis Green and Randolph C?wen Institute for Pediatric Neuroscience (IPN) sample will be a collection of past and present scans obtained from psychiatrically screened individuals ranging in age from 6 to 55 years old.
-The initial release consists of datasets from 49 psychiatrically neurotypical adults, with age, gender and intelligence quotient (IQ) information provided.
-Future releases will include more comprehensive phenotypic information, and child and adolescent datasets, as well as individuals from clinical populations.
-%
-%Each subject consists of the following data:
-%\begin{itemize}
-%\item At least one 6-minute resting state fMRI scan (R-fMRI)%
-%\footnote{
-% Most participants have 2 R-fMRI scans, collected less than 1 hour apart in the same scanning session. Rest$_1$ is always collected first.
-%}
-%\item One high-resolution T1-weighted mprage, defaced to protect patient confidentiality
-%\item Two 64-direction diffusion tensor imaging scans
-%\item Demographic information (age, gender) and IQ-measures (Verbal, Performance, and Composite; Weschler Abbreviated Scale of Intelligence [WASI])
-%\end{itemize}
-%
-
-\subsection{FMRIB Software Library (FSL)}
-
-FSL hosts a large number of algorithms for computational neuroanalysis.
-These include, among others, well-known tools for image segmentation (FAST---\citep{Zhang2001}), linear (FLIRT---\citep{Jenkinson2001}) and nonlinear (FNIRT) image registration, brain extraction (BET---\citep{Smith2002}), and TBSS and made publicly available by the Oxford Centre for Functional MRI of the Brain (FMRIB).
-Based solely on the number of citations in the literature,%
-\footnote{
-As of the beginning of the year 2011, the original Neuroimage 2006 TBSS paper has been cited more than 360 times making it the most cited in the journal's library.
-}
-usage of TBSS has become a de facto quantitative tool for DTI analysis based, at least in part, on its accessibility and ease of use.
-Due to its large presence in the neuroimaging community, many of the comparisons that are performed in the next section use the FMRIB58 template (shown in the upper left of Figure \ref{fig:workflow}).
-This template serves as the final coordinate system for most of the comparisons whether it be based on direct normalization or T1 normalization to the MNI152 template which resides in the FMRIB58 space.
-We also used the normalization component of TBSS composed of FLIRT and FNIRT registration steps optimized for normalization to FMRIB58 to provide an external comparative baseline (which we denote in subsequent sections as ``tbss'').
-
-%For the comparative evaluation studies used in this manuscript, TBSSv1.2%
-%\footnote{
-%Build 416 downloaded from the FMRIB web site http://www.fmrib.ox.ac.uk/fsldownloads/ on November 18, 2010.
-%}
-%was used.  The actual code pipeline constituting TBSS prior to statistical analysis consists of the following 4 shell scripts:
-%\begin{description}
-%  \item[tbss\_1\_preproc]  Input is a directory of FA images which are slightly eroded and then organized into a separate FA directory.
-A mask is also generated for each input FA image for use during the spatial normalization process.
-%  \item[tbss\_2\_reg]  The organized FA data and corresponding masks produced in the previous step are warped to one of three possible coordinate systems using FLIRT/ FNIRT:  1)  the atlas included with the FSL download---FMRIB58---(recommended), 2) a user-specified target image or 3) an exhaustive search for the ``optimal'' target image.
-In \citep{Smith2006}, page 1491 section {\it Identifying the target for alignment}, the latter option was advocated based on the observation that average FA images tend to be blurred in contrast to the FA of a single subject.
-However, even though it would be faster to simply choose a single subject at random, the complexities of nonrigid image registration would seemingly promote the exhaustive search over random selection.
-As a compromise between these two options, the first option in the script is to use FMRIB58 which also resides in the space of the MNI152 T1 template.
-It should also be noted that this particular script invokes another FSL script, {\bf fsl\_reg}.
-An option associated with {\bf fsl\_reg} permits the use of an FNIRT configuration file which has been optimized for use with the FMRIB58 atlas.
-%  \item[tbss\_3\_postreg] The mean FA image is generated from the aligned FA images which is subsequently skeletonized.
-%  \item[tbss\_4\_prestats] The skeleton mask is created by thresholding the mean FA skeleton.  The skeleton distance map is then generated and all FA data are projected onto the skeleton.
-%\end{description}
-
-\section{Results}
-%Results should be clear and concise.
-
-\subsection{TBI Cohort}
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-%  \includegraphics[width=180mm]{tbiTesting.pdf}
-  \includegraphics[width=130mm]{tbiTesting.pdf}
-\end{tabular}
-\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).
-Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn.
-(\ref{eq:variance}), namely the SSD and Demons metric which shrinks drastically for the Wilcoxon rank-sum test.
-Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.
-TBSS and the 4 ANTs metric
-results were all obtained using direct registration involving subject FA and the FMRIB58 template.
-The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
-\label{fig:tbi_testing}
-\end{center}        
-\end{figure*}
-
-We constructed a patient template and a control template from the T1-weighted images of the two groups.  These two 
-templates were then used to create a third common template.  
-To compare with TBSS normalization, the common template was
-registered to the MNI152 template.  Thus, the transformation
-which takes the FA image of an individual subject to the FMRIB58
-template is the composition of transforms: 
-subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template.
-We evaluated this strategy relative to direct FA-to-FA
-ANTs registration with the common similarity metrics used in
-the simulated data example described earlier.
-Standard TBSS alignment was also performed.
-
-Cumulative distribution functions (one-tailed, $\alpha = 0.95$)
-are shown in Fig. \ref{fig:tbi_testing}.
-Similar to previous findings with the simulated data, the SSD and 
-Demons metric show elevated statistical power using the Student's 
-$t$-test over the other alignment protocols which all exhibit similar 
-strength.  However, this discrepancy decreases significantly with the 
-nonparametric testing.  The Demons and SSD power decreases while the others
-remain fairly similar.  Interestingly, the ANTs template-based approach
-trends towards better performance relative to the remaining 3 strategies.
-It should be noted that although TBSS uses the SSD metric, the fixed and moving images are smoothed and subsampled even at the highest resolution level which substantially decreases the local effect of the metric.
-
-\begin{table}
-\begin{center}
-  \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
-  \label{table:lom}
-  \begin{tabular*}{85mm}{@{\extracolsep{\fill}} cccccc}
-  \hline
-%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-%  \cline{2-5}
-  {Norm.} & Target & Jaccard & Dice  & False $-$ & False $+$\\
-  \hline
-  \hline
-  antsT1 & 0.917 & 0.621 & 0.767 & 0.082 & 0.341 \\
-  antsCC & 0.780 & 0.570 & 0.727 & 0.220 & 0.320 \\
-  antsMI & 0.776 & 0.569 & 0.725 & 0.223 & 0.320 \\
-  antsDemons & 0.780 & 0.566 & 0.723 & 0.220 & 0.326 \\
-  antsSSD & 0.783 & 0.562 & 0.720 &  0.217 & 0.334 \\
-  tbss & 0.716 & 0.550 & 0.710 & 0.283 & 0.295 \\
-  \hline
-  \end{tabular*}
-\end{center}  
-\end{table}
-
-%\begin{table}
-%\begin{center}
-%  \caption{Overlap measures of thresholded mean FA and MNI152 white matter.}
-%  \label{table:lom}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccccc}
-%  \hline
-%  {Norm.} & Target & Jaccard & Dice  & False $-$ & False $+$\\
-%  \hline
-%  \hline
-%  antsT1 & 0.917 & 0.621 & 0.767 & 0.082 & 0.341 \\
-%  antsCC & 0.780 & 0.570 & 0.727 & 0.220 & 0.320 \\
-%  antsMI & 0.776 & 0.569 & 0.725 & 0.223 & 0.320 \\
-%  antsDemons & 0.780 & 0.566 & 0.723 & 0.220 & 0.326 \\
-%  antsSSD & 0.783 & 0.562 & 0.720 &  0.217 & 0.334 \\
-%  tbss & 0.716 & 0.550 & 0.710 & 0.283 & 0.295 \\
-%  \hline
-%  \end{tabular*}
-%\end{center}  
-%\end{table}
-
-The CDF of the voxelwise FA coefficient of variation (CV) of the aligned
-images is shown for each case in Figure \ref{fig:tbi_cov}.  
-The CV at each voxel is defined as the variance
-divided by the mean over the population of FA values.  Thus,
-those metrics which best minimize the voxelwise variance would
-trend towards greater alignment as quantified by the CV.
-
-%This is very much related to earlier findings in that strategies evincing
-%greater alignment, as quantified by the CV,
-%will trend towards smaller variance thus exhibiting a left-shifted
-%CDF.
-
-Although direct FA-to-FA registration using ANTs with common
-similarity metrics gives ``better alignment'' such alignment is 
-produced at a cost of significant Type 1 errors.  Common subsampling and
-smoothing strategies and local neighborhood metrics such as CC and MI reduce the 
-Type 1 error problem (as evidenced in Fig. \ref{fig:tbi_testing}) 
-but also show a decrease in alignment.  This is further supported
-by the overlap measures \citep{Crum2006} given in Table \ref{table:lom}.
-For each alignment strategy, we thresholded the derived mean FA image (FA $
-\geq$ 0.2) and compared it with white matter mask of the MNI152 template
-segmented using Atropos \citep{Avants2011a}.  The template-based
-strategy showed the best overall overlap with the white matter MNI152
-segmentation.
-
-Finally, qualitative assessment is provided in Fig. \ref{fig:meanFA} 
-in which the mean FA image derived from the template-based approach
-and the post-normalization component of the TBSS pipeline (Fig. \ref{fig:meanFA}(c)) 
-is juxtaposed 
-with the mean FA image produced by the TBSS pipeline alone (Fig.\ref{fig:meanFA}(a)).
-The skeleton masks were produced by thresholding the skeleton values at 
-FA $\geq 0.2$ and it would appear that the improved alignment produces 
-a more extensive skeletal network for greater localization testing of 
-significant differences, particularly in the outer white matter regions
-(compare Fig. \ref{fig:meanFA}(b) and Fig. \ref{fig:meanFA}(d)).
-%Given that TBSS is intrinsically a statistical method for finding local 
-%differences, improved alignment would presumably provide increased regional sensitivity to differences.
-
-\begin{figure}
-  \begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{tbiCoV.pdf}
-\end{tabular}
-\caption{Comparison of the different alignment strategies in normalizing the TBI data to the FMRIB58 template.  Assessment is quantitated using the voxelwise CV of the FA.  TBSS and the 4 ANTs metric
-results were all obtained using direct registration involving subject FA and the FMRIB58 template.
-The curve denoted as ``antsT1'' was also obtained in FMRIB58 space with the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$
-common T1 template $\rightarrow$ MNI152 (FMRIB58) template. }
-\label{fig:tbi_cov}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{meanFA.pdf}
-\end{tabular}
-\caption{Qualitative comparison of mean FA images used in TBSS taken from the TBI data described in Section 2.
-The mean FA image illustrated in (a) is created by mapping the sample FA population to the default FMRIB58 template which is the standard protocol for TBSS.
-In comparison, the mean template represented in (c) is created from alignment of the population sample to the optimal T1 template as described in the paper.
-The respective skeleton masks in (b) and (d) are created by thresholding the resulting skeleton values $\geq 0.2$.
-}
-\label{fig:meanFA}
-\end{center}
-\end{figure}
-
-%
-%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Fig.
-\ref{fig:hoonData}.
-The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.
-
-%As an additional quantitative comparison, we calculated the volumetric ratio between areas of significance and the volumetric brain mask for both $p = 0.01$ and $p = 0.05$.
-For standard TBSS, the MNI 152 "first brain mask" was used and for the ANTs template, Atropos was used to obtain a mask by skull stripping of the final template.
-These values are given in Table \ref{table:hoon}.
-%
-%
-%\begin{table}
-%  \caption{Volumetric ratio differences between ANTsTBSS and TBSS.}
-%  \label{table:hoon}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} cccc}
-%  \hline
-%  {} & $\frac{V(\mathrm{mean\,FA\,mask})}{V(\mathrm{brain\,mask})}$ & $\frac{V(\mathrm{sig.\,areas})}{V(\mathrm{mean\,FA\,mask})}$ & $\frac{V(\mathrm{sig.\,areas})}{V(\mathrm{brain\,mask})}$ \\
-%  \hline
-%  $p \leq 0.01$ & {} & {} & {} \\
-%  $p \leq 0.05$ & {} & {} & {} \\
-%  \hline
-%  \hline
-%  \end{tabular*}
-%\end{table}
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{hoonData2.pdf}
-%\end{tabular}
-%\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).
-Axial slices were chosen based on visual assessment of correspondence in both the ANTs template and MNI 152 template.
-%On the left is the thresholded skeleton overlaid on the mean FA image.  
-%On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).
-The yellow arrow points to an area of significant difference in the peripheral white matter found using ANTs-TBSS.
-%}
-%\label{fig:hoonData}
-%\end{figure}
-
-\subsection{INDI NKI/Rockland Sample}
-In this section, further exploration of relevant issues employs 
-the NKI/Rockland sample of the INDI database.  
-The first is one of experimental set-up consideration in which 
-we investigate how template-based
-alignment is affected by the number of subject images used to 
-create the template.  Both FA and mean diffusivity (MD) images were used.
-This is followed by statistical analysis
-of discriminating power of gender FA differences using 
-the various alignment strategies.
-
-\subsubsection{Alignment Assessment of Template Group Size}
-
- \begin{figure}
-  \begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{NKI2.pdf}
-\end{tabular}
-\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.
-(a) The standard FMRIB58 template.
-(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
-(c) and (d) The mean FA image produced using the ANTs template-based strategy employing 5 and 30 randomly chosen subjects, respectively, warped to the space of (a) followed by T1 registration of all subjects to the resulting anatomical template.
-Note the visual alignment difference particularly in the peripheral white matter complex.
-}
-\label{fig:NKI}
-\end{center}
-\end{figure}
-
-Four templates were generated from $\{5,10,20,30\}$ randomly selected subjects 
-to see if the number of subjects comprising the template affected the alignment 
-performance. In addition to the composite transforms
- to take each subject's FA or MD image to the population-specific anatomical template, we also registered
-each ANTs T1 template to the MNI152 template.  
-This permitted a
-standard space for qualitative and quantitative comparison with TBSS-aligned
-FA and MD images.  
-After template construction, the remaining
-images were registered to the resulting template and, ultimately, 
-to the MNI152 (FMRIB58) template via the template-to-MNI152 transformations.
-
-Sample axial slices for the FA FMRIB58-aligned mean images for the NKI data 
-set of template sizes of 5 and 30 images are provided in Fig. \ref{fig:NKI}.  
-We also include the same axial slice  of the FMRIB58 template in Fig. \ref{fig:NKI} 
-(a) for comparison.  In both sets of sample slices, visual inspection would  
-indicate better 
-performance with the ANTs-based FA mean images compared with the standard 
-TBSS-produced FA mean image.  Also, there is little discernible difference
- between the different ANTs results.  Alignment differences between the 
- approaches are particularly apparent in the peripheral white matter where 
- the anatomical structure in the ANTs template-based FA mean images is visually differentiable.
-
-Relative performance is quantified in the plots produced in Fig. \ref{fig:NKIplots}. 
-The mean FA image derived from the standard TBSS approach was thresholded at FA $\geq 0.2$. 
- This binary mask obtained in the NKI cohort provided the standard group mask for 
- fair comparisons for all evaluations.    For all FMRIB58-aligned images, the CV 
- was calculated in a voxelwise fashion within the mask.  Intuitively, 
- one would expect a smaller variance across better aligned sets of voxel values thus 
- producing a lower CV for a given mean value.  By calculating the CDF of the set of CV values within the mask, quantitative assessment of alignment performance can be visualized.  
-% In addition to producing CDF plots for voxels within the entire brain mask, we separated the brain into inner and outer sections to investigate the alignment difference between these two regions.
-This division was accomplished by applying a distance transform \citep{Maurer2003} to the FSL-provided MNI152 brain mask and thresholding the result such that the outer 30 mm form the outer mask and the subtracted binary image forms the inner mask.
-Note that only those voxels that were also within the thresholded FA mask were considered in the inner/outer analysis.
-
-There is a visible improvement in alignment for both data sets using the ANTs-based approach over individual subject FA registration to the FMRIB template.
-There is little difference between the ANTs results with only a slight trend in improved performance with the larger template subject sample sizes.
-%This was the case for all three regional CV calculations for both data sets.
-Also, as expected, there is greater misalignment in the outer region versus the inner core of the white matter mask regardless of methodology.
-However, the performance tally was the same in both the inner and outer regions as was the case with the entire white matter mask.% although the disparity in performance was much less for the NYU data set as the NKI data set most likely due to the differences in resolution of the original DTI.
-
-%The improvements seen in the CDF plots are also supported by comparative quantities derived from the skeletons of the different mean FA images.
-Since the skeletonization process is meant to determine the center of the various white matter tracts \citep{Smith2006}, a greater volume of skeletonization voxels would indicate identification of more centers of white matter tracts.
-In addition, as described in \cite{Zalesky2011}, the number of connected components of the white matter skeleton is directly related to the contiguity and thus inversely related to the level of structure.
-Table \ref{table:vol} provides the volume of the mean FA skeletons (thresholded at FA $\geq 0.2$) for standard TBSS protocol using the FMRIB template and the ANTs approach using the various templates.
-In agreement with the CDF plots illustrating improved alignment with ANTs, the voxel volumes of the skeletons are greater using ANTs templates with the greatest difference seen in the outer portion of the white matter.
-%Related, in Table \ref{table:cc}, we provide the number of connected components for both data sets for all isolated components greater in size than 1 voxel and 2 voxels.
-Even though the skeletal volumes associated with the ANTs approach are greater, the number of connected components is less which would also support the observation of improved alignment.
-
-%\begin{table}
-%  \caption{Volume (in voxels) of thresholded mean FA skeletons.}
-%  \label{table:vol}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
-%  \hline
-%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-%  \cline{2-5}
-%  {} & inner & outer & inner & outer \\
-%  \hline
-%  FMRIB & 84899 & 122306 & 72765 & 76769 \\
-%  ANTs$_5$ & 87572 & 134421 & 73650 & 94177 \\
-%  ANTs$_{10}$ & 86908 & 130569 & 73419 & 98014 \\
-%  ANTs$_{20}$ & 86523 & 129729 & 73577 & 95484 \\
-%  ANTs$_{30}$ & 86971 & 128892 & 73443 & 94862 \\
-%  \hline
-%  \hline
-%  \end{tabular*}
-%\end{table} 
-%
-%\begin{table}
-%  \caption{Number of connected components of thresholded mean FA skeletons.}
-%  \label{table:cc}
-%  \begin{tabular*}{88mm}{@{\extracolsep{\fill}} ccccc}
-%  \hline
-%  {} & \multicolumn{2}{c}{NKI} & \multicolumn{2}{c}{NYU} \\
-%  \cline{2-5}
-%  {} & $\geq$ 1 voxel & $\geq$ 2 voxels & $\geq$ 1 voxel & $\geq$ 2 voxels \\
-%  \hline
-%  FMRIB & 4340 & 1802 & 2431 & 993 \\
-%  ANTs$_5$ & 3561 & 1384 & 2278 & 916 \\
-%  ANTs$_{10}$ & 3413 & 1302 & 2232 & 899 \\
-%  ANTs$_{20}$ & 3283 & 1252 & 2093 & 885 \\
-%  ANTs$_{30}$ & 3114 & 1243 & 2128 & 867 \\
-%  \hline
-%  \hline
-%  \end{tabular*}
-%\end{table}
-
-%Since our proposed modification of TBSS impacts strictly the
-%critical alignment step, the first set of evaluations demonstrates
-%the difference between the two methods, i.e. spatial normalization 
-%via registration of the FA maps to the FMRIB58 template
-%or spatial normalization via ANTs template construction.
-
-%It is important to note that we did not consider normalization either to a random subject
-%or to the subject best representative of the mean as defined in 
-%\citep{Smith2006}. The former was not investigated due to the possible bias
-%in selecting a random subject as described in \cite{Smith2006}.  
-%Neither was the latter option pursued due to the computational complexity
-%required for such an $\OO{n^2}$ search over the relatively large NKI data set.  In addition, normalization to
-%the FMRIB58 template is both the default and recommended option
-%with the parameters of FNIRT optimally tuned for said option.
-
-%To warp the MD images following FA registration we used FSL's {\bf applywarp} function using the FA-derived transform.
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{NYU.pdf}
-%\end{tabular}
-%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.
-(a) The standard FMRIB58\_FA\_1mm template.
-(b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).
-%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).
-Note the visual alignment difference particularly in the peripheral white matter complex.
-%}
-%\label{fig:NYU}
-%\end{figure}
-
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{inner.png}
-%\end{tabular}
-%\caption{
-%}
-%\label{fig:inner}
-%\end{figure}
-%
-%\begin{figure}
-%\begin{tabular}{c}
-%  \includegraphics[width=80mm]{outer.png}
-%\end{tabular}
-%\caption{Three-canonical view of the inner
-%}
-%\label{fig:outer}
-%\end{figure}
-
-%\begin{figure*}
-%\begin{tabular}{c}
-%  \includegraphics[width=180mm]{NYUplots2.pdf}
-%\end{tabular}
-%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
-\ref{fig:NKI}(b)).
-In addition to the whole brain analysis (first column), the CV was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-%}
-%\label{fig:NYUplots}
-%\end{figure*}
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-  \includegraphics[width=85mm]{NKIplots3.pdf}
-\end{tabular}
-\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig.
-\ref{fig:NKI}(b)).
-The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
-}
-\label{fig:NKIplots}
-\end{center}
-\end{figure}
-
-\subsubsection{Statistical Power Related to Gender-Based FA Differences}
-
-For completeness, the second NKI data subset was used to explore the 
-power of the different alignment strategies in discriminating FA 
-differences in a control group divided by gender.  As described 
-earlier, this group consisted of 27 subjects 
-comprised of 9 females (age: $29.3 \pm 6.6$ years) and 18 males (age: $26.6 
-\pm 6.9$ years).  Each subject was registered to the template made from 
-30 randomly chosen subjects, described in the previous section, and
-subsequently warped to the MNI152 template space.
-
-Results are shown in Fig. \ref{fig:nki_testing} in which all but the more
-robust metrics (MI and CC) exhibit similar power.  Note that the magnitude
-of the effects is much smaller than for TBI.  Again, the template-based approach
-performs comparatively well without the negative effects associated with
-the issues perviously described.
-
-\begin{figure*}
-\begin{center}
-\begin{tabular}{c}
-%  \includegraphics[width=180mm]{nkiTesting.pdf}
-  \includegraphics[width=130mm]{nkiTesting.pdf}
-\end{tabular}
-\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results 
-(one-tailed, $\alpha = 0.95$) for the NKI FA data aligned to the FMRIB58 template 
-adjusted for multiple comparisons (FDR).  
-Similar statistical power is seen in all alignment scenarios except for the more robust
-metrics (CC and MI) in the $t$-test which show decreased power.   }
-\label{fig:nki_testing}
-\end{center}        
-\end{figure*}
-
-%\section{Discussion}
-%This should explore the significance of the results of the work, not repeat them.
-A combined Results and Discussion section is often appropriate.
-Avoid extensive citations and discussion of published literature.
-
-\section{Discussion} 
-Prior evaluation work using anatomically labeled
-data indicates that mutual information, cross-correlation, Demons, and
-SSD metrics are all capable of producing high quality anatomical
-alignment (e.g. \citep{Klein2009}).  All of this work, though, compares
-the quality of alignment at a relatively coarse scale of major gyri,
-lobes and regions.  At the voxel level, it is nearly impossible to
-determine the ground truth correspondence.  Indeed, there is no reason
-to believe that subtle differences between a set of alignment
-strategies (induced, in the experiments of this paper, by changing the
-similarity metrics used in image registration) that all apparently
-work well should be considered significant.  This work, however,
-highlights a dramatic impact of similarity metric choice on detection
-power in template-based FA studies.  Our contention is that this
-dramatic ``improvement'' in detection power is not due to better
-anatomical alignment.  Rather, it is a result of the
-circularity/nonindependence of the normalization and statistical
-estimation strategy.  Consequently, we recommend that future work
-should take greater care when pairing the normalization strategy with
-the statistical framework.  Furthermore, we recommend that researchers
-should diligently seek to maintain independence between these two
-critical stages of a population study.  That is, the features that
-drive the minimization of a similarity metric should be independent of
-the features that will be used in hypothesis testing.  In particular,
-one should avoid combining the SSD metric for normalizing FA with
-Student's $t$-test for FA differences.  A final point is that
-cumulative distribution functions and other power assessment methods
-cannot be regarded as meaningful comparison techniques in the presence
-of normalization/statistical nonindependence.  Similar issues are
-raised in \cite{Kriegeskorte2010}.
-
-Caveats with our analysis include a lack of ground truth in the
-clinical component of the study.   Additionally, we are not addressing
-the biological plausibility of our results nor do we
-intend to (in this paper) enter into discussion on the mechanisms of
-TBI, gender difference and DTI that may lead to the detected results.
-Rather, we focus on the technical issue of circularity caused by
-normalizing the image feature of interest with functions that maximize
-the statistic used in hypothesis testing.  We also cannot argue that our analysis
-invalidates previous studies in which circularity might be an issue.  
-Finally, we do not address the very promising work concerning
-whole tensor DTI normalization \citep{Zhang2007,Hecke2007}, DTI template-based
-strategies \citep{Mori2009,Hecke2011}, {\color{red}{or other related methods \cite{jbabdi2010}}}.  Although investigation is beyond the scope
-of this work, these whole-tensor approaches might provide the necessary
-anatomical information to overcome the deficiencies addressed in this
-work.
-
-Instead, we encourage the community to
-reexamine FA normalization and analysis methods and to consider the
-potential confounds highlighted in this paper.  Rank tests
-provide a more conservative alternative as does the methodology
-presented here based on T1 to DT intrasubject mapping and T1
-normalization.  To our surprise, we could not find a prior publication that directly
-addresses this issue.  Two reasons may be that the majority of
-technical work using DTI has focused on more cutting-edge issues while
-the clinical work using DTI is more concerned with generating
-statistically significant results.  An additional reason may be that
-only recently have toolkits emerged to enable easy comparison of
-statistical techniques (via {\bf R} \citep{RDevelopmentCoreTeam2011}), normalization methods (ANTs), and
-diffusion tensor processing (Camino).
-
-\section{Conclusions}
-%The main conclusions of the study may be presented in a short Conclusions section, which may stand alone or form a subsection of a Discussion or Results and Discussion section.
-
-%The popularity of FSL's TBSS is due, in large part, to its ease of use.
-Anybody interested in performing a neuroimaging study in which DTI-derived measures are compared across populations can simply download FSL and immediately apply it to their data.
-The proposed work is a substantial improvement to the core TBSS framework which dovetails with FSL.
-In addition, it is comparably easy to download and apply what we have proposed in the form of ANTs-TBSS.
-
-With the increasing use of DTI for white matter assessment and the
-increasing number of researchers that are taking advantage of the numerous
-software tools that are available to facilitate this research, guidance
-is necessary to ensure the proper use of these tools and the
-calculation and reporting of findings \citep{Ridgway2008}.  The normalization 
-circularity issue described in this work is the inducement of substantial Type 1
-errors caused by FA-to-FA normalization using standard similarity metrics.
-To reduce these types of errors, we advocate the use of an anatomical template-building
-strategy whereby the transformation of FA images to a common space is performed
-using each individual's anatomical image.
-This proposed improvement potentially has many benefits:
-%which were listed in the 
-%introduction.  Such benefits range from the theoretical in the form of eliminating 
-%the confounding effects of registering data which is to be statistically 
-%differentiated to the practical, e.g. reusing the ANTs-based template for other 
-%neuroimaging analyses.  
-\begin{itemize}
-\item As explained in \cite{Avants2010}, ANTs optimal template construction creates a true Frech\'et mean from the population sample with high resolution which facilitates within-group registration.
-This obviates the need for an externally-derived template or an $\OO{n^2}$ search for an approximation to the mean subject or the bias resulting from random subject selection.
-\item Since the creation of the standardized coordinate system is performed using the anatomical images, nonrigid registration of the DTI-derived scalar images is 
-minimized thus mitigating the confounds described previously.
-Each DTI-based scalar image is rigidly registered to its corresponding T1 image followed by a small number of nonrigid registration iterations to account for echo planar imaging (EPI) distortion \citep{Wu2008}.
-\item Although the total number of registrations to be performed is increased, this strategy makes the composite registration between an individual DTI-derived image to the template more robust as the shape distance between registrations is minimized.
-%\item Instead of Gaussian smoothing as with SPM, TBSS uses the warping process to FMRIB58 by default to provide an indirect smoothing.
-Such resampling occurs in ANTs-TBSS as a by-product of the registration of the DTI-derived image to the corresponding T1 image in a subject-specific manner.
-%\item Instead of using thresholding to create the white matter segmentation from which the skeleton is created, an anatomically-based segmentation, either in the template space or in each subject space followed by a label fusion technique (e.g.
-majority voting), can be used to ultimately provide a more accurate skeleton.
-\item Oftentimes FA analysis is only one of several components performed in a population-based study.
-Other investigations, such as voxel- and tensor-based morphometry, cortical thickness measurements, etc.
-can take advantage of the anatomical template.
-\item The proposed modification is compatible with other frameworks such as TBSS where the normalization step is
-performed using ANTs-based tools followed by skeletonization and projection using the FMRIB software (see Fig. \ref{fig:workflow}).
-\end{itemize}
-
-All software components described 
-in this paper are publicly available as open source primarily through the 
-Advanced Normalization Tools (ANTs) repository.%
-\footnote{
-http://www.picsl.upenn.edu/ANTs
-}
-In addition to registration and template building, ANTs also includes brain 
-extraction \citep{Avants2010a}, bias correction \citep{Tustison2010}, $n$-tissue 
-segmentation \citep{Avants2011a}, and cortical thickness estimation solutions \citep{Das2009}.
-
-%In performing the various analyses described in the work, it was pleasantly surprising to discover how well the alignments were for the ANTs template data considering that alignment was based on FA-to-FA correspondence and no such image registrations were performed to align FA directly with FA as with TBSS.
-Furthermore, since TBSS is a local quantification method with increased statistical power through the use of the skeletonization and projection step, improved alignments can increase the regional sensitivity to local white matter differences particularly in the peripheral tracts.
-
 %% The Appendices part is started with the command \appendix;
@@
 
+%% main text
+
+\input{intro}
+
+\input{methods}
+
+\section{Evaluation}
+
+\section{Discussion}
+
+\section{Conclusions}
+
 %% References with bibTeX database:
@@
 \section*{Acknowledgments}
-All visualizations were performed using ITK-SNAP%
-\footnote{
-http://www.itksnap.org/
-}
-\citep{Yushkevich2006} and
-DTI-TK.%
-\footnote{
-http://www.nitrc.org/projects/dtitk/
-}
-We also gratefully acknowledge Dr. Niels van Strien of the Norwegian University of Science and Technology
-who assisted in packaging the template construction algorithm in the very useful script \verb#buildtemplateparallel.sh#
-which is publicly available in ANTs.
 

### 2012-06-15 KapowskiChronicles/paper.tex
Abba:  Fernando.

@@
 (ANTS) cortical thickness automated pipeline
-comprised entirely of well-vetted components developed as part of the ANTS 
+comprising well-vetted components developed as part of the ANTS 
 toolkit such as multivariate template construction,

### 2012-06-15 KapowskiChronicles/paper.tex
Abba:  Fernando.

@@
 (ANTS) cortical thickness automated pipeline
-comprising well-vetted components developed as part of the ANTS 
-toolkit such as multivariate template construction,
+comprising well-vetted components  such as SyGN (multivariate template construction),
 SyN (image registration), N4 (bias

### 2012-06-24 KapowskiChronicles/methods.tex
ENH:  Added template figure.
ENH:  Luis Miguel, Con tus besos.

@@
 
-The first 20 subjects were used to create a population-specific
+All 40 subjects were used to create a population-specific
 unbiased average template \cite{avants2010}.  The brain masks corresponding 
@@
 for each of the 16 subjects using Atropos.  Similar to the LPBA40
-data set, a NIREP template was created from 8 subjects and each the
+data set, a NIREP template was created from all 16 subjects and each the
 warped labels were used to create probabilistic estimates of the 
@@
 thickness measured using the ANTs pipeline.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=90mm]{Figures/template.pdf}
+  \caption{Sample multivariate template constructed from a subset of the IXI data (female, age 40--50).  Axial slices of five of the 37 total subjects from this cohort are shown. }
+  \label{fig:template}
+\end{figure}

### 2012-07-01 KapowskiChronicles/intro.tex
ENH: Brian's edits.

@@
 
-Neuroscientific investigations into cortical morphological changes/differences have illuminated
-interesting correlations with normal and pathological neurodevelopment in 
-addition to cognitive function.
+% Neuroscientific investigations into cortical morphological
+% changes/differences have illuminated interesting correlations with
+% normal and pathological neurodevelopment in  addition to cognitive function.
 Historically rooted in the meticulous work of von Economo \citep{economo2008},
-computational methods for modeling the
-cortical gray matter have resulted in numerous studies 
-exploring such relationships.  
-These analysis techniques
-have been instrumental in demonstrating 
-the presence of regional cortical abnormalities in such conditions as 
+imaging-based structural analysis of the brain now plays a fundamental role
+in identifying the relationship between cortical morphology, disease and cognition.
+Quantitative cortical measures are able to identify abnormalities in such conditions as 
 Huntington's disease \citep{rosas2002,rosas2005,selemon2004}, 
@@
 disorder \citep{shin2007}, ADHD \citep{almeida-montes2012}, obesity \citep{raji2010}, and heritable \citep{peterson2009}
-and elderly \citep{ballmaier2004} depression.  Further study 
-topics exploring correlations with cortical thickness include
-normal aging \citep{kochunov2011},
-gender differences \citep{luders2006a},
-anatomical asymmetries 
+and elderly \citep{ballmaier2004} depression.  Cortical thickness also
+varies normally as a function of age \citep{kochunov2011},
+gender \citep{luders2006a}, handedness
 \citep{luders2006,amunts2007}, intelligence \citep{shaw2006}, athletic
 ability \citep{wei2011}, musical ability \citep{bermudez2009,foster2010}, 
-white collar criminals \citep{raine2011}, and Tetris-playing
+tendency toward criminality \citep{raine2011}, and Tetris-playing
 ability in female adolescents \citep{haier2009}.  Additionally,
-recent studies have demonstrated possible functional 
+recent studies demonstrate possible functional 
 connectivity relationships using cortical thickness measures
 \cite{worsley2005,lerch2006,he2007}.
-Although these study findings
+Although these findings
 are subject to debate and interpretation \citep{gernsbacher2007}, 
@@
 
-Broadly characterized as surface mesh-based or volumetric \citep{scott2009,clarkson2011}, such 
-techniques have been introduced and extensively developed over 
-the last $10+$ years.  Representative of the former is the
+Computational methods for analyzing the cortex may be 
+broadly characterized as surface mesh-based or volumetric \citep{scott2009,clarkson2011}.  Representative of the former is the
 Freesurfer%
@@
 
-Image volumetric (or meshless) techniques are varied both in algorithmic terms as well as
+Image volumetric (or meshless) techniques vary both in algorithmic terms as well as
 the underlying definition of cortical thickness.  An early, foundational technique is the 
@@
 are difficult as the quantified distances 
-are not necessarily Euclidean.  Other volumetric algorithms have employed coupled
+are not necessarily Euclidean.  Other volumetric algorithms employ coupled
 level sets \citep{zeng1999}, model-free intelligent search strategies either normal to 
@@
 white and pial matter surfaces is used to propagate thickness values 
-through the cortical gray matter.
+through the cortical gray matter.  A unique benefit of DiReCT is that it
+naturally estimates the boundaries of buried sulci by employing a
+diffeomorphic constraint on the probabilistic estimate of the gray
+matter and cerebrospinal fluid interface.
 
-Despite the numerous techniques that have been proposed in the literature (of which
-only a fraction has been cited), several requisite processing components are 
-common to many of them which are prior to quantification of cortical thickness.
-These include inhomogeneity correction, skull stripping, and $n$-tissue segmentation 
-for differentiating the gray and white matters.  For post-processing statistical analysis 
-across populations, construction of population-specific unbiased templates
-is also  beneficial.
+Despite the variety of techniques for estimating cortical thickness
+from imaging data (of which
+only a fraction are cited), several common pre-processing components
+may be identified.
+The most fundamental of these include inhomogeneity correction, skull stripping, and $n$-tissue segmentation 
+for differentiating the gray and white matter.  For statistical analysis 
+across large populations, construction of population-specific unbiased templates
+is also beneficial \cite{}.
 In addition, some of these steps might include a crucial registration component, e.g. 
@@
 implemented solely by the authors of the evaluation (not the actual software developers) using 
-unspecified parameters making the comparisons less than ideal.
+unspecified parameters making the comparisons less than
+ideal. \textcolor{blue}{might also note the Klein 2011 evaluation that
+  raises the difficulty of comparing volumetric and surface-based
+  methods and the bias induced by the domain representation.}
 
-In this work we describe our cortical thickness software pipeline which 
-we have made open source as part of our Advanced Normalization Tools (ANTs) 
-repository.  This includes all the necessary preprocessing steps consisting
+In this work, we describe our C$++$ based cortical thickness pipeline
+which is freely available as part of the Advanced Normalization Tools
+(ANTs) software package.  This includes all the necessary preprocessing steps consisting
 of well-vetted previously published algorithms for bias correction \cite{tustison2010},
@@
 these pipeline components and provide a set of useful parameters
-which are used to analyze the publicly available IXI data set. This is 
-encapsulated in a well-documented shell script which is also available in ANTS. 
+which are employed to analyze the publicly available IXI data set. The
+full pipeline and parameter set is 
+encapsulated in a well-documented shell script which is also available in ANTs. 
 Furthermore, we provide all the derived image data and other processing scripts on 
-a github repository specifically meant for this publication.  This permits
-the set of results described in this work to be fully reproducible thus
-permitting other researchers to use a complete volumetric pipeline for measuring
-cortical thickness.
+a github repository specifically meant for this publication.  The
+availability of both the code and data permits
+the set of results described in this work to be fully reproducible.  This
+permits other researchers to contrast their own results against
+this baseline processing and to adapt the given volumetric pipeline for measuring
+cortical thickness in their own datasets.
 

### 2012-07-01 KapowskiChronicles/methods.tex
ENH: Brian's edits.

@@
   \item initial N4 bias correction on input anatomical MRI,
+    \textcolor{blue}{do you usually bias correct the raw image?  phil
+      had some problems with the automated (otsu) masking employed
+      therein .... esp with elderly brain}
   \item brain extraction using a hybrid segmentation/template-based strategy,
@@
 SyN pairwise registration \citep{avants2011} and an
-optimized sharpening/averaging of the template appearance, Symmetric Group
+optimized shape-based sharpening/averaging of the template appearance, Symmetric Group
 Normalization (SyGN) is a powerful framework for producing optimal population-specific
@@
 in the corresponding data section.
+
+\textcolor{blue}{what is the motivation of the multivar template here?
+  it would seem useful to add an FA component to the segmentation step
+  to aid in cortical / wm delineation which would motivate the
+  multivar template approach but might be difficult to establish that
+  this actually helps .... though we could compare both ways : univar
+  and multivar }
 
@@
 region.  Each of the three tissue masks undergo specific morphological operations which are then
-combined to create a brain extraction mask for use in the rest of the cortical thickness workflow.
+combined to create a brain extraction mask for use in the rest of the
+cortical thickness workflow.  \textcolor{blue}{does there need to be a
+  bit more technical detail here?  while you can refer to the script,
+  why perform these operations?  there are numerous references, most
+  germane probably some recent stuff from J Prince ( i think ) and the
+  freesurfer watershed approach which came from ... i forget ... maybe
+  one of the french groups.}
 
-A comparison using open access brain data with publicly available brain extraction algorithms 
-including  AFNI's \verb#3dIntracranial# \citep{ward1999}, FSL's \verb#BET2# \citep{smith2002}, 
-Freesurfer's \verb#mri_watershed# \citep{segonne2004}, and BrainSuite \citep{dogdas2005}  demonstrated
-that our combined registration/segmentation approach \citep{avants2010a} performs at the top level
-alongside BrainSuite (tuned) and FreeSurfer.
+A comparison using open access brain data with publicly available
+brain extraction algorithms including AFNI's \verb#3dIntracranial#
+\citep{ward1999}, FSL's \verb#BET2# \citep{smith2002}, Freesurfer's
+\verb#mri_watershed# \citep{segonne2004}, and BrainSuite
+\citep{dogdas2005} demonstrated that our combined
+registration/segmentation approach \citep{avants2010a} performs at the
+top level alongside BrainSuite (tuned) and FreeSurfer.
+\textcolor{blue}{ok you have the segonne ref here ... }
 
@@
 was introduced in \cite{das2009}, we have made several 
-improvements (perhaps the most significant being that this particular 
+improvements.  Perhaps the most significant advance is that this particular 
 ITK-compatible implementation has been significantly multi-threaded,

### 2012-07-01 KapowskiChronicles/paper.tex
ENH: Brian's edits.

@@
                 Philadelphia, PA}
-\address[label3]{Department of Circulation and Medical Imaging, Norwegian University of Science and Technology, Trondheim, Norway}
+\address[label3]{Department of Circulation and Medical Imaging,
+  Norwegian University of Science and Technology, Trondheim,
+  Norway---\textcolor{blue}{should be netherlands not sure of specifics}}
 
@@
 Advanced Normalization Tools
-(ANTS) cortical thickness automated pipeline
+(ANT\textcolor{blue}{s}) cortical thickness automated pipeline
 comprising well-vetted components such as SyGN (multivariate template construction),
@@
 Atropos ($n$-tissue segmentation), and DiReCT (cortical thickness) 
-all developed as part of the ANTS open science effort.  
-Complementing the open source aspect of ANTS we confirm 
+all developed as part of the ANT\textcolor{blue}{s} open science effort.  
+Complementing the open source aspect of ANT\textcolor{blue}{s} we confirm 
 previous findings of

### 2012-07-09 KapowskiChronicles/intro.tex
ENH:  More Brian's edits.

@@
 unspecified parameters making the comparisons less than
-ideal. \textcolor{blue}{might also note the Klein 2011 evaluation that
-  raises the difficulty of comparing volumetric and surface-based
-  methods and the bias induced by the domain representation.}
+ideal. Further complicating comparisons is distinct processing domains between
+volumetric and surface-based techniques and the potential for the introduction
+of bias \cite{klein2010}.
 

### 2012-07-09 KapowskiChronicles/paper.tex
ENH:  More Brian's edits.

--- 2012-06-15 KapowskiChronicles/paper.tex
@@
                 Philadelphia, PA}
-\address[label3]{Department of Circulation and Medical Imaging, Norwegian University of Science and Technology, Trondheim, Norway}
+\address[label3]{Department of Circulation and Medical Imaging,
+  Norwegian University of Science and Technology, Trondheim,
+  Norway}
 
@@
 Advanced Normalization Tools
-(ANTS) cortical thickness automated pipeline
+(ANTs) cortical thickness automated pipeline
 comprising well-vetted components such as SyGN (multivariate template construction),
@@
 Atropos ($n$-tissue segmentation), and DiReCT (cortical thickness) 
-all developed as part of the ANTS open science effort.  
-Complementing the open source aspect of ANTS we confirm 
+all developed as part of the ANTs open science effort.  
+Complementing the open source aspect of ANTs we confirm 
 previous findings of

### 2013-02-08 KapowskiChronicles/intro.tex
ENH:  Updating.

@@
 Historically rooted in the meticulous work of von Economo \citep{economo2008},
-imaging-based structural analysis of the brain now plays a fundamental role
+imaging-based structural analysis of the brain plays a fundamental role
 in identifying the relationship between cortical morphology, disease and cognition.
@@
 schizophrenia \citep{nesvag2008}, bipolar disorder \cite{lyoo2006}, Alzheimer's disease and frontotemporal
-dementia \citep{du2007,dickerson2009}, Parkinson's disease \citep{jubault2011},
+dementia \citep{du2007,dickerson2009}, Parkinson's disease \citep{jubault2011}, Williams syndrome \cite{thompson2005},
 multiple sclerosis \citep{ramasamy2009}, autism \citep{chung2005,hardan2006},
@@
 varies normally as a function of age \citep{kochunov2011},
-gender \citep{luders2006a}, handedness
+gender \citep{luders2006a}, untreated transsexuality \citep{zubiaurre-elorza2012},  handedness
 \citep{luders2006,amunts2007}, intelligence \citep{shaw2006}, athletic
@@
 method of \cite{jones2000} in which the inner and outer surface geometry is used to determine the
-solution to Laplace's equation in which thickness is measured by integrating along the 
+solution to Laplace's equation where thickness is measured by integrating along the 
 tangents of the resulting field lines spanning the boundary surfaces.  Subsequent contributions
@@
 non-crossing correspondence paths, direct correlative assessments with histology
-are difficult as the quantified distances 
+are potentially problematic as the quantified distances 
 are not necessarily Euclidean.  Other volumetric algorithms employ coupled
@@
 Most relevant to this work is the DiReCT (Diffeomorphic Registration-based 
-Cortical Thickness) algorithm proposed by \cite{das2009} in which a 
+Cortical Thickness) algorithm proposed by \cite{das2009} in which 
 a registration-based approach proposed where the derived diffeomorphic mapping between the 
@@
 across large populations, construction of population-specific unbiased templates
-is also beneficial \cite{}.
+is also beneficial \cite{evans2012}.
 In addition, some of these steps might include a crucial registration component, e.g. 
@@
 (specifically in terms of suggested parameters); both volumetric methods were 
-implemented solely by the authors of the evaluation (not the actual software developers) using 
+implemented solely by the authors of the evaluation (not the actual algorithmic developers) using 
 unspecified parameters making the comparisons less than
@@
 
-In this work, we describe our C$++$ based cortical thickness pipeline
+In this work, we describe our Insight Toolkit (ITK)-based cortical thickness pipeline
 which is freely available as part of the Advanced Normalization Tools
@@
 We also describe improvements made to the original DiReCT algorithm \cite{das2009}.
-Equally as important, we demonstrate how to coordinate
-these pipeline components and provide a set of useful parameters
+Equally as important, we provide explicit coordination between
+these pipeline components complete with a set of useful parameters
 which are employed to analyze the publicly available IXI data set. The
@@
 Furthermore, we provide all the derived image data and other processing scripts on 
-a github repository specifically meant for this publication.  The
+the NITRC repository specifically meant for this publication.  The
 availability of both the code and data permits
@@
 this baseline processing and to adapt the given volumetric pipeline for measuring
-cortical thickness in their own datasets.
+cortical thickness with their own datasets.
 

### 2013-02-08 KapowskiChronicles/methods.tex
ENH:  More Brian's edits.
ENH:  Updating.

@@
   \item initial N4 bias correction on input anatomical MRI,
-    \textcolor{blue}{do you usually bias correct the raw image?  phil
-      had some problems with the automated (otsu) masking employed
-      therein .... esp with elderly brain}
   \item brain extraction using a hybrid segmentation/template-based strategy,
@@
   \centering
-  \includegraphics[width=130mm]{Figures/Kapowski_pipeline.pdf}
+  \includegraphics[width=170mm]{Figures/Kapowski_pipeline2.pdf}
   \caption{The ANTs T1 processing workflow containing all elements for 
@@
 16 (8 male and 8 female) high resolution skull-stripped brain 
-data with 32 cortical labels manually drawn using published protocol.
+data with 32 cortical labels (cf Table \ref{table:nirep_labels}) manually drawn using a 
+published protocol.%
+\footnote{
+http://www.nirep.org/
+}
 Given the gray matter labels, the white matter and CSF were identified 
 for each of the 16 subjects using Atropos.  Similar to the LPBA40
-data set, a NIREP template was created from all 16 subjects and each the
+data set, a NIREP template was created from all 16 subjects and each of the
 warped labels were used to create probabilistic estimates of the 
@@
 
+\begin{table}
+\centering
+\begin{tabular*}{0.475\textwidth}{@{\extracolsep{\fill}} l l}
+\toprule
+  1) L occipital lobe & 2) R occipital lobe \\
+  3) L cingulate gyrus & 4) R cingulate gyrus \\
+  5) L insula gyrus & 6) R insula gyrus \\
+  7) L temporal lobe & 8) R temporal lobe \\
+  9) L superior temporal gyr. & 10) R superior temporal gyr. \\
+  11) L infero temporal region & 12) R infero temporal region \\
+  13) L parahippocampal gyr. & 14) R parahippocampal gyr. \\
+  15) L frontal pole & 16) R frontal pole \\
+  17) L superior frontal gyrus & 18) R superior frontal gyrus \\
+  19) L middle frontal gyrus & 20) R middle frontal gyrus \\
+  21) L inferior gyrus & 22) R inferior gyrus \\
+  23) L orbital frontal gyrus & 24) R orbital frontal gyrus \\
+  25) L precentral gyrus & 26) R precentral gyrus \\
+  27) L superior parietal lobule & 28) R superior parietal lobule \\
+  29) L inferior parietal lobule & 30) R inferior parietal lobule \\
+  31) L postcentral gyrus & 32)   R postcentral gyrus \\  
+\bottomrule
+\end{tabular*}
+\caption{The 32 cortical NIREP labels.}
+\label{table:nirep_labels}
+\end{table}
+
 \subsection{IXI Data for Pipeline Evaluation}
@@
 }
-consists of approximately 600 healthy subjects imaged at three sites 
+used for the evaluation consists of 577 healthy subjects imaged at three sites 
 using several modalities (T1-weighted, T2-weighted, proton density, magnetic 
@@
   \centering
-  \includegraphics[width=90mm]{Figures/template.pdf}
-  \caption{Sample multivariate template constructed from a subset of the IXI data (female, age 40--50).  Axial slices of five of the 37 total subjects from this cohort are shown. }
+  \includegraphics[width=90mm]{Figures/template50_60.pdf}
+  \caption{Sample multivariate template constructed from a subset of the IXI data (female, age 50--60).  Axial slices of five of the 37 total subjects from this cohort are shown. }
   \label{fig:template}

### 2013-02-08 KapowskiChronicles/paper.tex
ENH:  Updating.

@@
 \usepackage{color}
+\usepackage{rccol}
+\usepackage[table]{xcolor}
 
@@
 
-\graphicspath{{./Figures/}
+\graphicspath{
+             {./Figures/}
+             {../SPIE2013/KK/IXI_results/} 
              }
@@
 structure and brain development, cognitive 
-function, and functional connectivity.
+function and functional connectivity.
 The highly convoluted cortical topography makes manual measurements
@@
 
-\section{Evaluation}
+\input{results}
 
-\section{Discussion}
-
-\section{Conclusions}
+\section{Discussion and Conclusions}
 

### 2013-02-08 SPIE2013/KK/Kirby_results/presentation.tex
ENH:  Updating.

\documentclass[xcolor=dvipsnames,serif,10pt]{beamer}

\usepackage{beamerthemesplit}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
%\usefonttheme{professionalfonts}
%\usepackage{times}
\usepackage{tikz}
\usepackage{amsmath,cancel}
\usepackage{mathtools}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{listings}
\usepackage[]{geometry}
\usepackage{ifxetex}
\ifxetex{%
  \usepackage{fontspec}
  \setmainfont{Linux Libertine O} % or any font on your system
  \newfontfamily\quotefont[Ligatures=TeX]{Linux Libertine O} % or any font on your system
\else
  \usepackage[T1]{fontenc}
  \usepackage{libertine} % or any other font package (or none)
  \newcommand*\quotefont{\fontfamily{fxl}} % selects Libertine for quote font
\fi
\usepackage{framed}

\usepackage{soul}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{calc,shapes.callouts,shapes.arrows}
\makeatletter

\newcommand{\defhighlighter}[3][]{%
  \tikzset{every highlighter/.style={color=#2, fill opacity=#3, #1}}%
}

\defhighlighter{yellow}{.5}

\newcommand{\highlight@DoHighlight}{
  \fill [ decoration = {random steps, amplitude=1pt, segment length=15pt}
        , outer sep = -15pt, inner sep = 0pt, decorate
        , every highlighter, this highlighter ]
        ($(begin highlight)+(0,8pt)$) rectangle ($(end highlight)+(0,-3pt)$) ;
}

\newcommand{\highlight@BeginHighlight}{
  \coordinate (begin highlight) at (0,0) ;
}

\newcommand{\highlight@EndHighlight}{
  \coordinate (end highlight) at (0,0) ;
}

\newdimen\highlight@previous
\newdimen\highlight@current

\DeclareRobustCommand*\highlight[1][]{%
  \tikzset{this highlighter/.style={#1}}%
  \SOUL@setup
  %
  \def\SOUL@preamble{%
    \begin{tikzpicture}[overlay, remember picture]
      \highlight@BeginHighlight
      \highlight@EndHighlight
    \end{tikzpicture}%
  }%
  %
  \def\SOUL@postamble{%
    \begin{tikzpicture}[overlay, remember picture]
      \highlight@EndHighlight
      \highlight@DoHighlight
    \end{tikzpicture}%
  }%
  %
  \def\SOUL@everyhyphen{%
    \discretionary{%
      \SOUL@setkern\SOUL@hyphkern
      \SOUL@sethyphenchar
      \tikz[overlay, remember picture] \highlight@EndHighlight ;%
    }{%
    }{%
      \SOUL@setkern\SOUL@charkern
    }%
  }%
  %
  \def\SOUL@everyexhyphen##1{%
    \SOUL@setkern\SOUL@hyphkern
    \hbox{##1}%
    \discretionary{%
      \tikz[overlay, remember picture] \highlight@EndHighlight ;%
    }{%
    }{%
      \SOUL@setkern\SOUL@charkern
    }%
  }%
  %
  \def\SOUL@everysyllable{%
    \begin{tikzpicture}[overlay, remember picture]
      \path let \p0 = (begin highlight), \p1 = (0,0) in \pgfextra
        \global\highlight@previous=\y0
        \global\highlight@current =\y1
      \endpgfextra (0,0) ;
      \ifdim\highlight@current < \highlight@previous
        \highlight@DoHighlight
        \highlight@BeginHighlight
      \fi
    \end{tikzpicture}%
    \the\SOUL@syllable
    \tikz[overlay, remember picture] \highlight@EndHighlight ;%
  }%
  \SOUL@
}
\makeatother

% wrap everything in its own environment
\newenvironment{shadequote}%
{\begin{quote}\openquote}
{\hfill\closequote\end{quote}}

\newcommand{\arrowthis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};
        \node[overlay,single arrow,draw=none,fill=red!50,anchor=tip,rotate=60] 
        at (#1.south) {#2};}%
    }%

\newcommand{\speechthis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};\node[overlay,ellipse callout,fill=blue!50] 
        at ($(#1.north)+(-.5cm,0.8cm)$) {#2};}%
    }%

\newcommand{\bubblethis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};\node[overlay,cloud callout,callout relative pointer={(0.2cm,-0.7cm)},%
        aspect=2.5,fill=yellow!90] at ($(#1.north)+(-0.5cm,1.6cm)$) {#2};}%
    }%

\newcommand{\pointthis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};\node[overlay,rectangle callout,%
        callout relative pointer={(0.2cm,0.7cm)},fill=green!50] at ($(#1.north)+(-.5cm,-1.4cm)$) {#2};}%
        }%

\usetikzlibrary{mindmap,backgrounds}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.9}
    \definecolor{listwhite}{gray}{1.0}

\usetheme[secheader]{Boadilla}
\useoutertheme{miniframes}
\useinnertheme{circles}
\usecolortheme{myct}
\usepackage[latin1]{inputenc}

\graphicspath{{Figures/}{../Figures/}}

\setlength{\tabcolsep}{1pt}

\newcommand\fontvi{\fontsize{6}{8}\selectfont}
\newcommand{\ColIndent}{\hspace{\labelwidth}\hspace{\labelsep}\hspace{\labelsep}\hspace{\labelsep}\hspace{\labelsep}}

\usetikzlibrary{backgrounds}
\usetikzlibrary{positioning}
\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Title slide
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% -My goal is to not get you excited about {\em my} research and the research I'm doing with
%  my collaborators.  Rather I want to get you excited about resources here at UVa for doing
%  {\em your} research, particularly if it involves computational image analysis and moderate
%  to large sized cohorts

\title[Big$\times$(Science $+$ Imaging) at UVA]{$\text{Big}\times(\text{Science}+\underbracket{\cancel{\text{Data}}}_{\small \text{Imaging}})$ @ UVA}
\author[N. Tustison]{%
  Nick Tustison, DSc
  }
\date[SOM Research Retreat]{}
\institute[ntustison@virginia.edu]{}

%\pgfdeclaremask{fsu}{fsu_logo_ybkgrd}
%\pgfdeclareimage[width=1cm]{logo}{ants_logo}
%\logo{\vbox{\vskip0.1cm\hbox{\pgfuseimage{logo}}}}

%\logo{
%  \includegraphics[height=0.8cm]{brain2.png}
%  \includegraphics[height=0.8cm]{itkLogo.jpg}
%  }

\begin{document}

\tikzstyle{every picture}+=[remember picture]

\frame{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  * PNAS VBM study of London cab drivers.
%%  * 465 data points age ~ 1 + cortical_atrophy
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% (Transition)  Let's start off with an interesting study that was done about 10 years ago.

% These computational techniques have proven invaluable in exploring fundamental, and perhaps 
% not so fundamental, neuroscience questions.

\footnotesize

\begin{frame}{Hippocampal growth in London taxi drivers}

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
    \begin{center}  
      \includegraphics[width=5cm]{London_taxi.jpg}
    \end{center}
    \end{column}  
    \begin{column}{0.4\textwidth}
    \begin{center}  
      \vspace{-0.5cm}
      \pause
      \includegraphics[width=5cm]{hippocampus.png} \\
      \pause
        \begin{align*}
			  SPM (1995):\,\,&   6418\,\,\text{citations}\\
			  VBM (2000):\,\,&   3593\,\,\text{citations}\\
			  Opt. VBM (2001):\,\,&  2353\,\,\text{citations}\\
			  TBSS (2006):\,\,&  1106\,\,\text{citations}\\
			  \end{align*}
      \vspace{0cm}
			  \highlight[yellow!100!black]{$6418 + 3593 + 2353 + 1106 = 13470\,\,\text{citations}$}
    \end{center}
    \end{column}  
  \end{columns}    
\end{frame}

\normalsize

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Studies at UVA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% (Transition)  Let's fast forward about 10 years to a young researcher at UVa who.
% Needless to say it was an interesting two year journey which finally ended in November
% when our paper was finally published in Human Brain Mapping which shows how these 
% techniques are susceptible to a methodological flaw known as logical circularity.

{
\usebackgroundtemplate{\includegraphics[width=1.175\paperwidth]{breachers.jpg}}%
\begin{frame}{TBI in breachers---{\em neuroimaging correlates?}}
   \begin{center}
   \vspace{0cm}
  \pause
   \color{black}{Investigate white matter injury \\
   {\scriptsize controls/students/instructors}} \\
   \vspace{1cm}
  \pause
   \color{black}{\em Logical circularity in\\ SPM, VBM, \& TBSS} \\
   \end{center}
\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Intro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Make commands for the quotes
\newcommand*{\openquote}{\tikz[remember picture,overlay,xshift=-20pt,yshift=-5pt]
     \node (OQ) {\quotefont\fontsize{60}{60}\selectfont``};\kern0pt}
\newcommand*{\closequote}{\tikz[remember picture,overlay,xshift=-30pt,yshift=-10pt]
     \node (CQ) {\quotefont\fontsize{60}{60}\selectfont''};}
% select a colour for the shading

\begin{frame}{}
  \begin{center}

  \begin{columns}[c]
    \begin{column}{0.6\textwidth}
	    \begin{center}  
					\begin{shadequote}
					{\large {\em Maximum efficiency with minimum} wasted {\em effort}}.\par{Jigaro Kano$'$}
					\end{shadequote}
			\end{center}		
    \end{column}
    \hspace{-2.5cm}
    \begin{column}{0.2\textwidth}
		  \begin{center}
		    \includegraphics[height=4cm]{Kano_Jigoro.png}
		  \end{center}
		\end{column}  
  \end{columns}		
  \vspace{1cm}
  \pause
  {\LARGE \em There are \highlight[green, draw=blue]{life-altering} and \highlight[red!100!black]{free} UVA resources that you are probably not using!}
  \end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  What is data analysis? (neuroscience highlighted)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\tikzset{level 1 concept/.append style={font=\sf, sibling angle=135,level distance = 25mm}}
\tikzset{level 2 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
\tikzset{level 3 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
\tikzset{every node/.append style={scale=0.5}}

\begin{center}
\resizebox{9cm}{7.5cm}{
\begin{tikzpicture}
  \path[mindmap,concept color=red!50!black,text=white]
    node[concept,scale=0.9] {\Large Data Analysis}
    [clockwise from=45]
    child[concept color=green!50!black, yshift=-0.5cm] {
      node[concept,scale=1.1] {Imaging Applications}
      [clockwise from=120]
      child[concept color=green!80!black] { node[concept] (ns) {Neuro\-science} }
      child { node[concept] {Func. Lung Imaging} }
      child { node[concept] {Breast} }
      child { node[concept] {Cardiac} }
    }  
    child[concept color=blue!80!black, yshift=0.25cm] {
      node[concept,scale=1.1] (stat) {Statistics}
      [clockwise from=-60]
      child { node[concept] {Linear Modeling} }
      child { node[concept] (ml) {Machine Learning} }
    }
    child[concept color=orange, yshift=-0.5cm] { 
      node[concept,scale=1.1] {Computer Science} 
        [clockwise from=-145]
        child { node[concept] (ad) {Algorithm Development}
           [clockwise from=-60]
           child { node[concept] (antsr) {\footnotesize ANTsR} }
           child { node[concept] {ANTs} }
           child { node[concept] {ITK} }
           }
        child { node[concept] (cp) {Computer Programming} }
        child { node[concept] (cc) {Cluster Computing} }
    };

		\newcommand{\connbluetoorange}{to[circle connection bar switch color=from (blue!80!black) to (orange)]}
		\newcommand{\connorangetoorange}{to[circle connection bar switch color=from (orange) to (orange)]}
		\newcommand{\conngreentoorange}{to[circle connection bar switch color=from (green!80!black) to (orange)]}

  \begin{pgfonlayer}{background}
      \path (ad) \connorangetoorange (cp);
      \path (stat) \connbluetoorange (antsr);
      \path (ns) \conngreentoorange (cc);
  \end{pgfonlayer}

%  \begin{pgfonlayer}{background}
%      \path (ml) \connbluetoorange (cs);
%  \end{pgfonlayer}

\end{tikzpicture}
}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Wintermark studies
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{}
\setbeamercovered{invisible}

\tikzstyle{mybox} = [draw=black, fill=bottomcolour, thick,
    rectangle, rounded corners, inner sep=10pt, inner ysep=10pt]
\tikzstyle{mybox2} = [draw=black, fill=middlecolour, thick,
    rectangle, rounded corners, inner sep=5pt, inner ysep=5pt, 
    minimum height=4cm]
\tikzstyle{fancytitle} =[fill=bottomcolour, text=textcolour, rectangle]
\tikzstyle{line} = [-latex',very thick,textcolour]

%\begin{center}
%\begin{tikzpicture}
%};
%\end{tikzpicture}%
%\end{center}
%\vspace{5mm}

\begin{center}
\begin{tikzpicture}[node distance = 4cm]

\node [shift={(4cm,4cm)},mybox] (max){%
  Misc. neuroimaging projects};

\uncover<1->{
\node [mybox2, text width = 3.5cm, opacity = 1] (SkullThickness) {%  
  \begin{center}
  \begin{tabular}{c}
  \includegraphics[height=2cm]{skullThickness.png} 
  \vspace{0.25cm}
  \end{tabular}
  \scriptsize{Surrogate CT Skull Thickness Measurements with MRI}
  \end{center}
  };
\node[fancytitle] (SkullThickness_title) at (SkullThickness.north) {\scriptsize Skull Thickness};
\path[line] (max.south) edge [out= 230, in= 90] (SkullThickness_title.north);
}

\uncover<2->{
\node [mybox2, text width = 3.5cm, right of=SkullThickness] (ET) {%
  \begin{center}
  \begin{tabular}{c}
  \includegraphics[height=1.8cm]{ET.png} 
  \vspace{0.25cm}
  \end{tabular}
  \scriptsize{Longitudinal FA correlations following surgical treatment for ET}
  \end{center}
  };
\node[fancytitle](ET_title) at (ET.north) {\scriptsize Essential Tremor};
 \path[line] (max.south) edge [out= 270, in= 90] (ET_title.north);
 }

\uncover<3->{
\node [mybox2, text width = 3.5cm, right of=ET] (ASL) {%
  \begin{center}
  \begin{tabular}{c}
  \includegraphics[height=2.75cm]{ASL.png} 
  \vspace{0.15cm}
  \end{tabular}
  \scriptsize{Automatic generation of clinical CBF maps from ASL}
  \end{center}
  };
\node[fancytitle](ASL_title) at (ASL.north) {\scriptsize ASL$\rightarrow$CBF};
\path[line] (max.south) edge [out= 310, in= 90] (ASL_title.north);
}

\end{tikzpicture}

%\begin{tikzpicture}
%  \tikz[overlay]\path[line] (max.south) edge [out= 230, in= 90] (SkullThickness_title.north);
%  \tikz[overlay]\path[line] (max.south) edge [out= 270, in= 90] (ET_title.north);
%  \tikz[overlay]\path[line] (max.south) edge [out= 310, in= 90] (ASL_title.north);
%\end{tikzpicture}
\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  ANTs pipeline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{ANTs pipeline for cortical thickness estimation}

\tikzstyle{format} = [draw, thick, fill=middlecolour]
\tikzstyle{format2} = [draw, thin, fill=middlecolour]
\tikzstyle{medium} = [ellipse, draw, thin, fill=green!20, minimum height=2.5em]
\tikzstyle{line} = [-latex',very thick,textcolour]

\usetikzlibrary{calc,decorations.pathmorphing,patterns}
\pgfdeclaredecoration{penciline}{initial}{
    \state{initial}[width=+\pgfdecoratedinputsegmentremainingdistance,
    auto corner on length=1mm,]{
        \pgfpathcurveto%
        {% From
            \pgfqpoint{\pgfdecoratedinputsegmentremainingdistance}
                      {\pgfdecorationsegmentamplitude}
        }
        {%  Control 1
        \pgfmathrand
        \pgfpointadd{\pgfqpoint{\pgfdecoratedinputsegmentremainingdistance}{0pt}}
                    {\pgfqpoint{-\pgfdecorationsegmentaspect
                     \pgfdecoratedinputsegmentremainingdistance}%
                               {\pgfmathresult\pgfdecorationsegmentamplitude}
                    }
        }
        {%TO 
        \pgfpointadd{\pgfpointdecoratedinputsegmentlast}{\pgfpoint{1pt}{1pt}}
        }
    }
    \state{final}{}
}

\begin{center}
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick,decoration=penciline, decorate]
    % We need to set at bounding box first. Otherwise the diagram
    % will change position for each frame.
    \path[use as bounding box] (-1,0) rectangle (10,0);

\uncover<1->{
    \path[->,rounded corners=0.1cm] node[format, yshift=2cm, text width=1.5cm ] (ixi) { 
        \begin{center}
			  \includegraphics[height=1.75cm]{IXI221.png} \\
        \tiny{Input MRI}
        \end{center}
			  };
    \path[->] node[format, right of=ixi, xshift=0.5cm, text width=1.25cm] (n4) {
          \begin{center}
            \tiny{Initial N4 bias correction}
          \end{center}
        };
    \draw[->] (ixi.0) --  (n4.180);    
    \path[->] node[format, below of=n4, yshift = 1.25cm, text width=1.25cm] (be) {
          \begin{center}
            \tiny{Brain extraction}
          \end{center}
        };
    \draw[->] (n4.-90) -- (be.90);    
    \path[->] node[format, below of=be, yshift = 1.25cm, text width=1.25cm] (n4_2) {
          \begin{center}
            \tiny{N4 using WM estimate}
          \end{center}
        };
    \draw[->] (be.-90) -- (n4_2.90);    
    \path[->] node[format, below of=n4_2, yshift=1.25cm, text width=1.25cm] (atropos) {
          \begin{center}
            \tiny{3-tissue segmentation}
          \end{center}
        };
    \draw[->] (n4_2.-90) -- (atropos.90);    
    \path[->] node[diamond, draw, thick, fill=middlecolour, left of=atropos, xshift=1.75cm, yshift=0.875cm, text width=0.5cm] (conv) {
          \begin{center}
            \tiny{Conv.?}
          \end{center}
        };
    \draw[->] (atropos.180) -| (conv.270);    
    \draw[->] (conv.90) |- node {\tiny no} (n4_2.180);

    \path[->] node[format, left of=atropos, xshift=-0.5cm, text width=1.5cm ] (kk) {
        \begin{center}
        \tiny{Cortical thickness estimation}
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[format, above of=kk, yshift=-1cm, text width=1.5cm ] (kkmap) {
        \begin{center}
			  \includegraphics[height=1.75cm]{IXI221_direct.png} \\
        \tiny{Thickness map}
        \end{center}
			  };

    \draw[->] (conv.180) -- node {\tiny yes} ([yshift=0cm,xshift=-0.45cm]conv.180) |- (kk.0);    
    \draw[<-] (kkmap.270) -- (kk.90);    
}

\uncover<2->{
    \path[->,rounded corners=0.1cm] 
      node[format, right of=n4, xshift=1.15cm, yshift=-2.5cm,text width=4.75cm] (template) { 
        \begin{center}
        \tiny{Oasis template (35 normals)}
        \begin{tabular}{cc}
          \includegraphics[width=1.5cm]{MICCAI_T1.png} &
          \includegraphics[width=1.5cm]{MICCAI_probMask.png} \\
          \tiny{T1} & \tiny{Whole brain} 
        \end{tabular} \\
        \begin{tabular}{ccc}
          \includegraphics[width=1.5cm]{MICCAI_probCSF.png} &
          \includegraphics[width=1.5cm]{MICCAI_probGM.png} &
          \includegraphics[width=1.5cm]{MICCAI_probWM.png} \\
          \tiny{CSF} & \tiny{GM} & \tiny{WM}
        \end{tabular} \\
        \end{center}
			  };
    \draw[->,dashed] (template.160) -- ([yshift=0cm,xshift=-0.5cm]template.160) |- (be.0); 
    \draw[->,dashed] (template.200) -- ([yshift=0cm,xshift=-0.5cm]template.200) |- (atropos.0); 
}

\uncover<3->{
  \node[decorate,draw,minimum height=7.25cm,minimum width=5.65cm,dashed,rounded corners=0.2cm, thin, color = yellow] (c) at (1.7,-0.35) {};
}  
\uncover<4->{
  \node[decorate,draw,minimum height=5.25cm,minimum width=5.55cm,dashed,rounded corners=0.2cm, thin, color = red] (c) at (7.7,-0.45) {};
}
\end{tikzpicture}

\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  IXI Thickness
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{$\texttt{thickness} \sim 1 + \texttt{age} + \texttt{I(age\^{}2)}$}

\tikzstyle{format} = []

\begin{center}
\begin{tikzpicture}[node distance=4cm, auto,>=latex', thin]
    % We need to set at bounding box first. Otherwise the diagram
    % will change position for each frame.
    \path[use as bounding box] (-1,0) rectangle (10,0);

\uncover<1->{    
    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm,xshift=1.5cm,yshift=-0.5cm ] (nirep) { 
        \begin{center}
			  \includegraphics[width=5.5cm]{NIREPLabels.png} \\
        \end{center}
			  };
}			  
\uncover<2->{    
    \path[->,rounded corners=0.1cm] 
      node[format, right of=nirep, xshift=2.25cm, yshift=1.85cm,text width=4.5cm] (rp) {
        \begin{center}
			  \includegraphics[width=4.5cm]{yylabel26_results.pdf}
        \end{center}
			  };
    \path[->,thick, color = yellow] (rp.west) edge [out= 180, in= 90] (1.25,0);			  
}			  
\uncover<3->{    
    \path[->,rounded corners=0.1cm] 
      node[format, right of=nirep, xshift=2.25cm, yshift=-1.85cm,text width=4.5cm] (rt) {
        \begin{center}
			  \includegraphics[width=4.5cm]{yylabel8_results.pdf} \\
        \end{center}
			  };
    \path[->,thick, color = yellow] (rt.west) edge [out= 180, in= 270] (1.4,-2.);
}			  
\uncover<4->{    
    \path[->,rounded corners=0.1cm] node[format,below of=nirep, text width = 4.0cm, yshift=1cm ] (wow) { 
        \begin{center}    
        {\color{yellow}{\em It takes $\sim$20--30 hours per subject!!}}
        \end{center}    
			  };
}			  
\end{tikzpicture}

%\begin{tikzpicture}
%  \tikz[overlay]\path[->,thick, color = yellow] (rp.west) edge [out= 180, in= 90] (-2.6,5);
%  \tikz[overlay]\path[->,thick, color = yellow] (rt.west) edge [out= 180, in= 270] (-2.5,2.75);
%\end{tikzpicture}

\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Open source software
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% - The visible human project (1999)---one of the earliest big imaging data/big science challenges.
%

\begin{frame}{Big data $\rightarrow$ open source software}

\begin{center}
\begin{tikzpicture}[node distance=3.5cm, auto,>=latex', thin]
    % We need to set at bounding box first. Otherwise the diagram
    % will change position for each frame.
    \path[use as bounding box] (-1,0) rectangle (10,0);

\uncover<1->{
    \path[->,rounded corners=0.1cm] node[text width=5cm,xshift = 1cm] (nlm) { 
        \begin{center}
			  \includegraphics[width=4.755cm]{visible_human_project.png} \\
        \end{center}
			  };
}			  
\uncover<2->{
    \path[->,rounded corners=0.1cm] 
        node[text width=6cm, right of=nlm, xshift=3cm, yshift=1.25cm] (logos) { 
        \begin{center}
			  \includegraphics[width=6cm]{logos.png} \\
        \end{center}
			  };
}			  
\uncover<3->{
    \path[->,rounded corners=0.1cm] node[text width=3cm, yshift=2cm, below of=logos] (itk) { 
        \begin{center}
			  \includegraphics[width=3.cm]{ITK.png} \\
        \end{center}
			  };
}			  
\uncover<4->{
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift=1.2cm, below of=itk] (ants) { 
        \begin{center}
			  \includegraphics[width=2.1cm]{ANTsLogo.png} \\
        \end{center}
			  };
}			  
\uncover<5->{
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift=1.2cm, below of=itk] (ants) { 
        \begin{center}
			  \includegraphics[width=2.1cm]{ANTsLogo.png} \\
        \end{center}
			  };
}

\end{tikzpicture}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  UVACSE/Linux cluster
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{UVACSE/computational cluster}

\begin{center}
\begin{tikzpicture}[node distance=2.5cm, auto,>=latex', thin]
    % We need to set at bounding box first. Otherwise the diagram
    % will change position for each frame.
    \path[use as bounding box] (-1,0) rectangle (10,0);

\uncover<1->{			  
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift = 2cm] (r16) { 
        \begin{center}
			  \includegraphics[width=2cm]{r16slice.png} \\
			  \vspace{-0.5cm}
			  {\footnotesize \texttt{387\_oasis}}
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, right of=r16] (r27) { 
        \begin{center}
			  \includegraphics[width=2cm]{r27slice.png} \\
			  \vspace{-0.5cm}
			  {\footnotesize \texttt{392\_oasis}}
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of=r16] (r30) { 
        \begin{center}
			  \includegraphics[width=2cm]{r30slice.png} \\
			  \vspace{-0.5cm}
			  {\footnotesize \texttt{406\_oasis}}
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of=r27] (r62) { 
        \begin{center}
			  \includegraphics[width=2cm]{r62slice.png} \\
			  \vspace{-0.5cm}
			  {\footnotesize \texttt{411\_oasis}}
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of = r30] (r64) { 
        \begin{center}
			  \includegraphics[width=2cm]{r64slice.png} \\
			  \vspace{-0.5cm}
			  {\footnotesize \texttt{425\_oasis}}
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of=r62] (r85) { 
        \begin{center}
			  \includegraphics[width=2cm]{r85slice.png} \\
			  \vspace{-0.5cm}
			  {\footnotesize \texttt{430\_oasis}}
        \end{center}
			  };
}			  
\uncover<2->{			  
    \path[->,rounded corners=0.1cm] node[text width=6cm, xshift=2cm,right of=r62] (qstat) { 
        \begin{center}
			  \includegraphics[width=5.75cm]{qstat.png} \\
        \end{center}
			  };		
}

\uncover<3->{			  
  \path[->,thick, color = yellow, dashed] (0.5,2.3) edge [out=35,in= 145] (4.5,2.3);
  \path[->,thick, color = yellow, dashed] (3.1,1.55) edge [out=-35,in=-145] (4.5,1.55);
  \path[->,thick, color = yellow, dashed] (0.5,-0.25) edge [out=35,in= 145] (4.5,-0.25);
  \path[->,thick, color = yellow, dashed] (3.1,-0.9) edge [out=-35,in=-145] (4.5,-0.9);
  \path[->,thick, color = yellow, dashed] (0.5,-2.8) edge [out=35,in= 145] (4.5,-2.8);
}  
\uncover<4->{			  
  \path[->,thick, color = red, dashed] (3.1,-3.45) edge [out=-35,in=-145] (4.5,-3.45);
}

\end{tikzpicture}

%\begin{tikzpicture}
%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-4,2.75) edge [out=35,in= 145] (0,2.75);
%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-1.4,2.1) edge [out=-35,in=-145] (0,2.1);
%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-4,0.2) edge [out=35,in= 145] (0,0.2);
%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-1.4,-0.45) edge [out=-35,in=-145] (0,-0.45);
%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-4,-2.35) edge [out=35,in= 145] (0,-2.35);
%  \tikz[overlay]\path[->,thick, color = red, dashed] (-1.4,-3.0) edge [out=-35,in=-145] (0,-3.0);
%\end{tikzpicture}

\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Closing slide
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Further information}

\begin{tikzpicture}[node distance=1cm, auto,>=latex', thin]
    % We need to set at bounding box first. Otherwise the diagram
    % will change position for each frame.
    \path[->,rounded corners=0.1cm] node[text width=2cm,xshift=-2cm] (ants) { 
        \begin{center}
			  \includegraphics[width=3.5cm]{ants2.png} \\
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=6cm, xshift=5cm, right of=ants] { 
        \begin{center}
        \begin{itemize}
          \item Insight Toolkit:  http://www.itk.org
          \item ANTs/ANTsR:  http://www.picsl.upenn.edu/ANTs/
          \item UVACSE:  http://www.uvacse.virginia.edu
        \end{itemize}
        \end{center}
			  };

\end{tikzpicture}
\end{frame}

\end{document}

### 2013-02-08 SPIE2013/KK/Presentation/presentation.tex
ENH:  Updating.

--- 2013-02-08 SPIE2013/KK/Kirby_results/presentation.tex
@@
 \documentclass[xcolor=dvipsnames,serif,10pt]{beamer}
+\usepackage{etex}
 
@@
 %\usepackage{times}
+\usepackage{rotating}
 \usepackage{tikz}
@@
 \usetikzlibrary{decorations.pathmorphing}
+\usetikzlibrary{decorations.text}
 \usetikzlibrary{calc,shapes.callouts,shapes.arrows}
@@
 
-% -My goal is to not get you excited about {\em my} research and the research I'm doing with
-%  my collaborators.  Rather I want to get you excited about resources here at UVa for doing
-%  {\em your} research, particularly if it involves computational image analysis and moderate
-%  to large sized cohorts
-
-\title[Big$\times$(Science $+$ Imaging) at UVA]{$\text{Big}\times(\text{Science}+\underbracket{\cancel{\text{Data}}}_{\small \text{Imaging}})$ @ UVA}
+\title[The ANTs Cortical Thickness Pipeline]{The ANTs Cortical Thickness Pipeline}
 \author[N. Tustison]{%
-  Nick Tustison, DSc
+  N.J. Tustison$^{1}$, B.B. Avants$^{2}$, P.A. Cook$^{2}$, G. Song$^{2}$, S. Das$^{2}$, N. van Strien$^{3}$, J.R. Stone$^{1}$, J.C. Gee$^{2}$
   }
-\date[SOM Research Retreat]{}
-\institute[ntustison@virginia.edu]{}
+\date[SPIE 2013:  Orlando, FL]{}
+\institute[ntustison@virginia.edu]{
+  $^{1}$Department of Radiology and Medical Imaging, University of Virginia \\
+  $^{2}$Penn Image and Computing Science Laboratory, University of Pennsylvania \\
+  $^{3}$Department of Circulation and Medical Imaging, Norwegian University of Science and Technology}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  * PNAS VBM study of London cab drivers.
-%%  * 465 data points age ~ 1 + cortical_atrophy
+%%  Cortical thickness uses
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% (Transition)  Let's start off with an interesting study that was done about 10 years ago.
-
-% These computational techniques have proven invaluable in exploring fundamental, and perhaps 
-% not so fundamental, neuroscience questions.
-
-\footnotesize
-
-\begin{frame}{Hippocampal growth in London taxi drivers}
-
+\begin{frame}{Cortical thickness studies}
 \begin{columns}[c]
-    \begin{column}{0.5\textwidth}
-    \begin{center}  
-      \includegraphics[width=5cm]{London_taxi.jpg}
-    \end{center}
+  \begin{column}{0.45\textwidth}
+    \begin{itemize}
+    \item Huntington's disease
+    \item schizophrenia
+    \item bipolar disorder
+    \item Alzheimer's disease
+    \item Fronto-temporal dementia
+    \item Parkinson's disease
+    \item Williams syndrome
+    \item multiple sclerosis
+    \item autism
+    \item migraines
+    \item chronic smoking
+    \end{itemize}
   \end{column}  
-    \begin{column}{0.4\textwidth}
-    \begin{center}  
-      \vspace{-0.5cm}
-      \pause
-      \includegraphics[width=5cm]{hippocampus.png} \\
-      \pause
-        \begin{align*}
-			  SPM (1995):\,\,&   6418\,\,\text{citations}\\
-			  VBM (2000):\,\,&   3593\,\,\text{citations}\\
-			  Opt. VBM (2001):\,\,&  2353\,\,\text{citations}\\
-			  TBSS (2006):\,\,&  1106\,\,\text{citations}\\
-			  \end{align*}
-      \vspace{0cm}
-			  \highlight[yellow!100!black]{$6418 + 3593 + 2353 + 1106 = 13470\,\,\text{citations}$}
-    \end{center}
+  \begin{column}{0.45\textwidth}
+    \begin{itemize}
+    \item alcholism
+    \item cocaine addiction
+    \item Tourette syndrome in children
+    \item scoliosis in female adolescents
+    \item obsessive compulsive disorder
+    \item attention deficit hyperactivity disorder
+    \item obesity
+    \item heritable depression
+    \item elderly depression
+    \end{itemize}
   \end{column}  
@@
 
-\normalsize
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  Studies at UVA
+%%  Cortical thickness uses 2
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% (Transition)  Let's fast forward about 10 years to a young researcher at UVa who.
-% Needless to say it was an interesting two year journey which finally ended in November
-% when our paper was finally published in Human Brain Mapping which shows how these 
-% techniques are susceptible to a methodological flaw known as logical circularity.
-
-{
-\usebackgroundtemplate{\includegraphics[width=1.175\paperwidth]{breachers.jpg}}%
-\begin{frame}{TBI in breachers---{\em neuroimaging correlates?}}
-   \begin{center}
-   \vspace{0cm}
-  \pause
-   \color{black}{Investigate white matter injury \\
-   {\scriptsize controls/students/instructors}} \\
-   \vspace{1cm}
-  \pause
-   \color{black}{\em Logical circularity in\\ SPM, VBM, \& TBSS} \\
-   \end{center}
-\end{frame}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  Intro
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Make commands for the quotes
-\newcommand*{\openquote}{\tikz[remember picture,overlay,xshift=-20pt,yshift=-5pt]
-     \node (OQ) {\quotefont\fontsize{60}{60}\selectfont``};\kern0pt}
-\newcommand*{\closequote}{\tikz[remember picture,overlay,xshift=-30pt,yshift=-10pt]
-     \node (CQ) {\quotefont\fontsize{60}{60}\selectfont''};}
-% select a colour for the shading
-
-\begin{frame}{}
-  \begin{center}
-
+\begin{frame}{Cortical thickness studies (cont.)}
 \begin{columns}[c]
-    \begin{column}{0.6\textwidth}
-	    \begin{center}  
-					\begin{shadequote}
-					{\large {\em Maximum efficiency with minimum} wasted {\em effort}}.\par{Jigaro Kano$'$}
-					\end{shadequote}
-			\end{center}		
+  \begin{column}[t]{0.45\textwidth}
+    \begin{itemize}
+    \item Correlated with:
+      \begin{itemize}
+		    \item age
+		    \item gender
+		    \item untreated transsexuality
+		    \item handedness
+		    \item intelligence
+		    \item athletic ability
+		    \item musical ability
+		    \item tendency towards criminality
+		    \item Tetris-playing ability in female adolescents
+		  \end{itemize}    
+    \end{itemize}
   \end{column}  
-    \hspace{-2.5cm}
-    \begin{column}{0.2\textwidth}
-		  \begin{center}
-		    \includegraphics[height=4cm]{Kano_Jigoro.png}
-		  \end{center}
+  \begin{column}[t]{0.45\textwidth}
+    \begin{itemize}
+     \item functional connectivity relationships
+    \end{itemize}
   \end{column}  
 \end{columns}
-  \vspace{1cm}
-  \pause
-  {\LARGE \em There are \highlight[green, draw=blue]{life-altering} and \highlight[red!100!black]{free} UVA resources that you are probably not using!}
-  \end{center}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  What is data analysis? (neuroscience highlighted)
+%%  Motivation
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}
+% Mention having a complete volumetric pipeline analogous to
+% the entire Freesurfer pipeline.  Also talk about Clarkson's
+% paper.
 
-\tikzset{level 1 concept/.append style={font=\sf, sibling angle=135,level distance = 25mm}}
-\tikzset{level 2 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
-\tikzset{level 3 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
-\tikzset{every node/.append style={scale=0.5}}
+\begin{frame}{}
 
   \begin{center}
-\resizebox{9cm}{7.5cm}{
-\begin{tikzpicture}
-  \path[mindmap,concept color=red!50!black,text=white]
-    node[concept,scale=0.9] {\Large Data Analysis}
-    [clockwise from=45]
-    child[concept color=green!50!black, yshift=-0.5cm] {
-      node[concept,scale=1.1] {Imaging Applications}
-      [clockwise from=120]
-      child[concept color=green!80!black] { node[concept] (ns) {Neuro\-science} }
-      child { node[concept] {Func. Lung Imaging} }
-      child { node[concept] {Breast} }
-      child { node[concept] {Cardiac} }
-    }  
-    child[concept color=blue!80!black, yshift=0.25cm] {
-      node[concept,scale=1.1] (stat) {Statistics}
-      [clockwise from=-60]
-      child { node[concept] {Linear Modeling} }
-      child { node[concept] (ml) {Machine Learning} }
-    }
-    child[concept color=orange, yshift=-0.5cm] { 
-      node[concept,scale=1.1] {Computer Science} 
-        [clockwise from=-145]
-        child { node[concept] (ad) {Algorithm Development}
-           [clockwise from=-60]
-           child { node[concept] (antsr) {\footnotesize ANTsR} }
-           child { node[concept] {ANTs} }
-           child { node[concept] {ITK} }
-           }
-        child { node[concept] (cp) {Computer Programming} }
-        child { node[concept] (cc) {Cluster Computing} }
-    };
+    \includegraphics[height=8cm]{Freesurfer.png}
+   \end{center}
 
-		\newcommand{\connbluetoorange}{to[circle connection bar switch color=from (blue!80!black) to (orange)]}
-		\newcommand{\connorangetoorange}{to[circle connection bar switch color=from (orange) to (orange)]}
-		\newcommand{\conngreentoorange}{to[circle connection bar switch color=from (green!80!black) to (orange)]}
-
-  \begin{pgfonlayer}{background}
-      \path (ad) \connorangetoorange (cp);
-      \path (stat) \connbluetoorange (antsr);
-      \path (ns) \conngreentoorange (cc);
-  \end{pgfonlayer}
-
-%  \begin{pgfonlayer}{background}
-%      \path (ml) \connbluetoorange (cs);
-%  \end{pgfonlayer}
-
-\end{tikzpicture}
-}
-\end{center}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  Wintermark studies
+%%  Materials
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}{}
-\setbeamercovered{invisible}
+\begin{frame}{Data}
+\begin{columns}[c]
+  \begin{column}[c]{0.4\textwidth}
+      \begin{itemize}
+		    \item IXI (Imperial College)
+		      \begin{itemize}
+		        \item 581 subjects
+		        \item T1, T2, PD, DWI
+		        \item 3 sites, 1.5 \& 3 T
+		      \end{itemize}
+		    \item Oasis (WUSTL)
+		      \begin{itemize}
+		        \item 416 subjects
+		        \item 100 mild AD
+		        \item defaced
+		      \end{itemize}
+		    \item NKI/Rockland
+		      \begin{itemize}
+		        \item 188 subjects
+		        \item T1, T2, R-fMRI, DTI
+		        \item defaced
+		      \end{itemize}
+		    \item Kirby (VU)
+		      \begin{itemize}
+		        \item 41 subjects
+		        \item T1, T2, DTI,\ldots
+		      \end{itemize}
+		  \end{itemize}    
+  \end{column}  
+ \begin{column}[c]{0.4\textwidth}
+   \begin{tabular}{cc}
+     \vspace{-0.05cm}
+     \includegraphics[width=1.75cm]{IXI048_axial.png} &  
+     \includegraphics[width=1.75cm]{IXI048_sag.png} \\
+     \vspace{-0.05cm}
+     \includegraphics[width=1.75cm]{OAS1_0039_axial.png} &  
+     \includegraphics[width=1.75cm]{OAS1_0039_sag.png} \\
+     \vspace{-0.05cm}
+     \includegraphics[width=1.75cm]{NKI_2861923_axial.png} &  
+     \includegraphics[width=1.75cm]{NKI_2861923_sag.png} \\
+     \vspace{-0.05cm}
+     \includegraphics[width=1.75cm]{KKI2009-25_axial.png} &  
+     \includegraphics[width=1.75cm]{KKI2009-25_sag.png} 
+   \end{tabular}
+ \end{column}
+\end{columns}
 
-\tikzstyle{mybox} = [draw=black, fill=bottomcolour, thick,
-    rectangle, rounded corners, inner sep=10pt, inner ysep=10pt]
-\tikzstyle{mybox2} = [draw=black, fill=middlecolour, thick,
-    rectangle, rounded corners, inner sep=5pt, inner ysep=5pt, 
-    minimum height=4cm]
-\tikzstyle{fancytitle} =[fill=bottomcolour, text=textcolour, rectangle]
-\tikzstyle{line} = [-latex',very thick,textcolour]
+\end{frame}
 
-%\begin{center}
-%\begin{tikzpicture}
-%};
-%\end{tikzpicture}%
-%\end{center}
-%\vspace{5mm}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  Algorithms
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}{Algorithms}
 
 \begin{center}
-\begin{tikzpicture}[node distance = 4cm]
+  \begin{itemize}
+    \item \texttt{N4BiasFieldCorrection.cxx}\tikz[remember picture] \node (a) {\vphantom{X}};
+    \item \texttt{antsRegistration.cxx}\tikz[remember picture] \node (b) {\vphantom{X}};
+    \item \texttt{antsAffineInitializer.cxx}
+    \item \texttt{Atropos.cxx}\tikz[remember picture] \node (c) {\vphantom{X}};
+    \item \texttt{KellyKapowski.cxx}\tikz[remember picture] \node (d) {\vphantom{X}};
+    \item \texttt{ImageMath.cxx}
+    \item \texttt{abp.sh}
+  \end{itemize}    
+\end{center}
 
-\node [shift={(4cm,4cm)},mybox] (max){%
-  Misc. neuroimaging projects};
+\begin{tikzpicture}[remember picture,overlay]
+\path<2> (a.east) ++(0,1) node[anchor=west,ellipse callout,fill=red!50,opacity=1, callout absolute pointer={(a.mid)},text =black]  {IEEE-TMI 2010};
+\end{tikzpicture}
 
-\uncover<1->{
-\node [mybox2, text width = 3.5cm, opacity = 1] (SkullThickness) {%  
-  \begin{center}
-  \begin{tabular}{c}
-  \includegraphics[height=2cm]{skullThickness.png} 
-  \vspace{0.25cm}
-  \end{tabular}
-  \scriptsize{Surrogate CT Skull Thickness Measurements with MRI}
-  \end{center}
-  };
-\node[fancytitle] (SkullThickness_title) at (SkullThickness.north) {\scriptsize Skull Thickness};
-\path[line] (max.south) edge [out= 230, in= 90] (SkullThickness_title.north);
+\begin{tikzpicture}[remember picture,overlay]
+\path<3> (b.east) ++(0,1) node[anchor=west,ellipse callout,fill=yellow!50,opacity=1, callout absolute pointer={(b.mid)},text =black]  {MIA 2008, IEEE-TIP 2009};
+\end{tikzpicture}
+
+\begin{tikzpicture}[remember picture,overlay]
+\path<4> (c.east) ++(0,1) node[anchor=west,ellipse callout,fill=green!50,opacity=1, callout absolute pointer={(c.mid)},text =black]  {Neuroinformatics 2011};
+\end{tikzpicture}
+
+\begin{tikzpicture}[remember picture,overlay]
+\path<5> (d.east) ++(0,1) node[anchor=west,ellipse callout,fill=orange!50,opacity=1, callout absolute pointer={(d.mid)},text =black]  {Neuroimage 2009};
+\end{tikzpicture}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  abp.sh
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\lstset{frame = htb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{black}},
+        basicstyle=\fontfamily{pcr}\selectfont\tiny\color{green},
+        keywordstyle = {\ttfamily\color{green}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{green}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=bash,
+        floatplacement=!h,
+%        caption={Representative script used for the LPBA40 evaluation.},
+        captionpos=b,
+        label=listing:script
         }
 
-\uncover<2->{
-\node [mybox2, text width = 3.5cm, right of=SkullThickness] (ET) {%
-  \begin{center}
-  \begin{tabular}{c}
-  \includegraphics[height=1.8cm]{ET.png} 
-  \vspace{0.25cm}
-  \end{tabular}
-  \scriptsize{Longitudinal FA correlations following surgical treatment for ET}
-  \end{center}
-  };
-\node[fancytitle](ET_title) at (ET.north) {\scriptsize Essential Tremor};
- \path[line] (max.south) edge [out= 270, in= 90] (ET_title.north);
- }
+\begin{frame}[fragile]
+\frametitle{\texttt{abp.sh}}
 
-\uncover<3->{
-\node [mybox2, text width = 3.5cm, right of=ET] (ASL) {%
-  \begin{center}
-  \begin{tabular}{c}
-  \includegraphics[height=2.75cm]{ASL.png} 
-  \vspace{0.15cm}
-  \end{tabular}
-  \scriptsize{Automatic generation of clinical CBF maps from ASL}
-  \end{center}
-  };
-\node[fancytitle](ASL_title) at (ASL.north) {\scriptsize ASL$\rightarrow$CBF};
-\path[line] (max.south) edge [out= 310, in= 90] (ASL_title.north);
-}
+\begin{lstlisting}
+$ sh ~/Pkg/Utilities/scripts/abp.sh
 
-\end{tikzpicture}
+This script, apb.sh, performs T1 anatomical brain processing where the following
+steps are currently applied:
 
-%\begin{tikzpicture}
-%  \tikz[overlay]\path[line] (max.south) edge [out= 230, in= 90] (SkullThickness_title.north);
-%  \tikz[overlay]\path[line] (max.south) edge [out= 270, in= 90] (ET_title.north);
-%  \tikz[overlay]\path[line] (max.south) edge [out= 310, in= 90] (ASL_title.north);
-%\end{tikzpicture}
-\end{center}
+  1. Brain extraction
+  2. Brain 3-tissue segmentation
+  3. Cortical thickness
+  4. (Optional) registration to a template
+
+Usage:
+
+abp.sh -d imageDimension
+       -a anatomicalImage.nii.gz
+       -e brainExtractionTemplate
+       -m brainExtractionProbabilityMask
+       -l brainSegmentationTemplate
+       -p brainSegmentationPriors
+       <OPTARGS>
+       -o outputPrefix
+
+Example:
+
+  bash /Users/ntustison/Pkg/Utilities/scripts/abp.sh -d 3 -i t1.nii.gz -e brainWithSkullTemplate.nii.gz
+  -m brainPrior.nii.gz -l segmentationTemplate.nii.gz -p segmentationPriors%d.nii.gz -o output
+
+Required arguments:
+   ...
+
+Optional arguments:
+   ...
+
+\end{lstlisting}
 
@@
 
-\uncover<1->{
-    \path[->,rounded corners=0.1cm] node[format, yshift=2cm, text width=1.5cm ] (ixi) { 
+%\uncover<1->{
+    \path[->,rounded corners=0.1cm] node[format, yshift=2cm, text width=1.5cm, text height=0.0cm ] (ixi) { 
         \begin{center}
@@
 			  };
-    \path[->] node[format, right of=ixi, xshift=0.5cm, text width=1.25cm] (n4) {
+    \path[->] node[format, right of=ixi, xshift=0.5cm, text width=1.25cm, text height=-0.35cm ] (n4) {
           \begin{center}
@@
     \draw[->] (ixi.0) --  (n4.180);    
-    \path[->] node[format, below of=n4, yshift = 1.25cm, text width=1.25cm] (be) {
+    \path[->] node[format, below of=n4, yshift = 1.25cm, text width=1.25cm, text height=-0.35cm ] (be) {
           \begin{center}
@@
     \draw[->] (n4.-90) -- (be.90);    
-    \path[->] node[format, below of=be, yshift = 1.25cm, text width=1.25cm] (n4_2) {
+    \path[->] node[format, below of=be, yshift = 1.25cm, text width=1.25cm, text height=-0.35cm ] (n4_2) {
           \begin{center}
@@
     \draw[->] (be.-90) -- (n4_2.90);    
-    \path[->] node[format, below of=n4_2, yshift=1.25cm, text width=1.25cm] (atropos) {
+    \path[->] node[format, below of=n4_2, yshift=1.25cm, text width=1.25cm, text height=-0.35cm ] (atropos) {
           \begin{center}
@@
     \draw[->] (n4_2.-90) -- (atropos.90);    
-    \path[->] node[diamond, draw, thick, fill=middlecolour, left of=atropos, xshift=1.75cm, yshift=0.875cm, text width=0.5cm] (conv) {
+    \path[->] node[diamond, draw, thick, fill=middlecolour, left of=atropos, xshift=1.75cm, yshift=0.875cm, text width=0.5cm, text height=-0.65cm ] (conv) {
           \begin{center}
@@
 
-    \path[->] node[format, left of=atropos, xshift=-0.5cm, text width=1.5cm ] (kk) {
+    \path[->] node[format, left of=atropos, xshift=-0.5cm, text width=1.5cm, text height=-0.35cm  ] (kk) {
         \begin{center}
@@
 			  };
-    \path[->,rounded corners=0.1cm] node[format, above of=kk, yshift=-1cm, text width=1.5cm ] (kkmap) {
+    \path[->,rounded corners=0.1cm] node[format, above of=kk, yshift=-1cm, text width=1.5cm, text height=-0.3cm  ] (kkmap) {
         \begin{center}
@@
     \draw[<-] (kkmap.270) -- (kk.90);    
-}
+%}
 
-\uncover<2->{
+%\uncover<2->{
     \path[->,rounded corners=0.1cm] 
-      node[format, right of=n4, xshift=1.15cm, yshift=-2.5cm,text width=4.75cm] (template) { 
+      node[format, right of=n4, xshift=1.15cm, yshift=-2.5cm,text width=4.75cm,text height=-0.3cm  ] (template) { 
         \begin{center}
@@
     \draw[->,dashed] (template.200) -- ([yshift=0cm,xshift=-0.5cm]template.200) |- (atropos.0); 
-}
+%}
 
-\uncover<3->{
+%\uncover<3->{
   \node[decorate,draw,minimum height=7.25cm,minimum width=5.65cm,dashed,rounded corners=0.2cm, thin, color = yellow] (c) at (1.7,-0.35) {};
-}  
-\uncover<4->{
+%}  
+%\uncover<4->{
   \node[decorate,draw,minimum height=5.25cm,minimum width=5.55cm,dashed,rounded corners=0.2cm, thin, color = red] (c) at (7.7,-0.45) {};
-}
+%}
 \end{tikzpicture}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  Brain extraction
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}{Brain extraction}
+
+  \tikzstyle{format} = []
+
+		\begin{tikzpicture}[node distance=4cm, auto,>=latex', thin]
+		    % We need to set at bounding box first. Otherwise the diagram
+		    % will change position for each frame.
+		    \path[use as bounding box] (-1,0) rectangle (10,0);
+
+\uncover<1->{    
+		    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm, xshift=2cm,yshift=1.5cm] (nodots) { 
+		        \begin{center}
+       	      \includegraphics[width = 4.5cm]{nodots.pdf}
+		        \end{center}
+					  };
+		    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm, below of=nodots, xshift=1.1cm,yshift=2.8cm] { 
+		    {\scriptsize template}
+		    };
+		    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm, below of=nodots, xshift=3.5cm,yshift=2.8cm] { 
+		    {\scriptsize subject}
+		    };
+}					  
+\uncover<2->{    
+		    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm, xshift=2cm,yshift=1.5cm] (dots) { 
+		        \begin{center}
+       	      \includegraphics[width = 4.5cm]{dots.pdf}
+		        \end{center}
+					  };
+}					  
+\uncover<3->{    
+		    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm, text height=-0.3cm, below of = dots] (pair) { 
+		        \begin{center}
+		          \begin{tabular}{c}
+         	      \includegraphics[width = 2.5cm]{laplacian.pdf} \\
+         	      \includegraphics[width = 2.5cm]{nodots.pdf}
+	            \end{tabular}
+		        \end{center}
+					  };
+        \path[draw, -latex'] (nodots) -- node{\scriptsize Affine init.} (pair);			  
+}
+
+\uncover<4->{    
+    \path[->,rounded corners=0.1cm] node[format,text width=5.5cm, right of = dots, xshift=2cm, yshift=-2cm] (fails) { 
+        \begin{center}
+    	      {Extraction failures} \\
+     \vspace{0.5cm}
+   \begin{tabular}{cc}
+     \begin{sideways}\texttt{IXI\_191}\end{sideways} & \includegraphics[width=4cm]{IXI191_failure.png} \\
+     \begin{sideways}\texttt{IXI\_370}\end{sideways} & \includegraphics[width=4cm]{IXI370_failure.png} \\
+     \begin{sideways}\texttt{IXI\_594}\end{sideways} & \includegraphics[width=4cm]{IXI594_failure.png} 
+   \end{tabular}
+        \end{center}
+
+			  };
+}
+
+\end{tikzpicture}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  N4 Atropos brain segmentation
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}{N4 $\leftrightarrow$ Atropos}
+
+\tikzstyle{format} = []
+
+\begin{center}
+\begin{tikzpicture}[node distance=4cm, auto,>=latex', thin]
+  % We need to set at bounding box first. Otherwise the diagram
+ 	% will change position for each frame.
+  \path[use as bounding box] (-1,0) rectangle (10,0);
+
+\uncover<1->{ 
+  \path[->,rounded corners=0.1cm] node[format,text width=1.75cm, xshift = -0.35cm] (extraction) { 
+      \begin{center}
+          \begin{tabular}{cc}
+    	        \includegraphics[width = 1.75cm]{extraction.png} &
+    	        \includegraphics[width = 1.75cm]{weightmask.png} \\
+	           {\scriptsize brain mask} & 
+	           {\scriptsize weight mask}
+	           \end{tabular}
+      \end{center}
+	};
+}	
+\uncover<2->{ 
+  \path[->,rounded corners=0.1cm] 
+    node[format,text width=1.75cm, yshift=2cm, xshift=1.75cm,right of=extraction] (n4) { 
+      \begin{center}
+        \includegraphics[width = 1.75cm]{n4.png} \\
+        {\scriptsize N4}
+      \end{center}
+	};
+  \path[->,thick] (extraction.45) edge [out=45, in=180] (n4.180);			  
+}  
+\uncover<3->{ 
+  \path[->,rounded corners=0.1cm] 
+    node[format,text width=1.75cm, yshift=-2cm, right of=n4] (atropos) { 
+      \begin{center}
+        \includegraphics[width = 1.75cm]{segmentation.png} \\
+	        {\scriptsize Atropos}
+      \end{center}
+	};
+  \path[->,thick] (n4.0) edge [out=0, in=135] (atropos.90);			  
+}
+
+\uncover<4->{ 
+  \path[->,rounded corners=0.1cm] 
+    node[format,text width=2.5cm, below of=n4] (conv) { 
+      \begin{center}
+        \includegraphics[width = 2.5cm]{convergence.pdf} \\
+	        {\scriptsize Convergence?}
+      \end{center}
+	};
+  \path[->,thick] (atropos.-90) edge [out=-135, in=0] (conv.0);			  
+}
+
+\uncover<5->{ 
+  \path[->,thick] (conv.180) edge [out=180, in=-45] (extraction.-45);			  
+  \draw[->,thick,dashed] (atropos.180) -- node {\scriptsize update weight mask} ([xshift=1.75cm]extraction.0);			  
+}
+
+\end{tikzpicture}
+\end{center}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  DiReCT
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{frame}{DiReCT}
+%
+%\end{frame}
 
@@
 }			  
-\uncover<4->{    
-    \path[->,rounded corners=0.1cm] node[format,below of=nirep, text width = 4.0cm, yshift=1cm ] (wow) { 
-        \begin{center}    
-        {\color{yellow}{\em It takes $\sim$20--30 hours per subject!!}}
-        \end{center}    
-			  };
-}			  
 \end{tikzpicture}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  Open source software
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% - The visible human project (1999)---one of the earliest big imaging data/big science challenges.
-%
-
-\begin{frame}{Big data $\rightarrow$ open source software}
-
-\begin{center}
-\begin{tikzpicture}[node distance=3.5cm, auto,>=latex', thin]
-    % We need to set at bounding box first. Otherwise the diagram
-    % will change position for each frame.
-    \path[use as bounding box] (-1,0) rectangle (10,0);
-
-\uncover<1->{
-    \path[->,rounded corners=0.1cm] node[text width=5cm,xshift = 1cm] (nlm) { 
-        \begin{center}
-			  \includegraphics[width=4.755cm]{visible_human_project.png} \\
-        \end{center}
-			  };
-}			  
-\uncover<2->{
-    \path[->,rounded corners=0.1cm] 
-        node[text width=6cm, right of=nlm, xshift=3cm, yshift=1.25cm] (logos) { 
-        \begin{center}
-			  \includegraphics[width=6cm]{logos.png} \\
-        \end{center}
-			  };
-}			  
-\uncover<3->{
-    \path[->,rounded corners=0.1cm] node[text width=3cm, yshift=2cm, below of=logos] (itk) { 
-        \begin{center}
-			  \includegraphics[width=3.cm]{ITK.png} \\
-        \end{center}
-			  };
-}			  
-\uncover<4->{
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift=1.2cm, below of=itk] (ants) { 
-        \begin{center}
-			  \includegraphics[width=2.1cm]{ANTsLogo.png} \\
-        \end{center}
-			  };
-}			  
-\uncover<5->{
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift=1.2cm, below of=itk] (ants) { 
-        \begin{center}
-			  \includegraphics[width=2.1cm]{ANTsLogo.png} \\
-        \end{center}
-			  };
-}
-
-\end{tikzpicture}
-\end{center}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  UVACSE/Linux cluster
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}{UVACSE/computational cluster}
-
-\begin{center}
-\begin{tikzpicture}[node distance=2.5cm, auto,>=latex', thin]
-    % We need to set at bounding box first. Otherwise the diagram
-    % will change position for each frame.
-    \path[use as bounding box] (-1,0) rectangle (10,0);
-
-\uncover<1->{			  
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift = 2cm] (r16) { 
-        \begin{center}
-			  \includegraphics[width=2cm]{r16slice.png} \\
-			  \vspace{-0.5cm}
-			  {\footnotesize \texttt{387\_oasis}}
-        \end{center}
-			  };
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, right of=r16] (r27) { 
-        \begin{center}
-			  \includegraphics[width=2cm]{r27slice.png} \\
-			  \vspace{-0.5cm}
-			  {\footnotesize \texttt{392\_oasis}}
-        \end{center}
-			  };
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of=r16] (r30) { 
-        \begin{center}
-			  \includegraphics[width=2cm]{r30slice.png} \\
-			  \vspace{-0.5cm}
-			  {\footnotesize \texttt{406\_oasis}}
-        \end{center}
-			  };
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of=r27] (r62) { 
-        \begin{center}
-			  \includegraphics[width=2cm]{r62slice.png} \\
-			  \vspace{-0.5cm}
-			  {\footnotesize \texttt{411\_oasis}}
-        \end{center}
-			  };
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of = r30] (r64) { 
-        \begin{center}
-			  \includegraphics[width=2cm]{r64slice.png} \\
-			  \vspace{-0.5cm}
-			  {\footnotesize \texttt{425\_oasis}}
-        \end{center}
-			  };
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, below of=r62] (r85) { 
-        \begin{center}
-			  \includegraphics[width=2cm]{r85slice.png} \\
-			  \vspace{-0.5cm}
-			  {\footnotesize \texttt{430\_oasis}}
-        \end{center}
-			  };
-}			  
-\uncover<2->{			  
-    \path[->,rounded corners=0.1cm] node[text width=6cm, xshift=2cm,right of=r62] (qstat) { 
-        \begin{center}
-			  \includegraphics[width=5.75cm]{qstat.png} \\
-        \end{center}
-			  };		
-}
-
-\uncover<3->{			  
-  \path[->,thick, color = yellow, dashed] (0.5,2.3) edge [out=35,in= 145] (4.5,2.3);
-  \path[->,thick, color = yellow, dashed] (3.1,1.55) edge [out=-35,in=-145] (4.5,1.55);
-  \path[->,thick, color = yellow, dashed] (0.5,-0.25) edge [out=35,in= 145] (4.5,-0.25);
-  \path[->,thick, color = yellow, dashed] (3.1,-0.9) edge [out=-35,in=-145] (4.5,-0.9);
-  \path[->,thick, color = yellow, dashed] (0.5,-2.8) edge [out=35,in= 145] (4.5,-2.8);
-}  
-\uncover<4->{			  
-  \path[->,thick, color = red, dashed] (3.1,-3.45) edge [out=-35,in=-145] (4.5,-3.45);
-}
-
-\end{tikzpicture}
-
-%\begin{tikzpicture}
-%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-4,2.75) edge [out=35,in= 145] (0,2.75);
-%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-1.4,2.1) edge [out=-35,in=-145] (0,2.1);
-%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-4,0.2) edge [out=35,in= 145] (0,0.2);
-%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-1.4,-0.45) edge [out=-35,in=-145] (0,-0.45);
-%  \tikz[overlay]\path[->,thick, color = yellow, dashed] (-4,-2.35) edge [out=35,in= 145] (0,-2.35);
-%  \tikz[overlay]\path[->,thick, color = red, dashed] (-1.4,-3.0) edge [out=-35,in=-145] (0,-3.0);
-%\end{tikzpicture}
-
-\end{center}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%  Closing slide
@@
 
-\begin{frame}{Further information}
+\begin{frame}{}
 
@@
 			  };
-    \path[->,rounded corners=0.1cm] node[text width=6cm, xshift=5cm, right of=ants] { 
+    \path[->,rounded corners=0.1cm] node[text width=6.5cm, xshift=5.5cm, right of=ants] { 
         \begin{center}
+        \begin{block}{Further information}
         \begin{itemize}
-          \item Insight Toolkit:  http://www.itk.org
-          \item ANTs/ANTsR:  http://www.picsl.upenn.edu/ANTs/
-          \item UVACSE:  http://www.uvacse.virginia.edu
+          \item \url{http://www.itk.org}
+          \item \url{http://www.picsl.upenn.edu/ANTs}
         \end{itemize}
+        \end{block}
         \end{center}

### 2013-02-08 SPIE2013/KK/article.tex
ENH:  Updating.

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage{amsmath}

\title{The ANTs cortical thickness processing pipeline}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Nicholas J. Tustison\supit{a}, Brian B. Avants\supit{b}, Philip A. Cook\supit{b}, Gang Song\supit{b}, Sandhitsu Das\supit{b}, Niels van Strien\supit{c}, James R. Stone\supit{a}, James C. Gee\supit{b}
\skiplinehalf
\supit{a}Dept. of Radiology and Medical Imaging, Univ. of Virginia, Charlottesville, Virginia, USA\\
\supit{b}PICSL, Univ. of Pennsylvania, Philadelphia, Pennsylvania, USA\\
\supit{c}Dept. of Circulation and Medical Imaging, Norwegian University of Science and Technology, Trondheim, Norway\\
}

%>>>> Further information about the authors, other than their 
%  institution and addresses, should be included as a footnote, 
%  which is facilitated by the \authorinfo{} command.

\authorinfo{Further author information: (Send correspondence to N.J.T.)\\
            N.J.T.: E-mail: njt4n@virginia.edu, Telephone: 1 434 924 7730}
%%>>>> when using amstex, you need to use @@ instead of @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

  \begin{document} 
  \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
Numerous studies have explored the relationship between cortical
structure and brain development, cognitive 
function, and functional connectivity.
The highly convoluted cortical topography makes manual measurements
arduous and often impractical given the population sizes necessary 
for sufficient statistical power.  Computational techniques 
have permitted large-scale studies as they
provide robust and reliable localized measurements characterizing
the cortex with little or no human intervention.  Particularly useful
to the neuroscience community are publicly available tools, such as the popular 
surface-based Freesurfer, which facilitate the testing and refinement of 
hypotheses.  In this paper, we introduce the volume-based 
Advanced Normalization Tools
(ANTs) cortical thickness automated pipeline
comprising well-vetted components such as SyGN (multivariate template construction),
SyN (image registration), N4 (bias
correction),
Atropos ($n$-tissue segmentation), and DiReCT (cortical thickness) 
all developed as part of the ANTs open science effort.  
Complementing the open source aspect of ANTs we demonstrate its
utility using the publicly available IXI data set.
\end{abstract}

%>>>> Include a list of keywords after the abstract

\keywords{advanced normalization tools, DiReCT, Insight Toolkit, KellyKapowski}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}
\label{sec:intro}  % \label{} allows reference to this section

Historically rooted in the meticulous work of von Economo \cite{economo2008},
imaging-based structural analysis of the brain plays a fundamental role
in identifying the relationship between cortical morphology, disease and cognition.
The availability of quantitative computational methods for extracting such information
has proven invaluable for developing and refining fundamental 
neuroscience hypotheses.  Computational methods for analyzing the cortex may be 
broadly characterized as surface mesh-based or volumetric.  
Representative of the former is the
Freesurfer%
\footnote{
http://surfer.nmr.mgh.harvard.edu/
}
cortical modeling software package \cite{dale1999,fischl1999}.
%which owes its popularity to public availability, excellent documentation, 
%good performance, and  integration with other toolkits.  
Similar to other surface
approaches, the pial
and white matter surfaces from individual subject MR data are modeled with polygonal meshes  
which are then used to determine local cortical thickness based on a specified correspondence between 
the surface models.
In contrast, volumetric approaches directly operate in the image space.  For example, 
DiReCT (Diffeomorphic Registration-based 
Cortical Thickness) algorithm\cite{das2009} is 
a registration-based approach where the derived diffeomorphic mapping between the 
white and pial matter surfaces is used to propagate thickness values 
through the cortical gray matter.  
%A unique benefit of DiReCT is that it
%naturally estimates the boundaries of buried sulci by employing a
%diffeomorphic constraint on the probabilistic estimate of the gray
%matter and cerebrospinal fluid interface.

Despite the variety of techniques for estimating cortical thickness
from imaging data, several common pre-processing components
can be identified.
The most fundamental of these include inhomogeneity correction, skull stripping, and $n$-tissue segmentation 
for differentiating the gray and white matter. 
In this work, we describe our C$++$ based cortical thickness pipeline
which is freely available as part of the Advanced Normalization Tools
(ANTs) software package based on the Insight Toolkit.  This includes 
all the necessary preprocessing steps consisting
of well-vetted previously published algorithms for bias correction \cite{tustison2010},
brain extraction \cite{avants2010a}, $n$-tissue segmentation \cite{avants2011a},
template construction \cite{avants2010}, and image normalization \cite{avants2011}.
Equally as important, we demonstrate how to coordinate
these pipeline components and provide a set of useful parameters
which are employed to analyze the publicly available IXI data set. The
full pipeline and parameter set is 
encapsulated in a well-documented shell script which is also available in ANTs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MATERIALS AND METHODS}

\begin{figure}
  \centering
  \includegraphics[width=120mm]{Kapowski_pipeline2.pdf}
  \caption{The ANTs T1 processing workflow containing all elements for 
  determining cortical thickness.   
  Not shown is the optional single subject
  to template registration.
  }
  \label{fig:pipeline}
\end{figure}

The ANTs-based cortical thickness estimation workflow is illustrated 
in Figure \ref{fig:pipeline}.  The steps are as follows: 
1) initial N4 bias correction on input anatomical MRI,
2) brain extraction using a hybrid segmentation/template-based strategy,
3) alternating between prior-based segmentation and white matter posterior
        probability weighted bias correction,
4) DiReCT-based cortical thickness estimation, and
5) optional normalization to specified template.
The coordination of all the algorithmic components is
encapsulated in the shell script \texttt{antsCorticalThickness.sh}.  
This includes
optimal parameters for each of the algorithmic components which were used to acquire the results 
described in this work.

\subsection{Description of Processing Pipeline Components}
\begin{description}
\item[N4:] We introduced an extension of N3 into ANTs, denoted as N4, which demonstrates 
improved performance and convergence behavior on a variety of data. This improvement 
is a result of an enhanced fitting routine and a modified optimization formulation. 
For our workflow, the additional possibility of specifying a weighted mask in N4 
permits the use of the current white matter probability map calculated during the 
segmentation pipeline for further improvement of bias field estimation.
\item[Atropos:]  Our $n$-tissue segmentation
software tool (which we denote as ÒAtroposÓ) attempts to distill 20+ years of active research 
in this area particularly some of its most seminal work. Specification of prior probabilities includes spatially varying Markov Random Field modeling, prior label maps, and prior probability maps typically derived from our template building process. Additional capabilities include handling of multivariate data, partial volume modeling, a memory-minimization mode, label propagation, a plug-n-play architecture for incorporation of novel likelihood models which includes both parametric and non-parametric models for both scalar and tensorial images, and alternative posterior formulations for different segmentation tasks.
\item[Brain extraction:]  Brain extraction using ANTs combines template building, high-performance brain image registration [5], and Atropos with topological refinements. An optimal template for brain extraction is generated offline using labeled brain data. For example, in this work we use the LPBA40 data for generating a brain extraction template and a corresponding brain probability mask.  A comparison\cite{avants2010a} using open access brain data with other publicly available brain extraction algorithms demonstrated that our combined registration/segmentation approach performs at the top level alongside BrainSuite (tuned) and FreeSurfer.
\item[DiReCT:] Although the basic formulation of DiReCT as reported in this work is as it was introduced, we have made several improvements. Perhaps the most significant advance is that this particular ITK-compatible implementation has been significantly multi-threaded, is written in ITK coding style, and has been made publicly available through ANTs complete with a unique user interface design developed specifically for ANTs tools.
\end{description}

\subsection{Description of Publicly Available Image Data Sets}

Several publicly available data sets were used in this work.  The IXI data%
\footnote{
http://biomedic.doc.ic.ac.uk/brain-development/
}
were processed to produce the cortical thickness maps.  
They were also used to construct the gender/age-based
multivariate templates consisting of T1, T2, proton density,
and FA components.  
%Sample axial slices from
%the female, age 50--60 subgroup template are shown in the middle of Fig. \ref{fig:template}.
The annotated NIREP NA0%
\footnote{
http://www.nirep.org
}
and LPBA40%
\footnote{
http://www.loni.ucla.edu/Atlases/LPBA40/
}
data sets were used, respectively, in the processing pipeline for 
 3-tissue segmentation and brain extraction
for each subject (as illustrated in Fig. \ref{fig:pipeline}).  Additionally,
we registered the NIREP template to the total template to propagate the NIREP
NA0 labels for tabulating thickness values in each of the 32 cortical labels
over the male and female IXI subjects.

%\paragraph{IXI:} The IXI data set consists of approximately 600 subjects.  Demographic
%information including gender and age is also available for each subject.  Further
%details can be found on the IXI website.%
%\footnote{
%}  
%Each subject was processed using the \texttt{antsCorticalThickness} script ultimately
%yielding cortical maps for each IXI subject.

%\begin{figure}
%  \centering
%  \includegraphics[width=170mm]{template50_60_long.pdf}
%  \caption{Multivariate template (T1, T2, PD, and FA) made from the IXI female, age 50--60 cohort.  
%  The set of all gender/age-based templates were then used to create a final template which
%  served as the normalized space for statistical comparison.
%  }
%  \label{fig:template}
%\end{figure}

%Age and gender-based multivariate templates\cite{avants2010} were also created from the IXI data
%consisting of T1, T2, proton-density (PD), and FA components.  Sample axial slices from
%the female, age 50--60 subgroup template are shown in the middle of Fig. \ref{fig:template}.  Each
%subject-specific set of T1, T2, PD, and FA images (of which 5 are shown) are weighted equally in constructing
%the multivariate template.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RESULTS}

The resulting thickness values for each subject were averaged within each
of the 32 labels subsequent to normalization of thickness values by the
ratio of the template volume to the individual subject volume.
Normalized cortical thickness plots (age vs. thickness) are given in Fig.
\ref{fig:nirep}.

%\begin{figure}
%  \centering
%  \begin{tabular}{cccc}
%  \includegraphics[width=51mm]{IXI_results/yylabel1_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel2_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel3_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel4_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel5_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel6_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel7_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel8_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel9_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel10_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel11_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel12_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel13_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel14_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel15_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel16_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel17_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel18_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel19_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel20_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel21_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel22_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel23_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel24_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel25_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel26_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel27_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel28_results.pdf} \\
%  \includegraphics[width=51mm]{IXI_results/yylabel29_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel30_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel31_results.pdf} &
%  \includegraphics[width=51mm]{IXI_results/yylabel32_results.pdf} 
%  \end{tabular}
%  \caption{Cortical thickness results (age vs. thickness) from the IXI data set where
%  each plot corresponds to one of the 32 cortical labels from the NIREP NA0 data set.  
%  Thickness values have been normalized corresponding to the ratio of the individual 
%  subject volume versus the total template volume.
%  }
%  \label{fig:nirep}
%\end{figure}

\begin{figure}
  \centering
  \begin{tabular}{ccc}
  \includegraphics[width=51mm]{IXI_results/yylabel1_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel2_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel3_results.pdf} \\
  \includegraphics[width=51mm]{IXI_results/yylabel4_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel5_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel6_results.pdf} \\
  \includegraphics[width=51mm]{IXI_results/yylabel7_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel8_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel9_results.pdf} \\
  \includegraphics[width=51mm]{IXI_results/yylabel10_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel11_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel12_results.pdf} \\
  \includegraphics[width=51mm]{IXI_results/yylabel13_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel14_results.pdf} &
  \includegraphics[width=51mm]{IXI_results/yylabel15_results.pdf} 
  \end{tabular}
  \caption{Cortical thickness results (age vs. thickness) from the IXI data set where
  each plot corresponds to one of the first 15 of 32 total cortical labels from the NIREP NA0 data set.  
  Thickness values have been normalized corresponding to the ratio of the individual 
  subject volume versus the total template volume.
  }
  \label{fig:nirep}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS}

Given the numerous studies that have used cortical thickness for 
making neuroscientific inferences, it is of vital importance that
1) there be multiple, robust computational tools for extracting such
quantitative information and 2) that these tool sets be publicly 
available for validating research findings.  The Freesurfer package
has proven to be exemplary in this respect.  Our proposed work is
meant to compliment such packages as Freesurfer by providing a 
volume-based, well-vetted, open source resource for extracting 
cortical thickness maps in large studies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\acknowledgments     %>>>> equivalent to \section*{ACKNOWLEDGMENTS}       
% 
%This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{references}   %>>>> bibliography data in report.bib
\bibliographystyle{spiebib}   %>>>> makes bibtex use spiebib.bst

\end{document}

### 2013-02-08 SPIE2013/LongitudinalHe3/Presentation/presentation.tex
ENH:  Updating.

\documentclass[xcolor=dvipsnames,serif,10pt]{beamer}

\usepackage{beamerthemesplit}
\usepackage{media9}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
%\usefonttheme{professionalfonts}
%\usepackage{times}
\usepackage{rotating}
\usepackage{tikz}
\usepackage{amsmath,cancel}
\usepackage{mathtools}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{listings}
\usepackage[]{geometry}
\usepackage{ifxetex}
\ifxetex{%
  \usepackage{fontspec}
  \setmainfont{Linux Libertine O} % or any font on your system
  \newfontfamily\quotefont[Ligatures=TeX]{Linux Libertine O} % or any font on your system
\else
  \usepackage[T1]{fontenc}
  \usepackage{libertine} % or any other font package (or none)
  \newcommand*\quotefont{\fontfamily{fxl}} % selects Libertine for quote font
\fi
\usepackage{framed}
\usepackage{booktabs}

\usepackage{soul}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.text}
\usetikzlibrary{calc,shapes.callouts,shapes.arrows,arrows}
\makeatletter

\newcommand{\defhighlighter}[3][]{%
  \tikzset{every highlighter/.style={color=#2, fill opacity=#3, #1}}%
}

\defhighlighter{yellow}{.5}

\newcommand{\highlight@DoHighlight}{
  \fill [ decoration = {random steps, amplitude=1pt, segment length=15pt}
        , outer sep = -15pt, inner sep = 0pt, decorate
        , every highlighter, this highlighter ]
        ($(begin highlight)+(0,8pt)$) rectangle ($(end highlight)+(0,-3pt)$) ;
}

\newcommand{\highlight@BeginHighlight}{
  \coordinate (begin highlight) at (0,0) ;
}

\newcommand{\highlight@EndHighlight}{
  \coordinate (end highlight) at (0,0) ;
}

\newdimen\highlight@previous
\newdimen\highlight@current

\DeclareRobustCommand*\highlight[1][]{%
  \tikzset{this highlighter/.style={#1}}%
  \SOUL@setup
  %
  \def\SOUL@preamble{%
    \begin{tikzpicture}[overlay, remember picture]
      \highlight@BeginHighlight
      \highlight@EndHighlight
    \end{tikzpicture}%
  }%
  %
  \def\SOUL@postamble{%
    \begin{tikzpicture}[overlay, remember picture]
      \highlight@EndHighlight
      \highlight@DoHighlight
    \end{tikzpicture}%
  }%
  %
  \def\SOUL@everyhyphen{%
    \discretionary{%
      \SOUL@setkern\SOUL@hyphkern
      \SOUL@sethyphenchar
      \tikz[overlay, remember picture] \highlight@EndHighlight ;%
    }{%
    }{%
      \SOUL@setkern\SOUL@charkern
    }%
  }%
  %
  \def\SOUL@everyexhyphen##1{%
    \SOUL@setkern\SOUL@hyphkern
    \hbox{##1}%
    \discretionary{%
      \tikz[overlay, remember picture] \highlight@EndHighlight ;%
    }{%
    }{%
      \SOUL@setkern\SOUL@charkern
    }%
  }%
  %
  \def\SOUL@everysyllable{%
    \begin{tikzpicture}[overlay, remember picture]
      \path let \p0 = (begin highlight), \p1 = (0,0) in \pgfextra
        \global\highlight@previous=\y0
        \global\highlight@current =\y1
      \endpgfextra (0,0) ;
      \ifdim\highlight@current < \highlight@previous
        \highlight@DoHighlight
        \highlight@BeginHighlight
      \fi
    \end{tikzpicture}%
    \the\SOUL@syllable
    \tikz[overlay, remember picture] \highlight@EndHighlight ;%
  }%
  \SOUL@
}
\makeatother

% wrap everything in its own environment
\newenvironment{shadequote}%
{\begin{quote}\openquote}
{\hfill\closequote\end{quote}}

\newcommand{\arrowthis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};
        \node[overlay,single arrow,draw=none,fill=red!50,anchor=tip,rotate=60] 
        at (#1.south) {#2};}%
    }%

\newcommand{\speechthis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};\node[overlay,ellipse callout,fill=blue!50] 
        at ($(#1.north)+(-.5cm,0.8cm)$) {#2};}%
    }%

\newcommand{\bubblethis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};\node[overlay,cloud callout,callout relative pointer={(0.2cm,-0.7cm)},%
        aspect=2.5,fill=yellow!90] at ($(#1.north)+(-0.5cm,1.6cm)$) {#2};}%
    }%

\newcommand{\pointthis}[2]{
        \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
        (#1) {\underline{#1}};\node[overlay,rectangle callout,%
        callout relative pointer={(0.2cm,0.7cm)},fill=green!50] at ($(#1.north)+(-.5cm,-1.4cm)$) {#2};}%
        }%

\usetikzlibrary{mindmap,backgrounds}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.9}
    \definecolor{listwhite}{gray}{1.0}

\usetheme[secheader]{Boadilla}
\useoutertheme{miniframes}
\useinnertheme{circles}
\usecolortheme{myct}
\usepackage[latin1]{inputenc}

\graphicspath{{Figures/}{../Figures/}}

\setlength{\tabcolsep}{1pt}

\newcommand\fontvi{\fontsize{6}{8}\selectfont}
\newcommand{\ColIndent}{\hspace{\labelwidth}\hspace{\labelsep}\hspace{\labelsep}\hspace{\labelsep}\hspace{\labelsep}}

\usetikzlibrary{backgrounds}
\usetikzlibrary{positioning}
\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Title slide
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Longitudinal assessment]{Longitudinal assessment of treatment effects on pulmonary ventilation using 1H/3He MRI multivariate templates}
\author[N. Tustison]{%
  N. J. Tustison$^{1}$, B. Contrella$^{1}$, T. A. Altes$^{1}$, B. B. Avants$^{2}$, E. E. de Lange$^{1}$, J. P. Mugler III$^{1}$
  }
\date[SPIE 2013:  Orlando, FL]{}
\institute[ntustison@virginia.edu]{
  $^{1}$Department of Radiology and Medical Imaging, University of Virginia \\
  $^{2}$Penn Image and Computing Science Laboratory, University of Pennsylvania
}
%\pgfdeclaremask{fsu}{fsu_logo_ybkgrd}
%\pgfdeclareimage[width=1cm]{logo}{ants_logo}
%\logo{\vbox{\vskip0.1cm\hbox{\pgfuseimage{logo}}}}

%\logo{
%  \includegraphics[height=0.8cm]{brain2.png}
%  \includegraphics[height=0.8cm]{itkLogo.jpg}
%  }

\begin{document}

\tikzstyle{every picture}+=[remember picture]

\frame{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Segmentation review
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Previously\ldots}

\begin{center}
  \begin{tabular*}{0.5\textwidth}{l@{\extracolsep{\fill}}c@{\extracolsep{\fill}}c}
    \toprule
    Rater & Sensitivity & Specificity\\
    \midrule
    Atropos & 0.898 & 0.905 \\
    Radiologist 1 & 0.743 & 0.897 \\
    Radiologist 2 & 0.501 & 0.985 \\
    Radiologist 3 & 0.898 & 0.848 \\
    First author & 0.600 & 0.984 \\
    K-means      & 0.973 & 0.734 \\
    \bottomrule
  \end{tabular*}
\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 4-D Atropos
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tikzset{
%Define standard arrow tip
>=stealth'
}

\begin{frame}{Longitudinal segmentation}

 \begin{tikzpicture}[scale=1]
   \path[use as bounding box] (0,6) rectangle (-1,0);

    % A grid can be useful when defining coordinates
    \draw[step=2mm, middlecolour, thin] (0,0) grid (9.8,5.8); 
%    \draw[step=5mm, black] (0,0) grid (10,6.5);

   \draw[<-|,thick] (3.25,-0.5) -- node[sloped,midway,above] {\footnotesize Time point} +(-4,4);
   \filldraw (2.85,-0.1) circle (2pt);
   \filldraw (2.05,0.7) circle (2pt);
   \filldraw (1.25,1.5) circle (2pt);
   \filldraw (0.45,2.3) circle (2pt);
   \filldraw (-0.35,3.1) circle (2pt);

    \node[left] at (-0.35,3.1) {\scriptsize $t=1$}; 
    \node[left] at (0.45,2.3) {\scriptsize $t=2$}; 
    \node[left] at (1.25,1.5) {\scriptsize $t=3$}; 
    \node[left] at (2.05,0.7) {\scriptsize $t=4$}; 
    \node[left] at (2.85,-0.1) {\scriptsize $t=5$};

   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=1.0cm,yshift=4.15cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint1.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=1.8cm,yshift=3.35cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint2.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=2.6cm,yshift=2.55cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint3.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=3.5cm,yshift=1.75cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint4.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=4.2cm,yshift=0.95cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint5.png}
			};

   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=7.0cm,yshift=4.15cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint1_atropos.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=7.8cm,yshift=3.35cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint2_atropos.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=8.6cm,yshift=2.55cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint3_atropos.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=9.5cm,yshift=1.75cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint4_atropos.png}
			};
   \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=10.2cm,yshift=0.95cm]  { 
  		  \includegraphics[width=1.5cm]{subject2_timePoint5_atropos.png}
			};

%   \draw[->, >=latex, blue!20!white, line width=72pt] (4, 16) Ñ node [black,sloped]            {Optimization} +(-45:17cm);

   \draw[->, thick, dashed] (3, 4) -- node [text width= 3cm,above, midway, xshift=0.25cm]  {\scriptsize Intensity normalization and 4-D segmentation} +(3,0);

    \end{tikzpicture}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Expected ventilation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Expected ventilation}

 \begin{tikzpicture}[scale=1]
   \path[use as bounding box] (0,6) rectangle (-1,0);

   \path[->,rounded corners=0.1cm] node[yshift = 5.5cm] (pos1) { 
  		  \includegraphics[width=1.5cm]{subject2_526_posterior1.png}
			};
   \path[->,rounded corners=0.1cm] node[below of=pos1, yshift=-0.5cm] (pos2) { 
  		  \includegraphics[width=1.5cm]{subject2_526_posterior2.png}
			};
   \path[->,rounded corners=0.1cm] node[below of=pos2, yshift=-0.5cm] (pos3) { 
  		  \includegraphics[width=1.5cm]{subject2_526_posterior3.png}
			};
   \path[->,rounded corners=0.1cm] node[below of=pos3, yshift=-0.5cm] (pos4) { 
  		  \includegraphics[width=1.5cm]{subject2_526_posterior4.png}
			};
   \path[->,rounded corners=0.1cm] node[right of=pos2, xshift=3.75cm, yshift=-0.75cm] (eqn) { 
    $EV_{v} = \sum_{l=1}^4 l \times \mathrm{Pr}( l | I_v )$ 
			};
   \path[-,rounded corners=0.1cm] node[right of=pos2, xshift=8cm, yshift=-0.75cm] (ev) { 
  		  \includegraphics[width=3cm]{subject2_526_ev.png}
			};

   \path[->,thick, color = red] (pos1.east) edge [out=0, in=180] node[xshift=-0.75cm,yshift=1.3cm] {\scriptsize $l=1$} (eqn.west);			  
   \path[->,thick, color = green] (pos2.east) edge [out=0, in=180] node[xshift=-0.75cm,yshift=0.55cm] {\scriptsize $l=2$} (eqn.west);			  
   \path[->,thick, color = blue] (pos3.east) edge [out=0, in=180] node[xshift=-0.75cm,yshift=-0.5cm] {\scriptsize $l=3$}(eqn.west);			  
   \path[->,thick, color = yellow] (pos4.east) edge [out=0, in=180] node[xshift=-0.75cm,yshift=-1.25cm] {\scriptsize $l=4$}(eqn.west);			  
   \path[->,thick] (eqn.east) edge [out=0, in=180] (ev.west);

 \end{tikzpicture}

%\begin{block}{}
%  \begin{align*} 
%    EV_{v} = \sum_{l=1}^L l \times \mathrm{Pr}( l | I_v ) \\
%  \end{align*}
%\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  template generation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Template generation}

Given $\{I_1^{1H},I_1^{3He},\ldots,I_5^{1H},I_5^{3He}\}$, template generation involves finding:
\begin{itemize}
  \item the set of paired diffeomorphic transformations $\left\{ \left(\phi_1,\phi_1^{-1}\right),\ldots, \left(\phi_5,\phi_5^{-1}\right) \right\}$,
  \item optimal appearance template, $J^{1H}$ and $J^{3He}$, and
  \item the corresponding coordinate system, $\psi(\mathbf{x})$
\end{itemize}
which minimize$^\dagger$
\begin{block}{}
\begin{align*}
C &= \sum_{i=1}^5 \left[ 
             D\left( \psi((\mathbf{x})), \phi(\mathbf{x}, 1)\right)
             + \Pi_1\left( I_m^{1H} \left( \phi_2^m (\mathbf{x},0.5 )\right), 
                         J^{1H}\left(\phi_1^m(\mathbf{x},0.5)\right)\right) 
             \right. \\
              &+ \left.\Pi_2\left( I_m^{3He} \left( \phi_2^m (\mathbf{x},0.5 )\right),
                          J^{3He}\left(\phi_2^m(\mathbf{x},0.5)\right)\right)
            \right]\\
\end{align*}
{ \scriptsize $^\dagger$implemented in \texttt{antsMultivariateTemplateConstruction.sh}}
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  He3/H1 template
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{He3/H1 longitudinal template}

\begin{center}
  \includegraphics[width=9cm]{He3H1template.png} 
\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Correlation analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tikzset{
%Define standard arrow tip
>=stealth',
%Define style for different line styles
help lines/.style={dashed, thick},
axis/.style={<->,thick},
important line/.style={thick},
connection/.style={thick, dotted},
}

\begin{frame}{Correlation analysis}

%\begin{tikzpicture}[node distance=1cm, auto,>=latex', thin]
%    % We need to set at bounding box first. Otherwise the diagram
%    % will change position for each frame.

 \begin{tikzpicture}[scale=1]
   \path[use as bounding box] (0,6) rectangle (-1,0);

    % A grid can be useful when defining coordinates
    \draw[step=2mm, middlecolour, thin] (0,0) grid (9.8,5.8); 
%    \draw[step=5mm, black] (0,0) grid (10,6.5);

    % Axis
    \coordinate (y) at (0,6);
    \coordinate (x) at (10,0);
    \draw[<->,thick] (y) node[above] {\scriptsize \em Expected ventilation} -- (0,0) --  (x) node[right]
    {\scriptsize \em Time};

    \coordinate (y1) at (0, 1.25);
    \coordinate (y2) at (0, 2.5);
    \coordinate (y3) at (0, 3.75);
    \coordinate (y4) at (0, 5);

    \node[left] at (y1) {1}; 
    \node[left] at (y2) {2}; 
    \node[left] at (y3) {3}; 
    \node[left] at (y4) {4};

    \coordinate (x1) at (1.75, 0);
    \coordinate (x2) at (3.5, 0);
    \coordinate (x3) at (5.25, 0);
    \coordinate (x4) at (7, 0);
    \coordinate (x5) at (8.75, 0);

    \node[below] at (x1) {1}; 
    \node[below] at (x2) {2}; 
    \node[below] at (x3) {3}; 
    \node[below] at (x4) {4}; 
    \node[below] at (x5) {5};

    \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=1.75cm,yshift=1.15cm] (t1) { 
			  \includegraphics[width=1.5cm]{timePoint1.png} \\
			  };
    \draw[thick,red] ([xshift=-0.35cm,yshift=1.5cm]x1) circle (1pt);

    \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=3.5cm,yshift=1.15cm] (t2) { 
			  \includegraphics[width=1.5cm]{timePoint2.png} \\
			  };
    \draw[thick,green] ([xshift=-0.35cm,yshift=1.5cm]x2) circle (1pt);

    \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=5.25cm,yshift=1.15cm] (t3) { 
			  \includegraphics[width=1.5cm]{timePoint3.png} \\
			  };
    \draw[thick,blue] ([xshift=-0.35cm,yshift=1.5cm]x3) circle (1pt);

    \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=7cm,yshift=1.15cm] (t4) { 
			  \includegraphics[width=1.5cm]{timePoint4.png} \\
			  };
    \draw[thick,orange] ([xshift=-0.35cm,yshift=1.5cm]x4) circle (1pt);

    \path[->,rounded corners=0.1cm] node[text width=1.5cm,xshift=8.75cm,yshift=1.15cm] (t5) { 
			  \includegraphics[width=1.5cm]{timePoint5.png} \\
			  };
    \draw[thick,violet] ([xshift=-0.35cm,yshift=1.5cm]x5) circle (1pt);

    % Let us define some coordinates
    \path
    coordinate (c0) at (1.75, 2.5)
    coordinate (c1) at (3.5, 2.5)
    coordinate (c2) at (5.25,5)
    coordinate (c3) at (7,5)
    coordinate (c4) at (8.75,3.75);

    \draw[dashed,very thick] (c0) -- (c1) -- (c2)  -- (c3) -- node[sloped,midway,below] {\tiny simplified hypothesis} (c4);

%    \draw[dashed,very thick] (c0) -- (c1) -- (c2) node[xshift=0.9cm,yshift=-0.25cm] {\tiny simplified hypothesis} -- (c3) -- (c4);

    \draw[very thick,red] ([yshift=1.2pt]c0) circle (3pt);
    \draw[very thick,green] ([yshift=17pt]c1) circle (3pt);
    \draw[very thick,blue] ([yshift=5pt]c2) circle (3pt);
    \draw[very thick,orange] ([yshift=12pt]c3) circle (3pt);
    \draw[very thick,violet] ([yshift=40pt]c4) circle (3pt);

    \draw[|<->|,dotted,thick] ([xshift=-10pt,yshift=17pt]c1) -- node[sloped,midway,above] {\tiny magnitude difference} +(0,2.1);

    \end{tikzpicture}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Subjects 1-4
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Subjects 1--4}
\begin{center}
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025001-3D.mov,
                  flashvars={%
                    src=VX770-025001-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}                  
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025002-3D.mov,
                  flashvars={%
                    src=VX770-025002-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}   \\               
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025003-3D.mov,
                  flashvars={%
                    src=VX770-025003-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}                  
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025004-3D.mov,
                  flashvars={%
                    src=VX770-025004-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}             
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Subjects 5-8
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Subjects 5--8}
\begin{center}
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025005-3D.mov,
                  flashvars={%
                    src=VX770-025005-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}                  
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025006-3D.mov,
                  flashvars={%
                    src=VX770-025006-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}   \\               
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025007-3D.mov,
                  flashvars={%
                    src=VX770-025007-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}                  
    \includemedia[activate=pageopen,
                  width=5.9325cm,
                  height=3.465cm,
                  addresource=VX770-025008-3D.mov,
                  flashvars={%
                    src=VX770-025008-3D.mov%
                    &autoPlay=true%
                    &loop=true%
                    }
                  ]{}{StrobeMediaPlayback.swf}             
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Test
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{frame}
%\begin{itemize}
%  \item This one is always shown
%  \item<1-> The first time (i.e. as soon as the slide loads)
%  \item<2-> The second time
%  \item<1-> Also the first time
%  \only<1-1> This one is shown at the first time, but it will hide soon (on the next event after the slide loads).
%\end{itemize}
%\end{frame}

%\begin{frame}
%        \frametitle{`Hidden higher-order concepts?'}
%        \begin{itemize}[<+->]
%        \item The truths of arithmetic which are independent of PA in some 
%        sense themselves `{contain} essentially {\color{blue}{hidden higher-order}},
%         or infinitary, concepts'???
%        \item `Truths in the language of arithmetic which \ldots
%        \item   That suggests stronger version of Isaacson's thesis. 
%        \end{itemize}
%\end{frame}
%
%\begin{frame}
% 
%   \begin{block}{This is a Block}
%      This is important information
%   \end{block}
% 
%   \begin{alertblock}{This is an Alert block}
%   This is an important alert
%   \end{alertblock}
% 
%   \begin{exampleblock}{This is an Example block}
%   This is an example 
%   \end{exampleblock}
% 
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Closing slide
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{}

\begin{tikzpicture}[node distance=1cm, auto,>=latex', thin]
    % We need to set at bounding box first. Otherwise the diagram
    % will change position for each frame.
    \path[->,rounded corners=0.1cm] node[text width=2cm,xshift=-2cm] (ants) { 
        \begin{center}
			  \includegraphics[width=3.5cm]{ants2.png} \\
        \end{center}
			  };
    \path[->,rounded corners=0.1cm] node[text width=6.5cm, xshift=5.5cm, right of=ants] { 
        \begin{center}
        \begin{block}{Further information}
        \begin{itemize}
          \item \url{http://www.itk.org}
          \item \url{http://www.picsl.upenn.edu/ANTs}
        \end{itemize}
        \end{block}
        \end{center}
			  };

\end{tikzpicture}
\end{frame}

\end{document}

### 2013-02-08 SPIE2013/LongitudinalHe3/article.tex
ENH:  Updating.

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage{amsmath}

\title{Longitudinal assessment of treatment effects on pulmonary ventilation using 1H/3He MRI multivariate templates}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Nicholas J. Tustison\supit{a}, Benjamin Contrella\supit{a}, Talissa A. Altes\supit{a}, Brian B. Avants\supit{b}, Eduard E. de Lange\supit{a}, John P. Mugler III\supit{a}
\skiplinehalf
\supit{a}Dept. of Radiology and Medical Imaging, Univ. of Virginia, Charlottesville, Virginia, USA\\
\supit{b}PICSL, Univ. of Pennsylvania, Philadelphia, Pennsylvania, USA
}

%>>>> Further information about the authors, other than their 
%  institution and addresses, should be included as a footnote, 
%  which is facilitated by the \authorinfo{} command.

\authorinfo{Further author information: (Send correspondence to N.J.T.)\\
            N.J.T.: E-mail: njt4n@virginia.edu, Telephone: 1 434 924 7730}
%%>>>> when using amstex, you need to use @@ instead of @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

  \begin{document} 
  \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}

The utility of pulmonary functional imaging techniques, such as
hyperpolarized 3He MRI, has encouraged their inclusion in research
studies for longitudinal assessment of disease progression and the study of 
treatment effects.  We present methodology for performing voxelwise
statistical analysis of ventilation maps derived from 
hyperpolarized 3He MRI which incorporates multivariate template
construction using simultaneous acquisition of 1H and
3He images.  Additional processing steps include intensity normalization,
bias correction,
4-D longitudinal segmentation, and generation of expected ventilation
maps prior to voxelwise regression analysis.  Analysis is demonstrated 
on a cohort of eight individuals with diagnosed cystic fibrosis (CF) 
undergoing treatment imaged five times every two weeks with a prescribed
treatment schedule.
\end{abstract}

\keywords{hyperpolarized 3He, pulmonary segmentation, regression analysis, ventilation defects}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}
\label{sec:intro}  % \label{} allows reference to this section

Due to the heterogeneous nature of pathologic lung processes, studies of progression and treatment are complex and best assessed through integrated longitudinal studies.  Many of the longitudinal pulmonary studies to date have used 
spirometry to evaluate lung function although imaging is rapidly being adopted given its increased
spatial sensitivity.  Over the past decade, hyperpolarized 3He gas has been successfully used to visualize ventilation defects in the lungs\cite{lange1999} with significant potential for computational analysis.\cite{tustison2011}  
Given the nature of longitudinal studies in which a single subject is imaged numerous times, we detail
a framework for aligning the subject's individual time point images to a common subject-specific 
reference space in preparation for
voxelwise statistical regression.  This, in turn, produces treatment effect maps which simultaneously condenses the amount of longitudinal information while retaining the spatial sensitivity of imaging.  These processes are illustrated with a cohort of eight individuals diagnosed with CF imaged every two weeks (five total time points) over the course of experimental treatment.  The scope of this work is limited to methodological description with detailed clinical hypotheses and findings reported elsewhere.\cite{altes2012}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MATERIALS AND METHODS}

\subsection{Image Analysis}

In our earlier work\cite{tustison2011}, we described a ventilation-based segmentation workflow applied to single time point 3He data and provided evidence of its excellent performance relative to trained human readers.  In contrast, this work is concerned with the additional complexity of tracking relative longitudinal changes over multiple image acquisitions within a single subject.   As a general overview, analysis of longitudinal data comprises the following basic steps:
\begin{itemize}
\item multimodal atlas construction for each subject (cf Fig. \ref{fig:template}), 
\item intensity normalization\cite{nyul2000} and bias correction\cite{tustison2010} across all time points,
\item ventilation-based segmentation,\cite{tustison2011} and
\item voxelwise regression analysis with the simplified hypothesis treatment effect (cf Fig. \ref{fig:corr}).
\end{itemize}
We briefly detail each step of the 4-D workflow below.  We also note that all software described in this work is available as open source from the ANTs repository.%
\footnote{
http://www.picsl.upenn.edu/ANTs
}

\subsubsection{Multivariate Template Construction}

\begin{figure}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=100mm]{H1He3Template.pdf}
\end{tabular}
\end{center}
\caption{Multivariate template construction using both 3He and 1H images.  To create a normalized subject-specific space for statistical analysis of the longitudinal data, a template is created from the 3He and 1H data from each of the five time points.   We show the alignment results of the simultaneous acquisition by overlaying the faux-color rendered 3He image over the grayscale 1H image.  The template is generated iteratively where the algorithm alternates between averaging the registered images and then registering each time point to the average image (i.e. template estimate).}
\label{fig:template}
\end{figure}

In order to perform voxelwise analysis on the longitudinal CF data described previously, we spatially normalize the 3He images to a common reference space (also known as a ÒtemplateÓ or ÒatlasÓ).  Although resources (both template construction algorithms and pre-constructed templates) are common in the neuroimaging community, they are less well-known in the pulmonary research literature.   %For example, one of the earliest and most widely used templates is the Talaraich atlas created by Dr. Jean Talairach (a French neurosurgeon) who identified landmarks on the post-mortem histological sections of a 60-year old French woman to define a universal, stereotactic cerebral space.  %The utility of a standardized anatomical coordinate system has prompted additional work such as the publicly available Montreal Neurological Institute brain template (Collins, Neelin et al. 1994) which is an average of hundreds of adult brain MRI all registered to the Talaraich atlas.
For the analysis in this work we generate subject-specific templates directly from the image data.  Given the variability in lung shape across populations and the lack of publicly available lung atlases, generating population- or subject-specific templates enhances the accuracy of various analyses including that which is reported in this work.   Applicable to pulmonary data is the template construction algorithm\cite{avants2010} applied to T1-weighted brain data.   However, the simultaneous acquisition of the 3He and 1H images lends itself to multimodal processing in which both modalities are used to simultaneously produce 3He and 1H templates.   This process is represented in Fig. \ref{fig:template} for a single subject.

Using the simultaneously acquired 3He and 1H data\cite{altes2012} for each of the five time points as input, generation of the subject-specific template(s) can be intuitively understood as iterating between averaging the current set of aligned images to create an estimate of the template and then registering the images to that template estimate (and then repeating for a given number iterations).  More technically, the SyN (Symmetric Normalization) pairwise registration algorithm\cite{avants2011} and an optimized Laplacian sharpening/averaging of the template estimate forms the core of what is denoted as the Symmetric Group Normalization (SyGN) algorithm\cite{avants2010} for template construction.   Given the set of five time point 1H/3He aligned image pairs from a single subject, denoted as $\left\{I_1^{\mathrm{1H}}, I_1^{\mathrm{3He}},\ldots,I_5^{\mathrm{1H}}, I_5^{\mathrm{3He}}\right\}$, template generation involves finding
the set of paired diffeomorphic transformations $\left\{\left(\phi_1,\phi_1^{-1}\right), \ldots,  \left(\phi_5,\phi_5^{-1}\right) \right\}$;
the optimal multimodal template appearance, $J^\mathrm{1H}$ and $J^\mathrm{3He}$; and 
the corresponding coordinate system (i.e. template), $\psi(\mathbf{x})$,
which minimize the cost function
\begin{align}
  \sum_{m=1}^5 \left[
  D\left(\psi(\mathbf{x},\phi_1^m(x,1)\right) + 
  \Pi_{1}\left(I_m^\mathrm{1H}\left(\phi_2^m (\mathbf{x},0.5) \right),J^\mathrm{1H}\left(\phi_1^m (\mathbf{x},0.5) \right)\right) +     
  \Pi_{2}\left(I_m^\mathrm{3He}\left(\phi_2^m (\mathbf{x},0.5) \right),J^\mathrm{3He}\left(\phi_1^m (\mathbf{x},0.5) \right)\right)   \right]
\end{align}
where $D$ is the diffeomorphic shape distance, $D\left( \phi(\mathbf{x},t_a), \phi(\mathbf{x},t_b) \right) = \int_{t_a}^{t_b} \|v(\mathbf{x},t\|_{L}dt$
dependent on the choice of linear operator, $L$, and $v$ is the velocity field, $v\left( \phi(\mathbf{x},t) \right) = \frac{d\phi(\mathbf{x},t)}{dt},\,\, \phi(\mathbf{x},0)=\mathbf{x}$.
Applied to each subject, atlas generation produced subject-specific 1H and 3He template images which constitute a mean shape/appearance model of the image data for all 5 time points.  In addition, each time point image was spatially normalized to the template space in preparation for longitudinal ventilated-based segmentation and subsequent statistical analysis as detailed below.

\subsubsection{Longitudinal ventilation-based segmentation}

Following template construction, preprocessing includes suppression of inhomogeneity artifacts through the application of N4 bias correction\cite{tustison2010} to each individual image.  This step is followed by intensity normalization\cite{nyul2000} where 
the intensities of all 5 time point images are normalized to a single intensity range (only 5 match points were employed to avoid excessive intensity distortion).  Following intensity correction and normalization, the image volumes from each time point are collated into a single spatio-temporal image which is then segmented into four ventilation classes using the Atropos segmentation tool.\cite{tustison2011}

\subsection{Statistical Processing}

The probabilistic formulation of the Atropos segmentation yields, in addition to the 4-class hard segmentation, a set of four posterior probability maps indicating the maximum a posteriori estimate for each class at each voxel.  These MAP estimates are used to rescale all images to a standard scale denoted as ``expected ventilation'' images.   These rescaled expected ventilation (ev) images are calculated at each voxel, $i$, as follows:
\begin{align}
ev_i = \sum_{l=1}^L l \times \mathrm{Pr}(l|I_i)
\end{align}
where $L$ is the total number of labels (as described above, we use the label set $l \in \{1,2,3,4\}$ to describe the
ventilation classes) and $\mathrm{Pr}(l|I_i)$ is the posterior probability at voxel $i$ (i.e. the probability that voxel $i$ has label $l$ given its intensity value $I$).  The derived expected ventilation maps are then smoothed to reduce slight misalignments.  Voxelwise expected ventilation values are then correlated with the simplified treatment hypothesis (each subject is administered a series of placebos and actual treatments which are hypothesized to result in improved ventilation according to the blue dashed line on the right in Fig. \ref{fig:corr}).

\begin{figure}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=100mm]{correlation.pdf}\\
\end{tabular}
\end{center}
\caption{Voxelwise regression analysis to determine image-based response to treatment.  Treatment effects are expected to follow the simplified treatment hypothesis illustrated with the dashed blue line in the plot on the right.  To explore how the longitudinal change in expected ventilation follows this treatment hypothesis with image data, we smooth the aligned expected ventilation maps (to account for potential voxelwise misalignments) and then quantify how the voxelwise intensities regress with the simplified treatment hypothesis. }
\label{fig:corr}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RESULTS}

\begin{figure}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=170mm]{3Drenderings.pdf}\\
\end{tabular}
\end{center}
\caption{Resulting signed magnitude maps for all eight subjects showing 
regions of significant positive (orange) and negative (blue) correlations
(where $|$correlation$|$ $\geq 0.50$ and $|ev_{day3} - ev_{day2}| \geq 1.0$) with expected treatment hypothesis.  
}
\label{fig:maps}
\end{figure}

The previously described analysis protocol was applied to each of the eight subjects
to determine spatial response to treatment and subsequently rendered in Fig. \ref{fig:maps}.
Regions of positive correlation (i.e. regions which responded positively to treatment as 
expected) are shown in orange whereas regions of negative correlation (i.e. regions
which became less ventilated during actual treatment) are rendered in blue.  The absolute
correlation threshold of these regions was 0.5.  Additionally, only those regions which 
demonstrated a magnitude expected ventilation value $\geq 1.0$ between days 2 and 3
were rendered.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS}

Increasing use of imaging for pulmonary longitudinal studies will require computational
techniques for quantifying and visually presenting the data which facilitates
exploration of fundamental clinical hypotheses.  We presented a framework 
for data preparation in single subject longitudinal studies whereby multivariate template
construction results in longitudinal alignment.  This permits such analysis as 
voxelwise statistical regression for determining spatial correlation with expected treatment 
effects.  This provides  a unique visualization technique for exploring clinical
pulmonary hypotheses.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\acknowledgments     %>>>> equivalent to \section*{ACKNOWLEDGMENTS}       
% 
%This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{references}   %>>>> bibliography data in report.bib
\bibliographystyle{spiebib}   %>>>> makes bibtex use spiebib.bst

\end{document}

### 2013-02-08 SPIE2013/VBMCircularity/article.tex
ENH:  Updating.

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage{amsmath}

\title{Statistical bias in optimized VBM}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Nicholas J. Tustison\supit{a}, Brian B. Avants\supit{b}, Philip A. Cook\supit{b}, James C. Gee\supit{b}, and James R. Stone\supit{a}
\skiplinehalf
\supit{a}Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, Virginia, USA\\
\supit{b}Penn Image Computing and Science Laboratory, University of Pennsylvania, Philadelphia, Pennsylvania, USA
}

%>>>> Further information about the authors, other than their 
%  institution and addresses, should be included as a footnote, 
%  which is facilitated by the \authorinfo{} command.

\authorinfo{Further author information: (Send correspondence to N.J.T.)\\
            N.J.T.: E-mail: njt4n@virginia.edu, Telephone: 1 434 924 7730}
%%>>>> when using amstex, you need to use @@ instead of @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

  \begin{document} 
  \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
The recent discovery of methodological flaws in experimental 
design and analysis in neuroscience research has raised concerns 
over the validity of certain 
techniques used in routine analyses and their corresponding 
findings. 
Such concerns have centered around selection bias whereby data is
inadvertently manipulated such that the resulting analysis produces falsely 
increased statistical significance, i.e. type I errors. This has been 
illustrated recently in fMRI studies \cite{vul2009}, with excessive flexibility 
in data collection \cite{simmons2011}, and general experimental design 
issues \cite{kriegeskorte2009}.  Current work from our group has shown
how this problem extends to generic voxel-based analysis (and certain 
technique derivatives such as tract-based spatial statistics\cite{Smith2006}) 
using fractional anisotropy images derived from diffusion tensor imaging \cite{tustison2012}.
In this work, we demonstrate how this circularity principle 
can potentially extend to the well-known
optimized voxel-based morphometry technique \cite{Good2001} for assessing
cortical density differences whereby the principal cause of experimental 
corruption is due to normalization strategy.  Specifically, the popular sum-of-squared-differences (SSD) metric
explicitly optimizes statistical findings potentially inflating type I errors.  
Additional experimentation demonstrates that 
this problem is not restricted to the SSD metric but extends to other commonly
used metrics such as mutual information, neighborhood cross correlation, and
Demons. 
%In the spirit of reproducibility, all experiments were carried out using
%the open source Advanced Normalization Tools (ANTs) software package and 
%a subset of the publicly available IXI dataset.
\end{abstract}

%TThe recent discovery of methodological flaws in experimental design and analysis in neuroscience research has raised concerns over the validity of certain techniques used in routine analyses and their corresponding findings.  Such concerns have centered around selection bias whereby data is inadvertently manipulated such that the resulting analysis produces falsely increased statistical significance, i.e. type I errors. This has been illustrated recently in fMRI studies, with excessive flexibility in data collection, and general experimental design issues.   We demonstrate how this circularity principle  can potentially extend to the well-known optimized voxel-based morphometry technique.

%>>>> Include a list of keywords after the abstract

\keywords{circularity, cortical thickness, normalization}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}
\label{sec:intro}  % \label{} allows reference to this section

Voxel-based morphometry (VBM) \cite{wright1995,Ashburner2000} has proven to be
an invaluable technique for the characterization of cortical structural differences 
between populations using neuroimaging data.  Subsequent work
using VBM has explored fundamental neuroscience hypotheses and
has, as a result, become a vital analysis technique in the
neuroscientist's computational toolbox.  
%The core technique of voxelwise comparisons
%in the gray matter between populations has revealed 
%such interesting insights as the relationship between anatomical changes and neuroplasticity 
%\cite{draganski2004},
%the degeneration associated with various neuropathologies \cite{tedeschi2009},
%and other clinically relevant correlations \cite{apkarian2004}.
Briefly, as originally proposed, the standard VBM preprocessing protocol aligns each image to a
T1-weighted template prior to probabilistic segmentation which delineates the white matter, gray matter, 
and cerebrospinal fluid image regions.  Spatial normalization is typically performed on
each subject's T1-weighted image using a global affine normalization procedure to the template
to account for global shape differences followed by deformable registration
to minimize local anatomical differences.  A modification of the standard VBM technique, 
known as ``optimized VBM'', was introduced with the work of Good et al.\cite{Good2001}.  
Instead of spatial normalization to the T1-weighted brain template, an initial 
segmentation occurs in the space of each subject.   The gray matter probability images are then 
normalized to a gray matter
probabilistic template.%
\footnote{
Note that these steps are also applicable to analysis of the white matter.
}  
The normalization parameters are then applied to the corresponding T1-weighted brain 
images.  The remaining portion of the workflow mirrors standard VBM.

The crucial distinction between the two methods is the normalization step and the images
used to find the transformation between each subject and the template. 
Implicitly quantifying the anatomical differences for many implementations
is the SSD metric.
% which is easily recognized as:
%\begin{align}
%  SSD(\mathcal{I},\mathcal{J},\mathcal{T}) = \frac{1}{N}\sum_{n=1}^N (\mathcal{I}(\mathcal{T}(x_n))-\mathcal{J}(x_n))^2
%\end{align}  
%where $\mathcal{I}$ is the moving image, $\mathcal{J}$ is the fixed image, 
%$\mathcal{T}$ is the transformation to be optimized, and $n$ indexes over the 
%region of interest.  
In fact, the SSD is the metric originally proposed in both methods
and is used in popular, publicly available methods such as SPM2%
\footnote{
http://www.fil.ion.ucl.ac.uk/spm/software/spm2/
}
and FSL-VBM.%
\footnote{
http://www.fmrib.ox.ac.uk/fsl/fslvbm/index.html
}

Closer inspection of the optimization problem of finding the 
set of transformations mapping the entire set of images to the gray
matter template using SSD reveals the circularity of the optimized approach.  
Optimization entails finding the set of $M$ optimal transforms, $\mathcal{T}^*
=\{\mathcal{T}_1, \ldots, \mathcal{T}_M\}$ which maps the $M$ gray 
matter probability images  $\{\mathcal{I}^{gm}_1, \ldots, \mathcal{I}^{gm}_M\}$
to the gray matter template $\mathcal{J}^{gm}$, i.e. 
\begin{align}
  \mathcal{T}^* = \underset{\mathcal{T}_1, \ldots, \mathcal{T}_M}{\operatorname{argmin}}\frac{1}{N}\sum_{m=1}^{M}\sum_{n=1}^N \left(\mathcal{I}^{gm}(\mathcal{T}_m(x_n))-\mathcal{J}^{gm}(x_n)\right)^2.
\end{align}
Switching the summations, 
\begin{align}
  \label{eq:circularity}
  \mathcal{T}^* = \underset{\mathcal{T}_1, \ldots, \mathcal{T}_M}{\operatorname{argmin}}\sum_{n=1}^N 
  \underbrace{\frac{1}{N}\sum_{m=1}^{M}(\mathcal{I}^{gm}\left(\mathcal{T}_m(x_n))-\mathcal{J}^{gm}(x_n)\right)^2}_{\approx average\,\,voxelwise\,\,variance},
\end{align}
 and recognizing that the template gray matter probability
image is a good approximation of the average gray matter probability image, it becomes
apparent that the optimization intended to minimize local shape differences instead, and
more problematically, explicitly minimizes the average variance which directly 
maximizes Student's $t$-test.  Thus, standard preprocessing normalization 
in optimized VBM, instead of maximizing anatomical alignment,  directly manipulates the 
data such that the statistical testing results are maximized which, by 
definition\cite{kriegeskorte2009}, is circular.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MATERIALS AND METHODS}

\begin{figure}
\begin{center}
\begin{tabular}{ccc}
  \includegraphics[width=50mm]{atlas.png}&
  \includegraphics[width=50mm]{atlas_gm.png}&
  \includegraphics[width=50mm]{atlas_mask.png} \\
  (a) & (b) & (c)
\end{tabular}
\end{center}
\caption{(a) The T1 template created from the set of $56+29=85$ T1 IXI images used in this study.
(b) The gray matter probability template created from the gray matter probability maps.
(c) Mask region defining the domain of the VBM calculations (average gray matter probability $\geq 0.5$).
 }
\label{fig:template}
\end{figure}

Although Eq. (\ref{eq:circularity}) demonstrates the susceptibility 
of the SSD difference metric to producing false positives, we illustrate such 
effects on public data for mutual information (MI), neighborhood cross
correlation (CC), Demons, and SSD.  The relationship of the Demons metric
to SSD would give prior indication that similar effects should 
be seen whereas the effect should be lesser for MI and CC given that the
similarity measures are less localized in addition to not being based on
absolute intensity differences.

\subsection{Images}

Although any of the several high quality public data sets could be used to demonstrate the
presence of circular analysis in optimized VBM, we use the IXI data set consisting
of approximately 600 subjects from three different locations.  Demographic
information including gender and age is also available for each subject.  Further
details can be found on the IXI website.%
\footnote{
http://biomedic.doc.ic.ac.uk/brain-development/
}
For the study in the paper, we used the T1 images from two male subgroups in the age
ranges of 20 to 30 years and 60 to 80 years.  The former subgroup consists of 29 total
images and the latter consists of 56 images.

\subsection{Generation of Aligned Cohort}

The combined set of T1 images were first used to create an average unbiased shape and 
intensity atlas\cite{Avants2010} (see Fig. \ref{fig:template}(a)). This process also produced a set of transformations
which warps each image to the template space.  Each T1 was then segmented in its native
space using the Atropos\cite{Avants2011a} tool in ANTs. The resulting gray matter
probability maps were then warped to the template space where they were averaged
to create an average gray probability map (denoted as the ``gray matter probability
template'' shown in Fig. \ref{fig:template}(b)).   The two aligned cohorts
were then used to create two sets
of  simulated, aligned gray matter probability maps reflecting the characteristic 
intensities of group. These probability maps were created by calculating the voxelwise
mean and standard deviation of the aligned gray matter probability maps within a
given cohort. The voxel intensity value of the simulated gray matter probability was
then randomly generated using these voxelwise Gaussian parameters.  For our
experiments, we generated 21 synthetic gray matter probability maps from the 20-to-30
group and 20 synthetic gray matter probability maps from the 60-to-80 group.

Using the four similarity metrics described above, we then spatially normalized
each synthetic gray matter probability map to the gray matter probability
template.%
\footnote{
We used the \texttt{antsRegistration} tool also available in ANTs.  Specific common command line
parameters were: \texttt{--convergence [10,0,10] --shrink-factors 1 --smoothing-sigmas 0
--transform SyN[0.5,3,0]}.
}
Despite the fact that each synthetic image was generated in situ 
in the coordinate system of the template, according to the discussion above
in deriving Eq. (\ref{eq:circularity}), optimizing the different similarity metrics will
generate different alignments yielding different VBM results with the SSD and Demons
metric falsely increasing the volume of statistically significant regions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RESULTS}

Following basic VBM protocol, voxelwise $t$-tests were performed between each
synthetic group over all four metrics and for different levels of smoothing
(0 mm, 4 mm, 8 mm, and 16 mm given in full width at half maximum (FWHM)).  
We then thresholded the resulting $p$-values at $p \leq 0.05$ to yield 
significant regions.  We then calculated the volume of significant regions
for each metric case and compared the volume to the results of
no normalization with the same smoothing level.  The results are shown in 
Table \ref{table:results}.  Positive values are indicative of 
increased statistical significance after spatial normalization.
Whereas the less localized metrics which are not dependent on absolute
intensity differences, i.e. CC and MI, demonstrate a decrease in statistical
significance with normalization, the Demons and SSD metrics 
trend towards a substantial increase in statistically significant volume
as predicted by Eq. (\ref{eq:circularity}).

\begin{table}
\begin{center}
\caption{Change in volume of significant regions relative to 
initial alignment}
\label{table:results}
\begin{tabular*}{0.75\textwidth}{@{\extracolsep{\fill}}ccccc}
{} & 0 mm & 4 mm & 8 mm & 16 mm\\
\hline
MI     & -62\% & -87\% & -91\% & -90\% \\
CC     & -46\% & -58\% & -46\% & -33\% \\
SSD    &  26\% &  14\% & -8\%  &  10\% \\
Demons &  16\% &  27\% & -3\%  &  34\% \\
\hline
\end{tabular*}
\end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS}

Spatial normalization is a crucial preprocessing step in many fundamental 
analysis techniques, including optimized VBM.  However, as we point out,
certain normalization strategies (which are commonly used in various 
analyses) are prone to statistical bias by falsely elevating statistical 
significance.  As we have shown, certain metrics such as Demons and SSD,
explicitly optimize the average voxelwise $t$-statistic instead of maximizing
anatomical alignment.  We encourage researchers to reexamine their own
workflow to avoid such problematic practices.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Given that the Student's $t$-statistic is estimated from both the group variance
and the difference in group means, an immediate question might concern
the possibility of a concomitant decrease in group mean difference thereby
negating the decrease in group difference caused by group normalization.  
It is important to 
note that it is the {\em average} group variance that goes down 
 and not necessarily the variance at a single voxel site.  Additionally, we can
look at the average group mean difference, $\overline{\Delta\mu}$, over the entire image domain of interest
and manipulate the summation ordering which yields
\begin{align}
  \overline{\Delta\mu}&=\frac{1}{N}\sum_{n=1}^N \left( \mu_1(x_n) - \mu_2(x_n) \right)  \\
  &=\frac{1}{N}\sum_{n=1}^N \left( \frac{1}{M_1}\sum_{m_1=1}^{M_1} \mathcal{I}^{gm}_{m_1}(T_{m_1}(x_n)) - \frac{1}{M_2}\sum_{m_2=1}^{M_2} \mathcal{I}^{gm}_{m_2}(T_{m_2}(x_n)) \right) \\
  &=\left( \frac{1}{M_1}\sum_{m_1=1}^{M_1} \frac{1}{N}\sum_{n=1}^N \mathcal{I}^{gm}_{m_1}\left(T_{m_1}(x_n)\right)\right)
   - \left(\frac{1}{M_2}\sum_{m_2=1}^{M_2} \frac{1}{N}\sum_{n=1}^N \mathcal{I}^{gm}_{m_2}\left(T_{m_2}(x_n)\right)\right).
\end{align}
The inner summations in the two terms make apparent that the average group mean 
difference is dependent solely on how the pixel intensity values are mapped 
within the coordinate space defined by the template. In other words,
$\overline{\Delta\mu}$ depends on the difference between the aggregate intensities
of group 1 and group 2 which is not being directly optimized via minimization of the 
similarity metric (unlike the average group variance). For example, if one combined 
an intensity-preserving transformation 
model with the SSD metric, although the average variance is guaranteed to decrease, the
average group mean difference will remain constant throughout the spatial normalization optimization.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\acknowledgments     %>>>> equivalent to \section*{ACKNOWLEDGMENTS}       
% 
%This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{references}   %>>>> bibliography data in report.bib
\bibliographystyle{spiebib}   %>>>> makes bibtex use spiebib.bst

\end{document}

### 2013-02-08 WBIR2012/paper.tex
ENH:  Updating.

@@
 \author{Nicholas J. Tustison\inst{1} \and Brian B. Avants\inst{2}}
-%
-\authorrunning{N. Tustison et al.} % abbreviated author list (for running head)
-%
 
+\authorrunning{N. J. Tustison and B. B.  Avants} % abbreviated author list (for running head)
 \institute{University of Virginia, Charlottesville VA 22903, USA,\\
@@
 The advancements of the proposed framework
-over previous work include a preconditioned gradient descent algorithm, 
-potential spatial weighting 
-of the gradient permitting (among other things) enforcement of 
-stationary boundary conditions, multiresolution in terms of the B-spline
-mesh, facilitating the incorporation of landmark information, 
-and the possibility of modeling temporal periodicity.  In addition to
+over previous work include a preconditioned gradient descent algorithm and 
+potential weighting 
+of the metric gradient permitting, among other things, enforcement of 
+stationary boundary conditions.  In addition to
 theoretical and practical discussions of our contribution, we also describe 
-its parallelized implementation as open source in the Insight Toolkit.  
+its parallelized implementation as open source in the Insight Toolkit and
+perform an evaluation on publicly available brain data.
 \keywords{B-splines, DMFFD, diffeomorphisms, image registration, ITK}
@@
 variant extensions, the 
-{\em directly manipulated free-form deformation} approach  
- addressed the hemstitching issue associated with the
-problematic energy topographies caused by the distribution of the uniform 
-B-spline shape functions over the transformation domain \cite{tustison2009}.
+{\em directly manipulated free-form deformation} approach  \cite{tustison2009} 
+ addressed the hemstitching issue associated with steepest descent traversal of
+problematic energy topographies during the course of optimization.
+%caused by the distribution of the uniform 
+%B-spline shape functions over the transformation domain
 
@@
 We also incorporate B-spline mesh multi-resolution capabilities
-for increased control during registration progress.  
+for increased control during registration progression.  
 Most importantly, we also describe
-our parallelized algorithmic implementation as open source available through the Insight Toolkit.%
+the parallelized algorithmic implementation as open source available through the Insight Toolkit.%
 \footnote{
@@
 of the various algorithmic elements coupled with implementation details
-where appropriate.  This is followed by evaluation on a public data set.
+where appropriate.  This is followed by an evaluation on publicly available
+brain data.
 
@@
 velocity fields can be used to produce diffeomorphisms through
-analytical integration involving the B-spline basis functions. 
+numerical integration involving the B-spline basis functions. 
 We then show how our previous work involving optimization in 
@@
 \end{align}
-
 In the case of 
@@
 \end{align}  
-which provides a less computationally expensive, more stable alternative than other numerical
-methods.
+which provides a more stable and reliable alternative than other numerical
+methods \cite{press1996}.
 
@@
 well-understood preconditioned gradient was proposed based on the work 
-described in \cite{tustison2006} which substantially 
-improves performance which we refer to as DMFFD image registration.  
-Similarly, we propose the following velocity field control
-point lattice preconditioned gradient,
+described in \cite{tustison2006} which we denote as directly manipulated
+free-form deformation (DMFFD) image registration.  This alternative demonstrates 
+improved performance over the steepest descent equivalent \cite{tustison2009}.
+
+Similarly, for the time-varying velocity field case we propose the following preconditioned gradient,
 $\delta v_{i_1,\ldots,i_d,i_t}$, given the similarity metric, $\Pi$,%
@@
 Current options include neighborhood cross correlation (CC), mutual information (MI), and
-Demons-style sum of squared differences (Demons).
+sum of squared differences (SSD).
 }
@@
 \footnote{
-Spline orders can be specified separately for each dimension but, for simplicity,
+In terms of implementation spline orders can be specified separately for each dimension but, for simplicity,
 we only specify a single spline order.  
@@
 gradient sample.
-Combining DMFFD optimization with the B-spline velocity field transform discussed in
-the previous subsection, we outline the iterative routine for optimizing the 
-velocity field control in Algorithm \ref{alg1}.
 
-\begin{algorithm}                    
-\caption{Gradient descent optimization of the B-spline velocity field}        
-\label{alg1}      
-\begin{algorithmic}[1]
-\REQUIRE images $\mathcal{I}$ and $\mathcal{J}$
-\REQUIRE  user-specified similarity metric $\Pi$ 
-\REQUIRE  velocity field mesh size and multiresolution schedule, $R$
-\REQUIRE  number of time point samples, $N_t$ (default: 4)
-\REQUIRE  spline order (default: 3)
-\REQUIRE  gradient step size, $\lambda$ (default: 0.25)
-\ENSURE velocity field control point lattice, $v$
-\STATE $v \gets \mathbf{0}$ \COMMENT{Initialize velocity field}
-\FOR{number of resolution levels} 
-\FOR{number of iterations for current level} 
-  \STATE $G \gets []$  \COMMENT{Initialize dense gradient storage array}
-\FOR{$t = 1 \to N_t$ } 
-  \STATE $t' \gets \frac{t-1}{N_t-1}$  \COMMENT{Scale time point to $[0,1]$}
-  \STATE $\mathcal{I}' \gets \mathcal{I} \circ \phi(\mathbf{x}, t')$  \COMMENT{Warp $\mathcal{I}$ to the current time point}
-  \STATE $\mathcal{J}' \gets \mathcal{J} \circ \phi^{-1}(\mathbf{x}, 1-t')$ \COMMENT{Warp $\mathcal{J}$ to the current time point}
-  \STATE $G[t] \gets \frac{\partial \Pi(\mathcal{I}',\mathcal{J}')}{\partial \mathbf{x}}$  \COMMENT{Store dense similarity metric gradient}
-\ENDFOR  
-  \STATE $\delta v \gets B(G)$ \COMMENT{Calculate velocity lattice gradient (Eqn. (\ref{eq:dmffd}))}
-  \STATE $v \gets v + \lambda\delta v$ \COMMENT{Take a step in the gradient direction}
-\ENDFOR  
-  \STATE $v \gets R(v)$         \COMMENT{Refine lattice based on schedule}
-%  \IF {converged}
-%    \RETURN $v$
-%  \ENDIF
-\ENDFOR  
-\end{algorithmic}
-\end{algorithm}
+Additionally, in \cite{tustison2006} it was shown that one could associate each 
+metric gradient sample, $\left( \frac{\partial \Pi}{\partial \mathbf{x}}\right)_c$ with a confidence
+weighting.  Thus, in order to enforce stationary boundaries, we 
+assign image boundary metric gradients a value of zero with a corresponding
+large confidence value.
+
+%Combining DMFFD optimization with the B-spline velocity field transform discussed in
+%the previous subsection, we outline the iterative routine for optimizing the 
+%velocity field control points in Algorithm \ref{alg1}.
+
+%\begin{algorithm}                    
+%\caption{Gradient descent optimization of the B-spline velocity field}        
+%\label{alg1}      
+%\begin{algorithmic}[1]
+%\REQUIRE images $\mathcal{I}$ and $\mathcal{J}$
+%\REQUIRE  user-specified similarity metric $\Pi$ 
+%\REQUIRE  velocity field mesh size and multiresolution schedule, $R$
+%\REQUIRE  number of time point samples, $N_t$ (default: 4)
+%\REQUIRE  spline order (default: 3)
+%\REQUIRE  gradient step size, $\lambda$ (default: 0.25)
+%\ENSURE velocity field control point lattice, $v$
+%\STATE $v \gets \mathbf{0}$ \COMMENT{Initialize velocity field}
+%\FOR{number of resolution levels} 
+%\FOR{number of iterations for current level} 
+%  \STATE $G \gets []$  \COMMENT{Initialize dense gradient storage array}
+%\FOR{$t = 1 \to N_t$ } 
+%  \STATE $t' \gets \frac{t-1}{N_t-1}$  \COMMENT{Scale time point to $[0,1]$}
+%  \STATE $\mathcal{I}' \gets \mathcal{I} \circ \phi(\mathbf{x}, t')$  \COMMENT{Warp $\mathcal{I}$ to the current time point}
+%  \STATE $\mathcal{J}' \gets \mathcal{J} \circ \phi^{-1}(\mathbf{x}, 1-t')$ \COMMENT{Warp $\mathcal{J}$ to the current time point}
+%  \STATE $G[t] \gets \frac{\partial \Pi(\mathcal{I}',\mathcal{J}')}{\partial \mathbf{x}}$  \COMMENT{Store dense similarity metric gradient}
+%\ENDFOR  
+%  \STATE $\delta v \gets B(G)$ \COMMENT{Calculate velocity lattice gradient (Eqn. (\ref{eq:dmffd}))}
+%  \STATE $v \gets v + \lambda\delta v$ \COMMENT{Take a step in the gradient direction}
+%\ENDFOR  
+%  \STATE $v \gets R(v)$         \COMMENT{Refine lattice based on schedule}
+%%  \IF {converged}
+%%    \RETURN $v$
+%%  \ENDIF
+%\ENDFOR  
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 
-As mentioned previously, the registration algorithm has been implemented in the
+As mentioned previously, the registration algorithm has been implemented 
+and introduced into the
 Insight Toolkit and consists of the following major classes: 
 \begin{itemize}
-  \item \verb#itk::TimeVaryingVelocityFieldIntegrationImageFilter#
-  \item \verb#itk::TimeVaryingBSplineVelocityFieldTransform#
-  \item \verb#itk::TimeVaryingBSplineVelocityFieldImageRegistrationMethod#
-  \item \verb#itk::TimeVaryingBSplineVelocityFieldTransformParametersAdaptor#
-  \item \verb#itk::BSplineScatteredDataPointSetToImageFilter#
-\end{itemize}
-The \verb#TimeVaryingVelocityFieldIntegrationImageFilter# class implements
-the Runge-Kutta integration described earlier.  Given a 
-velocity field control point lattice as 
+  \item \verb#itk::TimeVaryingVelocityFieldIntegrationImageFilter# ---
+Implementation of 
+the Runge-Kutta integration.  Given a 
+sampled velocity field derived from the control point lattice as 
 input and the lower and upper integration limits, integration is performed 
 in a multi-threaded fashion (since each point in the domain
-can be integrated separately).  The \verb#TimeVaryingBSplineVelocityFieldTransform# 
-is derived from the base transform class which handles the mapping of geometric primitives
-for warping images.  The DMFFD gradient calculation  is handled by the class \verb#BSplineScatteredDataPointSetToImageFilter# (cf. Eqn. \ref{eq:dmffd} and line 11
-of Algorithm \ref{alg1}) which takes as input the dense similarity metric and an optional
-weighting for each gradient sample.  This permits enforcement of stationary physical
-boundaries by specifying a zero gradient on the boundaries and a large weighting.
-Coordinating all the elements of image registration is the
-\verb#TimeVaryingBSplineVelocityFieldImageRegistrationMethod# class which
-encapsulates Algorithm \ref{alg1}.  The resolution scheduling is determined
-by the \verb#TimeVaryingBSplineVelocityFieldTransformParametersAdaptor# class.
-We provide access to the new ITK 
+can be integrated separately).  
+  \item \verb#itk::TimeVaryingBSplineVelocityFieldTransform#:
+Handles the mapping of geometric primitives
+using the transform described in this work.  It is defined by the velocity field
+control point lattice which is sampled prior to being integrated to yield the 
+resulting displacement field.  
+  \item \verb#itk::TimeVaryingBSplineVelocityFieldImageRegistrationMethod#: 
+  Coordinates the optimization
+  using gradient descent given an user-specified number of resolution levels, shrink
+  factors, and smoothing choices.  
+  \item \verb#itk::TimeVaryingBSplineVelocityFieldTransformParametersAdaptor#:
+  Handles the decomposition of the B-spline mesh resolution during the multi-resolution optimization
+  given a user-specified scheduling. 
+  \item \verb#itk::BSplineScatteredDataPointSetToImageFilter#:
+  Calculates the similarity metric gradient (cf. Eqn. \ref{eq:dmffd}) which takes as input the 
+  dense similarity metric gradient and an optional weighting for each gradient sample.  As 
+  described earlier, this permits enforcement of stationary physical boundaries. % by specifying a zero gradient on the boundaries and a large weighting. 
+\end{itemize}
+
+Access to the new ITK 
 registration framework (including the B-spline velocity field transform) 
-through the command line module \verb#antsRegistration# available both in ANTs%
+is facilitated
+via the command line module \verb#antsRegistration# available both in ANTs%
 \footnote{
@@
 }
-The interested reader should consult \cite{tustison2012} for further details on 
-actual usage.
+%The interested reader should consult \cite{tustison2012} for further details on 
+%actual usage.
 
@@
 
-Batch registration was facilitated by the inclusion of the proposed
-method in our ANTs interface to the new Insight Toolkit registration
-framework called \verb#antsRegistration# which is detailed in 
-\cite{tustison2012}.  An active development version is found in 
-the ANTS Sourceforge repository.%
-\footnote{
-http://sourceforge.net/projects/advants/
-}
+%Batch processing was facilitated by the inclusion of the proposed
+%method in the ANTs interface to the new Insight Toolkit registration
+%framework called \verb#antsRegistration# which is detailed in 
+%\cite{tustison2012}.  An active development version is found in 
+%the ANTS Sourceforge repository.%
+%\footnote{
+%http://sourceforge.net/projects/advants/
+%}
+
 A portion of the batch script used to produce the registrations is 
@@
 Batch processing was performed on the linux cluster at the 
-University of Virginia with each job allocated 4 nodes for
+University of Virginia with allocation of 4 nodes for each job
 multi-threading purposes.  Each registration took approximately
@@
 The numerical values of this assessment are given in Table 1.
-These values were projected onto subject 1 for display in Fig. 1
+These values were projected onto LBPA40 Subject 1 for display in Fig. 1
 which can be compared to a similar display given in \cite{Klein2009}
@@
 periodicity in dealing with the possibility of multiple images describing
-periodic motion (e.g. cardiac or pulmonary motion).  Although not discussed
-in this work, a future publication will explore these possibilities in 
-greater detail.
+periodic motion (e.g. cardiac or pulmonary motion). A future publication 
+will explore these possibilities in greater detail.
 

### 2013-02-08 SPIE2013/KK/IXI_NKI_Oasis_combinedResults/agePredict.R

@@ -1,0 +1,113 @@
+demographics <- read.csv( file = "demographics.csv", header = TRUE );
+# volumes <- read.csv( file = "volume.csv", header = TRUE );
+thickness <- read.csv( file = "thickness.csv", header = TRUE );
+
+indices <- which( demographics$AGE >= 0 & demographics$AGE <= 100 );
+thickness <- thickness[indices,];
+demographics <- demographics[indices,];
+
+cortical_labels <- c( "L occipital", "R occipital",
+                     "L cingulate", "R cingulate",
+                     "L insula",    "R insula",
+                     "L temporal pole",   "R temporal pole",
+                     "L superior temporal", "R superior temporal",
+                     "L infero temporal", "R infero temporal",
+                     "L parahippocampal", "R parahippocampal",
+                     "L frontal pole",    "R frontal pole",
+                     "L superior frontal","R superior frontal",
+                     "L middle frontal",  "R middle frontal",
+                     "L inferior",        "R inferior",
+                     "L orbital frontal", "R orbital frontal",
+                     "L precentral",      "R precentral",
+                     "L superior parietal", "R superior parietal",
+                     "L inferior parietal", "R inferior parietal",
+                     "L postcentral",       "R postcentral" );
+
+nSamples <- 100;
+male_color = "darkred";
+female_color = "navyblue";
+
+age <- demographics$AGE;
+id <- demographics$ID;
+gender <- cut( demographics$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
+
+ixi <- indices[which( indices <= 360 )];
+nki <- indices[which( indices >= 361 & indices <= 537 )];
+oasis <- indices[which( indices >= 537 )];
+
+ixiAge <- age[ixi];
+nkiAge <- age[nki];
+oasisAge <- age[oasis];
+
+ixiGender <- gender[ixi];
+nkiGender <- gender[nki];
+oasisGender <- gender[oasis];
+
+
+
+sampled.th <- seq( min( thickness ), max( thickness ) , length = nSamples );
+sampled.age <- seq( 0, 100, length = nSamples );
+
+for( i in 1:32 )
+  {
[…]

### 2013-02-08 SPIE2013/KK/IXI_results/agePredict.R

@@ -1,0 +1,107 @@
+library( ggplot2 )
+
+demographics <- read.csv( file = "demographics.csv", header = TRUE );
+# volumes <- read.csv( file = "volume.csv", header = TRUE );
+thickness <- read.csv( file = "thickness.csv", header = TRUE );
+
+cortical_labels <- c( "L occipital", "R occipital",
+                     "L cingulate", "R cingulate",
+                     "L insula",    "R insula",
+                     "L temporal pole",   "R temporal pole",
+                     "L superior temporal", "R superior temporal",
+                     "L infero temporal", "R infero temporal",
+                     "L parahippocampal", "R parahippocampal",
+                     "L frontal pole",    "R frontal pole",
+                     "L superior frontal","R superior frontal",
+                     "L middle frontal",  "R middle frontal",
+                     "L inferior",        "R inferior",
+                     "L orbital frontal", "R orbital frontal",
+                     "L precentral",      "R precentral",
+                     "L superior parietal", "R superior parietal",
+                     "L inferior parietal", "R inferior parietal",
+                     "L postcentral",       "R postcentral" );
+
+nSamples <- 100;
+male_color = "darkred";
+female_color = "navyblue";
+
+age <- demographics$AGE;
+gender <- cut( demographics$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
+
+sampled.th <- seq( min( thickness ), max( thickness ) , length = nSamples );
+sampled.age <- seq( 20, 90, length = nSamples );
+
+for( i in 1:32 )
+  {
+  th <- thickness[,i];
+
+#   n <- lm( formula = age ~ th + gender );
+#   n <- lm( formula = age ~ md + th + I(th^2) + vo + th:vo + gender );
+#   n <- lm( formula = age ~ md + th + vo + gender );
+
+    n <- lm( formula = th ~ age + I(age^2) + gender + age:gender );
+
+#   coeffs <- coef( n );
+#   pvalues <- anova( n )[,"Pr(>F)"];
+
+  male.data <- list();
+  female.data <- list();
+
[…]

### 2013-02-08 SPIE2013/KK/IXI_results/plotResults.R

@@ -1,0 +1,374 @@
+# resampler <- function( sp )
+#   {
+#   n <- nrow( sp.frame )
+#   resample.rows <- sample( 1:n,size = n, replace = TRUE )
+#   return( sp.frame[resample.rows,] )
+#   }
+#
+# sp.spline.estimator <- function( x, y, m=300 )
+#   {
+#   # Fit spline to data, with cross-validation to pick lambda
+#   fit <- smooth.spline( x, y, cv = TRUE );
+#   # Set up a grid of m evenly-spaced points on which to evaluate the spline
+#   eval.grid <- seq( from = min( x ), to = max( x ), length.out = m )
+#     # Slightly inefficient to re-define the same grid every time we call this,
+#     # but not a big overhead
+#   # Do the prediction and return the predicted values
+#   return( predict( fit, x = eval.grid)$y )  # We only want the predicted values
+#   }
+#
+# sp.spline.cis <- function( B, alpha, m = 300 ) {
+#   spline.main <- sp.spline.estimator( sp.frame, m = m )
+#   # Draw B boottrap samples, fit the spline to each
+#   spline.boots <- replicate( B, sp.spline.estimator( sp.resampler(), m = m ) )
+#     # Result has m rows and B columns
+#   cis.lower <- 2 * spline.main - apply( spline.boots, 1, quantile, probs = 1 - 0.5 * alpha )
+#   cis.upper <- 2 * spline.main - apply( spline.boots, 1, quantile, probs = 0.5 * alpha )
+#   return( list( main.curve = spline.main, lower.ci = cis.lower, upper.ci = cis.upper,
+#               x = seq( from = min( sp.today ), to = max( sp.today ), length.out = m ) ) )
+# }
+
+calculate.spline <- function( spline.data, m = 300 )
+  {
+  fit <- smooth.spline( spline.data$x, spline.data$y, df = 3 );
+  res <- residuals( fit );
+  eval.grid <- seq( min( spline.data$x ), to = max( spline.data$x ), length.out = m );
+  return( list( x = eval.grid, y = predict( fit, x = eval.grid )$y, res = res ) );
+  }
+
+resampler <- function( spline.data )
+  {
+  n <- length( spline.data$x );
+  resample <- sample( 1:n, size = n, replace = TRUE );
+
+  resampled.data <- list();
+  resampled.data$x = spline.data$x[resample];
[…]

### 2013-02-08 SPIE2013/KK/IXI_results/predict_age.R

@@ -1,0 +1,48 @@
+library(MASS)
+library(bestglm)
+md<-read.csv('meandiffusion.csv')
+vol<-read.csv('volume.csv')
+th<-read.csv('thickness.csv')
+demog<-read.csv('demographics.csv')
+# remove some bad data
+whold<-( demog$AGE > 50 & demog$AGE > 1  )
+whmid<-( demog$AGE > 40 & demog$AGE < 60  )
+whyoung<-( demog$AGE < 41 & demog$AGE > 1  )
+wh<-whold
+print(sum(wh)/nrow(demog))
+demog<-demog[wh,]
+vol<-vol[wh,]
+th<-th[wh,]
+md<-md[wh,]
+# training / testing data
+subs1<-c(1:(nrow(demog)/2))
+subs1<-subs1*2
+subs2<-subs1-1
+subs2<-subs2[1:(length(subs2)-1)] ; length(subs2) ; length(subs1)
+# make data frame with all training data
+Xy1<-data.frame(vol=vol[subs1,],th=th[subs1,],sex=demog$SEX[subs1],age=demog$AGE[subs1])
+Xy2<-data.frame(vol=vol[subs2,],th=th[subs2,],sex=demog$SEX[subs2],age=demog$AGE[subs2])
+Xy1<-data.frame(th=th[subs1,],sex=demog$SEX[subs1],age=demog$AGE[subs1])
+names(Xy1)<-c(paste("X",1:(ncol(Xy1)-1),sep=""),"y")
+Xy2<-data.frame(th=th[subs2,],sex=demog$SEX[subs2],age=demog$AGE[subs2])
+names(Xy2)<-c(paste("X",1:(ncol(Xy2)-1),sep=""),"y")
+Xy<-Xy1
+ans<-bestglm(Xy, IC="CV" , t=125 )
+# ans<-bestglm(Xy, IC="CV", t=10,CVArgs=list(Method="d", K=2, REP=10))
+# now apply to test data
+Xy<-Xy2
+predictedage<-predict(ans$BestModel,newdata=Xy)
+realage<-Xy$y
+print(paste('Mean:',mean(abs(predictedage-realage))))
+ranger<-c( min( c(realage,predictedage) ),  max( c(realage,predictedage) ) )
+pdf( file = "brian_results.pdf", 8, 8 );
+plot( ranger, ranger, type = "n",
+    xlab = expression( paste( bold( "Real Age (years)" ) ) ),
+    ylab = expression( paste( bold( "Predicted Age (years)" ) ) ),
[…]

### 2013-02-08 SPIE2013/KK/IXI_results/predict_age2.R

@@ -1,0 +1,47 @@
+library( "earth" )
+
+
+get.used.pred.names <- function( obj ) # obj is an earth object
+{
+  any1 <- function(x) any(x != 0)    # like any but no warning if x is double
+  names(which(apply(obj$dirs[obj$selected.terms, , drop=FALSE], 2, any1)))
+}
+
+rm( list = ls( all = TRUE ) )
+
+meandiffusion <- read.csv( 'meandiffusion.csv', header = TRUE )
+volumes <- read.csv( 'volume.csv', header = TRUE )
+thickness <- read.csv( 'thickness.csv', header = TRUE )
+demographics <- read.csv( 'demographics.csv', header = TRUE )
+
+meandiffusion <- meandiffusion[demographics$SEX == 2,];
+volumes <- volumes[demographics$SEX == 2,];
+thickness <- thickness[demographics$SEX == 2,];
+demographics <- demographics[demographics$SEX == 2,];
+
+
+agePredictionData <- data.frame( md = meandiffusion, th = thickness );
+
+trainingIndices <- createDataPartition( demographics$AGE, p = 1/2, list = FALSE );
+
+agePredictionData.training <- agePredictionData[trainingIndices,];
+agePredictionData.testing <- agePredictionData[-trainingIndices,]
+
+fit <- earth( x = agePredictionData.training, y = demographics$AGE[trainingIndices], degree = 2, pmethod = "exhaustive"  );
+
+age.real <- demographics$AGE[-trainingIndices];
+age.predict <- predict( fit, agePredictionData.testing );
+
+pdf( file = "brian_results_model.pdf", 8, 8 );
+plot( c( 20, 90 ), c( 20, 90 ), type = "n",
+    xlab = expression( paste( bold( "Real Age (years)" ) ) ),
+    ylab = expression( paste( bold( "Predicted Age (years)" ) ) ),
+    frame.plot = FALSE  );
+lines( c( 20, 90 ), c( 20, 90 ), lwd = 3, lty = 3, col = "black" );
+title( main = paste( "Cortical thickness prediction of age"), col.main = "black", font.main = 2 );
+axis( 1, tick = TRUE, lwd = 2, font = 2 );
+axis( 2, tick = TRUE, lwd = 2, font = 2 );
+points( x = age.real, y = age.predict, pch = 16, lwd = 1.5, cex = 1.0, col = "darkred" );
+dev.off();
+
+print( mean( abs( age.real - age.predict ), na.rm = TRUE ) );

### 2013-02-08 SPIE2013/KK/IXI_results/predict_age3.R

@@ -1,0 +1,92 @@
+library( MASS )
+library( bestglm )
+library( caret )
+
+meandiffusion <- read.csv( 'meandiffusion.csv' )
+volumes <- read.csv( 'volume.csv' )
+thickness <- read.csv( 'thickness.csv' )
+demographics <- read.csv( 'demographics.csv' )
+
+# create the total model
+
+trainingIndices <- createDataPartition( demographics$AGE, p = 1/2, list = FALSE );
+
+training.data <- data.frame( th = thickness[trainingIndices,],
+#                              md = meandiffusion[trainingIndices,],
+                             sex = demographics$SEX[trainingIndices],
+                             age = demographics$AGE[trainingIndices] );
+testing.data <- data.frame( th = thickness[-trainingIndices,],
+#                             md = meandiffusion[-trainingIndices,],
+                             sex = demographics$SEX[-trainingIndices],
+                             age = demographics$AGE[-trainingIndices] );
+
+model.total <- bestglm( training.data, IC = "CV" );
+
+# create the three sub-models
+
+subjects <- list();
+subjects[[1]] <- ( demographics$AGE[trainingIndices] >= 20 & demographics$AGE[trainingIndices] <= 40 );
+subjects[[2]] <- ( demographics$AGE[trainingIndices] >= 40 & demographics$AGE[trainingIndices] <= 60 );
+subjects[[3]] <- ( demographics$AGE[trainingIndices] >= 60 );
+
+models.sub <- list();
+for( i in seq( 1, 3, by = 1 ) )
+  {
+  cat( "constructing model ", i, "\n", sep = "" );
+  dem <- demographics[subjects[[i]],];
+  vol <- volumes[subjects[[i]],];
+  th <- thickness[subjects[[i]],];
+  md <- meandiffusion[subjects[[i]],];
+
+  training.data <- data.frame( th = th,
+#                              md = md,
+                               sex = dem$SEX,
+                               age = dem$AGE );
+  models.sub[[i]] <- bestglm( training.data, IC = "CV", t = 125 );
+  }
+
+pdf( file = "brian_results_model.pdf", 8, 8 );
+plot( c( 20, 90 ), c( 20, 90 ), type = "n",
[…]

### 2013-02-08 SPIE2013/KK/IXI_results/predict_age4.R

@@ -1,0 +1,111 @@
+library( MASS )
+library( glmulti )
+library( caret )
+
+rm( list = ls( all = TRUE ) )
+
+meandiffusion <- read.csv( 'meandiffusion.csv' )
+volumes <- read.csv( 'volume.csv' )
+thickness <- read.csv( 'thickness.csv' )
+demographics <- read.csv( 'demographics.csv' )
+
+mylabels <- c( "whole", "L occipital", "R occipital",
+                     "L cingulate", "R cingulate",
+                     "L insula",    "R insula",
+                     "L temporal pole",   "R temporal pole",
+                     "L superior temporal", "R superior temporal",
+                     "L infero temporal", "R infero temporal",
+                     "L parahippocampal", "R parahippocampal",
+                     "L frontal pole",    "R frontal pole",
+                     "L superior frontal","R superior frontal",
+                     "L middle frontal",  "R middle frontal",
+                     "L inferior",        "R inferior",
+                     "L orbital frontal", "R orbital frontal",
+                     "L precentral",      "R precentral",
+                     "L superior parietal", "R superior parietal",
+                     "L inferior parietal", "R inferior parietal",
+                     "L postcentral",       "R postcentral" );
+
+
+wholeThickness <- rowSums( as.matrix( volumes ) * as.matrix( thickness ) ) / rowSums( volumes );
+wholeMeanDiffusion <- rowSums( as.matrix( volumes ) * as.matrix( meandiffusion ) ) / rowSums( volumes );
+
+accuracy <- rep( 0, 33 );
+for( n in 0:100 )
+  {
+  cat( "Run ", n, "\n", sep = "" );
+  trainingIndices <- createDataPartition( demographics$AGE, p = 1/2, list = FALSE )
+
+  training.data <- data.frame( sex = demographics$SEX[trainingIndices],
+                               age = demographics$AGE[trainingIndices],
+                               wth = wholeThickness[trainingIndices],
+                               th = thickness[trainingIndices,],
+                               wmd = wholeMeanDiffusion[trainingIndices],
+                               md = meandiffusion[trainingIndices,]
+                             );
+
[…]

### 2013-02-08 SPIE2013/KK/Kirby_results/agePredict.R

@@ -1,0 +1,57 @@
+library( ggplot2 )
+
+priorWeights <- c( "0.0", "0.1", "0.25", "0.5" )
+
+
+for( n in 1:length( priorWeights ) )
+  {
+  results <- read.csv( file = paste( "labelresults", priorWeights[n], ".csv", sep = '' ), header = TRUE );
+
+  cortical_labels <- c( "L occipital", "R occipital",
+                       "L cingulate", "R cingulate",
+                       "L insula",    "R insula",
+                       "L temporal pole",   "R temporal pole",
+                       "L superior temporal", "R superior temporal",
+                       "L infero temporal", "R infero temporal",
+                       "L parahippocampal", "R parahippocampal",
+                       "L frontal pole",    "R frontal pole",
+                       "L superior frontal","R superior frontal",
+                       "L middle frontal",  "R middle frontal",
+                       "L inferior",        "R inferior",
+                       "L orbital frontal", "R orbital frontal",
+                       "L precentral",      "R precentral",
+                       "L superior parietal", "R superior parietal",
+                       "L inferior parietal", "R inferior parietal",
+                       "L postcentral",       "R postcentral" );
+
+  male_color = "darkred";
+  female_color = "navyblue";
+
+  age <- results$AGE;
+  gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
+
+  for( i in 1:32 )
+    {
+    th <- results[,i+3];
+
+    gender[which( gender == 1 )] <- 'male';
+    gender[which( gender == 2 )] <- 'female';
+
+    plotData <- data.frame( cbind( Age = age, Thickness = th, Gender = gender ) )
+    plotData <- transform( plotData, Gender = factor( Gender ) );
+
+    thickPlot <- ggplot( plotData, aes( x = Age, y = Thickness, group = Gender ) ) +
+                 stat_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x, method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fullrange = FALSE, fill = 'black', alpha = 0.5 ) +
[…]

### 2013-02-08 SPIE2013/KK/NKI_results/agePredict.R

@@ -1,0 +1,85 @@
+demographics <- read.csv( file = "demographics.csv", header = TRUE );
+# volumes <- read.csv( file = "volume.csv", header = TRUE );
+thickness <- read.csv( file = "thickness.csv", header = TRUE );
+
+cortical_labels <- c( "L occipital", "R occipital",
+                     "L cingulate", "R cingulate",
+                     "L insula",    "R insula",
+                     "L temporal pole",   "R temporal pole",
+                     "L superior temporal", "R superior temporal",
+                     "L infero temporal", "R infero temporal",
+                     "L parahippocampal", "R parahippocampal",
+                     "L frontal pole",    "R frontal pole",
+                     "L superior frontal","R superior frontal",
+                     "L middle frontal",  "R middle frontal",
+                     "L inferior",        "R inferior",
+                     "L orbital frontal", "R orbital frontal",
+                     "L precentral",      "R precentral",
+                     "L superior parietal", "R superior parietal",
+                     "L inferior parietal", "R inferior parietal",
+                     "L postcentral",       "R postcentral" );
+
+nSamples <- 100;
+male_color = "darkred";
+female_color = "navyblue";
+
+age <- demographics$AGE;
+gender <- cut( demographics$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
+
+sampled.th <- seq( min( thickness ), max( thickness ) , length = nSamples );
+sampled.age <- seq( 20, 90, length = nSamples );
+
+for( i in 1:32 )
+  {
+  th <- thickness[,i];
+
+#   n <- lm( formula = age ~ th + gender );
+#   n <- lm( formula = age ~ md + th + I(th^2) + vo + th:vo + gender );
+#   n <- lm( formula = age ~ md + th + vo + gender );
+
+    n <- lm( formula = th ~ age + I(age^2) + gender + age:gender );
+
+#   coeffs <- coef( n );
+#   pvalues <- anova( n )[,"Pr(>F)"];
+
+  male.data <- list();
+  female.data <- list();
+
+  male.data$x <- data.frame( age = sampled.age, th = sampled.th, gender = rep( "male", nSamples ) );
[…]

### 2013-02-08 WBIR2012/WBIRStudy/extractRunsInformation.pl

@@ -1,0 +1,141 @@
+#! /usr/bin/perl/ -w
+
+my $baseDir = '/Users/ntustison/Data/Public/LONI/LPBA40/WBIRStudy/';
+my $dataDir = $baseDir . $ARGV[0];
+
+my @elapsedTimes = ();
+my @metricValues = ();
+
+my @metricFileNames = ( "${dataDir}/metricValuesLevel0.csv", "${dataDir}/metricValuesLevel1.csv",
+  "${dataDir}/metricValuesLevel2.csv", "${dataDir}/metricValuesLevel3.csv" );
+
+for( my $i = 0; $i < @metricFileNames; $i++ )
+  {
+  open( FILE, ">$metricFileNames[$i]" );
+  close( FILE );
+  }
+
+for( my $i = 1; $i <= 10; $i++ )
+  {
+  for( my $j = 1; $j <= 10; $j++ )
+    {
+    if( $i == $j )
+      {
+      next;
+      }
+
+    my $i_idx = sprintf( "%02d", $i );
+    my $j_idx = sprintf( "%02d", $j );
+
+    my $filename = "S${j_idx}toS${i_idx}_log.txt";
+
+    if( ! -e "${dataDir}/${filename}" )
+      {
+      next;
+      }
+
+    print "Doing ${filename}\n";
+
+    open( FILE, "<${dataDir}/${filename}" );
+
+    my @filecontents = <FILE>;
+
+    close( FILE );
+    # first get start of stage line numbers
+
+    my @stageLineNumbers = ();
+
+    for( my $k = 200; $k < @filecontents; $k++ )
+      {
+      if( $filecontents[$k] =~ m/^Stage/ )
+        {
+        push( @stageLineNumbers, $k );
+        }
+      }
+
+    print "@stageLineNumbers\n";
+    my @iterations = ();
+
+    for( my $k = $stageLineNumbers[1]; $k < @filecontents; $k++ )
+      {
+      if( $filecontents[$k] =~ m/Elapsed time \(stage/ )
+        {
+        my @tmp = split( ': ', $filecontents[$k] );
+        chomp( $tmp[1] );
+        push( @elapsedTimes, scalar( $tmp[1] ) );
+        }
+      if( $filecontents[$k] =~ m/iterations/ && $filecontents[$k] =~ m/x/ )
+        {
+        my @tmp = split( ' = ', $filecontents[$k] );
+        print "$filecontents[$k]\n";
+        chomp( $tmp[1] );
[…]

### 2013-02-08 WBIR2012/WBIRStudy/plotResults.R

@@ -1,0 +1,256 @@
+syn_files <- list();
+
+# syn_files[[1]] <- list.files( "./output_syn_300_0/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[2]] <- list.files( "./output_syn_300_0_downsample/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[3]] <- list.files( "./output_syn_500_0/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[4]] <- list.files( "./output_syn_500_0_downsample/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[5]] <- list.files( "./output_syn_700_0/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[6]] <- list.files( "./output_syn_700_0_downsample/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[7]] <- list.files( "./output_bsplinesyn_10x12x10_order2/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[8]] <- list.files( "./output_bsplinesyn_10x12x10_order2_downsample/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[9]] <- list.files( "./output_bsplinesyn_10x12x10_order3/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[10]] <- list.files( "./output_bsplinesyn_10x12x10_order3_downsample/", full.names = TRUE, pattern = "*.csv" );
+#
+# syn_files[[11]] <- list.files( "./output_gauss_300_0/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[12]] <- list.files( "./output_dmffd_10x12x10_order2/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[13]] <- list.files( "./output_dmffd_10x12x10_order3/", full.names = TRUE, pattern = "*.csv" );
+# syn_files[[14]] <- list.files( "./output_original_SyN/", full.names = TRUE, pattern = "*.csv" );
+#
+#
+# legendEntries <- c( "SyN[3]", "SyNd[3]", "SyN[5]", "SyNd[5]", "SyN[7]", "SyNd[7]",
+#                     "BSyN[10x12x10,2]", "BSyNd[10x12x10,2]", "BSyN[10x12x10,3]", "BSyNd[10x12x10,3]",
[…]

### 2013-02-08 WBIR2012/WBIRStudy/warpLabels.pl

@@ -1,0 +1,72 @@
+#! /usr/bin/perl/ -w
+
+my $baseDir = '/Users/ntustison/Data/Public/LONI/LPBA40/WBIRStudy/';
+my $dataDir = $baseDir . $ARGV[0];
+
+
+my @labels = ();
+my @averageTargetOverlap = ();
+my $count = 0;
+
+for( my $i = 1; $i <= 10; $i++ )
+  {
+  for( my $j = 1; $j <= 10; $j++ )
+    {
+    if( $i == $j )
+      {
+      next;
+      }
+
+    my $i_idx = sprintf( "%02d", $i );
+    my $j_idx = sprintf( "%02d", $j );
+    print "Doing S${j_idx}toS${i_idx}\n";
+
+#    my @args = ( 'antsApplyTransforms',
+#                 '-d', 3,
+#                 '-i', "${baseDir}/native_labels_renamed/S${j_idx}.nii.gz",
+#                 '-r', "${baseDir}/native_labels_renamed/S${i_idx}.nii.gz",
+#                 '-o', "${baseDir}/WBIRStudy/S${j_idx}toS${i_idx}_WarpedLabels.nii.gz",
+#                 '-n', 'NearestNeighbor',
+#                 '-t', "${baseDir}/WBIRStudy/S${j_idx}toS${i_idx}_1Warp.nii.gz",
+#                 '-t', "${baseDir}/WBIRStudy/S${j_idx}toS${i_idx}_0Affine.mat",
+#                 '-v', 0
+#               );
+#    system( @args ) == 0 || die "Fail.\n";
+
+    my @out = `LabelOverlapMeasures 3 ${dataDir}S${j_idx}toS${i_idx}_WarpedLabels.nii.gz ${baseDir}/../native_labels_renamed/S${i_idx}.nii.gz`;
+
+    for( my $n = 5; $n < @out; $n++ )
+      {
+      my @labelStats = split( ' ', $out[$n] );;
+      if( $i == 1 && $j == 2 )
+        {
+        $averageTargetOverlap[$n-5] = $labelStats[1];
+        $labels[$n-5] = $labelStats[0];
+        }
+      else
+        {
+        $averageTargetOverlap[$n-5] += $labelStats[1];
+        if( $labels[$n-5] != $labelStats[0] )
+          {
+          print "Mismatched labels.\n";
+          exit;
+          }
+        }
+      }
[…]

### 2013-02-08 SOMResearchRetreat2013/Presentation/presentation.tex
ENH:  UVA school of medicine research retreat presentation.

--- 2013-02-08 SPIE2013/KK/Kirby_results/presentation.tex
@@
       \vspace{-0.5cm}
-      \pause
       \includegraphics[width=5cm]{hippocampus.png} \\
@@
 
-\begin{frame}
-
-\tikzset{level 1 concept/.append style={font=\sf, sibling angle=135,level distance = 25mm}}
-\tikzset{level 2 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
-\tikzset{level 3 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
-\tikzset{every node/.append style={scale=0.5}}
-
-\begin{center}
-\resizebox{9cm}{7.5cm}{
-\begin{tikzpicture}
-  \path[mindmap,concept color=red!50!black,text=white]
-    node[concept,scale=0.9] {\Large Data Analysis}
-    [clockwise from=45]
-    child[concept color=green!50!black, yshift=-0.5cm] {
-      node[concept,scale=1.1] {Imaging Applications}
-      [clockwise from=120]
-      child[concept color=green!80!black] { node[concept] (ns) {Neuro\-science} }
-      child { node[concept] {Func. Lung Imaging} }
-      child { node[concept] {Breast} }
-      child { node[concept] {Cardiac} }
-    }  
-    child[concept color=blue!80!black, yshift=0.25cm] {
-      node[concept,scale=1.1] (stat) {Statistics}
-      [clockwise from=-60]
-      child { node[concept] {Linear Modeling} }
-      child { node[concept] (ml) {Machine Learning} }
-    }
-    child[concept color=orange, yshift=-0.5cm] { 
-      node[concept,scale=1.1] {Computer Science} 
-        [clockwise from=-145]
-        child { node[concept] (ad) {Algorithm Development}
-           [clockwise from=-60]
-           child { node[concept] (antsr) {\footnotesize ANTsR} }
-           child { node[concept] {ANTs} }
-           child { node[concept] {ITK} }
-           }
-        child { node[concept] (cp) {Computer Programming} }
-        child { node[concept] (cc) {Cluster Computing} }
-    };
-
-		\newcommand{\connbluetoorange}{to[circle connection bar switch color=from (blue!80!black) to (orange)]}
-		\newcommand{\connorangetoorange}{to[circle connection bar switch color=from (orange) to (orange)]}
-		\newcommand{\conngreentoorange}{to[circle connection bar switch color=from (green!80!black) to (orange)]}
-
-  \begin{pgfonlayer}{background}
-      \path (ad) \connorangetoorange (cp);
-      \path (stat) \connbluetoorange (antsr);
-      \path (ns) \conngreentoorange (cc);
-  \end{pgfonlayer}
-
+%\begin{frame}
+%
+%
+%\tikzset{level 1 concept/.append style={font=\sf, sibling angle=135,level distance = 25mm}}
+%\tikzset{level 2 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
+%\tikzset{level 3 concept/.append style={font=\sf, sibling angle=60,level distance = 15mm}}
+%\tikzset{every node/.append style={scale=0.5}}
+%
+%\begin{center}
+%\resizebox{9cm}{7.5cm}{
+%\begin{tikzpicture}
+%  \path[mindmap,concept color=red!50!black,text=white]
+%    node[concept,scale=0.9] {\Large Data Analysis}
+%    [clockwise from=45]
+%    child[concept color=green!50!black, yshift=-0.5cm] {
+%      node[concept,scale=1.1] {Imaging Applications}
+%      [clockwise from=120]
+%      child[concept color=green!80!black] { node[concept] (ns) {Neuro\-science} }
+%      child { node[concept] {Func. Lung Imaging} }
+%      child { node[concept] {Breast} }
+%      child { node[concept] {Cardiac} }
+%    }  
+%    child[concept color=blue!80!black, yshift=0.25cm] {
+%      node[concept,scale=1.1] (stat) {Statistics}
+%      [clockwise from=-60]
+%      child { node[concept] {Linear Modeling} }
+%      child { node[concept] (ml) {Machine Learning} }
+%    }
+%    child[concept color=orange, yshift=-0.5cm] { 
+%      node[concept,scale=1.1] {Computer Science} 
+%        [clockwise from=-145]
+%        child { node[concept] (ad) {Algorithm Development}
+%           [clockwise from=-60]
+%           child { node[concept] (antsr) {\footnotesize ANTsR} }
+%           child { node[concept] {ANTs} }
+%           child { node[concept] {ITK} }
+%           }
+%        child { node[concept] (cp) {Computer Programming} }
+%        child { node[concept] (cc) {Cluster Computing} }
+%    };
+%    
+%		\newcommand{\connbluetoorange}{to[circle connection bar switch color=from (blue!80!black) to (orange)]}
+%		\newcommand{\connorangetoorange}{to[circle connection bar switch color=from (orange) to (orange)]}
+%		\newcommand{\conngreentoorange}{to[circle connection bar switch color=from (green!80!black) to (orange)]}
+%
 %  \begin{pgfonlayer}{background}
-%      \path (ml) \connbluetoorange (cs);
+%      \path (ad) \connorangetoorange (cp);
+%      \path (stat) \connbluetoorange (antsr);
+%      \path (ns) \conngreentoorange (cc);
 %  \end{pgfonlayer}
-
-\end{tikzpicture}
-}
-\end{center}
-\end{frame}
+%    
+%%  \begin{pgfonlayer}{background}
+%%      \path (ml) \connbluetoorange (cs);
+%%  \end{pgfonlayer}
+%      
+%\end{tikzpicture}
+%}
+%\end{center}
+%\end{frame}
 
@@
   \begin{tabular}{c}
-  \includegraphics[height=2.75cm]{ASL.png} 
+  \includegraphics[height=2.5cm]{ASL.png} 
   \vspace{0.15cm}
@@
     \path[->,thick, color = yellow] (rp.west) edge [out= 180, in= 90] (1.25,0);			  
-}			  
-\uncover<3->{    
     \path[->,rounded corners=0.1cm] 
@@
 }			  
-\uncover<4->{    
+\uncover<3->{    
     \path[->,rounded corners=0.1cm] node[format,below of=nirep, text width = 4.0cm, yshift=1cm ] (wow) { 
@@
 			  };
-}			  
-\uncover<3->{
     \path[->,rounded corners=0.1cm] node[text width=3cm, yshift=2cm, below of=logos] (itk) { 
@@
 }			  
-\uncover<4->{
-    \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift=1.2cm, below of=itk] (ants) { 
-        \begin{center}
-			  \includegraphics[width=2.1cm]{ANTsLogo.png} \\
-        \end{center}
-			  };
-}			  
-\uncover<5->{
+\uncover<3->{
     \path[->,rounded corners=0.1cm] node[text width=2.1cm, yshift=1.2cm, below of=itk] (ants) { 
@@
 			  };		
-}
-
-\uncover<3->{			  
   \path[->,thick, color = yellow, dashed] (0.5,2.3) edge [out=35,in= 145] (4.5,2.3);
@@
   \path[->,thick, color = yellow, dashed] (0.5,-2.8) edge [out=35,in= 145] (4.5,-2.8);
-}  
-\uncover<4->{			  
   \path[->,thick, color = red, dashed] (3.1,-3.45) edge [out=-35,in=-145] (4.5,-3.45);
@@
 			  };
-    \path[->,rounded corners=0.1cm] node[text width=6cm, xshift=5cm, right of=ants] { 
+    \path[->,rounded corners=0.1cm] node[text width=7cm, xshift=5.5cm, right of=ants] { 
         \begin{center}
+        \begin{block}{Further information}
         \begin{itemize}
-          \item Insight Toolkit:  http://www.itk.org
-          \item ANTs/ANTsR:  http://www.picsl.upenn.edu/ANTs/
-          \item UVACSE:  http://www.uvacse.virginia.edu
+          \item \url{http://www.itk.org}
+          \item \url{http://www.picsl.upenn.edu/ANTs}
+          \item \url{http://www.uvacse.virginia.edu}
         \end{itemize}
+        \end{block}
         \end{center}

### 2013-03-25 ANTsRS/paper.tex
ENH: Ozzy

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%

%\documentclass[preprint,authoryear,review,12pt]{elsarticle}
\documentclass[final,5p,times,twocolumn]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,two column; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{color}
\usepackage{multirow,booktabs,ctable,array}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{ulem}
\usepackage{setspace}
\usepackage{listings}
\usepackage{float}
\usepackage{listings}
\usepackage{color,colortbl}
\usepackage{rccol}
\usepackage[table]{xcolor}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.955}
    \definecolor{listwhite}{gray}{1.0}
    \definecolor{lightcyan}{rgb}{0.88,1,1}

\newcommand{\lstsetcpplong}
{
\lstset{frame = tb,
        framerule = 0.25pt,
        float,
        fontadjust,
        backgroundcolor={\color{listlightgray}},
        basicstyle = {\ttfamily\scriptsize},
        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
        identifierstyle = {\ttfamily},
        commentstyle = {\ttfamily\color{listcomment}\textit},
        stringstyle = {\ttfamily},
        showstringspaces = false,
        showtabs = false,
        numbers = none,
        numbersep = 6pt,
        numberstyle={\ttfamily\color{listnumbers}},
        tabsize = 2,
        language=,
        floatplacement=!h,
        caption={\small \baselineskip 12pt DiReCT long command line menu which is invoked using the `{\ttfamily {-}{-}help}' option.  The short command line menu is obtained by typing `{\ttfamily {-}h}'},
        captionpos=b,
        label=listing:long
        }
}

\floatstyle{plain}
\newfloat{command}{thp}{lop}
\floatname{command}{Command}

%\usepackage[nomarkers,notablist]{endfloat}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
% \usepackage{amsthm}

 \usepackage{makecell}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\providecommand{\OO}[1]{\operatorname{O}\bigl(#1\bigr)}

\graphicspath{
             {./Figures/}
             }

\long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

\journal{NeuroImage}

\begin{document}

\begin{frontmatter}

\title{Multivariate Neuroanalysis with ANTsR:  Application to Supervised Brain Segmentation}

\author[label1]{Nicholas J.~Tustison
  \fnref{label0}}
  \fntext[label0]{\scriptsize Corresponding author:  PO Box 801339, Charlottesville, VA 22908; T:  434-924-7730; email address:  ntustison@virginia.edu.
  }
\author[label2]{K.~L.~Shrinidhi}
\author[label2]{Jeffrey T.~Duda}
\author[label2]{Philip A.~Cook}
\author[label1]{Christopher Durst}
\author[label1]{James C.~Gee}
\author[label1]{Murray C.~Grossman}
\author[label1]{Max Wintermark}
\author[label2]{Brian B.~Avants}
\address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
\address[label2]{Penn Image Computing and Science Laboratory, University of Pennsylvania,
                Philadelphia, PA}

%\maketitle

\linenumbers

\begin{abstract} 
Improvements in image acquisition and increasingly sophisticated statistical techniques 
for neuroanalysis have spurred the development of advanced computational frameworks for studying the
brain---many of which
have been made publicly available.  One such popular toolkit is the open source Advanced
Normalization Tools (ANTs) which contains a suite of well-vetted processing tools for
core data transformation tasks such as image registration, image segmentation, 
inhomogeneity field correction and template building.  However, despite the numerous
solutions afforded by such tools, neuroscience (and data analysis in general) requires
a statistical platform for making inferences with respect to given hypotheses once the 
requisite data transformations have been performed.  In this paper, we describe the
coupling of ANTs with the widely-used R language---an open source environment for 
statistical processing and data visualization---which we denote as ANTsR.  One of the
many benefits from such an integration is the accessibility to advanced statistical 
and machine learning techniques.  To showcase the flexibility and power of ANTsR, we 
apply the combination of one such technique, Random Forests, and ANTs processing 
tools to the difficult problem of brain tumor segmentation.  This includes evaluation on 
the public data set from the MICCAI 2012 BRATS challenge consisting of both real and simulated
data.  To facilitate reproducibility, all scripts and data used in this paper are 
publicly available for download.  
\end{abstract}

\begin{keyword}
advanced normalization tools \sep BRATS \sep brain tumor segmentation \sep R project \sep random forests
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

\end{frontmatter}
%
%
\newpage

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \citep{key}          ==>>  [#]
%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
%%   \citet{key}         ==>>  Author [#]

%% main text

%\input{intro}
%
%\input{methods}

\section{Introduction}

O

\section{Materials and Methods}

\subsection{ANTsR:  An ANTs/R Interface}

\subsubsection{Installation}

The ANTsR package is publicly available on the github project hosting service.%
\footnote{
https://github.com/stnava/ANTsR
}
Prior to installation of ANTsR, several external R packages
need to be installed including: \verb#Rcpp#, \verb#signal#, \verb#timeSeries#, 
\verb#mFilter#, \verb#doParallel#, \verb#robust#, \verb#magic#, \verb#knitr#, \verb#pixmap#, 
\verb#rgl#, \verb#misc3d# which is facilitated by the 
\verb#install.packages()# mechanism.  Additionally, in order
to perform the supervised brain segmentation as described 
in later sections, one would need to also install 
\verb#randomForest#, \verb#snowfall#, \verb#rlecuyer#,
and \verb#ggplot2#.

CMake%
\footnote{
http://www.cmake.org/
}
is an open source tool for the management and building of 
large-scale software projects.  It is used
to coordinate the downloading of external packages,
such as the Insight Toolkit (ITK)%
\footnote{
http://www.itk.org/
}
and ANTs.  Detailed instructions for download and
installation can be found on the ANTsR github website.

\subsection{Supervised Brain Segmentation}

\subsubsection{Random Forests}

Building on the developments of Ho \cite{ho1995} and Amit and Geman \cite{amit1997},
Breiman introduced Random Forests for classification/regression \cite{breiman2001} which performs relatively well against other machine learning strategies such as support vector machines and neural networks.

\subsubsection{Multi-Modality Feature Image Preprocessing}

Although several studies have pointed out the importance of
intensity normalization and bias correction, our experience 
with the training data illustrated a degradation in performance
when one or both steps (using \cite{nyul2000} and \cite{tustison2010},
respectively) were performed due to the presence of the tumor/edema complex.  
Instead we simply windowed the image intensity
for all images to be between the quantiles $[0.01,0.99]$ and
subsequently rescaled to $[0,1]$.  This strategy seemed to provide 
the best classification (probably due to the fact that the intensity
corruption due to bias was relatively minimal in the data used).
However, improved results might be obtained
by a more sophisticated strategy in which an initial supervised segmentation 
is used to find the probabilistic estimates of the normal tissue
locations (as done in this work) which could then be followed by using
only the probabilistic maps of the normal tissues to construct a
weight, or confidence, mask for estimating the inhomogeneity bias
using N4 \cite{tustison2010} followed by a repeated application of 
the supervised segmentation strategy described.  Note that a similar
iterative-based strategy until convergence is implemented in the ANTs script \verb#antsAtroposN4.sh#.

\subsubsection{Multi-Modality Feature Image Generation}

Key to any supervised regression or classification protocol are the 
selected features for training and subsequent testing.  Based on previous
work and our own experience, we selected the following feature images
to showcase the supervised segmentation strategy developed in this work.

\begin{itemize}
  \item Per modality (FLAIR, T1, T1C, T2)
    \begin{itemize}
      \item First-order neighborhood statistical images:
            mean, variance, skewness, and entropy. 
            Neighborhood radius = 3.
    \item GMM-based posteriors: CSF, gray matter, white matter, edema, and tumor
    \item GMM connected component geometry features:  volume, volume to 
          surface area ratio, eccentricity, elongation
    \item Template-based:  symmetric template difference and contralateral difference.
          Gaussian smoothing ($\sigma = 4$mm).
    \end{itemize}
  \item Misc.
    \begin{itemize}
    \item Normalized Euclidean distance
    \item Log Jacobian  
    \end{itemize}
\end{itemize}

For each modality, we create four first-order statistical feature images,
five Gaussian mixture model (GMM)-based posterior probability feature images,
four geometry features generated from the GMM posterior probability images
based on connected components, and two difference images using symmetric template
construction for a total of 4 modalities $\times$ (4 + 5 + 4 + 2) feature images per modality $=$ 60 images.  We employ two additional images consisting of the 
Euclidean distance image \cite{maurer2003} created from the skull-stripped 
binary mask rescaled to the range $[0,1]$ and the log
Jacobian image derived from the spatial normalization of the symmetric multivariate template and individual subject images.  Given the intensity corrected images
the corresponding multivariate template images, and a brain mask for each subject
creation of all feature images is performed using the script \verb#createFeatureImages.sh#.

Prior  cluster centers for specific tissue types learned from training data \cite{reynolds2009} are used in the GMM to create multiple feature images.  
Given $M$ tissue types (e.g. CSF, gray matter,
white matter, edema, and tumor), a GMM formulates the 
probability distribution at each voxel, $\mathbf{x}$, as the
sum of Gaussian components, $\mathcal{N}(\mathbf{x}|\mu,\sigma)$, i.e.
\begin{align}
p\left(\mathbf{x}|\mu_m,\sigma_m,\lambda_m\right) = \sum_{i=1}^M \lambda_m \mathcal{N}(\mathbf{x}|\mu_m,\sigma_m)
\end{align}
where $\sum_{m=1}^M \lambda_m = 1$.  One popular method for 
determining the parameters of the GMM is maximum likelihood 
estimation which can performed using the Atropos segmentation 
tool \cite{avants2011}.  In contrast to previous generative
modeling approaches for multi-modal tumor segmentation 
(e.g. \cite{prastawa2003,zikic2012}), we do not use multivariate 
Gaussians to specify tissue probabilities but rather incorporate each
univariate probability map into the feature vector of the training
data.  As pointed out in \cite{menze2010}, multivariate modeling
might obscure the distinct biological information provided by each 
modality.  Instead, we let the random forest construction 
process determine the optimal combination of such multivariate
information.
Additionally, maximum posterior labeling from the GMM processing
is used to determine the connected components for each label.  
Geometric features (assigned voxel-wise) include the physical volumes 
of each connected component including the volume to surface area ratio, 
the elongation, and eccentricity.

In order to better characterize deviations from normal
multi-modal brain shape and appearance, several features were derived 
using population-specific multivariate template 
construction. A recent neuroimaging reproducibility study
by Landman et al. resulted in an open data cohort of 21
normal individuals, each imaged twice, comprising several
modalities including arterial spin labeling, 
fluid attenuated inversion recovery (FLAIR),
diffusion tensor imaging, functional imaging, T1, and T2 
\cite{landman2011}.

Given $K$ image modality types for $N$ subjects,  
${\mathbf I} = \{I_1,I_2,\ldots, I_K\}$, multivariate 
template construction iterates between optimizing the set 
of diffeomorphic transforms between the subjects and the 
template, 
$\left\{\left(\phi_1,\phi_1^{-1}\right),\ldots,\left(\phi_N,\phi_N^{-1}\right)\right\}$ 
and constructing the 
optimal multivariate template appearance 
$\mathbf{J}=\{J_1,J_2,\ldots, J_K\}$ to minimize the
following cost function:
\begin{align}
  \sum_{n=1}^N 
        \Bigg[ D &\left( \psi(\mathbf{x}),\phi_1^n(\mathbf{x},1)\right) \\ \nonumber 
        +& \sum_{k=1}^K \lambda_k \Pi_k \left(I_k^n\left(\phi_n(\mathbf{x},0.5)\right),J_k\left(\phi^{-1}_n(\mathbf{x},0.5)\right)\right)\Bigg]
\end{align}
where $D$ is the diffeomorphic shape distance,
\begin{align}
D\left( \phi( \mathbf{x},0),\phi( \mathbf{x},1)\right) = \int_0^1 \| \nu(\mathbf{x},t)\|_L dt
\end{align}
dependent on the choice of linear operator, $L$, and $\nu$
is the velocity field
\begin{align}
\nu\left( \phi(\mathbf{x},t) \right) = \frac{d\phi(\mathbf{x},t)}{dt},\,\,\, \phi(\mathbf{x},0) = \mathbf{x}.
\end{align}
Each pairwise registration employing the similarity metric $\Pi_k$ can 
be assigned a relative weighting, $\lambda_k$, to weight a particular
modality's influence in the construction process.  Further theoretical
details can be found in \cite{avants2008,avants2010}.
In terms of implementation, this algorithm is 
encapsulated in the script \verb#antsMultivariateTemplateConstruction.sh#,
available in the ANTs repository, which permits parallel processing on
an individual workstation or on a large computational cluster.

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=40mm]{Figures/S_templateFA_140.png} &
    \includegraphics[width=40mm]{Figures/S_templateFLAIR_140.png} \\
    (a) & (b) \\
    \includegraphics[width=40mm]{Figures/S_templateT1_140.png} &
    \includegraphics[width=40mm]{Figures/S_templateT2_140.png} \\
    (c) & (d) 
  \end{tabular}
  \caption{Multivariate symmetric template created from the Kirby 
           21 data described in \cite{landman2011}.  Shown are the
           (a) fractional anisotropy (FA), (b) FLAIR, (c) MPRAGE, 
           and (d) T2 template components.
          }
  \label{fig:symmetrictemplates}
\end{figure}

\subsubsection{\underline{Bra}in \underline{T}umor \underline{S}egmentation Challenge Data}

The Brain tumor segmentation
Associated with the 2012 International Conference on Medical Image Computing and Computer Assisted
Intervention (MICCAI),%
\footnote{
http://www.miccai2012.org
}
the

\subsubsection{MS lesion segmentation challenge Challenge Data}

\section{Results}

\begin{table*}
\caption{Dice scores from the MICCAI 2012 BRATs Study}
\begin{center}
\begin{tabular*}{0.975\textwidth}{@{\extracolsep{\fill} } c c c c c c c c c}
\toprule
{} & \multicolumn{2}{c}{High-grade (real)} & \multicolumn{2}{c}{Low-grade (real)} & \multicolumn{2}{c}{High-grade (simulated)} & \multicolumn{2}{c}{Low-grade (simulated)}\\
{\bf Method} & Edema & Tumor & Edema & Tumor & Edema & Tumor & Edema & Tumor\\
\midrule
\cite{zikic2012} & {$0.70 \pm 0.09$} & {$0.71 \pm 0.24$} & {$0.44 \pm 0.18$} & {$0.62 \pm 0.27$} & {$0.65 \pm 0.27$} & {$0.90 \pm 0.05$} & {$0.55 \pm 0.23$} & {$0.71 \pm 0.20$} \\
\cite{bauer2012} & {$0.61 \pm 0.15$} & {$0.62 \pm 0.27$} & {$0.35 \pm 0.18$} & {$0.49 \pm 0.26$} & {$0.68 \pm 0.26$} & {$0.90 \pm 0.06$} & {$0.57 \pm 0.24$} & {$0.74 \pm 0.10$} \\
\rowcolor{yellow}
ANTsR & {$0.64 \pm 0.18$} & {$0.65 \pm 0.30$} & {$0.48 \pm 0.16$} & {$0.69 \pm 0.21$} & {$0.75 \pm 0.25$} & {$0.94 \pm 0.04$} & {$0.65 \pm 0.25$} & {$0.83 \pm 0.09$}  \\
\bottomrule
\end{tabular*}
\end{center}
\end{table*}

\section{Discussion and Conclusions}

%% References with bibTeX database:

\section*{Acknowledgments}

\section*{References}

\bibliographystyle{elsarticle-harv}
\bibliography{references}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model1-num-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-1-num.tex'.

### 2013-03-25 KapowskiChronicles/methods.tex
ENH: Ozzy

@@
 
+In order to better integrate Atropos and N4, we use  
+a pure tissue probability weight mask generated from the 
+posterior probabilities generated by the segmentation 
+process.  Given $N$ labels and the corresponding $N$
+posterior probability maps $\{ P_1, \ldots, P_N\}$ produced
+during the segmentation process, the N4 weight mask is 
+created at each N4 $\Leftrightarrow$ Atropos iteration from:
+\begin{align}
+  P_{pure\,\,tissue}(\mathbf{x}) = \sum_{i=1}^N P_i(\mathbf{x}) \prod_{j=1, j \neq i}^N \left( 1 - P_j(\mathbf{x}) \right).
+\end{align}
+
 \subsubsection{Brain extraction}

### 2013-03-25 SPIE2013/KK/IXI_NKI_Oasis_combinedResults/agePredict.R

@@ -1,113 +1,0 @@
-demographics <- read.csv( file = "demographics.csv", header = TRUE );
-# volumes <- read.csv( file = "volume.csv", header = TRUE );
-thickness <- read.csv( file = "thickness.csv", header = TRUE );
-
-indices <- which( demographics$AGE >= 0 & demographics$AGE <= 100 );
-thickness <- thickness[indices,];
-demographics <- demographics[indices,];
-
-cortical_labels <- c( "L occipital", "R occipital",
-                     "L cingulate", "R cingulate",
-                     "L insula",    "R insula",
-                     "L temporal pole",   "R temporal pole",
-                     "L superior temporal", "R superior temporal",
-                     "L infero temporal", "R infero temporal",
-                     "L parahippocampal", "R parahippocampal",
-                     "L frontal pole",    "R frontal pole",
-                     "L superior frontal","R superior frontal",
-                     "L middle frontal",  "R middle frontal",
-                     "L inferior",        "R inferior",
-                     "L orbital frontal", "R orbital frontal",
-                     "L precentral",      "R precentral",
-                     "L superior parietal", "R superior parietal",
-                     "L inferior parietal", "R inferior parietal",
-                     "L postcentral",       "R postcentral" );
-
-nSamples <- 100;
-male_color = "darkred";
-female_color = "navyblue";
-
-age <- demographics$AGE;
-id <- demographics$ID;
-gender <- cut( demographics$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
-
-ixi <- indices[which( indices <= 360 )];
-nki <- indices[which( indices >= 361 & indices <= 537 )];
-oasis <- indices[which( indices >= 537 )];
-
-ixiAge <- age[ixi];
-nkiAge <- age[nki];
-oasisAge <- age[oasis];
-
-ixiGender <- gender[ixi];
-nkiGender <- gender[nki];
-oasisGender <- gender[oasis];
-
-
-
-sampled.th <- seq( min( thickness ), max( thickness ) , length = nSamples );
-sampled.age <- seq( 0, 100, length = nSamples );
-
-for( i in 1:32 )
-  {
[…]

### 2013-03-25 SPIE2013/KK/IXI_results/plotResults.R

@@ -1,374 +1,0 @@
-# resampler <- function( sp )
-#   {
-#   n <- nrow( sp.frame )
-#   resample.rows <- sample( 1:n,size = n, replace = TRUE )
-#   return( sp.frame[resample.rows,] )
-#   }
-#
-# sp.spline.estimator <- function( x, y, m=300 )
-#   {
-#   # Fit spline to data, with cross-validation to pick lambda
-#   fit <- smooth.spline( x, y, cv = TRUE );
-#   # Set up a grid of m evenly-spaced points on which to evaluate the spline
-#   eval.grid <- seq( from = min( x ), to = max( x ), length.out = m )
-#     # Slightly inefficient to re-define the same grid every time we call this,
-#     # but not a big overhead
-#   # Do the prediction and return the predicted values
-#   return( predict( fit, x = eval.grid)$y )  # We only want the predicted values
-#   }
-#
-# sp.spline.cis <- function( B, alpha, m = 300 ) {
-#   spline.main <- sp.spline.estimator( sp.frame, m = m )
-#   # Draw B boottrap samples, fit the spline to each
-#   spline.boots <- replicate( B, sp.spline.estimator( sp.resampler(), m = m ) )
-#     # Result has m rows and B columns
-#   cis.lower <- 2 * spline.main - apply( spline.boots, 1, quantile, probs = 1 - 0.5 * alpha )
-#   cis.upper <- 2 * spline.main - apply( spline.boots, 1, quantile, probs = 0.5 * alpha )
-#   return( list( main.curve = spline.main, lower.ci = cis.lower, upper.ci = cis.upper,
-#               x = seq( from = min( sp.today ), to = max( sp.today ), length.out = m ) ) )
-# }
-
-calculate.spline <- function( spline.data, m = 300 )
-  {
-  fit <- smooth.spline( spline.data$x, spline.data$y, df = 3 );
-  res <- residuals( fit );
-  eval.grid <- seq( min( spline.data$x ), to = max( spline.data$x ), length.out = m );
-  return( list( x = eval.grid, y = predict( fit, x = eval.grid )$y, res = res ) );
-  }
-
-resampler <- function( spline.data )
-  {
-  n <- length( spline.data$x );
-  resample <- sample( 1:n, size = n, replace = TRUE );
-
-  resampled.data <- list();
-  resampled.data$x = spline.data$x[resample];
[…]

### 2013-03-25 SPIE2013/KK/IXI_results/predict_age.R

@@ -1,48 +1,0 @@
-library(MASS)
-library(bestglm)
-md<-read.csv('meandiffusion.csv')
-vol<-read.csv('volume.csv')
-th<-read.csv('thickness.csv')
-demog<-read.csv('demographics.csv')
-# remove some bad data
-whold<-( demog$AGE > 50 & demog$AGE > 1  )
-whmid<-( demog$AGE > 40 & demog$AGE < 60  )
-whyoung<-( demog$AGE < 41 & demog$AGE > 1  )
-wh<-whold
-print(sum(wh)/nrow(demog))
-demog<-demog[wh,]
-vol<-vol[wh,]
-th<-th[wh,]
-md<-md[wh,]
-# training / testing data
-subs1<-c(1:(nrow(demog)/2))
-subs1<-subs1*2
-subs2<-subs1-1
-subs2<-subs2[1:(length(subs2)-1)] ; length(subs2) ; length(subs1)
-# make data frame with all training data
-Xy1<-data.frame(vol=vol[subs1,],th=th[subs1,],sex=demog$SEX[subs1],age=demog$AGE[subs1])
-Xy2<-data.frame(vol=vol[subs2,],th=th[subs2,],sex=demog$SEX[subs2],age=demog$AGE[subs2])
-Xy1<-data.frame(th=th[subs1,],sex=demog$SEX[subs1],age=demog$AGE[subs1])
-names(Xy1)<-c(paste("X",1:(ncol(Xy1)-1),sep=""),"y")
-Xy2<-data.frame(th=th[subs2,],sex=demog$SEX[subs2],age=demog$AGE[subs2])
-names(Xy2)<-c(paste("X",1:(ncol(Xy2)-1),sep=""),"y")
-Xy<-Xy1
-ans<-bestglm(Xy, IC="CV" , t=125 )
-# ans<-bestglm(Xy, IC="CV", t=10,CVArgs=list(Method="d", K=2, REP=10))
-# now apply to test data
-Xy<-Xy2
-predictedage<-predict(ans$BestModel,newdata=Xy)
-realage<-Xy$y
-print(paste('Mean:',mean(abs(predictedage-realage))))
-ranger<-c( min( c(realage,predictedage) ),  max( c(realage,predictedage) ) )
-pdf( file = "brian_results.pdf", 8, 8 );
-plot( ranger, ranger, type = "n",
-    xlab = expression( paste( bold( "Real Age (years)" ) ) ),
-    ylab = expression( paste( bold( "Predicted Age (years)" ) ) ),
[…]

### 2013-03-25 SPIE2013/KK/IXI_results/predict_age2.R

@@ -1,47 +1,0 @@
-library( "earth" )
-
-
-get.used.pred.names <- function( obj ) # obj is an earth object
-{
-  any1 <- function(x) any(x != 0)    # like any but no warning if x is double
-  names(which(apply(obj$dirs[obj$selected.terms, , drop=FALSE], 2, any1)))
-}
-
-rm( list = ls( all = TRUE ) )
-
-meandiffusion <- read.csv( 'meandiffusion.csv', header = TRUE )
-volumes <- read.csv( 'volume.csv', header = TRUE )
-thickness <- read.csv( 'thickness.csv', header = TRUE )
-demographics <- read.csv( 'demographics.csv', header = TRUE )
-
-meandiffusion <- meandiffusion[demographics$SEX == 2,];
-volumes <- volumes[demographics$SEX == 2,];
-thickness <- thickness[demographics$SEX == 2,];
-demographics <- demographics[demographics$SEX == 2,];
-
-
-agePredictionData <- data.frame( md = meandiffusion, th = thickness );
-
-trainingIndices <- createDataPartition( demographics$AGE, p = 1/2, list = FALSE );
-
-agePredictionData.training <- agePredictionData[trainingIndices,];
-agePredictionData.testing <- agePredictionData[-trainingIndices,]
-
-fit <- earth( x = agePredictionData.training, y = demographics$AGE[trainingIndices], degree = 2, pmethod = "exhaustive"  );
-
-age.real <- demographics$AGE[-trainingIndices];
-age.predict <- predict( fit, agePredictionData.testing );
-
-pdf( file = "brian_results_model.pdf", 8, 8 );
-plot( c( 20, 90 ), c( 20, 90 ), type = "n",
-    xlab = expression( paste( bold( "Real Age (years)" ) ) ),
-    ylab = expression( paste( bold( "Predicted Age (years)" ) ) ),
-    frame.plot = FALSE  );
-lines( c( 20, 90 ), c( 20, 90 ), lwd = 3, lty = 3, col = "black" );
-title( main = paste( "Cortical thickness prediction of age"), col.main = "black", font.main = 2 );
-axis( 1, tick = TRUE, lwd = 2, font = 2 );
-axis( 2, tick = TRUE, lwd = 2, font = 2 );
-points( x = age.real, y = age.predict, pch = 16, lwd = 1.5, cex = 1.0, col = "darkred" );
-dev.off();
-
-print( mean( abs( age.real - age.predict ), na.rm = TRUE ) );

### 2013-03-25 SPIE2013/KK/IXI_results/predict_age3.R

@@ -1,92 +1,0 @@
-library( MASS )
-library( bestglm )
-library( caret )
-
-meandiffusion <- read.csv( 'meandiffusion.csv' )
-volumes <- read.csv( 'volume.csv' )
-thickness <- read.csv( 'thickness.csv' )
-demographics <- read.csv( 'demographics.csv' )
-
-# create the total model
-
-trainingIndices <- createDataPartition( demographics$AGE, p = 1/2, list = FALSE );
-
-training.data <- data.frame( th = thickness[trainingIndices,],
-#                              md = meandiffusion[trainingIndices,],
-                             sex = demographics$SEX[trainingIndices],
-                             age = demographics$AGE[trainingIndices] );
-testing.data <- data.frame( th = thickness[-trainingIndices,],
-#                             md = meandiffusion[-trainingIndices,],
-                             sex = demographics$SEX[-trainingIndices],
-                             age = demographics$AGE[-trainingIndices] );
-
-model.total <- bestglm( training.data, IC = "CV" );
-
-# create the three sub-models
-
-subjects <- list();
-subjects[[1]] <- ( demographics$AGE[trainingIndices] >= 20 & demographics$AGE[trainingIndices] <= 40 );
-subjects[[2]] <- ( demographics$AGE[trainingIndices] >= 40 & demographics$AGE[trainingIndices] <= 60 );
-subjects[[3]] <- ( demographics$AGE[trainingIndices] >= 60 );
-
-models.sub <- list();
-for( i in seq( 1, 3, by = 1 ) )
-  {
-  cat( "constructing model ", i, "\n", sep = "" );
-  dem <- demographics[subjects[[i]],];
-  vol <- volumes[subjects[[i]],];
-  th <- thickness[subjects[[i]],];
-  md <- meandiffusion[subjects[[i]],];
-
-  training.data <- data.frame( th = th,
-#                              md = md,
-                               sex = dem$SEX,
-                               age = dem$AGE );
-  models.sub[[i]] <- bestglm( training.data, IC = "CV", t = 125 );
-  }
-
-pdf( file = "brian_results_model.pdf", 8, 8 );
-plot( c( 20, 90 ), c( 20, 90 ), type = "n",
[…]

### 2013-03-25 SPIE2013/KK/IXI_results/predict_age4.R

@@ -1,111 +1,0 @@
-library( MASS )
-library( glmulti )
-library( caret )
-
-rm( list = ls( all = TRUE ) )
-
-meandiffusion <- read.csv( 'meandiffusion.csv' )
-volumes <- read.csv( 'volume.csv' )
-thickness <- read.csv( 'thickness.csv' )
-demographics <- read.csv( 'demographics.csv' )
-
-mylabels <- c( "whole", "L occipital", "R occipital",
-                     "L cingulate", "R cingulate",
-                     "L insula",    "R insula",
-                     "L temporal pole",   "R temporal pole",
-                     "L superior temporal", "R superior temporal",
-                     "L infero temporal", "R infero temporal",
-                     "L parahippocampal", "R parahippocampal",
-                     "L frontal pole",    "R frontal pole",
-                     "L superior frontal","R superior frontal",
-                     "L middle frontal",  "R middle frontal",
-                     "L inferior",        "R inferior",
-                     "L orbital frontal", "R orbital frontal",
-                     "L precentral",      "R precentral",
-                     "L superior parietal", "R superior parietal",
-                     "L inferior parietal", "R inferior parietal",
-                     "L postcentral",       "R postcentral" );
-
-
-wholeThickness <- rowSums( as.matrix( volumes ) * as.matrix( thickness ) ) / rowSums( volumes );
-wholeMeanDiffusion <- rowSums( as.matrix( volumes ) * as.matrix( meandiffusion ) ) / rowSums( volumes );
-
-accuracy <- rep( 0, 33 );
-for( n in 0:100 )
-  {
-  cat( "Run ", n, "\n", sep = "" );
-  trainingIndices <- createDataPartition( demographics$AGE, p = 1/2, list = FALSE )
-
-  training.data <- data.frame( sex = demographics$SEX[trainingIndices],
-                               age = demographics$AGE[trainingIndices],
-                               wth = wholeThickness[trainingIndices],
-                               th = thickness[trainingIndices,],
-                               wmd = wholeMeanDiffusion[trainingIndices],
-                               md = meandiffusion[trainingIndices,]
-                             );
-
[…]

### 2013-03-25 SPIE2013/KK/Kirby_results/agePredict.R

@@ -1,57 +1,51 @@
 library( ggplot2 )
 
-priorWeights <- c( "0.0", "0.1", "0.25", "0.5" )
+results <- read.csv( file = paste( "labelresults", ".csv", sep = '' ), header = TRUE );
 
+cortical_labels <- c( "L occipital", "R occipital",
+                     "L cingulate", "R cingulate",
+                     "L insula",    "R insula",
+                     "L temporal pole",   "R temporal pole",
+                     "L superior temporal", "R superior temporal",
+                     "L infero temporal", "R infero temporal",
+                     "L parahippocampal", "R parahippocampal",
+                     "L frontal pole",    "R frontal pole",
+                     "L superior frontal","R superior frontal",
+                     "L middle frontal",  "R middle frontal",
+                     "L inferior",        "R inferior",
+                     "L orbital frontal", "R orbital frontal",
+                     "L precentral",      "R precentral",
+                     "L superior parietal", "R superior parietal",
+                     "L inferior parietal", "R inferior parietal",
+                     "L postcentral",       "R postcentral" );
 
-for( n in 1:length( priorWeights ) )
+male_color = "darkred";
+female_color = "navyblue";
+
+age <- results$AGE;
+gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
+
+for( i in 1:32 )
   {
-  results <- read.csv( file = paste( "labelresults", priorWeights[n], ".csv", sep = '' ), header = TRUE );
+  th <- results[,i+3];
 
-  cortical_labels <- c( "L occipital", "R occipital",
-                       "L cingulate", "R cingulate",
-                       "L insula",    "R insula",
-                       "L temporal pole",   "R temporal pole",
-                       "L superior temporal", "R superior temporal",
-                       "L infero temporal", "R infero temporal",
-                       "L parahippocampal", "R parahippocampal",
-                       "L frontal pole",    "R frontal pole",
-                       "L superior frontal","R superior frontal",
-                       "L middle frontal",  "R middle frontal",
[…]

### 2013-03-25 SPIE2013/KK/NKI_results/agePredict.R

@@ -1,6 +1,6 @@
-demographics <- read.csv( file = "demographics.csv", header = TRUE );
-# volumes <- read.csv( file = "volume.csv", header = TRUE );
-thickness <- read.csv( file = "thickness.csv", header = TRUE );
+library( ggplot2 )
+
+results <- read.csv( file = paste( "labelresults", ".csv", sep = '' ), header = TRUE );
 
 cortical_labels <- c( "L occipital", "R occipital",
                      "L cingulate", "R cingulate",
@@ -19,67 +19,33 @@ cortical_labels <- c( "L occipital", "R occipital",
                      "L inferior parietal", "R inferior parietal",
                      "L postcentral",       "R postcentral" );
 
-nSamples <- 100;
 male_color = "darkred";
 female_color = "navyblue";
 
-age <- demographics$AGE;
-gender <- cut( demographics$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
-
-sampled.th <- seq( min( thickness ), max( thickness ) , length = nSamples );
-sampled.age <- seq( 20, 90, length = nSamples );
+age <- results$AGE;
+gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
 
 for( i in 1:32 )
   {
-  th <- thickness[,i];
+  th <- results[,i+3];
 
-#   n <- lm( formula = age ~ th + gender );
-#   n <- lm( formula = age ~ md + th + I(th^2) + vo + th:vo + gender );
-#   n <- lm( formula = age ~ md + th + vo + gender );
+  gender[which( gender == 1 )] <- 'male';
+  gender[which( gender == 2 )] <- 'female';
 
-    n <- lm( formula = th ~ age + I(age^2) + gender + age:gender );
+  plotData <- data.frame( cbind( Age = age, Thickness = th, Gender = gender ) )
+  plotData <- transform( plotData, Gender = factor( Gender ) );
 
-#   coeffs <- coef( n );
-#   pvalues <- anova( n )[,"Pr(>F)"];
-
-  male.data <- list();
-  female.data <- list();
-
-  male.data$x <- data.frame( age = sampled.age, th = sampled.th, gender = rep( "male", nSamples ) );
[…]

### 2013-05-05 Tilting/paper.tex
ENH:  Adding tilting paper.

--- 2013-03-25 ANTsRS/paper.tex
@@
 
-\title{Multivariate Neuroanalysis with ANTsR:  Application to Supervised Brain Segmentation}
+\title{Tilting at Algorithmic Windmills}
 
@@
   }
-\author[label2]{K.~L.~Shrinidhi}
-\author[label2]{Jeffrey T.~Duda}
-\author[label2]{Philip A.~Cook}
-\author[label1]{Christopher Durst}
-\author[label1]{James C.~Gee}
-\author[label1]{Murray C.~Grossman}
-\author[label1]{Max Wintermark}
-\author[label2]{Brian B.~Avants}
+\author[label2]{Brian B.~Avants (combined first author, second author or senior author?)}
 \address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
@@
 \begin{abstract} 
-Improvements in image acquisition and increasingly sophisticated statistical techniques 
-for neuroanalysis have spurred the development of advanced computational frameworks for studying the
-brain---many of which
-have been made publicly available.  One such popular toolkit is the open source Advanced
-Normalization Tools (ANTs) which contains a suite of well-vetted processing tools for
-core data transformation tasks such as image registration, image segmentation, 
-inhomogeneity field correction and template building.  However, despite the numerous
-solutions afforded by such tools, neuroscience (and data analysis in general) requires
-a statistical platform for making inferences with respect to given hypotheses once the 
-requisite data transformations have been performed.  In this paper, we describe the
-coupling of ANTs with the widely-used R language---an open source environment for 
-statistical processing and data visualization---which we denote as ANTsR.  One of the
-many benefits from such an integration is the accessibility to advanced statistical 
-and machine learning techniques.  To showcase the flexibility and power of ANTsR, we 
-apply the combination of one such technique, Random Forests, and ANTs processing 
-tools to the difficult problem of brain tumor segmentation.  This includes evaluation on 
-the public data set from the MICCAI 2012 BRATS challenge consisting of both real and simulated
-data.  To facilitate reproducibility, all scripts and data used in this paper are 
-publicly available for download.  
+Exploration of neuroscience hypotheses have been greatly enhanced
+by the increased availability of high-performance computational resources, large-scale
+communal efforts such as the Insight Toolkit and the R statistical project and
+algorithmic advancements for data transformations and processing. 
+An integral component of the vetting process for novel algorithmic techniques
+includes comparison with other methods previously
+established within the community.  Public availability, including 
+open source distribution, of established packages has significantly 
+facilitated these comparative evaluations.  However, complementing the
+recent set of papers pointing to serious methodological and statistical
+bias considerations in neuroimaging research, we point out potential 
+issues associated with a type of measurement bias in comparative algorithmic
+evaluations and propose a set of guidelines for authors and reviewers
+to minimize this confound.
 \end{abstract}
-
 \begin{keyword}
-advanced normalization tools \sep BRATS \sep brain tumor segmentation \sep R project \sep random forests
+open science \sep reproducibility 
 %% keywords here, in the form: keyword \sep keyword
@@
 
-O
+%% References with bibTeX database:
 
-\section{Materials and Methods}
+%\section*{Acknowledgments}
 
-\subsection{ANTsR:  An ANTs/R Interface}
-
-\subsubsection{Installation}
-
-The ANTsR package is publicly available on the github project hosting service.%
-\footnote{
-https://github.com/stnava/ANTsR
-}
-Prior to installation of ANTsR, several external R packages
-need to be installed including: \verb#Rcpp#, \verb#signal#, \verb#timeSeries#, 
-\verb#mFilter#, \verb#doParallel#, \verb#robust#, \verb#magic#, \verb#knitr#, \verb#pixmap#, 
-\verb#rgl#, \verb#misc3d# which is facilitated by the 
-\verb#install.packages()# mechanism.  Additionally, in order
-to perform the supervised brain segmentation as described 
-in later sections, one would need to also install 
-\verb#randomForest#, \verb#snowfall#, \verb#rlecuyer#,
-and \verb#ggplot2#.
-
-CMake%
-\footnote{
-http://www.cmake.org/
-}
-is an open source tool for the management and building of 
-large-scale software projects.  It is used
-to coordinate the downloading of external packages,
-such as the Insight Toolkit (ITK)%
-\footnote{
-http://www.itk.org/
-}
-and ANTs.  Detailed instructions for download and
-installation can be found on the ANTsR github website.
-
-\subsection{Supervised Brain Segmentation}
-
-\subsubsection{Random Forests}
-
-Building on the developments of Ho \cite{ho1995} and Amit and Geman \cite{amit1997},
-Breiman introduced Random Forests for classification/regression \cite{breiman2001} which performs relatively well against other machine learning strategies such as support vector machines and neural networks.
-
-\subsubsection{Multi-Modality Feature Image Preprocessing}
-
-Although several studies have pointed out the importance of
-intensity normalization and bias correction, our experience 
-with the training data illustrated a degradation in performance
-when one or both steps (using \cite{nyul2000} and \cite{tustison2010},
-respectively) were performed due to the presence of the tumor/edema complex.  
-Instead we simply windowed the image intensity
-for all images to be between the quantiles $[0.01,0.99]$ and
-subsequently rescaled to $[0,1]$.  This strategy seemed to provide 
-the best classification (probably due to the fact that the intensity
-corruption due to bias was relatively minimal in the data used).
-However, improved results might be obtained
-by a more sophisticated strategy in which an initial supervised segmentation 
-is used to find the probabilistic estimates of the normal tissue
-locations (as done in this work) which could then be followed by using
-only the probabilistic maps of the normal tissues to construct a
-weight, or confidence, mask for estimating the inhomogeneity bias
-using N4 \cite{tustison2010} followed by a repeated application of 
-the supervised segmentation strategy described.  Note that a similar
-iterative-based strategy until convergence is implemented in the ANTs script \verb#antsAtroposN4.sh#.
-
-\subsubsection{Multi-Modality Feature Image Generation}
-
-Key to any supervised regression or classification protocol are the 
-selected features for training and subsequent testing.  Based on previous
-work and our own experience, we selected the following feature images
-to showcase the supervised segmentation strategy developed in this work.
-
+\section{Notes/Thoughts}
 \begin{itemize}
-  \item Per modality (FLAIR, T1, T1C, T2)
+  \item Several kinds of bias have cropped up in the neuroimaging
+  community recently:
   \begin{itemize}
-      \item First-order neighborhood statistical images:
-            mean, variance, skewness, and entropy. 
-            Neighborhood radius = 3.
-    \item GMM-based posteriors: CSF, gray matter, white matter, edema, and tumor
-    \item GMM connected component geometry features:  volume, volume to 
-          surface area ratio, eccentricity, elongation
-    \item Template-based:  symmetric template difference and contralateral difference.
-          Gaussian smoothing ($\sigma = 4$mm).
-    \end{itemize}
-  \item Misc.
+    \item{methodological bias}:  Kreikesgorte
+    \item{statistical bias}:  Vul2009, Tustison2012
+    \item  What about the dead salmon fmri study---multiple comparisons correction?
+    \item The single subject VBM study (assuming the single subject is representative of the population).
+    \item{overview}:  9 circles of scientific hell ---neuroskeptic %http://blogs.discovermagazine.com/neuroskeptic/?p=1205#.UYZxyJWO7Tx
+    \item Types of measurement bias: \cite{sackett1979}
     \begin{itemize}
-    \item Normalized Euclidean distance
-    \item Log Jacobian  
+      \item Instrument bias (code as a type of instrument which needs to be calibrated correctly).
+      \item Expectation bias (i.e. confirmation bias)
     \end{itemize}  
   \end{itemize}
-
-For each modality, we create four first-order statistical feature images,
-five Gaussian mixture model (GMM)-based posterior probability feature images,
-four geometry features generated from the GMM posterior probability images
-based on connected components, and two difference images using symmetric template
-construction for a total of 4 modalities $\times$ (4 + 5 + 4 + 2) feature images per modality $=$ 60 images.  We employ two additional images consisting of the 
-Euclidean distance image \cite{maurer2003} created from the skull-stripped 
-binary mask rescaled to the range $[0,1]$ and the log
-Jacobian image derived from the spatial normalization of the symmetric multivariate template and individual subject images.  Given the intensity corrected images
-the corresponding multivariate template images, and a brain mask for each subject
-creation of all feature images is performed using the script \verb#createFeatureImages.sh#.
-
-Prior  cluster centers for specific tissue types learned from training data \cite{reynolds2009} are used in the GMM to create multiple feature images.  
-Given $M$ tissue types (e.g. CSF, gray matter,
-white matter, edema, and tumor), a GMM formulates the 
-probability distribution at each voxel, $\mathbf{x}$, as the
-sum of Gaussian components, $\mathcal{N}(\mathbf{x}|\mu,\sigma)$, i.e.
-\begin{align}
-p\left(\mathbf{x}|\mu_m,\sigma_m,\lambda_m\right) = \sum_{i=1}^M \lambda_m \mathcal{N}(\mathbf{x}|\mu_m,\sigma_m)
-\end{align}
-where $\sum_{m=1}^M \lambda_m = 1$.  One popular method for 
-determining the parameters of the GMM is maximum likelihood 
-estimation which can performed using the Atropos segmentation 
-tool \cite{avants2011}.  In contrast to previous generative
-modeling approaches for multi-modal tumor segmentation 
-(e.g. \cite{prastawa2003,zikic2012}), we do not use multivariate 
-Gaussians to specify tissue probabilities but rather incorporate each
-univariate probability map into the feature vector of the training
-data.  As pointed out in \cite{menze2010}, multivariate modeling
-might obscure the distinct biological information provided by each 
-modality.  Instead, we let the random forest construction 
-process determine the optimal combination of such multivariate
-information.
-Additionally, maximum posterior labeling from the GMM processing
-is used to determine the connected components for each label.  
-Geometric features (assigned voxel-wise) include the physical volumes 
-of each connected component including the volume to surface area ratio, 
-the elongation, and eccentricity.
-
-In order to better characterize deviations from normal
-multi-modal brain shape and appearance, several features were derived 
-using population-specific multivariate template 
-construction. A recent neuroimaging reproducibility study
-by Landman et al. resulted in an open data cohort of 21
-normal individuals, each imaged twice, comprising several
-modalities including arterial spin labeling, 
-fluid attenuated inversion recovery (FLAIR),
-diffusion tensor imaging, functional imaging, T1, and T2 
-\cite{landman2011}.
-
-Given $K$ image modality types for $N$ subjects,  
-${\mathbf I} = \{I_1,I_2,\ldots, I_K\}$, multivariate 
-template construction iterates between optimizing the set 
-of diffeomorphic transforms between the subjects and the 
-template, 
-$\left\{\left(\phi_1,\phi_1^{-1}\right),\ldots,\left(\phi_N,\phi_N^{-1}\right)\right\}$ 
-and constructing the 
-optimal multivariate template appearance 
-$\mathbf{J}=\{J_1,J_2,\ldots, J_K\}$ to minimize the
-following cost function:
-\begin{align}
-  \sum_{n=1}^N 
-        \Bigg[ D &\left( \psi(\mathbf{x}),\phi_1^n(\mathbf{x},1)\right) \\ \nonumber 
-        +& \sum_{k=1}^K \lambda_k \Pi_k \left(I_k^n\left(\phi_n(\mathbf{x},0.5)\right),J_k\left(\phi^{-1}_n(\mathbf{x},0.5)\right)\right)\Bigg]
-\end{align}
-where $D$ is the diffeomorphic shape distance,
-\begin{align}
-D\left( \phi( \mathbf{x},0),\phi( \mathbf{x},1)\right) = \int_0^1 \| \nu(\mathbf{x},t)\|_L dt
-\end{align}
-dependent on the choice of linear operator, $L$, and $\nu$
-is the velocity field
-\begin{align}
-\nu\left( \phi(\mathbf{x},t) \right) = \frac{d\phi(\mathbf{x},t)}{dt},\,\,\, \phi(\mathbf{x},0) = \mathbf{x}.
-\end{align}
-Each pairwise registration employing the similarity metric $\Pi_k$ can 
-be assigned a relative weighting, $\lambda_k$, to weight a particular
-modality's influence in the construction process.  Further theoretical
-details can be found in \cite{avants2008,avants2010}.
-In terms of implementation, this algorithm is 
-encapsulated in the script \verb#antsMultivariateTemplateConstruction.sh#,
-available in the ANTs repository, which permits parallel processing on
-an individual workstation or on a large computational cluster.
-
-\begin{figure}
-  \centering
-  \begin{tabular}{cc}
-    \includegraphics[width=40mm]{Figures/S_templateFA_140.png} &
-    \includegraphics[width=40mm]{Figures/S_templateFLAIR_140.png} \\
-    (a) & (b) \\
-    \includegraphics[width=40mm]{Figures/S_templateT1_140.png} &
-    \includegraphics[width=40mm]{Figures/S_templateT2_140.png} \\
-    (c) & (d) 
-  \end{tabular}
-  \caption{Multivariate symmetric template created from the Kirby 
-           21 data described in \cite{landman2011}.  Shown are the
-           (a) fractional anisotropy (FA), (b) FLAIR, (c) MPRAGE, 
-           and (d) T2 template components.
-          }
-  \label{fig:symmetrictemplates}
-\end{figure}
-
-\subsubsection{\underline{Bra}in \underline{T}umor \underline{S}egmentation Challenge Data}
-
-The Brain tumor segmentation
-Associated with the 2012 International Conference on Medical Image Computing and Computer Assisted
-Intervention (MICCAI),%
-\footnote{
-http://www.miccai2012.org
-}
-the
-
-\subsubsection{MS lesion segmentation challenge Challenge Data}
-
-\section{Results}
-
-\begin{table*}
-\caption{Dice scores from the MICCAI 2012 BRATs Study}
-\begin{center}
-\begin{tabular*}{0.975\textwidth}{@{\extracolsep{\fill} } c c c c c c c c c}
-\toprule
-{} & \multicolumn{2}{c}{High-grade (real)} & \multicolumn{2}{c}{Low-grade (real)} & \multicolumn{2}{c}{High-grade (simulated)} & \multicolumn{2}{c}{Low-grade (simulated)}\\
-{\bf Method} & Edema & Tumor & Edema & Tumor & Edema & Tumor & Edema & Tumor\\
-\midrule
-\cite{zikic2012} & {$0.70 \pm 0.09$} & {$0.71 \pm 0.24$} & {$0.44 \pm 0.18$} & {$0.62 \pm 0.27$} & {$0.65 \pm 0.27$} & {$0.90 \pm 0.05$} & {$0.55 \pm 0.23$} & {$0.71 \pm 0.20$} \\
-\cite{bauer2012} & {$0.61 \pm 0.15$} & {$0.62 \pm 0.27$} & {$0.35 \pm 0.18$} & {$0.49 \pm 0.26$} & {$0.68 \pm 0.26$} & {$0.90 \pm 0.06$} & {$0.57 \pm 0.24$} & {$0.74 \pm 0.10$} \\
-\rowcolor{yellow}
-ANTsR & {$0.64 \pm 0.18$} & {$0.65 \pm 0.30$} & {$0.48 \pm 0.16$} & {$0.69 \pm 0.21$} & {$0.75 \pm 0.25$} & {$0.94 \pm 0.04$} & {$0.65 \pm 0.25$} & {$0.83 \pm 0.09$}  \\
-\bottomrule
-\end{tabular*}
-\end{center}
-\end{table*}
-
-\section{Discussion and Conclusions}
-
-%% References with bibTeX database:
-
-\section*{Acknowledgments}
+  We are concerned with the latter two.
+  \item How does one avoid confirmation bias?  If I'm comparing my
+  algorithm to somebody else's algorithm, I'm going to be naturally
+  predisposed to confirmatory evidence that my algorithm is better
+  while neglecting (most likely not with intent of doing so) evidence
+  that disconfirms my original belief (that my algorithm is better).
+  Richard Feynman quote (Cargo Cult Science):  
+  \begin{quote} 
+  The first principle [of science] is that you must not fool yourself—--and you are the easiest person to fool.
+So you have to be very careful about that.
+After you've not fooled yourself, it's easy not to fool other scientists.
+You just have to be honest in a conventional way after that.
+  \end{quote}
+  \item One particular aspect of the previous item is that it needs 
+  to be highlighted if somebody codes up the algorithm
+  to be compared themselves.  Writing bug-free code is difficult.  Look
+  at how many bug fixes are provided by the ITK community on a daily
+  basis.  
+  \item Also, is simply saying that one algorithm is better than 
+  another sufficient?  Should there be some theoretical explanation
+  for it.  For example, when I wrote my DMFFD image registration paper,
+  the comparison wasn't IRTK vs. DMFFD but rather FFD vs. DMFFD and
+  there were theoretical reasons why performance should be better
+  with the latter.  Also, I insisted to the reviewer on not using IRTK 
+  for comparison since there are also so many other performance-related
+  issues (type of interpolation, gradient step, metric, metric implementation, etc.) which would confound
+  comparisons.  Also, it probably helped that the theme had a more specific,
+  more verifiable focus, i.e. DMFFD produces a more efficient energy minimizer
+  since it acts as a preconditioner in the standard gradient descent
+  optimization.
+  \item Comparisons are best made with publicly available data and
+  results of the comparisons should be made available.
+  \item Operating system needs to be defined.  Freesurfer results
+  varied with MacOSx.
+  \item Ideally, the authors of the new algorithm would work
+  with the authors of the compared algorithm.  Given human
+  nature, this type of cooperation is not guaranteed.  However,
+  a minimum set of information is needed.
+  \item Arno Klein as an example of somebody who did it right.  Takes a
+  lot of work.
+  \item 
+  \item The source of the package software needs to be defined.  For 
+  example, N4 has been
+  instantiated in Slicer, ANTs, and c3d (also might be implemented
+  in one of Styner's NITRC projects).  However, each might use
+  different default parameters and have other tweaks which effects
+  performance.  In Vladimir Fonov's github repository containing 
+  various processing scripts for MINC, one can see from the history
+  how N4 was used but then the users switched back to N3MNI (due to 
+  performance issues?).  However, they used N4 out of the c3d package
+  which the original authors (N.T./B.A.) haven't touched in three years
+  so the parameters aren't optimal (shrink factors = 4, spline distance
+  = 100 with 3 levels: $100\times50\times50$).  Changing these parameters 
+  (which are crucial to performance) isn't
+  accessible to the user from c3d.  
+  \item Great resources for code-sharing
+  	\begin{itemize}
+		  \item github
+		  \item NITRC
+		  \item ITK
+		\end{itemize}
+\end{itemize}
 

### 2013-05-05 Tilting/paper.tex
WIP: notes on this paper

@@
 
-\usepackage{color}
+\usepackage{color,hyperref}
 \usepackage{multirow,booktabs,ctable,array}
@@
   accessible to the user from c3d.
+
+  \item Collaborators should help verify findings by repeating a
+(representative subset of a) study independently, from scratch,
+preferably on another computer perhaps even with a different operating
+system.  A scientist can do this him/herself as well.  Personal
+experience suggests that this procedure will often uncover coding
+errors sometimes simply by forcing one to reread code or scripts again
+or clarify documentation.
+
+  \item
+    \textcolor{blue}{\href{http://en.wikipedia.org/wiki/Accuracy_and_precision}{Accuracy
+        and precision (link)}} --- most studies ignore precision.  it
+    is critical in longitudinal studies.   precision is impacted by
+    parameter choices as well as computational architecture
+    (Freesurfer paper)
+
+   \item parameters impact accuracy - e.g. clarkson's paper vs our
+     recent tuning of DiReCT parameters w.r.t. expected thickness values.
+
+  \item  Biological plausibility is important---a more accurate method
+    may produce less biologically plausible results.
+    Cross-referencing with other fields is critical e.g. patterns of
+    MRI atrophy should match pathology distribution.  There is an art
+    to this, of course.
+
+  \item Prediction---an algorithm that yields a ``more significant''
+    result is not always better.  might mean ``finds more effected
+    areas'' or smaller p-values.  Testing prediction is complements
+    studies of significance and cumulative distribution functions.
+
+  \item Verification is supported by open science technology for
+    sharing data and code.
+
+   \item Great resources for data-sharing:
+  	\begin{itemize}
+	   \item figshare
+           \item MIDAS 
+           \item slicedrop
+           \item R 
+        \end{itemize}
+
   \item Great resources for code-sharing

### 2013-05-06 Tilting/paper.tex
WIP: notes on this paper
WIP: illustrate what we have in mind by an example
ENH:  Paramore new album

@@
 
-\usepackage{color,hyperref}
+\usepackage{color}
 \usepackage{multirow,booktabs,ctable,array}
@@
 \begin{abstract} 
-Exploration of neuroscience hypotheses have been greatly enhanced
+Exploration of neuroscience hypotheses have been  enhanced
 by the increased availability of high-performance computational resources, large-scale
@@
 
-%\input{intro}
-%
-%\input{methods}
+\section{Introduction}
+The neuroimaging community has benefited from the proliferation
+of imaging software.  Established packages such as SPM from the 
+Wellcome Trust Centre for Neuroimaging \cite{Ashburner2012},
+the FMRIB Software Library (FSL) \cite{Jenkinson2012}, and
+the AFNI toolkit \cite{cox2012} have aided neuroimaging researchers 
+around the world in performing complex analyses as part of 
+ongoing neuroscience research.  Parallel to this line of
+inquiry is continued algorithmic innovation for improvement
+in analysis tools.
 
-\section{Introduction}
+As fellow scientists who actively participate in this type of 
+research, we have noticed several publications in which a principle
+component involves the comparison of algorithms.  One of our
+concerns is the lack of detail with which these comparisons are
+presented and the corresponding possibility of {\it instrumentation
+bias} \cite{sackett1979} (considering software  as a type of 
+instrument requiring proper ``calibration'' for accurate measurements).  
+In this editorial we propose an initial set of guidelines for minimal
+reporting of algorithmic usage to minimize such bias understanding that 
+the discussion will hopefully 
+elicit a much more comprehensive response from the larger community.
+
+Although previous articles have discussed similar concerns, oftentimes
+within a much larger context (e.g. fMRI reporting \cite{poldrack2008}), 
+we are particularly interested in comparative evaluations of software
+and provide additional discussion through experience and actual 
+examples found within the literature.
+It is hoped that this commentary serves to raise awareness to both 
+authors and reviewers to such problematic issues (not unlike other
+recent articles detailing additional potential sources of methodological 
+bias \cite{kriegeskorte2009,vul2012,tustison2012}).
+
+\section{Guidelines}
+
+\subsection{Parameters should be included (perhaps the precise command line call)}
+Perhaps the most blatantly egregious sin is omitting
+
+\subsection{Hassle the developers for a set of default parameters}
+
+\subsection{Scripts used to invoke the algorithms should be posted}
+
+\subsection{Comparisons should be performed on publicly available data}
+
+\subsection{Corollary to the above:  Resulting images should be publicly posted}
+
+\subsection{Scripts used to create the plots should also be posted}
+
+\subsection{When possible, consultation should be made with the original authors}
+
+\subsection{Provide details of where the algorithm was obtained}
+
+\subsection{Provide system details of where the algorithm was run}
+
+\subsection{Avoid confirmation bias:  Don't code up your own version particularly if one
+is already available}
+
+\subsection{Co-authors should verify findings}
+
+\subsection{Properly contextualize comparisons}
+Accuracy is important but other considerations should be included such
+as biological plausibility.  Torsten's CURT algorithm could produce an ``accurate''
+registration but not one which is biologically plausible.
+
+\section{Conclusion}
+
+\section{Acknowledgments}
 
@@
   \item
-    \textcolor{blue}{\href{http://en.wikipedia.org/wiki/Accuracy_and_precision}{Accuracy
-        and precision (link)}} --- most studies ignore precision.  it
+    \textcolor{blue}%{\href{http://en.wikipedia.org/wiki/Accuracy_and_precision}{Accuracy and precision (link)}} 
+        wiki link--- most studies ignore precision.  it
     is critical in longitudinal studies.   precision is impacted by
@@
     result is not always better.  might mean ``finds more effected
-    areas'' or smaller p-values.  Testing prediction is complements
+    areas'' or smaller p-values.  Testing prediction complements
     studies of significance and cumulative distribution functions.

### 2013-05-08 ANTsRS/paper.tex
ENH:  Added figure.
ENH:  Amy

@@
 
-\title{Multivariate Neuroanalysis with ANTsR:  Application to Supervised Brain Segmentation}
+\title{Multivariate Neuroanalysis with ANTsR:  Application to Supervised Brain Segmentation with Concatenated Random Forests}
 
@@
 
-O
+ANTs (Advanced Normalization Tools) originated with the open source availing
+of state-of-the-art registration algorithms for neuroimage analysis
+\cite{avants2008a} built upon the mature and well-vetted Insight Toolkit
+of the National Institutes of Health.  Since then, the toolkit has grown to include 
+several algorithmic solutions necessary
+for robust medical image analysis including bias correction \cite{tustison2010}, 
+$n$-tissue multivariate segmentation \cite{avants2011}, template construction \cite{avants2010}, 
+and cortical thickness estimation \cite{das2009} (many of which have been
+introduced into ITK partially in an attempted leveraging of Linus's Law).%
+\footnote{
+``Given enough eyeballs, all bugs are shallow.'' --Linus Torvalds
+}   
+However, in the evolution of the toolkit, it became clear (as neuroimaging
+research certainly falls under the popular rubric of ``big data analysis'')
+that robust statistical machinery was lacking for proper inferences regarding
+the data produced by the various ANTs tools.
+ANTsR was developed
+specifically to provide an interface between a 
+powerful neuroimaging toolkit for producing reliable imaging data 
+transformations and the R project%
+\footnote{
+http://www.r-project.org
+}
+for statistical computing and visualization thus providing a complete
+set of tools for multivariate image analysis.
+
+In addition to describing ANTsR basics and how it can be generally
+used in multivariate neuroimaging studies, we showcase its use in
+a particularly salient application for performing supervised brain
+segmentation using random forests.  Although random forests have
+been proposed previously in the literature for supervised brain 
+segmentation (e.g. \cite{geremia2011,zikic2012}), we demonstrate that ANTsR
+significantly facilitates the construction of a fully functional, 
+parallelizable workflow for such a task and that performance 
+exceeds that of the top competitors in the recent Multimodal Brain
+Tumor Segmentation Challenge as part of the MICCAI 2012 conference.%
+\footnote{
+http://www2.imm.dtu.dk/projects/BRATS2012/  
+}
 
@@
 
+Given a set of training data consisting of 
+labeled (csf, gray matter, white matter, edema, and tumor),
+multimodal brain image data, 
+supervised brain segmentation entails the generation of a 
+statistical model from such training data which can
+then be extended to testing data, i.e. unlabeled 
+brain data.  In subsequent sections, we introduce the generic
+{\it random forest} modeling framework which takes as input
+a set of {\it feature images} (also described in a later section)
+and labels for each training subject 
+and outputs a statistical model.  This model, in turn, can then
+be used to provide a probabilistic estimate of the labels in an
+unlabeled subject.
+
+One of the core extensions that we provide in this work is
+the use of concatenated random forests for improved probabilistic estimation
+of the labels.
+As we demonstrate in the Results section, the set of feature
+images employed work sufficiently well for good performance
+on the training data (which exceeds that of what has been
+currently published in the literature).  However, we discovered
+that further improvements could be gained by using the probabilistic 
+label estimates as input to enhance preprocessing before a 
+second round of feature image generation including 
+modality-specific, prior-based $n$-tissue segmentation and 
+random forest model creation.
+
 \subsubsection{Random Forests}
 
-Building on the developments of Ho \cite{ho1995} and Amit and Geman \cite{amit1997},
-Breiman introduced Random Forests for classification/regression \cite{breiman2001} which performs relatively well against other machine learning strategies such as support vector machines and neural networks.
+Several machine learning concepts were integrated to create 
+the random forests framework first articulated in its entirety by Breiman
+et al. \cite{breiman2001} for performing classification/regression.  
+Although decision trees had been previously explored in the literature, 
+it was the success of ``boosting''-style machine learning 
+techniques, such as AdaBoost \cite{schapire1990,freund1997}, which influenced 
+the aggregation of such decision trees into ``forests'' 
+with randomized node optimization for improved
+classification/regression performance \cite{ho1995,amit1997}.
+The final element of bootstrap aggregating or ``bagging'' (i.e.
+random sampling of the training data) was
+introduced by Breiman \cite{breiman1996} to achieve improved
+accuracy.
 
-\subsubsection{Multi-Modality Feature Image Preprocessing}
+Early adoption \cite{viola2005} and success in the
+computer vision community
+has led to a recent surge within the medical image analysis
+community of using random forests for handling complex 
+classification/regression tasks including
+normal brain segmentation \cite{yi2009},
+MS lesion segmentation \cite{geremia2011}, 
+multimodal brain tumor segmentation
+\cite{zikic2012,geremia2012}, brain extraction \cite{iglesias2010}, 
+anatomy detection in computed tomography \cite{criminisi2013}, and
+segmentation of echocardiographic images \cite{verhoek2011}.
+A thorough introduction for those interested in delving deeper 
+into the more theoretical aspects of random forests can be found
+in \cite{criminisi2011}.
+
+One of the principal advantages of R is the extensive community of
+developers  who have contributed on the order of thousands of packages 
+extending R's capabilities beyond its core functionality.
+Most relevant for the work described in this paper
+is the \verb#randomForest# package developed from Breiman's original
+Fortran code by Liaw and Wiener (described briefly in \cite{liaw2002}).
+
+\subsubsection{Multi-Modal Feature Image Preprocessing}
 
@@
 with the training data illustrated a degradation in performance
-when one or both steps (using \cite{nyul2000} and \cite{tustison2010},
+when one or both steps (using \cite{nyul2000} and N4 \cite{tustison2010},
 respectively) were performed due to the presence of the tumor/edema complex.
-Instead we simply windowed the image intensity
+
+As a corrective, for the first stage we simply windowed the image intensity
 for all images to be between the quantiles $[0.01,0.99]$ and
-subsequently rescaled to $[0,1]$.  This strategy seemed to provide 
-the best classification (probably due to the fact that the intensity
-corruption due to bias was relatively minimal in the data used).
-However, improved results might be obtained
-by a more sophisticated strategy in which an initial supervised segmentation 
-is used to find the probabilistic estimates of the normal tissue
-locations (as done in this work) which could then be followed by using
-only the probabilistic maps of the normal tissues to construct a
-weight, or confidence, mask for estimating the inhomogeneity bias
-using N4 \cite{tustison2010} followed by a repeated application of 
-the supervised segmentation strategy described.  Note that a similar
-iterative-based strategy until convergence is implemented in the ANTs script \verb#antsAtroposN4.sh#.
+subsequently rescaled to $[0,1]$.  From these ``corrected'' images,
+the first set of feature images were derived.  For the training cohort, these
+data were used to create the random forest regression model for the first
+stage.  During the second stage, the probabilistic estimates
+of the white matter and gray matter labels were used to generate a
+``pure tissue weight mask'' to estimate the bias field 
+using N4 (although the resulting bias field estimation was used
+to correct the image within the entire cerebral mask).  Formally, this 
+involved generation of a probabilistic map defined as:
+\begin{align}
+  P_{pure\,\,tissue}(\mathbf{x}) = \sum_{i=1}^N P_i(\mathbf{x}) \prod_{j=1, j \neq i}^N \left( 1 - P_j(\mathbf{x}) \right)
+\end{align}
+where $N$ is the set of user-selected tissue labels (in our
+case $N=2$ consisting of the gray and white matter probability
+maps).
 
-\subsubsection{Multi-Modality Feature Image Generation}
+Both rescaling and weighted bias correction were applied to produce
+the ``corrected images'' for the second stage resulting in
+modified features images for the second stage.  Note that we
+perform a similar iterative scenario for normal brain 
+segmentation \cite{avants2011} (encapsulated in the ANTs script 
+\verb#antsAtroposN4.sh#).
+
+\subsubsection{Multi-Modal Feature Image Generation}
 
@@
 to showcase the supervised segmentation strategy developed in this work.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=180mm]{Figures/featuresImages.pdf}
+  \caption{Feature images from the BRATS\_HG0004 data set.}
+\end{figure*}
 
@@
     \end{itemize}
-  \item Misc.
+  \item Not modality specific
     \begin{itemize}
@@
     \item Log Jacobian  
+    \item T1C,T1 difference image
     \end{itemize}
@@
 based on connected components, and two difference images using symmetric template
-construction for a total of 4 modalities $\times$ (4 + 5 + 4 + 2) feature images per modality $=$ 60 images.  We employ two additional images consisting of the 
+construction for a total of 4 modalities $\times$ (4 + 5 + 4 + 2) feature images per modality $=$ 60 total
+feature images.  We employ two additional images consisting of the 
 Euclidean distance image \cite{maurer2003} created from the skull-stripped 
@@
 \cite{bauer2012} & {$0.61 \pm 0.15$} & {$0.62 \pm 0.27$} & {$0.35 \pm 0.18$} & {$0.49 \pm 0.26$} & {$0.68 \pm 0.26$} & {$0.90 \pm 0.06$} & {$0.57 \pm 0.24$} & {$0.74 \pm 0.10$} \\
-\rowcolor{yellow}
-ANTsR & {$0.64 \pm 0.18$} & {$0.65 \pm 0.30$} & {$0.48 \pm 0.16$} & {$0.69 \pm 0.21$} & {$0.75 \pm 0.25$} & {$0.94 \pm 0.04$} & {$0.65 \pm 0.25$} & {$0.83 \pm 0.09$}  \\
+ANTsR & {$0.65 \pm 0.15$} & {$0.66 \pm 0.28$} & {$0.49 \pm 0.16$} & {$0.65 \pm 0.21$} & {$0.68 \pm 0.25$} & {$0.91 \pm 0.08$} & {$0.61 \pm 0.25$} & {$0.84 \pm 0.09$} \\
+w/ Atropos & {$0.68 \pm 0.15$} & {$0.67 \pm 0.30$} & {$0.50 \pm 0.15$} & {$0.67 \pm 0.23$} & {$0.74 \pm 0.26$} & {$0.92 \pm 0.09$} & {$0.65 \pm 0.26$} & {$0.84\pm 0.08$} \\
 \bottomrule

### 2013-05-08 KapowskiChronicles/intro.tex
ENH:  Amy

@@
 in identifying the relationship between cortical morphology, disease and cognition.
-Quantitative cortical measures are able to identify abnormalities in such conditions as 
+Quantitative cortical measures have been demonstrated in conditional 
+abnormalities such as 
 Huntington's disease \citep{rosas2002,rosas2005,selemon2004}, 
@@
 Most relevant to this work is the DiReCT (Diffeomorphic Registration-based 
-Cortical Thickness) algorithm proposed by \cite{das2009} in which 
-a registration-based approach proposed where the derived diffeomorphic mapping between the 
-white and pial matter surfaces is used to propagate thickness values 
+Cortical Thickness) algorithm proposed in \cite{das2009} where generated
+diffeomorphic mappings between the 
+white and pial matter surfaces are used to propagate thickness values 
 through the cortical gray matter.  A unique benefit of DiReCT is that it
@@
 from imaging data (of which
-only a fraction are cited), several common pre-processing components
+only a fraction are cited), several common preprocessing components
 may be identified.
@@
 across large populations, construction of population-specific unbiased templates
-is also beneficial \cite{evans2012}.
-In addition, some of these steps might include a crucial registration component, e.g. 
-propagating template-based tissue priors for improved segmentation.  The requisite
-additional components coupled with the general lack of availability of published
-algorithms \cite{kovacevic2006} inhibits performing studies by external researchers 
+is also potentially beneficial \cite{evans2012}.
+In addition, intermediate steps might include a crucial registration component, e.g. 
+propagating template-based tissue priors for improved segmentation.
+
+  The requisite
+additional components apart from the actual cortical thickness estimation, 
+coupled with the general lack of availability of published
+algorithms \cite{kovacevic2006}, inhibits performing studies by external researchers 
 and makes comparative evaluations difficult.  For example, one recent evaluation 

### 2013-05-08 KapowskiChronicles/methods.tex
ENH:  Amy

--- 2013-02-08 KapowskiChronicles/methods.tex
@@
   \item brain extraction using a hybrid segmentation/template-based strategy,
-  \item alternating between prior-based segmentation and white matter posterior
+  \item alternation between prior-based segmentation and white matter posterior
         probability weighted bias correction,
@@
 Additionally, the coordination of all the algorithmic components is
-encapsulated in the shell script \verb#antsCorticalThickness.sh#.  This includes
+encapsulated in the shell scripts \verb#antsCorticalThickness.sh# with
+subcomponents delegated to \verb#antsBrainExtraction.sh# 
+and \verb#antsAtroposN4.sh#.  This includes
 optimal parameters for each of the algorithmic components which has worked
@@
   \includegraphics[width=170mm]{Figures/Kapowski_pipeline2.pdf}
-  \caption{The ANTs T1 processing workflow containing all elements for 
-  determining cortical thickness. Not shown is the optional single subject
-  to template registration.}
+  \caption{Illustration of the main components of the ANTs processing 
+  workflow containing all elements for determining cortical thickness. 
+  Not shown is the optional subject to template registration.}
   \label{fig:pipeline}
@@
 Normalization (SyGN) is a powerful framework for producing optimal population-specific
-templates \citep{avants2010} with arbitrary similarity metric choice.
+templates \citep{avants2010}.
 
@@
 The ANTs implementation of this technique is currently available as a shell script, 
-\verb#buildtemplateparallel.sh#, and a multivariate version,
-\verb#antsMultivariateTemplateConstruction.sh#, both of which are distributed as part of
+\verb#buildtemplateparallel.sh# (a multivariate version is also available as
+\verb#antsMultivariateTemplateConstruction.sh#).  Both scripts are distributed as part of
  the ANTs repository.
 The latter script permits the construction of multimodal templates (e.g. 
-T1-weighted, T2-weighted, and proton density MRI as described in the 
-Evaluation section).  Both scripts accommodate a variety of computational resources
+T1-weighted, T2-weighted, proton density MRI and fractional anisotropy).  
+Both scripts accommodate a variety of computational resources
 for facilitating template construction.  These computational resource possibilities include:
@@
 \end{itemize}
-Within this work multiple templates were created for all stages of 
-image processing and analysis.  The creation of these templates are described
-in the corresponding data section.
 
-\textcolor{blue}{what is the motivation of the multivar template here?
-  it would seem useful to add an FA component to the segmentation step
-  to aid in cortical / wm delineation which would motivate the
-  multivar template approach but might be difficult to establish that
-  this actually helps .... though we could compare both ways : univar
-  and multivar }
+For this work, database-specific templates were used during cortical thickness pipeline
+processing for both brain extraction and brain segmentation steps.  Multivariate templates 
+were constructed from the multimodal data sets.  However, their usage was based on the 
+fact that they had been built previously for other work and not because they provide a 
+discernible advantage over univariate templates (i.e. T1-only) for the proposed workflow.
 
@@
 
-In \cite{tustison2010}, we introduced an extension of N3, denoted as
-N4, which demonstrates improved performance and convergence behavior
+In \cite{tustison2010}, we introduced an improvement of N3, denoted as
+``N4'', which demonstrates a significant increase in performance and convergence behavior
 on a variety of data.  This improvement is a result of an enhanced 
@@
 formulation.  For our workflow, the additional possibility of specifying
-a weighted mask in N4 permits the use of the current white matter probability map 
+a weighted mask in N4 permits the use of a ``pure tissue'' probability map 
+(described below)
 calculated during the segmentation pipeline for further improvement of 
@@
 the background from the foreground using Otsu's algorithm \cite{otsu1979}.
-Following brain extraction, the three-tissue segmentation involves iterating
-between bias field correction using the current white matter posterior 
-probability as a weight mask and then using that bias corrected image
-as input to the Atropos segmentation step (described in subsequent sections).
+Following brain extraction, three-tissue segmentation involves iterating
+between bias field correction using the current pure tissue 
+probability map as a weight mask and then using that bias corrected image
+as input to the Atropos segmentation step (described in the next section).
 
@@
 images, and alternative posterior formulations for different segmentation tasks.
+
+Due to the important interplay between segmentation and bias correction,
+we perform multiple N4 $\Leftrightarrow$ Atropos iterations.
+In order to better integrate Atropos and N4, we use  
+a pure tissue probability weight mask generated from the 
+posterior probabilities generated from the segmentation 
+process.  Given $N$ labels and the corresponding $N$
+posterior probability maps $\{ P_1, \ldots, P_N\}$ produced
+during the segmentation process, the N4 weight mask is 
+created at each N4 $\Leftrightarrow$ Atropos iteration from
+\begin{align}
+  P_{pure\,\,tissue}(\mathbf{x}) = \sum_{i=1}^N P_i(\mathbf{x}) \prod_{j=1, j \neq i}^N \left( 1 - P_j(\mathbf{x}) \right).
+\end{align}
+One of the key insights of the original N3 algorithm is the
+observation that inhomogeneities cause the intensity values of
+pure tissue peaks to spread in the intensity histogram as though
+convolved with a Gaussian.  A core contribution of N3 is the
+proposed corrective of deconvolving the intensity histogram to 
+accentuate the tissue peaks coupled with a spatial smoothing 
+constraint. The pure tissue probability mask
+weights more heavily the voxels corresponding to pure tissue 
+types (as determined by the segmentation) during the deconvolution process 
+while minimizing the contribution of regions such as the gray/white matter 
+interface where peak membership is ambiguous.
 
@@
 
-\subsubsection{DiReCT Cortical Thickness Estimation}
+\subsubsection{DiReCT (aka KellySlater/KellyKapowski) Cortical Thickness Estimation}
 
-Although the basic formulation of DiReCT as reported in this work is as it 
-was introduced in \cite{das2009}, we have made several 
-improvements.  Perhaps the most significant advance is that this particular 
-ITK-compatible implementation has been significantly multi-threaded,
-is written in ITK coding style, and has been made publicly available through 
-ANTs complete with a unique user interface design developed specifically for 
-ANTs tools.
+DiReCT was introduced 
+in \cite{das2009} and made available in ANTs with the program \verb#KellySlater#.
+Since then several improvements have been made and incorporated into the program
+\verb#KellyKapowski#.%
+\footnote{
+Traditional academic discourse encountered in the published literature
+rarely contextualizes peculiarities such as algorithmic nomenclature.
+We briefly mention that
+this was the source of a rare disagreement between the first two authors
+based, as many disagreements are, on a simple misunderstanding and not an
+affronting existential statement concerning a certain favorite sitcom
+of the author's youth. 
+}
+Among the most significant advancements is that the more recent
+implementation is multi-threaded, written in rigorous ITK coding style, and 
+has been made publicly available through  ANTs complete with a unique user 
+interface design developed specifically for ANTs tools.
 
@@
 
-\subsubsection{LPBA40 Data for Skull Stripping}
+%\subsubsection{LPBA40 Data for Skull Stripping}
+%
+%For the brain extraction step
+%we used the data from the LPBA40 repository \citep{shattuck2008}.
+%These data consist of 40 high-resolution 3D Spoiled Gradient Echo
+%(SPGR) MRI acquisitions which were manually labeled delineating
+%56 brain structures.  Additional post processing 
+%included automated brain extraction using FSL's brain extraction tool 
+%\citep{smith2002} which was followed by manual corrections.  These
+%40 brain masks are also included with the database.  
+%
+%All 40 subjects were used to create a population-specific
+%unbiased average template \cite{avants2010}.  The brain masks corresponding 
+%to the 20 subjects were warped to the template space using the 
+%transforms derived during the template building process.  A template 
+%probability mask was created by averaging the warped brain masks.
+%For brain extraction of any single individual this ANTS-based LPBA40 template 
+%is coarsely registered to the single subject brain.  The template
+%probability brain mask is warped to the individual subject and 
+%is used as the initial brain mask estimate.  As mentioned previously,
+%Atropos and binary morphological operations are used to refine the
+%brain mask estimate.
 
-For the brain extraction step
-we used the data from the LPBA40 repository \citep{shattuck2008}.
-These data consist of 40 high-resolution 3D Spoiled Gradient Echo
-(SPGR) MRI acquisitions which were manually labeled delineating
-56 brain structures.  Additional post processing 
-included automated brain extraction using FSL's brain extraction tool 
-\citep{smith2002} which was followed by manual corrections.  These
-40 brain masks are also included with the database.
+\subsubsection{NIREP Data for Cortical Labelings}
 
-All 40 subjects were used to create a population-specific
-unbiased average template \cite{avants2010}.  The brain masks corresponding 
-to the 20 subjects were warped to the template space using the 
-transforms derived during the template building process.  A template 
-probability mask was created by averaging the warped brain masks.
-For brain extraction of any single individual this ANTS-based LPBA40 template 
-is coarsely registered to the single subject brain.  The template
-probability brain mask is warped to the individual subject and 
-is used as the initial brain mask estimate.  As mentioned previously,
-Atropos and binary morphological operations are used to refine the
-brain mask estimate.
-
-\subsubsection{NIREP Data for 3-Tissue Segmentation}
-
-The nonrigid image registration evaluation project is an ongoing 
-framework for evaluating image registration algorithms \citep{christensen2006}.
+The Nonrigid Image Registration Evaluation Project (NIREP%
+\footnote{
+http://www.nirep.org/
+}) 
+is an ongoing framework for evaluating image registration algorithms \citep{christensen2006}.
 The initial data set introduced into the project consists of 
 16 (8 male and 8 female) high resolution skull-stripped brain 
-data with 32 cortical labels (cf Table \ref{table:nirep_labels}) manually drawn using a 
-published protocol.%
-\footnote{
-http://www.nirep.org/
-}
-Given the gray matter labels, the white matter and CSF were identified 
-for each of the 16 subjects using Atropos.  Similar to the LPBA40
-data set, a NIREP template was created from all 16 subjects and each of the
-warped labels were used to create probabilistic estimates of the 
-labeled region boundaries. These probability maps were used as 
-spatial prior probabilities during the 3-tissue segmentation component
-of the pipeline.  Using SyN, the NIREP template is registered to the
-extracted individual subject brain which is followed by a warping of the 
-NIREP priors to the space of the individual subject.  The initial warped 
-white matter probability map is used as the weighted confidence mask 
-in the follow-up bias correction step.
+data with 32 cortical labels (cf. Table \ref{table:nirep_labels}) manually drawn using a 
+published protocol.
+
+Cortical label propagation to each individual subject for all data sets
+described below is performed using the PICSL multi-atlas joint label fusion
+algorithm described in \cite{wang2013} which recently topped the competition
+at the MICCAI 2012 Grand Challenge on Multi-Atlas Labeling and is also 
+distributed with the ANTs toolkit.  .
+
+%Given the gray matter labels, the white matter and CSF were identified 
+%for each of the 16 subjects using Atropos.  Similar to the LPBA40
+%data set, a NIREP template was created from all 16 subjects and each of the
+%warped labels were used to create probabilistic estimates of the 
+%labeled region boundaries. These probability maps were used as 
+%spatial prior probabilities during the 3-tissue segmentation component
+%of the pipeline.  Using SyN, the NIREP template is registered to the
+%extracted individual subject brain which is followed by a warping of the 
+%NIREP priors to the space of the individual subject.  The initial warped 
+%white matter probability map is used as the weighted confidence mask 
+%in the follow-up bias correction step.
 
@@
 
-\subsection{IXI Data for Pipeline Evaluation}
+\subsubsection{Data for Pipeline Evaluation}
 
-The IXI data%
+\paragraph{IXI} The IXI data%
 \footnote{
@@
 \end{figure}
+
+\paragraph{Oasis}

### 2013-05-08 KapowskiChronicles/paper.tex
ENH: Ozzy
ENH:  Amy

@@
 hypotheses.  In this paper, we introduce the volume-based 
-Advanced Normalization Tools
-(ANTs) cortical thickness automated pipeline
+Advanced Normalization Tools (ANTs) cortical thickness automated pipeline
 comprising well-vetted components such as SyGN (multivariate template construction),
-SyN (image registration), N4 (bias
-correction),
-Atropos ($n$-tissue segmentation), and DiReCT (cortical thickness) 
-all developed as part of the ANTs open science effort.  
-Complementing the open source aspect of ANTs we confirm 
-previous findings of
-gender differences, age correlation, and hemispherical asymmetry 
-in cortical thickness using the publicly available IXI data
-consisting of multimodal structural MRI of approximately 600 subjects.
+SyN (image registration), N4 (bias correction), Atropos ($n$-tissue segmentation), 
+and DiReCT (cortical thickness) all developed as part of the ANTs open science effort.  
+Complementing the open source aspect of ANTs we investigate several hypotheses
+explored previously in the literature using four different publicly 
+available data sets totaling approximately 1200 subjects.  
 To further promote open science and use of the proposed tools,
@@
 \begin{keyword}
-advanced normalization tools \sep bias correction \sep cortical thickness \sep open source software \sep segmentation \sep skull stripping
+advanced normalization tools \sep cortical thickness \sep open science
 %% keywords here, in the form: keyword \sep keyword
@@
 
-\input{results}
+%\input{results}
 

### 2013-05-08 BSplineDiff/Results/bsyn.R

@@ -1,0 +1,38 @@
+library( ggplot2 )
+
+stats <- read.csv( "overlapMeasures3.csv", header = TRUE )
+
+# difference <- stats$bsyn - stats$syn
+# average <- 0.5 * ( stats$syn + stats$bsyn )
+#
+# dmean <- data.frame( x = -c( -Inf, Inf ), y = mean( difference ) )
+# pStd <- data.frame( x = -c( -Inf, Inf ), y = mean( difference ) + 1.96 * sd( difference ) )
+# mStd <- data.frame( x = -c( -Inf, Inf ), y = mean( difference ) - 1.96 * sd( difference ) )
+#
+# plotData <- data.frame( cbind( SyN = stats$syn, BSyN = stats$bsyn, Difference = difference, Average = average  ) )
+#
+# myPlot <- ggplot( plotData, aes( x = Average, y = Difference ) ) +
+#           geom_line( aes( x, y ), linetype = 1, dmean ) +
+#           geom_line( aes( x, y ), linetype = 2, pStd ) +
+#           geom_line( aes( x, y ), linetype = 2, mStd ) +
+#           geom_point( aes( colour = Difference ), size = 3 ) +
+#           geom_point( colour = "grey90", size = 1 ) +
+#           scale_x_continuous( "Average Dice", breaks = seq( 0.6, 0.8, by = 0.05 ), labels = seq( 0.6, 0.8, by = 0.05 ), limits = c( 0.6, 0.8 ) ) +
+#           scale_y_continuous( "BSyN Dice - SyN Dice", breaks = seq( -0.025, 0.05, by = 0.025 ), labels = seq( -0.025, 0.05, by = 0.025 ), limits = c( -0.025, 0.05 ) ) +
+# #           theme( legend.justification = c( 0, 0 ), legend.position = c( 0, 0 ) ) +
+#           scale_shape( solid = FALSE ) +
+#           ggtitle( paste(  "SyN vs. BSyN", sep = "" ) )
+# ggsave( filename = paste( "syn.pdf", sep = "" ), plot = myPlot, width = 8, height = 5, units = 'in' )
+
+
+plotLabels <- c( rep.int( "SyN", length( stats$syn ) ), rep.int( "BSplineSyN", length( stats$bsyn ) ) )
+
+plotData <- data.frame( Type = plotLabels, Dice = c( stats$syn, stats$bsyn ) )
+
[…]

### 2013-05-08 SPIE2013/KK/Kirby_results/agePredict.R

@@ -27,7 +27,7 @@ gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "fem
 
 for( i in 1:32 )
   {
-  th <- results[,i+3];
+  th <- results[,i+4];
 
   gender[which( gender == 1 )] <- 'male';
   gender[which( gender == 2 )] <- 'female';

### 2013-05-08 SPIE2013/KK/NKI_results/agePredict.R

@@ -27,7 +27,7 @@ gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "fem
 
 for( i in 1:32 )
   {
-  th <- results[,i+3];
+  th <- results[,i+4];
 
   gender[which( gender == 1 )] <- 'male';
   gender[which( gender == 2 )] <- 'female';
@@ -41,7 +41,7 @@ for( i in 1:32 )
 #                geom_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x + I(x^2), method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fill = 'black', alpha = 0.5 ) +
                geom_point( data = plotData, aes( colour = Gender, shape = Gender ), size = 3 ) +
                scale_x_continuous( "Age (years)", breaks = seq( 20, 90, by = 10 ), labels = seq( 20, 90, by = 10 ), limits = c( 20, 90 ) ) +
-               scale_y_continuous( "Thickness (mm)", breaks = seq( 0, 7, by = 1 ), labels = seq( 0, 7, by = 1 ), limits = c( 0, 7 ) ) +
+               scale_y_continuous( "Thickness (mm)", breaks = seq( 0, 6, by = 1 ), labels = seq( 0, 6, by = 1 ), limits = c( 0, 6 ) ) +
                scale_colour_manual( values = c( "navyblue", "darkred" ), breaks = c( 1, 2 ), labels = c( "Male", "Female" ) ) +
                scale_shape_manual( values = c( 18, 16 ), breaks = c( 1, 2 ), labels = c( "Male", "Female" ) ) +
 #                scale_shape_discrete( breaks = c( 1, 2 ), labels = c( "Male", "Female" ) ) +

### 2013-05-11 Tilting/Rpaper.Rnw
ENH:  Amy
ENH: Silence.

--- 2013-05-05 Tilting/paper.tex
@@
 
+\usepackage{epigraph}
 \usepackage{color}
@@
     \definecolor{lightcyan}{rgb}{0.88,1,1}
-
-\newcommand{\lstsetcpplong}
-{
-\lstset{frame = tb,
-        framerule = 0.25pt,
-        float,
-        fontadjust,
-        backgroundcolor={\color{listlightgray}},
-        basicstyle = {\ttfamily\scriptsize},
-        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
-        identifierstyle = {\ttfamily},
-        commentstyle = {\ttfamily\color{listcomment}\textit},
-        stringstyle = {\ttfamily},
-        showstringspaces = false,
-        showtabs = false,
-        numbers = none,
-        numbersep = 6pt,
-        numberstyle={\ttfamily\color{listnumbers}},
-        tabsize = 2,
-        language=,
-        floatplacement=!h,
-        caption={\small \baselineskip 12pt DiReCT long command line menu which is invoked using the `{\ttfamily {-}{-}help}' option.
-The short command line menu is obtained by typing `{\ttfamily {-}h}'},
-        captionpos=b,
-        label=listing:long
-        }
-}
 
@@
 evaluations and propose a set of guidelines for authors and reviewers
-to minimize this confound.
+to minimize this confound.  We also illustrate these guidelines
+through the use of a recent real-world scenario.
 \end{abstract}
@@
 
-%\input{intro}
-%
-%\input{methods}
+\section{Introduction}
+The neuroscience community has significantly benefited from the proliferation
+of imaging software.  Established packages such as SPM \cite{Ashburner2012},
+the FMRIB Software Library (FSL) \cite{Jenkinson2012}, and
+the AFNI toolkit \cite{cox2012} have aided neuroimaging researchers 
+around the world in performing complex analyses as part of 
+ongoing neuroscience research.  Parallel to this line of
+inquiry is continued algorithmic innovation for improvement
+in analysis tools.
 
-\section{Introduction}
+As fellow scientists who actively participate in this 
+research through our many contributions to the Insight
+Toolkit and our own Advanced Normalization Tools (ANTs),%
+\footnote{
+http://www.picsl.upenn.edu/ANTs
+} 
+we have noticed several publications in which a principle
+component involves the comparison of algorithms.  One of our
+concerns is the lack of detail with which these comparisons are
+presented and the corresponding possibility of {\it instrumentation
+bias} \cite{sackett1979} (considering software  as a type of 
+instrument requiring proper ``calibration'' for accurate measurements).  
+Based on our experience (including mistakes that we have made), 
+we propose an initial set of guidelines for minimal
+reporting of algorithmic usage to minimize such bias, understanding that 
+the discussion will hopefully 
+elicit a much more comprehensive response from the larger community.
+
+Although previous articles have discussed similar concerns, oftentimes
+within a much larger context (e.g. fMRI reporting \cite{poldrack2008}), 
+we are particularly interested in comparative evaluations of software
+and provide additional discussion through experience and actual 
+examples found within the literature in addition to our own small,
+yet important, real world scenario.
+It is hoped that this commentary serves to raise awareness to both 
+authors and reviewers to such problematic issues (not unlike other
+recent articles detailing additional potential sources of methodological 
+bias \cite{kriegeskorte2009,vul2012,tustison2012}).
+
+\section{Guidelines}
+
+\subsection{Include parameters.}
+A best practice exemplar of comparative evaluation studies
+is the work by Klein et al. \cite{Klein2009} in which
+he compared 14 nonlinear registration algorithms using
+four publicly available labeled brain data sets.  
+As part of the study, the authors themselves tuned 
+the parameters for their algorithm to ensure good
+performance.  As
+part of this work, Prof. Klein availed the scripts
+used for the study on his website.%
+\footnote{
+http://mindboggle.info/papers/evaluation\_NeuroImage2009.php
+}
+This practice promotes reproducibility, fairness, and transparency in 
+the evaluation and is perhaps most fundamental in these types
+of studies for establishing conclusions regarding the relative
+performance of algorithms.
+
+In contrast, not specifying parameters 
+leaves one susceptible to criticisms of confirmation and/or 
+instrumentation bias \cite{sackett1979}.  For example, in 
+a recent paper \cite{haegelen2013} (similar issues can 
+be found in \cite{luo2013,wu2013}, the authors compared
+their ANIMAL registration algorithm with SyN \cite{avants2011} and determined
+that ``[r]egistration with ANIMAL was better than with SyN for 
+the left thalamus'' in a cohort of Parkinson's disease patients.
+One of the stated reasons for using ANIMAL was ``because of our
+extensive experience with this technique.''  However, for SyN,
+the authors provide no equivalent assurance---no contact
+was made with the primary developers of ANTs (SyN) nor
+were any parameters listed for independent verification 
+by reviewers or readers of the article.%
+\footnote{
+This is in addition to the authors' apparent neglect in  
+removing the possible confound of using two different linear 
+normalization strategies for the two algorithms.
+}
+Given the authors' relationship with one of the algorithms
+and likely predisposition towards its relative performance,
+particular caution has to be provided against possible confirmation
+bias.  One of the easiest ways is to provide the set of parameters
+used to run both algorithms so that others can ensure that 
+the parameters selected are sufficient to warrant the conclusions
+of the study.  Even better would be to include  the parameter
+files or command line calls used in the study.
+
+\subsection{Hassle the developers for a set of default parameters.}
+Dovetailing with the previous item, developers should also facilitate
+comparisons by providing default parameter files, sample scripts, 
+or possible command line calls.  For example, the developers of
+the image registration program Elastix \cite{klein2010} provide an
+assortment of parameter files on the wiki page%
+\footnote{
+  http://elastix.bigr.nl/wiki/index.php/Parameter\_file\_database
+}
+listed in tabular format complete with  short
+description (including applied modality and object of interest) and 
+any publications which used that specific parameter file.  Additionally,
+other packages are associated with discussion forums where users can
+query regarding usage.
+
+\subsection{Do not code up your own version of the algorithm particularly 
+if one is available from the original authors.  If you do, be sure to make
+your implementation available.}
+
+Collective experience in developing software has taught us the 
+difficulty of translating descriptions found in
+articles into functional code and that much is left unstated
+in presenting core algorithmic ideas.  Ideally, the authors,
+in conjunction with the manuscript, should provide an instantiation
+of the code.  Unfortunately, as observed in \cite{kovacevic2006}
+this is often not the case (even in terms of pseudocode).
+As a result, comparative evaluations are often carried out using 
+code developed not by the original authors but rather the 
+group doing the comparison.  This is not ideal as it questionable
+whether or not the authors have sufficient motivation to ensure that
+their instantiation of the algorithm is working correctly. For
+example, in \cite{clarkson2011}, the authors compared three algorithms
+for estimating cortical thickness.  Two of the algorithms were
+coded by the authors (even though one was publicly available from 
+the original authors) although not made public while the third, 
+Freesurfer \cite{fischl2012}, was used ``off the shelf'' having been extensively 
+tuned and well-documented by the original authors.  
+Thus, a natural question to ask is whether the performance difference
+is due to the algorithm itself or the implementation.
+
+\subsection{Provide details as to the source of the algorithm}
+As a corollary to the previous guideline, source of origin
+should be provided for any code or package used during the
+evaluation.  For example, N4 \cite{tustison2010} is a well-known 
+bias correction algorithm first made available as a tech report
+in the Insight Journal \cite{tustison2009a} for subsequent introduction
+into the Insight Toolkit for public dissemination.  However, since
+its inclusion, different instantiations have been made available.  
+Apart from the source code that accompanied the original tech
+report, N4 is also available in ANTs (the only version directly maintained
+by the original authors),
+it is also available as a module in Slicer,%
+\footnote{
+http://www.slicer.org/slicerWiki/index.php/Documentation/4.2/Modules/N4ITKBiasFieldCorrection
+}
+a wrapper of the slicer module in nipype,%
+\footnote{
+http://www.mit.edu/~satra/nipype-nightly/interfaces/generated/nipype.interfaces.slicer.filtering.n4itkbiasfieldcorrection.html
+}
+a module in \verb#c3d#,%
+\footnote{
+http://www.itksnap.org8/pmwiki/pmwiki.php?n=Convert3D.Documentation
+}
+and as a plugin in the BRAINS suite.%
+\footnote{
+http://www.nitrc.org/plugins/mwiki/index.php/brains:N4ITK
+}
+While each version is dependent on the original source
+code, there are subtle variations which can affect performance.
+For example, the parameters governing the spline distance and 
+number of iterations directly affect outcome and should be tuned
+based on application.  However, as a particular case, the \verb#c3d# 
+implementation hard codes
+these values with no access to modify them by the user potentially
+leading to suboptimal performance.
+
+\subsection{Provide computational platform details of the evaluation}
+An recent article \cite{gronenschild2012} pointed out significant
+differences in Freesurfer output that varied with Freesurfer release
+version and with operating systems.  While the former is to be
+expected given upgrades and big fixes which occur between releases,
+the latter underscores both the need for consistency in study processing
+as well as the reporting of computational details for reproducibility.
+
+\subsection{Pre- and post-processing steps need to be supplied.}
+In addition to disclosure of all parameters associated with the
+methods to be compared, all processing steps from the raw 
+to the final processed images in the workflow
+need to be specified.    Justification for any deviation
+of steps between algorithms needs to be provided.  Additionally,
+failures during processing should be mentioned and whether
+or not these data were excluded from subsequent analysis.
+
+\subsection{Comparisons should be performed on publicly available data.}`
+For reasons of reproducibility and transparency, evaluations should be performed
+using publicly availabile data sets.  Given the rather large number
+of such institutional efforts including 
+IXI,%
+\footnote{
+http://www.brain-development.org
+}
+NKI,%
+\footnote{
+http://fcon\_1000.projects.nitrc.org/indi/pro/nki.html
+}
+OASIS,%
+\footnote{
+http://www.oasis-brains.org
+}
+Kirby,%
+\footnote{
+http://mri.kennedykrieger.org/databases.html
+},
+LONI,%
+\footnote{
+http://www.loni.ucla.edu/Research/Databases/
+}, and
+others, evaluations should include (if not be exhausted by)
+comparisons using such data.  While evaluation on specific abnormal cohorts
+might exclude such possibilities, such specific evaluations should
+be extensively motivated in the introduction and/or discussion.
+For example, if a particular algorithm with general application is 
+found to perform better
+on a private cohort of Parkinson's disease subject data, reasons
+for performance disparity should be offered and supplemented with
+analysis on public data.
+
+\subsection{Resulting images should be publicly posted}
+
+The current publishing paradigm limits the quantity of
+results that can be posted.  There are only so many pages
+allowed for a particular publication and displaying every
+slice of every processed image, for example, is not feasible.  
+This results in possible selection bias where results provided
+in the manuscript are selected by the authors for demonstrating 
+the effect postulated at the onset of the study.  Thus, differences
+in performance assessment tend to be exaggerated based strictly
+on visual representations in the paper.  Resources for data
+sharing include MIDAS,%
+\footnote{
+http://www.midasplatform.org
+}
+XNAT,%
+\footnote{
+http://xnat.org
+}
+and slicedrop.%
+\footnote{
+http://slicedrop.com
+}
+
+\subsection{Properly contextualize comparisons and motivate performance differences.}
+In addition to algorithmic and study specifics, it is important to 
+discuss potential limitations concerning qualitative and/or quantitative 
+assessment metrics.  In \cite{rohlfing2012}, Rohlfing pointed out 
+deficiencies in using standard overlap measures and image similarity
+metrics in quantifying performance.  Other issues, such as biological
+plausibility of the resulting transforms, need to also be considered.
+
+Also important for inclusion is discussion on the possible reasons for performance
+disparity.  If one algorithm outperforms another, reporting of those findings 
+would be much more significant if the authors discussed possible reasons
+for relative  performance levels.  This simply constitutes evidence for
+the claims of the paper.
+
+\subsection{Co-authors should verify findings.}
+
+Although different journals have varying guidelines for determining 
+co-authorship, there is at least an implied sense of responsibility for
+an article's contents assumed by each of the co-authors.  Strategies 
+taken by journal editorial boards are being used to reduce undeserving
+authorship attribution such as requiring the listing of the specific 
+contributions of each co-author.  Additional proposals have included
+signed statements of responsibility for the contents of the article 
+\cite{nature2007}.  We suggest that at least one co-author
+independently verify a subset of the results by running the processing
+on their own computer.
+
+\section{Evaluation Example:  ANTS vs. antsRegistration}
+
+\subsection{Introduction}
+As part of the recent Medical Image Computing and Computer-Aided Intervention 
+(MICCAI) 2012 international meeting, several groups participated in the 
+Multi-Atlas Labeling challenge%
+\footnote{
+https://masi.vuse.vanderbilt.edu/workshop2012
+}
+where participants proposed algorithmic workflows to propagate
+labels from expertly annotated brains to testing data.  One of 
+the overall conclusions drawn from the meeting was the significance 
+of spatial normalization and so the organizers (Profs. Bennet Landman and Simon
+Warfield) suggested rerunning the competition using a set
+of transforms derived from a single registration algorithm.
+Since three of the top four performing algorithms all used
+the \verb#ANTS# registration tool \cite{avants2011} (available
+in the Advanced Normalization Tools open source 
+toolkit%
+\footnote{
+http://picsl.upenn.edu/ANTS
+}) the organizers approached one of the principle authors of
+the ANTs package (B.A.) to provide the ``canonical'' 
+\verb#ANTS#-derived transforms.  
+However, as part of the recent ITKv4 effort, an improved
+version for image registration, known as \verb#antsRegistration#
+had been created and also released within ANTs \cite{avants2012}.
+Motivated by this request and also to provide an example of the 
+guidelines previously presented, we performed a comparison of 
+ \verb#ANTS# and \verb#antsRegistration#.  The principle hypothesis
+ to be explored with relevance to the challenge is:  {\it Does
+ \verb#antsRegistration# outperform \verb#ANTS# in propagating
+ labels for the MICCAI challenge training data?}  Important 
+ corrollary hypotheses involve similar assessments of both the
+ affine and deformable (i.e. SyN) components of the respective
+ algorithms.
+
+\subsection{Materials and Methods}
+
+\subsubsection{Images}
+15 (10 females, 5 males) defaced T1-weighted MRI subjects 
+were sampled from the 
+publicly available 
+Open Access Series of Imaging Studies (OASIS) 
+project%
+\footnote{
+  http://www.oasis-brains.org
+} 
+which formed the ``training set'' for the challenge.  
+Ages ranged between 19 and 34 (23 $\pm$ 4.3) years.
+The complete
+list of sampled subjects can be found in the challenge proceedings.
+The corresponding labels
+were provided by Neuromorphometrics, Inc.%
+\footnote{
+http://neuromorphemetrics.com/
+} under academic subscription (released under the Creative Commons 
+Atribution-Non Commercial license).
+
+\subsubsection{Registration Protocols}
+ANTs%
+\footnote{
+https://github.com/stnava/ANTs
+}
+is distributed using the github web-hosting service
+facilitating developer collaboration.  The version
+of the ANTs toolkit that was used for this study 
+was updated on Friday May 3, 2013.%
+\footnote{
+git tag: 2e907d081e21514d2edc736f8894d991f79cbab7 
+}   
+The computational platform for evaluation for both 
+\verb#ANTS# and \verb#antsRegistration# is the Linux 
+cluster managed by the University of Virginia for 
+Computational Science and Engineering%
+\footnote{
+http://www.uvacse.virginia.edu
+}
+running the Portable Batch System (PBS)%
+\footnote{
+http://www.pbsworks.com
+}
+for queue management.  Each ``job'' consisted of a single
+pairwise registration.  Since we performed pairwise registrations
+for all possible pairings, this consisted of $15\times14=210$
+registrations for each of the algorithms.  To send all 420 
+jobs to the cluster
+a perl script was created, \verb#runTiltingRegistrations.pl#,%
+\footnote{
+Note that a special github repository has been created for this
+manuscript which includes all scripts for the evaluation.
+}
+which created a shell script for each pairwise registration.
+Given a fixed and moving image pair denoted, respectively, as
+\verb#$fixed# and \verb#$moving# and the ANTs path (\verb#$ANTsPATH#), 
+the core command line calls
+for each algorithm is given in Listing \ref{listing:ANTS} 
+(\verb#ANTS#) and Listing \ref{listing:antsRegistration}
+(\verb#antsRegistration#).  Note that the parameters were
+constructed by two of the principle authors of ANTs 
+(B.A. and N.T.) to maximize performance for both algorithms.
+
+\lstset{frame = tb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{listlightgray}},
+        basicstyle = {\ttfamily\scriptsize},
+        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{listcomment}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false,
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=,
+        floatplacement=!h,
+        caption={\small \baselineskip 12pt ANTS command line call used for the evaluation}.,
+        captionpos=b,
+        label=listing:ANTS
+        }
+\begin{lstlisting}
+${ANTsPATH}/ImageMath 3 $fixedTruncated \
+  TruncateImageIntensity $fixed 0.01 0.99
+${ANTsPATH}/ImageMath 3 $movingTruncated \
+  TruncateImageIntensity $moving 0.01 0.99
+
+${ANTsPATH}/ANTS 3 -o ${outputPrefix} \
+  -m CC[${fixedTruncated},${movingTruncated},1,4] \
+  -t SyN[0.1] -r Gauss[3,0] -i 100x100x70x20 \
+  --use-Histogram-Matching
+\end{lstlisting}
+
+\lstset{frame = tb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{listlightgray}},
+        basicstyle = {\ttfamily\scriptsize},
+        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{listcomment}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false,
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=,
+        floatplacement=!h,
+        caption={\small \baselineskip 12pt antsRegistration command line call used for the evaluation}.,
+        captionpos=b,
+        label=listing:antsRegistration
+        }
+\begin{lstlisting}  
+${ANTsPATH}/antsRegistration -d 3 -o ${outputPrefix} \
+  -u 1 -w [0.01,0.99], -r [${fixed},${moving},1] \
+  -t Rigid[0.1] -c [1000x500x250x100,1e-8,10] \
+  -s 4x2x1x0 -f 8x4x2x1 \
+  -m MI[${fixed},${moving},1,32,Regular,0.25] \
+  -t Affine[0.1] -c [1000x500x250x100,1e-8,10] \
+  -s 4x2x1x0 -f 8x4x2x1 \
+  -m MI[${fixed},${moving},1,32,Regular,0.25] \
+  -t SyN[0.1,3,0] -c [100x100x70x20,1e-9,15] \
+  -s 3x2x1x0 -f 6x4x2x1 \
+  -m CC[${fixed},${moving},1,4]
+\end{lstlisting}
+
+Following registration, the output was used
+to warp the corresponding labels of the moving
+image to the fixed image using \verb#antsApplyTransforms#:
+
+\lstset{frame = tb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{listlightgray}},
+        basicstyle = {\ttfamily\scriptsize},
+        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{listcomment}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false,
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=,
+        floatplacement=!h,
+        caption={\small \baselineskip 12pt antsApplyTransforms command line call used to warp
+        the moving label images to the fixed image for subsequent label overlap assessment}.,
+        captionpos=b,
+        label=listing:antsApplyTransforms
+        }
+\begin{lstlisting}  
+${ANTsPATH}/antsApplyTransforms -d 3 \
+  -i $movingLabels -r $fixed \
+  -n NearestNeighbor \
+  -o ${outputPrefix}LabelsWarped.nii.gz \
+  -t $outputWarp -t $outputAffine
+\end{lstlisting}
+Label overlap assessment for each registration was quantified
+using the \verb#LabelOverlapMeasures# available through ANTs and 
+implemented in ITK\cite{tustison2009}.
+It should noted that only the affine transform files are used
+to assess affine algorithmic performance. In order to evaluate
+the SyN component in isolation from the affine portion, the 
+affine transforms from the better-performing algorithm (strictly
+in terms of affine performance) were used as input for the other
+algorithm which was then rerun for all data sets.
+
+\subsection{Results}
+
+\subsection{Discussion}
+
+\section{Conclusion}
+\epigraph{By honest I don't mean that you only tell what's true.
+But you make clear the entire situation.
+You make clear all the information that is required for somebody else who is intelligent to make up their mind.}{Professor Richard Feynman}
+
+\section{Acknowledgments}
 
@@
 
-\section{Notes/Thoughts}
-\begin{itemize}
-  \item Several kinds of bias have cropped up in the neuroimaging
-  community recently:
-  \begin{itemize}
-    \item{methodological bias}:  Kreikesgorte
-    \item{statistical bias}:  Vul2009, Tustison2012
-    \item  What about the dead salmon fmri study---multiple comparisons correction?
-    \item The single subject VBM study (assuming the single subject is representative of the population).
-    \item{overview}:  9 circles of scientific hell ---neuroskeptic %http://blogs.discovermagazine.com/neuroskeptic/?p=1205#.UYZxyJWO7Tx
-    \item Types of measurement bias: \cite{sackett1979}
-    \begin{itemize}
-      \item Instrument bias (code as a type of instrument which needs to be calibrated correctly).
-      \item Expectation bias (i.e. confirmation bias)
-    \end{itemize}  
-  \end{itemize}
-  We are concerned with the latter two.
-  \item How does one avoid confirmation bias?  If I'm comparing my
-  algorithm to somebody else's algorithm, I'm going to be naturally
-  predisposed to confirmatory evidence that my algorithm is better
-  while neglecting (most likely not with intent of doing so) evidence
-  that disconfirms my original belief (that my algorithm is better).
-  Richard Feynman quote (Cargo Cult Science):  
-  \begin{quote} 
-  The first principle [of science] is that you must not fool yourself—--and you are the easiest person to fool.
+%\section{Notes/Thoughts}
+%\begin{itemize}
+%  \item Several kinds of bias have cropped up in the neuroimaging
+%  community recently:
+%  \begin{itemize}
+%    \item{methodological bias}:  Kreikesgorte
+%    \item{statistical bias}:  Vul2009, Tustison2012
+%    \item  What about the dead salmon fmri study---multiple comparisons correction?
+%    \item The single subject VBM study (assuming the single subject is representative of the population).
+%    \item{overview}:  9 circles of scientific hell ---neuroskeptic %http://blogs.discovermagazine.com/neuroskeptic/?p=1205#.UYZxyJWO7Tx
+%    \item Types of measurement bias: \cite{sackett1979}
+%    \begin{itemize}
+%      \item Instrument bias (code as a type of instrument which needs to be calibrated correctly).
+%      \item Expectation bias (i.e. confirmation bias)
+%    \end{itemize}  
+%  \end{itemize}
+%  We are concerned with the latter two.
+%  \item How does one avoid confirmation bias?  If I'm comparing my
+%  algorithm to somebody else's algorithm, I'm going to be naturally
+%  predisposed to confirmatory evidence that my algorithm is better
+%  while neglecting (most likely not with intent of doing so) evidence
+%  that disconfirms my original belief (that my algorithm is better).
+%  Richard Feynman quote (Cargo Cult Science):  
+%  \begin{quote} 
+%  The first principle [of science] is that you must not fool yourself—--and you are the easiest person to fool.
 So you have to be very careful about that.
@@
 You just have to be honest in a conventional way after that.
-  \end{quote}
-  \item One particular aspect of the previous item is that it needs 
-  to be highlighted if somebody codes up the algorithm
-  to be compared themselves.  Writing bug-free code is difficult.  Look
-  at how many bug fixes are provided by the ITK community on a daily
-  basis.  
-  \item Also, is simply saying that one algorithm is better than 
-  another sufficient?  Should there be some theoretical explanation
-  for it.  For example, when I wrote my DMFFD image registration paper,
-  the comparison wasn't IRTK vs. DMFFD but rather FFD vs. DMFFD and
-  there were theoretical reasons why performance should be better
-  with the latter.  Also, I insisted to the reviewer on not using IRTK 
-  for comparison since there are also so many other performance-related
-  issues (type of interpolation, gradient step, metric, metric implementation, etc.) which would confound
-  comparisons.  Also, it probably helped that the theme had a more specific,
-  more verifiable focus, i.e. DMFFD produces a more efficient energy minimizer
-  since it acts as a preconditioner in the standard gradient descent
-  optimization.
-  \item Comparisons are best made with publicly available data and
-  results of the comparisons should be made available.
-  \item Operating system needs to be defined.  Freesurfer results
-  varied with MacOSx.
-  \item Ideally, the authors of the new algorithm would work
-  with the authors of the compared algorithm.  Given human
-  nature, this type of cooperation is not guaranteed.  However,
-  a minimum set of information is needed.
-  \item Arno Klein as an example of somebody who did it right.  Takes a
-  lot of work.
-  \item 
-  \item The source of the package software needs to be defined.  For 
-  example, N4 has been
-  instantiated in Slicer, ANTs, and c3d (also might be implemented
-  in one of Styner's NITRC projects).  However, each might use
-  different default parameters and have other tweaks which effects
-  performance.  In Vladimir Fonov's github repository containing 
-  various processing scripts for MINC, one can see from the history
-  how N4 was used but then the users switched back to N3MNI (due to 
-  performance issues?).  However, they used N4 out of the c3d package
-  which the original authors (N.T./B.A.) haven't touched in three years
-  so the parameters aren't optimal (shrink factors = 4, spline distance
-  = 100 with 3 levels: $100\times50\times50$).  Changing these parameters 
-  (which are crucial to performance) isn't
-  accessible to the user from c3d.  
-  \item Great resources for code-sharing
-  	\begin{itemize}
-		  \item github
-		  \item NITRC
-		  \item ITK
-		\end{itemize}
-\end{itemize}
+%  \end{quote}
+%  \item In ‘The Meaning of It All’ Feynman said, ``By honest I don’t mean that you only tell what’s true.
+But you make clear the entire situation.
+You make clear all the information that is required for somebody else who is intelligent to make up their mind.''
+%  \item One particular aspect of the previous item is that it needs 
+%  to be highlighted if somebody codes up the algorithm
+%  to be compared themselves.  Writing bug-free code is difficult.  Look
+%  at how many bug fixes are provided by the ITK community on a daily
+%  basis.  
+%  \item Also, is simply saying that one algorithm is better than 
+%  another sufficient?  Should there be some theoretical explanation
+%  for it.  For example, when I wrote my DMFFD image registration paper,
+%  the comparison wasn't IRTK vs. DMFFD but rather FFD vs. DMFFD and
+%  there were theoretical reasons why performance should be better
+%  with the latter.  Also, I insisted to the reviewer on not using IRTK 
+%  for comparison since there are also so many other performance-related
+%  issues (type of interpolation, gradient step, metric, metric implementation, etc.) which would confound
+%  comparisons.  Also, it probably helped that the theme had a more specific,
+%  more verifiable focus, i.e. DMFFD produces a more efficient energy minimizer
+%  since it acts as a preconditioner in the standard gradient descent
+%  optimization.
+%  \item Comparisons are best made with publicly available data and
+%  results of the comparisons should be made available.
+%  \item Operating system needs to be defined.  Freesurfer results
+%  varied with MacOSx.
+%  \item Ideally, the authors of the new algorithm would work
+%  with the authors of the compared algorithm.  Given human
+%  nature, this type of cooperation is not guaranteed.  However,
+%  a minimum set of information is needed.
+%  \item Arno Klein as an example of somebody who did it right.  Takes a
+%  lot of work.
+%  \item 
+%  \item The source of the package software needs to be defined.  For 
+%  example, N4 has been
+%  instantiated in Slicer, ANTs, and c3d (also might be implemented
+%  in one of Styner's NITRC projects).  However, each might use
+%  different default parameters and have other tweaks which effects
+%  performance.  In Vladimir Fonov's github repository containing 
+%  various processing scripts for MINC, one can see from the history
+%  how N4 was used but then the users switched back to N3MNI (due to 
+%  performance issues?).  However, they used N4 out of the c3d package
+%  which the original authors (N.T./B.A.) haven't touched in three years
+%  so the parameters aren't optimal (shrink factors = 4, spline distance
+%  = 100 with 3 levels: $100\times50\times50$).  Changing these parameters 
+%  (which are crucial to performance) isn't
+%  accessible to the user from c3d.  
+%
+% \item Collaborators should help verify findings by repeating a
+%(representative subset of a) study independently, from scratch,
+%preferably on another computer perhaps even with a different operating
+%system.  A scientist can do this him/herself as well.  Personal
+%experience suggests that this procedure will often uncover coding
+%errors sometimes simply by forcing one to reread code or scripts again
+%or clarify documentation.
+% 
+%  \item
+%    \textcolor{blue}%{\href{http://en.wikipedia.org/wiki/Accuracy_and_precision}{Accuracy and precision (link)}} 
+%        wiki link--- most studies ignore precision.  it
+%    is critical in longitudinal studies.   precision is impacted by
+%    parameter choices as well as computational architecture
+%    (Freesurfer paper)
+%
+%   \item parameters impact accuracy - e.g. clarkson's paper vs our
+%     recent tuning of DiReCT parameters w.r.t. expected thickness values.
+%
+%  \item  Biological plausibility is important---a more accurate method
+%    may produce less biologically plausible results.
+%    Cross-referencing with other fields is critical e.g. patterns of
+%    MRI atrophy should match pathology distribution.  There is an art
+%    to this, of course.
+%
+%  \item Prediction---an algorithm that yields a ``more significant''
+%    result is not always better.  might mean ``finds more effected
+%    areas'' or smaller p-values.  Testing prediction complements
+%    studies of significance and cumulative distribution functions.
+%
+%  \item Verification is supported by open science technology for
+%    sharing data and code.
+%
+%   \item Great resources for data-sharing:
+%  	\begin{itemize}
+%	   \item figshare
+%           \item MIDAS 
+%           \item slicedrop
+%           \item R 
+%        \end{itemize}
+%
+%  \item Great resources for code-sharing
+%  	\begin{itemize}
+%		  \item github
+%		  \item NITRC
+%		  \item ITK
+%		\end{itemize}
+%\end{itemize}
 

### 2013-05-11 SPIE2013/KK/Combined_results/agePredict.R

@@ -1,6 +1,13 @@
 library( ggplot2 )
 
-results <- read.csv( file = paste( "labelresults", ".csv", sep = '' ), header = TRUE );
+resultsX <- list()
+resultsX[[1]] <- read.csv( file = "../IXI_results/labelresults.csv", header = TRUE )
+resultsX[[2]] <- read.csv( file = "../Kirby_results/labelresults.csv", header = TRUE )
+resultsX[[3]] <- read.csv( file = "../NKI_results/labelresults.csv", header = TRUE )
+resultsX[[4]] <- read.csv( file = "../Oasis_results/labelresults.csv", header = TRUE )
+
+
+results <- rbind( resultsX[[1]], resultsX[[2]], resultsX[[3]], resultsX[[4]] )
 
 cortical_labels <- c( "L occipital", "R occipital",
                      "L cingulate", "R cingulate",
@@ -27,7 +34,7 @@ gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "fem
 
 for( i in 1:32 )
   {
-  th <- results[,i+3];
+  th <- results[,i+4];
 
   gender[which( gender == 1 )] <- 'male';
   gender[which( gender == 2 )] <- 'female';
@@ -39,9 +46,9 @@ for( i in 1:32 )
 #                stat_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x, method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fullrange = FALSE, fill = 'black', alpha = 0.5 ) +
                stat_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x + I(x^2), method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fullrange = TRUE, fill = 'black', alpha = 0.5 ) +
 #                geom_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x + I(x^2), method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fill = 'black', alpha = 0.5 ) +
-               geom_point( data = plotData, aes( colour = Gender, shape = Gender ), size = 3 ) +
+               geom_point( data = plotData, aes( colour = Gender, shape = Gender ), size = 3, alpha = 0.5 ) +
[…]

### 2013-05-11 SPIE2013/KK/Oasis_results/agePredict.R

@@ -41,7 +41,7 @@ for( i in 1:32 )
 #                geom_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x + I(x^2), method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fill = 'black', alpha = 0.5 ) +
                geom_point( data = plotData, aes( colour = Gender, shape = Gender ), size = 3 ) +
                scale_x_continuous( "Age (years)", breaks = seq( 20, 90, by = 10 ), labels = seq( 20, 90, by = 10 ), limits = c( 20, 90 ) ) +
-               scale_y_continuous( "Thickness (mm)", breaks = seq( 0, 7, by = 1 ), labels = seq( 0, 7, by = 1 ), limits = c( 0, 7 ) ) +
+               scale_y_continuous( "Thickness (mm)", breaks = seq( 0, 6, by = 1 ), labels = seq( 0, 6, by = 1 ), limits = c( 0, 6 ) ) +
                scale_colour_manual( values = c( "navyblue", "darkred" ), breaks = c( 1, 2 ), labels = c( "Male", "Female" ) ) +
                scale_shape_manual( values = c( 18, 16 ), breaks = c( 1, 2 ), labels = c( "Male", "Female" ) ) +
 #                scale_shape_discrete( breaks = c( 1, 2 ), labels = c( "Male", "Female" ) ) +

### 2013-05-14 Tilting/Rpaper.Rnw
WIP: adding note from satra
ENH:  Downpressor man.

@@
   }
-\author[label2]{Brian B.~Avants (combined first author, second author or senior author?)}
+\author[label2]{Brian B.~Avants}
 \address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
@@
 concerns is the lack of detail with which these comparisons are
-presented and the corresponding possibility of {\it instrumentation
+often presented and the corresponding possibility of {\it instrumentation
 bias} \cite{sackett1979} (considering software  as a type of 
@@
 
-\subsection{Comparisons should be performed on publicly available data.}`
+\subsection{Comparisons should be performed on publicly available data.}
 For reasons of reproducibility and transparency, evaluations should be performed
@@
 
-\subsection{Introduction}
+\subsection{ANTS vs. antsRegistration:  Introduction}
 As part of the recent Medical Image Computing and Computer-Aided Intervention 
@@
 \verb#ANTS#-derived transforms.  
-However, as part of the recent ITKv4 effort, an improved
-version for image registration, known as \verb#antsRegistration#
-had been created and also released within ANTs \cite{avants2012}.
+However, as part of the recent ITKv4 effort, a new
+program for image registration, known as \verb#antsRegistration#
+has been created and also released within ANTs \cite{avants2012}
+where it continues to be refined (in contrast to \verb#ANTS# which
+has been more or less stable for quite some time).
+
 Motivated by this request and also to provide an example of the 
@@
  affine and deformable (i.e. SyN) components of the respective
- algorithms.
+ programs.
 
-\subsection{Materials and Methods}
+\subsection{ANTS vs. antsRegistration:  Materials and Methods}
 
@@
         floatplacement=!h,
-        caption={\small \baselineskip 12pt ANTS command line call used for the evaluation}.,
+        caption={\small \baselineskip 12pt ANTS command line call used for the evaluation. The help menu can be invoked by the command `ANTS --help'.},
         captionpos=b,
@@
   -t SyN[0.1] -r Gauss[3,0] -i 100x100x70x20 \
+  --number-of-affine-iterations 0 \
   --use-Histogram-Matching
@@
         floatplacement=!h,
-        caption={\small \baselineskip 12pt antsRegistration command line call used for the evaluation}.,
+        caption={\small \baselineskip 12pt antsRegistration command line call used for the evaluation.  The help menu can be invoked by the command `antsRegistration --help'.},
         captionpos=b,
@@
 using the \verb#LabelOverlapMeasures# available through ANTs and 
-implemented in ITK\cite{tustison2009}.
+implemented in ITK \cite{tustison2009}.
 It should noted that only the affine transform files are used
@@
 
-\subsection{Results}
+\subsection{ANTS vs. antsRegistration:  Results}
 
-\subsection{Discussion}
+Upon completion of all pairwise registrations, the labels for the
+moving image were warped to the fixed image using \verb#antsApplyTransforms#
+as described in Listing \ref{listing:antsApplyTransforms} using 
+nearest neighbor interpolation.  This was performed 
+for both affine and affine + deformable scenarios.  The
+corresponding box plots are given in Figure \ref{fig:results}.
+
+\begin{center}
+\begin{figure}
+  \includegraphics[width=85mm]{tilting.pdf}
+  \caption{Combined Dice values for all labels of the evaluation between the two ANTs-based registration programs including affine-only transformations.
+\textcolor{red}{Perhaps we should also plot Dice but with the white matter and cerebellum labels removed since they're so large (could they be masking more noticeable differences?).}
+  }
+  \label{fig:results}
+\end{figure}
+\end{center}
+
+Pairwise Students t-tests were performed for each of the three
+comparisons with Bonferroni correction.
+For the affine-only Dice value comparison, \verb#antsRegistration# 
+only slightly output performed \verb#ANTS# ($q = 0.053$, mean of the 
+differences = 0.004).  For the comprehensive affine + SyN comparison
+\verb#antsRegistration# yielded higher Dice values ($q = 0.053$, 
+mean of the differences = 0.009).
+Due to the difference in affine-only outcome, the \verb#ANTS# 
+SyN deformable component was run again with initialization from
+the affine output of the \verb#antsRegistration# runs.  This
+latter set is denoted in Figure \ref{fig:results} as ``ANTS with
+other affine.''  This scenario also showed \verb#antsRegistration# SyN 
+producing slightly higher values than its ANTs counterpart 
+($q = 0.053$, mean of the differences = 0.006).
+
+%\begin{table}
+%\begin{tabular*}{0.45\textwidth}{@{\extracolsep{\fill}}lr}
+%  Comparison & q-value \\
+%  \toprule
+%  ANTS vs. antsRegistration (affine only) & 0.053 \\
+%  ANTS vs. antsRegistration (affine + SyN) & 2.7e-14 \\
+%  ANTS vs. antsRegistration (affine' + SyN) & 3.4e-13 \\
+%  \bottomrule
+%\end{tabular*}
+%\end{table}
+
+\subsection{ANTS vs. antsRegistration:  Discussion}
+Strictly in terms of the Dice overlap metric, \verb#antsRegistration# yields
+improved results over \verb#ANTS#.  Although the SyN components are theoretically
+the same, implementation with the newer program was constrained to conform
+with the rigorous ITK coding standards and the plug-in architecture of
+the generic registration framework resulting in additional possible
+configurations not in \verb#ANTS#. Additionally, the newer program benefitted
+from ITK's crowdsourcing of the contributed code (a great example of Linus's Law
+in action---"with enough eyeballs, all bugs are shallow").  In contrast to the
+SyN component, the affine components between the two programs differ substantially.
+The parameters governing the affine registration in \verb#ANTS# were specifically 
+tailored for brain registration and much of it is hard-coded internally within the
+code.  In contrast, with \verb#antsRegistration# we took a much more general approach
+by permitting the user to build up any number of linear ``registration stages'' 
+to produce a final generic affine transform.
+
+Other differences to be considered include the use of real types between the
+two programs and the use of multi-threading.  As regards the former, due to 
+the constraints of certain  internals of ITK,  \verb#antsRegistration#
+is currently implemented using \verb#double# instead of \verb#float#.  This
+results in a larger memory footprint in addition to timing between the two
+programs (single-threaded on a single core on the UVA cluster \verb#antsRegistration#
+can run 3-4 times as long as \verb#ANTS#).  However, multi-threading capabilities
+are much better developed in \verb#antsRegistration#.
 
 \section{Conclusion}
-\epigraph{By honest I don't mean that you only tell what's true.
-But you make clear the entire situation.
-You make clear all the information that is required for somebody else who is intelligent to make up their mind.}{Professor Richard Feynman}
+\epigraph{By honest I don't mean that you only tell what's true. But you make clear the 
+entire situation. You make clear all the information that is required for somebody else 
+who is intelligent to make up their mind.}{Professor Richard Feynman}
 
-\section{Acknowledgments}
+Considering the permeation of the computational sciences in doing 
+nueroimaging research, certain safeguards should be in place to prevent
+(or at least minimize) potential biases and errors that can
+unknowingly effect study outcomes.  There is no vetting agency for
+ensuring that analysis programs used for research are reasonably error-free.
+In addition, these software packages are simply ``black
+boxes'' to many researchers who are not formally trained to debug
+code.  And even to those of us who are, debugging other people's code
+is oftentimes very difficult.  To this end, algorithmic comparisons are 
+a very good way of evaluating general performance.  It is hoped that the
+proposed set of guidelines set forth in this editorial aid the community
+for future comparative assessments.
 

### 2013-05-14 Tilting/Rpaper.Rnw
WIP: reorganize slightly

@@
 \section{Guidelines}
+A solid evaluation is essential to a paper's longevity and
+impact.  This requires effort in both planning the evaluation (before any
+experiment is performed) and in publishing the evaluation.
+We suggest general guidelines for both of these steps most of which
+derive from basic scientific principles of clarity and reproducibility.
 
-\subsection{Include parameters.}
+\subsection{Designing the Evaluation Study}
+
+\subsubsection{Define the Algorithm's Context}
+The only algorithm that is capable of outperforming all others on all
+problems is exhaustive search.  Therefore, many algorithms employ
+prior knowledge such as regularization or other heuristics that
+restrict the search space.  The very idea that one, for instance,
+registration algorithm should be better than all other algorithms on
+all types of data is fundamentally flawed.  Indeed, the No Free Lunch
+Theorems provide bounds on solution quality that show otherwise
+i.e. specifically that ``improvement of performance in problem-solving hinges on using prior information to match procedures to problems.''  One should keep this in
+mind when following the next guideline.
+
+\subsubsection{Hassle the developers for a set of default parameters.}
+Dovetailing with the previous item, developers should also facilitate
+comparisons by providing default parameter files, sample scripts, 
+or possible command line calls.  For example, the developers of
+the image registration program Elastix \cite{klein2010} provide an
+assortment of parameter files on the wiki page%
+\footnote{
+  http://elastix.bigr.nl/wiki/index.php/Parameter\_file\_database
+}
+listed in tabular format complete with  short
+description (including applied modality and object of interest) and 
+any publications which used that specific parameter file.  Additionally,
+other packages are associated with discussion forums where users can
+query regarding usage.
+
+\subsubsection{Do not code up your own version of the algorithm particularly 
+if one is available from the original authors.  If you do, be sure to make
+your implementation available.}
+
+Collective experience in developing software has taught us the 
+difficulty of translating descriptions found in
+articles into functional code and that much is left unstated
+in presenting core algorithmic ideas.  Ideally, the authors,
+in conjunction with the manuscript, should provide an instantiation
+of the code.  Unfortunately, as observed in \cite{kovacevic2006}
+this is often not the case (even in terms of pseudocode).
+As a result, comparative evaluations are often carried out using 
+code developed not by the original authors but rather the 
+group doing the comparison.  This is not ideal as it questionable
+whether or not the authors have sufficient motivation to ensure that
+their instantiation of the algorithm is working correctly. For
+example, in \cite{clarkson2011}, the authors compared three algorithms
+for estimating cortical thickness.  Two of the algorithms were
+coded by the authors (even though one was publicly available from 
+the original authors) although not made public while the third, 
+Freesurfer \cite{fischl2012}, was used ``off the shelf'' having been extensively 
+tuned and well-documented by the original authors.  
+Thus, a natural question to ask is whether the performance difference
+is due to the algorithm itself or the implementation.
+
+\subsubsection{Co-authors should verify findings.}
+
+Although different journals have varying guidelines for determining 
+co-authorship, there is at least an implied sense of responsibility for
+an article's contents assumed by each of the co-authors.  Strategies 
+taken by journal editorial boards are being used to reduce undeserving
+authorship attribution such as requiring the listing of the specific 
+contributions of each co-author.  Additional proposals have included
+signed statements of responsibility for the contents of the article 
+\cite{nature2007}.  We suggest that at least one co-author
+independently verify a subset of the results by running the processing
+on their own computer.
+
+\subsubsection{Comparisons should be performed on publicly available data.}
+For reasons of reproducibility and transparency, evaluations should be performed
+using publicly availabile data sets.  Given the rather large number
+of such institutional efforts including 
+IXI,%
+\footnote{
+http://www.brain-development.org
+}
+NKI,%
+\footnote{
+http://fcon\_1000.projects.nitrc.org/indi/pro/nki.html
+}
+OASIS,%
+\footnote{
+http://www.oasis-brains.org
+}
+Kirby,%
+\footnote{
+http://mri.kennedykrieger.org/databases.html
+},
+LONI,%
+\footnote{
+http://www.loni.ucla.edu/Research/Databases/
+}, and
+others, evaluations should include (if not be exhausted by)
+comparisons using such data.  While evaluation on specific abnormal cohorts
+might exclude such possibilities, such specific evaluations should
+be extensively motivated in the introduction and/or discussion.
+For example, if a particular algorithm with general application is 
+found to perform better
+on a private cohort of Parkinson's disease subject data, reasons
+for performance disparity should be offered and supplemented with
+analysis on public data.
+
+\subsection{Publishing the Evaluation}
+
+\subsubsection{Include parameters.}
 A best practice exemplar of comparative evaluation studies
@@
 
-\subsection{Hassle the developers for a set of default parameters.}
-Dovetailing with the previous item, developers should also facilitate
-comparisons by providing default parameter files, sample scripts, 
-or possible command line calls.  For example, the developers of
-the image registration program Elastix \cite{klein2010} provide an
-assortment of parameter files on the wiki page%
-\footnote{
-  http://elastix.bigr.nl/wiki/index.php/Parameter\_file\_database
-}
-listed in tabular format complete with  short
-description (including applied modality and object of interest) and 
-any publications which used that specific parameter file.  Additionally,
-other packages are associated with discussion forums where users can
-query regarding usage.
-
-\subsection{Do not code up your own version of the algorithm particularly 
-if one is available from the original authors.  If you do, be sure to make
-your implementation available.}
-
-Collective experience in developing software has taught us the 
-difficulty of translating descriptions found in
-articles into functional code and that much is left unstated
-in presenting core algorithmic ideas.  Ideally, the authors,
-in conjunction with the manuscript, should provide an instantiation
-of the code.  Unfortunately, as observed in \cite{kovacevic2006}
-this is often not the case (even in terms of pseudocode).
-As a result, comparative evaluations are often carried out using 
-code developed not by the original authors but rather the 
-group doing the comparison.  This is not ideal as it questionable
-whether or not the authors have sufficient motivation to ensure that
-their instantiation of the algorithm is working correctly. For
-example, in \cite{clarkson2011}, the authors compared three algorithms
-for estimating cortical thickness.  Two of the algorithms were
-coded by the authors (even though one was publicly available from 
-the original authors) although not made public while the third, 
-Freesurfer \cite{fischl2012}, was used ``off the shelf'' having been extensively 
-tuned and well-documented by the original authors.  
-Thus, a natural question to ask is whether the performance difference
-is due to the algorithm itself or the implementation.
-
-\subsection{Provide details as to the source of the algorithm}
+\subsubsection{Provide details as to the source of the algorithm}
 As a corollary to the previous guideline, source of origin
@@
 
-\subsection{Provide computational platform details of the evaluation}
+\subsubsection{Provide computational platform details of the evaluation}
 An recent article \cite{gronenschild2012} pointed out significant
@@
 
-\subsection{Pre- and post-processing steps need to be supplied.}
+\subsubsection{Pre- and post-processing steps need to be supplied.}
 In addition to disclosure of all parameters associated with the
@@
 
-\subsection{Comparisons should be performed on publicly available data.}
-For reasons of reproducibility and transparency, evaluations should be performed
-using publicly availabile data sets.  Given the rather large number
-of such institutional efforts including 
-IXI,%
-\footnote{
-http://www.brain-development.org
-}
-NKI,%
-\footnote{
-http://fcon\_1000.projects.nitrc.org/indi/pro/nki.html
-}
-OASIS,%
-\footnote{
-http://www.oasis-brains.org
-}
-Kirby,%
-\footnote{
-http://mri.kennedykrieger.org/databases.html
-},
-LONI,%
-\footnote{
-http://www.loni.ucla.edu/Research/Databases/
-}, and
-others, evaluations should include (if not be exhausted by)
-comparisons using such data.  While evaluation on specific abnormal cohorts
-might exclude such possibilities, such specific evaluations should
-be extensively motivated in the introduction and/or discussion.
-For example, if a particular algorithm with general application is 
-found to perform better
-on a private cohort of Parkinson's disease subject data, reasons
-for performance disparity should be offered and supplemented with
-analysis on public data.
-
-\subsection{Resulting images should be publicly posted}
+\subsubsection{Resulting images should be publicly posted}
 
@@
 
-\subsection{Properly contextualize comparisons and motivate performance differences.}
+figshare
+
+\subsubsection{Properly contextualize comparisons and motivate performance differences.}
 In addition to algorithmic and study specifics, it is important to 
@@
 
-\subsection{Co-authors should verify findings.}
+\subsection{Post-Publication: Publications Age}
 
-Although different journals have varying guidelines for determining 
-co-authorship, there is at least an implied sense of responsibility for
-an article's contents assumed by each of the co-authors.  Strategies 
-taken by journal editorial boards are being used to reduce undeserving
-authorship attribution such as requiring the listing of the specific 
-contributions of each co-author.  Additional proposals have included
-signed statements of responsibility for the contents of the article 
-\cite{nature2007}.  We suggest that at least one co-author
-independently verify a subset of the results by running the processing
-on their own computer.
+quoting satra ``a published evaluation is old the day the evaluation was done, as most software keeps changing.
+for example, arno's results were fine at that time with those versions and parameterization of the software, but might be a different story currently.''
 

### 2013-05-14 Tilting/Rpaper.Rnw
WIP: small comments on why co-authors should run your code

@@
 \section{Introduction}
-The neuroscience community has significantly benefited from the proliferation
+The neuroscience community has significantly benefitted from the proliferation
 of imaging software.  Established packages such as SPM \cite{Ashburner2012},
@@
 independently verify a subset of the results by running the processing
-on their own computer.
+on their own computer.   The point of this excercise is to verify not only
+reproducibility but also that the process can be explained in
+sufficient detail.
 
@@
 version and with operating systems.  While the former is to be
-expected given upgrades and big fixes which occur between releases,
+expected given upgrades and bug fixes which occur between releases,
 the latter underscores both the need for consistency in study processing

### 2013-05-15 Tilting/tilting.R
ENH:  ToniBraxton.

@@ -1,26 +1,0 @@
-library( ggplot2 )
-
-stats <- read.csv( "overlapMeasuresTilting.csv", header = TRUE )
-
-plotLabels <- c(
-                rep.int( "ANTS", length( stats$ANTS ) ),
-                rep.int( "ANTS with other affine", length( stats$ANTSWithOtherAffine ) ),
-                rep.int( "antsRegistration", length( stats$antsRegistration ) ),
-                rep.int( "Affine ANTS", length( stats$ANTSAffineOnly ) ),
-                rep.int( "Affine antsRegistration", length( stats$antsRegistrationAffineOnly ) )
-                )
-
-dice <- c( stats$ANTS, stats$ANTSWithOtherAffine, stats$antsRegistration, stats$ANTSAffineOnly, stats$antsRegistrationAffineOnly )
-
-plotData <- data.frame( Type = plotLabels, Dice = dice )
-
-myPlot <- ggplot( plotData,  aes( x = factor( Type ), y = Dice ) ) +
-          geom_boxplot( aes( fill = Type ) ) +
-          ggtitle( "MICCAI 2012 Multi-Label Atlas Challenge Data" ) +
-          theme( legend.position = "none" ) +
-          scale_x_discrete( "" )
-#           scale_x_discrete( breaks = c( "ANTS", "ANTS with other affine", "antsRegistration", "Affine ANTS", "Affine antsRegistration" ),
-#                             labels = c( "ANTS", "ANTS with other affine", "antsRegistration", "Affine ANTS", "Affine antsRegistration" ) )
-
-ggsave( filename = paste( "tilting.pdf", sep = "" ), plot = myPlot, width = 8, height = 5, units = 'in' )
-

### 2013-05-26 KapowskiChronicles/analytics/gender_study2.R
ENH:  Added comments/ggplots

@@ -1,0 +1,288 @@
+# load required libraries
+library( boot )
+library( ANTsR )
+library( ggplot2 )
+library( scales )
+library( igraph )
+library( pheatmap )
+library( randomForest )
+library( reshape )
+library( e1071 )
+
+sigma <- 3
+ages <- seq( 10, 80, by = 2 )
+
+#################################################################
+##
+##   function definitions
+##     * makeGraph - returns a vector of scores or weights for
+##                   each vertex in the network based on
+##     * calculateCorrelationMatrix -
+##
+#################################################################
+
+makeGraph <- function( myrsfnetworkcorrs )
+  {
+  correlationThreshold <- 0.000001
+
+  numberOfNeighbors <- nrow( myrsfnetworkcorrs )
+  if( numberOfNeighbors == 0 )
+    {
+    return( 0 )
+    }
+
+  edgeWeights <- c( 0 )
+  for( x in c( 1:numberOfNeighbors ) )
+    {
+    for( y in c( x:numberOfNeighbors ) )
+      {
+      if( myrsfnetworkcorrs[x,y] > correlationThreshold & myrsfnetworkcorrs[x,y] < 1 )
+        {
+        edgeWeights <- c( edgeWeights, myrsfnetworkcorrs[x,y] )
+        }
+      }
+    }
+
+  edgeWeights <- edgeWeights[2:length( edgeWeights )]
+  adjacencyMatrix <- as.matrix(
+    ( myrsfnetworkcorrs > correlationThreshold & myrsfnetworkcorrs < 1 ),
+    nrow = numberOfNeighbors, ncol = nnumberOfNeighbors )
+  g1 <- graph.adjacency( adjacencyMatrix, mode = c( "undirected" ) )
+
+#   gmetric <- betweenness( g1, normalized = T, weights = 1/edgeWeights )
+#   gmetric <- closeness( g1, normalized = T, weights = 1/edgeWeights )
+#   gmetric <- page.rank( g1, weights = 1/edgeWeights )$vector
+#   gmetric <- degree( g1 )
+
+  gmetric <- transitivity( g1, type = "local", isolates = c( "zero" ) )
+  return( gmetric )
+  }
+
+calculateCorrelationMatrix <- function( mat, weights, nuis )
+  {
+  if( missing( mat ) | missing( weights ) )
+    {
+    print( args( calculateCorrelationMatrix ) )
+    return( 1 )
+    }
[…]

### 2013-05-26 KapowskiChronicles/analytics/gender_study2.R
ENH:  Removed unnecessary items.

@@ -115,12 +115,10 @@ thicknessColumns <- 6:ncol( resultsCombined )
 weightedAges <- rep( NA, length( ages ) )
 corrs <- rep( NA, numberOfAges )
 
-thicknessDifferenceMatrix <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 tstatisticMatrix <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 pvalueMatrix <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
-
-networkMales <- thicknessDifferenceMatrix
-networkFemales <- thicknessDifferenceMatrix
+networkMales <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
+networkFemales <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 
 count <- 1
 for( age in ages )
@@ -190,12 +188,6 @@ for( age in ages )
 
   for( corticalLabel in 1:length( corticalLabels ) )
     {
-    myth <- residuals( lm( thicknessValues[,corticalLabel] ~ resultsSubsetBasedOnAgeDifference$SITE ) )
-
-    thicknessDifferenceMatrix[corticalLabel,count] <-
-      sum( myth[ resultsSubsetBasedOnAgeDifference$SEX == 2] * cweights[resultsSubsetBasedOnAgeDifference$SEX == 2] )
-      - sum( myth[ resultsSubsetBasedOnAgeDifference$SEX == 1] * cweights[resultsSubsetBasedOnAgeDifference$SEX == 1] )
-
     genderTest <- summary( lm( thicknessValues[,corticalLabel] ~ SEX + SITE + VOLUME, weights = cweights, data = resultsSubsetBasedOnAgeDifference ) )
 
     # get t-statistic and p-value on SEX significance (respectively)

### 2013-05-26 KapowskiChronicles/analytics/gender_study2.R
ENH:  changed to heat colormap.

@@ -211,8 +211,6 @@ for( age in ages )
 #
 ##################################
 
-jet.colors <- colorRampPalette( c( "#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000" ) )
-
 # (weighted) age vs. average thickness network plot
 
 corrsPlotData <- data.frame( weightedAges = weightedAges, correlationValues = corrs )
@@ -236,7 +234,7 @@ qvalueData$CorticalLabels <- factor( corticalLabels, levels = rev( corticalLabel
 
 qvaluePlot <- ggplot( melt( qvalueData ), aes( x = variable, y = CorticalLabels, fill = value ) ) +
               geom_tile( colour = "darkred" ) +
-              scale_fill_gradientn( name = "q-value", colours = jet.colors( 7 ) ) +
+              scale_fill_gradientn( name = "q-value", colours = heat.colors( 7 ) ) +
               scale_x_discrete( 'Age' ) +
               scale_y_discrete( 'Cortical Labels' )
 ggsave( filename = "qvalueHeatMap.pdf", plot = qvaluePlot, width = 10, height = 6, units = 'in' )
@@ -248,7 +246,7 @@ tstatisticData$CorticalLabels <- factor( corticalLabels, levels = rev( corticalL
 
 tstatisticPlot <- ggplot( melt( tstatisticData ), aes( x = variable, y = CorticalLabels, fill = value ) ) +
               geom_tile( colour = "darkred" ) +
-              scale_fill_gradientn( name = "t-statistic", colours = jet.colors( 7 ) ) +
+              scale_fill_gradientn( name = "t-statistic", colours = heat.colors( 7 ) ) +
               scale_x_discrete( 'Age' ) +
               scale_y_discrete( 'Cortical Labels' )
 ggsave( filename = "tstatisticHeatMap.pdf", plot = tstatisticPlot, width = 10, height = 6, units = 'in' )
@@ -260,7 +258,7 @@ networkMaleData$CorticalLabels <- factor( corticalLabels, levels = rev( cortical
 
 networkMalePlot <- ggplot( melt( networkMaleData ), aes( x = variable, y = CorticalLabels, fill = value ) ) +
                geom_tile( colour = "darkred" ) +
-               scale_fill_gradientn( name = "correlation\nvalues", colours = jet.colors( 7 ) ) +
+               scale_fill_gradientn( name = "correlation\nvalues", colours = heat.colors( 7 ) ) +
[…]

### 2013-05-28 KapowskiChronicles/methods.tex
ENH:  James C. Edwards

@@
   \centering
-  \includegraphics[width=90mm]{Figures/template50_60.pdf}
-  \caption{Sample multivariate template constructed from a subset of the IXI data (female, age 50--60).  Axial slices of five of the 37 total subjects from this cohort are shown. }
+  \includegraphics[width=90mm]{Figures/templates.pdf}
+  \caption{Population-specific templates for each of the four public data sets used 
+  for cortical thickness 
+  estimation generated using the script antsMultivariateTemplateConstruction.sh. 
+  The benefit of using such templates is obvious considering the variability in
+  acquisition and data preparation (e.g. defacing protocols).
+  }
   \label{fig:template}

### 2013-05-29 KapowskiChronicles/methods.tex
ENH:  Lil John

@@
   \centering
-  \includegraphics[width=170mm]{Figures/Kapowski_pipeline2.pdf}
+  \includegraphics[width=180mm]{Figures/Kapowski_pipeline2.pdf}
   \caption{Illustration of the main components of the ANTs processing 
   workflow containing all elements for determining cortical thickness. 
+  We also included the domain of operations for select scripts also available
+  in ANTs.
   Not shown is the optional subject to template registration.}
@@
 combined to create a brain extraction mask for use in the rest of the
-cortical thickness workflow.  \textcolor{blue}{does there need to be a
-  bit more technical detail here?  while you can refer to the script,
-  why perform these operations?  there are numerous references, most
-  germane probably some recent stuff from J Prince ( i think ) and the
-  freesurfer watershed approach which came from ... i forget ... maybe
-  one of the french groups.}
+cortical thickness workflow.
 
@@
 top level alongside BrainSuite (tuned) and FreeSurfer.
-\textcolor{blue}{ok you have the segonne ref here ... }
 
@@
 
-\paragraph{IXI} The IXI data%
+In total, we processed 1202 T1-weighted images from four different
+public data sets to obtain the corresponding cortical thickness maps.                           
+The four data sets are described below:
+
+\paragraph{IXI}
+The IXI data%
 \footnote{
@@
 
+\paragraph{Oasis}
+The Open Access Series of Imaging Studies (OASIS)%
+\footnote{
+http://www.oasis-brains.org/
+}
+ project consists of 416
+(160 males and 256 females) T1-weighted images spanning 
+the ages between 18 and 96 with 100 of these subjects having been 
+diagnosed with very mild to moderate Alzheimer's disease.  
+All subjects are right-handed.  Additional processing is included
+with the distribution including bias corrected and segmented images
+(gray, white and CSF matters).
+
+\paragraph{NKI}
+In support of open science, the 1,000 Functional Connectomes Project%
+\footnote{ 
+http://fcon\_1000.projects.nitrc.org
+}
+was initiated on December 11, 2009 by various members of the MRI community
+seeking to form collaborative partnerships with imaging institutions for 
+sharing well-documented multimodal image sets accompanied by phenotypic data.
+This has resulted in the The Nathan Klein Institute (NKI)/Rockland sample
+consisting of (at the time of download, September 22, 2012), 186 T1-weighted
+images (87 females and 99 males).
+
+\paragraph{Kirby}
+The Multi-Modal MRI Reproducibility Resource%
+\footnote{
+http://www.nitrc.org/projects/multimodal/
+}, 
+or more informally the Kirby
+data set, was originally described in \cite{landman2011} consisting of 
+21 subjects (10 female and 11 male) and featuring a rich multiple modality and 
+repeated acquisition schedule.
+
 \begin{figure}
@@
 \end{figure}
-
-\paragraph{Oasis}

### 2013-05-29 KapowskiChronicles/paper.tex
ENH:  Lil John

@@
 
+\section{Comments from Brian}
+\begin{itemize}
+  \item contributions
+  \begin{itemize}
+    \item establish population variability internationally ... reference thickness values .... i.e. look up table for thickness results ....
+    \item 7 sites 
+    \item report novel gender diff over age 
+    \item play up large - scale reliability  multisite 
+    \item largest N on public multisite data?  reference peele ...
+    \item quantify site effect .... is it low?
+    \item allows us to quantify gender differences over a large age range ....
+  \end{itemize}
+  \item misc.
+    \begin{itemize}
+      \item figures  
+      \item ask peele for feedback?
+      \item questions - can we get by on large N and multisite alone 
+      \item do we need to establish that we are more reliable than FSurf or FSL or SPM?
+      \item we set the bar high by doing X number of images
+      \item in discussion, say that future work will quantify SPM ....  not nec. by us. 
+    \end{itemize}
+\end{itemize}
+
 \input{intro}

### 2013-05-30 KapowskiChronicles/analytics/gender_study2.R
ENH:  Added thicknessplots.

@@ -219,7 +219,8 @@ corrsPlot <- ggplot( corrsPlotData, aes( x = weightedAges, y = correlationValues
              stat_smooth( colour = "navyblue", formula = y ~ 1 + x + I(x^2) + I(x^3) + I(x^4), method = "lm",
                           size = 1, n = 1000, level = 0.95, se = TRUE, fullrange = TRUE, fill = 'black', alpha = 0.25 ) +
              scale_x_continuous( "Age", breaks = seq( 10, 80, by = 10 ), labels = seq( 10, 80, by = 10 ), limits = c( 10, 80 ) ) +
-             scale_y_continuous( "Average thickness network correlation", breaks = seq( 0.7, 1, by = 0.05 ), labels = seq( 0.7, 1, by = 0.05 ), limits = c( 0.7, 1 ) )
+             scale_y_continuous( "Correlation", breaks = seq( 0.7, 1, by = 0.05 ), labels = seq( 0.7, 1, by = 0.05 ), limits = c( 0.685, 1 ) ) +
+             ggtitle( "Average thickness network correlation vs. age" )
 ggsave( filename = paste( "averageThicknessNetworkWithAge.pdf", sep = "" ), plot = corrsPlot, width = 8, height = 6, units = 'in' )
 
 # (weighted) age vs. average thickness network plot
@@ -244,8 +245,8 @@ tstatisticData <- data.frame( tstatisticMatrix )
 colnames( tstatisticData ) <- ages
 tstatisticData$CorticalLabels <- factor( corticalLabels, levels = rev( corticalLabels ) )
 
-tstatisticPlot <- ggplot( melt( tstatisticData ), aes( x = variable, y = CorticalLabels, fill = value ) ) +
-              geom_tile( colour = "darkred" ) +
+tstatisticPlot <- ggplot( melt( tstatisticData ) ) +
+              geom_tile( aes( x = variable, y = CorticalLabels, fill = value ), colour = "gray50", size = 0 ) +
               scale_fill_gradientn( name = "t-statistic", colours = heat.colors( 7 ) ) +
               scale_x_discrete( 'Age' ) +
               scale_y_discrete( 'Cortical Labels' )
@@ -256,8 +257,8 @@ networkMaleData <- data.frame( networkMales )
 colnames( networkMaleData ) <- ages
 networkMaleData$CorticalLabels <- factor( corticalLabels, levels = rev( corticalLabels ) )
 
[…]

### 2013-06-04 KapowskiChronicles/methods.tex
ENH:  Added captions.

@@
 Cortical label propagation to each individual subject for all data sets
-described below is performed using the PICSL multi-atlas joint label fusion
+described below was performed using the PICSL multi-atlas joint label fusion
 algorithm described in \cite{wang2013} which recently topped the competition
 at the MICCAI 2012 Grand Challenge on Multi-Atlas Labeling and is also 
-distributed with the ANTs toolkit.  .
+distributed with the ANTs toolkit.  Cortical thickness values are averaged
+within each label using only the non-zero voxels from the cortical thickness map.
 
@@
 
-In total, we processed 1202 T1-weighted images from four different
+In total, we processed 1205 T1-weighted images from four different
 public data sets to obtain the corresponding cortical thickness maps.                           
-The four data sets are described below:
+The four data sets are:
 
@@
 }
- used for the evaluation consists of 577 healthy subjects imaged at three sites 
-using several modalities (T1-weighted, T2-weighted, proton density, magnetic 
+consists of 561 healthy subjects (312 females, 249 males) imaged at three sites 
+with several modalities acquired (T1-weighted, T2-weighted, proton density, magnetic 
 resonance angiography, and diffusion tensor imaging).  The 
-database also consists of  demographic information such as age, weight,
+database also consists of  demographic information such as date of birth, date
+of scan, weight,
 height, ethnicity, occupation category, educational level, and marital status.
-The number of subjects spanning a range of demographic characteristics makes
-this a rich data set for validating and exploring correlations with cortical 
-thickness measured using the ANTs pipeline.
+Several subjects were removed during the analysis since certain
+demographic information was missing preventing an accurate estimate of 
+the age at the time of image acquisition.
 
@@
  project consists of 416
-(160 males and 256 females) T1-weighted images spanning 
+(160 males, 256 females) T1-weighted images spanning 
 the ages between 18 and 96 with 100 of these subjects having been 
@@
 sharing well-documented multimodal image sets accompanied by phenotypic data.
-This has resulted in the The Nathan Klein Institute (NKI)/Rockland sample
+This resulted in the Nathan Klein Institute (NKI)/Rockland sample
 consisting of (at the time of download, September 22, 2012), 186 T1-weighted
-images (87 females and 99 males).
+images (87 females, 99 males).
 
@@
 data set, was originally described in \cite{landman2011} consisting of 
-21 subjects (10 female and 11 male) and featuring a rich multiple modality and 
+21 subjects (10 females, 11 males) and featuring a rich multiple modality and 
 repeated acquisition schedule.

### 2013-06-04 KapowskiChronicles/analytics/gender_study2.R

@@ -170,7 +170,7 @@ for( age in ages )
     networkFemales[,count] <- makeGraph( subnet0$network )
 
     networkDifference <- abs( networkMales[,count] -  networkFemales[,count] )
-    if ( permutation == 0 | permutation == maximumNumberOfPermutations )
+    if( permutation == 0 | permutation == maximumNumberOfPermutations )
       {
       initialNetworkDifference <- networkDifference
       } else {

### 2013-06-05 KapowskiChronicles/intro.tex
ENH: adding stuff.

@@
 varies normally as a function of age \citep{kochunov2011},
-gender \citep{luders2006a}, untreated transsexuality \citep{zubiaurre-elorza2012},  handedness
+gender \citep{luders2006a}, untreated
+male-to-female transsexuality \cite{luders2012},  handedness
 \citep{luders2006,amunts2007}, intelligence \citep{shaw2006}, athletic

### 2013-06-07 KapowskiChronicles/conclusions.tex
ENH:  Adding conclusion section.

\section{Discussion and Conclusions}

Imaging biomarkers such as cortical thickness play an 
important role in neuroscience research.  In this work we 
detailed our open source strategy for estimating
cortical thickness directly from T1 images and demonstrated
its utility on a large collection of public brain data from
multiple databases acquired at multiple sites.  To our knowledge
this cohort constitutes the largest collection of cortical
thickness data processed in a single study.

In this work, we only explored a portion of the potentially
interesting investigations possible with these data.  However,
since all these data are publicly available, further work can
be easily pursued by us or even other interested groups.  We 
note that in addition to making the scripts, csv result files,
and cortical thickness maps available to researchers, 
posterior probability images are also offered permitting additional
quantification such as voxel-based morphometry \cite{ashburner2000}.

### 2013-06-13 KapowskiChronicles/analytics/gender_study2.R
ENH:  Adding Brian's changes.

@@ -9,8 +9,8 @@ library( randomForest )
 library( reshape )
 library( e1071 )
 
-sigma <- 3
-ages <- seq( 10, 80, by = 2 )
+sigma <- 5
+ages <- seq( 10, 80, by = 1 )
 
 #################################################################
 ##
@@ -119,7 +119,9 @@ tstatisticMatrix <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = num
 pvalueMatrix <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 networkMales <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 networkFemales <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
-
+networkAll <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
+rownames(networkAll)<-corticalLabels
+colnames(networkAll)<-ages
 count <- 1
 for( age in ages )
   {
@@ -145,7 +147,7 @@ for( age in ages )
   #
   ############################################
 
-  maximumNumberOfPermutations <- 5
+  maximumNumberOfPermutations <- 1
 
   initialNetworkDifference <- c();
 
@@ -159,6 +161,10 @@ for( age in ages )
       samplesSex <- sample( resultsSubsetBasedOnAgeDifference$SEX )
       }
 
+    temp <- calculateCorrelationMatrix( myth, cweights)
+    subnet0 <- reduceNetwork( temp, N = 200 )
+    networkAll[,count] <- makeGraph( subnet0$network )
+    
     males <- samplesSex == 1
     temp <- calculateCorrelationMatrix( myth[males,], cweights[males] )
     subnet0 <- reduceNetwork( temp, N = 200 )
@@ -170,6 +176,7 @@ for( age in ages )
     networkFemales[,count] <- makeGraph( subnet0$network )
 
     networkDifference <- abs( networkMales[,count] -  networkFemales[,count] )
+    
     if( permutation == 0 | permutation == maximumNumberOfPermutations )
       {
       initialNetworkDifference <- networkDifference
@@ -277,3 +284,15 @@ networkFemalePlot <- ggplot( melt( networkFemaleData ) ) +
                scale_y_discrete( 'Cortical Labels' ) +
[…]

### 2013-06-14 KapowskiChronicles/analytics/gender_study2.R
ENH:  Adding plot for network (both genders)

@@ -120,8 +120,10 @@ pvalueMatrix <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberO
 networkMales <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 networkFemales <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
 networkAll <- matrix( rep( NA, numberOfLabels * numberOfAges ), ncol = numberOfAges )
-rownames(networkAll)<-corticalLabels
-colnames(networkAll)<-ages
+
+rownames( networkAll ) <- corticalLabels
+colnames( networkAll ) <- ages
+
 count <- 1
 for( age in ages )
   {
@@ -164,7 +166,7 @@ for( age in ages )
     temp <- calculateCorrelationMatrix( myth, cweights)
     subnet0 <- reduceNetwork( temp, N = 200 )
     networkAll[,count] <- makeGraph( subnet0$network )
-    
+
     males <- samplesSex == 1
     temp <- calculateCorrelationMatrix( myth[males,], cweights[males] )
     subnet0 <- reduceNetwork( temp, N = 200 )
@@ -176,7 +178,7 @@ for( age in ages )
     networkFemales[,count] <- makeGraph( subnet0$network )
 
     networkDifference <- abs( networkMales[,count] -  networkFemales[,count] )
-    
+
     if( permutation == 0 | permutation == maximumNumberOfPermutations )
       {
       initialNetworkDifference <- networkDifference
@@ -243,7 +245,7 @@ qvalueData$CorticalLabels <- factor( corticalLabels, levels = rev( corticalLabel
 qvaluePlot <- ggplot( melt( qvalueData ), aes( x = variable, y = CorticalLabels, fill = value ) ) +
               geom_tile( colour = "darkred" ) +
               scale_fill_gradientn( name = "q-value", colours = heat.colors( 7 ) ) +
-              scale_x_discrete( 'Age' ) +
+              scale_x_discrete( 'Age', labels = seq( from = ages[1], to = ages[length( ages )], by = 5 ), breaks = seq( from = ages[1], to = ages[length( ages )], by = 5 ) ) +
               scale_y_discrete( 'Cortical Labels' )
 ggsave( filename = "qvalueHeatMap.pdf", plot = qvaluePlot, width = 10, height = 6, units = 'in' )
 
[…]

### 2013-06-15 ANTsRS/BRATS2013/paper.tex
McLovin

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithmic,eqparbox,array}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{multirow,booktabs,ctable,array}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.955}
    \definecolor{listwhite}{gray}{1.0}

%\floatname{algorithm}{Procedure}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\renewcommand{\algorithmiccomment}[1]{\hfill\eqparbox{COMMENT}{// #1}}
%
\begin{document}
%
\frontmatter          % for the preliminaries
%

\mainmatter              % start of the contributions
%
\title{ANTs and \'{A}rboles}
%\title{ANTs, \'{A}rboles, BRATS, and Brains}
%
\titlerunning{}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nicholas J. Tustison\inst{1} \and Max Wintermark\inst{1} \and Christopher Durst\inst{1} \and Brian B. Avants\inst{2}}

\authorrunning{N. J. Tustison, M. Wintermark, C. Durst, and B. B. Avants} % abbreviated author list (for running head)
\institute{University of Virginia, Charlottesville VA 22903, USA\\
\and
University of Pennsylvania,
Philadelphia PA  18104,USA
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Given the success of random forest approaches for segmentation, particularly for 
the BRATS 2012 tumor segmentation challenge, we implemented
a variant framework for our own research. 
The innovation of our methodology and implementation is characterized by 
the following four-fold contribution:
1) generation of
novel feature images in addition to what has been previously 
reported which significantly enhances classification,  2) 
concatenated application of random forest models for improved performance, 
3) the use of ANTsR (a packaging of the ANTs library plus 
additional analysis tools for
the R statistical project) for direct access to robust 
random forest functionality with parallelization, and 4) 
public availability of all scripts (coupled with the 
open source ANTs and R distributions) to recreate the 
leave-one-out evaluation study performed with the provided 
training data.
\keywords{ANTsR, Atropos, BRATS, N4, R, random forests, tumor segmentation}
\end{abstract}

\section{Introduction}

Adoption of the random forest framework \cite{breiman1996} for learning computer vision tasks (e.g. \cite{viola2005}) has led to recent use within 
the medical image analysis community for handling complex classification/regression tasks including normal brain segmentation \cite{yi2009},
MS lesion segmentation \cite{geremia2011}, multimodal brain tumor segmentation
\cite{zikic2012,geremia2012}, brain extraction \cite{iglesias2010}, 
anatomy detection in computed tomography \cite{criminisi2013}, and
segmentation of echocardiographic images \cite{verhoek2011}.
A thorough introduction delving deeper into the more theoretical aspects 
of random forests can be found in \cite{criminisi2011}.

The success of random forest-based approaches in the BRATS 2012 challenge, our 
own clinical research needs, and the lack of publicly available tools for such processing motivated the implementation that we describe herein.   Although we
borrowed many ideas from related previous work, our approach expands on this previous work in algorithmic and implementation terms both of which rely 
heavily on our Advanced Normalization Tools (ANTs)%
\footnote{
http://stnava.github.io/ANTs
} including its R packaging known as ANTsR.  
This includes concatenated application of random forest models (one based on Gaussian mixture modeling (GMM), similar to previous efforts, used as input to the succeeding one based on maximum a priori estimation and Markov random fields (MAP-MRF)).  We then refine the resulting labeling using a STAPLE-based approach \cite{warfield2004}.

We first provide an overview of the workflow followed by a description of the different multi-modal feature images used to construct the random forest models.  We then detail the implementation including descriptions of the various scripts used to perform the evaluation study with the provided training data.  Finally, we provide the resulting label overlap metrics for the leave-one-out study performed for each training data cohort.

\section{Methods}

The proposed workflow for estimating tumor-based labeling from multi-modal images involves the following steps:
\begin{enumerate}
  \item Symmetric template construction
  \item Multi-modal image preprocessing.
  \item Stage 1 processing:
  \begin{itemize}
    \item generation of feature images,
    \item construction of random forest model, and
    \item creation of random forest label probability images.
  \end{itemize}
  \item Stage 2 processing:
  \begin{itemize}
    \item generation of MAP-MRF specific feature images,
    \item construction of random forest model, and
    \item creation of random forest labelings.
  \end{itemize}
  \item Refinement of stage 2 labels.
\end{enumerate}

\subsection{Symmetric Template Construction}

In order to better characterize deviations from normal
multi-modal brain shape and appearance, several features were derived 
using population-specific multivariate symmetric templates.

Given $K$ image modality types for $N$ subjects,  
${\mathbf I} = \{I_1,I_2,\ldots, I_K\}$, multivariate 
template construction iterates between optimizing the set 
of diffeomorphic transforms between the subjects and the 
template, 
$\left\{\left(\phi_1,\phi_1^{-1}\right),\ldots,\left(\phi_N,\phi_N^{-1}\right)\right\}$ 
and constructing the 
optimal multivariate template appearance 
$\mathbf{J}=\{J_1,\ldots, J_K\}$ to minimize the
cost function
\begin{align}
  \sum_{n=1}^N 
        \left[ D \left( \psi(\mathbf{x}),\phi_1^n(\mathbf{x},1)\right)
        + \sum_{k=1}^K \lambda_k \Pi_k \left(I_k^n\left(\phi_n(\mathbf{x},0.5)\right),J_k\left(\phi^{-1}_n(\mathbf{x},0.5)\right)\right)\right]
\end{align}
where $D$ is the shape distance,
$D\left( \phi( \mathbf{x},0),\phi( \mathbf{x},1)\right) = \int_0^1 \| \nu(\mathbf{x},t)\|_L dt$
dependent on the choice of linear operator, $L$, and $\nu$
is the velocity field
$\nu\left( \phi(\mathbf{x},t) \right) = \frac{d\phi(\mathbf{x},t)}{dt},\,\,\, \phi(\mathbf{x},0) = \mathbf{x}$.
Each pairwise registration employing the similarity metric $\Pi_k$ can 
be assigned a relative weighting, $\lambda_k$, to weight a particular
modality's influence in the construction process.  Further theoretical
details can be found in \cite{avants2008,avants2010}.

We used the publicly available cohort associated with the multi-modal
reproducibility study produced by Landman et al. \cite{landman2011}
comprised of repeated acquisitions of several modalities for 21 normal 
individuals.  Shown in Figure \ref{fig:symmetrictemplates} are slices
from the resulting FLAIR, MPRAGE and T2 symmetric template components.

\begin{figure}
  \centering
  \begin{tabular}{ccc}
    \includegraphics[width=38mm]{Figures/S_templateFLAIR_140.png} &
    \includegraphics[width=38mm]{Figures/S_templateT1_140.png} &
    \includegraphics[width=38mm]{Figures/S_templateT2_140.png} \\
        (a) & (b) & (c)
  \end{tabular}
  \caption{Multivariate symmetric template created from the Kirby 
           data described in \cite{landman2011}.  Shown are the
           (a) FLAIR, (b) MPRAGE, 
           and (c) T2 template components.
          }
  \label{fig:symmetrictemplates}
\end{figure}

\subsection{Multi-Modal Image Preprocessing}

Although several studies have pointed out the importance of
intensity normalization and bias correction, our experience 
with the training data demonstrated a degradation in performance
when one or both steps (using \cite{nyul2000} and N4 \cite{tustison2010},
respectively) were performed.%
\footnote{
However, as we describe later, N4 is used in an interative scheme during the second
stage in calculating the MAP-MRF segmentation for each subject.  
} 
Therefore, for the first stage we simply windowed the image intensity
for all images to be between the quantiles $[0.01,0.99]$ and
subsequently rescaled the resulting intensity range to $[0,1]$.  
The set of feature images described in the next section are derived 
from these ``corrected'' images.

\subsection{Multi-Modal Feature Image Generation}

Key to any supervised regression or classification protocol are the 
selected features for training and subsequent testing.  Based on previous
work and our own experience, we selected the following feature images
for our supervised segmentation framework (cf Figure \ref{fig:featureImages}):
\begin{itemize}
  \item Per modality (FLAIR, T1, T1C, T2)
    \begin{itemize}
      \item First-order neighborhood statistical images:
            mean, variance, skewness, and entropy. 
            Neighborhood radius $\in \{1,3\}$.
    \item GMM (stage 1) and MAP-MRF (stage 2) posteriors: CSF, gray matter, white 
          matter, necrosis, edema, non-enhancing tumor and enhancing tumor (or just 
          CSF, gray matter, white 
          matter, edema, and tumor for the simulated cohorts).
    \item GMM (stage 1) and MAP-MRF (stage 2) connected component geometry 
          features:  distance to tumor core label, volume, volume to surface area ratio, eccentricity, and elongation
    \item Template-based:  symmetric template difference and 
          contralateral difference with Gaussian smoothing ($\sigma = 4$mm).
    \end{itemize}
  \item Miscellaneous
    \begin{itemize}
    \item Normalized Euclidean distance
    \item Log Jacobian image
    \item (T1C - T1) difference image
    \end{itemize}
\end{itemize}

\begin{figure}
  \centerline{
  \includegraphics[width=125mm]{Figures/featuresImages2.pdf}
  }
  \caption{Sample feature images from the BRATS\_HG0004 data set.}
  \label{fig:featureImages}
\end{figure}

For each modality, we create eight first-order statistical feature images,
seven (or five for simulated data) posterior probability feature images, five geometry features generated from the GMM posterior probability images
based on connected components, and two difference images using symmetric template
construction.  Additionally, we create one normalized distance image 
(the Euclidean distance image \cite{maurer2003} created from the skull-stripped 
binary mask rescaled to the range $[0,1]$), a log Jacobian image based on normalization to the symmetric template, and a difference image between the T1 and T1 contrast image for a total of 4 modalities $\times (8 + 7 + 5 + 2) + 3$ feature images per modality $=$ 91 total
feature images.

Prior  cluster centers for specific tissue types learned from training data \cite{reynolds2009} are used in the first stage to construction multiple 
GMM-based feature images.  One popular method for 
determining the parameters of the GMM is maximum likelihood 
estimation which can performed using the Atropos segmentation 
tool \cite{avants2011}.  In contrast to previous generative
modeling approaches for multi-modal tumor segmentation 
(e.g. \cite{prastawa2003,zikic2012}), we do not use multivariate 
Gaussians to specify tissue probabilities but rather incorporate each
univariate probability map into the feature vector of the training
data.  As pointed out in \cite{menze2010}, multivariate modeling
might obscure the distinct biological information provided by each 
modality.  Instead, we let the random forest construction 
process determine the optimal combination of such multivariate
information.

During the second stage, the probabilistic estimates
of the white matter and gray matter labels were used to generate a
``pure tissue weight mask'' to estimate the bias field 
using N4 (although the resulting bias field estimation was used
to correct the image within the entire cerebral mask).  Formally, this 
involved generation of a probabilistic map defined as:
\begin{align}
  P_{pure\,\,tissue}(\mathbf{x}) = \sum_{i=1}^N P_i(\mathbf{x}) \prod_{j=1, j \neq i}^N \left( 1 - P_j(\mathbf{x}) \right)
\end{align}
where $N$ is the set of user-selected tissue labels (in our
case $N=2$ consisting of the gray and white matter probability
maps).
Both rescaling and weighted bias correction were applied to produce
the ``corrected images'' for the second stage resulting in
modified features images for the second stage.  Note that we
perform a similar iterative scenario for normal brain 
segmentation \cite{avants2011} (encapsulated in the ANTs script 
\verb#antsAtroposN4.sh#).

%Additionally, maximum posterior labeling from the GMM 
%processing  is used to determine the connected components for each label.  
%Geometric features (assigned voxel-wise) include the physical volumes 
%of each connected component including the volume to surface area ratio, 
%the elongation, and eccentricity. 
%Similarly, at the second stage, the probability maps determined from
%application of the stage 1 random forest model are used as spatial 
%priors for to construct MAP-MRF-based feature images analogous to the 
%stage 1 GMM feature images.  This is also performed using the Atropos
%segmentation tool available in ANTs.

ANTs registration capabilities are also used to determine the transform from 
each subject to the symmetric template.  This transform provides three 
sets of feature images:  the log Jacobian determinant image (assuming that the 
presence of tumor causes abnormal displacements), voxelwise intensity differences between each modality of each subject and the corresponding symmetric template component, and voxelwise contralateral differences.  The latter is calculated by warping each modality to the symmetric template, flipping the image contralaterally 
and warping it back to the subject space, and then calculating the difference image
between the image and its contralateral counterpart.  This is very much influenced by earlier symmetric-based features described in \cite{geremia2011} although our approach accounts for a potentially deformed mid-sagittal plane caused by tumor growth.

\subsection{Implementation}

As mentioned previously, motivating this work were the limited public resources
for performing multi-modal tumor segmentation.  As a partial corrective, all the
software and scripts used to create the evaluation performed in the next section have been made publicly available.

The ANTsR package (which also contains ANTs) is publicly available on the github project hosting service.%
\footnote{
https://github.com/stnava/ANTsR
}
Prior to installation of ANTsR, several external R packages
need to be installed including: \verb#Rcpp#, \verb#signal#, \verb#timeSeries#, 
\verb#mFilter#, \verb#doParallel#, \verb#robust#, \verb#magic#, \verb#knitr#, \verb#pixmap#, 
\verb#rgl#, \verb#misc3d# which is facilitated by the 
\verb#install.packages()# mechanism.  Additionally, in order
to perform the supervised brain segmentation as described 
in this work, one would need to also install 
\verb#randomForest#, \verb#snowfall#, \verb#rlecuyer#, and \verb#ggplot2#. 
The multivariate template construction is done using \verb#antsMultivariateTemplateConstruction.sh#, available in the ANTs \verb#Scripts# directory which permits parallel processing on
an individual workstation or on a large computational cluster.

Additional scripts necessary to perform the evaluation can be found in the 
first author's github repository%
\footnote{
https://github.com/ntustison/Utilities
}
and are listed as follows: 
\begin{itemize}
  \item \verb#createTruthLabels.pl# -- performs a 3-tissue segmentation of the tumor training data.  These three labels are then combined with the given labels. 
  \item \verb#createNormalizedImagesForCohort.pl# -- windows and rescales the images (commented out are previous attempts at N4 bias correction and intensity normalization).
  \item \verb#createFeatureImagesForCohort.pl# -- calculates the features images by calling \verb#createFeatureImages.sh# for each subject.
  \item \verb#runLeaveOneOutCrossValidation.pl#  -- calls \verb#createRandomForestModel.pl# for each subject using the training 
  data from the other subjects.
  \item \verb#createRandomForestModel.pl# -- calls the R script \verb#createModel.R#.
  \item \verb#applyTumorSegmentationModelForCohort.pl# -- creates the random
  forest probability maps for each label using \verb#applyTumorSegmentationModel.sh#.
  \item \verb#refineTumorSegmentationResultsForCohort.pl# -- refines the final labels from the random forest model using STAPLE.
  \item \verb#createFeatureImages.sh#  -- creates features images for a specific
  subject.
  \item \verb#applyTumorSegmentationModel.sh# -- given a random forest model (.RData), produces the subject-specific random forest probability images.
  \item \verb#createModel.R# -- R script interface to the \verb#randomForest# R package.  Provides optional parallelization with the \verb#snowfall# package.
  \item \verb#applyModel.R# -- R script interface to the \verb#randomForest# R package.  Provides optional parallelization with the \verb#snowfall# package.
  \item \verb#createCSVFileFromModel.R# -- produces a csv file containing a data frame of feature images.
  \item \verb#plotVariableImportance.R# -- plots feature importance given a random forest model.
\end{itemize}
Although the perl scripts are specific to running jobs on the cluster at the 
University of Virginia, they can be easily adapted to run on one's own platform.

\section{Evaluation}

\begin{figure}
  \centerline{
  \includegraphics[width=150mm]{Figures/BRATS_HG0004GMM.pdf}
  }
  \caption{Relative variable importance plot of the different features used to construct the stage 1 random forest model for subject BRATS\_HG0004.}
  \label{fig:importance}
\end{figure}

A leave-one-out evaluation strategy was adopted for each of the four cohorts (high vs. low grade and real vs. simulated data).  The processing workflow described previously was applied to each subject whereby the remaining cohort members were used to train the random forest models which were then applied to that subject.  One of the benefits of random forests is that the training process can be used to produce a relative importance weighting of the individual features.  For example, we show the importance plot for the stage 1 model for subject BRATS\_HG0004 in Figure \ref{fig:importance}.

The metrics for performance assessment were given by the organizers and include 
combining labels to assess overlap of complete tumor (labels 1--4), tumor core (labels 1,3, and 4), and enhancing tumor (label 4).  The corresponding Dice metrics were calculated using open source software \cite{tustison2009} and are provided in Table \ref{table:brats}.

\begin{table*}
\caption{Dice scores from the MICCAI 2013 BRATs Competition}
\label{table:brats}
\centerline{
\begin{tabular*}{0.975\textwidth}{@{\extracolsep{\fill} } c c c c c c c c c}
\toprule
{} & \multicolumn{2}{c}{High-grade (real)} & \multicolumn{2}{c}{Low-grade (real)} & \multicolumn{2}{c}{High-grade (simulated)} & \multicolumn{2}{c}{Low-grade (simulated)}\\
{\bf Method} & Edema & Tumor & Edema & Tumor & Edema & Tumor & Edema & Tumor\\
\midrule
\cite{zikic2012} & {$0.70 \pm 0.09$} & {$0.71 \pm 0.24$} & {$0.44 \pm 0.18$} & {$0.62 \pm 0.27$} & {$0.65 \pm 0.27$} & {$0.90 \pm 0.05$} & {$0.55 \pm 0.23$} & {$0.71 \pm 0.20$} \\
\cite{bauer2012} & {$0.61 \pm 0.15$} & {$0.62 \pm 0.27$} & {$0.35 \pm 0.18$} & {$0.49 \pm 0.26$} & {$0.68 \pm 0.26$} & {$0.90 \pm 0.06$} & {$0.57 \pm 0.24$} & {$0.74 \pm 0.10$} \\
ANTsR & {$0.65 \pm 0.15$} & {$0.66 \pm 0.28$} & {$0.49 \pm 0.16$} & {$0.65 \pm 0.21$} & {$0.68 \pm 0.25$} & {$0.91 \pm 0.08$} & {$0.61 \pm 0.25$} & {$0.84 \pm 0.09$} \\
w/ Atropos & {$0.68 \pm 0.15$} & {$0.67 \pm 0.30$} & {$0.50 \pm 0.15$} & {$0.67 \pm 0.23$} & {$0.74 \pm 0.26$} & {$0.92 \pm 0.09$} & {$0.65 \pm 0.26$} & {$0.84\pm 0.08$} \\
\bottomrule
\end{tabular*}
}
\end{table*}

\section{Discussion and Conclusions}

Our experience matches with previous research demonstrating good segmentation performance using random forests.  We found that additional feature images to what has been proposed previously can significantly improve classification.  For example, the most discriminative feature image at the top of the plot in Figure \ref{fig:importance} is a GMM-derived image based on connected components.  Immediately following in importance is the FLAIR contralateral difference image generated from the symmetric template normalization.  Other novel contributions which appear to be significant are the normalized distance and log Jacobian images.  
Aside from feature image contributions, we provide our implementation as open source.

\bibliographystyle{splncs03}
\bibliography{references}

\end{document}

### 2013-06-15 ANTsRS/paper.tex
McLovin

@@
 \author[label1]{Christopher Durst}
+\author[label1]{Max Wintermark}
 \author[label1]{James C.~Gee}
 \author[label1]{Murray C.~Grossman}
-\author[label1]{Max Wintermark}
 \author[label2]{Brian B.~Avants}
@@
 then be extended to testing data, i.e. unlabeled 
-brain data.  In subsequent sections, we introduce the generic
+brain data.  In subsequent sections, we detail the generic
 {\it random forest} modeling framework which takes as input

### 2013-07-07 KapowskiChronicles/intro.tex
ENH:  Adding evaluation section.

@@
 connectivity relationships using cortical thickness measures
-\cite{worsley2005,lerch2006,he2007}.
-Although these findings
+\citep{worsley2005,lerch2006,he2007,chen2008}.
+Although all these findings
 are subject to debate and interpretation \citep{gernsbacher2007}, 
@@
 Equally as important, we provide explicit coordination between
-these pipeline components complete with a set of useful parameters
-which are employed to analyze the publicly available IXI data set. The
-full pipeline and parameter set is 
-encapsulated in a well-documented shell script which is also available in ANTs. 
+these pipeline components within a set of well-documented shell scripts which 
+is also available in the ANTs repository where parameters have been tuned
+by the original authors and provide good performance across a number of data sets.
 Furthermore, we provide all the derived image data and other processing scripts on 

### 2013-07-07 KapowskiChronicles/methods.tex
ENH:  Adding evaluation section.

@@
   \item brain extraction using a hybrid segmentation/template-based strategy,
-  \item alternation between prior-based segmentation and white matter posterior
-        probability weighted bias correction,
+  \item alternation between prior-based segmentation and ``pure tissue'' 
+        posterior probability weighted bias correction,
   \item DiReCT-based cortical thickness estimation, and
@@
 %}
-Additionally, the coordination of all the algorithmic components is
+The coordination of all the algorithmic components is
 encapsulated in the shell scripts \verb#antsCorticalThickness.sh# with
@@
 optimal parameters for each of the algorithmic components which has worked
-well for our processing and which are used to acquire the results 
+well for our processing and are used to acquire the results 
 described in this work.
@@
   \centering
-  \includegraphics[width=180mm]{Figures/Kapowski_pipeline2.pdf}
+  \includegraphics[width=140mm]{Figures/Kapowski_pipeline2.pdf}
   \caption{Illustration of the main components of the ANTs processing 
   workflow containing all elements for determining cortical thickness. 
-  We also included the domain of operations for select scripts also available
-  in ANTs.
+  We also included the domain of operations for the selected scripts.
   Not shown is the optional subject to template registration.}
@@
 The ANTs implementation of this technique is currently available as a shell script, 
-\verb#buildtemplateparallel.sh# (a multivariate version is also available as
+\verb#buildtemplateparallel.sh# (a generalized, multivariate version is also available as
 \verb#antsMultivariateTemplateConstruction.sh#).  Both scripts are distributed as part of
@@
 Due to the important interplay between segmentation and bias correction,
-we perform multiple N4 $\Leftrightarrow$ Atropos iterations.
+we perform multiple N4 $\rightleftharpoons$ Atropos iterations.
 In order to better integrate Atropos and N4, we use  
 a pure tissue probability weight mask generated from the 
-posterior probabilities generated from the segmentation 
+posterior probabilities derived from the segmentation 
 process.  Given $N$ labels and the corresponding $N$
@@
 during the segmentation process, the N4 weight mask is 
-created at each N4 $\Leftrightarrow$ Atropos iteration from
+created at each N4 $\rightleftharpoons$ Atropos iteration from
 \begin{align}
@@
 DiReCT was introduced 
-in \cite{das2009} and made available in ANTs with the program \verb#KellySlater#.
+in \cite{das2009} and made available in ANTs as the program \verb#KellySlater#.
 Since then several improvements have been made and incorporated into the program
@@
 \centering
-\begin{tabular*}{0.475\textwidth}{@{\extracolsep{\fill}} l l}
+\begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}} l l}
 \toprule
@@
 height, ethnicity, occupation category, educational level, and marital status.
-Several subjects were removed during the analysis since certain
+Several subjects were removed after cortical thickness map calculation during 
+the post-analysis 
+since certain
 demographic information was missing preventing an accurate estimate of 
@@
 with the distribution including bias corrected and segmented images
-(gray, white and CSF matters).
+(gray, white and CSF matters).  20 subjects had repeat scans which
+were used for evaluation of reproducibility.
 

### 2013-07-07 KapowskiChronicles/paper.tex
ENH:  Added thicknessplots.
ENH:  Adding conclusion section.
ENH:  Adding evaluation section.

--- 2013-05-08 KapowskiChronicles/paper.tex
@@
 
-%\documentclass[preprint,authoryear,review,12pt]{elsarticle}
-\documentclass[final,5p,times,twocolumn]{elsarticle}
+\documentclass[preprint,authoryear,review,12pt]{elsarticle}
+% \documentclass[final,5p,times,twocolumn]{elsarticle}
 
@@
 and DiReCT (cortical thickness) all developed as part of the ANTs open science effort.  
-Complementing the open source aspect of ANTs we investigate several hypotheses
-explored previously in the literature using four different publicly 
-available data sets totaling approximately 1200 subjects.  
+Four open data sets (IXI, Kirby, NKI, and Oasis), comprising approximately 
+1200 subjects, were processed using the proposed pipeline.  Reproducibility scores
+from a subset of these data, along with a complementary age prediction
+evaluation, demonstrate robust performance.  Additionally, we present novel
+gender differences in predicting structural connectivity across the lifespan.
 To further promote open science and use of the proposed tools,
@@
 
-%\input{results}
+\input{results}
 
-\section{Discussion and Conclusions}
+\input{conclusions}
 

### 2013-07-07 KapowskiChronicles/analytics/age_study.R

@@ -1,3 +1,4 @@
+library( ANTsR )
 
 #####################################################################################
 mysurf <- renderSurfaceFunction(list(template,brain), alphasurf=c(0.3,0.3), surfval=0.5, smoothsval=1.5,  alphafunc=1, mycol="cadetblue1")
@@ -25,7 +26,7 @@ for( permutation in 0:maximumNumberOfPermutations )
     {
       samplesAge <- sample( resultsCombined$AGE )
     }
-    # localtransitivity betweeness pagerank 
+    # localtransitivity betweeness pagerank
   gdensity <- 0.25
   ageDifference1 <- abs( samplesAge - testages[1] )
   resultsSubsetBasedOnAgeDifference1 <- subset( resultsCombined, ageDifference1 <= 10 )
@@ -74,11 +75,11 @@ for( permutation in 0:maximumNumberOfPermutations )
       dev.off()
       }
   setTxtProgressBar(pb, permutation )
-} # end permutation 
+} # end permutation
 pvals <-  permutationCount / maximumNumberOfPermutations
 pvageMat[,ntests]<-pvals
 cat( "Ages: ", testages, "\n", sep = ' ' );
-for ( ff in 1:length(mysign) ) if (  pvals[ff ] < 0.05 ) print(paste(corticalLabels[ff],initialNetworkDifference[ff],pvals[ff]))#  
+for ( ff in 1:length(mysign) ) if (  pvals[ff ] < 0.05 ) print(paste(corticalLabels[ff],initialNetworkDifference[ff],pvals[ff]))#
 }
 
 qq<-pvageMat

### 2013-07-07 KapowskiChronicles/analytics/brainAgeAnalysis.R

@@ -1,46 +1,46 @@
 library( kernlab )
+library( randomForest )
 library( ggplot2 )
 
-trainingData <- read.csv( "PCAVariation0.99/trainingBrainSegmentationPosteriors2Projections.csv" )
-testingData <- read.csv( "PCAVariation0.99/testingBrainSegmentationPosteriors2Projections.csv" )
+# trainingData <- read.csv( "PCAVariation0.99/trainingBrainSegmentationPosteriors2Projections.csv" )
+# testingData <- read.csv( "PCAVariation0.99/testingBrainSegmentationPosteriors2Projections.csv" )
 # trainingData <- read.csv( "PCAVariation0.99/trainingCorticalThicknessProjections.csv" )
 # testingData <- read.csv( "PCAVariation0.99/testingCorticalThicknessProjections.csv" )
+trainingData <- read.csv( "PCAVariation0.99/trainingTiledProjections.csv" )
+testingData <- read.csv( "PCAVariation0.99/testingTiledProjections.csv" )
 
 # Remove ID, gender, and site
-#  drops <- c( "ID", "SEX", "SITE" )
-#  trainingData <- trainingData[, !( names( trainingData ) %in% drops )]
-#  testingData <- testingData[, !( names( trainingData ) %in% drops )]
 
-trainingData<-trainingData[,3:ncol(testingData)]
-testingData<-testingData[,3:ncol(testingData)]
+trainingData <- trainingData[, 3:ncol( testingData )]
+testingData <- testingData[, 3:ncol( testingData )]
 
 brainAgeVM <- rvm( AGE ~ ., data = trainingData, type = "regression",
                     kernel = "laplacedot",
                     verbosity = 1, tol = .Machine$double.eps,
                     minmaxdiff = 1e-3, cross = 0, fit = TRUE,
-                    na.action = na.omit , iterations = 200 )
-
+                    na.action = na.omit , iterations = 1000 )
 predictedAge <- predict( brainAgeVM, testingData )
 
-correlation <- cor( testingData$AGE, predictedAge )
-meanerr<-(  mean( abs(  testingData$AGE - predictedAge ) ) )
-cat( "Correlation (true vs. predicted age): ", correlation, " mean err " , meanerr , "\n", sep = '' )
[…]

### 2013-07-07 KapowskiChronicles/analytics/repeatability.R

@@ -1,3 +1,6 @@
+library( psych )
+library( ggplot2 )
+
 kirby <- read.csv( "labelResultsK_pairwise.csv" )
 oasis <- read.csv( "labelResultsO_pairwise.csv" )
 
@@ -37,21 +40,86 @@ repeatabilityDataFrame <- data.frame( rbind( kirby, oasis ) )
 repeatabilityDataFrame$ID <- as.factor( subjectID )
 repeatabilityDataFrame$ScanOrder <- as.factor( scanOrder )
 
-qvalues <- c()
-meanDifferences <- c()
-for( i in 6:37 )
-  {
-  myTest <- t.test( repeatabilityDataFrame[which( repeatabilityDataFrame$ScanOrder == "First"),i],
-                    repeatabilityDataFrame[which( repeatabilityDataFrame$ScanOrder == "Repeat"),i],
-                    paired = TRUE, conf.int = TRUE )
-  qvalues[i-5] <- myTest$p.value
-  meanDifferences[i-5] <- myTest$estimate
-  }
-qvalues <- p.adjust( qvalues, method = "fdr" )
+iccData <- data.frame( First = as.vector( data.matrix( repeatabilityDataFrame[which( repeatabilityDataFrame$ScanOrder == "First" ), 6:37] ) ),
+                       Repeat = as.vector( data.matrix( repeatabilityDataFrame[which( repeatabilityDataFrame$ScanOrder == "Repeat" ), 6:37] ) ) )
+
+iccResults <- ICC( iccData )
+
+# > ICC( iccData)
+# Call: ICC(x = iccData)
+#
+# Intraclass correlation coefficients
+#                          type  ICC  F  df1  df2 p lower bound upper bound
+# Single_raters_absolute   ICC1 0.95 40 1279 1280 0        0.95        0.96
+# Single_random_raters     ICC2 0.95 41 1279 1279 0        0.94        0.96
+# Single_fixed_raters      ICC3 0.95 41 1279 1279 0        0.95        0.96
+# Average_raters_absolute ICC1k 0.98 40 1279 1280 0        0.97        0.98
+# Average_random_raters   ICC2k 0.98 41 1279 1279 0        0.97        0.98
+# Average_fixed_raters    ICC3k 0.98 41 1279 1279 0        0.97        0.98
+#
+#  Number of subjects = 1280     Number of Judges =  2
+#
+# > iccResults$summary
[…]

### 2013-07-12 KapowskiChronicles/analytics/gender_study2.R
ENH:  Added phylo plot.

@@ -1,42 +1,19 @@
 # load required libraries
 library( boot )
-library( ANTsR )
 library( ggplot2 )
 library( scales )
 library( igraph )
-library( pheatmap )
-library( randomForest )
 library( reshape )
-library( e1071 )
-rgl.bg( color="white")
-labels <- antsImageRead('nirep.nii.gz',3)
-centroids <- LabelImageCentroids( labels, physical=TRUE )
-template <- antsImageRead('glasshead.nii.gz', 3)
-brain <- antsImageRead('glassbrain.nii.gz', 3)
-leftright <- antsImageRead('leftright.nii.gz', 3)
-# template <- maskImage(template, leftright, 1)
-# brain <- maskImage(brain, leftright, 1)
-id<-rotationMatrix( 0, 0, 1, 0) 
-lateralLeft <- rotationMatrix(pi/2, 0, -1, 0) %*% rotationMatrix(pi/2, -1, 0, 0)
-frontal <- rotationMatrix(pi*3/2, 1,  0, 0)  # %*% rotationMatrix( pi/2, -1, 0, 0)
-lateralRigt <- rotationMatrix(pi/2, 0,  1, 0) %*% rotationMatrix( pi/2, -1, 0, 0)
-par3d(userMatrix=id, windowRect=c(0,0,512,512), zoom=0.7)
-mysurf <- renderSurfaceFunction(list(template,brain), alphasurf=c(0.3,0.3), surfval=0.5, smoothsval=1.5,  alphafunc=1, mycol="cadetblue1")
-par3d(userMatrix=lateralLeft, windowRect=c(0,0,512,512), zoom=0.7)
-par3d(userMatrix=id, windowRect=c(0,0,512,512), zoom=0.7)
-par3d(userMatrix=lateralRigt, windowRect=c(0,0,512,512), zoom=0.7)
+library( ANTsR )
+library( ape )
 
-maximumNumberOfPermutations <- 1000
-sigma <- 5
-ages <- seq( 10, 80, by = 5 )
 #################################################################
 ##
 ##   function definitions
 ##
-##     * calculateCorrelationMatrix -
+##     * calculateCorrelationMatrix
 ##
 #################################################################
-
 
 calculateCorrelationMatrix <- function( mat, weights, nuis )
   {
@@ -61,6 +38,10 @@ calculateCorrelationMatrix <- function( mat, weights, nuis )
 ##   main routine
 ##
[…]

### 2013-07-12 SPIE2013/KK/Combined_results/agePredict.R
DEL:  Remove SPIE2013

@@ -1,58 +1,0 @@
-library( ggplot2 )
-
-resultsX <- list()
-resultsX[[1]] <- read.csv( file = "../IXI_results/labelresults.csv", header = TRUE )
-resultsX[[2]] <- read.csv( file = "../Kirby_results/labelresults.csv", header = TRUE )
-resultsX[[3]] <- read.csv( file = "../NKI_results/labelresults.csv", header = TRUE )
-resultsX[[4]] <- read.csv( file = "../Oasis_results/labelresults.csv", header = TRUE )
-
-
-results <- rbind( resultsX[[1]], resultsX[[2]], resultsX[[3]], resultsX[[4]] )
-
-cortical_labels <- c( "L occipital", "R occipital",
-                     "L cingulate", "R cingulate",
-                     "L insula",    "R insula",
-                     "L temporal pole",   "R temporal pole",
-                     "L superior temporal", "R superior temporal",
-                     "L infero temporal", "R infero temporal",
-                     "L parahippocampal", "R parahippocampal",
-                     "L frontal pole",    "R frontal pole",
-                     "L superior frontal","R superior frontal",
-                     "L middle frontal",  "R middle frontal",
-                     "L inferior",        "R inferior",
-                     "L orbital frontal", "R orbital frontal",
-                     "L precentral",      "R precentral",
-                     "L superior parietal", "R superior parietal",
-                     "L inferior parietal", "R inferior parietal",
-                     "L postcentral",       "R postcentral" );
-
-male_color = "darkred";
-female_color = "navyblue";
-
-age <- results$AGE;
-gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
-
-for( i in 1:32 )
-  {
-  th <- results[,i+4];
-
-  gender[which( gender == 1 )] <- 'male';
-  gender[which( gender == 2 )] <- 'female';
-
-  plotData <- data.frame( cbind( Age = age, Thickness = th, Gender = gender ) )
-  plotData <- transform( plotData, Gender = factor( Gender ) );
-
-  thickPlot <- ggplot( plotData, aes( x = Age, y = Thickness, group = Gender ) ) +
[…]

### 2013-07-12 SPIE2013/KK/Combined_results/asymmetry.R

@@ -1,50 +1,0 @@
-resultsX <- list()
-resultsX[[1]] <- read.csv( file = "../IXI_results/labelresults.csv", header = TRUE )
-resultsX[[2]] <- read.csv( file = "../Kirby_results/labelresults.csv", header = TRUE )
-resultsX[[3]] <- read.csv( file = "../NKI_results/labelresults.csv", header = TRUE )
-resultsX[[4]] <- read.csv( file = "../Oasis_results/labelresults.csv", header = TRUE )
-
-
-results <- rbind( resultsX[[1]], resultsX[[2]], resultsX[[3]], resultsX[[4]] )
-
-cortical_labels <- c( "L occipital", "R occipital",
-                      "L cingulate", "R cingulate",
-                      "L insula",    "R insula",
-                      "L temporal pole",   "R temporal pole",
-                      "L superior temporal", "R superior temporal",
-                      "L infero temporal", "R infero temporal",
-                      "L parahippocampal", "R parahippocampal",
-                      "L frontal pole",    "R frontal pole",
-                      "L superior frontal","R superior frontal",
-                      "L middle frontal",  "R middle frontal",
-                      "L inferior",        "R inferior",
-                      "L orbital frontal", "R orbital frontal",
-                      "L precentral",      "R precentral",
-                      "L superior parietal", "R superior parietal",
-                      "L inferior parietal", "R inferior parietal",
-                      "L postcentral",       "R postcentral" )
-
-
-
-for( i in seq( 1, 32, by = 2 ) )
-  {
-  left <- results[,4+i];
-  right <- results[,4+i+1];
-
-  asym <- t.test( left, right, alternative = "two.sided", paired = TRUE )
-
-  meandiff <- mean( left - right )
-
-  prefix <- ''
-  if( asym$p.value <= 0.01 )
-    {
-    prefix <- '**'
-    } else if( asym$p.value <= 0.1 )
-    {
-    prefix <- '*'
-    }
-
-
-  cat( prefix, sub( "L ", '', cortical_labels[i] ), ": p-value = ", asym$p.value,
-    ", mean difference = ", meandiff, "\n", sep = "" )
-  }

### 2013-07-12 SPIE2013/KK/IXI_results/agePredict.R

@@ -1,51 +1,0 @@
-library( ggplot2 )
-
-results <- read.csv( file = paste( "labelresults", ".csv", sep = '' ), header = TRUE );
-
-cortical_labels <- c( "L occipital", "R occipital",
-                     "L cingulate", "R cingulate",
-                     "L insula",    "R insula",
-                     "L temporal pole",   "R temporal pole",
-                     "L superior temporal", "R superior temporal",
-                     "L infero temporal", "R infero temporal",
-                     "L parahippocampal", "R parahippocampal",
-                     "L frontal pole",    "R frontal pole",
-                     "L superior frontal","R superior frontal",
-                     "L middle frontal",  "R middle frontal",
-                     "L inferior",        "R inferior",
-                     "L orbital frontal", "R orbital frontal",
-                     "L precentral",      "R precentral",
-                     "L superior parietal", "R superior parietal",
-                     "L inferior parietal", "R inferior parietal",
-                     "L postcentral",       "R postcentral" );
-
-male_color = "darkred";
-female_color = "navyblue";
-
-age <- results$AGE;
-gender <- cut( results$SEX, breaks = c( 0.5, 1.5, 2.5 ), label = c( "male", "female" ) );
-
-for( i in 1:32 )
-  {
-  th <- results[,i+4];
-
-  gender[which( gender == 1 )] <- 'male';
-  gender[which( gender == 2 )] <- 'female';
-
-  plotData <- data.frame( cbind( Age = age, Thickness = th, Gender = gender ) )
-  plotData <- transform( plotData, Gender = factor( Gender ) );
-
-  thickPlot <- ggplot( plotData, aes( x = Age, y = Thickness, group = Gender ) ) +
-#                stat_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x, method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fullrange = FALSE, fill = 'black', alpha = 0.5 ) +
-               stat_smooth( aes( group = Gender, colour = Gender ), formula = y ~ 1 + x + I(x^2), method = "lm", size = 1, n = 1000, level = 0.95, se = TRUE, fullrange = TRUE, fill = 'black', alpha = 0.5 ) +
[…]
